### 2019-07-25 manuscript.tex
Initial Overleaf Import

\documentclass[crop]{CSLB}%%%%where CSL is the template name

%The authors can define any packages after the \documentclass{CSL} command.
\usepackage{amsmath} %for dealing with mathematics,
\usepackage{amsthm} %for dealing with theorem environments,
\usepackage{cite} %%%for dealing with citations
\usepackage{hyperref} %for linking the cross references
\usepackage{graphics} %for dealing with figures.
\usepackage{algorithmic} %for describing algorithms
\usepackage{subfig} %for getting the subfigures e.g., "Figure 1a and 1b" etc.
\usepackage{url} %It provides better support for handling and breaking URLs.

%\usepackage{natbib}
\usepackage{boites}

%\usepackage{graphicx}
%\usepackage{float}
%\usepackage{psfig}
%\usepackage{amsfonts,amssymb}
%\usepackage{color}
%\usepackage{textcomp}
%\usepackage[T1]{fontenc}
%\usepackage{aecompl}
%\usepackage{times}
%\usepackage{txfonts}
%\usepackage[english]{babel}
%\usepackage[utf8x]{inputenc}
%\usepackage{amsmath}
%\usepackage{caption}
%\usepackage{siunitx}

\jname{International Journal of Astrobiology}%

\newtheorem{theorem}{Theorem}
\newtheorem{condition}{Condition}

\begin{document}

\bibliographystyle{mn2e}

\supertitle{Research Paper}

\title[Probabilities of SETI contacts]{Probability of causal
contact between interstellar civilizations through MonteCarlo simulations}

\author[Lares \& Funes]{Lares M$^{1, 2}$, Funes J$^{1, 3}$}

\corres{\name{Lares M} \email{marcelo.lares@unc.edu.ar}}

\address{
   %
   \add{1}{CONICET, Argentina}

   \add{2}{Universidad Nacional de C\'ordoba, Observatorio
   Astron\'omico de C\'ordoba, Argentina}

   \add{3}{Universidad Cat\'olica de C\'ordoba, Argentina}
   %
}

\begin{abstract}
%
The abundance of intelligent civilizations in the galaxy is a
longstanding question, which is often conceptualized as the problem
of the lack of received communication of the Fermi paradox.
%
Most efforts on the estimation of a number of intelligent
civilizations are centered on the Drake equation, although 
its factors are affected by large uncertainties, and it lacks a
temporal nature.
%
Here we present numerical simulations of stochastic processes of
emergence of civilizations with communication capabilities, and
the causal contacts among them.
%
We analyze the rate of causal contacts as a function of the mean
number of civilizations and the mean lifetime span distribution.
%
Our results indicate that, given the large distances involved, in
the odds to obtain a contact in a few years of monitoring, assuming a
perfect detection rate, are very rare.                                 
%
\end{abstract}

\selfcitation{Lares M, Funes J (XXXX).  Estimating the probability of
causal contact between civilizations through MonteCarlo simulations,
International Journal of Astrobiology}

\received{xx xxxx xxxx}
\revised{xx xxxx xxxx}
\accepted{xx xxxx xxxx}

\keywords{SETI, Computer simulations, Statistics}

\JELclassification{XX; XX}
\MSCcodes{XX; XX}

\Abbreviations{CETI: Communicating Extraterrestrial Intelligence, DE:
Discrete Event, GHZ: Galactic Habitable Zone}

\maketitle

\Fpagebreak

\section{Motivations: Uncertainties and the non detection problem for
ETI contact}
%{{{

A common approach to discuss the possible number of extraterrestrial
intelligent civilizations (ETI) has been the use of the Drake equation
\citep{Gleiser2010, Prantzos2013, Haqq-Misra2017}.
%
However, the uncertainties in the factors make it less prone to a
formal application to define searching strategies or to compute
actual numbers, but is used to organize the discussion instead.
%
Some modifications to the original idea of the equation have tried to
imprint a stochastic nature, or to propose a probabilistic approach,
or to consider the temporal structure which is missing in the
equation.
%
Temporal aspects of the distribution of communicating civilizations
and their contacts have also been explored by several authors
\citep{Fogg1987, Forgan2011, Balbi2018},
%The Impact of the Temporal Distribution of Communicating Civilizations on Their Detectability
%Temporal dispersion of the emergence of intelligence: an inter-arrival time analysis
%The spatiotemporal aspects of SETI
%
as well as efforts on considering the stochastic nature of the Drake equation
\citep{Glade2011}.
%
The simulation approach has also benn considered
\citep{Forgan2008, Forgan2010}.
%New numerical determination of habitability in the Galaxy: the SETI connection

Since the factors in the Drake equation are uncertain, we propose to
avoid the frequentist approach of this equation
and to explore a parameter space, where instead of computing a final
number, we provide a statistical distribution that gives conditional
probabilities.
%
The only fact that can be stated with certainty is that for the number
of years SETI projects have been working we have not received any
signal within the conditions stablished by SETI.

The discrete events method for simulating a stochastic process is an
approximation that allows to study the behaviour of complex
systems, by considering a sequence of well defined discrete events.
%
The simulation os carried out by following all the variables that
describe the system, that constitute the state of the system.
%
The evolution of the process, then, is described as a set of changes
in the state of the system.
%
In this context, an event produces a specific change in the state,
that can be triggered by random variables that encode the stochastic
nature of the physical phenomenon.
%
The process involves following the changes on the state of the system,
definig the initial and final states, defining a method that allows to
keep track of time progress, and maintaining a list of relevant
events.

In a recent work, \citep{Balbi2018} use a statistical model to analyze
the occurrence of causal contacts between civilizations in the Galaxy.
%
The author highlights the effect of evolutionary processes when
attempting to estimate the number of communicating civilizations that
might be in causal contact with an observer located on the Earth.

\citet{cirkovic_temporal_2004} also emphasize the lack of temporal
structure in the Drake equation.

%Our results highlight the necessity, which has been noted
%elsewhere (Ćirković, 2004), of modeling 
%
%
%While the Drake equation remains a very useful
%guide in identifying and analyzing the various factors in-
%volved in the problem and their relative probabilities, our
%approach could provide a better framework with which to
%perform a statistical analysis when taking into account evo-
%lutionary processes. In particular, we have shown that the
%Drake equation can misestimate the number of detectable
%communicating species if spatial and temporal dependencies
%are important. This should suggest some caution in the un-
%critical application of such an equation. In this paper, we offer
%an exemplificative illustration of the effect of different
%probability distributions for the typical timescales involved
%in the problem. Ideally, it would be desirable to arrive at a
%mathematical description of such distributions based on a
%modeling of the underlying astrophysical and biological
%processes. This points out a possible future direction of in-
%vestigation in which an integrated evolutionary (and coevo-
%lutionary) approach to the problem would be a central focus
%of interest (for more on this perspective, see, e.g., Chaisson

In this work we address the problem of the temporal and spatial
structure of the distribution of communicating civilizations, by
exploring the hypothesis space over a minimal set of parameters.
%
In Sec.~\ref{S_methods} we introduce the methods and discuss the candidate
distributions for the statistical aspects of the times involved in the
communication process.
%
Then, we present our results in Sec~\ref{S_results}, with special emphasis on
the statistical distributions of the duration of causal contacts in one or
both directions, the possible differences on the position of a
civilization on the Galaxy and the distribution of time intervals for
the waiting of the first contact, always as a function of the three
simulation parameters.
%
Finally, in Sec.~\ref{S_discussion} we discuss our results and future research
directions.

%}}}

\section{Methods: accounting for causal contacts with 
discrete event simulation process}\label{S_methods}
%{{{

Some complex stochastic processes can be efficiently modeled with the
discrete--event (DE) simulation approach.
%
Simulations are suitable tools to analyze systems that evolve with
time and involve randomness.
%
In general, simulations require less assumptions and simplifications
than theoretical approaches, and can be applied to systems where a
theoretical model can not be found.
%
A system described with the DE paradigm is characterized by a set of
actors and events, where actors interact causally through a series of
events on a time line and process these events in chronological order
\citep{ptolemaeus_system_2014, chung_simulation_2003,
simulation_ross_2012}.
%
On the particular case of the difusion of signals of intelligence in
the Galaxy, the discrete events method can be performed taking into
account a small number of variables.
%
The temporal structure is defined by two distribution parameters that
represent the mean time interval that an intelligent civilization can
emit and receive signals, and the mean time interval between the
emergence of an intelligent comunicating civilization and the next
one.
%
The spatial structure of the simulation is given by the size and shape
of the Galactic Habitable Zone and the maximum distance a signal can
travel to be detected.
%
The parameters for the temporal distributions also determine the
spatial properties.
%
For example, the density of active cetis in the Galaxy depend on these
two parameters.
%
Also, some hypothesis must be made in order to complete the
simulation.
%
Firstly, we assume that the distribution of the appeareance of new
CETIs is a stationary Poisson distribution.
%
Regarding the duration of a CETI, we propose that the distribution of
the duration of CETIs is a stationary exponential distribution.
%
\citet{Balbi2018} stress the fact that it would be desirable to arrive
at a theoretical statistical distribution of the lifespan of
civilizations, preferentially on the basis of the underlying
astrophysical and biological processes.
%
\citet{Maccone} argues that this distribution should be a log normal.
(EXPAND)

There is no a temporal window for the model, i.e., the simulation
starts when the process is already stable, and ends before any
galactic effect could modify or affect of fixed variables.
%
The probability for the rise of a CETI is homogeneous over the GHZ.
%
Although the Galaxy has a well known spiral structure, the assumed
sparcity of civilizations makes this assumption very reasonably.
%
Otherwise, if the distribution of civilizations is not sparce, it
could be the case that the spiral arms would host most of the
civilizations, and contacts are frecuent between closely located
civilizations.
%
In such a case we could expect that the interpretation of our results
could be considered as pesimist for the stablishment of a contact with
our planet.
%
We also limit the posibilities of life or other types of civilizations
to the hypothesis usually stated for the definition of the GHZ.
%
This means that we set aside possibel civilizations that could survive
in severe conditions or unstable systems, which would prevent the
appeareance of life as we know it.

\setlength{\tabcolsep}{10pt}
\begin{table}
\centering
\begin{tabular}{cl}
\hline
   $\tau_{a}$ & Mean temporal separation between consecutive awakenings \\
   $\tau_{s}$ & Mean lifetime of a CETI \\
   D$_{max}$ & Maximum reach of a message \\
\hline
\end{tabular}
\caption{Definition of variables that are part of the simulation.}
\label{T_simu}
\end{table}

While the aforementioned assumptions are basic conditions for most of
the stochastic processes observed in nature, 
we need to make stronger assumptions related to the nature of the
message.
%
The most simple assumption about the message itself, is that it
travels at light speed.
%
With this, we are considering messages sent through, for example,
electromagnetic radiation or even gravitational radiation, but we set
appart messages sent with mechanical means or physical objects, or
through some unknown technology that violates the known so far laws of
physics.
% 
For the communication of CETIs through messages sent isotropically, 
we assume that the capacity to emit signals and to receive
signals ocurr at the same time.
%
This means that the ability to find another civilization develops at
the same time that the ability and intention to emit a signal in all
direction, intented to be detected by an unknown civilization.
%
Although there are several reasons to think that this could not be the
case, at large time scales it can be considered that both abilities
occurr roughly at coincident epochs.
%
Another essential assumption is that all CETIs use the same signal power, so that there is a maximum distance out to which it can be detected.
%
It is worth mentioning that we are considering in our experimental
setup a system composed by several emmiters and receivers across the
Galaxy, under the hypothesis proposed before.
%
In such a system we compute probabilities of a CETI making contact
with another CETI, and the results must be interpreted as for a
randomly chosen CETI, given certain conditions of homogeneityon an
ensemble of equivalent galaxies.
%
We stress the fact that we are not centered on the probabilities for
the Earth itselft, but on a generalized CETI.
%
Regarding the extent of the signals, we know that the distance from
which a signal from Earth could be detected using the current
technology is of a few hundreds light years, given that the signal was
sent to a specific direction.

%https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
%https://www.engadget.com/2017/05/25/listening-to-starlight-our-ongoing-search-for-alien-intelligenc/]
%[from https://en.wikipedia.org/wiki/Fermi_paradox]
%
It is straightforward to propose and implement a distribution of
maximum distances, although this would increase the model complexity
at the cost of a larger uncertainty or of another dimension on the
hypothesis space.
%
This proposal is based on a pronoid scenario (REF), and we do not
consider in this work the ocurrence of paranoid or partially paranoid
civilizations.
%
If that would be the case, the results obtained here can be taken as
upper limits.

In our simulations we also assume the simplest situations for the
growth of ETIs.
%
For example, we discard the possibility of stellar colonization
(Walters 1980) and assume that the communication aim is performed to
all directions with the same power and the same probability, i.e.,
strongly assume isotropic communication in all cases.
%
More detailed simulations could be produced considering different 
efficiency of communication or detection methods.
%
Also, a mixed paranoid/pronoid population of intelligent civilizations
could be implemented easily.
%
However, in this case we also would complicate the experimental setup,
make the results less evident.
%
The results are independent of the nature of life (organic or
artificial).
%
The lifetime of a civilization can be caused by autodestruction or by
external factors.

\subsection{Power laws vs. exponential laws}

The exponential distribution of lifespan and waiting times is
justified by considering that the process of appeareance of life in
the galaxy is homogeneous and stationary.
%
This means that there is no preferred location within the GHZ for the
spontaneos appeareance of life, and that the emergence of a
civilization is independent of the existence of previous civilizations 
in the galaxy.
%            
This is equivalent to proposing a Poisson process for the emergence of
CETIs, since there is a close relation between the number of events in
time or space and the waiting time or separation, respectively.
%
That is, these are two alternative approaches to describing the same
process, a Poisson distribution for the number of events implies an
exponential distribution for their sepations, and viceversa.
%
It should be emphasized that the exponential laws used in this work
are assumed, although instead of analyzing results from a particular
parameter chosen ad--hoc, we explore the hypothesis space and analyze
the impact of the values of these parameters on the results.
%

The power law and exponential statistical distributions are among the
most common patterns found in natural phenomena.
%
For example, the distribution of the frecuency of words in many
languajes is known to follow a Zipf law (which is a power law).
%
Zipf law also describes population ranks of cities in various
countries, corporation sizes, income rankings, ranks of number of
people watching the same TV channel (Zipf's law entry on Wikipedia).
%
The magnitudes of earthquakes, hurricanes, volcanic eruptions and
floods; the sizes of meteorites or the losses caused by business
interruptions from accidents, are also well described by power laws
(Sornette 2006).
%
The power law behaviour has been observed in a variety of systems,
including for example stock market fluctuations, sizes of computer
files and word frequency in languages \citep{mitzenmacher_brief_2004,
Newman_power_2005, simkin_theory_2006}
%
Power laws have also been widely used in biological sciences, e.g., in
analysing connectivity patterns in metabolic networks
\citep{jeong_large_2000} and in the number of species observed per
unit area in ecology \citep{garcia_origin_2006, frank_common_2009}.

\subsection{Age distribution of civilizations}

\subsection{Rate of appeareance of civilizations}

\subsection{Exploration of the hypothesis space}

Zipf law.

The adopted values for the GHZ are XX and XX \citep{Lineweaver2004}.

\setlength{\tabcolsep}{10pt}
\begin{table}
\centering
\begin{tabular}{cccc}
\hline
   fixed parameter & \multicolumn{3}{c}{adopted value}\hfill \\
\hline
   inner GHZ radius\hfill & \multicolumn{3}{c}{20000 ly}\\
   outer GHZ radius & \multicolumn{3}{c}{60000 ly}\\ \\
\hline
   variable parameter & min & max & N \\
\hline
   D$_{max}$ & 1000 ly & 20000 ly & 20\\
   $\tau_{a}$ & 1000 y & 20000 y & 20\\
   $\tau_{s}$ & 1000 y & 20000 y & 20\\
\hline
\end{tabular}
\caption{Adopted values for simulation parameters. Fixed parameters
   define the GHZ.  Variable parameters define the temporal structure
   of the process and the reach of messages.}
\label{T_parameters}
\end{table}

The main variables that follow the evolution of the simulation are:

\begin{itemize}
   \item Position of stars.  Sampled randomly within the GHZ
   \item Time of the awakening of each CETI
   \item Time of the vanishing of each CETI
\end{itemize}

The variables that can be deduced from the previous ones are:

\begin{itemize}
   \item Number of CETIs in casual contact with at least another CETI
      at a given time.
   \item Number of CETIs as a function of time
   \item Number of CETIs that receive a message at least one time
   \item Number of CETIs that receive a message at least one time and
      successfully deliver an answer.
   \item Number distribution of waiting time to receive a message
\end{itemize}

%========================================================================
%========================================================================
%========================================================================

\begin{figure}
   \centering
   \includegraphics[width=0.5\textwidth]{growingsphere.pdf}
   \caption{Schematic representation of the growing communicating
   sphere, over the space--time diagrams, where time is represented on
   the horizontal direction, and space is represented in the vertical
   direction.  In (a) the sphere is
   growing as the surface of first contact has not reached the maximum
   distance.  In (b) it has reached the maximum distance, so that it
   remains at the same size.  After a Doomsday event, the signals can
   still be observed, but the surface of last contact grows }
   \label{F_sphere}
\end{figure}

\begin{figure}
   \centering
   \includegraphics[width=0.5\textwidth]{abcd.pdf}
   \caption{Schematic representation of two emitters, $E_1$ and $E_2$,
   that reach each other at different times.  The time span for $E_i$
   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
   emitter $j$ between $C_{ij}$ and $B_{ij}$.
   The type and length of causal contact in both directions depend on
   the distance and time lag between the awakening events, the maximum
   distance that a signal can reach and the time period in which each
   emitter is active.}
   \label{F_abcd}
\end{figure}

Possible hypothesis improvements are:

\begin{itemize}
   \item Dmax is different for different CETIs.  For example a power
      law where powerfull emmisions are rare and weak emssions are
      common.
   \item The probability of the rising of a new CETI vary within the
      GHZ.
   \item Correction by coverage ratio in the detection and by
      targetting ratio in the emission.
   \item Spiritual factor.  Explore the role of the S factor: message
      contents, influence on the lifespan of a CETI that receives a
      message, mean lifespan of a CETI that emits a message.
\end{itemize}

\subsection{Model for CETIs}

Since the method is based on events, the first step is to define an
architecture of events, the relationships between them and how events
trigger changes on the state of the system.
%
We consider a simulated system that represents the galactic habitable
zone (GHZ).
%
On a first approach, the GHZ is a 2-dimensional annular region.
%
This simple model does not take into accont the variations in stellar
density given by the spiral structure.
%
Although there are several possible approaches, we chose to follow the
evolution of the system according to the following events:

\begin{enumerate}
   \item[(A)] A new CETI appears (Awakening)
   \item[(D)] An old CETI dissappears (Doomsday)
   \item[(C)] A new causal contact is stablished (Contact)
   \item[(B)] An existing causal contact is interrupted (Blackout)
\end{enumerate}

The event of type B is produced when one of the two CETIs halts in its
capability of emmiting and receiving signals.

%======================================================================
%======================================================================
%======================================================================

The system is updated each time an event is produced.
%
The time marks for each event (A and D for each CETI and C and B for
each causal contact) are stored as a result of the simulation.
%
Also, the list of active CETIs is obtained as a function of time.
%
Some fixed variables must be set in order to carry out the simulation,
namely:

\begin{itemize}
   \item Size of the Galactic Habitable Zone,
   \item mean lifetime of a CETI, $\lambda_1$,
   \item mean waiting time for the appeareance of another CETI,
      $\lambda_2$, and
   \item maximum distance a message can be detected.
\end{itemize}

The Galactic Habitable Zone is the only one of the four parameters
that is barely known, is set at fixed values of the inner and
outer radii, radial symetry is assumed, and the with of the
galactic disk is neglected compared to the radial size.
%
The two radii are measured in light years, with the aim to maintain a
single and comprehensive unit for both space and time coordinates.

The simulation was implemented on a python code, which is publically
available\footnote{https://github.com/mlares/simu\_contact}

%Variables en el programa
%\begin{itemize}
%   \item Nact: Número de CETIs activas
%   \item Nsee: Número de contactos por cada CETI
%   \item Nccc: CETIs en contacto causal. Notar que una CETI puede tener múltiples
%contactos de otras CETIs. Para cada una de ellas, guardar:
%   \item ID: ID de la CETI emisora
%   \item ID: ID de la CETI receptora
%   \item t\_hola: tiempo del primer contacto
%   \item t\_chau: tiempo del último contacto
%\end{itemize}
%
%
%Caso 1: aparece una nueva CETI
%Modificaciones al estado del sistema:
%\begin{itemize}
%   \item Agregar la CETI “A” a la lista de CETIs activas
%   \item Sortear la duración de la actividad de comunicación de la CETI “A”
%   \item Sortear el intervalo de tiempo hasta la aparición de la próxima CETI:
%t\_WakeUp\_next
%   \item Agregar la CETI “A” al árbol
%   \item Identificar todas las CETIs que están a menos de D max de la nueva CETI, y
%actualizar la lista de tiempos
%\end{itemize}
%
%Caso 2: desaparece una CETI
%Modificaciones al estado del sistema:
%\begin{itemize}
%   \item Eliminar la CETI a la lista de CETIs activas
%\end{itemize}
%
%
%Caso 3: una nueva CETI entra en contacto causal
%Modificaciones al estado del sistema:
%\begin{itemize}
%   \item Agregar la CETI B a la lista de contactos de la CETI A
%\end{itemize}
%

%}}}

\begin{figure*}
   \centering
   \includegraphics[width=\textwidth]{Messages_01.pdf}
   \caption{Schematic representation of the possible cases in which a
   ceti A can be in causal contact with a ceti B which appears later.
   The duration of the causal contact in one direction depends on
   several factors, mainly $\Delta t_A$, $\Delta t_C$, and $\Delta X$.}
   \label{F_messages}
\end{figure*}

\section{Results: exploring the parameter space}\label{S_results}
%{{{

We implemented the simulation of a regular grid of models varying over the
hypothesis space, which covers XXX different models.
%
For each model, we simulated XXX realizations with different random
seeds, in order to estimate the variance.
%
The simulated parameters, $\lambda_1$, $\lambda_2$ and $D_{max}$ cover
the range XX-XX, XX-XX and XX-XX, respectively, with a partition of XX
values for each parameter.

As a product of the simulations, several quantities are obtained.
%
Some quantities are directly derived from the discrete events, namely,
the ID of emitting CETI, the ID of emitting (receiving) CETI, the
position in the galaxy, the time of Awakening (Contact) and the time
of Doomsday  (Blackout) for each CETI.
%
This structure is always for the A-D cycle of each emitting CETI, and
repeats for each contact (C-B).
%
We can also derive the additional quantities.
%
The age of a CETI is the time elapsed from A to D.
%
%
%given time, ceti\_e is the CETI emitting signals (emiter), ceti\_r is
%the CETI receiving signals (receiver), ceti\_c is the ceti that
%listen at least another ceti (citizen) and ceti\_h is the CETI ceti
%that is lestened by at least another CETI.
%
The time span of a CETI listening another or being listened by another
can also be easily derived from the results of a realization.
%
This way we can also compute the distribution in the galaxy of cetis
that reach contact and the waiting time until the first contact or the
waiting time until the next contact.
%
Regarding the properties of the population of CETIs and its evolution,
we analyze the fraction of awaken time a ceti is listening at least
another ceti, the age of contacted CETIs at first contact, the
distribution of time to wait until next contact, the fraction of cetis
where the first contact is given at the awakening, the distribution of
the number of contacts for each CETI, the distribution of the maximum
number of contacts for each CETI, the distribution of the number of
contacts as a function of CETI age, the number of contacts as a
function of time in the galaxy, the rate of CETIs that succeed in
contact, and the distribution of distances between contacted CETIs.
%
Usefull derived quantity is the duration of two-way communication
channels the or the fraction of contacts that admit a response.
% 
We also analyze the relations between the distance to ceti vs. time of
double communication, the distance to ceti vs. age of contacted ceti,
the age of a ceti and the maximum number of contacted cetis before D,
or the lifespan of a ceti vs. max number of contacts.
%
All these quantities can be analyzed as a function of the simulation
parameters.
%

We chose eight models that are on fairly oposed regions of the
explored parameter space, and cover short/long lifetimes, short/long
waiting times for the next CETI to appear in the Galaxy and
short/large range reach of the signal, including all possible
combinations.
%
The details of each model are summarized in the
Table~\ref{T_SelectedModels}.

In Fig.~\ref{F_res_1} we show the Empirical cumulative distributions
of the number contacts for CETIs in four different samplas (M5, M6,
M7, M8) with long range signal reach (D$_{max}$=10000). 
%
The lifetime is more determinant than the awakening rate in increasing
the number of contacts.
%
As expected, of the four samples, the sample M7, which corresponds to
a dense awakening in the timeline and a long lifetime, is the model
that maximizes the number of contacts, reaching a maximum of order 100
contacts for a single CETI.
%
This is considerable larger than any model with a shorter lifetime,
which produce a number of contacts of at most the order of ten
contacts per CETI.
%
A couple of comments are worth to mention about this plot.
%
First, it has a logarithmic scale on the x-axis, and second it is the
cumulative, not differential, empirical distribution.

In what follows, we focus on the analysis of the duration of contacts
(Sec. X), the waiting time for the first contact (Sec. X) the
variations of contact posibilities as a function of the position of
the CETI in the Galaxy (Sec. X) and the occurrence of reciprocal
communication (Sec. X).

%}}}

\subsection{Duration of contacts}
%{{{

The duration of each contact for a single receiving CETI is given by the
time difference between a C event and a B event, which are direct
outputs for each simulation experiment.

%}}}

\subsection{Waiting time for a first contact}
%{{{

In this section we analyze the distribution of the waiting times for
a first contact.
%
Such distribution can be considered to compute the probability of
reaching a first contact within the first XX years.

%}}}

\subsection{Reciprocal communication}
%{{{

%}}}

\subsection{Location within the Galaxy}
%{{{

%}}}

\setlength{\tabcolsep}{10pt}
\begin{table*}
\centering
\begin{tabular}{cccccc}
\hline
Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & size & description  \\
M1 & $<$5000 & $<$15000 & $<$15000 &800  &dense awakening and short lifetime\\
M2 & $<$5000 & $>$20000 & $>$50000 &1050 &sparse awakening and long lifetime\\
M3 & $<$5000 & $<$5000  & $>$15000 &1350 &dense awakening and long lifetime \\
M4 & $<$5000 & $>$15000 & $<$26000 &1200 &sparse awakening and short lifetime\\

M5 & $>$20000 & $<$15000 & $<$15000 &800  &dense awakening and short lifetime\\
M6 & $>$20000 & $>$20000 & $>$50000 &1050 &sparse awakening and long lifetime\\
M7 & $>$20000 & $<$5000  & $>$15000 &1350 &dense awakening and long lifetime \\
M8 & $>$20000 & $>$15000 & $<$26000 &1200 &sparse awakening and short lifetime\\
\hline
\end{tabular}
\caption{Selected models to analize the behevior of some variables.}
\label{T_SelectedModels}
\end{table*}

\begin{figure} \centering
   %
   \includegraphics[width=0.5\textwidth]{Figure_1.png}
   %
   \caption{Empirical cumulative distributions of the number contacts
   for CETIs in four different samplas (M5, M6, M7, M8) with long
   range signal reach (D$_{max}$=10000).
   %
   }
   % 
\label{F_res_1} \end{figure}

\begin{figure} % Distribution
   \centering
   \includegraphics[width=0.5\textwidth]{Figure_2.png}
   \caption{Rate of CETIs that make no contact as a function of
   $\tau_a$ and $\tau_s$}
   \label{F_res_2}
\end{figure}

\begin{figure}
   \centering
   \includegraphics[width=0.5\textwidth]{Figure_3.png}
   \caption{Distribution of number of contacts as a function of
   galactocentric distance.}
   \label{F_res_5}
\end{figure}

\begin{figure} % 2D color plot
   \centering
   \includegraphics[width=0.5\textwidth]{Figure_4.png}
   \caption{}
   \label{F_res_4}
\end{figure}

 \begin{figure} % 2D color plot
   \centering
   \includegraphics[width=0.5\textwidth]{Figure_5.png}
   \caption{Rate of CETIs that listen at least one time, and make its
   first contact at the moment of the awakening, as a funcion of
   $\tau_a$ and $\tau_s$}
   \label{F_res_3}
\end{figure}

\section{Discusion}\label{S_discussion}
%{{{

We have implemented a simulation approach to explore the hypothesis
space about the causal connections between communicating
extraterrestrial civilizations.
%
The different models can be generated with a minimal number of free
parameters and some fixed assumptions.
%
We argue that three parameters can be used to describe a variety of
situations, ranging from a Galaxy where an intelligent civilization is
very rare, to a Galaxy populated with plenty of civilizations in
causal contact.
%
We explored the consecuences of different models using mensurable
quantities.
%
With this approach, we can estimate several quantities as a function
of the parameters on the hypothesis space.
%
For instance, it is possible to estimate the timescale for an
uninterrupted SETI effort to reach success, in terms of different
model parameters that reflect different, so far unknown, scenarios
for the spread if intelligent life on the Galaxy.

Under the hypotheses of our experiments, we conclude that
a causal contact is extremely unlikely unless the Galaxy is heavily
populated by intelligent civilizations.
%
Even so, we must consider that in order to stablish a contact between
any two entities, a minimum degree of compatibility must be
accomplished without any previous agreement, 
making the posibility of a contact with a message that
could be deciphered highly rare.

Future work can be performed following this framework, in order to
explore possible implications of the results for more detailed 
configurations of the experiment.
%
For example, the communication method (isotropic, colimated,
serendipitous) can affect the observables.
%
Also the nature of the message carrier, electromagnetic or another.
%
It is also possible to explore the effects of alignments or the use of
stars as sources or amplifiers.
%
%energetic markers
%Spiritual markers
%
As a future work, we will explore several improvements to the
simulation, and analyze the impact on the final results of these
changes.
%
They include a spatial distribution that actually resembles the shape
of the Galaxy, different distribution functions for the mean lifetime
of CETIs and different efficiencies for the sphere of the causal
region of each CETI, resembling different searching strategies.

%}}}

\ack[Acknowledgement]{
%
This work was partially supported by the Consejo Nacional de
Investigaciones Cient\'{\i}ficas y T\'ecnicas (CONICET, Argentina),
the Secretar\'{\i}a de Ciencia y Tecnolog\'{\i}a, Universidad Nacional
de C\'ordoba, Argentina, and the Universidad Católica de Córdoba,
Argentina.
%
This research has made use of NASA's Astrophysics Data System.
%
Plots and simulations were made with software developed by the authors
using R and python languajes, plots were postprocessed with inkscape.
%
}

\bibliography{biblio_seti}

%%%%%%%%%%%% Biography text %%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Author Biography}

\aubio{ML.jpeg}{\textbf{Marcelo Lares} received a degree in physics from the University
of A in 1998 and received his Ph.D. degree in communication engineering in
2002. He now holds a research chair at the B institute. His main research
interests are design and optimization of high power microwave power
amplifiers.}

\aubio{JF.jpeg}{\textbf{Jos\'e Funes} received her Diploma in 1997 from the University of
C and received her Ph.D. at the D University in 2001. She became a full
university professor in 2009. She is currently serving as chair of the E
project and is active in the study of Si power electronics.}

\end{document}

DEVELOPMENT DEVELOPMENT  DEVELOPMENT  DEVELOPMENT  DEVELOPMENT  DEVELOPMENT

% Papers para citar y discutir:

%## GHZ

%## D_max

%## exponential distribution

%## Lifetime

%## Simulations: cosmological

Forgan2015

%## Simulations: stochastic

grimaldi_signal_2017

%## Temporal aspects
Temporal aspects of the interaction among the first galactic civilizations: The 'interdict hypothesis'
by M.J. Fogg

%## Drake equation

%## Detalles de la simulacion
Galactic civilizations: Population dynamics and interstellar diffusion
by W.I. Newman, C. Sagan

### 2019-07-25 softdoc.tex
Initial Overleaf Import

--- 2019-07-25 manuscript.tex
@@
-\documentclass[crop]{CSLB}%%%%where CSL is the template name
+\documentclass[useAMS,usenatbib]{mnras}
+% A Monthly Notices of the Royal Astronomical Society sample TeX file.
+% The teX file has had all publication information removed and been left intentionally blank
+\usepackage{graphicx}
+\usepackage{float}
+\usepackage{graphicx}
+\usepackage{psfig}
+\usepackage{amsfonts,amssymb,amsmath}
+\usepackage{hyperref,color}
+\usepackage{textcomp}
+\usepackage[T1]{fontenc}
+\usepackage{aecompl}
+\usepackage{times}
+\usepackage[colorinlistoftodos]{todonotes}
+\usepackage{txfonts}
+\usepackage[english]{babel}
+\usepackage[utf8x]{inputenc}
+\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{siunitx}
+%\usepackage{booktabs}
 
-%The authors can define any packages after the \documentclass{CSL} command.
-\usepackage{amsmath} %for dealing with mathematics,
-\usepackage{amsthm} %for dealing with theorem environments,
-\usepackage{cite} %%%for dealing with citations
-\usepackage{hyperref} %for linking the cross references
-\usepackage{graphics} %for dealing with figures.
-\usepackage{algorithmic} %for describing algorithms
-\usepackage{subfig} %for getting the subfigures e.g., "Figure 1a and 1b" etc.
-\usepackage{url} %It provides better support for handling and breaking URLs.
+\usepackage[colorinlistoftodos]{todonotes}
 
-%\usepackage{natbib}
-\usepackage{boites}
+%\usepackage[authoryear]{natbib}
+%\graphicspath{ {images/} }
 
-%\usepackage{graphicx}
-%\usepackage{float}
-%\usepackage{psfig}
-%\usepackage{amsfonts,amssymb}
-%\usepackage{color}
-%\usepackage{textcomp}
-%\usepackage[T1]{fontenc}
-%\usepackage{aecompl}
-%\usepackage{times}
-%\usepackage{txfonts}
-%\usepackage[english]{babel}
-%\usepackage[utf8x]{inputenc}
-%\usepackage{amsmath}
-%\usepackage{caption}
-%\usepackage{siunitx}
+\def\aap{Astronomy \& Astrophysics}
+\def\apj{The Astrophysical Journal}
+\def\apjs{The Astrophysical Journal Supplement}
+\def\mnras{Monthly Notices of the Royal Astronomical Society}
+\def\pasp{Publications of the Astronomical Society of the Pacific}
+%
+\title{Estimating the probability of causal contact between
+civilizations through MonteCarlo simulations}
 
-\jname{International Journal of Astrobiology}%
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{condition}{Condition}
-
+% Group authors per affiliation:
+\author[M. Lares et al.]
+{M. Lares\thanks{E-mail: marcelo.lares@unc.edu.ar}\footnotemark[0] $^{1,2,3}$,
+J. Funes $^{3, 4}$
+\\
+% institutions
+$^{1}$Instituto de Astronom\'{\i}a Te\'orica y Experimental (CCT C\'ordoba, CONICET, UNC), Argentina.\\
+$^{2}$Observatorio Astron\'omico de C\'ordoba, Universidad Nacional de C\'ordoba, Argentina.\\
+$^{3}$Consejo Nacional de Investigaciones Cient\'{\i}ficas y
+T\'ecnicas, Argentina.\\
+$^{4}$Universidad Católica de Córdoba, Argentina
+}
+%
+%\hypersetup{draft}
 \begin{document}
 
-\bibliographystyle{mn2e}
+\date{Accepted . Received ; in original form }
 
-\supertitle{Research Paper}
-
-\title[Probabilities of SETI contacts]{Probability of causal
-contact between interstellar civilizations through MonteCarlo simulations}
-
-\author[Lares \& Funes]{Lares M$^{1, 2}$, Funes J$^{1, 3}$}
-
-\corres{\name{Lares M} \email{marcelo.lares@unc.edu.ar}}
-
-\address{
-   %
-   \add{1}{CONICET, Argentina}
-
-   \add{2}{Universidad Nacional de C\'ordoba, Observatorio
-   Astron\'omico de C\'ordoba, Argentina}
-
-   \add{3}{Universidad Cat\'olica de C\'ordoba, Argentina}
-   %
-}
-
-\begin{abstract}
-%
-The abundance of intelligent civilizations in the galaxy is a
-longstanding question, which is often conceptualized as the problem
-of the lack of received communication of the Fermi paradox.
-%
-Most efforts on the estimation of a number of intelligent
-civilizations are centered on the Drake equation, although 
-its factors are affected by large uncertainties, and it lacks a
-temporal nature.
-%
-Here we present numerical simulations of stochastic processes of
-emergence of civilizations with communication capabilities, and
-the causal contacts among them.
-%
-We analyze the rate of causal contacts as a function of the mean
-number of civilizations and the mean lifetime span distribution.
-%
-Our results indicate that, given the large distances involved, in
-the odds to obtain a contact in a few years of monitoring, assuming a
-perfect detection rate, are very rare.                                 
-%
-\end{abstract}
-
-\selfcitation{Lares M, Funes J (XXXX).  Estimating the probability of
-causal contact between civilizations through MonteCarlo simulations,
-International Journal of Astrobiology}
-
-\received{xx xxxx xxxx}
-\revised{xx xxxx xxxx}
-\accepted{xx xxxx xxxx}
-
-\keywords{SETI, Computer simulations, Statistics}
-
-\JELclassification{XX; XX}
-\MSCcodes{XX; XX}
-
-\Abbreviations{CETI: Communicating Extraterrestrial Intelligence, DE:
-Discrete Event, GHZ: Galactic Habitable Zone}
+%\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2014}
 
@@
 
-\Fpagebreak
+\label{firstpage}
+
+\begin{abstract}
+
+\end{abstract}
+%
+\begin{keywords}
+keyword1, keyword2
+\end{keywords}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Introduction
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %{{{
-
-A common approach to discuss the possible number of extraterrestrial
-intelligent civilizations (ETI) has been the use of the Drake equation
-\citep{Gleiser2010, Prantzos2013, Haqq-Misra2017}.
-%
-However, the uncertainties in the factors make it less prone to a
-formal application to define searching strategies or to compute
-actual numbers, but is used to organize the discussion instead.
-%
-Some modifications to the original idea of the equation have tried to
-imprint a stochastic nature, or to propose a probabilistic approach,
-or to consider the temporal structure which is missing in the
-equation.
-%
-Temporal aspects of the distribution of communicating civilizations
-and their contacts have also been explored by several authors
-\citep{Fogg1987, Forgan2011, Balbi2018},
-%The Impact of the Temporal Distribution of Communicating Civilizations on Their Detectability
-%Temporal dispersion of the emergence of intelligence: an inter-arrival time analysis
-%The spatiotemporal aspects of SETI
-%
-as well as efforts on considering the stochastic nature of the Drake equation
-\citep{Glade2011}.
-%
-The simulation approach has also benn considered
-\citep{Forgan2008, Forgan2010}.
-%New numerical determination of habitability in the Galaxy: the SETI connection
 
@@
 events.
-
-In a recent work, \citep{Balbi2018} use a statistical model to analyze
-the occurrence of causal contacts between civilizations in the Galaxy.
-%
-The author highlights the effect of evolutionary processes when
-attempting to estimate the number of communicating civilizations that
-might be in causal contact with an observer located on the Earth.
-
-\citet{cirkovic_temporal_2004} also emphasize the lack of temporal
-structure in the Drake equation.
-
-%Our results highlight the necessity, which has been noted
-%elsewhere (Ćirković, 2004), of modeling 
-%
-%
-%While the Drake equation remains a very useful
-%guide in identifying and analyzing the various factors in-
-%volved in the problem and their relative probabilities, our
-%approach could provide a better framework with which to
-%perform a statistical analysis when taking into account evo-
-%lutionary processes. In particular, we have shown that the
-%Drake equation can misestimate the number of detectable
-%communicating species if spatial and temporal dependencies
-%are important. This should suggest some caution in the un-
-%critical application of such an equation. In this paper, we offer
-%an exemplificative illustration of the effect of different
-%probability distributions for the typical timescales involved
-%in the problem. Ideally, it would be desirable to arrive at a
-%mathematical description of such distributions based on a
-%modeling of the underlying astrophysical and biological
-%processes. This points out a possible future direction of in-
-%vestigation in which an integrated evolutionary (and coevo-
-%lutionary) approach to the problem would be a central focus
-%of interest (for more on this perspective, see, e.g., Chaisson
-
-In this work we address the problem of the temporal and spatial
-structure of the distribution of communicating civilizations, by
-exploring the hypothesis space over a minimal set of parameters.
-%
-In Sec.~\ref{S_methods} we introduce the methods and discuss the candidate
-distributions for the statistical aspects of the times involved in the
-communication process.
-%
-Then, we present our results in Sec~\ref{S_results}, with special emphasis on
-the statistical distributions of the duration of causal contacts in one or
-both directions, the possible differences on the position of a
-civilization on the Galaxy and the distribution of time intervals for
-the waiting of the first contact, always as a function of the three
-simulation parameters.
-%
-Finally, in Sec.~\ref{S_discussion} we discuss our results and future research
-directions.
-
 %}}}
@@
 \section{Methods: accounting for causal contacts with 
-discrete event simulation process}\label{S_methods}
+discrete event simulation process}
 %{{{
 
-Some complex stochastic processes can be efficiently modeled with the
-discrete--event (DE) simulation approach.
-%
-Simulations are suitable tools to analyze systems that evolve with
-time and involve randomness.
-%
-In general, simulations require less assumptions and simplifications
-than theoretical approaches, and can be applied to systems where a
-theoretical model can not be found.
-%
-A system described with the DE paradigm is characterized by a set of
-actors and events, where actors interact causally through a series of
-events on a time line and process these events in chronological order
-\citep{ptolemaeus_system_2014, chung_simulation_2003,
-simulation_ross_2012}.
-%
 On the particular case of the difusion of signals of intelligence in
 the Galaxy, the discrete events method can be performed taking into
-account a small number of variables.
-%
-The temporal structure is defined by two distribution parameters that
-represent the mean time interval that an intelligent civilization can
-emit and receive signals, and the mean time interval between the
-emergence of an intelligent comunicating civilization and the next
-one.
-%
+account a small number of variables.  The temporal structure is
+defined by two distribution parameters that represent the mean
+time interval that an intelligent civilization can emit and receive
+signals, and the mean time interval between the emergence of an
+intelligent comunicating civilization and the next.
 The spatial structure of the simulation is given by the size and shape
@@
 travel to be detected.
-%
-The parameters for the temporal distributions also determine the
-spatial properties.
-%
-For example, the density of active cetis in the Galaxy depend on these
-two parameters.
-%
+
 Also, some hypothesis must be made in order to complete the
-simulation.
-%
-Firstly, we assume that the distribution of the appeareance of new
-CETIs is a stationary Poisson distribution.
-%
-Regarding the duration of a CETI, we propose that the distribution of
-the duration of CETIs is a stationary exponential distribution.
-%
-\citet{Balbi2018} stress the fact that it would be desirable to arrive
-at a theoretical statistical distribution of the lifespan of
-civilizations, preferentially on the basis of the underlying
-astrophysical and biological processes.
-%
-\citet{Maccone} argues that this distribution should be a log normal.
-(EXPAND)
+simulation, namely:
 
-There is no a temporal window for the model, i.e., the simulation
-starts when the process is already stable, and ends before any
-galactic effect could modify or affect of fixed variables.
-%
-The probability for the rise of a CETI is homogeneous over the GHZ.
-%
-Although the Galaxy has a well known spiral structure, the assumed
-sparcity of civilizations makes this assumption very reasonably.
-%
-Otherwise, if the distribution of civilizations is not sparce, it
-could be the case that the spiral arms would host most of the
-civilizations, and contacts are frecuent between closely located
-civilizations.
-%
-In such a case we could expect that the interpretation of our results
-could be considered as pesimist for the stablishment of a contact with
-our planet.
-%
-We also limit the posibilities of life or other types of civilizations
-to the hypothesis usually stated for the definition of the GHZ.
-%
-This means that we set aside possibel civilizations that could survive
-in severe conditions or unstable systems, which would prevent the
-appeareance of life as we know it.
+\begin{itemize}
+   \item The capacity to emit signals and to receive signals ocurr at
+      the same time.
+   \item All CETIs use the same signal power, so that there is a
+      maximum distance out to which it can be detected.
+   \item The distribution of the appeareance of new CETIs
+      is a stationary Poisson distribution.
+   \item The distribution of duration of CETIs
+      is a stationary exponential distribution.
+   \item There is no a temporal window 
+   \item The probability for the rise of a CETI is homogeneous
+      over the GHZ.
+   \item The message travels at light speed.
 
-\setlength{\tabcolsep}{10pt}
-\begin{table}
-\centering
-\begin{tabular}{cl}
-\hline
-   $\tau_{a}$ & Mean temporal separation between consecutive awakenings \\
-   $\tau_{s}$ & Mean lifetime of a CETI \\
-   D$_{max}$ & Maximum reach of a message \\
-\hline
-\end{tabular}
-\caption{Definition of variables that are part of the simulation.}
-\label{T_simu}
-\end{table}
+\end{itemize}
 
-While the aforementioned assumptions are basic conditions for most of
-the stochastic processes observed in nature, 
-we need to make stronger assumptions related to the nature of the
-message.
-%
-The most simple assumption about the message itself, is that it
-travels at light speed.
-%
-With this, we are considering messages sent through, for example,
-electromagnetic radiation or even gravitational radiation, but we set
-appart messages sent with mechanical means or physical objects, or
-through some unknown technology that violates the known so far laws of
-physics.
-% 
-For the communication of CETIs through messages sent isotropically, 
-we assume that the capacity to emit signals and to receive
-signals ocurr at the same time.
-%
-This means that the ability to find another civilization develops at
-the same time that the ability and intention to emit a signal in all
-direction, intented to be detected by an unknown civilization.
-%
-Although there are several reasons to think that this could not be the
-case, at large time scales it can be considered that both abilities
-occurr roughly at coincident epochs.
-%
-Another essential assumption is that all CETIs use the same signal power, so that there is a maximum distance out to which it can be detected.
-%
-It is worth mentioning that we are considering in our experimental
-setup a system composed by several emmiters and receivers across the
-Galaxy, under the hypothesis proposed before.
-%
-In such a system we compute probabilities of a CETI making contact
-with another CETI, and the results must be interpreted as for a
-randomly chosen CETI, given certain conditions of homogeneityon an
-ensemble of equivalent galaxies.
-%
-We stress the fact that we are not centered on the probabilities for
-the Earth itselft, but on a generalized CETI.
-%
-Regarding the extent of the signals, we know that the distance from
-which a signal from Earth could be detected using the current
-technology is of a few hundreds light years, given that the signal was
-sent to a specific direction.
+The fixed variables of the simulation are not explored:
 
-%https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
-%https://www.engadget.com/2017/05/25/listening-to-starlight-our-ongoing-search-for-alien-intelligenc/]
-%[from https://en.wikipedia.org/wiki/Fermi_paradox]
-%
-It is straightforward to propose and implement a distribution of
-maximum distances, although this would increase the model complexity
-at the cost of a larger uncertainty or of another dimension on the
-hypothesis space.
-%
-This proposal is based on a pronoid scenario (REF), and we do not
-consider in this work the ocurrence of paranoid or partially paranoid
-civilizations.
-%
-If that would be the case, the results obtained here can be taken as
-upper limits.
+\begin{itemize}
+   \item Size of the GHZ
+\end{itemize}
 
-In our simulations we also assume the simplest situations for the
-growth of ETIs.
-%
-For example, we discard the possibility of stellar colonization
-(Walters 1980) and assume that the communication aim is performed to
-all directions with the same power and the same probability, i.e.,
-strongly assume isotropic communication in all cases.
-%
-More detailed simulations could be produced considering different 
-efficiency of communication or detection methods.
-%
-Also, a mixed paranoid/pronoid population of intelligent civilizations
-could be implemented easily.
-%
-However, in this case we also would complicate the experimental setup,
-make the results less evident.
-%
-The results are independent of the nature of life (organic or
-artificial).
-%
-The lifetime of a civilization can be caused by autodestruction or by
-external factors.
+The variables that are explored with the simulation are:
 
-\subsection{Power laws vs. exponential laws}
-
-The exponential distribution of lifespan and waiting times is
-justified by considering that the process of appeareance of life in
-the galaxy is homogeneous and stationary.
-%
-This means that there is no preferred location within the GHZ for the
-spontaneos appeareance of life, and that the emergence of a
-civilization is independent of the existence of previous civilizations 
-in the galaxy.
-%            
-This is equivalent to proposing a Poisson process for the emergence of
-CETIs, since there is a close relation between the number of events in
-time or space and the waiting time or separation, respectively.
-%
-That is, these are two alternative approaches to describing the same
-process, a Poisson distribution for the number of events implies an
-exponential distribution for their sepations, and viceversa.
-%
-It should be emphasized that the exponential laws used in this work
-are assumed, although instead of analyzing results from a particular
-parameter chosen ad--hoc, we explore the hypothesis space and analyze
-the impact of the values of these parameters on the results.
-%
-
-The power law and exponential statistical distributions are among the
-most common patterns found in natural phenomena.
-%
-For example, the distribution of the frecuency of words in many
-languajes is known to follow a Zipf law (which is a power law).
-%
-Zipf law also describes population ranks of cities in various
-countries, corporation sizes, income rankings, ranks of number of
-people watching the same TV channel (Zipf's law entry on Wikipedia).
-%
-The magnitudes of earthquakes, hurricanes, volcanic eruptions and
-floods; the sizes of meteorites or the losses caused by business
-interruptions from accidents, are also well described by power laws
-(Sornette 2006).
-%
-The power law behaviour has been observed in a variety of systems,
-including for example stock market fluctuations, sizes of computer
-files and word frequency in languages \citep{mitzenmacher_brief_2004,
-Newman_power_2005, simkin_theory_2006}
-%
-Power laws have also been widely used in biological sciences, e.g., in
-analysing connectivity patterns in metabolic networks
-\citep{jeong_large_2000} and in the number of species observed per
-unit area in ecology \citep{garcia_origin_2006, frank_common_2009}.
-
-\subsection{Age distribution of civilizations}
-
-\subsection{Rate of appeareance of civilizations}
-
-\subsection{Exploration of the hypothesis space}
-
-Zipf law.
-
-The adopted values for the GHZ are XX and XX \citep{Lineweaver2004}.
-
-\setlength{\tabcolsep}{10pt}
-\begin{table}
-\centering
-\begin{tabular}{cccc}
-\hline
-   fixed parameter & \multicolumn{3}{c}{adopted value}\hfill \\
-\hline
-   inner GHZ radius\hfill & \multicolumn{3}{c}{20000 ly}\\
-   outer GHZ radius & \multicolumn{3}{c}{60000 ly}\\ \\
-\hline
-   variable parameter & min & max & N \\
-\hline
-   D$_{max}$ & 1000 ly & 20000 ly & 20\\
-   $\tau_{a}$ & 1000 y & 20000 y & 20\\
-   $\tau_{s}$ & 1000 y & 20000 y & 20\\
-\hline
-\end{tabular}
-\caption{Adopted values for simulation parameters. Fixed parameters
-   define the GHZ.  Variable parameters define the temporal structure
-   of the process and the reach of messages.}
-\label{T_parameters}
-\end{table}
+\begin{itemize}
+   \item Mean number of CETIs (Drake equation), or equivalently, the
+      mean time between the awakening of two consecutive CETIs,
+      $\tau_N$
+   \item Mean lifetime of a CETI, $\tau_a$
+   \item Maximum reach of CETI messages, $D_{max}$
+\end{itemize}
 
@@
 %========================================================================
-
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{growingsphere.pdf}
-   \caption{Schematic representation of the growing communicating
-   sphere, over the space--time diagrams, where time is represented on
-   the horizontal direction, and space is represented in the vertical
-   direction.  In (a) the sphere is
-   growing as the surface of first contact has not reached the maximum
-   distance.  In (b) it has reached the maximum distance, so that it
-   remains at the same size.  After a Doomsday event, the signals can
-   still be observed, but the surface of last contact grows }
-   \label{F_sphere}
-\end{figure}
-
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{abcd.pdf}
-   \caption{Schematic representation of two emitters, $E_1$ and $E_2$,
-   that reach each other at different times.  The time span for $E_i$
-   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
-   emitter $j$ between $C_{ij}$ and $B_{ij}$.
-   The type and length of causal contact in both directions depend on
-   the distance and time lag between the awakening events, the maximum
-   distance that a signal can reach and the time period in which each
-   emitter is active.}
-   \label{F_abcd}
-\end{figure}
 
@@
 
-The system is updated each time an event is produced.
-%
-The time marks for each event (A and D for each CETI and C and B for
-each causal contact) are stored as a result of the simulation.
+The system is updated each time an event is produced.  The time marks
+for each event (A and D for each CETI and C and B for each causal
+contact) are stored as a result of the simulation.
 %
 Also, the list of active CETIs is obtained as a function of time.
-%
+
 Some fixed variables must be set in order to carry out the simulation,
@@
 \begin{itemize}
-   \item Size of the Galactic Habitable Zone,
-   \item mean lifetime of a CETI, $\lambda_1$,
-   \item mean waiting time for the appeareance of another CETI,
-      $\lambda_2$, and
-   \item maximum distance a message can be detected.
+   \item Size of the Galactic Habitable Zone.  Radial symetry is
+      assumed, so that the GHZ is determined by two radii, the inner
+      radius (GHZ_inner) and the outer radius (GHZ_outer).  They are
+      measured in light years, with the aim to maintain y single and
+      comprehensive unit for both space and time coordinates.
+   \item Mean lifetime of a CETI
+   \item Mean waiting time for the appeareance of another CETI
+   \item Maximum distance a message can be detected
 \end{itemize}
-
-The Galactic Habitable Zone is the only one of the four parameters
-that is barely known, is set at fixed values of the inner and
-outer radii, radial symetry is assumed, and the with of the
-galactic disk is neglected compared to the radial size.
-%
-The two radii are measured in light years, with the aim to maintain a
-single and comprehensive unit for both space and time coordinates.
-
-The simulation was implemented on a python code, which is publically
-available\footnote{https://github.com/mlares/simu\_contact}
 
@@
 
+\section{Results: exploring the parameter space}
+%{{{
+
+As the result of the simulation, several quantities are obtained.
+
+%#
+%# CONTENTS OF THE <CETIS> PICKLE:
+%# . . . . . . . . . . . . . . . . . . . . . . . .
+%# 0. ID of emitting CETI
+%# 1. ID of emitting [receiving] CETI
+%# 2. x position in the galaxy
+%# 3. y position in the galaxy
+%# 4. time of Awakening [Contact]
+%# 5. time of Doomsday  [Blackout]
+%# . . . . . . . . . . . . . . . . . . . . . . . .
+%#
+%# This structure is always for the A-D cycle of each emitting CETI,
+%# and repeats for each contact (C-B).
+%#
+%# The time span of a CETI is t_D - t_A for CETIs[k][0]
+%# The time span of a contact is t_B - t_C for CETIs[k][i], with i>0
+%#
+%#
+%# age: time elapsed from A to a given time
+%# ceti_e: ceti emitting signals (emiter)
+%# ceti_r: ceti receiving signals (receiver)
+%# ceti_c: ceti that listen at least another ceti (citizen)
+%# ceti_h: ceti that is lestened by at least another ceti
+%#
+%
+%# for Check purposes
+%#-------------------
+%# duration of a civilization (exponential distribution by costruction)
+%# time from the appeareance of a CETI to the next
+%#
+%# 
+%# derived quantities
+%#-------------------
+%# [time]
+%# time span of a ceti listening another
+%# time span of a ceti being listened by another
+%# duration of two-way communication channels
+%# waiting time until the first contact
+%# fraction of awaken time a ceti is listening at least another ceti
+%# age of contacted cetis_e at first C
+%# distribution of time to wait until next C
+%# fraction of cetis where the first contact is given at the A
+%
+%# [numbers]
+%# distribution of the number of contacts for each ceti
+%# distribution of the maximum number of contacts for each ceti
+%# distribution of the number of contacts as a function of age
+%# number of contacts as a function of time in the galaxy
+%# rate of cetis that succeed in contact 
+%# fraction of contacts that admit a response
+%#
+%# [distances]
+%# distribution of distances between contacted cetis
+%#
+%# [correlations] relation between {RB}
+%# RB distance to ceti and time of double communication
+%# RB distance to ceti and age of contacted ceti
+%# RB the age of a ceti and the maximum number of contacted cetis before D
+%# RB lifespan of a ceti and max number of contacts
+%#
+%# distribution in the galaxy of cetis that reach contact
+%
+%# ... everything as a function of parameters!
+
+Output
+\begin{itemize}
+   \item La salida de la simulación está guardad en una variable llamada CETI.
+   \item La misma es una lista de objetos que son también listas.
+   \item El elemento cero de cada lista (i.e., cada elemento de CETI) contiene:
+   \item coordenada X
+   \item coordenada Y
+   \item tiempo de inicio de la capacidad de comunicación
+   \item tiempo de final de la capacidad de comunicación
+   \item y los demás elementos contienen:
+   \item coordenada X de la CETI
+   \item coordenada Y de la CETI
+   \item tiempo del primer contacto
+   \item tiempo del último contacto
+\end{itemize}
+
+%}}}
+
+\section{Results}
+%{{{
+
+Resultados cuantitativos de la simulación
+\begin{itemize}
+   \item distribucion del numero de contactos
+   \item distribucion de la duracion de los contactos
+   \item distribucion del tiempo que hay que esperar para el proximo contacto
+\end{itemize}
+
+\begin{figure}
+   \centering
+   \includegraphics[width=0.5\textwidth]{growingsphere.pdf}
+   \caption{Schematic representation of the growing communicating
+   sphere, over the space--time diagrams.  In (a) the sphere is
+   growing as the surface of first contact has not reached the maximum
+   distance.  In (b) it has reached the maximum distance, so that it
+   remains at the same size.  After a Doomsday event, the signals can
+   still be observed, but the surface of last contact grows }
+   \label{F_sphere}
+\end{figure}
+
+\begin{figure}
+   \centering
+   \includegraphics[width=0.5\textwidth]{abcd.pdf}
+   \caption{Schematic representation of two emitters, $E_1$ and $E_2$,
+   that reach each other at different times.  The time span for $E_i$
+   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
+   emitter $j$ between $C_{ij}$ and $B_{ij}$.
+   The type and length of causal contact in both directions depend on
+   the distance and time lag between the awakening events, the maximum
+   distance that a signal can reach and the time period in which each
+   emitter is active.}
+   \label{F_abcd}
+\end{figure}
+
 \begin{figure*}
@@
 
-\section{Results: exploring the parameter space}\label{S_results}
-%{{{
-
-We implemented the simulation of a regular grid of models varying over the
-hypothesis space, which covers XXX different models.
-%
-For each model, we simulated XXX realizations with different random
-seeds, in order to estimate the variance.
-%
-The simulated parameters, $\lambda_1$, $\lambda_2$ and $D_{max}$ cover
-the range XX-XX, XX-XX and XX-XX, respectively, with a partition of XX
-values for each parameter.
-
-As a product of the simulations, several quantities are obtained.
-%
-Some quantities are directly derived from the discrete events, namely,
-the ID of emitting CETI, the ID of emitting (receiving) CETI, the
-position in the galaxy, the time of Awakening (Contact) and the time
-of Doomsday  (Blackout) for each CETI.
-%
-This structure is always for the A-D cycle of each emitting CETI, and
-repeats for each contact (C-B).
-%
-We can also derive the additional quantities.
-%
-The age of a CETI is the time elapsed from A to D.
-%
-%
-%given time, ceti\_e is the CETI emitting signals (emiter), ceti\_r is
-%the CETI receiving signals (receiver), ceti\_c is the ceti that
-%listen at least another ceti (citizen) and ceti\_h is the CETI ceti
-%that is lestened by at least another CETI.
-%
-The time span of a CETI listening another or being listened by another
-can also be easily derived from the results of a realization.
-%
-This way we can also compute the distribution in the galaxy of cetis
-that reach contact and the waiting time until the first contact or the
-waiting time until the next contact.
-%
-Regarding the properties of the population of CETIs and its evolution,
-we analyze the fraction of awaken time a ceti is listening at least
-another ceti, the age of contacted CETIs at first contact, the
-distribution of time to wait until next contact, the fraction of cetis
-where the first contact is given at the awakening, the distribution of
-the number of contacts for each CETI, the distribution of the maximum
-number of contacts for each CETI, the distribution of the number of
-contacts as a function of CETI age, the number of contacts as a
-function of time in the galaxy, the rate of CETIs that succeed in
-contact, and the distribution of distances between contacted CETIs.
-%
-Usefull derived quantity is the duration of two-way communication
-channels the or the fraction of contacts that admit a response.
-% 
-We also analyze the relations between the distance to ceti vs. time of
-double communication, the distance to ceti vs. age of contacted ceti,
-the age of a ceti and the maximum number of contacted cetis before D,
-or the lifespan of a ceti vs. max number of contacts.
-%
-All these quantities can be analyzed as a function of the simulation
-parameters.
-%
-
-We chose eight models that are on fairly oposed regions of the
-explored parameter space, and cover short/long lifetimes, short/long
-waiting times for the next CETI to appear in the Galaxy and
-short/large range reach of the signal, including all possible
-combinations.
-%
-The details of each model are summarized in the
-Table~\ref{T_SelectedModels}.
-
-In Fig.~\ref{F_res_1} we show the Empirical cumulative distributions
-of the number contacts for CETIs in four different samplas (M5, M6,
-M7, M8) with long range signal reach (D$_{max}$=10000). 
-%
-The lifetime is more determinant than the awakening rate in increasing
-the number of contacts.
-%
-As expected, of the four samples, the sample M7, which corresponds to
-a dense awakening in the timeline and a long lifetime, is the model
-that maximizes the number of contacts, reaching a maximum of order 100
-contacts for a single CETI.
-%
-This is considerable larger than any model with a shorter lifetime,
-which produce a number of contacts of at most the order of ten
-contacts per CETI.
-%
-A couple of comments are worth to mention about this plot.
-%
-First, it has a logarithmic scale on the x-axis, and second it is the
-cumulative, not differential, empirical distribution.
-
-In what follows, we focus on the analysis of the duration of contacts
-(Sec. X), the waiting time for the first contact (Sec. X) the
-variations of contact posibilities as a function of the position of
-the CETI in the Galaxy (Sec. X) and the occurrence of reciprocal
-communication (Sec. X).
+Posibles implicaciones
+\begin{itemize}
+   \item Análisis del método de comunicación: isotrópico, colimado, accidental
+   \item Naturaleza del transporte del mensaje: radiación electromagnética u otros
+   \item Efectos de fenómenos de alineación (tránsitos)
+   \item Marcadores energéticos
+   \item Marcadores espirituales
+   \item Uso de estrellas como amplificadores o fuentes
+\end{itemize}
 
 %}}}
-
-\subsection{Duration of contacts}
-%{{{
-
-The duration of each contact for a single receiving CETI is given by the
-time difference between a C event and a B event, which are direct
-outputs for each simulation experiment.
-
-%}}}
-
-\subsection{Waiting time for a first contact}
-%{{{
-
-In this section we analyze the distribution of the waiting times for
-a first contact.
-%
-Such distribution can be considered to compute the probability of
-reaching a first contact within the first XX years.
-
-%}}}
-
-\subsection{Reciprocal communication}
-%{{{
-
-%}}}
-
-\subsection{Location within the Galaxy}
-%{{{
-
-%}}}
-
-\setlength{\tabcolsep}{10pt}
-\begin{table*}
-\centering
-\begin{tabular}{cccccc}
-\hline
-Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & size & description  \\
-M1 & $<$5000 & $<$15000 & $<$15000 &800  &dense awakening and short lifetime\\
-M2 & $<$5000 & $>$20000 & $>$50000 &1050 &sparse awakening and long lifetime\\
-M3 & $<$5000 & $<$5000  & $>$15000 &1350 &dense awakening and long lifetime \\
-M4 & $<$5000 & $>$15000 & $<$26000 &1200 &sparse awakening and short lifetime\\
-
-M5 & $>$20000 & $<$15000 & $<$15000 &800  &dense awakening and short lifetime\\
-M6 & $>$20000 & $>$20000 & $>$50000 &1050 &sparse awakening and long lifetime\\
-M7 & $>$20000 & $<$5000  & $>$15000 &1350 &dense awakening and long lifetime \\
-M8 & $>$20000 & $>$15000 & $<$26000 &1200 &sparse awakening and short lifetime\\
-\hline
-\end{tabular}
-\caption{Selected models to analize the behevior of some variables.}
-\label{T_SelectedModels}
-\end{table*}
-
-\begin{figure} \centering
-   %
-   \includegraphics[width=0.5\textwidth]{Figure_1.png}
-   %
-   \caption{Empirical cumulative distributions of the number contacts
-   for CETIs in four different samplas (M5, M6, M7, M8) with long
-   range signal reach (D$_{max}$=10000).
-   %
-   }
-   % 
-\label{F_res_1} \end{figure}
-
-\begin{figure} % Distribution
-   \centering
-   \includegraphics[width=0.5\textwidth]{Figure_2.png}
-   \caption{Rate of CETIs that make no contact as a function of
-   $\tau_a$ and $\tau_s$}
-   \label{F_res_2}
-\end{figure}
-
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{Figure_3.png}
-   \caption{Distribution of number of contacts as a function of
-   galactocentric distance.}
-   \label{F_res_5}
-\end{figure}
-
-\begin{figure} % 2D color plot
-   \centering
-   \includegraphics[width=0.5\textwidth]{Figure_4.png}
-   \caption{}
-   \label{F_res_4}
-\end{figure}
-
- \begin{figure} % 2D color plot
-   \centering
-   \includegraphics[width=0.5\textwidth]{Figure_5.png}
-   \caption{Rate of CETIs that listen at least one time, and make its
-   first contact at the moment of the awakening, as a funcion of
-   $\tau_a$ and $\tau_s$}
-   \label{F_res_3}
-\end{figure}
-
-\section{Discusion}\label{S_discussion}
-%{{{
-
-We have implemented a simulation approach to explore the hypothesis
-space about the causal connections between communicating
-extraterrestrial civilizations.
-%
-The different models can be generated with a minimal number of free
-parameters and some fixed assumptions.
-%
-We argue that three parameters can be used to describe a variety of
-situations, ranging from a Galaxy where an intelligent civilization is
-very rare, to a Galaxy populated with plenty of civilizations in
-causal contact.
-%
-We explored the consecuences of different models using mensurable
-quantities.
-%
-With this approach, we can estimate several quantities as a function
-of the parameters on the hypothesis space.
-%
-For instance, it is possible to estimate the timescale for an
-uninterrupted SETI effort to reach success, in terms of different
-model parameters that reflect different, so far unknown, scenarios
-for the spread if intelligent life on the Galaxy.
 
@@
 
-Future work can be performed following this framework, in order to
-explore possible implications of the results for more detailed 
-configurations of the experiment.
-%
-For example, the communication method (isotropic, colimated,
-serendipitous) can affect the observables.
-%
-Also the nature of the message carrier, electromagnetic or another.
-%
-It is also possible to explore the effects of alignments or the use of
-stars as sources or amplifiers.
-%
-%energetic markers
-%Spiritual markers
-%
-As a future work, we will explore several improvements to the
-simulation, and analyze the impact on the final results of these
-changes.
-%
-They include a spatial distribution that actually resembles the shape
-of the Galaxy, different distribution functions for the mean lifetime
-of CETIs and different efficiencies for the sphere of the causal
-region of each CETI, resembling different searching strategies.
-
-%}}}
-
-\ack[Acknowledgement]{
-%
-This work was partially supported by the Consejo Nacional de
-Investigaciones Cient\'{\i}ficas y T\'ecnicas (CONICET, Argentina),
-the Secretar\'{\i}a de Ciencia y Tecnolog\'{\i}a, Universidad Nacional
-de C\'ordoba, Argentina, and the Universidad Católica de Córdoba,
-Argentina.
-%
-This research has made use of NASA's Astrophysics Data System.
-%
-Plots and simulations were made with software developed by the authors
-using R and python languajes, plots were postprocessed with inkscape.
-%
-}
-
+\bibliographystyle{mn2e}
 \bibliography{biblio_seti}
-
-%%%%%%%%%%%% Biography text %%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Author Biography}
-
-\aubio{ML.jpeg}{\textbf{Marcelo Lares} received a degree in physics from the University
-of A in 1998 and received his Ph.D. degree in communication engineering in
-2002. He now holds a research chair at the B institute. His main research
-interests are design and optimization of high power microwave power
-amplifiers.}
-
-\aubio{JF.jpeg}{\textbf{Jos\'e Funes} received her Diploma in 1997 from the University of
-C and received her Ph.D. at the D University in 2001. She became a full
-university professor in 2009. She is currently serving as chair of the E
-project and is active in the study of Si power electronics.}
-
 \end{document}
-
-DEVELOPMENT DEVELOPMENT  DEVELOPMENT  DEVELOPMENT  DEVELOPMENT  DEVELOPMENT
-
-% Papers para citar y discutir:
-
-%## GHZ
-
-%## D_max
-
-%## exponential distribution
-
-%## Lifetime
-
-%## Simulations: cosmological
-
-Forgan2015
-
-%## Simulations: stochastic
-
-grimaldi_signal_2017
-
-%## Temporal aspects
-Temporal aspects of the interaction among the first galactic civilizations: The 'interdict hypothesis'
-by M.J. Fogg
-
-%## Drake equation
-
-%## Detalles de la simulacion
-Galactic civilizations: Population dynamics and interstellar diffusion
-by W.I. Newman, C. Sagan

### 2019-07-25 tmanuscript.tex
Initial Overleaf Import

--- 2019-07-25 manuscript.tex
@@
-\documentclass[crop]{CSLB}%%%%where CSL is the template name
+\documentclass[crop]{CSLB}%%%% Modified CSL template tom include Bibliography
 
-%The authors can define any packages after the \documentclass{CSL} command.
 \usepackage{amsmath} %for dealing with mathematics,
@@
 
-%\usepackage{natbib}
-\usepackage{boites}
-
-%\usepackage{graphicx}
-%\usepackage{float}
-%\usepackage{psfig}
-%\usepackage{amsfonts,amssymb}
-%\usepackage{color}
-%\usepackage{textcomp}
-%\usepackage[T1]{fontenc}
-%\usepackage{aecompl}
-%\usepackage{times}
-%\usepackage{txfonts}
-%\usepackage[english]{babel}
-%\usepackage[utf8x]{inputenc}
-%\usepackage{amsmath}
-%\usepackage{caption}
-%\usepackage{siunitx}
-
 \jname{International Journal of Astrobiology}%
@@
 
-\bibliographystyle{mn2e}
-
 \supertitle{Research Paper}
@@
 
-\author[Lares \& Funes]{Lares M$^{1, 2}$, Funes J$^{1, 3}$}
+\author[verso running head]{Lares M$^{1, 2}$, Funes J$^{1, 3}$}
 
@@
 and their contacts have also been explored by several authors
-\citep{Fogg1987, Forgan2011, Balbi2018},
+\citep{Fogg1987, Forgan2011, Balb2018},
 %The Impact of the Temporal Distribution of Communicating Civilizations on Their Detectability
@@
 
-In a recent work, \citep{Balbi2018} use a statistical model to analyze
+In a recent work, \citep{Balb2018} use a statistical model to analyze
 the occurrence of causal contacts between civilizations in the Galaxy.
@@
 
-\section{Methods: accounting for causal contacts with 
-discrete event simulation process}\label{S_methods}
-%{{{
-
-Some complex stochastic processes can be efficiently modeled with the
-discrete--event (DE) simulation approach.
-%
-Simulations are suitable tools to analyze systems that evolve with
-time and involve randomness.
-%
-In general, simulations require less assumptions and simplifications
-than theoretical approaches, and can be applied to systems where a
-theoretical model can not be found.
-%
-A system described with the DE paradigm is characterized by a set of
-actors and events, where actors interact causally through a series of
-events on a time line and process these events in chronological order
-\citep{ptolemaeus_system_2014, chung_simulation_2003,
-simulation_ross_2012}.
-%
-On the particular case of the difusion of signals of intelligence in
-the Galaxy, the discrete events method can be performed taking into
-account a small number of variables.
-%
-The temporal structure is defined by two distribution parameters that
-represent the mean time interval that an intelligent civilization can
-emit and receive signals, and the mean time interval between the
-emergence of an intelligent comunicating civilization and the next
-one.
-%
-The spatial structure of the simulation is given by the size and shape
-of the Galactic Habitable Zone and the maximum distance a signal can
-travel to be detected.
-%
-The parameters for the temporal distributions also determine the
-spatial properties.
-%
-For example, the density of active cetis in the Galaxy depend on these
-two parameters.
-%
-Also, some hypothesis must be made in order to complete the
-simulation.
-%
-Firstly, we assume that the distribution of the appeareance of new
-CETIs is a stationary Poisson distribution.
-%
-Regarding the duration of a CETI, we propose that the distribution of
-the duration of CETIs is a stationary exponential distribution.
-%
-\citet{Balbi2018} stress the fact that it would be desirable to arrive
-at a theoretical statistical distribution of the lifespan of
-civilizations, preferentially on the basis of the underlying
-astrophysical and biological processes.
-%
-\citet{Maccone} argues that this distribution should be a log normal.
-(EXPAND)
-
-There is no a temporal window for the model, i.e., the simulation
-starts when the process is already stable, and ends before any
-galactic effect could modify or affect of fixed variables.
-%
-The probability for the rise of a CETI is homogeneous over the GHZ.
-%
-Although the Galaxy has a well known spiral structure, the assumed
-sparcity of civilizations makes this assumption very reasonably.
-%
-Otherwise, if the distribution of civilizations is not sparce, it
-could be the case that the spiral arms would host most of the
-civilizations, and contacts are frecuent between closely located
-civilizations.
-%
-In such a case we could expect that the interpretation of our results
-could be considered as pesimist for the stablishment of a contact with
-our planet.
-%
-We also limit the posibilities of life or other types of civilizations
-to the hypothesis usually stated for the definition of the GHZ.
-%
-This means that we set aside possibel civilizations that could survive
-in severe conditions or unstable systems, which would prevent the
-appeareance of life as we know it.
-
-\setlength{\tabcolsep}{10pt}
-\begin{table}
-\centering
-\begin{tabular}{cl}
-\hline
-   $\tau_{a}$ & Mean temporal separation between consecutive awakenings \\
-   $\tau_{s}$ & Mean lifetime of a CETI \\
-   D$_{max}$ & Maximum reach of a message \\
-\hline
-\end{tabular}
-\caption{Definition of variables that are part of the simulation.}
-\label{T_simu}
-\end{table}
-
-While the aforementioned assumptions are basic conditions for most of
-the stochastic processes observed in nature, 
-we need to make stronger assumptions related to the nature of the
-message.
-%
-The most simple assumption about the message itself, is that it
-travels at light speed.
-%
-With this, we are considering messages sent through, for example,
-electromagnetic radiation or even gravitational radiation, but we set
-appart messages sent with mechanical means or physical objects, or
-through some unknown technology that violates the known so far laws of
-physics.
-% 
-For the communication of CETIs through messages sent isotropically, 
-we assume that the capacity to emit signals and to receive
-signals ocurr at the same time.
-%
-This means that the ability to find another civilization develops at
-the same time that the ability and intention to emit a signal in all
-direction, intented to be detected by an unknown civilization.
-%
-Although there are several reasons to think that this could not be the
-case, at large time scales it can be considered that both abilities
-occurr roughly at coincident epochs.
-%
-Another essential assumption is that all CETIs use the same signal power, so that there is a maximum distance out to which it can be detected.
-%
-It is worth mentioning that we are considering in our experimental
-setup a system composed by several emmiters and receivers across the
-Galaxy, under the hypothesis proposed before.
-%
-In such a system we compute probabilities of a CETI making contact
-with another CETI, and the results must be interpreted as for a
-randomly chosen CETI, given certain conditions of homogeneityon an
-ensemble of equivalent galaxies.
-%
-We stress the fact that we are not centered on the probabilities for
-the Earth itselft, but on a generalized CETI.
-%
-Regarding the extent of the signals, we know that the distance from
-which a signal from Earth could be detected using the current
-technology is of a few hundreds light years, given that the signal was
-sent to a specific direction.
-
-%https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
-%https://www.engadget.com/2017/05/25/listening-to-starlight-our-ongoing-search-for-alien-intelligenc/]
-%[from https://en.wikipedia.org/wiki/Fermi_paradox]
-%
-It is straightforward to propose and implement a distribution of
-maximum distances, although this would increase the model complexity
-at the cost of a larger uncertainty or of another dimension on the
-hypothesis space.
-%
-This proposal is based on a pronoid scenario (REF), and we do not
-consider in this work the ocurrence of paranoid or partially paranoid
-civilizations.
-%
-If that would be the case, the results obtained here can be taken as
-upper limits.
-
-In our simulations we also assume the simplest situations for the
-growth of ETIs.
-%
-For example, we discard the possibility of stellar colonization
-(Walters 1980) and assume that the communication aim is performed to
-all directions with the same power and the same probability, i.e.,
-strongly assume isotropic communication in all cases.
-%
-More detailed simulations could be produced considering different 
-efficiency of communication or detection methods.
-%
-Also, a mixed paranoid/pronoid population of intelligent civilizations
-could be implemented easily.
-%
-However, in this case we also would complicate the experimental setup,
-make the results less evident.
-%
-The results are independent of the nature of life (organic or
-artificial).
-%
-The lifetime of a civilization can be caused by autodestruction or by
-external factors.
-
-\subsection{Power laws vs. exponential laws}
-
-The exponential distribution of lifespan and waiting times is
-justified by considering that the process of appeareance of life in
-the galaxy is homogeneous and stationary.
-%
-This means that there is no preferred location within the GHZ for the
-spontaneos appeareance of life, and that the emergence of a
-civilization is independent of the existence of previous civilizations 
-in the galaxy.
-%            
-This is equivalent to proposing a Poisson process for the emergence of
-CETIs, since there is a close relation between the number of events in
-time or space and the waiting time or separation, respectively.
-%
-That is, these are two alternative approaches to describing the same
-process, a Poisson distribution for the number of events implies an
-exponential distribution for their sepations, and viceversa.
-%
-It should be emphasized that the exponential laws used in this work
-are assumed, although instead of analyzing results from a particular
-parameter chosen ad--hoc, we explore the hypothesis space and analyze
-the impact of the values of these parameters on the results.
-%
-
-The power law and exponential statistical distributions are among the
-most common patterns found in natural phenomena.
-%
-For example, the distribution of the frecuency of words in many
-languajes is known to follow a Zipf law (which is a power law).
-%
-Zipf law also describes population ranks of cities in various
-countries, corporation sizes, income rankings, ranks of number of
-people watching the same TV channel (Zipf's law entry on Wikipedia).
-%
-The magnitudes of earthquakes, hurricanes, volcanic eruptions and
-floods; the sizes of meteorites or the losses caused by business
-interruptions from accidents, are also well described by power laws
-(Sornette 2006).
-%
-The power law behaviour has been observed in a variety of systems,
-including for example stock market fluctuations, sizes of computer
-files and word frequency in languages \citep{mitzenmacher_brief_2004,
-Newman_power_2005, simkin_theory_2006}
-%
-Power laws have also been widely used in biological sciences, e.g., in
-analysing connectivity patterns in metabolic networks
-\citep{jeong_large_2000} and in the number of species observed per
-unit area in ecology \citep{garcia_origin_2006, frank_common_2009}.
-
-\subsection{Age distribution of civilizations}
-
-\subsection{Rate of appeareance of civilizations}
-
-\subsection{Exploration of the hypothesis space}
-
-Zipf law.
-
-The adopted values for the GHZ are XX and XX \citep{Lineweaver2004}.
-
-\setlength{\tabcolsep}{10pt}
-\begin{table}
-\centering
-\begin{tabular}{cccc}
-\hline
-   fixed parameter & \multicolumn{3}{c}{adopted value}\hfill \\
-\hline
-   inner GHZ radius\hfill & \multicolumn{3}{c}{20000 ly}\\
-   outer GHZ radius & \multicolumn{3}{c}{60000 ly}\\ \\
-\hline
-   variable parameter & min & max & N \\
-\hline
-   D$_{max}$ & 1000 ly & 20000 ly & 20\\
-   $\tau_{a}$ & 1000 y & 20000 y & 20\\
-   $\tau_{s}$ & 1000 y & 20000 y & 20\\
-\hline
-\end{tabular}
-\caption{Adopted values for simulation parameters. Fixed parameters
-   define the GHZ.  Variable parameters define the temporal structure
-   of the process and the reach of messages.}
-\label{T_parameters}
-\end{table}
-
-The main variables that follow the evolution of the simulation are:
-
-\begin{itemize}
-   \item Position of stars.  Sampled randomly within the GHZ
-   \item Time of the awakening of each CETI
-   \item Time of the vanishing of each CETI
-\end{itemize}
-
-The variables that can be deduced from the previous ones are:
-
-\begin{itemize}
-   \item Number of CETIs in casual contact with at least another CETI
-      at a given time.
-   \item Number of CETIs as a function of time
-   \item Number of CETIs that receive a message at least one time
-   \item Number of CETIs that receive a message at least one time and
-      successfully deliver an answer.
-   \item Number distribution of waiting time to receive a message
-\end{itemize}
-
-%========================================================================
-%========================================================================
-%========================================================================
-
 \begin{figure}
@@
 \end{figure}
-
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{abcd.pdf}
-   \caption{Schematic representation of two emitters, $E_1$ and $E_2$,
-   that reach each other at different times.  The time span for $E_i$
-   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
-   emitter $j$ between $C_{ij}$ and $B_{ij}$.
-   The type and length of causal contact in both directions depend on
-   the distance and time lag between the awakening events, the maximum
-   distance that a signal can reach and the time period in which each
-   emitter is active.}
-   \label{F_abcd}
-\end{figure}
-
-Possible hypothesis improvements are:
-
-\begin{itemize}
-   \item Dmax is different for different CETIs.  For example a power
-      law where powerfull emmisions are rare and weak emssions are
-      common.
-   \item The probability of the rising of a new CETI vary within the
-      GHZ.
-   \item Correction by coverage ratio in the detection and by
-      targetting ratio in the emission.
-   \item Spiritual factor.  Explore the role of the S factor: message
-      contents, influence on the lifespan of a CETI that receives a
-      message, mean lifespan of a CETI that emits a message.
-\end{itemize}
-
-\subsection{Model for CETIs}
-
-Since the method is based on events, the first step is to define an
-architecture of events, the relationships between them and how events
-trigger changes on the state of the system.
-%
-We consider a simulated system that represents the galactic habitable
-zone (GHZ).
-%
-On a first approach, the GHZ is a 2-dimensional annular region.
-%
-This simple model does not take into accont the variations in stellar
-density given by the spiral structure.
-%
-Although there are several possible approaches, we chose to follow the
-evolution of the system according to the following events:
-
-\begin{enumerate}
-   \item[(A)] A new CETI appears (Awakening)
-   \item[(D)] An old CETI dissappears (Doomsday)
-   \item[(C)] A new causal contact is stablished (Contact)
-   \item[(B)] An existing causal contact is interrupted (Blackout)
-\end{enumerate}
-
-The event of type B is produced when one of the two CETIs halts in its
-capability of emmiting and receiving signals.
-
-%======================================================================
-%======================================================================
-%======================================================================
-
-The system is updated each time an event is produced.
-%
-The time marks for each event (A and D for each CETI and C and B for
-each causal contact) are stored as a result of the simulation.
-%
-Also, the list of active CETIs is obtained as a function of time.
-%
-Some fixed variables must be set in order to carry out the simulation,
-namely:
-
-\begin{itemize}
-   \item Size of the Galactic Habitable Zone,
-   \item mean lifetime of a CETI, $\lambda_1$,
-   \item mean waiting time for the appeareance of another CETI,
-      $\lambda_2$, and
-   \item maximum distance a message can be detected.
-\end{itemize}
-
-The Galactic Habitable Zone is the only one of the four parameters
-that is barely known, is set at fixed values of the inner and
-outer radii, radial symetry is assumed, and the with of the
-galactic disk is neglected compared to the radial size.
-%
-The two radii are measured in light years, with the aim to maintain a
-single and comprehensive unit for both space and time coordinates.
-
-The simulation was implemented on a python code, which is publically
-available\footnote{https://github.com/mlares/simu\_contact}
-
-%Variables en el programa
-%\begin{itemize}
-%   \item Nact: Número de CETIs activas
-%   \item Nsee: Número de contactos por cada CETI
-%   \item Nccc: CETIs en contacto causal. Notar que una CETI puede tener múltiples
-%contactos de otras CETIs. Para cada una de ellas, guardar:
-%   \item ID: ID de la CETI emisora
-%   \item ID: ID de la CETI receptora
-%   \item t\_hola: tiempo del primer contacto
-%   \item t\_chau: tiempo del último contacto
-%\end{itemize}
-%
-%
-%Caso 1: aparece una nueva CETI
-%Modificaciones al estado del sistema:
-%\begin{itemize}
-%   \item Agregar la CETI “A” a la lista de CETIs activas
-%   \item Sortear la duración de la actividad de comunicación de la CETI “A”
-%   \item Sortear el intervalo de tiempo hasta la aparición de la próxima CETI:
-%t\_WakeUp\_next
-%   \item Agregar la CETI “A” al árbol
-%   \item Identificar todas las CETIs que están a menos de D max de la nueva CETI, y
-%actualizar la lista de tiempos
-%\end{itemize}
-%
-%Caso 2: desaparece una CETI
-%Modificaciones al estado del sistema:
-%\begin{itemize}
-%   \item Eliminar la CETI a la lista de CETIs activas
-%\end{itemize}
-%
-%
-%Caso 3: una nueva CETI entra en contacto causal
-%Modificaciones al estado del sistema:
-%\begin{itemize}
-%   \item Agregar la CETI B a la lista de contactos de la CETI A
-%\end{itemize}
-%
-
-%}}}
-
-\begin{figure*}
-   \centering
-   \includegraphics[width=\textwidth]{Messages_01.pdf}
-   \caption{Schematic representation of the possible cases in which a
-   ceti A can be in causal contact with a ceti B which appears later.
-   The duration of the causal contact in one direction depends on
-   several factors, mainly $\Delta t_A$, $\Delta t_C$, and $\Delta X$.}
-   \label{F_messages}
-\end{figure*}
-
-\section{Results: exploring the parameter space}\label{S_results}
-%{{{
-
-We implemented the simulation of a regular grid of models varying over the
-hypothesis space, which covers XXX different models.
-%
-For each model, we simulated XXX realizations with different random
-seeds, in order to estimate the variance.
-%
-The simulated parameters, $\lambda_1$, $\lambda_2$ and $D_{max}$ cover
-the range XX-XX, XX-XX and XX-XX, respectively, with a partition of XX
-values for each parameter.
-
-As a product of the simulations, several quantities are obtained.
-%
-Some quantities are directly derived from the discrete events, namely,
-the ID of emitting CETI, the ID of emitting (receiving) CETI, the
-position in the galaxy, the time of Awakening (Contact) and the time
-of Doomsday  (Blackout) for each CETI.
-%
-This structure is always for the A-D cycle of each emitting CETI, and
-repeats for each contact (C-B).
-%
-We can also derive the additional quantities.
-%
-The age of a CETI is the time elapsed from A to D.
-%
-%
-%given time, ceti\_e is the CETI emitting signals (emiter), ceti\_r is
-%the CETI receiving signals (receiver), ceti\_c is the ceti that
-%listen at least another ceti (citizen) and ceti\_h is the CETI ceti
-%that is lestened by at least another CETI.
-%
-The time span of a CETI listening another or being listened by another
-can also be easily derived from the results of a realization.
-%
-This way we can also compute the distribution in the galaxy of cetis
-that reach contact and the waiting time until the first contact or the
-waiting time until the next contact.
-%
-Regarding the properties of the population of CETIs and its evolution,
-we analyze the fraction of awaken time a ceti is listening at least
-another ceti, the age of contacted CETIs at first contact, the
-distribution of time to wait until next contact, the fraction of cetis
-where the first contact is given at the awakening, the distribution of
-the number of contacts for each CETI, the distribution of the maximum
-number of contacts for each CETI, the distribution of the number of
-contacts as a function of CETI age, the number of contacts as a
-function of time in the galaxy, the rate of CETIs that succeed in
-contact, and the distribution of distances between contacted CETIs.
-%
-Usefull derived quantity is the duration of two-way communication
-channels the or the fraction of contacts that admit a response.
-% 
-We also analyze the relations between the distance to ceti vs. time of
-double communication, the distance to ceti vs. age of contacted ceti,
-the age of a ceti and the maximum number of contacted cetis before D,
-or the lifespan of a ceti vs. max number of contacts.
-%
-All these quantities can be analyzed as a function of the simulation
-parameters.
-%
-
-We chose eight models that are on fairly oposed regions of the
-explored parameter space, and cover short/long lifetimes, short/long
-waiting times for the next CETI to appear in the Galaxy and
-short/large range reach of the signal, including all possible
-combinations.
-%
-The details of each model are summarized in the
-Table~\ref{T_SelectedModels}.
-
-In Fig.~\ref{F_res_1} we show the Empirical cumulative distributions
-of the number contacts for CETIs in four different samplas (M5, M6,
-M7, M8) with long range signal reach (D$_{max}$=10000). 
-%
-The lifetime is more determinant than the awakening rate in increasing
-the number of contacts.
-%
-As expected, of the four samples, the sample M7, which corresponds to
-a dense awakening in the timeline and a long lifetime, is the model
-that maximizes the number of contacts, reaching a maximum of order 100
-contacts for a single CETI.
-%
-This is considerable larger than any model with a shorter lifetime,
-which produce a number of contacts of at most the order of ten
-contacts per CETI.
-%
-A couple of comments are worth to mention about this plot.
-%
-First, it has a logarithmic scale on the x-axis, and second it is the
-cumulative, not differential, empirical distribution.
-
-In what follows, we focus on the analysis of the duration of contacts
-(Sec. X), the waiting time for the first contact (Sec. X) the
-variations of contact posibilities as a function of the position of
-the CETI in the Galaxy (Sec. X) and the occurrence of reciprocal
-communication (Sec. X).
-
-%}}}
-
-\subsection{Duration of contacts}
-%{{{
-
-The duration of each contact for a single receiving CETI is given by the
-time difference between a C event and a B event, which are direct
-outputs for each simulation experiment.
-
-%}}}
-
-\subsection{Waiting time for a first contact}
-%{{{
-
-In this section we analyze the distribution of the waiting times for
-a first contact.
-%
-Such distribution can be considered to compute the probability of
-reaching a first contact within the first XX years.
-
-%}}}
-
-\subsection{Reciprocal communication}
-%{{{
-
-%}}}
-
-\subsection{Location within the Galaxy}
-%{{{
-
-%}}}
-
-\setlength{\tabcolsep}{10pt}
-\begin{table*}
-\centering
-\begin{tabular}{cccccc}
-\hline
-Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & size & description  \\
-M1 & $<$5000 & $<$15000 & $<$15000 &800  &dense awakening and short lifetime\\
-M2 & $<$5000 & $>$20000 & $>$50000 &1050 &sparse awakening and long lifetime\\
-M3 & $<$5000 & $<$5000  & $>$15000 &1350 &dense awakening and long lifetime \\
-M4 & $<$5000 & $>$15000 & $<$26000 &1200 &sparse awakening and short lifetime\\
-
-M5 & $>$20000 & $<$15000 & $<$15000 &800  &dense awakening and short lifetime\\
-M6 & $>$20000 & $>$20000 & $>$50000 &1050 &sparse awakening and long lifetime\\
-M7 & $>$20000 & $<$5000  & $>$15000 &1350 &dense awakening and long lifetime \\
-M8 & $>$20000 & $>$15000 & $<$26000 &1200 &sparse awakening and short lifetime\\
-\hline
-\end{tabular}
-\caption{Selected models to analize the behevior of some variables.}
-\label{T_SelectedModels}
-\end{table*}
-
-\begin{figure} \centering
-   %
-   \includegraphics[width=0.5\textwidth]{Figure_1.png}
-   %
-   \caption{Empirical cumulative distributions of the number contacts
-   for CETIs in four different samplas (M5, M6, M7, M8) with long
-   range signal reach (D$_{max}$=10000).
-   %
-   }
-   % 
-\label{F_res_1} \end{figure}
-
-\begin{figure} % Distribution
-   \centering
-   \includegraphics[width=0.5\textwidth]{Figure_2.png}
-   \caption{Rate of CETIs that make no contact as a function of
-   $\tau_a$ and $\tau_s$}
-   \label{F_res_2}
-\end{figure}
-
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{Figure_3.png}
-   \caption{Distribution of number of contacts as a function of
-   galactocentric distance.}
-   \label{F_res_5}
-\end{figure}
-
-\begin{figure} % 2D color plot
-   \centering
-   \includegraphics[width=0.5\textwidth]{Figure_4.png}
-   \caption{}
-   \label{F_res_4}
-\end{figure}
-
- \begin{figure} % 2D color plot
-   \centering
-   \includegraphics[width=0.5\textwidth]{Figure_5.png}
-   \caption{Rate of CETIs that listen at least one time, and make its
-   first contact at the moment of the awakening, as a funcion of
-   $\tau_a$ and $\tau_s$}
-   \label{F_res_3}
-\end{figure}
-
-\section{Discusion}\label{S_discussion}
-%{{{
-
-We have implemented a simulation approach to explore the hypothesis
-space about the causal connections between communicating
-extraterrestrial civilizations.
-%
-The different models can be generated with a minimal number of free
-parameters and some fixed assumptions.
-%
-We argue that three parameters can be used to describe a variety of
-situations, ranging from a Galaxy where an intelligent civilization is
-very rare, to a Galaxy populated with plenty of civilizations in
-causal contact.
-%
-We explored the consecuences of different models using mensurable
-quantities.
-%
-With this approach, we can estimate several quantities as a function
-of the parameters on the hypothesis space.
-%
-For instance, it is possible to estimate the timescale for an
-uninterrupted SETI effort to reach success, in terms of different
-model parameters that reflect different, so far unknown, scenarios
-for the spread if intelligent life on the Galaxy.
-
-Under the hypotheses of our experiments, we conclude that
-a causal contact is extremely unlikely unless the Galaxy is heavily
-populated by intelligent civilizations.
-%
-Even so, we must consider that in order to stablish a contact between
-any two entities, a minimum degree of compatibility must be
-accomplished without any previous agreement, 
-making the posibility of a contact with a message that
-could be deciphered highly rare.
-
-Future work can be performed following this framework, in order to
-explore possible implications of the results for more detailed 
-configurations of the experiment.
-%
-For example, the communication method (isotropic, colimated,
-serendipitous) can affect the observables.
-%
-Also the nature of the message carrier, electromagnetic or another.
-%
-It is also possible to explore the effects of alignments or the use of
-stars as sources or amplifiers.
-%
-%energetic markers
-%Spiritual markers
-%
-As a future work, we will explore several improvements to the
-simulation, and analyze the impact on the final results of these
-changes.
-%
-They include a spatial distribution that actually resembles the shape
-of the Galaxy, different distribution functions for the mean lifetime
-of CETIs and different efficiencies for the sphere of the causal
-region of each CETI, resembling different searching strategies.
-
-%}}}
 
@@
 
+\bibliographystyle{mn2e}
 \bibliography{biblio_seti}
 
-%%%%%%%%%%%% Biography text %%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Author Biography}
-
-\aubio{ML.jpeg}{\textbf{Marcelo Lares} received a degree in physics from the University
-of A in 1998 and received his Ph.D. degree in communication engineering in
-2002. He now holds a research chair at the B institute. His main research
-interests are design and optimization of high power microwave power
-amplifiers.}
-
-\aubio{JF.jpeg}{\textbf{Jos\'e Funes} received her Diploma in 1997 from the University of
-C and received her Ph.D. at the D University in 2001. She became a full
-university professor in 2009. She is currently serving as chair of the E
-project and is active in the study of Si power electronics.}
-
 \end{document}
-
-DEVELOPMENT DEVELOPMENT  DEVELOPMENT  DEVELOPMENT  DEVELOPMENT  DEVELOPMENT
-
-% Papers para citar y discutir:
-
-%## GHZ
-
-%## D_max
-
-%## exponential distribution
-
-%## Lifetime
-
-%## Simulations: cosmological
-
-Forgan2015
-
-%## Simulations: stochastic
-
-grimaldi_signal_2017
-
-%## Temporal aspects
-Temporal aspects of the interaction among the first galactic civilizations: The 'interdict hypothesis'
-by M.J. Fogg
-
-%## Drake equation
-
-%## Detalles de la simulacion
-Galactic civilizations: Population dynamics and interstellar diffusion
-by W.I. Newman, C. Sagan

### 2019-07-30 manuscript.tex
github connection test
tratando de cambiar el estilo de latex

@@
 
-\jname{International Journal of Astrobiology}%
+\jname{Astrobiology}%
 
@@
 
-\received{xx xxxx xxxx}
-\revised{xx xxxx xxxx}
-\accepted{xx xxxx xxxx}
-
 \keywords{SETI, Computer simulations, Statistics}
@@
 equation.
+
 %
@@
 
-%%%%%%%%%%%% Biography text %%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Author Biography}
-
-\aubio{ML.jpeg}{\textbf{Marcelo Lares} received a degree in physics from the University
-of A in 1998 and received his Ph.D. degree in communication engineering in
-2002. He now holds a research chair at the B institute. His main research
-interests are design and optimization of high power microwave power
-amplifiers.}
-
-\aubio{JF.jpeg}{\textbf{Jos\'e Funes} received her Diploma in 1997 from the University of
-C and received her Ph.D. at the D University in 2001. She became a full
-university professor in 2009. She is currently serving as chair of the E
-project and is active in the study of Si power electronics.}
-
 \end{document}

### 2019-07-30 manuscript.tex
secciones con numeros

@@
 
-\section{Motivations: Uncertainties and the non detection problem for
-ETI contact}
+\section{Motivations}
+%\section{Motivations: Uncertainties and the non detection problem for ETI contact}
 %{{{
@@
 %New numerical determination of habitability in the Galaxy: the SETI connection
-
+%
 Since the factors in the Drake equation are uncertain, we propose to
@@
 
-\section{Methods: accounting for causal contacts with 
-discrete event simulation process}\label{S_methods}
+\section{Methods and working hypotheses}
+
+%: accounting for causal contacts with 
+%discrete event simulation process}\label{S_methods}
 %{{{
@@
 
-\subsection{Exploration of the hypothesis space}
+\subsection{Discrete event process}
 
-Zipf law.
-
-The adopted values for the GHZ are XX and XX \citep{Lineweaver2004}.
+Given a set of parameters $(\tau_a, \tau_s, D_{max})$, the functional
+forms of the statistical distributions and fixed values
+for the hypotheses, a discrete
+event simulation is performed by keeping track of a set of variables
+that change each time a meaningful event is produced.
 
@@
 \end{itemize}
+
+\subsection{Exploration of the hypothesis space}
 
@@
 On a first approach, the GHZ is a 2-dimensional annular region.
+%
+The adopted values for the GHZ are XX and XX \citep{Lineweaver2004}.
 %

### 2019-07-31 manuscript.tex
un poco mas de texto

@@
 equation.
-
 %
@@
 
-%Our results highlight the necessity, which has been noted
-%elsewhere (Ćirković, 2004), of modeling 
-%
-%
 %While the Drake equation remains a very useful
@@
 %
-On the particular case of the difusion of signals of intelligence in
-the Galaxy, the discrete events method can be performed taking into
-account a small number of variables.
+This method is well suited for the 
+particular case of the difusion of intelligence signals in
+the Galaxy.
 %
-The temporal structure is defined by two distribution parameters that
+Assuming some simple hypotheses, 
+the discrete events method can be performed taking into
+account a small number of variables, which allows to analyze the
+results as a function of those variables.
+
+In what follows, we describe the experimental setup chosen to estimate
+the probabilities of causal contacts and several derived quantities in
+terms of three independent parameters, namely, the mean time span
+between the appeareance of consecutive communicating civilizations
+($\tau_a$), their mean lifetime ($\tau_s$), and the maximum distance a
+signal can be detected by another civilization ($D_{max}$).
+%
+Intuitively, the shorter $\tau_a$ and the larger $\tau_s$ and
+$D_{max}$, there is a greater probability of the existence of causal
+contacts between pairs of intelligent civilizations in the Galaxy
+(Table \ref{T_simu}).
+%
+The causal connection can be produced without the need of the two
+civilizations being concurrent or active at the same time.
+%
+Each CETI will fill an annular region of the Galaxy, limited by two
+surfaces.
+%
+The leading front, or surface of first contact (SFC) grows from the
+central CETI until it reaches the distance $D_{max}$.
+%
+The trailing front, or surface of last contact (SLC) also grows from
+the central CETI, with a delay with respect to the SFC equivalent to
+the lifetime of the CETI, and produces an annular region that shrinks
+until it reaches $D_{max}$, when all signal from the central CETI is
+lost.
+%
+In the 2D simulation only the intersection of these spheres with the
+plane of the galaxy is relevant, and produce the corresponding
+circumferences.
+
+The temporal structure of the process is defined by two distribution parameters that
 represent the mean time interval that an intelligent civilization can
@@
 The parameters for the temporal distributions also determine the
-spatial properties.
-%
-For example, the density of active cetis in the Galaxy depend on these
+spatial properties since the density of active cetis in the Galaxy depend on these
 two parameters.
 %
-Also, some hypothesis must be made in order to complete the
+For example, a small $\tau_a$ and a large $\tau_s$ will produce a
+Galaxy densely populated woth CETIs, and viceversa.
+%
+Also, some hypotheses must be made in order to complete the
 simulation.
@@
 \hline
-   $\tau_{a}$ & Mean temporal separation between consecutive awakenings \\
+   \multicolumn{2}{l}{independent variables} \\
+\hline
+   $\tau_{a}$ & Mean temporal separation between\\
+              & consecutive awakenings \\
    $\tau_{s}$ & Mean lifetime of a CETI \\
@@
 \hline
+\multicolumn{2}{l}{fixed variables and assumptions} \\
+\hline
+   & All CETIs share the same statistical properties\\
+   & The process is homogeneous \\
+   & The scan of the sly is fully efficient\\
+   & Absence of panspermia or colonization \\
+   & The GHZ is a 2--dimensional ring \\
+20000ly    & Inner radius of the GHZ \\
+60000ly    & Outer radius of the GHZ \\
+1.e6 y    & Time span of the simulation \\
+
+\hline
 \end{tabular}
-\caption{Definition of variables that are part of the simulation.}
+\caption{Definition of independent variables and fixed paramenters 
+   that are part of the simulation.}
 \label{T_simu}
@@
 unit area in ecology \citep{garcia_origin_2006, frank_common_2009}.
-
-\subsection{Age distribution of civilizations}
-
-\subsection{Rate of appeareance of civilizations}
 

### 2019-08-08 manuscript.tex
analisis del waiting time

@@
 \centering
-\begin{tabular}{cccccc}
-\hline
-Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & size & description  \\
-M1 & $<$5000 & $<$15000 & $<$15000 &800  &dense awakening and short lifetime\\
-M2 & $<$5000 & $>$20000 & $>$50000 &1050 &sparse awakening and long lifetime\\
-M3 & $<$5000 & $<$5000  & $>$15000 &1350 &dense awakening and long lifetime \\
-M4 & $<$5000 & $>$15000 & $<$26000 &1200 &sparse awakening and short lifetime\\
-
-M5 & $>$20000 & $<$15000 & $<$15000 &800  &dense awakening and short lifetime\\
-M6 & $>$20000 & $>$20000 & $>$50000 &1050 &sparse awakening and long lifetime\\
-M7 & $>$20000 & $<$5000  & $>$15000 &1350 &dense awakening and long lifetime \\
-M8 & $>$20000 & $>$15000 & $<$26000 &1200 &sparse awakening and short lifetime\\
+\begin{tabular}{ccccc}
+\hline
+Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & description  \\
+\hline
+M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
+M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
+M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
+M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
+
+M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
+M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
+M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
+M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
+
 \hline
@@
    %
-   \includegraphics[width=0.5\textwidth]{Figure_1.png}
+   \includegraphics[width=0.5\textwidth]{F1.pdf}
    %
@@
    \centering
-   \includegraphics[width=0.5\textwidth]{Figure_2.png}
+   \includegraphics[width=0.5\textwidth]{F2.pdf}
    \caption{Rate of CETIs that make no contact as a function of
@@
 
+\begin{figure*} % 2D color plot
+   \centering
+   \includegraphics[width=0.3\textwidth]{m1d2.png}%
+   \includegraphics[width=0.3\textwidth]{m1d3.png}%
+   \includegraphics[width=0.3\textwidth]{m1d4.png}
+   \caption{Rate of CETIs that never make contact (listening) as a
+   function of $\tau_a$ and $\tau_s$, for 
+   $D_{max}$=10000 (left panel),
+   $D_{max}$=40000 (middle panel), and
+   $D_{max}$=80000 (right panel)}
+   \label{F_res_3}
+\end{figure*}
+
+\begin{figure*} % 2D color plot
+   \centering
+   \includegraphics[width=0.3\textwidth]{m2d2.png}%
+   \includegraphics[width=0.3\textwidth]{m2d3.png}%
+   \includegraphics[width=0.3\textwidth]{m2d4.png}
+   \caption{Rate of CETIs that listen at the moment of the
+   awakening, as a funcion of $\tau_a$ and $\tau_s$}
+   \label{F_res_4}
+\end{figure*}
+
 \begin{figure}
    \centering
-   \includegraphics[width=0.5\textwidth]{Figure_3.png}
-   \caption{Distribution of number of contacts as a function of
+   \includegraphics[width=0.5\textwidth]{radial.png}
+   \caption{Fraction of CETIs making cero, one, two or three contacts as a function of
    galactocentric distance.}
    \label{F_res_5}
-\end{figure}
-
-\begin{figure} % 2D color plot
-   \centering
-   \includegraphics[width=0.5\textwidth]{Figure_4.png}
-   \caption{}
-   \label{F_res_4}
-\end{figure}
-
- \begin{figure} % 2D color plot
-   \centering
-   \includegraphics[width=0.5\textwidth]{Figure_5.png}
-   \caption{Rate of CETIs that listen at least one time, and make its
-   first contact at the moment of the awakening, as a funcion of
-   $\tau_a$ and $\tau_s$}
-   \label{F_res_3}
 \end{figure}
@@
 
+\begin{figure*}
+   \centering
+   \includegraphics[width=0.33\textwidth]{waiting_a_acum_log.png}%
+   \includegraphics[width=0.33\textwidth]{waiting_a_acum_lin.png}%
+   \includegraphics[width=0.33\textwidth]{waiting_a_dif.png}
+
+   \includegraphics[width=0.33\textwidth]{waiting_s_acum_log.png}%
+   \includegraphics[width=0.33\textwidth]{waiting_s_acum_lin.png}%
+   \includegraphics[width=0.33\textwidth]{waiting_s_dif.png}
+
+   \includegraphics[width=0.33\textwidth]{prob1_10k.png}%
+   \includegraphics[width=0.33\textwidth]{prob1_40k.png}%
+   \includegraphics[width=0.33\textwidth]{prob1_80k.png}
+
+   \caption{Fraction of the CETIs that have made at least one contact  
+   as a funtion of the elapsed time since the awakening, given that
+   they make at least one contact on their lifetime.}
+   \label{F_}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+   \caption{}
+   \label{F_}
+\end{figure*}
+
 \ack[Acknowledgement]{

### 2019-08-12 manuscript.tex
manuscript

@@
 
-%}}}
-
-\subsection{Reciprocal communication}
-%{{{
+In the Fig.~\ref{F_waiting} we show the fraction of the CETIs that
+have made at least one contact  as a funtion of the elapsed time since
+the awakening, given that they make at least one contact on their
+lifetime.
+%
+From a frequentist approach, the normalizaed cumulative counts are
+related to the estimation of the probability of observing a given time
+with no success.
+%             
+XX panels show the histograms of the waiting times for the first
+contact, for all the CETIs that make at least one contact along their
+lifetime.
+%
+We show separately the histograms for several values of $\tau_a$ and a
+fixed value of $(\tau_s, D_{max})$, and for several values of $\tau_s$ and a
+fixed value of $(\tau_a, D_{max})$.
+%
+For different values of $\tau_a$, the shape of the distribution is
+similar, so that a normalization ... but not in the case of different values of $\tau_s$.
 
@@
 
+   \caption{Fraction of the CETIs that have made at least one contact  
+   as a funtion of the elapsed time since the awakening, given that
+   they make at least one contact on their lifetime.
+   %
+   From a frequentist approach, this fraction is the estimation of the
+   probability
+
+   }
+   \label{F_waiting}
+\end{figure*}
+
+\begin{figure*}
+   \centering
    \includegraphics[width=0.33\textwidth]{prob1_10k.png}%
@@
 
-   \caption{Fraction of the CETIs that have made at least one contact  
-   as a funtion of the elapsed time since the awakening, given that
-   they make at least one contact on their lifetime.}
-   \label{F_}
-\end{figure*}
-
-\begin{figure*}
-   \centering
-   \caption{}
-   \label{F_}
+   \caption{F_}
+   \label{F_waiting}
 \end{figure*}

### 2019-08-16 manuscript.tex
waiting time...

@@
 Such distribution can be considered to compute the probability of
-reaching a first contact within the first XX years.
+reaching a first contact within the first XX years of a SETI project,
+assuming the methods are correct and the hypotheses of the experiment.
 
@@
 the awakening, given that they make at least one contact on their
-lifetime.
+lifetime, for different models.
 %
@@
 For different values of $\tau_a$, the shape of the distribution is
-similar, so that a normalization ... but not in the case of different values of $\tau_s$.
+similar, and the main difference given by the different number of
+contacts on the different models.
+%
+Conversely, 
+ For different values of $\tau_a$, the shape of the distribution is
+similar, and the main difference given by the different number of
+contacts on the different models.
+
+... but not in the case of different values of $\tau_s$.
 
@@
 
+\section{Discusion}\label{S_discussion}
+%{{{
+
+We have implemented a simulation approach to explore the hypothesis
+space about the causal connections between communicating
+extraterrestrial civilizations.
+%
+The different models can be generated with a minimal number of free
+parameters and some fixed assumptions.
+%
+We argue that three parameters can be used to describe a variety of
+situations, ranging from a Galaxy where an intelligent civilization is
+very rare, to a Galaxy populated with plenty of civilizations in
+causal contact.
+%
+We explored the consecuences of different models using mensurable
+quantities.
+%
+With this approach, we can estimate several quantities as a function
+of the parameters on the hypothesis space.
+%
+For instance, it is possible to estimate the timescale for an
+uninterrupted SETI effort to reach success, in terms of different
+model parameters that reflect different, so far unknown, scenarios
+for the spread if intelligent life on the Galaxy.
+
+Under the hypotheses of our experiments, we conclude that
+a causal contact is extremely unlikely unless the Galaxy is heavily
+populated by intelligent civilizations.
+%
+Even so, we must consider that in order to stablish a contact between
+any two entities, a minimum degree of compatibility must be
+accomplished without any previous agreement, 
+making the posibility of a contact with a message that
+could be deciphered highly rare.
+
+Future work can be performed following this framework, in order to
+explore possible implications of the results for more detailed 
+configurations of the experiment.
+%
+For example, the communication method (isotropic, colimated,
+serendipitous) can affect the observables.
+%
+Also the nature of the message carrier, electromagnetic or another.
+%
+It is also possible to explore the effects of alignments or the use of
+stars as sources or amplifiers.
+%
+%energetic markers
+%Spiritual markers
+%
+As a future work, we will explore several improvements to the
+simulation, and analyze the impact on the final results of these
+changes.
+%
+They include a spatial distribution that actually resembles the shape
+of the Galaxy, different distribution functions for the mean lifetime
+of CETIs and different efficiencies for the sphere of the causal
+region of each CETI, resembling different searching strategies.
+
+%}}}
+
+\begin{figure*}
+   \centering
+   \includegraphics[width=0.33\textwidth]{waiting_a_acum_log.png}%
+   \includegraphics[width=0.33\textwidth]{waiting_a_acum_lin.png}%
+   \includegraphics[width=0.33\textwidth]{waiting_a_dif.png}
+
+   \includegraphics[width=0.33\textwidth]{waiting_s_acum_log.png}%
+   \includegraphics[width=0.33\textwidth]{waiting_s_acum_lin.png}%
+   \includegraphics[width=0.33\textwidth]{waiting_s_dif.png}
+
+   \caption{Fraction of the CETIs that have made at least one contact  
+   as a funtion of the elapsed time since the awakening, given that
+   they make at least one contact on their lifetime.
+   %
+   From a frequentist approach, this fraction is the estimation of the
+   probability    
+   }
+   \label{F_waiting}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+   \includegraphics[width=0.5\textwidth]{Figure_taua.png}%
+   \includegraphics[width=0.5\textwidth]{Figure_taus.png}%
+   %
+   \caption{Histograms of the mean waiting times for the first contact, for
+   several models.
+   %
+   Left panel shows the histograms for several values of $\tau_a$, and
+   $\tau_s$ in the range 17000-24000 yr.
+   %
+   Right panel shows the histograms for several values of $\tau_s$, and
+   $\tau_a$ in the range 68000-100000 yr.
+   %
+   The shape of the histograms does not change when $\tau_a$ is
+   modified, but significantly changes when $\tau_s$ is modified.
+   }
+   \label{F_waiting}
+\end{figure*}
+
+\begin{figure*}
+   \centering
+   \includegraphics[width=0.33\textwidth]{prob1_10k.png}%
+   \includegraphics[width=0.33\textwidth]{prob1_40k.png}%
+   \includegraphics[width=0.33\textwidth]{prob1_80k.png}
+   %
+   \caption{F}
+   \label{F_waiting}
+\end{figure*}
+
 \begin{figure}
@@
 
-\section{Discusion}\label{S_discussion}
-%{{{
-
-We have implemented a simulation approach to explore the hypothesis
-space about the causal connections between communicating
-extraterrestrial civilizations.
-%
-The different models can be generated with a minimal number of free
-parameters and some fixed assumptions.
-%
-We argue that three parameters can be used to describe a variety of
-situations, ranging from a Galaxy where an intelligent civilization is
-very rare, to a Galaxy populated with plenty of civilizations in
-causal contact.
-%
-We explored the consecuences of different models using mensurable
-quantities.
-%
-With this approach, we can estimate several quantities as a function
-of the parameters on the hypothesis space.
-%
-For instance, it is possible to estimate the timescale for an
-uninterrupted SETI effort to reach success, in terms of different
-model parameters that reflect different, so far unknown, scenarios
-for the spread if intelligent life on the Galaxy.
-
-Under the hypotheses of our experiments, we conclude that
-a causal contact is extremely unlikely unless the Galaxy is heavily
-populated by intelligent civilizations.
-%
-Even so, we must consider that in order to stablish a contact between
-any two entities, a minimum degree of compatibility must be
-accomplished without any previous agreement, 
-making the posibility of a contact with a message that
-could be deciphered highly rare.
-
-Future work can be performed following this framework, in order to
-explore possible implications of the results for more detailed 
-configurations of the experiment.
-%
-For example, the communication method (isotropic, colimated,
-serendipitous) can affect the observables.
-%
-Also the nature of the message carrier, electromagnetic or another.
-%
-It is also possible to explore the effects of alignments or the use of
-stars as sources or amplifiers.
-%
-%energetic markers
-%Spiritual markers
-%
-As a future work, we will explore several improvements to the
-simulation, and analyze the impact on the final results of these
-changes.
-%
-They include a spatial distribution that actually resembles the shape
-of the Galaxy, different distribution functions for the mean lifetime
-of CETIs and different efficiencies for the sphere of the causal
-region of each CETI, resembling different searching strategies.
-
-%}}}
-
-\begin{figure*}
-   \centering
-   \includegraphics[width=0.33\textwidth]{waiting_a_acum_log.png}%
-   \includegraphics[width=0.33\textwidth]{waiting_a_acum_lin.png}%
-   \includegraphics[width=0.33\textwidth]{waiting_a_dif.png}
-
-   \includegraphics[width=0.33\textwidth]{waiting_s_acum_log.png}%
-   \includegraphics[width=0.33\textwidth]{waiting_s_acum_lin.png}%
-   \includegraphics[width=0.33\textwidth]{waiting_s_dif.png}
-
-   \caption{Fraction of the CETIs that have made at least one contact  
-   as a funtion of the elapsed time since the awakening, given that
-   they make at least one contact on their lifetime.
-   %
-   From a frequentist approach, this fraction is the estimation of the
-   probability
-
-   }
-   \label{F_waiting}
-\end{figure*}
-
-\begin{figure*}
-   \centering
-   \includegraphics[width=0.33\textwidth]{prob1_10k.png}%
-   \includegraphics[width=0.33\textwidth]{prob1_40k.png}%
-   \includegraphics[width=0.33\textwidth]{prob1_80k.png}
-
-   \caption{F_}
-   \label{F_waiting}
-\end{figure*}
-
 \ack[Acknowledgement]{

### 2019-08-16 manuscript.tex
un poco mas de texto

@@
 %
-Such distribution can be considered to compute the probability of
-reaching a first contact within the first XX years of a SETI project,
-assuming the methods are correct and the hypotheses of the experiment.
+Such distribution can be considered to compute the probability for a
+random CETI to carry out a SETI project and spend a given time until
+the first contact is made.
+%
+Reaching a first contact within the first $t_1$ years of a SETI project,
+requires assuming that the methods are correct and the hypotheses of
+the experiment fullfil.
+%
+This values can be considered as upper limits, since the efficiency of
+a SETI initiative could be smaller than the ideal.
+%
+Also, a civilization could consider itself as "awakened" before it
+discovers the right technology to perform a communication.
 
@@
 %
-From a frequentist approach, the normalizaed cumulative counts are
+From a frequentist approach, the normalized cumulative counts are
 related to the estimation of the probability of observing a given time
@@
 %             
-XX panels show the histograms of the waiting times for the first
+Both panels show the histograms of the waiting times for the first
 contact, for all the CETIs that make at least one contact along their
@@
 
+%    \caption{Fraction of the CETIs that have made at least one contact  
+%   as a funtion of the elapsed time since the awakening, given that they make at least one contact on their lifetime.
+%   %
+%   From a frequentist approach, this fraction is the estimation of the
+%   probability}
+
 %}}}
@@
 %{{{
+
+On the Figure~\ref{F_radial} we show the fraction of CETIs that make
+zero, one, two or three contacts, as a function of the galactocentric
+distance.
+%
+For this model, which corresponds to XXX, the majority of CETIs do not
+make any contact during all the time period in which they are active.
+%
+As expected, the fractions of CETIs making at least one contact
+diminishes as the number of contacts rises.
+%
+The shape of this distribution is a result of the geometry of the GHZ.
+%
+In general, a closer position to the center of the Galaxy slightly favours 
+the posibility of a contact.
+%
+This would indicate that the position of the Earth is suitable for the
+odds of making a contact.
 
@@
 \end{tabular}
-\caption{Selected models to analize the behevior of some variables.}
+\caption{Selected models to analize the behevior of simulation outputs
+   and their dependence on simulation outputs and their dependence on
+   simulation parameters..}
 \label{T_SelectedModels}
@@
 
+\begin{figure}
+   \centering
+   \includegraphics[width=0.5\textwidth]{radial.png}
+   \caption{Fraction of CETIs making cero, one, two or three contacts as a function of
+   galactocentric distance.}
+   \label{F_radial}
+\end{figure}
+
 \section{Discusion}\label{S_discussion}
@@
    \centering
-   \includegraphics[width=0.33\textwidth]{waiting_a_acum_log.png}%
-   \includegraphics[width=0.33\textwidth]{waiting_a_acum_lin.png}%
-   \includegraphics[width=0.33\textwidth]{waiting_a_dif.png}
-
-   \includegraphics[width=0.33\textwidth]{waiting_s_acum_log.png}%
-   \includegraphics[width=0.33\textwidth]{waiting_s_acum_lin.png}%
-   \includegraphics[width=0.33\textwidth]{waiting_s_dif.png}
-
-   \caption{Fraction of the CETIs that have made at least one contact  
-   as a funtion of the elapsed time since the awakening, given that
-   they make at least one contact on their lifetime.
-   %
-   From a frequentist approach, this fraction is the estimation of the
-   probability    
-   }
-   \label{F_waiting}
-\end{figure*}
-
-\begin{figure*}
-   \centering
    \includegraphics[width=0.5\textwidth]{Figure_taua.png}%
-   \includegraphics[width=0.5\textwidth]{Figure_taus.png}%
+   \includegraphics[width=0.5\textwidth]{Figure_taus.png}
    %
@@
 
-\begin{figure*}
-   \centering
-   \includegraphics[width=0.33\textwidth]{prob1_10k.png}%
-   \includegraphics[width=0.33\textwidth]{prob1_40k.png}%
-   \includegraphics[width=0.33\textwidth]{prob1_80k.png}
-   %
-   \caption{F}
-   \label{F_waiting}
-\end{figure*}
-
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{radial.png}
-   \caption{Fraction of CETIs making cero, one, two or three contacts as a function of
-   galactocentric distance.}
-   \label{F_res_5}
-\end{figure}
-
 \ack[Acknowledgement]{
@@
 the Secretar\'{\i}a de Ciencia y Tecnolog\'{\i}a, Universidad Nacional
-de C\'ordoba, Argentina, and the Universidad Católica de Córdoba,
+de C\'ordoba, Argentina, and the Universidad Cat\'olica de C\'ordoba,
 Argentina.

### 2019-08-20 manuscript.tex
figura de primer contacto con make up

@@
    \centering
-   \includegraphics[width=\textwidth]{Messages_01.pdf}
+   \includegraphics[width=0.8\textwidth]{Messages_01.pdf}
    \caption{Schematic representation of the possible cases in which a
@@
 %
+%Such distribution can be considered to compute the probability of
+%reaching a first contact within the first XX years of a SETI project,
+%assuming the methods are correct and the hypotheses of the experiment.
 Such distribution can be considered to compute the probability for a
@@
 the first contact is made.
-%
-Reaching a first contact within the first $t_1$ years of a SETI project,
-requires assuming that the methods are correct and the hypotheses of
-the experiment fullfil.
-%
-This values can be considered as upper limits, since the efficiency of
-a SETI initiative could be smaller than the ideal.
-%
-Also, a civilization could consider itself as "awakened" before it
-discovers the right technology to perform a communication.
 
@@
 
-... but not in the case of different values of $\tau_s$.
-
-%    \caption{Fraction of the CETIs that have made at least one contact  
-%   as a funtion of the elapsed time since the awakening, given that they make at least one contact on their lifetime.
-%   %
-%   From a frequentist approach, this fraction is the estimation of the
-%   probability}
-
 %}}}
@@
    and their dependence on simulation outputs and their dependence on
-   simulation parameters..}
+   simulation parameters.}
 \label{T_SelectedModels}
@@
 
+\section{Discusion}\label{S_discussion}
+%{{{
+
+We have implemented a simulation approach to explore the hypothesis
+space about the causal connections between communicating
+extraterrestrial civilizations.
+%
+The different models can be generated with a minimal number of free
+parameters and some fixed assumptions.
+%
+We argue that three parameters can be used to describe a variety of
+situations, ranging from a Galaxy where an intelligent civilization is
+very rare, to a Galaxy populated with plenty of civilizations in
+causal contact.
+%
+We explored the consecuences of different models using mensurable
+quantities.
+%
+With this approach, we can estimate several quantities as a function
+of the parameters on the hypothesis space.
+%
+For instance, it is possible to estimate the timescale for an
+uninterrupted SETI effort to reach success, in terms of different
+model parameters that reflect different, so far unknown, scenarios
+for the spread if intelligent life on the Galaxy.
+
+Under the hypotheses of our experiments, we conclude that
+a causal contact is extremely unlikely unless the Galaxy is heavily
+populated by intelligent civilizations.
+%
+Even so, we must consider that in order to stablish a contact between
+any two entities, a minimum degree of compatibility must be
+accomplished without any previous agreement, 
+making the posibility of a contact with a message that
+could be deciphered highly rare.
+
+Future work can be performed following this framework, in order to
+explore possible implications of the results for more detailed 
+configurations of the experiment.
+%
+For example, the communication method (isotropic, colimated,
+serendipitous) can affect the observables.
+%
+Also the nature of the message carrier, electromagnetic or another.
+%
+It is also possible to explore the effects of alignments or the use of
+stars as sources or amplifiers.
+%
+%energetic markers
+%Spiritual markers
+%
+As a future work, we will explore several improvements to the
+simulation, and analyze the impact on the final results of these
+changes.
+%
+They include a spatial distribution that actually resembles the shape
+of the Galaxy, different distribution functions for the mean lifetime
+of CETIs and different efficiencies for the sphere of the causal
+region of each CETI, resembling different searching strategies.
+
+%}}}
+
+\begin{figure}
+   \centering
+   %\includegraphics[width=0.5\textwidth]{Figure_taua.png}%
+   %\includegraphics[width=0.5\textwidth]{Figure_taus.png}
+   \includegraphics[width=0.5\textwidth]{waiting_s+a_dif_ylog.pdf}
+   %
+   \caption{Histograms of the mean waiting times for the first contact, for
+   several models.
+   %
+   Left panel shows the histograms for several values of $\tau_a$, and
+   $\tau_s$ in the range 17000-24000 yr.
+   %
+   Right panel shows the histograms for several values of $\tau_s$, and
+   $\tau_a$ in the range 68000-100000 yr.
+   %
+   The shape of the histograms does not change when $\tau_a$ is
+   modified, but significantly changes when $\tau_s$ is modified.
+   }
+   \label{F_waiting}
+\end{figure}
+
 \begin{figure}
@@
    galactocentric distance.}
-   \label{F_radial}
+   \label{F_res_5}
 \end{figure}
-
-\section{Discusion}\label{S_discussion}
-%{{{
-
-We have implemented a simulation approach to explore the hypothesis
-space about the causal connections between communicating
-extraterrestrial civilizations.
-%
-The different models can be generated with a minimal number of free
-parameters and some fixed assumptions.
-%
-We argue that three parameters can be used to describe a variety of
-situations, ranging from a Galaxy where an intelligent civilization is
-very rare, to a Galaxy populated with plenty of civilizations in
-causal contact.
-%
-We explored the consecuences of different models using mensurable
-quantities.
-%
-With this approach, we can estimate several quantities as a function
-of the parameters on the hypothesis space.
-%
-For instance, it is possible to estimate the timescale for an
-uninterrupted SETI effort to reach success, in terms of different
-model parameters that reflect different, so far unknown, scenarios
-for the spread if intelligent life on the Galaxy.
-
-Under the hypotheses of our experiments, we conclude that
-a causal contact is extremely unlikely unless the Galaxy is heavily
-populated by intelligent civilizations.
-%
-Even so, we must consider that in order to stablish a contact between
-any two entities, a minimum degree of compatibility must be
-accomplished without any previous agreement, 
-making the posibility of a contact with a message that
-could be deciphered highly rare.
-
-Future work can be performed following this framework, in order to
-explore possible implications of the results for more detailed 
-configurations of the experiment.
-%
-For example, the communication method (isotropic, colimated,
-serendipitous) can affect the observables.
-%
-Also the nature of the message carrier, electromagnetic or another.
-%
-It is also possible to explore the effects of alignments or the use of
-stars as sources or amplifiers.
-%
-%energetic markers
-%Spiritual markers
-%
-As a future work, we will explore several improvements to the
-simulation, and analyze the impact on the final results of these
-changes.
-%
-They include a spatial distribution that actually resembles the shape
-of the Galaxy, different distribution functions for the mean lifetime
-of CETIs and different efficiencies for the sphere of the causal
-region of each CETI, resembling different searching strategies.
-
-%}}}
-
-\begin{figure*}
-   \centering
-   \includegraphics[width=0.5\textwidth]{Figure_taua.png}%
-   \includegraphics[width=0.5\textwidth]{Figure_taus.png}
-   %
-   \caption{Histograms of the mean waiting times for the first contact, for
-   several models.
-   %
-   Left panel shows the histograms for several values of $\tau_a$, and
-   $\tau_s$ in the range 17000-24000 yr.
-   %
-   Right panel shows the histograms for several values of $\tau_s$, and
-   $\tau_a$ in the range 68000-100000 yr.
-   %
-   The shape of the histograms does not change when $\tau_a$ is
-   modified, but significantly changes when $\tau_s$ is modified.
-   }
-   \label{F_waiting}
-\end{figure*}
 

### 2019-08-20 manuscript.tex
complete las llamadas a las figuras, falta ordenar

@@
 longstanding question, which is often conceptualized as the problem
-of the lack of received communication of the Fermi paradox.
+of the lack of received communication or the Fermi paradox.
 %
@@
 %
+A different approach uses detailed numerical simulations of the galaxy
+and recipes for the rates of the formation of stars and planets, and
+even for the origin of life.
+%
 Here we present numerical simulations of stochastic processes of
-emergence of civilizations with communication capabilities, and
-the causal contacts among them.
-%
-We analyze the rate of causal contacts as a function of the mean
-number of civilizations and the mean lifetime span distribution.
-%
-Our results indicate that, given the large distances involved, in
-the odds to obtain a contact in a few years of monitoring, assuming a
-perfect detection rate, are very rare.                                 
+emergence of civilizations with communication capabilities, based
+on minimal asumptions, and analyze the causal connections among them.
+%
+The analysis of the rate of causal contacts as a function of the mean
+number of civilizations, the mean lifetime span distribution and the
+maximum distance a civilization can send signals is presented, and
+used as a framework to discuss the spatial and temporal structure
+of a populated galaxy within several different scenarios.
+%
+Our results indicate that, given the large distances involved, causal
+contacts between civilizations are very rare.
+%
+Additionally, the odds to make a contact in a few years of monitoring, 
+assuming a perfect detection rate, are low for most models, with the
+exception of models which propose a galaxy densely populated with
+long lived civilizations.
+%
+The probability of causal contacts increases with the increase of the 
+mean lifetime of civilizations much more signifficantly than with
+the increase in the mean number of active civilizations for a given
+time window.
 %
@@
 
-\section{Motivations}
-%\section{Motivations: Uncertainties and the non detection problem for ETI contact}
+\section{Motivations}\label{S_motivations}
 %{{{
@@
 %
-In Sec.~\ref{S_methods} we introduce the methods and discuss the candidate
-distributions for the statistical aspects of the times involved in the
-communication process.
-%
-Then, we present our results in Sec~\ref{S_results}, with special emphasis on
-the statistical distributions of the duration of causal contacts in one or
-both directions, the possible differences on the position of a
-civilization on the Galaxy and the distribution of time intervals for
-the waiting of the first contact, always as a function of the three
-simulation parameters.
-%
-Finally, in Sec.~\ref{S_discussion} we discuss our results and future research
-directions.
+In Sec.~\ref{S_methods} we introduce the methods and discuss the
+candidate distributions for the statistical aspects of the times
+involved in the communication process.
+%
+Then, we present our results in Sec~\ref{S_results}, with special
+emphasis on the statistical distributions of the duration of causal
+contacts in one or both directions, the possible differences on the
+position of a civilization on the Galaxy and the distribution of time
+intervals for the waiting of the first contact, always as a function
+of the three simulation parameters.
+%
+Finally, in Sec.~\ref{S_discussion} we discuss our results and future
+research directions.
 
@@
 
-\section{Methods and working hypotheses}
-
-%: accounting for causal contacts with 
-%discrete event simulation process}\label{S_methods}
+\section{Methods and working hypotheses}\label{S_methods}
 %{{{
@@
 %
-This method is well suited for the 
-particular case of the difusion of intelligence signals in
-the Galaxy.
-%
-Assuming some simple hypotheses, 
-the discrete events method can be performed taking into
-account a small number of variables, which allows to analyze the
-results as a function of those variables.
+This method is well suited for the particular case of the difusion of
+intelligence signals in the Galaxy.
+%
+Assuming some simple hypotheses, the discrete events method can be
+performed taking into account a small number of variables, which
+allows to analyze the results as a function of those variables.
 
@@
 contacts between pairs of intelligent civilizations in the Galaxy
-(Table \ref{T_simu}).
+(Table \ref{T_simu_hypotheses}).
 %
@@
 circumferences.
-
-The temporal structure of the process is defined by two distribution parameters that
-represent the mean time interval that an intelligent civilization can
-emit and receive signals, and the mean time interval between the
-emergence of an intelligent comunicating civilization and the next
-one.
+%
+We illustrate in the Fig.~\ref{F_growing_sphere} the schematic
+representation of the growing communicating sphere, over the
+space–time diagrams, where time is represented on the horizontal
+direction, and space is represented in the vertical direction.
+%
+In the upper panel (a), the sphere is growing as the surface of first
+contact has not reached the maximum distance.
+%
+In the middle panel (b) it has reached the maximum distance, so that
+it remains at the same size.
+%
+After a Doomsday event (bottom panel, b), the signals can still be
+observed, but the surface of last contact grows.
+%
+In the Fig.~\ref{F_abcd}, we show a schematic representation of two
+emitters, E$_1$ and E$_2$, that reach each other at different times.
+%
+The time span for E$_i$ is (A$_i$ , D$_i$) , for $i = {1, 2}$.
+%
+Emitter $i$ can listen to emitter $j$ between C$_{ij}$ and B$_{ij}$.
+%
+The type and length of causal contact in both directions depend on the
+distance and time lag between the awakening events, the maximum
+distance that a signal can reach and the time period in which each
+emitter is active.
+%
+In the Fig.~\ref{F_messages} we show a schematic representation of the
+possible cases in which a CETI A can be in causal contact with a CETI
+B which appears later.
+%
+The duration of the causal contact in one direction depends on several
+factors, mainly $\Delta$t~A , $\Delta$t~C , and $\Delta$X.
+
+The temporal structure of the process is defined by two distribution
+parameters that represent the mean time interval that an intelligent
+civilization can emit and receive signals, and the mean time interval
+between the emergence of an intelligent comunicating civilization and
+the next one.
 %
@@
 The parameters for the temporal distributions also determine the
-spatial properties since the density of active cetis in the Galaxy depend on these
-two parameters.
+spatial properties since the density of active cetis in the Galaxy
+depend on these two parameters.
 %
@@
    that are part of the simulation.}
-\label{T_simu}
+\label{T_simu_hypotheses}
 \end{table}
@@
    of the process and the reach of messages.}
-\label{T_parameters}
+\label{T_range_parameters}
 \end{table}
@@
    still be observed, but the surface of last contact grows }
-   \label{F_sphere}
+   \label{F_growing_sphere}
 \end{figure}
@@
 The details of each model are summarized in the
-Table~\ref{T_SelectedModels}.
-
-In Fig.~\ref{F_res_1} we show the Empirical cumulative distributions
-of the number contacts for CETIs in four different samplas (M5, M6,
-M7, M8) with long range signal reach (D$_{max}$=10000). 
+Table~\ref{T_selected_models}.
+
+In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
+distributions of the number contacts for CETIs in four different
+samples (M5, M6, M7, M8) with long range signal reach
+(D$_{max}$=10000). 
 %
@@
 
-\subsection{Duration of contacts}
+\subsection{Multiplicity and duration of contacts}\label{SS_multiplicity}
 %{{{
@@
 
-\subsection{Waiting time for a first contact}
+\subsection{Waiting time for a first contact}\label{SS_waiting}
 %{{{
@@
 
-In the Fig.~\ref{F_waiting} we show the fraction of the CETIs that
-have made at least one contact as a funtion of the elapsed time since
-the awakening, given that they make at least one contact on their
-lifetime, for different models.
+In the Fig.~\ref{F_waiting_time} we show the fraction of the CETIs
+that have made at least one contact as a funtion of the elapsed time
+since the awakening, given that they make at least one contact on
+their lifetime, for different models.
 %
@@
 We show separately the histograms for several values of $\tau_a$ and a
-fixed value of $(\tau_s, D_{max})$, and for several values of $\tau_s$ and a
-fixed value of $(\tau_a, D_{max})$.
+fixed value of $(\tau_s, D_{max})$, and for several values of $\tau_s$
+and a fixed value of $(\tau_a, D_{max})$.
 %
@@
 %
-Conversely, 
- For different values of $\tau_a$, the shape of the distribution is
-similar, and the main difference given by the different number of
-contacts on the different models.
+Conversely, for different values of $\tau_a$, the shape of the
+distribution is similar, and the main difference given by the
+different number of contacts on the different models. 
+%
+We show the 2D maps of the rate of CETIs that never make contact and
+the rate of CETIs that make contact at the moment of the awakening,
+inf the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
+as a function of $\tau_s$ and $\tau_s$.
+%
+In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
+waiting times for the first contact, for several models.
+%
+Left panel shows the histograms for several values of $\tau_a$, and
+$\tau_s$ in the range 17000-24000 yr.
+%
+Right panel shows the histograms for several values of $\tau_s$, and
+$\tau_a$ in the range 68000-100000 yr.
+%
+The shape of the histograms does not change when $\tau_a$ is modified,
+but significantly changes when $\tau_s$ is modified.
 
@@
 
-\subsection{Location within the Galaxy}
+\subsection{Location within the Galaxy}\label{SS_location}
 %{{{
 
-On the Figure~\ref{F_radial} we show the fraction of CETIs that make
+On the Figure~\ref{F_radial_frac} we show the fraction of CETIs that make
 zero, one, two or three contacts, as a function of the galactocentric
@@
 M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
-
+%
 M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
@@
 M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
-
+%
 \hline
@@
    simulation parameters.}
-\label{T_SelectedModels}
+\label{T_selected_models}
 \end{table*}
@@
    \caption{Empirical cumulative distributions of the number contacts
-   for CETIs in four different samplas (M5, M6, M7, M8) with long
+   for CETIs in four different samples (M5, M6, M7, M8) with long
    range signal reach (D$_{max}$=10000).
@@
    % 
-\label{F_res_1} \end{figure}
+\label{F_number_of_contacts} \end{figure}
 
@@
    $\tau_a$ and $\tau_s$}
-   \label{F_res_2}
+   \label{F_waiting_time}
 \end{figure}
@@
    $D_{max}$=80000 (right panel)}
-   \label{F_res_3}
+   \label{F_never_contact}
 \end{figure*}
@@
    awakening, as a funcion of $\tau_a$ and $\tau_s$}
-   \label{F_res_4}
+   \label{F_C_at_A}
 \end{figure*}
@@
 \begin{figure}
+   %
    \centering
-   %\includegraphics[width=0.5\textwidth]{Figure_taua.png}%
-   %\includegraphics[width=0.5\textwidth]{Figure_taus.png}
+   %
    \includegraphics[width=0.5\textwidth]{waiting_s+a_dif_ylog.pdf}
@@
    }
-   \label{F_waiting}
+   %
+   \label{F_waiting_for_1C}
+   %
 \end{figure}
@@
 \begin{figure}
+   %
    \centering
+   %
    \includegraphics[width=0.5\textwidth]{radial.png}
-   \caption{Fraction of CETIs making cero, one, two or three contacts as a function of
-   galactocentric distance.}
-   \label{F_res_5}
+   %
+   \caption{Fraction of CETIs making cero, one, two or three contacts
+   as a function of galactocentric distance.}
+   %
+   \label{F_radial_frac}
+   %
 \end{figure}

### 2019-08-21 manuscript.tex
detalles

@@
 %
-The adopted values for the GHZ are XX and XX \citep{Lineweaver2004}.
+The adopted values for the GHZ are 20 kly and 60 kly, for the inner
+and outer radius, respectively \citep{Lineweaver2004}.
+% ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
+% mas chico...
 %
@@
 
-We implemented the simulation of a regular grid of models varying over the
-hypothesis space, which covers XXX different models.
-%
-For each model, we simulated XXX realizations with different random
+We implemented the simulation of a regular grid of models varying over
+the hypothesis space, which covers 15000 different models.
+%
+For each model, we simulated 50 realizations with different random
 seeds, in order to estimate the variance.
 %
-The simulated parameters, $\lambda_1$, $\lambda_2$ and $D_{max}$ cover
-the range XX-XX, XX-XX and XX-XX, respectively, with a partition of XX
+The simulated parameters, $\tau_a$, $\tau_s$ cover the ranges
+4000-200000 and 10000-500000, respectively, with a partition of 50
 values for each parameter.
+%
+For the D$_{max}$ parameter, we chose the values 500, 1000, 10000,
+20000, 40000 and 80000 lyr.
+%
+This setup makes a total of 750000 simulations, each covering a time
+range of one million years.
 
@@
 %
-
 We chose eight models that are on fairly oposed regions of the
@@
 In what follows, we focus on the analysis of the duration of contacts
-(Sec. X), the waiting time for the first contact (Sec. X) the
+(Sec.~\ref{SS_multiplicity}), the waiting time for the first contact
+(Sec.~\ref{SS_waiting}) and the
 variations of contact posibilities as a function of the position of
-the CETI in the Galaxy (Sec. X) and the occurrence of reciprocal
-communication (Sec. X).
+the CETI in the Galaxy (Sec.~\ref{SS_location}).
 
@@
 zero, one, two or three contacts, as a function of the galactocentric
-distance.
-%
-For this model, which corresponds to XXX, the majority of CETIs do not
+distance, for model M1.
+% VERIFICAR!  M1?
+%
+For this model, the majority of CETIs do not
 make any contact during all the time period in which they are active.

### 2019-08-22 manuscript.tex
make up de las ultimas figuras
detalles del texto

@@
 might be in causal contact with an observer located on the Earth.
-
+%
 \citet{cirkovic_temporal_2004} also emphasize the lack of temporal
@@
 %
-This method is well suited for the particular case of the difusion of
+In the case of study, the actors are the communicating civilizations
+(CETIs), and the events are all meaningfull that produce a change in
+the simulation variables, i.e., in the arrangement of CETIs and their
+causal contacts.
+%
+Since it allows to explore several models easily, 
+this method is well suited for the particular case of the difusion of
 intelligence signals in the Galaxy.
@@
 performed taking into account a small number of variables, which
-allows to analyze the results as a function of those variables.
+allow to analyze the variation of the results in the model parameter space.
 
@@
 %
-The causal connection can be produced without the need of the two
-civilizations being concurrent or active at the same time.
-%
-Each CETI will fill an annular region of the Galaxy, limited by two
+We also need to propose theoretical distribution functions for both the
+distribution of lifespans and the distribution of the number of CETIs
+per unit time.
+%
+The shapes of these distributions are discussed in Sec.~\ref{SS_PDF_shape}.
+%
+A causal connection can be produced without the need of the two
+involved civilizations being concurrent or active at the same time.
+%
+This is a fundamental property of the system which arises as a
+consecuence of the large spatial and temporal scales, where a message
+is transmited at light speed.
+%
+As a consecuence of their finite lifetime,
+each CETI will fill a spherical shell region of the Galaxy, limited by two
 surfaces.
@@
 The leading front, or surface of first contact (SFC) grows from the
-central CETI until it reaches the distance $D_{max}$.
+central CETI until it reaches the maximum distance $D_{max}$.
+%
+Although a CETI could be old enough to transmit a message at large
+distances, the limited power of the message and the dilution that runs
+as the squared distance from the source, imposes a
+detectability limit.
+%
+Thus, the volume reached causally by a CETI is initially a growing
+sphere.
+%
+Following the end of the civilization, there is still a region which
+is covered with the emited signals.
 %
@@
 the central CETI, with a delay with respect to the SFC equivalent to
-the lifetime of the CETI, and produces an annular region that shrinks
-until it reaches $D_{max}$, when all signal from the central CETI is
-lost.
+the lifetime of the CETI, and produces a spherical shell region.
+%
+Any other CETI within this region will be aware of the originating CETI,
+even if it has disapeared before the time of contact.
+%
+This region will grow if the SFC has not reached $D_{max}$, and
+will shrink otherwise, until the SLC reaches $D_{max}$.
+%
+A that point, all signal from the central CETI is lost.
 %
@@
 plane of the galaxy is relevant, and produce the corresponding
-circumferences.
+circles for the filled spheres or annular regions for the spherical
+shells.
 %
 We illustrate in the Fig.~\ref{F_growing_sphere} the schematic
-representation of the growing communicating sphere, over the
-space–time diagrams, where time is represented on the horizontal
+representation of the region causally connected to the central CETI.
+%
+The initially growing communicating sphere, is shown over
+space--time diagrams, where time is represented on the horizontal
 direction, and space is represented in the vertical direction.
@@
 factors, mainly $\Delta$t~A , $\Delta$t~C , and $\Delta$X.
-
-The temporal structure of the process is defined by two distribution
-parameters that represent the mean time interval that an intelligent
-civilization can emit and receive signals, and the mean time interval
-between the emergence of an intelligent comunicating civilization and
-the next one.
-%
-The spatial structure of the simulation is given by the size and shape
-of the Galactic Habitable Zone and the maximum distance a signal can
-travel to be detected.
-%
-The parameters for the temporal distributions also determine the
-spatial properties since the density of active cetis in the Galaxy
-depend on these two parameters.
-%
-For example, a small $\tau_a$ and a large $\tau_s$ will produce a
-Galaxy densely populated woth CETIs, and viceversa.
-%
-Also, some hypotheses must be made in order to complete the
-simulation.
-%
-Firstly, we assume that the distribution of the appeareance of new
-CETIs is a stationary Poisson distribution.
-%
-Regarding the duration of a CETI, we propose that the distribution of
-the duration of CETIs is a stationary exponential distribution.
-%
-\citet{Balbi2018} stress the fact that it would be desirable to arrive
-at a theoretical statistical distribution of the lifespan of
-civilizations, preferentially on the basis of the underlying
-astrophysical and biological processes.
-%
-\citet{Maccone} argues that this distribution should be a log normal.
-(EXPAND)
-
-There is no a temporal window for the model, i.e., the simulation
-starts when the process is already stable, and ends before any
-galactic effect could modify or affect of fixed variables.
-%
-The probability for the rise of a CETI is homogeneous over the GHZ.
-%
-Although the Galaxy has a well known spiral structure, the assumed
-sparcity of civilizations makes this assumption very reasonably.
-%
-Otherwise, if the distribution of civilizations is not sparce, it
-could be the case that the spiral arms would host most of the
-civilizations, and contacts are frecuent between closely located
-civilizations.
-%
-In such a case we could expect that the interpretation of our results
-could be considered as pesimist for the stablishment of a contact with
-our planet.
-%
-We also limit the posibilities of life or other types of civilizations
-to the hypothesis usually stated for the definition of the GHZ.
-%
-This means that we set aside possibel civilizations that could survive
-in severe conditions or unstable systems, which would prevent the
-appeareance of life as we know it.
-
-\setlength{\tabcolsep}{10pt}
-\begin{table}
-\centering
-\begin{tabular}{cl}
-\hline
-   \multicolumn{2}{l}{independent variables} \\
-\hline
-   $\tau_{a}$ & Mean temporal separation between\\
-              & consecutive awakenings \\
-   $\tau_{s}$ & Mean lifetime of a CETI \\
-   D$_{max}$ & Maximum reach of a message \\
-\hline
-\multicolumn{2}{l}{fixed variables and assumptions} \\
-\hline
-   & All CETIs share the same statistical properties\\
-   & The process is homogeneous \\
-   & The scan of the sly is fully efficient\\
-   & Absence of panspermia or colonization \\
-   & The GHZ is a 2--dimensional ring \\
-20000ly    & Inner radius of the GHZ \\
-60000ly    & Outer radius of the GHZ \\
-1.e6 y    & Time span of the simulation \\
-
-\hline
-\end{tabular}
-\caption{Definition of independent variables and fixed paramenters 
-   that are part of the simulation.}
-\label{T_simu_hypotheses}
-\end{table}
-
-While the aforementioned assumptions are basic conditions for most of
-the stochastic processes observed in nature, 
-we need to make stronger assumptions related to the nature of the
-message.
-%
-The most simple assumption about the message itself, is that it
-travels at light speed.
-%
-With this, we are considering messages sent through, for example,
-electromagnetic radiation or even gravitational radiation, but we set
-appart messages sent with mechanical means or physical objects, or
-through some unknown technology that violates the known so far laws of
-physics.
-% 
-For the communication of CETIs through messages sent isotropically, 
-we assume that the capacity to emit signals and to receive
-signals ocurr at the same time.
-%
-This means that the ability to find another civilization develops at
-the same time that the ability and intention to emit a signal in all
-direction, intented to be detected by an unknown civilization.
-%
-Although there are several reasons to think that this could not be the
-case, at large time scales it can be considered that both abilities
-occurr roughly at coincident epochs.
-%
-Another essential assumption is that all CETIs use the same signal power, so that there is a maximum distance out to which it can be detected.
-%
-It is worth mentioning that we are considering in our experimental
-setup a system composed by several emmiters and receivers across the
-Galaxy, under the hypothesis proposed before.
-%
-In such a system we compute probabilities of a CETI making contact
-with another CETI, and the results must be interpreted as for a
-randomly chosen CETI, given certain conditions of homogeneityon an
-ensemble of equivalent galaxies.
-%
-We stress the fact that we are not centered on the probabilities for
-the Earth itselft, but on a generalized CETI.
-%
-Regarding the extent of the signals, we know that the distance from
-which a signal from Earth could be detected using the current
-technology is of a few hundreds light years, given that the signal was
-sent to a specific direction.
-
-%https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
-%https://www.engadget.com/2017/05/25/listening-to-starlight-our-ongoing-search-for-alien-intelligenc/]
-%[from https://en.wikipedia.org/wiki/Fermi_paradox]
-%
-It is straightforward to propose and implement a distribution of
-maximum distances, although this would increase the model complexity
-at the cost of a larger uncertainty or of another dimension on the
-hypothesis space.
-%
-This proposal is based on a pronoid scenario (REF), and we do not
-consider in this work the ocurrence of paranoid or partially paranoid
-civilizations.
-%
-If that would be the case, the results obtained here can be taken as
-upper limits.
-
-In our simulations we also assume the simplest situations for the
-growth of ETIs.
-%
-For example, we discard the possibility of stellar colonization
-(Walters 1980) and assume that the communication aim is performed to
-all directions with the same power and the same probability, i.e.,
-strongly assume isotropic communication in all cases.
-%
-More detailed simulations could be produced considering different 
-efficiency of communication or detection methods.
-%
-Also, a mixed paranoid/pronoid population of intelligent civilizations
-could be implemented easily.
-%
-However, in this case we also would complicate the experimental setup,
-make the results less evident.
-%
-The results are independent of the nature of life (organic or
-artificial).
-%
-The lifetime of a civilization can be caused by autodestruction or by
-external factors.
-
-\subsection{Power laws vs. exponential laws}
-
-The exponential distribution of lifespan and waiting times is
-justified by considering that the process of appeareance of life in
-the galaxy is homogeneous and stationary.
-%
-This means that there is no preferred location within the GHZ for the
-spontaneos appeareance of life, and that the emergence of a
-civilization is independent of the existence of previous civilizations 
-in the galaxy.
-%            
-This is equivalent to proposing a Poisson process for the emergence of
-CETIs, since there is a close relation between the number of events in
-time or space and the waiting time or separation, respectively.
-%
-That is, these are two alternative approaches to describing the same
-process, a Poisson distribution for the number of events implies an
-exponential distribution for their sepations, and viceversa.
-%
-It should be emphasized that the exponential laws used in this work
-are assumed, although instead of analyzing results from a particular
-parameter chosen ad--hoc, we explore the hypothesis space and analyze
-the impact of the values of these parameters on the results.
-%
-
-The power law and exponential statistical distributions are among the
-most common patterns found in natural phenomena.
-%
-For example, the distribution of the frecuency of words in many
-languajes is known to follow a Zipf law (which is a power law).
-%
-Zipf law also describes population ranks of cities in various
-countries, corporation sizes, income rankings, ranks of number of
-people watching the same TV channel (Zipf's law entry on Wikipedia).
-%
-The magnitudes of earthquakes, hurricanes, volcanic eruptions and
-floods; the sizes of meteorites or the losses caused by business
-interruptions from accidents, are also well described by power laws
-(Sornette 2006).
-%
-The power law behaviour has been observed in a variety of systems,
-including for example stock market fluctuations, sizes of computer
-files and word frequency in languages \citep{mitzenmacher_brief_2004,
-Newman_power_2005, simkin_theory_2006}
-%
-Power laws have also been widely used in biological sciences, e.g., in
-analysing connectivity patterns in metabolic networks
-\citep{jeong_large_2000} and in the number of species observed per
-unit area in ecology \citep{garcia_origin_2006, frank_common_2009}.
-
-\subsection{Discrete event process}
-
-Given a set of parameters $(\tau_a, \tau_s, D_{max})$, the functional
-forms of the statistical distributions and fixed values
-for the hypotheses, a discrete
-event simulation is performed by keeping track of a set of variables
-that change each time a meaningful event is produced.
-
-\setlength{\tabcolsep}{10pt}
-\begin{table}
-\centering
-\begin{tabular}{cccc}
-\hline
-   fixed parameter & \multicolumn{3}{c}{adopted value}\hfill \\
-\hline
-   inner GHZ radius\hfill & \multicolumn{3}{c}{20000 ly}\\
-   outer GHZ radius & \multicolumn{3}{c}{60000 ly}\\ \\
-\hline
-   variable parameter & min & max & N \\
-\hline
-   D$_{max}$ & 1000 ly & 20000 ly & 20\\
-   $\tau_{a}$ & 1000 y & 20000 y & 20\\
-   $\tau_{s}$ & 1000 y & 20000 y & 20\\
-\hline
-\end{tabular}
-\caption{Adopted values for simulation parameters. Fixed parameters
-   define the GHZ.  Variable parameters define the temporal structure
-   of the process and the reach of messages.}
-\label{T_range_parameters}
-\end{table}
-
-The main variables that follow the evolution of the simulation are:
-
-\begin{itemize}
-   \item Position of stars.  Sampled randomly within the GHZ
-   \item Time of the awakening of each CETI
-   \item Time of the vanishing of each CETI
-\end{itemize}
-
-The variables that can be deduced from the previous ones are:
-
-\begin{itemize}
-   \item Number of CETIs in casual contact with at least another CETI
-      at a given time.
-   \item Number of CETIs as a function of time
-   \item Number of CETIs that receive a message at least one time
-   \item Number of CETIs that receive a message at least one time and
-      successfully deliver an answer.
-   \item Number distribution of waiting time to receive a message
-\end{itemize}
-
-\subsection{Exploration of the hypothesis space}
-
-%========================================================================
-%========================================================================
-%========================================================================
 
@@
 
+The temporal structure of the process is defined by two distribution
+parameters that represent the mean time interval that an intelligent
+civilization can emit and receive signals, and the mean time interval
+between the emergence of an intelligent comunicating civilization and
+the next one.
+%
+The spatial structure of the simulation is given by the size and shape
+of the Galactic Habitable Zone and the maximum distance a signal can
+travel to be detected.
+%
+The parameters for the temporal distributions also determine the
+spatial properties since the density of active cetis in the Galaxy
+depend on these two parameters.
+%
+For example, a small $\tau_a$ and a large $\tau_s$ will produce a
+Galaxy densely populated woth CETIs, and viceversa.
+%
+Also, some hypotheses must be made in order to complete the
+simulation.
+%
+Firstly, we assume that the distribution of the appeareance of new
+CETIs is a stationary Poisson distribution.
+%
+Regarding the duration of a CETI, we propose that the distribution of
+the duration of CETIs is a stationary exponential distribution.
+%
+\citet{Balbi2018} stress the fact that it would be desirable to arrive
+at a theoretical statistical distribution of the lifespan of
+civilizations, preferentially on the basis of the underlying
+astrophysical and biological processes.
+%
+\citet{Maccone} argues that this distribution should be a log normal.
+(EXPAND)
+
+There is no a temporal window for the model, i.e., the simulation
+starts when the process is already stable, and ends before any
+galactic effect could modify or affect of fixed variables.
+%
+The probability for the rise of a CETI is homogeneous over the GHZ.
+%
+Although the Galaxy has a well known spiral structure, the assumed
+sparcity of civilizations makes this assumption very reasonably.
+%
+Otherwise, if the distribution of civilizations is not sparce, it
+could be the case that the spiral arms would host most of the
+civilizations, and contacts are frecuent between closely located
+civilizations.
+%
+In such a case we could expect that the interpretation of our results
+could be considered as pesimist for the stablishment of a contact with
+our planet.
+%
+We also limit the posibilities of life or other types of civilizations
+to the hypothesis usually stated for the definition of the GHZ.
+%
+This means that we set aside possibel civilizations that could survive
+in severe conditions or unstable systems, which would prevent the
+appeareance of life as we know it.
+
+\setlength{\tabcolsep}{10pt}
+\begin{table*}
+\centering
+\begin{tabular}{clccc}
+\hline
+   \multicolumn{2}{l}{independent variables} &min&max&Nbins\\
+\hline
+   $\tau_{a}$ & Mean temporal separation between consecutive awakenings & 4000 y & 200000 y & 50\\ 
+   $\tau_{s}$ & Mean lifetime of a CETI                                 & 10000 y & 500000 y& 50\\ 
+   D$_{max}$ & Maximum reach of a message  &  \multicolumn{3}{l}{500, 1000, 10000, 20000, 40000, 80000 lyr}  \\
+\hline
+   \multicolumn{4}{l}{fixed variables and assumptions} &value \\
+\hline
+   & statistical properties of all CETIs &equally distributed&&\\
+   & Point process for the distribution in time & homogeneous &&\\
+   f$_s$ & The scan of the sky & fully efficient&&1\\
+   f$_p$ & panspermia or colonization &absent&&0\\
+   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
+   R$_{SFC}$   & Inner radius of the GHZ     &&&20000 ly\\
+   R$_{SLC}$   & Outer radius of the GHZ     &&&60000 ly\\
+   t$_{max}$ & Time span of the simulation  &&&1.e6 y\\
+\hline
+   \multicolumn{5}{l}{discrete events}\\
+\hline
+   A & Awakening: a civilization starts a SETI program  &&&\\
+   B & Blackout: the end of the communication chanel stops &&&\\
+   C & Contact: a new causal contact is produced &&&\\
+   D & Doomsday: a civilization ends its communication capabilities&&&\\
+\hline
+
+\hline
+\end{tabular}
+\caption{Definition of independent variables and adopted values for 
+   fixed paramenters 
+   that are part of the simulation.  Variable parameters define the
+   spatial and temporal structure
+   of the process and the maximum reach of the messages.}
+\label{T_simu_hypotheses}
+\end{table*}
+
+While the aforementioned assumptions are basic conditions for most of
+the stochastic processes observed in nature, we need to make stronger
+assumptions related to the nature of the message.
+%
+The most simple assumption about the message itself, is that it
+travels at light speed.
+%
+With this, we are considering messages sent through, for example,
+electromagnetic radiation or even gravitational radiation, but we set
+appart messages sent with mechanical means or physical objects, or
+through some unknown technology that violates the known so far laws of
+physics.
+% 
+For the communication of CETIs through messages sent isotropically, we
+assume that the capacity to emit signals and to receive signals ocurr
+at the same time.
+%
+This means that the ability to find another civilization develops at
+the same time that the ability and intention to emit a signal in all
+direction, intented to be detected by an unknown civilization.
+%
+Although there are several reasons to think that this could not be the
+case, at large time scales it can be considered that both abilities
+occurr roughly at coincident epochs.
+%
+Another essential assumption is that all CETIs use the same signal
+power, so that there is a maximum distance out to which it can be
+detected.
+%
+It is worth mentioning that we are considering in our experimental
+setup a system composed by several emmiters and receivers across the
+Galaxy, under the hypothesis proposed before.
+%
+In such a system we compute probabilities of a CETI making contact
+with another CETI, and the results must be interpreted as for a
+randomly chosen CETI, given certain conditions of homogeneityon an
+ensemble of equivalent galaxies.
+%
+We stress the fact that we are not centered on the probabilities for
+the Earth itselft, but on a generalized CETI.
+%
+Regarding the extent of the signals, we know that the distance from
+which a signal from Earth could be detected using the current
+technology is of a few hundreds light years, given that the signal was
+sent to a specific direction.
+
+\begin{figure*}
+   \centering
+   \includegraphics[width=0.8\textwidth]{Messages_01.pdf}
+   \caption{Schematic representation of the possible cases in which a
+   ceti A can be in causal contact with a ceti B which appears later.
+   The duration of the causal contact in one direction depends on
+   several factors, mainly $\Delta t_A$, $\Delta t_C$, and $\Delta X$.}
+   \label{F_messages}
+\end{figure*}
+
+%https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
+%https://www.engadget.com/2017/05/25/listening-to-starlight-our-ongoing-search-for-alien-intelligenc/]
+%[from https://en.wikipedia.org/wiki/Fermi_paradox]
+%
+It is straightforward to propose and implement a distribution of
+maximum distances, although this would increase the model complexity
+at the cost of a larger uncertainty or of another dimension on the
+hypothesis space.
+%
+This proposal is based on a pronoid scenario (REF), and we do not
+consider in this work the ocurrence of paranoid or partially paranoid
+civilizations.
+%
+If that would be the case, the results obtained here can be taken as
+upper limits.
+
+In our simulations we also assume the simplest situations for the
+growth of ETIs.
+%
+For example, we discard the possibility of stellar colonization
+(Walters 1980) and assume that the communication aim is performed to
+all directions with the same power and the same probability, i.e.,
+strongly assume isotropic communication in all cases.
+%
+More detailed simulations could be produced considering different 
+efficiency of communication or detection methods.
+%
+Also, a mixed paranoid/pronoid population of intelligent civilizations
+could be implemented easily.
+%
+However, in this case we also would complicate the experimental setup,
+make the results less evident.
+%
+The results are independent of the nature of life (organic or
+artificial).
+%
+The lifetime of a civilization can be caused by autodestruction or by
+external factors.
+%}}}
+
+\subsection{Power laws vs. exponential laws}\label{SS_PDF_shape}
+%{{{
+
+The exponential distribution of lifespan and waiting times is
+justified by considering that the process of appeareance of life in
+the galaxy is homogeneous and stationary.
+%
+This means that there is no preferred location within the GHZ for the
+spontaneos appeareance of life, and that the emergence of a
+civilization is independent of the existence of previous civilizations 
+in the galaxy.
+%            
+This is equivalent to proposing a Poisson process for the emergence of
+CETIs, since there is a close relation between the number of events in
+time or space and the waiting time or separation, respectively.
+%
+That is, these are two alternative approaches to describing the same
+process, a Poisson distribution for the number of events implies an
+exponential distribution for their sepations, and viceversa.
+%
+It should be emphasized that the exponential laws used in this work
+are assumed, although instead of analyzing results from a particular
+parameter chosen ad--hoc, we explore the hypothesis space and analyze
+the impact of the values of these parameters on the results.
+%
+
+The power law and exponential statistical distributions are among the
+most common patterns found in natural phenomena.
+%
+For example, the distribution of the frecuency of words in many
+languajes is known to follow a Zipf law (which is a power law).
+%
+Zipf law also describes population ranks of cities in various
+countries, corporation sizes, income rankings, ranks of number of
+people watching the same TV channel (Zipf's law entry on Wikipedia).
+%
+The magnitudes of earthquakes, hurricanes, volcanic eruptions and
+floods; the sizes of meteorites or the losses caused by business
+interruptions from accidents, are also well described by power laws
+(Sornette 2006).
+%
+The power law behaviour has been observed in a variety of systems,
+including for example stock market fluctuations, sizes of computer
+files and word frequency in languages \citep{mitzenmacher_brief_2004,
+Newman_power_2005, simkin_theory_2006}
+%
+Power laws have also been widely used in biological sciences, e.g., in
+analysing connectivity patterns in metabolic networks
+\citep{jeong_large_2000} and in the number of species observed per
+unit area in ecology \citep{garcia_origin_2006, frank_common_2009}.
+
+%}}}
+
+\subsection{Discrete event process}
+%{{{
+
+Given a parametric model $M(\tau_a, \tau_s, D_{max})$, which includes the
+functional forms of the statistical distributions and fixed values for the
+hypotheses, a discrete event simulation is performed by keeping track of a set
+of variables that change each time a meaningful event is produced.
+%
+The main variables that follow the evolution of the simulation are the positions
+of stars, which are sampled randomly within the GHZ, the time of the awakening
+of each CETI (A event), and the time of the vanishing of each CETI (D event).
+%
+The variables that can be deduced from the previous ones include the number of
+CETIs in casual contact with at least another CETI at a given time, the number
+of CETIs as a function of time, the number of CETIs that receive a message at
+least one time, the number of CETIs that receive a message at least one time and
+successfully deliver an answer and the number distribution of waiting time to
+receive a message.
+%
+All these quantities are updated each time one of the four events (A, B, C, D)
+is produced.
+
+\begin{figure} \centering
+   %
+   \includegraphics[width=0.5\textwidth]{F1.pdf}
+   %
+   \caption{Empirical cumulative distributions of the number contacts
+   for CETIs in four different samples (M5, M6, M7, M8) with long
+   range signal reach (D$_{max}$=10000).
+   %
+   }
+   % 
+\label{F_number_of_contacts} \end{figure}
+
+\begin{figure} % Distribution
+   \centering
+   \includegraphics[width=0.5\textwidth]{F2.pdf}
+   \caption{Rate of CETIs that make no contact as a function of
+   $\tau_a$ and $\tau_s$}
+   \label{F_waiting_time}
+\end{figure}
+
+%}}}
+
+%\subsection{Exploration of the hypothesis space}
+
+\subsection{Model for CETIs}
+%{{{
+
+Since the method is based on events, the first step is to define an
+architecture of events, the relationships between them and how events
+trigger changes on the state of the system.
+%
+We consider a simulated system that represents the galactic habitable
+zone (GHZ).
+%
+On a first approach, the GHZ is a 2-dimensional annular region.
+%
+The adopted values for the GHZ are 20 kly and 60 kly, for the inner
+and outer radius, respectively \citep{Lineweaver2004}.
+% ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
+% mas chico...
+%
+This simple model does not take into accont the variations in stellar
+density given by the spiral structure.
+%
+Although there are several possible approaches, we chose to follow the
+evolution of the system according to the following events:
+
+\begin{enumerate}
+   \item[(A)] A new CETI appears (Awakening)
+   \item[(D)] An old CETI dissappears (Doomsday)
+   \item[(C)] A new causal contact is stablished (Contact)
+   \item[(B)] An existing causal contact is interrupted (Blackout)
+\end{enumerate}
+
+The event of type B is produced when one of the two CETIs halts in its
+capability of emmiting and receiving signals.
+
+\begin{figure*} % 2D color plot
+   \centering
+   \includegraphics[width=\textwidth]{matrix_1.pdf}
+   \caption{Rate of CETIs that never make contact (listening) as a
+   function of $\tau_a$ and $\tau_s$, for 
+   $D_{max}$=10000 (left panel),
+   $D_{max}$=40000 (middle panel), and
+   $D_{max}$=80000 (right panel)}
+   \label{F_never_contact}
+\end{figure*}
+
+\begin{figure*} % 2D color plot
+   \centering
+   \includegraphics[width=\textwidth]{matrix_2.pdf}
+   \caption{Rate of CETIs that listen at the moment of the
+   awakening, as a funcion of $\tau_a$ and $\tau_s$, for
+   $D_{max}$=10000 (left panel),
+   $D_{max}$=40000 (middle panel), and
+   $D_{max}$=80000 (right panel)}
+   \label{F_C_at_A}
+\end{figure*}
+
+The system is updated each time an event is produced.
+%
+The time marks for each event (A and D for each CETI and C and B for each causal
+contact) are stored as a result of the simulation.
+%
+Also, the list of active CETIs is obtained as a function of time.
+%
+Some fixed variables must be set in order to carry out the simulation: the size
+of the Galactic Habitable Zone, the mean lifetime of a CETI, $\tau_s$, the mean
+waiting time for the appeareance of another CETI, $\tau_a$, and the maximum
+distance a message can be detected.
+
+The Galactic Habitable Zone is the only one of the four parameters that is
+barely known, is set at fixed values of the inner and outer radii, radial
+symetry is assumed, and the with of the galactic disk is neglected compared to
+the radial size.
+%
+The two radii are measured in light years, with the aim to maintain a single and
+comprehensive unit for both space and time coordinates.
+
+The simulation was implemented on a python 3 code, which is publically
+available\footnote{https://github.com/mlares/simu\_contact}
+
+%}}}
+
+\begin{figure}
+   %
+   \centering
+   %
+   \includegraphics[width=0.5\textwidth]{waiting_s+a_dif_ylog.pdf}
+   %
+   \caption{Histograms of the mean waiting times for the first contact, for
+   several models.
+   %
+   Left panel shows the histograms for several values of $\tau_a$, and
+   $\tau_s$ in the range 17000-24000 yr.
+   %
+   Right panel shows the histograms for several values of $\tau_s$, and
+   $\tau_a$ in the range 68000-100000 yr.
+   %
+   The shape of the histograms does not change when $\tau_a$ is
+   modified, but significantly changes when $\tau_s$ is modified.
+   }
+   %
+   \label{F_waiting_for_1C}
+   %
+\end{figure}
+
+\section{Results: exploring the parameter space}\label{S_results}
+%{{{
+
+We implemented the simulation of a regular grid of models varying over
+the hypothesis space, which covers 15000 different models.
+%
+For each model, we simulated 50 realizations with different random
+seeds, in order to estimate the variance.
+%
+The simulated parameters, $\tau_a$, $\tau_s$ cover the ranges
+4000-200000 and 10000-500000, respectively, with a partition of 50
+values for each parameter.
+%
+For the D$_{max}$ parameter, we chose the values 500, 1000, 10000,
+20000, 40000 and 80000 lyr.
+%
+This setup makes a total of 750000 simulations, each covering a time
+range of one million years.
+
+As a product of the simulations, several quantities are obtained.
+%
+Some quantities are directly derived from the discrete events, namely,
+the ID of emitting CETI, the ID of emitting (receiving) CETI, the
+position in the galaxy, the time of Awakening (Contact) and the time
+of Doomsday  (Blackout) for each CETI.
+%
+This structure is always for the A-D cycle of each emitting CETI, and
+repeats for each contact (C-B).
+%
+We can also derive the additional quantities.
+%
+The age of a CETI is the time elapsed from A to D.
+%
+%
+%given time, ceti\_e is the CETI emitting signals (emiter), ceti\_r is
+%the CETI receiving signals (receiver), ceti\_c is the ceti that
+%listen at least another ceti (citizen) and ceti\_h is the CETI ceti
+%that is lestened by at least another CETI.
+%
+The time span of a CETI listening another or being listened by another
+can also be easily derived from the results of a realization.
+%
+This way we can also compute the distribution in the galaxy of cetis
+that reach contact and the waiting time until the first contact or the
+waiting time until the next contact.
+%
+Regarding the properties of the population of CETIs and its evolution,
+we analyze the fraction of awaken time a ceti is listening at least
+another ceti, the age of contacted CETIs at first contact, the
+distribution of time to wait until next contact, the fraction of cetis
+where the first contact is given at the awakening, the distribution of
+the number of contacts for each CETI, the distribution of the maximum
+number of contacts for each CETI, the distribution of the number of
+contacts as a function of CETI age, the number of contacts as a
+function of time in the galaxy, the rate of CETIs that succeed in
+contact, and the distribution of distances between contacted CETIs.
+%
+Usefull derived quantity is the duration of two-way communication
+channels the or the fraction of contacts that admit a response.
+% 
+We also analyze the relations between the distance to ceti vs. time of
+double communication, the distance to ceti vs. age of contacted ceti,
+the age of a ceti and the maximum number of contacted cetis before D,
+or the lifespan of a ceti vs. max number of contacts.
+%
+All these quantities can be analyzed as a function of the simulation
+parameters.
+%
+We chose eight models that are on fairly oposed regions of the
+explored parameter space, and cover short/long lifetimes, short/long
+waiting times for the next CETI to appear in the Galaxy and
+short/large range reach of the signal, including all possible
+combinations.
+%
+The details of each model are summarized in the
+Table~\ref{T_selected_models}.
+
+In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
+distributions of the number contacts for CETIs in four different
+samples (M5, M6, M7, M8) with long range signal reach
+(D$_{max}$=10000). 
+%
+The lifetime is more determinant than the awakening rate in increasing
+the number of contacts.
+%
+As expected, of the four samples, the sample M7, which corresponds to
+a dense awakening in the timeline and a long lifetime, is the model
+that maximizes the number of contacts, reaching a maximum of order 100
+contacts for a single CETI.
+%
+This is considerable larger than any model with a shorter lifetime,
+which produce a number of contacts of at most the order of ten
+contacts per CETI.
+%
+A couple of comments are worth to mention about this plot.
+%
+First, it has a logarithmic scale on the x-axis, and second it is the
+cumulative, not differential, empirical distribution.
+
+In what follows, we focus on the analysis of the duration of contacts
+(Sec.~\ref{SS_multiplicity}), the waiting time for the first contact
+(Sec.~\ref{SS_waiting}) and the
+variations of contact posibilities as a function of the position of
+the CETI in the Galaxy (Sec.~\ref{SS_location}).
+
+%}}}
+
+\subsection{Multiplicity and duration of contacts}\label{SS_multiplicity}
+%{{{
+
+The duration of each contact for a single receiving CETI is given by the
+time difference between a C event and a B event, which are direct
+outputs for each simulation experiment.
+
+%}}}
+
+\subsection{Waiting time for a first contact}\label{SS_waiting}
+%{{{
+
+In this section we analyze the distribution of the waiting times for
+a first contact.
+%
+%Such distribution can be considered to compute the probability of
+%reaching a first contact within the first XX years of a SETI project,
+%assuming the methods are correct and the hypotheses of the experiment.
+Such distribution can be considered to compute the probability for a
+random CETI to carry out a SETI project and spend a given time until
+the first contact is made.
+
+In the Fig.~\ref{F_waiting_time} we show the fraction of the CETIs
+that have made at least one contact as a funtion of the elapsed time
+since the awakening, given that they make at least one contact on
+their lifetime, for different models.
+%
+From a frequentist approach, the normalized cumulative counts are
+related to the estimation of the probability of observing a given time
+with no success.
+%             
+Both panels show the histograms of the waiting times for the first
+contact, for all the CETIs that make at least one contact along their
+lifetime.
+%
+We show separately the histograms for several values of $\tau_a$ and a
+fixed value of $(\tau_s, D_{max})$, and for several values of $\tau_s$
+and a fixed value of $(\tau_a, D_{max})$.
+%
+For different values of $\tau_a$, the shape of the distribution is
+similar, and the main difference given by the different number of
+contacts on the different models.
+%
+Conversely, for different values of $\tau_a$, the shape of the
+distribution is similar, and the main difference given by the
+different number of contacts on the different models. 
+%
+We show the 2D maps of the rate of CETIs that never make contact and
+the rate of CETIs that make contact at the moment of the awakening,
+inf the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
+as a function of $\tau_s$ and $\tau_s$.
+%
+In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
+waiting times for the first contact, for several models.
+%
+Left panel shows the histograms for several values of $\tau_a$, and
+$\tau_s$ in the range 17000-24000 yr.
+%
+Right panel shows the histograms for several values of $\tau_s$, and
+$\tau_a$ in the range 68000-100000 yr.
+%
+The shape of the histograms does not change when $\tau_a$ is modified,
+but significantly changes when $\tau_s$ is modified.
+
+%}}}
+
+\begin{figure}
+   %
+   \centering
+   %
+   \includegraphics[width=0.45\textwidth]{galactocentric_frac.pdf}
+   %
+   \caption{Fraction of CETIs making cero, one, two or three contacts
+   as a function of galactocentric distance.}
+   %
+   \label{F_radial_frac}
+   %
+\end{figure}
+
+\subsection{Location within the Galaxy}\label{SS_location}
+%{{{
+
+On the Figure~\ref{F_radial_frac} we show the fraction of CETIs that make
+zero, one, two or three contacts, as a function of the galactocentric
+distance, for model M1.
+% VERIFICAR!  M1?
+%
+For this model, the majority of CETIs do not
+make any contact during all the time period in which they are active.
+%
+As expected, the fractions of CETIs making at least one contact
+diminishes as the number of contacts rises.
+%
+The shape of this distribution is a result of the geometry of the GHZ.
+%
+In general, a closer position to the center of the Galaxy slightly favours 
+the posibility of a contact.
+%
+This would indicate that the position of the Earth is suitable for the
+odds of making a contact.
+
+%}}}
+
+\setlength{\tabcolsep}{10pt}
+\begin{table*}
+\centering
+\begin{tabular}{ccccc}
+\hline
+Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & description  \\
+\hline
+M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
+M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
+M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
+M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
+%
+M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
+M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
+M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
+M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
+%
+\hline
+\end{tabular}
+\caption{Selected models to analize the behevior of simulation outputs
+   and their dependence on simulation outputs and their dependence on
+   simulation parameters.}
+\label{T_selected_models}
+\end{table*}
+
+\section{Discusion}\label{S_discussion}
+%{{{
+
+We have implemented a simulation approach to explore the hypothesis
+space about the causal connections between communicating
+extraterrestrial civilizations.
+%
+The different models can be generated with a minimal number of free
+parameters and some fixed assumptions.
+%
+We argue that three parameters can be used to describe a variety of
+situations, ranging from a Galaxy where an intelligent civilization is
+very rare, to a Galaxy populated with plenty of civilizations in
+causal contact.
+%
+We explored the consecuences of different models using mensurable
+quantities.
+%
+With this approach, we can estimate several quantities as a function
+of the parameters on the hypothesis space.
+%
+For instance, it is possible to estimate the timescale for an
+uninterrupted SETI effort to reach success, in terms of different
+model parameters that reflect different, so far unknown, scenarios
+for the spread if intelligent life on the Galaxy.
+
+Under the hypotheses of our experiments, we conclude that
+a causal contact is extremely unlikely unless the Galaxy is heavily
+populated by intelligent civilizations.
+%
+Even so, we must consider that in order to stablish a contact between
+any two entities, a minimum degree of compatibility must be
+accomplished without any previous agreement, 
+making the posibility of a contact with a message that
+could be deciphered highly rare.
+
+Future work can be performed following this framework, in order to
+explore possible implications of the results for more detailed 
+configurations of the experiment.
+%
+For example, the communication method (isotropic, colimated,
+serendipitous) can affect the observables.
+%
+Also the nature of the message carrier, electromagnetic or another.
+%
+It is also possible to explore the effects of alignments or the use of
+stars as sources or amplifiers.
+%
+%energetic markers
+%Spiritual markers
+%
+As a future work, we will explore several improvements to the
+simulation, and analyze the impact on the final results of these
+changes.
+%
+They include a spatial distribution that actually resembles the shape
+of the Galaxy, different distribution functions for the mean lifetime
+of CETIs and different efficiencies for the sphere of the causal
+region of each CETI, resembling different searching strategies.
+
 Possible hypothesis improvements are:
@@
 
-\subsection{Model for CETIs}
-
-Since the method is based on events, the first step is to define an
-architecture of events, the relationships between them and how events
-trigger changes on the state of the system.
-%
-We consider a simulated system that represents the galactic habitable
-zone (GHZ).
-%
-On a first approach, the GHZ is a 2-dimensional annular region.
-%
-The adopted values for the GHZ are 20 kly and 60 kly, for the inner
-and outer radius, respectively \citep{Lineweaver2004}.
-% ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
-% mas chico...
-%
-This simple model does not take into accont the variations in stellar
-density given by the spiral structure.
-%
-Although there are several possible approaches, we chose to follow the
-evolution of the system according to the following events:
-
-\begin{enumerate}
-   \item[(A)] A new CETI appears (Awakening)
-   \item[(D)] An old CETI dissappears (Doomsday)
-   \item[(C)] A new causal contact is stablished (Contact)
-   \item[(B)] An existing causal contact is interrupted (Blackout)
-\end{enumerate}
-
-The event of type B is produced when one of the two CETIs halts in its
-capability of emmiting and receiving signals.
-
-%======================================================================
-%======================================================================
-%======================================================================
-
-The system is updated each time an event is produced.
-%
-The time marks for each event (A and D for each CETI and C and B for
-each causal contact) are stored as a result of the simulation.
-%
-Also, the list of active CETIs is obtained as a function of time.
-%
-Some fixed variables must be set in order to carry out the simulation,
-namely:
-
-\begin{itemize}
-   \item Size of the Galactic Habitable Zone,
-   \item mean lifetime of a CETI, $\lambda_1$,
-   \item mean waiting time for the appeareance of another CETI,
-      $\lambda_2$, and
-   \item maximum distance a message can be detected.
-\end{itemize}
-
-The Galactic Habitable Zone is the only one of the four parameters
-that is barely known, is set at fixed values of the inner and
-outer radii, radial symetry is assumed, and the with of the
-galactic disk is neglected compared to the radial size.
-%
-The two radii are measured in light years, with the aim to maintain a
-single and comprehensive unit for both space and time coordinates.
-
-The simulation was implemented on a python code, which is publically
-available\footnote{https://github.com/mlares/simu\_contact}
-
-%Variables en el programa
-%\begin{itemize}
-%   \item Nact: Número de CETIs activas
-%   \item Nsee: Número de contactos por cada CETI
-%   \item Nccc: CETIs en contacto causal. Notar que una CETI puede tener múltiples
-%contactos de otras CETIs. Para cada una de ellas, guardar:
-%   \item ID: ID de la CETI emisora
-%   \item ID: ID de la CETI receptora
-%   \item t\_hola: tiempo del primer contacto
-%   \item t\_chau: tiempo del último contacto
-%\end{itemize}
-%
-%
-%Caso 1: aparece una nueva CETI
-%Modificaciones al estado del sistema:
-%\begin{itemize}
-%   \item Agregar la CETI “A” a la lista de CETIs activas
-%   \item Sortear la duración de la actividad de comunicación de la CETI “A”
-%   \item Sortear el intervalo de tiempo hasta la aparición de la próxima CETI:
-%t\_WakeUp\_next
-%   \item Agregar la CETI “A” al árbol
-%   \item Identificar todas las CETIs que están a menos de D max de la nueva CETI, y
-%actualizar la lista de tiempos
-%\end{itemize}
-%
-%Caso 2: desaparece una CETI
-%Modificaciones al estado del sistema:
-%\begin{itemize}
-%   \item Eliminar la CETI a la lista de CETIs activas
-%\end{itemize}
-%
-%
-%Caso 3: una nueva CETI entra en contacto causal
-%Modificaciones al estado del sistema:
-%\begin{itemize}
-%   \item Agregar la CETI B a la lista de contactos de la CETI A
-%\end{itemize}
-%
-
 %}}}
-
-\begin{figure*}
-   \centering
-   \includegraphics[width=0.8\textwidth]{Messages_01.pdf}
-   \caption{Schematic representation of the possible cases in which a
-   ceti A can be in causal contact with a ceti B which appears later.
-   The duration of the causal contact in one direction depends on
-   several factors, mainly $\Delta t_A$, $\Delta t_C$, and $\Delta X$.}
-   \label{F_messages}
-\end{figure*}
-
-\section{Results: exploring the parameter space}\label{S_results}
-%{{{
-
-We implemented the simulation of a regular grid of models varying over
-the hypothesis space, which covers 15000 different models.
-%
-For each model, we simulated 50 realizations with different random
-seeds, in order to estimate the variance.
-%
-The simulated parameters, $\tau_a$, $\tau_s$ cover the ranges
-4000-200000 and 10000-500000, respectively, with a partition of 50
-values for each parameter.
-%
-For the D$_{max}$ parameter, we chose the values 500, 1000, 10000,
-20000, 40000 and 80000 lyr.
-%
-This setup makes a total of 750000 simulations, each covering a time
-range of one million years.
-
-As a product of the simulations, several quantities are obtained.
-%
-Some quantities are directly derived from the discrete events, namely,
-the ID of emitting CETI, the ID of emitting (receiving) CETI, the
-position in the galaxy, the time of Awakening (Contact) and the time
-of Doomsday  (Blackout) for each CETI.
-%
-This structure is always for the A-D cycle of each emitting CETI, and
-repeats for each contact (C-B).
-%
-We can also derive the additional quantities.
-%
-The age of a CETI is the time elapsed from A to D.
-%
-%
-%given time, ceti\_e is the CETI emitting signals (emiter), ceti\_r is
-%the CETI receiving signals (receiver), ceti\_c is the ceti that
-%listen at least another ceti (citizen) and ceti\_h is the CETI ceti
-%that is lestened by at least another CETI.
-%
-The time span of a CETI listening another or being listened by another
-can also be easily derived from the results of a realization.
-%
-This way we can also compute the distribution in the galaxy of cetis
-that reach contact and the waiting time until the first contact or the
-waiting time until the next contact.
-%
-Regarding the properties of the population of CETIs and its evolution,
-we analyze the fraction of awaken time a ceti is listening at least
-another ceti, the age of contacted CETIs at first contact, the
-distribution of time to wait until next contact, the fraction of cetis
-where the first contact is given at the awakening, the distribution of
-the number of contacts for each CETI, the distribution of the maximum
-number of contacts for each CETI, the distribution of the number of
-contacts as a function of CETI age, the number of contacts as a
-function of time in the galaxy, the rate of CETIs that succeed in
-contact, and the distribution of distances between contacted CETIs.
-%
-Usefull derived quantity is the duration of two-way communication
-channels the or the fraction of contacts that admit a response.
-% 
-We also analyze the relations between the distance to ceti vs. time of
-double communication, the distance to ceti vs. age of contacted ceti,
-the age of a ceti and the maximum number of contacted cetis before D,
-or the lifespan of a ceti vs. max number of contacts.
-%
-All these quantities can be analyzed as a function of the simulation
-parameters.
-%
-We chose eight models that are on fairly oposed regions of the
-explored parameter space, and cover short/long lifetimes, short/long
-waiting times for the next CETI to appear in the Galaxy and
-short/large range reach of the signal, including all possible
-combinations.
-%
-The details of each model are summarized in the
-Table~\ref{T_selected_models}.
-
-In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
-distributions of the number contacts for CETIs in four different
-samples (M5, M6, M7, M8) with long range signal reach
-(D$_{max}$=10000). 
-%
-The lifetime is more determinant than the awakening rate in increasing
-the number of contacts.
-%
-As expected, of the four samples, the sample M7, which corresponds to
-a dense awakening in the timeline and a long lifetime, is the model
-that maximizes the number of contacts, reaching a maximum of order 100
-contacts for a single CETI.
-%
-This is considerable larger than any model with a shorter lifetime,
-which produce a number of contacts of at most the order of ten
-contacts per CETI.
-%
-A couple of comments are worth to mention about this plot.
-%
-First, it has a logarithmic scale on the x-axis, and second it is the
-cumulative, not differential, empirical distribution.
-
-In what follows, we focus on the analysis of the duration of contacts
-(Sec.~\ref{SS_multiplicity}), the waiting time for the first contact
-(Sec.~\ref{SS_waiting}) and the
-variations of contact posibilities as a function of the position of
-the CETI in the Galaxy (Sec.~\ref{SS_location}).
-
-%}}}
-
-\subsection{Multiplicity and duration of contacts}\label{SS_multiplicity}
-%{{{
-
-The duration of each contact for a single receiving CETI is given by the
-time difference between a C event and a B event, which are direct
-outputs for each simulation experiment.
-
-%}}}
-
-\subsection{Waiting time for a first contact}\label{SS_waiting}
-%{{{
-
-In this section we analyze the distribution of the waiting times for
-a first contact.
-%
-%Such distribution can be considered to compute the probability of
-%reaching a first contact within the first XX years of a SETI project,
-%assuming the methods are correct and the hypotheses of the experiment.
-Such distribution can be considered to compute the probability for a
-random CETI to carry out a SETI project and spend a given time until
-the first contact is made.
-
-In the Fig.~\ref{F_waiting_time} we show the fraction of the CETIs
-that have made at least one contact as a funtion of the elapsed time
-since the awakening, given that they make at least one contact on
-their lifetime, for different models.
-%
-From a frequentist approach, the normalized cumulative counts are
-related to the estimation of the probability of observing a given time
-with no success.
-%             
-Both panels show the histograms of the waiting times for the first
-contact, for all the CETIs that make at least one contact along their
-lifetime.
-%
-We show separately the histograms for several values of $\tau_a$ and a
-fixed value of $(\tau_s, D_{max})$, and for several values of $\tau_s$
-and a fixed value of $(\tau_a, D_{max})$.
-%
-For different values of $\tau_a$, the shape of the distribution is
-similar, and the main difference given by the different number of
-contacts on the different models.
-%
-Conversely, for different values of $\tau_a$, the shape of the
-distribution is similar, and the main difference given by the
-different number of contacts on the different models. 
-%
-We show the 2D maps of the rate of CETIs that never make contact and
-the rate of CETIs that make contact at the moment of the awakening,
-inf the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
-as a function of $\tau_s$ and $\tau_s$.
-%
-In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
-waiting times for the first contact, for several models.
-%
-Left panel shows the histograms for several values of $\tau_a$, and
-$\tau_s$ in the range 17000-24000 yr.
-%
-Right panel shows the histograms for several values of $\tau_s$, and
-$\tau_a$ in the range 68000-100000 yr.
-%
-The shape of the histograms does not change when $\tau_a$ is modified,
-but significantly changes when $\tau_s$ is modified.
-
-%}}}
-
-\subsection{Location within the Galaxy}\label{SS_location}
-%{{{
-
-On the Figure~\ref{F_radial_frac} we show the fraction of CETIs that make
-zero, one, two or three contacts, as a function of the galactocentric
-distance, for model M1.
-% VERIFICAR!  M1?
-%
-For this model, the majority of CETIs do not
-make any contact during all the time period in which they are active.
-%
-As expected, the fractions of CETIs making at least one contact
-diminishes as the number of contacts rises.
-%
-The shape of this distribution is a result of the geometry of the GHZ.
-%
-In general, a closer position to the center of the Galaxy slightly favours 
-the posibility of a contact.
-%
-This would indicate that the position of the Earth is suitable for the
-odds of making a contact.
-
-%}}}
-
-\setlength{\tabcolsep}{10pt}
-\begin{table*}
-\centering
-\begin{tabular}{ccccc}
-\hline
-Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & description  \\
-\hline
-M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
-M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
-M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
-M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
-%
-M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
-M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
-M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
-M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
-%
-\hline
-\end{tabular}
-\caption{Selected models to analize the behevior of simulation outputs
-   and their dependence on simulation outputs and their dependence on
-   simulation parameters.}
-\label{T_selected_models}
-\end{table*}
-
-\begin{figure} \centering
-   %
-   \includegraphics[width=0.5\textwidth]{F1.pdf}
-   %
-   \caption{Empirical cumulative distributions of the number contacts
-   for CETIs in four different samples (M5, M6, M7, M8) with long
-   range signal reach (D$_{max}$=10000).
-   %
-   }
-   % 
-\label{F_number_of_contacts} \end{figure}
-
-\begin{figure} % Distribution
-   \centering
-   \includegraphics[width=0.5\textwidth]{F2.pdf}
-   \caption{Rate of CETIs that make no contact as a function of
-   $\tau_a$ and $\tau_s$}
-   \label{F_waiting_time}
-\end{figure}
-
-\begin{figure*} % 2D color plot
-   \centering
-   \includegraphics[width=0.3\textwidth]{m1d2.png}%
-   \includegraphics[width=0.3\textwidth]{m1d3.png}%
-   \includegraphics[width=0.3\textwidth]{m1d4.png}
-   \caption{Rate of CETIs that never make contact (listening) as a
-   function of $\tau_a$ and $\tau_s$, for 
-   $D_{max}$=10000 (left panel),
-   $D_{max}$=40000 (middle panel), and
-   $D_{max}$=80000 (right panel)}
-   \label{F_never_contact}
-\end{figure*}
-
-\begin{figure*} % 2D color plot
-   \centering
-   \includegraphics[width=0.3\textwidth]{m2d2.png}%
-   \includegraphics[width=0.3\textwidth]{m2d3.png}%
-   \includegraphics[width=0.3\textwidth]{m2d4.png}
-   \caption{Rate of CETIs that listen at the moment of the
-   awakening, as a funcion of $\tau_a$ and $\tau_s$}
-   \label{F_C_at_A}
-\end{figure*}
-
-\section{Discusion}\label{S_discussion}
-%{{{
-
-We have implemented a simulation approach to explore the hypothesis
-space about the causal connections between communicating
-extraterrestrial civilizations.
-%
-The different models can be generated with a minimal number of free
-parameters and some fixed assumptions.
-%
-We argue that three parameters can be used to describe a variety of
-situations, ranging from a Galaxy where an intelligent civilization is
-very rare, to a Galaxy populated with plenty of civilizations in
-causal contact.
-%
-We explored the consecuences of different models using mensurable
-quantities.
-%
-With this approach, we can estimate several quantities as a function
-of the parameters on the hypothesis space.
-%
-For instance, it is possible to estimate the timescale for an
-uninterrupted SETI effort to reach success, in terms of different
-model parameters that reflect different, so far unknown, scenarios
-for the spread if intelligent life on the Galaxy.
-
-Under the hypotheses of our experiments, we conclude that
-a causal contact is extremely unlikely unless the Galaxy is heavily
-populated by intelligent civilizations.
-%
-Even so, we must consider that in order to stablish a contact between
-any two entities, a minimum degree of compatibility must be
-accomplished without any previous agreement, 
-making the posibility of a contact with a message that
-could be deciphered highly rare.
-
-Future work can be performed following this framework, in order to
-explore possible implications of the results for more detailed 
-configurations of the experiment.
-%
-For example, the communication method (isotropic, colimated,
-serendipitous) can affect the observables.
-%
-Also the nature of the message carrier, electromagnetic or another.
-%
-It is also possible to explore the effects of alignments or the use of
-stars as sources or amplifiers.
-%
-%energetic markers
-%Spiritual markers
-%
-As a future work, we will explore several improvements to the
-simulation, and analyze the impact on the final results of these
-changes.
-%
-They include a spatial distribution that actually resembles the shape
-of the Galaxy, different distribution functions for the mean lifetime
-of CETIs and different efficiencies for the sphere of the causal
-region of each CETI, resembling different searching strategies.
-
-%}}}
-
-\begin{figure}
-   %
-   \centering
-   %
-   \includegraphics[width=0.5\textwidth]{waiting_s+a_dif_ylog.pdf}
-   %
-   \caption{Histograms of the mean waiting times for the first contact, for
-   several models.
-   %
-   Left panel shows the histograms for several values of $\tau_a$, and
-   $\tau_s$ in the range 17000-24000 yr.
-   %
-   Right panel shows the histograms for several values of $\tau_s$, and
-   $\tau_a$ in the range 68000-100000 yr.
-   %
-   The shape of the histograms does not change when $\tau_a$ is
-   modified, but significantly changes when $\tau_s$ is modified.
-   }
-   %
-   \label{F_waiting_for_1C}
-   %
-\end{figure}
-
-\begin{figure}
-   %
-   \centering
-   %
-   \includegraphics[width=0.5\textwidth]{radial.png}
-   %
-   \caption{Fraction of CETIs making cero, one, two or three contacts
-   as a function of galactocentric distance.}
-   %
-   \label{F_radial_frac}
-   %
-\end{figure}
 
@@
 Plots and simulations were made with software developed by the authors
-using R and python languajes, plots were postprocessed with inkscape.
+using R and python languajes. Plots were postprocessed with inkscape.
 %

### 2019-08-23 manuscript.tex
hacia la version 0

@@
 signal within the conditions stablished by SETI.
+%
+This, by some astonishing result, has lead to the Fermi paradox, which
+states an apparent contradiction between the high estimation of the
+abundance of life in the Galaxy and the lack of their detection
+\citep{Anchordoqui2017}.
 
@@
 
-\begin{figure} % Distribution
-   \centering
-   \includegraphics[width=0.5\textwidth]{F2.pdf}
-   \caption{Rate of CETIs that make no contact as a function of
-   $\tau_a$ and $\tau_s$}
-   \label{F_waiting_time}
-\end{figure}
-
 %}}}
@@
 
-\begin{figure}
-   %
-   \centering
-   %
-   \includegraphics[width=0.45\textwidth]{galactocentric_frac.pdf}
-   %
-   \caption{Fraction of CETIs making cero, one, two or three contacts
-   as a function of galactocentric distance.}
-   %
-   \label{F_radial_frac}
-   %
-\end{figure}
+%   \begin{figure}
+%      %
+%      \centering
+%      %
+%      \includegraphics[width=0.45\textwidth]{galactocentric_frac.pdf}
+%      %
+%      \caption{Fraction of CETIs making cero, one, two or three contacts
+%      as a function of galactocentric distance.}
+%      %
+%      \label{F_radial_frac}
+%      %
+%   \end{figure}
 
@@
 
-On the Figure~\ref{F_radial_frac} we show the fraction of CETIs that make
-zero, one, two or three contacts, as a function of the galactocentric
-distance, for model M1.
+%On the Figure~\ref{F_radial_frac} we show the fraction of CETIs that make
+%zero, one, two or three contacts, as a function of the galactocentric
+%distance, for model M1.
 % VERIFICAR!  M1?
@@
 
-We have implemented a simulation approach to explore the hypothesis
-space about the causal connections between communicating
-extraterrestrial civilizations.
+We have implemented a suite of simulations following a stochastic
+approach, to explore the hypothesis
+space of a simple model that accounts for
+the causal connections between communicating
+civilizations on an simplified galaxy.
 %
@@
 We explored the consecuences of different models using mensurable
-quantities.
+quantities, that arise as by-products of the simulations.
 %
@@
 model parameters that reflect different, so far unknown, scenarios
-for the spread if intelligent life on the Galaxy.
+for the spread of intelligent life on the galaxy.
+
+Although the simulations use a number of speculative assumptions, we
+argue that the degree of knowledge about the origin and persistence of
+life in the Galaxy does not worth the implementation of more detailed
+or sophisticated models.
+%
+Instead, we take advantadge of the simplicity of the model to explore
+the hypothesis space, i.e., the three--dimensional parameter space of
+the parametric model, in order to gain insight on the consecuences of
+different scenarios for the search of intelligent life.
+%
+Our analysis is not centered in obtaining the odds for the Earth
+making contact with another extraterrestrial civilization, but in 
+obtaining a statistical, parameter dependent description of the
+communication system consisting of a group of broadcasting
+civilizations which arise sparsely both in time and space in the
+Galaxy.
 
@@
 
-Future work can be performed following this framework, in order to
+Taking into account the caveat mentioned previously about the
+trade--off between the simplicity and complexity of the experiment,
+future work can be performed following this framework in order to
 explore possible implications of the results for more detailed
-configurations of the experiment.
+configurations.
 %
@@
 %
+Since we analyze the causal contacts, effective contacts depend on the
+efficiency of both the emitter and the receiver.
+%
+A correction by coverage ratio in the detection and by a targetting
+ratio in the emission could be easily implemented in the simulation,
+although the effect would be to reduce the probability of contact by
+an efficiency ratio.
+%
+Thus, such implementation is not really necessary, since the values of
+probabilities could modified by a constant correction factor equal to
+the combined emission/reception efficiency.
+%
+If we consider a statistical distribution for this factor, then the
+results could change, and this could be studied with the DE approach.
+%
 Also the nature of the message carrier, electromagnetic or another.
@@
 %
-%energetic markers
-%Spiritual markers
+%energetic markers Spiritual markers
 %
@@
       GHZ.
-   \item Correction by coverage ratio in the detection and by
-      targetting ratio in the emission.
    \item Spiritual factor.  Explore the role of the S factor: message
@@
 \end{document}
-
-DEVELOPMENT DEVELOPMENT  DEVELOPMENT  DEVELOPMENT  DEVELOPMENT  DEVELOPMENT
-
-% Papers para citar y discutir:
-
-%## GHZ
-
-%## D_max
-
-%## exponential distribution
-
-%## Lifetime
-
-%## Simulations: cosmological
-
-Forgan2015
-
-%## Simulations: stochastic
-
-grimaldi_signal_2017
-
-%## Temporal aspects
-Temporal aspects of the interaction among the first galactic civilizations: The 'interdict hypothesis'
-by M.J. Fogg
-
-%## Drake equation
-
-%## Detalles de la simulacion
-Galactic civilizations: Population dynamics and interstellar diffusion
-by W.I. Newman, C. Sagan

### 2019-08-23 purgatorio.tex
hacia la version 0

Alien civilizations: Good or bad?.  The spiritual factor

Indicators of the level of spirituality.

Since it is not possible to directly measure the spiritual factor, the
strategy would be to search for signs of the consequences of a moral or
spiritual nature in alien civilizations.

Lets call the indicator as "E-factor", standing for "spiritual factor".

In a first approach, civilizations can be spiritual based (E+) or with absence of spirituality (E-).

HOW DOES THE E-FACTOR AFFECTS THE SURVIVAL OF A CIVILIZATION?

Is it possible that an E+ civilization has more chances to survive
than an E- civilization?

CHARACTERISTICS

*** Civilizations with spiritual values (E+)

+ Seek for other civilizations to deepen its understanding of "divine plans"
+ Are tolerant to other emerging civilizations
+ Are tolerant to other ideas or organizations (?)
+ Are eager to share its vision about spirituality.

*** Civilizations without spiritual values (E-)

+ Pursue survival at all cost, even when it requires que conquer or enslave other species
+ Aggressive?

*** Characteristics independent of the E-factor

+ Use of natural resources
+ Development of science
+ Development of technology
+ Exploration of space

+ Is it possible for a E+ type civilization to have some of the E- type characteristics?   And viceversa?

+ Does the lifetime depend on the E-type?

+ Is it possible that an E+ civilization discovers a communication channel
that is unknown for E- civilizations?

WHY SEARCHING FOR ALIEN LIFE? (LISTENING)

+ Spiritual or religious reasons
+ Search for knowledge
+ Need of habitable locations or resources
+ monitor developing civilizations as a "galactic censorship"

WHY CALLING FOR ALIEN LIFE?

+ Spiritual or religious reasons
+ Search for knowledge
+ Need of habitable locations or resources
+ S.O.S. calls
+ Warnings for other civilizations about threats
+ Use messages as decoys to identify emerging civilizations

WHY NOT CALLING FOR ALIEN LIFE?

+ Fear of more powerfull and aggressive species looking for resources

COMMUNICATION CHANNELS

The emmision, reception and decodification are all necessary to complete qye
communication process.

+ What if all are listening but no one is talking?
+ Is it possible for an advanced civilization to avoid emmisions?

A STOCHASTIC PROCESS APPROACH

Hipothesis 1:  Aggressive spreading

Les supose that there are plenty of advanced civilizations in the
galaxy, some of them of E+ type and som of E- type.  If one type is
aggressive (E-) and the other is not (E+), would it mean (according to
"characteristics") that after some time E- types will prevail?

May be an aggressive civilization can not reach another distant
planet, although it knows its location and value.   Taking this into
account, the growth of aggressive species requires a fast travel
device.  Otherwise, its possible military power is not usefull.

If these civilizations discover a device to quickly travel to target
planets, the galaxy would be fully populated with this species and we
would have already made contact.   This evidence would then disfavour
this hypothesis?

Hipothesis 2:  The evil assumption

Given the total lack of knowledge about other species, a conservative
position would be to be a "quiet" planet, as long as it is possible,
just in case nearby civilizations are looking for planets with
valuable resources.   The fact that we have not detected any signs of
life in the universe outside Earth would favour this hypothesis but
can not be used to discard it or its opposite.

Hipothesis 3:  Spiritual independence

It is also possible that the survival of a civilization does not
depend at all on the development of a spiritual awareness.

It is also possible that the degree of hostility does not depend on
the E-factor, or that it is a consequence of a lack of undestanding or
sympathy towards less advanced species.  For example, humankind would
exploit resources of a solar system planet, even if it implies the
destruction of a less advanced "intelligent civilization" that we can
not undertand.  We would see them as "animals" and would wipe them out
if nececessary, albeit we can be considered as E+ type.

% Papers para citar y discutir:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

https://www.theguardian.com/science/2016/sep/23/stephen-hawking-warns-against-seeking-out-aliens-in-new-film

http://adsabs.harvard.edu/abs/2016arXiv161102586T

http://adsabs.harvard.edu/abs/1959Natur.184..844C

Searching for Interstellar Communications
by Cocconi, Giuseppe; Morrison, Philip

http://adsabs.harvard.edu/abs/2002PASP..114..416R

Optical Search for Extraterrestrial Intelligence: A Spectroscopic Search for Laser Emission from Nearby Stars
by Reines, Amy E.; Marcy, Geoffrey W.

http://adsabs.harvard.edu/abs/2001ARA%26A..39..511T

The Search for Extraterrestrial Intelligence (SETI)
by Tarter, Jill

%## GHZ

Morrison & Gowanlock
Extending Galactic Habitable Zone Modelling to Include the Emergence of Intelligent Life
https://ui.adsabs.harvard.edu/abs/2014ebi..conf..5.4M/abstract

Javaux & Dehant
Habitability: from stars to cells
https://ui.adsabs.harvard.edu/abs/2010A%26ARv..18..383J/abstract

Schwieterman et al.
A Limited Habitable Zone for Complex Life
https://ui.adsabs.harvard.edu/abs/2019ApJ...878...19S/abstract

%## Simulation parameters

%  D_max

%  tau_a

%  tau_s

Legassick 2015
The Age Distribution of Potential Intelligent Life in the Milky Way
https://ui.adsabs.harvard.edu/abs/2015arXiv150902832L/abstract

% hypotheses

W.I. Newman, C. Sagan 
Galactic civilizations: Population dynamics and interstellar diffusion
https://www.sciencedirect.com/science/article/pii/0019103581901354

%## SETI strategy: from causal contact to communication

% stars as amplifiers:

E.F. Borra & E. Trottier
Discovery of peculiar periodic spectral modulations in a small fraction of solar type stars
https://arxiv.org/abs/1610.03031

by Tamburini, Fabrizio; Licata, Ignazio
Can the periodic spectral modulations of the 236 SETI candidate Sloan Survey stars be due to Dark Matter effects?

Borra, Ermanno F.
Searching for Extraterrestrial Intelligence Signals in Astronomical Spectra, Including Existing Data
http://adsabs.harvard.edu/abs/2012AJ....144..181B

### 2019-08-27 manuscript.tex
ordenando la bibliografia

@@
 intelligent civilizations (ETI) has been the use of the Drake equation
-\citep{Gleiser2010, Prantzos2013, Haqq-Misra2017}.
+\citep{gleiser_drake_2010, prantzos_joint_2013, haqq-misra_drake_2017}.
 %
@@
 and their contacts have also been explored by several authors
-\citep{Fogg1987, Forgan2011, Balbi2018},
+\citep{fogg_temporal_1987,forgan_spatiotemporal_2011, balbi_impact_2018},
 %The Impact of the Temporal Distribution of Communicating Civilizations on Their Detectability
@@
 as well as efforts on considering the stochastic nature of the Drake equation
-\citep{Glade2011}.
+\citep{glade_stochastic_2011}.
 %
 The simulation approach has also benn considered
-\citep{Forgan2008, Forgan2010}.
+\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
+murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017}
 %New numerical determination of habitability in the Galaxy: the SETI connection
@@
 abundance of life in the Galaxy and the lack of their detection
-\citep{Anchordoqui2017}.
+\citep{anchordoqui_isthere_2017}.
 
@@
 
-In a recent work, \citep{Balbi2018} use a statistical model to analyze
+In a recent work, \citep{balbi_impact_2018} use a statistical model to analyze
 the occurrence of causal contacts between civilizations in the Galaxy.
@@
 \citep{ptolemaeus_system_2014, chung_simulation_2003,
-simulation_ross_2012}.
+ross_simulation_2012}.
 %
@@
 %
-\citet{Balbi2018} stress the fact that it would be desirable to arrive
+\citet{balbi_impact_2018} stress the fact that it would be desirable to arrive
 at a theoretical statistical distribution of the lifespan of
@@
 %
-\citet{Maccone} argues that this distribution should be a log normal.
+\citet{maccone_lognormals_2014} argues that this distribution should be a log normal.
 (EXPAND)
@@
 files and word frequency in languages \citep{mitzenmacher_brief_2004,
-Newman_power_2005, simkin_theory_2006}
+newman_power_2005, simkin_theory_2006}
 %
@@
 \citep{jeong_large_2000} and in the number of species observed per
-unit area in ecology \citep{garcia_origin_2006, frank_common_2009}.
+unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
 
@@
 The adopted values for the GHZ are 20 kly and 60 kly, for the inner
-and outer radius, respectively \citep{Lineweaver2004}.
+and outer radius, respectively \citep{lineweaver_galactic_2004}.
 % ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
@@
 \end{tabular}
-\caption{Selected models to analize the behevior of simulation outputs
+\caption{Selected models to analize the behavior of simulation outputs
    and their dependence on simulation outputs and their dependence on
@@
 
+\nocite{barabasi_scale_2009}
+\nocite{barlow_galactic_2012}
+\nocite{bloetscher_using_2019}
+\nocite{burchell_whither_2006}
+\nocite{conway_three_2018}
+\nocite{dayal_habitability_2016}
+\nocite{enriquez_breakthrough_2017}
+\nocite{forgan_exoplanet_2017}
+\nocite{forgan_possibility_2013}
+\nocite{forgan_collimated_2014}
+\nocite{forgan_numerical_2010}
+\nocite{funes_searching_2019}
+\nocite{glade_stochastic_2012}
+\nocite{gonzalez_habitable_2005}
+\nocite{gonzalez_galactic_2001}
+\nocite{grimaldi_signal_2017}
+\nocite{haqq-misra_evolution_2019}
+\nocite{harp_application_2018}
+\nocite{hinkel_interdisciplinary_2019}
+\nocite{hippke_interstellar_2017}
+\nocite{horvat_calculating_2007}
+\nocite{horvat_impact_2011}
+\nocite{lampton_information_2013}
+\nocite{lingam_relative_2019}
+\nocite{loeb_relative_2016}
+\nocite{loeb_eavesdropping_2006}
+\nocite{maccone_SETI_2013}
+\nocite{maccone_evolution_2014}
+\nocite{maccone_mathematical_2011}
+\nocite{maccone_statistical_2015}
+\nocite{maccone_KLT_2010}
+\nocite{maccone_statistical_2010}
+\nocite{maccone_SETI_2011}
+\nocite{morrison_extending_2015}
+\nocite{morrison_extending_2015}
+\nocite{newman_galactic_1981}
+\nocite{peters_outer_2018}
+\nocite{rahvar_cosmic_2016}
+\nocite{seager_search_2018}
+\nocite{shostak_what_2010}
+\nocite{smith_broadcasting_2009}
+\nocite{solomonides_probabilistic_2016}
+\nocite{starling_virulence_2013}
+\nocite{tarter_search_2001}
+\nocite{tarter_advancing_2009}
+\nocite{vukotic_astrobiological_2012}
+\nocite{walters_interstellar_1980}
+\nocite{wright_theGsearch_2015}
+\nocite{anchordoqui_upper_2019}
+\nocite{balb_spatiotemporal_2018}
+\nocite{forgan_galactic_2016}
+\nocite{gobat_evolution_2016}
+
 \bibliography{biblio_seti}

### 2019-08-28 manuscript.tex
biblio...

@@
 %
+The abundance of intelligent civilizations in the galaxy is a
+longstanding question, which is often conceptualized as the problem of
+the lack of received communication or the Fermi paradox
+\citep{barlow_galactic_2012, Sotos_biotechnology_2019, 
+forgan_galactic_2016}.
+There are many proposals intended to resolve this paradox, which make
+use of statistical \citep{solomonides_probabilistic_2016,
+vanhouten_isthere_2017, horvat_calculating_2007, maccone_statistical_2015} or
+stochastic approaches \citep{forgan_numerical_2009,
+bloetscher_using_2019, glade_stochastic_2011, forgan_numerical_2010},
+analytical interpretations of the Drake equation \citep{prantzos_joint_2013, smith_broadcasting_2009}
+or more speculative proposals
+\citep{barlow_galactic_2013, lampton_information_2013, conway_three_2018, forgan_galactic_2017}.
+%
 However, the uncertainties in the factors make it less prone to a
 formal application to define searching strategies or to compute
-actual numbers, but is used to organize the discussion instead.
+numbers, but is used to organize the discussion instead
+\citep{hinkel_interdisciplinary_2019}.
 %
@@
 and their contacts have also been explored by several authors
-\citep{fogg_temporal_1987,forgan_spatiotemporal_2011, balbi_impact_2018},
+\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
+balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011},
 %The Impact of the Temporal Distribution of Communicating Civilizations on Their Detectability
@@
 Since the factors in the Drake equation are uncertain, we propose to
-avoid the frequentist approach of this equation
-and to explore a parameter space, where instead of computing a final
-number, we provide a statistical distribution that gives conditional
-probabilities.
+avoid the frequentist approach of this equation and to explore a
+parameter space, where instead of computing a final number, we provide
+a statistical distribution that gives conditional probabilities.
 %
@@
 of years SETI projects have been working we have not received any
-signal within the conditions stablished by SETI.
-%
+signal within the conditions stablished by SETI \citep{tarter_search_2001}.
+%
+This lack of detection has lead to a number of alternative proposals
+for new SETI strategies \citep{forgan_exoplanet_2017,
+balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010,
+tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016,
+maccone_SETI_2011,  lingam_relative_2019, wright_theGsearch_2015,
+maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018,
+forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}
+
 This, by some astonishing result, has lead to the Fermi paradox, which
@@
 abundance of life in the Galaxy and the lack of their detection
-\citep{anchordoqui_isthere_2017}.
+\citep{vanhouten_isthere_2017}.
 
@@
 distribution of lifespans and the distribution of the number of CETIs
-per unit time.
+per unit time \citep{maccone_evolution_2014,Sotos_biotechnology_2019}.
 %
@@
 We also limit the posibilities of life or other types of civilizations
-to the hypothesis usually stated for the definition of the GHZ.
+to the hypothesis usually stated for the definition of the GHZ
+\citep{dayal_habitability_2016,gonzalez_galactic_2001,lineweaver_galactic_2004,gonzalez_habitable_2005,morrison_extending_2015,haqq-misra_evolution_2019,rahvar_cosmic_2016,gobat_evolution_2016,
+rahvar_cosmic_2016}.
+%
+\citet{vukotic_astrobiological_2012} propose probabilistic cellular
+automata model to explain the astrobiological history of the Milky
+Way.
 %
@@
 For example, we discard the possibility of stellar colonization
-(Walters 1980) and assume that the communication aim is performed to
+\citep[e.g.][]{newman_galactic_1981, walters_interstellar_1980,
+starling_virulence_2013, barlow_galactic_2012, jeong_large_2000, maccone_mathematical_2011}
+and assume that the communication aim is performed to
 all directions with the same power and the same probability, i.e.,
@@
 unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
+% 
+More examples can be found in the literature
+\citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009,
+maccone_evolution_2014, maccone_lognormals_2014,
+benguigui_classificacion_2016}
 
@@
 %}}}
-
-%\subsection{Exploration of the hypothesis space}
 
@@
 
-In the Fig.~\ref{F_waiting_time} we show the fraction of the CETIs
+In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the CETIs
 that have made at least one contact as a funtion of the elapsed time
@@
 probabilities could modified by a constant correction factor equal to
-the combined emission/reception efficiency.
+the combined emission/reception efficiency
+\citep{smith_broadcasting_2009, anchordoqui_upper_2019, forgan_collimated_2014}.
 %
@@
 %
-%energetic markers Spiritual markers
+%energetic markers
+%Spiritual markers
 %
@@
 of CETIs and different efficiencies for the sphere of the causal
-region of each CETI, resembling different searching strategies.
+region of each CETI, resembling different searching strategies
+\citep{hippke_interstellar_2017}.
 
@@
       GHZ.
+   \item Correction by coverage ratio in the detection and by
+      targetting ratio in the emission.
    \item Spiritual factor.  Explore the role of the S factor: message
@@
 
-\nocite{barabasi_scale_2009}
-\nocite{barlow_galactic_2012}
-\nocite{bloetscher_using_2019}
-\nocite{burchell_whither_2006}
-\nocite{conway_three_2018}
-\nocite{dayal_habitability_2016}
-\nocite{enriquez_breakthrough_2017}
-\nocite{forgan_exoplanet_2017}
-\nocite{forgan_possibility_2013}
-\nocite{forgan_collimated_2014}
-\nocite{forgan_numerical_2010}
-\nocite{funes_searching_2019}
-\nocite{glade_stochastic_2012}
-\nocite{gonzalez_habitable_2005}
-\nocite{gonzalez_galactic_2001}
-\nocite{grimaldi_signal_2017}
-\nocite{haqq-misra_evolution_2019}
-\nocite{harp_application_2018}
-\nocite{hinkel_interdisciplinary_2019}
-\nocite{hippke_interstellar_2017}
-\nocite{horvat_calculating_2007}
-\nocite{horvat_impact_2011}
-\nocite{lampton_information_2013}
-\nocite{lingam_relative_2019}
-\nocite{loeb_relative_2016}
-\nocite{loeb_eavesdropping_2006}
-\nocite{maccone_SETI_2013}
-\nocite{maccone_evolution_2014}
-\nocite{maccone_mathematical_2011}
-\nocite{maccone_statistical_2015}
-\nocite{maccone_KLT_2010}
-\nocite{maccone_statistical_2010}
-\nocite{maccone_SETI_2011}
-\nocite{morrison_extending_2015}
-\nocite{morrison_extending_2015}
-\nocite{newman_galactic_1981}
-\nocite{peters_outer_2018}
-\nocite{rahvar_cosmic_2016}
-\nocite{seager_search_2018}
-\nocite{shostak_what_2010}
-\nocite{smith_broadcasting_2009}
-\nocite{solomonides_probabilistic_2016}
-\nocite{starling_virulence_2013}
-\nocite{tarter_search_2001}
-\nocite{tarter_advancing_2009}
-\nocite{vukotic_astrobiological_2012}
-\nocite{walters_interstellar_1980}
-\nocite{wright_theGsearch_2015}
-\nocite{anchordoqui_upper_2019}
-\nocite{balb_spatiotemporal_2018}
-\nocite{forgan_galactic_2016}
-\nocite{gobat_evolution_2016}
-
 \bibliography{biblio_seti}

### 2019-08-29 manuscript.tex
puliendo las referencias y modificado el plot 1

@@
 contact between interstellar civilizations through MonteCarlo simulations}
+
+%% Stochastic estimation of the waiting time for the first causal contact
+%% with an extraterrestrial intelligence
 
@@
 
+HACER DISCUSION FUERTE DE ESTOS PAPERS: \cite{balbi_impact_2018,cirkovic_temporal_2004,grimaldi_signal_2017,smith_broadcasting_2009}
+
 \section{Methods and working hypotheses}\label{S_methods}
@@
 %
-We illustrate in the Fig.~\ref{F_growing_sphere} the schematic
+We illustrate in the Fig.~\ref{F_scheme} the schematic
 representation of the region causally connected to the central CETI.
+%
+%
+%
+% Schematic representation of the growing communicating
+%   sphere, over the space--time diagrams, where time is represented on
+%   the horizontal direction, and space is represented in the vertical
+%   direction.  In (a) the sphere is
+%   growing as the surface of first contact has not reached the maximum
+%   distance.  In (b) it has reached the maximum distance, so that it
+%   remains at the same size.  After a Doomsday event, the signals can
+%   still be observed, but the surface of last contact grows.
+%   %
+%   Schematic representation of two emitters, $E_1$ and $E_2$,
+%   that reach each other at different times.  The time span for $E_i$
+%   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
+%   emitter $j$ between $C_{ij}$ and $B_{ij}$.
+%   The type and length of causal contact in both directions depend on
+%   the distance and time lag between the awakening events, the maximum
+%   distance that a signal can reach and the time period in which each
+%   emitter is active.
+%
+%
 %
@@
 %
-In the Fig.~\ref{F_abcd}, we show a schematic representation of two
+We also show in this Figure a schematic representation of two
 emitters, E$_1$ and E$_2$, that reach each other at different times.
@@
 
+%\begin{figure}
+%   \centering
+%   \includegraphics[width=0.5\textwidth]{growingsphere.pdf}
+%   \caption{Schematic representation of the growing communicating
+%   sphere, over the space--time diagrams, where time is represented on
+%   the horizontal direction, and space is represented in the vertical
+%   direction.  In (a) the sphere is
+%   growing as the surface of first contact has not reached the maximum
+%   distance.  In (b) it has reached the maximum distance, so that it
+%   remains at the same size.  After a Doomsday event, the signals can
+%   still be observed, but the surface of last contact grows }
+%   \label{F_growing_sphere}
+%\end{figure}
+%
+%  
+%\begin{figure}
+%   \centering
+%   \includegraphics[width=0.5\textwidth]{abcd.pdf}
+%   \caption{Schematic representation of two emitters, $E_1$ and $E_2$,
+%   that reach each other at different times.  The time span for $E_i$
+%   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
+%   emitter $j$ between $C_{ij}$ and $B_{ij}$.
+%   The type and length of causal contact in both directions depend on
+%   the distance and time lag between the awakening events, the maximum
+%   distance that a signal can reach and the time period in which each
+%   emitter is active.}
+%   \label{F_abcd}
+%\end{figure}
+
 \begin{figure}
    \centering
-   \includegraphics[width=0.5\textwidth]{growingsphere.pdf}
-   \caption{Schematic representation of the growing communicating
-   sphere, over the space--time diagrams, where time is represented on
-   the horizontal direction, and space is represented in the vertical
-   direction.  In (a) the sphere is
-   growing as the surface of first contact has not reached the maximum
-   distance.  In (b) it has reached the maximum distance, so that it
-   remains at the same size.  After a Doomsday event, the signals can
-   still be observed, but the surface of last contact grows }
-   \label{F_growing_sphere}
-\end{figure}
-
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{abcd.pdf}
-   \caption{Schematic representation of two emitters, $E_1$ and $E_2$,
-   that reach each other at different times.  The time span for $E_i$
-   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
-   emitter $j$ between $C_{ij}$ and $B_{ij}$.
-   The type and length of causal contact in both directions depend on
-   the distance and time lag between the awakening events, the maximum
-   distance that a signal can reach and the time period in which each
-   emitter is active.}
-   \label{F_abcd}
+   \includegraphics[width=0.5\textwidth]{scheme.pdf}
+   \caption{Space--time diagrams showing a schematic representation of the 
+   different stages in the development of a communicating
+   civilization, the growing communicating sphere (a) and the temporal
+   intervals for the communication between two civilizations (b).
+   %
+   In the upper panel (a), the sphere the encloses the region which is
+   causally connected to the emitter grows until it reaches the
+   $D_{max}$ distance.  After the $D$ event, that region is limited by
+   the surface of last contact.
+   %
+   The scheme in the lower panel (b) represents
+   two communicating civilizations that have
+   causal contact in both directions}   
+   \label{F_scheme}
 \end{figure}
@@
 %
-Firstly, we assume that the distribution of the appeareance of new
-CETIs is a stationary Poisson distribution.
+\citet{forgan_spatiotemporal_2011} argues that the times at which
+different civilizations become intelligent 
+follow a Gaussian distribution, and then the distribution of
+inter-arrival times is an inverse exponential.
+%
+We chose to assume that the distribution of the times of A events is a
+stationary Poisson process, and then the distribution of the times between the
+appeareance of new consecutive CETIs is exponential.
 %
@@
 
+%%% comparar con el stochsatic cellular automata
+% ver libro Solving Fermi paradox, sec. 27.3.1
+
 \begin{figure} \centering

### 2019-10-28 manuscript.tex
agregue algunas referencias de las notas de mendeley

@@
 
+\citet{bloetscher_using_2019} uses a probabilistic approach, though
+still strongly motivated by the Drake equation, to obtain a
+propbabilistic measure for the number of civilizations in the Galaxy.
+To that end, perform MonteCarlo Markov Chains over each factor of the
+Drake equation, and combine the means to obtain a probabilistic
+result. It is worth mentioning that the author proposes a lof-normal
+target distribution to compute the posterior probabilities. This
+approach, which assigns probability distributions to the Drake
+factors. This study concludes that there is a very low probability for
+Earth to make contact with another extraterrestrail civilization.
+
+\citet{smith_broadcasting_2009} uses a analytical model to compute the
+probabilities of contact between two randomly located civilizations,
+and the waiting time for the first contact, assuming a fixed maximum
+broadcasting distance.
+
+\citet{balbi_impact_2018} investigate the chance of communicating
+civilizations making causal contact within a volume surrounding the
+location of the Earth, using a statistical model to compute the
+occurrence of causal contacts. The author stresses the fact that the
+causal contact requirement involves the distance between
+civilizations, their lifespan and their times of appeareance.  This is
+important since the time it takes the light to travel across the
+Galaxy is much lesser than the lifetime of the Galaxy.
+
+\citet{balbi_impact_2018} fix the total number of civilizations and
+explore the parameter space that comprise three parameters, namely,
+the distance to the Earth, the time of appeareance and the lifespan of
+the communicating civilization. Each of these three variables are
+drawn from a random distribution. The distribution of the distances
+results from a uniform distribution of civilization within the plane
+of the Galaxy. For the distribution of the characteristic time of
+appeareance, the author explores and exponential and a trucated
+Gaussian distributions. For the distribution of the lifespans, the
+author choses an exponential distribution. The radius of influence is
+set to 1000 ly.
+
+\citet{balbi_impact_2018} finds that the fraction of communicating
+civilizations vary with the choice of the statistical distribution for
+the time of appeareance. For all analyzed distributions, the fraction
+corresponding to a mean lifetime of 10$^9$ years and for a maximu
+radius for the detection of the signal of 1000 ys, is roughly 0.1.
+
+\citet{vukotic_astrobiological_2012} propose a simulation approach
+based on probabilistic cellular automata (PCA) modeling. In this
+framework, a complex system is modeled by a lattice of cells which
+evolve at discrete time steps, acording to transition rules that take
+into account for each cell the states of its neighbour cells.
+
+The authors implement a PCA model to a network of cells which
+represent life complexity on a 2-dimensional annular ring resembling
+the GHZ. The authors set the GHZ between a minimum radius of 6 kpc and
+a maximum radius of 10 kpc. Within this framework, the authors also
+make several MonteCarlo simulations and analyze ensemble-averaged
+results.
+
+This work aims at analyzing the evolution of life, although it does
+not account for the network of causal contacts among technological
+civilizations.
+
+%- - - -
+
 In this work we address the problem of the temporal and spatial
@@
 
-HACER DISCUSION FUERTE DE ESTOS PAPERS: \cite{balbi_impact_2018,cirkovic_temporal_2004,grimaldi_signal_2017,smith_broadcasting_2009}
+% HACER DISCUSION FUERTE DE ESTOS PAPERS: \cite{balbi_impact_2018,cirkovic_temporal_2004,grimaldi_signal_2017,smith_broadcasting_2009}
 

### 2019-11-28 manuscript.tex
retoque de la intro

@@
 
-A common approach to discuss the possible number of extraterrestrial
-intelligent civilizations (ETI) has been the use of the Drake equation
-\citep{gleiser_drake_2010, prantzos_joint_2013, haqq-misra_drake_2017}.
+A common approach to discussing the estimations of the number of
+extraterrestrial intelligent civilizations (ETI) in the Galaxy has
+been the use of the Drake equation \citep{gleiser_drake_2010,
+prantzos_joint_2013, haqq-misra_drake_2017}.
 %
@@
 forgan_galactic_2016}.
-There are many proposals intended to resolve this paradox, which make
-use of statistical \citep{solomonides_probabilistic_2016,
-vanhouten_isthere_2017, horvat_calculating_2007, maccone_statistical_2015} or
-stochastic approaches \citep{forgan_numerical_2009,
-bloetscher_using_2019, glade_stochastic_2011, forgan_numerical_2010},
-analytical interpretations of the Drake equation \citep{prantzos_joint_2013, smith_broadcasting_2009}
-or more speculative proposals
-\citep{barlow_galactic_2013, lampton_information_2013, conway_three_2018, forgan_galactic_2017}.
-%
-However, the uncertainties in the factors make it less prone to a
-formal application to define searching strategies or to compute
-numbers, but is used to organize the discussion instead
-\citep{hinkel_interdisciplinary_2019}.
+%
+There are many proposals aimed at solving this paradox, which make use
+of statistical \citep{solomonides_probabilistic_2016,
+vanhouten_isthere_2017, horvat_calculating_2007,
+maccone_statistical_2015} or stochastic approaches
+\citep{forgan_numerical_2009, bloetscher_using_2019,
+glade_stochastic_2011, forgan_numerical_2010}, analytical
+interpretations of the Drake equation \citep{prantzos_joint_2013,
+smith_broadcasting_2009} or more speculative proposals
+\citep{barlow_galactic_2013, lampton_information_2013,
+conway_three_2018, forgan_galactic_2017}.
+%
+The uncertainties in the factors of the Drake equation make it less
+prone to a formal application in order to define searching strategies
+or to compute the actual number of ETIs, but is a key tool to organize
+the discussion around the problem of the population of ETIs in the
+Galaxy \citep{hinkel_interdisciplinary_2019}, which is noticeable
+dominated by the fact that despite decades of effort not a single ETI,
+if any, has been detected so far.
 %
@@
 \citep{forgan_evaluating_2015, vukotic_grandeur_2016,
-murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017}
+murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017},
+although with a similar problem that consists on the large number of
+parameters which are either unknown or largely uncertain.
+%
 %New numerical determination of habitability in the Galaxy: the SETI connection
 %
-Since the factors in the Drake equation are uncertain, we propose to
-avoid the frequentist approach of this equation and to explore a
+%Since the factors in the Drake equation are uncertain,
+%
+In this work, we propose to
+avoid the frequentist approach of the Drake equation and to explore a
 parameter space, where instead of computing a final number, we provide
@@
 %
+This is an exploratory analysis that aims at providing a numerical 
+tool to discuss not the several theoretical problems summarized by the
+Drake equation factors, 
+but the different scenarios on the basis of statistical heuristics.
+%
+The approach proposed here shold be considered as a compromise between
+the uncertainties of the frequentist approach and the detailed recipes
+required on the simulation approach.
+%
+We then provide a numerical framework to explore via simulations a
+parameter space of unknown observables in order to discuss different
+scenarios and their consecuences in terms of the probability con
+making contacts.
+%
 The only fact that can be stated with certainty is that for the number
 of years SETI projects have been working we have not received any
-signal within the conditions stablished by SETI \citep{tarter_search_2001}.
-%
-This lack of detection has lead to a number of alternative proposals
+signal, at least
+within the technical posibilities and the 
+conditions stablished by SETI \citep{tarter_search_2001}.
+%
+This arguably astonishing result has lead to the Fermi paradox, which
+states an apparent contradiction between the high estimation of the
+abundance of life in the Galaxy and the lack of their detection
+\citep{vanhouten_isthere_2017}.
+%
+The lack of detection has also ispired to a number of alternative proposals
 for new SETI strategies \citep{forgan_exoplanet_2017,
@@
 maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018,
-forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}
-
-This, by some astonishing result, has lead to the Fermi paradox, which
-states an apparent contradiction between the high estimation of the
-abundance of life in the Galaxy and the lack of their detection
-\citep{vanhouten_isthere_2017}.
+forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
 

### 2019-12-10 manuscript.tex
descripcion de la figura 1

@@
 
-Some complex stochastic processes can be efficiently modeled with the
-discrete--event (DE) simulation approach.
-%
 Simulations are suitable tools to analyze systems that evolve with
@@
 theoretical model can not be found.
+%
+Certain complex stochastic processes can be efficiently modeled with
+the discrete--event (hereafter, DE) simulation approach.
 %
@@
 %
-In the case of study, the actors are the communicating civilizations
-(CETIs), and the events are all meaningfull that produce a change in
-the simulation variables, i.e., in the arrangement of CETIs and their
-causal contacts.
-%
-Since it allows to explore several models easily, 
-this method is well suited for the particular case of the difusion of
-intelligence signals in the Galaxy.
+We follow this approach to simulate the statistical properties of a
+set of intelligent civilizations (CETIs) in the Galaxy that have the
+ability of emitting and communicating messages at light speed.
+%
+Since this method to explore several models easily, it is well suited
+for the particular case of the difusion of intelligence signals in the
+Galaxy and its dependence on the statistical properties proposed for
+the rise and fall of the communicating capacity in the galactic
+intelligent civilizations.
+%
+The system is defined by a number of actors that represent CETIs and
+appear at different instants in time, generating events that produce
+meaningful changes in the variables that describe the system, i.e., in
+the arrangement of CETIs and their network of causal contacts.
 %
 Assuming some simple hypotheses, the discrete events method can be
-performed taking into account a small number of variables, which
-allow to analyze the variation of the results in the model parameter space.
+performed taking into account a small number of variables, which allow
+to analyze the variation of the results in the model parameter space.
 
@@
 %
-We also need to propose theoretical distribution functions for both the
-distribution of lifespans and the distribution of the number of CETIs
-per unit time \citep{maccone_evolution_2014,Sotos_biotechnology_2019}.
-%
-The shapes of these distributions are discussed in Sec.~\ref{SS_PDF_shape}.
+We also need to propose theoretical distribution functions for both
+the distribution of lifespans and the distribution of the number of
+CETIs per unit time \citep{maccone_evolution_2014,
+Sotos_biotechnology_2019}.
+%
+The shapes of these distributions are discussed in
+Sec.~\ref{SS_PDF_shape}.
 %
@@
 %
-As a consecuence of their finite lifetime,
-each CETI will fill a spherical shell region of the Galaxy, limited by two
-surfaces.
+Although a CETI could be old enough to transmit a message at large
+distances, the limited power of the message and the dilution that runs
+as the squared distance from the source, imposes a detectability
+limit.
+%
+As a consecuence of this limitation and of their finite lifetime,
+considered as the period of time between the aquisition and loss of
+communicating capacity, each CETI will fill a spherical shell region
+of the Galaxy, limited by two surfaces.
 %
@@
 %
-Although a CETI could be old enough to transmit a message at large
-distances, the limited power of the message and the dilution that runs
-as the squared distance from the source, imposes a
-detectability limit.
-%
 Thus, the volume reached causally by a CETI is initially a growing
@@
 Following the end of the civilization, there is still a region which
-is covered with the emited signals.
+is filled with the emited signals.
 %
@@
 %
-Any other CETI within this region will be aware of the originating CETI,
-even if it has disapeared before the time of contact.
-%
-This region will grow if the SFC has not reached $D_{max}$, and
-will shrink otherwise, until the SLC reaches $D_{max}$.
-%
-A that point, all signal from the central CETI is lost.
-%
-In the 2D simulation only the intersection of these spheres with the
-plane of the galaxy is relevant, and produce the corresponding
-circles for the filled spheres or annular regions for the spherical
-shells.
-%
-We illustrate in the Fig.~\ref{F_scheme} the schematic
-representation of the region causally connected to the central CETI.
-%
-%
+Any other CETI within this region will be aware of the originating
+CETI, even if it has disapeared before the time of contact.
+%
+This region will grow if the surface of first contact has not yet
+reached the maximum distance $D_{max}$, and will shrink otherwise
+until the surface of last contact reaches $D_{max}$, and producing as
+a result the loss of all signal from the central CETI.
+%
+In our approach, we consider a model Galaxy where the width of the
+disk is negligigle with respect to the radius of the disk, so that
+CETIs are placed on a 2D plane disk.
+%
+In the 2D simulation only the intersection of the communicating
+spherical shells with the plane of the galaxy is relevant, and produce
+the corresponding circles or rings for the filled spheres or annular
+regions of the spherical shells, respectively.
+%
+We illustrate in the Fig.~\ref{F_scheme} the schematic representation
+of the region causally connected to the central CETI.
+%
+To this end, we use space--time diagrams, where time is represented on
+the horizontal axis, and space is represented in the vertical axis.
+%
+A growing surface of first contact in this diagram is represented as a
+growing circle, centered on a point in the time axis, which shifts
+towards the right at the same rate the radius of the circle increases.
+%
+The maximum distance of the message limits the radius of the circles,
+so that once this distance is reached, the circle continues its shift
+towards the right, but maintaining a constant radius.
+%
+At the end of the lifetime of a CETI, a growing surface of last
+contact is represented by a circle that covers the inner part of the
+causal region, producing a ring of diminishig area (panel (a)).
+%
+In this scheme, the entire lifetime of a CETI can then be represented
+by a polygon, limited in the spatial axis by the maximum distance of
+the signal $D_{max}$ and in the time axis by the time span between the
+starting time (concave vertex) and the ending time (convex vertex).
+%
+This representation allows to easily visualize the important events
+that result from the existence and communications of two CETIs (panel (b)).
+%
+The points in time where the CETIs aquire their communicating
+capacity (CC) are dubbed ''awakening'' events ($A_1$ and $A_2$ in the
+Figure).
+%
+Simmilarly, the points in time where the CETIs loss their communicating
+capacity are dubbed ''doomsday'' events ($D_1$ and $D_2$ in the
+Figure).
+%
+The points in space--time where the first contact is produced for each
+one of the CETIs, are defined as ''contact'' events, shown as $C_1$
+and $C_2$ in the Figure.
+%
+Finally, the two points where the contact is lost for each one the
+CETIs, are denominated ''blackout'' events, shown as $B_1$ and $B_2$
+in the Figure.
+%
+It is important to notice that there is a time delay between the contact
+events of the two CETIs involved in this analysis, and also a time
+delay between the two blackout events, so that the time range when a
+bidirectional contact is possible occurs between the maximum time of
+the contact points and the minimum time between the blackout points.
+
 %

### 2019-12-13 manuscript.tex
un poco de texto

@@
 
-\Abbreviations{CETI: Communicating Extraterrestrial Intelligence, DE:
+\Abbreviations{MPL: Communicating Extraterrestrial Intelligence, DE:
 Discrete Event, GHZ: Galactic Habitable Zone}
@@
 %
-In general, simulations require less assumptions and simplifications
-than theoretical approaches, and can be applied to systems where a
-theoretical model can not be found.
-%
-Certain complex stochastic processes can be efficiently modeled with
-the discrete--event (hereafter, DE) simulation approach.
+An advantadge of simulations compared to theoretical approaches, is
+that the former usually require less assumptions and simplifications,
+and can be applied to systems where a theoretical model can not be
+found.
+%
+In particular, many complex stochastic processes that can be described
+by the evolution of the state of a system, can be efficiently modeled
+with the discrete--event (hereafter, DE) simulation approach.
 %
@@
 actors and events, where actors interact causally through a series of
-events on a time line and process these events in chronological order
+events on a timeline and process these events in chronological order
 \citep{ptolemaeus_system_2014, chung_simulation_2003,
@@
 %
-We follow this approach to simulate the statistical properties of a
-set of intelligent civilizations (CETIs) in the Galaxy that have the
-ability of emitting and communicating messages at light speed.
-%
-Since this method to explore several models easily, it is well suited
-for the particular case of the difusion of intelligence signals in the
-Galaxy and its dependence on the statistical properties proposed for
-the rise and fall of the communicating capacity in the galactic
-intelligent civilizations.
-%
-The system is defined by a number of actors that represent CETIs and
+Each event produces changes in the values of the variables that define
+the process, and thus the corresponding change in the state of the
+simulated system.
+%
+This method is well suited for the particular case of the difusion of
+intelligent signals in the Galaxy, and allows to explore several
+models easily.
+%
+We simulate the statistical properties of a set of points in space and
+time that have the ability of emitting and communicating messages at
+light speed, hereafter dubbed ''message processing locations'' (MPL).
+%
+These locations have the capability of emitting and receiving messages
+in all directions.
+%
+It is worth mentioning that this is a general approach, and not
+necessarily a MPL is the host of an intelligent civilization.
+%
+It can be associated with a set of planets where life has developed,
+became intelligent, reached the skills required to find the right
+communication channel, sustain a search and stablish a contact.
+%
+Alternative message processing entities could be considered, for
+example interstellar beacons where life has ceased to exist but
+continue with its emission or communication stations stablished by
+probes or left intentionally by intelligent beings \citep[see, e.g.,
+][]{peters_outer_2018, barlow_galactic_2013}.
+%
+In principle, these strategies could affect our results since it would
+be easier to configure a cluster of MPLs that spread in time.
+%
+However, we do not consider this highly speculative alternatives at
+this point.
+%
+% have been proposed
+%
+For the purpose of this analysis, only the communication capability is
+relevant, since we study the causal contacts between the locations.
+%
+% communicating, civilization, intelligent, message, processing,
+% handling, locations CALL, MPI, MPL, MPL
+%
+The system is defined by a number of actors that represent MPLs and
 appear at different instants in time, generating events that produce
 meaningful changes in the variables that describe the system, i.e., in
-the arrangement of CETIs and their network of causal contacts.
+the arrangement of MPLs and their network of causal contacts.
+%
+For example, the appeareance of a new MPL in a region which is filled
+with a signal emitted by another MPL, will increase the number of
+active MPLs and the number of pairs of MPLs in causal contact.
 %
@@
 terms of three independent parameters, namely, the mean time span
-between the appeareance of consecutive communicating civilizations
-($\tau_a$), their mean lifetime ($\tau_s$), and the maximum distance a
-signal can be detected by another civilization ($D_{max}$).
-%
-Intuitively, the shorter $\tau_a$ and the larger $\tau_s$ and
-$D_{max}$, there is a greater probability of the existence of causal
-contacts between pairs of intelligent civilizations in the Galaxy
-(Table \ref{T_simu_hypotheses}).
+between the appeareance of consecutive MPLs, ($<\tau_a>$), their mean
+lifetime ($<\tau_s>$), and the maximum distance a signal can be
+detected by another MPL ($D_{max}$).
+%
+Intuitively, the shorter the $\tau_a$ parameter and the larger the
+$\tau_s$ and $D_{max}$ parameters are, the greater the probability of
+the existence of causal contacts between pairs of MPLs would be.
 %
 We also need to propose theoretical distribution functions for both
-the distribution of lifespans and the distribution of the number of
-CETIs per unit time \citep{maccone_evolution_2014,
-Sotos_biotechnology_2019}.
-%
-The shapes of these distributions are discussed in
-Sec.~\ref{SS_PDF_shape}.
-%
+the distribution of lifespans ($\tau_s$) and the distribution of the
+number of MPLs per unit time \citep{maccone_evolution_2014,
+Sotos_biotechnology_2019}, related to the time span between the
+appeareance of consecutive MPLs (since when $<\tau_a>$ is shorter, it
+produces a greater density of MPLs).
+%
+The shapes of these distributions are set to a fixed law, as discussed
+in Sec.~\ref{SS_PDF_shape}.
+
 A causal connection can be produced without the need of the two
-involved civilizations being concurrent or active at the same time.
+involved locations being concurrent or active at the same time.
 %
@@
 %
-Although a CETI could be old enough to transmit a message at large
-distances, the limited power of the message and the dilution that runs
-as the squared distance from the source, imposes a detectability
-limit.
+Although a MPL could be active for enough time so as to transmit a
+message at large distances, the limited power of the message and the
+dilution that runs as the squared distance from the source, imposes a
+detectability limit.
 %
@@
 considered as the period of time between the aquisition and loss of
-communicating capacity, each CETI will fill a spherical shell region
-of the Galaxy, limited by two surfaces.
+communicating capacity, each MPL will fill a spherical shell region of
+the Galaxy, limited by two concentric spherical surfaces.
 %
 The leading front, or surface of first contact (SFC) grows from the
-central CETI until it reaches the maximum distance $D_{max}$.
-%
-Thus, the volume reached causally by a CETI is initially a growing
+central MPL until it reaches the maximum distance $D_{max}$.
+%
+Thus, the volume reached causally by a MPL is initially a growing
 sphere.
@@
 The trailing front, or surface of last contact (SLC) also grows from
-the central CETI, with a delay with respect to the SFC equivalent to
-the lifetime of the CETI, and produces a spherical shell region.
-%
-Any other CETI within this region will be aware of the originating
-CETI, even if it has disapeared before the time of contact.
+the central MPL, with a delay with respect to the SFC equivalent to
+the lifetime of the MPL, and produces a spherical shell region.
+%
+Any other MPL within this region will be aware of the originating MPL,
+even if it has disapeared before the time of contact.
 %
@@
 reached the maximum distance $D_{max}$, and will shrink otherwise
-until the surface of last contact reaches $D_{max}$, and producing as
-a result the loss of all signal from the central CETI.
+until the surface of last contact reaches $D_{max}$, producing as a
+result the loss of all signal from the central MPL.
 %
 In our approach, we consider a model Galaxy where the width of the
-disk is negligigle with respect to the radius of the disk, so that
-CETIs are placed on a 2D plane disk.
+disk is negligigle with respect to the radius of the disk, so that the
+MPLs are placed on a 2D plane disk.
 %
@@
 We illustrate in the Fig.~\ref{F_scheme} the schematic representation
-of the region causally connected to the central CETI.
+of the region causally connected to the central MPL.
 %
@@
 %
-At the end of the lifetime of a CETI, a growing surface of last
+At the end of the lifetime of a MPL, a growing surface of last
 contact is represented by a circle that covers the inner part of the
@@
 %
-In this scheme, the entire lifetime of a CETI can then be represented
+In this scheme, the entire lifetime of a MPL can then be represented
 by a polygon, limited in the spatial axis by the maximum distance of
@@
 This representation allows to easily visualize the important events
-that result from the existence and communications of two CETIs (panel (b)).
-%
-The points in time where the CETIs aquire their communicating
-capacity (CC) are dubbed ''awakening'' events ($A_1$ and $A_2$ in the
-Figure).
-%
-Simmilarly, the points in time where the CETIs loss their communicating
-capacity are dubbed ''doomsday'' events ($D_1$ and $D_2$ in the
-Figure).
+that result from the existence and communications of two MPLs (panel
+(b)).
+%
+The points in time where the MPLs aquire their communicating capacity
+(CC) are dubbed ''awakening'' events ($A_1$ and $A_2$ in the Figure).
+%
+Simmilarly, the points in time where the MPLs loss their
+communicating capacity are dubbed ''doomsday'' events ($D_1$ and $D_2$
+in the Figure).
 %
 The points in space--time where the first contact is produced for each
-one of the CETIs, are defined as ''contact'' events, shown as $C_1$
+one of the MPLs, are defined as ''contact'' events, shown as $C_1$
 and $C_2$ in the Figure.
@@
 Finally, the two points where the contact is lost for each one the
-CETIs, are denominated ''blackout'' events, shown as $B_1$ and $B_2$
+MPLs, are denominated ''blackout'' events, shown as $B_1$ and $B_2$
 in the Figure.
 %
-It is important to notice that there is a time delay between the contact
-events of the two CETIs involved in this analysis, and also a time
-delay between the two blackout events, so that the time range when a
-bidirectional contact is possible occurs between the maximum time of
-the contact points and the minimum time between the blackout points.
+It is important to notice that there is a time delay between the
+contact events of the two MPLs involved in this analysis, and also a
+time delay between the two blackout events, so that the time range
+when a bidirectional contact is possible occurs between the maximum
+time of the contact points and the minimum time between the blackout
+points.
 
@@
 In the Fig.~\ref{F_messages} we show a schematic representation of the
-possible cases in which a CETI A can be in causal contact with a CETI
+possible cases in which a MPL A can be in causal contact with a MPL
 B which appears later.
@@
 For example, a small $\tau_a$ and a large $\tau_s$ will produce a
-Galaxy densely populated woth CETIs, and viceversa.
+Galaxy densely populated with MPLs, and viceversa.
 %
@@
 stationary Poisson process, and then the distribution of the times between the
-appeareance of new consecutive CETIs is exponential.
-%
-Regarding the duration of a CETI, we propose that the distribution of
-the duration of CETIs is a stationary exponential distribution.
+appeareance of new consecutive MPLs is exponential.
+%
+Regarding the duration of a MPL, we propose that the distribution of
+the duration of MPLs is a stationary exponential distribution.
 %
@@
 %
-The probability for the rise of a CETI is homogeneous over the GHZ.
+The probability for the rise of a MPL is homogeneous over the GHZ.
 %
@@
 %
-This means that we set aside possibel civilizations that could survive
+This means that we set aside possible civilizations that could survive
 in severe conditions or unstable systems, which would prevent the
@@
    $\tau_{a}$ & Mean temporal separation between consecutive awakenings & 4000 y & 200000 y & 50\\ 
-   $\tau_{s}$ & Mean lifetime of a CETI                                 & 10000 y & 500000 y& 50\\ 
+   $\tau_{s}$ & Mean lifetime of a MPL                                 & 10000 y & 500000 y& 50\\ 
    D$_{max}$ & Maximum reach of a message  &  \multicolumn{3}{l}{500, 1000, 10000, 20000, 40000, 80000 lyr}  \\
@@
 \hline
-   & statistical properties of all CETIs &equally distributed&&\\
+   & statistical properties of all MPLs &equally distributed&&\\
    & Point process for the distribution in time & homogeneous &&\\
@@
 % 
-For the communication of CETIs through messages sent isotropically, we
+For the communication of MPLs through messages sent isotropically, we
 assume that the capacity to emit signals and to receive signals ocurr
@@
 %
-Another essential assumption is that all CETIs use the same signal
+Another essential assumption is that all MPLs use the same signal
 power, so that there is a maximum distance out to which it can be
@@
 %
-In such a system we compute probabilities of a CETI making contact
-with another CETI, and the results must be interpreted as for a
-randomly chosen CETI, given certain conditions of homogeneityon an
+In such a system we compute probabilities of a MPL making contact
+with another MPL, and the results must be interpreted as for a
+randomly chosen MPL, given certain conditions of homogeneityon an
 ensemble of equivalent galaxies.
@@
 We stress the fact that we are not centered on the probabilities for
-the Earth itselft, but on a generalized CETI.
+the Earth itselft, but on a generalized MPL.
 %
@@
 This is equivalent to proposing a Poisson process for the emergence of
-CETIs, since there is a close relation between the number of events in
+MPLs, since there is a close relation between the number of events in
 time or space and the waiting time or separation, respectively.
@@
 of stars, which are sampled randomly within the GHZ, the time of the awakening
-of each CETI (A event), and the time of the vanishing of each CETI (D event).
+of each MPL (A event), and the time of the vanishing of each MPL (D event).
 %
 The variables that can be deduced from the previous ones include the number of
-CETIs in casual contact with at least another CETI at a given time, the number
-of CETIs as a function of time, the number of CETIs that receive a message at
-least one time, the number of CETIs that receive a message at least one time and
+MPLs in casual contact with at least another MPL at a given time, the number
+of MPLs as a function of time, the number of MPLs that receive a message at
+least one time, the number of MPLs that receive a message at least one time and
 successfully deliver an answer and the number distribution of waiting time to
@@
    \caption{Empirical cumulative distributions of the number contacts
-   for CETIs in four different samples (M5, M6, M7, M8) with long
+   for MPLs in four different samples (M5, M6, M7, M8) with long
    range signal reach (D$_{max}$=10000).
@@
 
-\subsection{Model for CETIs}
+\subsection{Model for MPLs}
 %{{{
@@
 \begin{enumerate}
-   \item[(A)] A new CETI appears (Awakening)
-   \item[(D)] An old CETI dissappears (Doomsday)
+   \item[(A)] A new MPL appears (Awakening)
+   \item[(D)] An old MPL dissappears (Doomsday)
    \item[(C)] A new causal contact is stablished (Contact)
@@
 
-The event of type B is produced when one of the two CETIs halts in its
+The event of type B is produced when one of the two MPLs halts in its
 capability of emmiting and receiving signals.
@@
    \includegraphics[width=\textwidth]{matrix_1.pdf}
-   \caption{Rate of CETIs that never make contact (listening) as a
+   \caption{Rate of MPLs that never make contact (listening) as a
    function of $\tau_a$ and $\tau_s$, for 
@@
    \includegraphics[width=\textwidth]{matrix_2.pdf}
-   \caption{Rate of CETIs that listen at the moment of the
+   \caption{Rate of MPLs that listen at the moment of the
    awakening, as a funcion of $\tau_a$ and $\tau_s$, for
@@
 %
-The time marks for each event (A and D for each CETI and C and B for each causal
+The time marks for each event (A and D for each MPL and C and B for each causal
 contact) are stored as a result of the simulation.
 %
-Also, the list of active CETIs is obtained as a function of time.
+Also, the list of active MPLs is obtained as a function of time.
 %
 Some fixed variables must be set in order to carry out the simulation: the size
-of the Galactic Habitable Zone, the mean lifetime of a CETI, $\tau_s$, the mean
-waiting time for the appeareance of another CETI, $\tau_a$, and the maximum
+of the Galactic Habitable Zone, the mean lifetime of a MPL, $\tau_s$, the mean
+waiting time for the appeareance of another MPL, $\tau_a$, and the maximum
 distance a message can be detected.
@@
 range of one million years.
+%
+In Table \ref{T_simu_hypotheses} we show the range of values of these parameters explored 
+in our numerical experiments.
 
@@
 Some quantities are directly derived from the discrete events, namely,
-the ID of emitting CETI, the ID of emitting (receiving) CETI, the
+the ID of emitting MPL, the ID of emitting (receiving) MPL, the
 position in the galaxy, the time of Awakening (Contact) and the time
-of Doomsday  (Blackout) for each CETI.
-%
-This structure is always for the A-D cycle of each emitting CETI, and
+of Doomsday  (Blackout) for each MPL.
+%
+This structure is always for the A-D cycle of each emitting MPL, and
 repeats for each contact (C-B).
@@
 %
-The age of a CETI is the time elapsed from A to D.
-%
-%
-%given time, ceti\_e is the CETI emitting signals (emiter), ceti\_r is
-%the CETI receiving signals (receiver), ceti\_c is the ceti that
-%listen at least another ceti (citizen) and ceti\_h is the CETI ceti
-%that is lestened by at least another CETI.
-%
-The time span of a CETI listening another or being listened by another
+The age of a MPL is the time elapsed from A to D.
+%
+%
+%given time, ceti\_e is the MPL emitting signals (emiter), ceti\_r is
+%the MPL receiving signals (receiver), ceti\_c is the ceti that
+%listen at least another ceti (citizen) and ceti\_h is the MPL ceti
+%that is lestened by at least another MPL.
+%
+The time span of a MPL listening another or being listened by another
 can also be easily derived from the results of a realization.
@@
 %
-Regarding the properties of the population of CETIs and its evolution,
+Regarding the properties of the population of MPLs and its evolution,
 we analyze the fraction of awaken time a ceti is listening at least
-another ceti, the age of contacted CETIs at first contact, the
+another ceti, the age of contacted MPLs at first contact, the
 distribution of time to wait until next contact, the fraction of cetis
 where the first contact is given at the awakening, the distribution of
-the number of contacts for each CETI, the distribution of the maximum
-number of contacts for each CETI, the distribution of the number of
-contacts as a function of CETI age, the number of contacts as a
-function of time in the galaxy, the rate of CETIs that succeed in
-contact, and the distribution of distances between contacted CETIs.
+the number of contacts for each MPL, the distribution of the maximum
+number of contacts for each MPL, the distribution of the number of
+contacts as a function of MPL age, the number of contacts as a
+function of time in the galaxy, the rate of MPLs that succeed in
+contact, and the distribution of distances between contacted MPLs.
 %
@@
 explored parameter space, and cover short/long lifetimes, short/long
-waiting times for the next CETI to appear in the Galaxy and
+waiting times for the next MPL to appear in the Galaxy and
 short/large range reach of the signal, including all possible
@@
 In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
-distributions of the number contacts for CETIs in four different
+distributions of the number contacts for MPLs in four different
 samples (M5, M6, M7, M8) with long range signal reach
@@
 that maximizes the number of contacts, reaching a maximum of order 100
-contacts for a single CETI.
+contacts for a single MPL.
 %
@@
 which produce a number of contacts of at most the order of ten
-contacts per CETI.
+contacts per MPL.
 %
@@
 variations of contact posibilities as a function of the position of
-the CETI in the Galaxy (Sec.~\ref{SS_location}).
+the MPL in the Galaxy (Sec.~\ref{SS_location}).
 
@@
 
-The duration of each contact for a single receiving CETI is given by the
+The duration of each contact for a single receiving MPL is given by the
 time difference between a C event and a B event, which are direct
@@
 Such distribution can be considered to compute the probability for a
-random CETI to carry out a SETI project and spend a given time until
+random MPL to carry out a SETI project and spend a given time until
 the first contact is made.
 
-In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the CETIs
+In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the MPLs
 that have made at least one contact as a funtion of the elapsed time
@@
 Both panels show the histograms of the waiting times for the first
-contact, for all the CETIs that make at least one contact along their
+contact, for all the MPLs that make at least one contact along their
 lifetime.
@@
 %
-We show the 2D maps of the rate of CETIs that never make contact and
-the rate of CETIs that make contact at the moment of the awakening,
+We show the 2D maps of the rate of MPLs that never make contact and
+the rate of MPLs that make contact at the moment of the awakening,
 inf the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
@@
 %      %
-%      \caption{Fraction of CETIs making cero, one, two or three contacts
+%      \caption{Fraction of MPLs making cero, one, two or three contacts
 %      as a function of galactocentric distance.}
@@
 
-%On the Figure~\ref{F_radial_frac} we show the fraction of CETIs that make
-%zero, one, two or three contacts, as a function of the galactocentric
-%distance, for model M1.
-% VERIFICAR!  M1?
-%
-For this model, the majority of CETIs do not
-make any contact during all the time period in which they are active.
-%
-As expected, the fractions of CETIs making at least one contact
+%On the Figure~\ref{F_radial_frac} we show the fraction of MPLs that
+%make zero, one, two or three contacts, as a function of the
+%galactocentric distance, for model M1.  VERIFICAR!  M1?
+%
+For this model, the majority of MPLs do not make any contact during
+all the time period in which they are active.
+%
+As expected, the fractions of MPLs making at least one contact
 diminishes as the number of contacts rises.
@@
 %
-In general, a closer position to the center of the Galaxy slightly favours 
-the posibility of a contact.
+In general, a closer position to the center of the Galaxy slightly
+favours the posibility of a contact.
 %
@@
 We have implemented a suite of simulations following a stochastic
-approach, to explore the hypothesis
-space of a simple model that accounts for
-the causal connections between communicating
+approach, to explore the hypothesis space of a simple model that
+accounts for the causal connections between communicating
 civilizations on an simplified galaxy.
@@
 uninterrupted SETI effort to reach success, in terms of different
-model parameters that reflect different, so far unknown, scenarios
-for the spread of intelligent life on the galaxy.
+model parameters that reflect different, so far unknown, scenarios for
+the spread of intelligent life on the galaxy.
 
@@
 
-Under the hypotheses of our experiments, we conclude that
-a causal contact is extremely unlikely unless the Galaxy is heavily
-populated by intelligent civilizations.
+Under the hypotheses of our experiments, we conclude that a causal
+contact is extremely unlikely unless the Galaxy is heavily populated
+by intelligent civilizations.
 %
@@
 any two entities, a minimum degree of compatibility must be
-accomplished without any previous agreement, 
-making the posibility of a contact with a message that
-could be deciphered highly rare.
+accomplished without any previous agreement, making the posibility of
+a contact with a message that could be deciphered highly rare.
 
@@
 the combined emission/reception efficiency
-\citep{smith_broadcasting_2009, anchordoqui_upper_2019, forgan_collimated_2014}.
+\citep{smith_broadcasting_2009, anchordoqui_upper_2019,
+forgan_collimated_2014}.
 %
@@
 of the Galaxy, different distribution functions for the mean lifetime
-of CETIs and different efficiencies for the sphere of the causal
-region of each CETI, resembling different searching strategies
+of MPLs and different efficiencies for the sphere of the causal region
+of each MPL, resembling different searching strategies
 \citep{hippke_interstellar_2017}.
@@
 \begin{itemize}
-   \item Dmax is different for different CETIs.  For example a power
+   \item D$_{max}$ is different for different MPLs.  For example a power
       law where powerfull emmisions are rare and weak emssions are
       common.
-   \item The probability of the rising of a new CETI vary within the
-      GHZ.
+   \item The probability of the rising of a new MPL vary within the GHZ.
    \item Correction by coverage ratio in the detection and by
@@
    \item Spiritual factor.  Explore the role of the S factor: message
-      contents, influence on the lifespan of a CETI that receives a
-      message, mean lifespan of a CETI that emits a message.
+      contents, influence on the lifespan of a MPL that receives a
+      message, mean lifespan of a MPL that emits a message.
 \end{itemize}

### 2019-12-19 captions.tex
reestructuracion

--- 2019-11-28 manuscript.tex
@@
-\documentclass[crop]{CSLB}%%%%where CSL is the template name
+% #1 is the caption name, #2 is the caption text
+\newcommand{\definecaption}[2]{%
+   \expandafter\newcommand\csname caption@#1\endcsname{#2}}
 
-%The authors can define any packages after the \documentclass{CSL} command.
-\usepackage{amsmath} %for dealing with mathematics,
-\usepackage{amsthm} %for dealing with theorem environments,
-\usepackage{cite} %%%for dealing with citations
-\usepackage{hyperref} %for linking the cross references
-\usepackage{graphics} %for dealing with figures.
-\usepackage{algorithmic} %for describing algorithms
-\usepackage{subfig} %for getting the subfigures e.g., "Figure 1a and 1b" etc.
-\usepackage{url} %It provides better support for handling and breaking URLs.
+\newcommand{\usecaption}[1]{%
+   \caption{\csname caption@#1\endcsname\label{#1}}}
 
-%\usepackage{natbib}
-\usepackage{boites}
+% Now the captions
 
-%\usepackage{graphicx}
-%\usepackage{float}
-%\usepackage{psfig}
-%\usepackage{amsfonts,amssymb}
-%\usepackage{color}
-%\usepackage{textcomp}
-%\usepackage[T1]{fontenc}
-%\usepackage{aecompl}
-%\usepackage{times}
-%\usepackage{txfonts}
-%\usepackage[english]{babel}
-%\usepackage[utf8x]{inputenc}
-%\usepackage{amsmath}
-%\usepackage{caption}
-%\usepackage{siunitx}
-
-\jname{Astrobiology}%
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{condition}{Condition}
-
-\begin{document}
-
-\bibliographystyle{mn2e}
-
-\supertitle{Research Paper}
-
-\title[Probabilities of SETI contacts]{Probability of causal
-contact between interstellar civilizations through MonteCarlo simulations}
-
-%% Stochastic estimation of the waiting time for the first causal contact
-%% with an extraterrestrial intelligence
-
-\author[Lares \& Funes]{Lares M$^{1, 2}$, Funes J$^{1, 3}$}
-
-\corres{\name{Lares M} \email{marcelo.lares@unc.edu.ar}}
-
-\address{
+%_______________________________________________________________
+\definecaption{F_scheme}{%
 %
-   \add{1}{CONICET, Argentina}
-
-   \add{2}{Universidad Nacional de C\'ordoba, Observatorio
-   Astron\'omico de C\'ordoba, Argentina}
-
-   \add{3}{Universidad Cat\'olica de C\'ordoba, Argentina}
+Space--time diagrams showing a schematic representation of the
+different stages in the development of a communicating civilization,
+the growing communicating sphere (a) and the temporal intervals for
+the communication between two civilizations (b).
+%
+In the upper panel (a), the sphere the encloses the region which is
+causally connected to the emitter grows until it reaches the
+$D_{max}$ distance.  After the $D$ event, that region is limited by
+the surface of last contact.
+%
+The scheme in the lower panel (b) represents two communicating
+civilizations that have causal contact in both directions
 %
@@
 
-\begin{abstract}
+%_______________________________________________________________
+\definecaption{F_messages}{%
 %   
-The abundance of intelligent civilizations in the galaxy is a
-longstanding question, which is often conceptualized as the problem
-of the lack of received communication or the Fermi paradox.
+Schematic representation of the possible cases in which a
+ceti A can be in causal contact with a ceti B which appears later.
 %   
-Most efforts on the estimation of a number of intelligent
-civilizations are centered on the Drake equation, although 
-its factors are affected by large uncertainties, and it lacks a
-temporal nature.
+The duration of the causal contact in one direction depends on
+several factors, mainly $\Delta t_A$, $\Delta t_C$, and $\Delta X$.
 %   
-A different approach uses detailed numerical simulations of the galaxy
-and recipes for the rates of the formation of stars and planets, and
-even for the origin of life.
+}
+
+%_______________________________________________________________
+\definecaption{F_number_of_contacts}{%
 %
-Here we present numerical simulations of stochastic processes of
-emergence of civilizations with communication capabilities, based
-on minimal asumptions, and analyze the causal connections among them.
+Empirical cumulative distributions of the number contacts
+for MPLs in four different samples (M5, M6, M7, M8) with long
+range signal reach (D$_{max}$=10000).
 %
-The analysis of the rate of causal contacts as a function of the mean
-number of civilizations, the mean lifetime span distribution and the
-maximum distance a civilization can send signals is presented, and
-used as a framework to discuss the spatial and temporal structure
-of a populated galaxy within several different scenarios.
+}
+
+%_______________________________________________________________
+\definecaption{F_never_contact}{%
 %
-Our results indicate that, given the large distances involved, causal
-contacts between civilizations are very rare.
+Rate of MPLs that never make contact (listening) as a
+function of $\tau_a$ and $\tau_s$, for 
+$D_{max}$=10000 (left panel),
+$D_{max}$=40000 (middle panel), and
+$D_{max}$=80000 (right panel)
 %
-Additionally, the odds to make a contact in a few years of monitoring, 
-assuming a perfect detection rate, are low for most models, with the
-exception of models which propose a galaxy densely populated with
-long lived civilizations.
+}
+
+%_______________________________________________________________
+\definecaption{F_C_at_A}{%
 %
-The probability of causal contacts increases with the increase of the 
-mean lifetime of civilizations much more signifficantly than with
-the increase in the mean number of active civilizations for a given
-time window.
+Rate of MPLs that listen at the moment of the
+awakening, as a funcion of $\tau_a$ and $\tau_s$, for
+$D_{max}$=10000 (left panel),
+$D_{max}$=40000 (middle panel), and
+$D_{max}$=80000 (right panel)
 %
-\end{abstract}
+}
 
-\selfcitation{Lares M, Funes J (XXXX).  Estimating the probability of
-causal contact between civilizations through MonteCarlo simulations,
-International Journal of Astrobiology}
+%_______________________________________________________________
+\definecaption{F_waiting_for_1C}{%
+%
+Histograms of the mean waiting times for the first contact, for
+several models.
+%
+Left panel shows the histograms for several values of $\tau_a$, and
+$\tau_s$ in the range 17000-24000 yr.
+%
+Right panel shows the histograms for several values of $\tau_s$, and
+$\tau_a$ in the range 68000-100000 yr.
+%
+The shape of the histograms does not change when $\tau_a$ is
+modified, but significantly changes when $\tau_s$ is modified.
+}
 
-\keywords{SETI, Computer simulations, Statistics}
-
-\JELclassification{XX; XX}
-\MSCcodes{XX; XX}
-
-\Abbreviations{CETI: Communicating Extraterrestrial Intelligence, DE:
-Discrete Event, GHZ: Galactic Habitable Zone}
-
-\maketitle
-
-\Fpagebreak
-
-\section{Motivations}\label{S_motivations}
 %{{{ 
-
-A common approach to discussing the estimations of the number of
-extraterrestrial intelligent civilizations (ETI) in the Galaxy has
-been the use of the Drake equation \citep{gleiser_drake_2010,
-prantzos_joint_2013, haqq-misra_drake_2017}.
-%
-The abundance of intelligent civilizations in the galaxy is a
-longstanding question, which is often conceptualized as the problem of
-the lack of received communication or the Fermi paradox
-\citep{barlow_galactic_2012, Sotos_biotechnology_2019,
-forgan_galactic_2016}.
-%
-There are many proposals aimed at solving this paradox, which make use
-of statistical \citep{solomonides_probabilistic_2016,
-vanhouten_isthere_2017, horvat_calculating_2007,
-maccone_statistical_2015} or stochastic approaches
-\citep{forgan_numerical_2009, bloetscher_using_2019,
-glade_stochastic_2011, forgan_numerical_2010}, analytical
-interpretations of the Drake equation \citep{prantzos_joint_2013,
-smith_broadcasting_2009} or more speculative proposals
-\citep{barlow_galactic_2013, lampton_information_2013,
-conway_three_2018, forgan_galactic_2017}.
-%
-The uncertainties in the factors of the Drake equation make it less
-prone to a formal application in order to define searching strategies
-or to compute the actual number of ETIs, but is a key tool to organize
-the discussion around the problem of the population of ETIs in the
-Galaxy \citep{hinkel_interdisciplinary_2019}, which is noticeable
-dominated by the fact that despite decades of effort not a single ETI,
-if any, has been detected so far.
-%
-Some modifications to the original idea of the equation have tried to
-imprint a stochastic nature, or to propose a probabilistic approach,
-or to consider the temporal structure which is missing in the
-equation.
-%
-Temporal aspects of the distribution of communicating civilizations
-and their contacts have also been explored by several authors
-\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
-balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011},
-%The Impact of the Temporal Distribution of Communicating Civilizations on Their Detectability
-%Temporal dispersion of the emergence of intelligence: an inter-arrival time analysis
-%The spatiotemporal aspects of SETI
-%
-as well as efforts on considering the stochastic nature of the Drake equation
-\citep{glade_stochastic_2011}.
-%
-The simulation approach has also benn considered
-\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
-murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017},
-although with a similar problem that consists on the large number of
-parameters which are either unknown or largely uncertain.
-%
-%New numerical determination of habitability in the Galaxy: the SETI connection
-%
-%Since the factors in the Drake equation are uncertain,
-%
-In this work, we propose to
-avoid the frequentist approach of the Drake equation and to explore a
-parameter space, where instead of computing a final number, we provide
-a statistical distribution that gives conditional probabilities.
-%
-This is an exploratory analysis that aims at providing a numerical 
-tool to discuss not the several theoretical problems summarized by the
-Drake equation factors, 
-but the different scenarios on the basis of statistical heuristics.
-%
-The approach proposed here shold be considered as a compromise between
-the uncertainties of the frequentist approach and the detailed recipes
-required on the simulation approach.
-%
-We then provide a numerical framework to explore via simulations a
-parameter space of unknown observables in order to discuss different
-scenarios and their consecuences in terms of the probability con
-making contacts.
-%
-The only fact that can be stated with certainty is that for the number
-of years SETI projects have been working we have not received any
-signal, at least
-within the technical posibilities and the 
-conditions stablished by SETI \citep{tarter_search_2001}.
-%
-This arguably astonishing result has lead to the Fermi paradox, which
-states an apparent contradiction between the high estimation of the
-abundance of life in the Galaxy and the lack of their detection
-\citep{vanhouten_isthere_2017}.
-%
-The lack of detection has also ispired to a number of alternative proposals
-for new SETI strategies \citep{forgan_exoplanet_2017,
-balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010,
-tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016,
-maccone_SETI_2011,  lingam_relative_2019, wright_theGsearch_2015,
-maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018,
-forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
-
-The discrete events method for simulating a stochastic process is an
-approximation that allows to study the behaviour of complex
-systems, by considering a sequence of well defined discrete events.
-%
-The simulation os carried out by following all the variables that
-describe the system, that constitute the state of the system.
-%
-The evolution of the process, then, is described as a set of changes
-in the state of the system.
-%
-In this context, an event produces a specific change in the state,
-that can be triggered by random variables that encode the stochastic
-nature of the physical phenomenon.
-%
-The process involves following the changes on the state of the system,
-definig the initial and final states, defining a method that allows to
-keep track of time progress, and maintaining a list of relevant
-events.
-
-In a recent work, \citep{balbi_impact_2018} use a statistical model to analyze
-the occurrence of causal contacts between civilizations in the Galaxy.
-%
-The author highlights the effect of evolutionary processes when
-attempting to estimate the number of communicating civilizations that
-might be in causal contact with an observer located on the Earth.
-%
-\citet{cirkovic_temporal_2004} also emphasize the lack of temporal
-structure in the Drake equation.
-
-%While the Drake equation remains a very useful
-%guide in identifying and analyzing the various factors in-
-%volved in the problem and their relative probabilities, our
-%approach could provide a better framework with which to
-%perform a statistical analysis when taking into account evo-
-%lutionary processes. In particular, we have shown that the
-%Drake equation can misestimate the number of detectable
-%communicating species if spatial and temporal dependencies
-%are important. This should suggest some caution in the un-
-%critical application of such an equation. In this paper, we offer
-%an exemplificative illustration of the effect of different
-%probability distributions for the typical timescales involved
-%in the problem. Ideally, it would be desirable to arrive at a
-%mathematical description of such distributions based on a
-%modeling of the underlying astrophysical and biological
-%processes. This points out a possible future direction of in-
-%vestigation in which an integrated evolutionary (and coevo-
-%lutionary) approach to the problem would be a central focus
-%of interest (for more on this perspective, see, e.g., Chaisson
-
-\citet{bloetscher_using_2019} uses a probabilistic approach, though
-still strongly motivated by the Drake equation, to obtain a
-propbabilistic measure for the number of civilizations in the Galaxy.
-To that end, perform MonteCarlo Markov Chains over each factor of the
-Drake equation, and combine the means to obtain a probabilistic
-result. It is worth mentioning that the author proposes a lof-normal
-target distribution to compute the posterior probabilities. This
-approach, which assigns probability distributions to the Drake
-factors. This study concludes that there is a very low probability for
-Earth to make contact with another extraterrestrail civilization.
-
-\citet{smith_broadcasting_2009} uses a analytical model to compute the
-probabilities of contact between two randomly located civilizations,
-and the waiting time for the first contact, assuming a fixed maximum
-broadcasting distance.
-
-\citet{balbi_impact_2018} investigate the chance of communicating
-civilizations making causal contact within a volume surrounding the
-location of the Earth, using a statistical model to compute the
-occurrence of causal contacts. The author stresses the fact that the
-causal contact requirement involves the distance between
-civilizations, their lifespan and their times of appeareance.  This is
-important since the time it takes the light to travel across the
-Galaxy is much lesser than the lifetime of the Galaxy.
-
-\citet{balbi_impact_2018} fix the total number of civilizations and
-explore the parameter space that comprise three parameters, namely,
-the distance to the Earth, the time of appeareance and the lifespan of
-the communicating civilization. Each of these three variables are
-drawn from a random distribution. The distribution of the distances
-results from a uniform distribution of civilization within the plane
-of the Galaxy. For the distribution of the characteristic time of
-appeareance, the author explores and exponential and a trucated
-Gaussian distributions. For the distribution of the lifespans, the
-author choses an exponential distribution. The radius of influence is
-set to 1000 ly.
-
-\citet{balbi_impact_2018} finds that the fraction of communicating
-civilizations vary with the choice of the statistical distribution for
-the time of appeareance. For all analyzed distributions, the fraction
-corresponding to a mean lifetime of 10$^9$ years and for a maximu
-radius for the detection of the signal of 1000 ys, is roughly 0.1.
-
-\citet{vukotic_astrobiological_2012} propose a simulation approach
-based on probabilistic cellular automata (PCA) modeling. In this
-framework, a complex system is modeled by a lattice of cells which
-evolve at discrete time steps, acording to transition rules that take
-into account for each cell the states of its neighbour cells.
-
-The authors implement a PCA model to a network of cells which
-represent life complexity on a 2-dimensional annular ring resembling
-the GHZ. The authors set the GHZ between a minimum radius of 6 kpc and
-a maximum radius of 10 kpc. Within this framework, the authors also
-make several MonteCarlo simulations and analyze ensemble-averaged
-results.
-
-This work aims at analyzing the evolution of life, although it does
-not account for the network of causal contacts among technological
-civilizations.
-
-%- - - -
-
-In this work we address the problem of the temporal and spatial
-structure of the distribution of communicating civilizations, by
-exploring the hypothesis space over a minimal set of parameters.
-%
-In Sec.~\ref{S_methods} we introduce the methods and discuss the
-candidate distributions for the statistical aspects of the times
-involved in the communication process.
-%
-Then, we present our results in Sec~\ref{S_results}, with special
-emphasis on the statistical distributions of the duration of causal
-contacts in one or both directions, the possible differences on the
-position of a civilization on the Galaxy and the distribution of time
-intervals for the waiting of the first contact, always as a function
-of the three simulation parameters.
-%
-Finally, in Sec.~\ref{S_discussion} we discuss our results and future
-research directions.
-
-%}}}
-
-% HACER DISCUSION FUERTE DE ESTOS PAPERS: \cite{balbi_impact_2018,cirkovic_temporal_2004,grimaldi_signal_2017,smith_broadcasting_2009}
-
-\section{Methods and working hypotheses}\label{S_methods}
-%{{{
-
-Some complex stochastic processes can be efficiently modeled with the
-discrete--event (DE) simulation approach.
-%
-Simulations are suitable tools to analyze systems that evolve with
-time and involve randomness.
-%
-In general, simulations require less assumptions and simplifications
-than theoretical approaches, and can be applied to systems where a
-theoretical model can not be found.
-%
-A system described with the DE paradigm is characterized by a set of
-actors and events, where actors interact causally through a series of
-events on a time line and process these events in chronological order
-\citep{ptolemaeus_system_2014, chung_simulation_2003,
-ross_simulation_2012}.
-%
-In the case of study, the actors are the communicating civilizations
-(CETIs), and the events are all meaningfull that produce a change in
-the simulation variables, i.e., in the arrangement of CETIs and their
-causal contacts.
-%
-Since it allows to explore several models easily, 
-this method is well suited for the particular case of the difusion of
-intelligence signals in the Galaxy.
-%
-Assuming some simple hypotheses, the discrete events method can be
-performed taking into account a small number of variables, which
-allow to analyze the variation of the results in the model parameter space.
-
-In what follows, we describe the experimental setup chosen to estimate
-the probabilities of causal contacts and several derived quantities in
-terms of three independent parameters, namely, the mean time span
-between the appeareance of consecutive communicating civilizations
-($\tau_a$), their mean lifetime ($\tau_s$), and the maximum distance a
-signal can be detected by another civilization ($D_{max}$).
-%
-Intuitively, the shorter $\tau_a$ and the larger $\tau_s$ and
-$D_{max}$, there is a greater probability of the existence of causal
-contacts between pairs of intelligent civilizations in the Galaxy
-(Table \ref{T_simu_hypotheses}).
-%
-We also need to propose theoretical distribution functions for both the
-distribution of lifespans and the distribution of the number of CETIs
-per unit time \citep{maccone_evolution_2014,Sotos_biotechnology_2019}.
-%
-The shapes of these distributions are discussed in Sec.~\ref{SS_PDF_shape}.
-%
-A causal connection can be produced without the need of the two
-involved civilizations being concurrent or active at the same time.
-%
-This is a fundamental property of the system which arises as a
-consecuence of the large spatial and temporal scales, where a message
-is transmited at light speed.
-%
-As a consecuence of their finite lifetime,
-each CETI will fill a spherical shell region of the Galaxy, limited by two
-surfaces.
-%
-The leading front, or surface of first contact (SFC) grows from the
-central CETI until it reaches the maximum distance $D_{max}$.
-%
-Although a CETI could be old enough to transmit a message at large
-distances, the limited power of the message and the dilution that runs
-as the squared distance from the source, imposes a
-detectability limit.
-%
-Thus, the volume reached causally by a CETI is initially a growing
-sphere.
-%
-Following the end of the civilization, there is still a region which
-is covered with the emited signals.
-%
-The trailing front, or surface of last contact (SLC) also grows from
-the central CETI, with a delay with respect to the SFC equivalent to
-the lifetime of the CETI, and produces a spherical shell region.
-%
-Any other CETI within this region will be aware of the originating CETI,
-even if it has disapeared before the time of contact.
-%
-This region will grow if the SFC has not reached $D_{max}$, and
-will shrink otherwise, until the SLC reaches $D_{max}$.
-%
-A that point, all signal from the central CETI is lost.
-%
-In the 2D simulation only the intersection of these spheres with the
-plane of the galaxy is relevant, and produce the corresponding
-circles for the filled spheres or annular regions for the spherical
-shells.
-%
-We illustrate in the Fig.~\ref{F_scheme} the schematic
-representation of the region causally connected to the central CETI.
-%
-%
-%
-% Schematic representation of the growing communicating
-%   sphere, over the space--time diagrams, where time is represented on
-%   the horizontal direction, and space is represented in the vertical
-%   direction.  In (a) the sphere is
-%   growing as the surface of first contact has not reached the maximum
-%   distance.  In (b) it has reached the maximum distance, so that it
-%   remains at the same size.  After a Doomsday event, the signals can
-%   still be observed, but the surface of last contact grows.
-%   %
-%   Schematic representation of two emitters, $E_1$ and $E_2$,
-%   that reach each other at different times.  The time span for $E_i$
-%   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
-%   emitter $j$ between $C_{ij}$ and $B_{ij}$.
-%   The type and length of causal contact in both directions depend on
-%   the distance and time lag between the awakening events, the maximum
-%   distance that a signal can reach and the time period in which each
-%   emitter is active.
-%
-%
-%
-The initially growing communicating sphere, is shown over
-space--time diagrams, where time is represented on the horizontal
-direction, and space is represented in the vertical direction.
-%
-In the upper panel (a), the sphere is growing as the surface of first
-contact has not reached the maximum distance.
-%
-In the middle panel (b) it has reached the maximum distance, so that
-it remains at the same size.
-%
-After a Doomsday event (bottom panel, b), the signals can still be
-observed, but the surface of last contact grows.
-%
-We also show in this Figure a schematic representation of two
-emitters, E$_1$ and E$_2$, that reach each other at different times.
-%
-The time span for E$_i$ is (A$_i$ , D$_i$) , for $i = {1, 2}$.
-%
-Emitter $i$ can listen to emitter $j$ between C$_{ij}$ and B$_{ij}$.
-%
-The type and length of causal contact in both directions depend on the
-distance and time lag between the awakening events, the maximum
-distance that a signal can reach and the time period in which each
-emitter is active.
-%
-In the Fig.~\ref{F_messages} we show a schematic representation of the
-possible cases in which a CETI A can be in causal contact with a CETI
-B which appears later.
-%
-The duration of the causal contact in one direction depends on several
-factors, mainly $\Delta$t~A , $\Delta$t~C , and $\Delta$X.
-
 %\begin{figure}
@@
 
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{scheme.pdf}
-   \caption{Space--time diagrams showing a schematic representation of the 
-   different stages in the development of a communicating
-   civilization, the growing communicating sphere (a) and the temporal
-   intervals for the communication between two civilizations (b).
-   %
-   In the upper panel (a), the sphere the encloses the region which is
-   causally connected to the emitter grows until it reaches the
-   $D_{max}$ distance.  After the $D$ event, that region is limited by
-   the surface of last contact.
-   %
-   The scheme in the lower panel (b) represents
-   two communicating civilizations that have
-   causal contact in both directions}   
-   \label{F_scheme}
-\end{figure}
-
-The temporal structure of the process is defined by two distribution
-parameters that represent the mean time interval that an intelligent
-civilization can emit and receive signals, and the mean time interval
-between the emergence of an intelligent comunicating civilization and
-the next one.
-%
-The spatial structure of the simulation is given by the size and shape
-of the Galactic Habitable Zone and the maximum distance a signal can
-travel to be detected.
-%
-The parameters for the temporal distributions also determine the
-spatial properties since the density of active cetis in the Galaxy
-depend on these two parameters.
-%
-For example, a small $\tau_a$ and a large $\tau_s$ will produce a
-Galaxy densely populated woth CETIs, and viceversa.
-%
-Also, some hypotheses must be made in order to complete the
-simulation.
-%
-\citet{forgan_spatiotemporal_2011} argues that the times at which
-different civilizations become intelligent 
-follow a Gaussian distribution, and then the distribution of
-inter-arrival times is an inverse exponential.
-%
-We chose to assume that the distribution of the times of A events is a
-stationary Poisson process, and then the distribution of the times between the
-appeareance of new consecutive CETIs is exponential.
-%
-Regarding the duration of a CETI, we propose that the distribution of
-the duration of CETIs is a stationary exponential distribution.
-%
-\citet{balbi_impact_2018} stress the fact that it would be desirable to arrive
-at a theoretical statistical distribution of the lifespan of
-civilizations, preferentially on the basis of the underlying
-astrophysical and biological processes.
-%
-\citet{maccone_lognormals_2014} argues that this distribution should be a log normal.
-(EXPAND)
-
-There is no a temporal window for the model, i.e., the simulation
-starts when the process is already stable, and ends before any
-galactic effect could modify or affect of fixed variables.
-%
-The probability for the rise of a CETI is homogeneous over the GHZ.
-%
-Although the Galaxy has a well known spiral structure, the assumed
-sparcity of civilizations makes this assumption very reasonably.
-%
-Otherwise, if the distribution of civilizations is not sparce, it
-could be the case that the spiral arms would host most of the
-civilizations, and contacts are frecuent between closely located
-civilizations.
-%
-In such a case we could expect that the interpretation of our results
-could be considered as pesimist for the stablishment of a contact with
-our planet.
-%
-We also limit the posibilities of life or other types of civilizations
-to the hypothesis usually stated for the definition of the GHZ
-\citep{dayal_habitability_2016,gonzalez_galactic_2001,lineweaver_galactic_2004,gonzalez_habitable_2005,morrison_extending_2015,haqq-misra_evolution_2019,rahvar_cosmic_2016,gobat_evolution_2016,
-rahvar_cosmic_2016}.
-%
-\citet{vukotic_astrobiological_2012} propose probabilistic cellular
-automata model to explain the astrobiological history of the Milky
-Way.
-%
-This means that we set aside possibel civilizations that could survive
-in severe conditions or unstable systems, which would prevent the
-appeareance of life as we know it.
-
-\setlength{\tabcolsep}{10pt}
-\begin{table*}
-\centering
-\begin{tabular}{clccc}
-\hline
-   \multicolumn{2}{l}{independent variables} &min&max&Nbins\\
-\hline
-   $\tau_{a}$ & Mean temporal separation between consecutive awakenings & 4000 y & 200000 y & 50\\ 
-   $\tau_{s}$ & Mean lifetime of a CETI                                 & 10000 y & 500000 y& 50\\ 
-   D$_{max}$ & Maximum reach of a message  &  \multicolumn{3}{l}{500, 1000, 10000, 20000, 40000, 80000 lyr}  \\
-\hline
-   \multicolumn{4}{l}{fixed variables and assumptions} &value \\
-\hline
-   & statistical properties of all CETIs &equally distributed&&\\
-   & Point process for the distribution in time & homogeneous &&\\
-   f$_s$ & The scan of the sky & fully efficient&&1\\
-   f$_p$ & panspermia or colonization &absent&&0\\
-   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
-   R$_{SFC}$   & Inner radius of the GHZ     &&&20000 ly\\
-   R$_{SLC}$   & Outer radius of the GHZ     &&&60000 ly\\
-   t$_{max}$ & Time span of the simulation  &&&1.e6 y\\
-\hline
-   \multicolumn{5}{l}{discrete events}\\
-\hline
-   A & Awakening: a civilization starts a SETI program  &&&\\
-   B & Blackout: the end of the communication chanel stops &&&\\
-   C & Contact: a new causal contact is produced &&&\\
-   D & Doomsday: a civilization ends its communication capabilities&&&\\
-\hline
-
-\hline
-\end{tabular}
-\caption{Definition of independent variables and adopted values for 
-   fixed paramenters 
-   that are part of the simulation.  Variable parameters define the
-   spatial and temporal structure
-   of the process and the maximum reach of the messages.}
-\label{T_simu_hypotheses}
-\end{table*}
-
-While the aforementioned assumptions are basic conditions for most of
-the stochastic processes observed in nature, we need to make stronger
-assumptions related to the nature of the message.
-%
-The most simple assumption about the message itself, is that it
-travels at light speed.
-%
-With this, we are considering messages sent through, for example,
-electromagnetic radiation or even gravitational radiation, but we set
-appart messages sent with mechanical means or physical objects, or
-through some unknown technology that violates the known so far laws of
-physics.
-% 
-For the communication of CETIs through messages sent isotropically, we
-assume that the capacity to emit signals and to receive signals ocurr
-at the same time.
-%
-This means that the ability to find another civilization develops at
-the same time that the ability and intention to emit a signal in all
-direction, intented to be detected by an unknown civilization.
-%
-Although there are several reasons to think that this could not be the
-case, at large time scales it can be considered that both abilities
-occurr roughly at coincident epochs.
-%
-Another essential assumption is that all CETIs use the same signal
-power, so that there is a maximum distance out to which it can be
-detected.
-%
-It is worth mentioning that we are considering in our experimental
-setup a system composed by several emmiters and receivers across the
-Galaxy, under the hypothesis proposed before.
-%
-In such a system we compute probabilities of a CETI making contact
-with another CETI, and the results must be interpreted as for a
-randomly chosen CETI, given certain conditions of homogeneityon an
-ensemble of equivalent galaxies.
-%
-We stress the fact that we are not centered on the probabilities for
-the Earth itselft, but on a generalized CETI.
-%
-Regarding the extent of the signals, we know that the distance from
-which a signal from Earth could be detected using the current
-technology is of a few hundreds light years, given that the signal was
-sent to a specific direction.
-
-\begin{figure*}
-   \centering
-   \includegraphics[width=0.8\textwidth]{Messages_01.pdf}
-   \caption{Schematic representation of the possible cases in which a
-   ceti A can be in causal contact with a ceti B which appears later.
-   The duration of the causal contact in one direction depends on
-   several factors, mainly $\Delta t_A$, $\Delta t_C$, and $\Delta X$.}
-   \label{F_messages}
-\end{figure*}
-
-%https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
-%https://www.engadget.com/2017/05/25/listening-to-starlight-our-ongoing-search-for-alien-intelligenc/]
-%[from https://en.wikipedia.org/wiki/Fermi_paradox]
-%
-It is straightforward to propose and implement a distribution of
-maximum distances, although this would increase the model complexity
-at the cost of a larger uncertainty or of another dimension on the
-hypothesis space.
-%
-This proposal is based on a pronoid scenario (REF), and we do not
-consider in this work the ocurrence of paranoid or partially paranoid
-civilizations.
-%
-If that would be the case, the results obtained here can be taken as
-upper limits.
-
-In our simulations we also assume the simplest situations for the
-growth of ETIs.
-%
-For example, we discard the possibility of stellar colonization
-\citep[e.g.][]{newman_galactic_1981, walters_interstellar_1980,
-starling_virulence_2013, barlow_galactic_2012, jeong_large_2000, maccone_mathematical_2011}
-and assume that the communication aim is performed to
-all directions with the same power and the same probability, i.e.,
-strongly assume isotropic communication in all cases.
-%
-More detailed simulations could be produced considering different 
-efficiency of communication or detection methods.
-%
-Also, a mixed paranoid/pronoid population of intelligent civilizations
-could be implemented easily.
-%
-However, in this case we also would complicate the experimental setup,
-make the results less evident.
-%
-The results are independent of the nature of life (organic or
-artificial).
-%
-The lifetime of a civilization can be caused by autodestruction or by
-external factors.
-%}}}
-
-\subsection{Power laws vs. exponential laws}\label{SS_PDF_shape}
-%{{{
-
-The exponential distribution of lifespan and waiting times is
-justified by considering that the process of appeareance of life in
-the galaxy is homogeneous and stationary.
-%
-This means that there is no preferred location within the GHZ for the
-spontaneos appeareance of life, and that the emergence of a
-civilization is independent of the existence of previous civilizations 
-in the galaxy.
-%            
-This is equivalent to proposing a Poisson process for the emergence of
-CETIs, since there is a close relation between the number of events in
-time or space and the waiting time or separation, respectively.
-%
-That is, these are two alternative approaches to describing the same
-process, a Poisson distribution for the number of events implies an
-exponential distribution for their sepations, and viceversa.
-%
-It should be emphasized that the exponential laws used in this work
-are assumed, although instead of analyzing results from a particular
-parameter chosen ad--hoc, we explore the hypothesis space and analyze
-the impact of the values of these parameters on the results.
-%
-
-The power law and exponential statistical distributions are among the
-most common patterns found in natural phenomena.
-%
-For example, the distribution of the frecuency of words in many
-languajes is known to follow a Zipf law (which is a power law).
-%
-Zipf law also describes population ranks of cities in various
-countries, corporation sizes, income rankings, ranks of number of
-people watching the same TV channel (Zipf's law entry on Wikipedia).
-%
-The magnitudes of earthquakes, hurricanes, volcanic eruptions and
-floods; the sizes of meteorites or the losses caused by business
-interruptions from accidents, are also well described by power laws
-(Sornette 2006).
-%
-The power law behaviour has been observed in a variety of systems,
-including for example stock market fluctuations, sizes of computer
-files and word frequency in languages \citep{mitzenmacher_brief_2004,
-newman_power_2005, simkin_theory_2006}
-%
-Power laws have also been widely used in biological sciences, e.g., in
-analysing connectivity patterns in metabolic networks
-\citep{jeong_large_2000} and in the number of species observed per
-unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
-% 
-More examples can be found in the literature
-\citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009,
-maccone_evolution_2014, maccone_lognormals_2014,
-benguigui_classificacion_2016}
-
-%}}}
-
-\subsection{Discrete event process}
-%{{{
-
-Given a parametric model $M(\tau_a, \tau_s, D_{max})$, which includes the
-functional forms of the statistical distributions and fixed values for the
-hypotheses, a discrete event simulation is performed by keeping track of a set
-of variables that change each time a meaningful event is produced.
-%
-The main variables that follow the evolution of the simulation are the positions
-of stars, which are sampled randomly within the GHZ, the time of the awakening
-of each CETI (A event), and the time of the vanishing of each CETI (D event).
-%
-The variables that can be deduced from the previous ones include the number of
-CETIs in casual contact with at least another CETI at a given time, the number
-of CETIs as a function of time, the number of CETIs that receive a message at
-least one time, the number of CETIs that receive a message at least one time and
-successfully deliver an answer and the number distribution of waiting time to
-receive a message.
-%
-All these quantities are updated each time one of the four events (A, B, C, D)
-is produced.
-
-%%% comparar con el stochsatic cellular automata
-% ver libro Solving Fermi paradox, sec. 27.3.1
-
-\begin{figure} \centering
-   %
-   \includegraphics[width=0.5\textwidth]{F1.pdf}
-   %
-   \caption{Empirical cumulative distributions of the number contacts
-   for CETIs in four different samples (M5, M6, M7, M8) with long
-   range signal reach (D$_{max}$=10000).
-   %
-   }
-   % 
-\label{F_number_of_contacts} \end{figure}
-
-%}}}
-
-\subsection{Model for CETIs}
-%{{{
-
-Since the method is based on events, the first step is to define an
-architecture of events, the relationships between them and how events
-trigger changes on the state of the system.
-%
-We consider a simulated system that represents the galactic habitable
-zone (GHZ).
-%
-On a first approach, the GHZ is a 2-dimensional annular region.
-%
-The adopted values for the GHZ are 20 kly and 60 kly, for the inner
-and outer radius, respectively \citep{lineweaver_galactic_2004}.
-% ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
-% mas chico...
-%
-This simple model does not take into accont the variations in stellar
-density given by the spiral structure.
-%
-Although there are several possible approaches, we chose to follow the
-evolution of the system according to the following events:
-
-\begin{enumerate}
-   \item[(A)] A new CETI appears (Awakening)
-   \item[(D)] An old CETI dissappears (Doomsday)
-   \item[(C)] A new causal contact is stablished (Contact)
-   \item[(B)] An existing causal contact is interrupted (Blackout)
-\end{enumerate}
-
-The event of type B is produced when one of the two CETIs halts in its
-capability of emmiting and receiving signals.
-
-\begin{figure*} % 2D color plot
-   \centering
-   \includegraphics[width=\textwidth]{matrix_1.pdf}
-   \caption{Rate of CETIs that never make contact (listening) as a
-   function of $\tau_a$ and $\tau_s$, for 
-   $D_{max}$=10000 (left panel),
-   $D_{max}$=40000 (middle panel), and
-   $D_{max}$=80000 (right panel)}
-   \label{F_never_contact}
-\end{figure*}
-
-\begin{figure*} % 2D color plot
-   \centering
-   \includegraphics[width=\textwidth]{matrix_2.pdf}
-   \caption{Rate of CETIs that listen at the moment of the
-   awakening, as a funcion of $\tau_a$ and $\tau_s$, for
-   $D_{max}$=10000 (left panel),
-   $D_{max}$=40000 (middle panel), and
-   $D_{max}$=80000 (right panel)}
-   \label{F_C_at_A}
-\end{figure*}
-
-The system is updated each time an event is produced.
-%
-The time marks for each event (A and D for each CETI and C and B for each causal
-contact) are stored as a result of the simulation.
-%
-Also, the list of active CETIs is obtained as a function of time.
-%
-Some fixed variables must be set in order to carry out the simulation: the size
-of the Galactic Habitable Zone, the mean lifetime of a CETI, $\tau_s$, the mean
-waiting time for the appeareance of another CETI, $\tau_a$, and the maximum
-distance a message can be detected.
-
-The Galactic Habitable Zone is the only one of the four parameters that is
-barely known, is set at fixed values of the inner and outer radii, radial
-symetry is assumed, and the with of the galactic disk is neglected compared to
-the radial size.
-%
-The two radii are measured in light years, with the aim to maintain a single and
-comprehensive unit for both space and time coordinates.
-
-The simulation was implemented on a python 3 code, which is publically
-available\footnote{https://github.com/mlares/simu\_contact}
-
-%}}}
-
-\begin{figure}
-   %
-   \centering
-   %
-   \includegraphics[width=0.5\textwidth]{waiting_s+a_dif_ylog.pdf}
-   %
-   \caption{Histograms of the mean waiting times for the first contact, for
-   several models.
-   %
-   Left panel shows the histograms for several values of $\tau_a$, and
-   $\tau_s$ in the range 17000-24000 yr.
-   %
-   Right panel shows the histograms for several values of $\tau_s$, and
-   $\tau_a$ in the range 68000-100000 yr.
-   %
-   The shape of the histograms does not change when $\tau_a$ is
-   modified, but significantly changes when $\tau_s$ is modified.
-   }
-   %
-   \label{F_waiting_for_1C}
-   %
-\end{figure}
-
-\section{Results: exploring the parameter space}\label{S_results}
-%{{{
-
-We implemented the simulation of a regular grid of models varying over
-the hypothesis space, which covers 15000 different models.
-%
-For each model, we simulated 50 realizations with different random
-seeds, in order to estimate the variance.
-%
-The simulated parameters, $\tau_a$, $\tau_s$ cover the ranges
-4000-200000 and 10000-500000, respectively, with a partition of 50
-values for each parameter.
-%
-For the D$_{max}$ parameter, we chose the values 500, 1000, 10000,
-20000, 40000 and 80000 lyr.
-%
-This setup makes a total of 750000 simulations, each covering a time
-range of one million years.
-
-As a product of the simulations, several quantities are obtained.
-%
-Some quantities are directly derived from the discrete events, namely,
-the ID of emitting CETI, the ID of emitting (receiving) CETI, the
-position in the galaxy, the time of Awakening (Contact) and the time
-of Doomsday  (Blackout) for each CETI.
-%
-This structure is always for the A-D cycle of each emitting CETI, and
-repeats for each contact (C-B).
-%
-We can also derive the additional quantities.
-%
-The age of a CETI is the time elapsed from A to D.
+% Schematic representation of the growing communicating
+%   sphere, over the space--time diagrams, where time is represented on
+%   the horizontal direction, and space is represented in the vertical
+%   direction.  In (a) the sphere is
+%   growing as the surface of first contact has not reached the maximum
+%   distance.  In (b) it has reached the maximum distance, so that it
+%   remains at the same size.  After a Doomsday event, the signals can
+%   still be observed, but the surface of last contact grows.
+%   %
+%   Schematic representation of two emitters, $E_1$ and $E_2$,
+%   that reach each other at different times.  The time span for $E_i$
+%   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
+%   emitter $j$ between $C_{ij}$ and $B_{ij}$.
+%   The type and length of causal contact in both directions depend on
+%   the distance and time lag between the awakening events, the maximum
+%   distance that a signal can reach and the time period in which each
+%   emitter is active.
 %
 %
-%given time, ceti\_e is the CETI emitting signals (emiter), ceti\_r is
-%the CETI receiving signals (receiver), ceti\_c is the ceti that
-%listen at least another ceti (citizen) and ceti\_h is the CETI ceti
-%that is lestened by at least another CETI.
 %
-The time span of a CETI listening another or being listened by another
-can also be easily derived from the results of a realization.
-%
-This way we can also compute the distribution in the galaxy of cetis
-that reach contact and the waiting time until the first contact or the
-waiting time until the next contact.
-%
-Regarding the properties of the population of CETIs and its evolution,
-we analyze the fraction of awaken time a ceti is listening at least
-another ceti, the age of contacted CETIs at first contact, the
-distribution of time to wait until next contact, the fraction of cetis
-where the first contact is given at the awakening, the distribution of
-the number of contacts for each CETI, the distribution of the maximum
-number of contacts for each CETI, the distribution of the number of
-contacts as a function of CETI age, the number of contacts as a
-function of time in the galaxy, the rate of CETIs that succeed in
-contact, and the distribution of distances between contacted CETIs.
-%
-Usefull derived quantity is the duration of two-way communication
-channels the or the fraction of contacts that admit a response.
-% 
-We also analyze the relations between the distance to ceti vs. time of
-double communication, the distance to ceti vs. age of contacted ceti,
-the age of a ceti and the maximum number of contacted cetis before D,
-or the lifespan of a ceti vs. max number of contacts.
-%
-All these quantities can be analyzed as a function of the simulation
-parameters.
-%
-We chose eight models that are on fairly oposed regions of the
-explored parameter space, and cover short/long lifetimes, short/long
-waiting times for the next CETI to appear in the Galaxy and
-short/large range reach of the signal, including all possible
-combinations.
-%
-The details of each model are summarized in the
-Table~\ref{T_selected_models}.
-
-In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
-distributions of the number contacts for CETIs in four different
-samples (M5, M6, M7, M8) with long range signal reach
-(D$_{max}$=10000). 
-%
-The lifetime is more determinant than the awakening rate in increasing
-the number of contacts.
-%
-As expected, of the four samples, the sample M7, which corresponds to
-a dense awakening in the timeline and a long lifetime, is the model
-that maximizes the number of contacts, reaching a maximum of order 100
-contacts for a single CETI.
-%
-This is considerable larger than any model with a shorter lifetime,
-which produce a number of contacts of at most the order of ten
-contacts per CETI.
-%
-A couple of comments are worth to mention about this plot.
-%
-First, it has a logarithmic scale on the x-axis, and second it is the
-cumulative, not differential, empirical distribution.
-
-In what follows, we focus on the analysis of the duration of contacts
-(Sec.~\ref{SS_multiplicity}), the waiting time for the first contact
-(Sec.~\ref{SS_waiting}) and the
-variations of contact posibilities as a function of the position of
-the CETI in the Galaxy (Sec.~\ref{SS_location}).
-
-%}}}
-
-\subsection{Multiplicity and duration of contacts}\label{SS_multiplicity}
-%{{{
-
-The duration of each contact for a single receiving CETI is given by the
-time difference between a C event and a B event, which are direct
-outputs for each simulation experiment.
-
-%}}}
-
-\subsection{Waiting time for a first contact}\label{SS_waiting}
-%{{{
-
-In this section we analyze the distribution of the waiting times for
-a first contact.
-%
-%Such distribution can be considered to compute the probability of
-%reaching a first contact within the first XX years of a SETI project,
-%assuming the methods are correct and the hypotheses of the experiment.
-Such distribution can be considered to compute the probability for a
-random CETI to carry out a SETI project and spend a given time until
-the first contact is made.
-
-In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the CETIs
-that have made at least one contact as a funtion of the elapsed time
-since the awakening, given that they make at least one contact on
-their lifetime, for different models.
-%
-From a frequentist approach, the normalized cumulative counts are
-related to the estimation of the probability of observing a given time
-with no success.
-%             
-Both panels show the histograms of the waiting times for the first
-contact, for all the CETIs that make at least one contact along their
-lifetime.
-%
-We show separately the histograms for several values of $\tau_a$ and a
-fixed value of $(\tau_s, D_{max})$, and for several values of $\tau_s$
-and a fixed value of $(\tau_a, D_{max})$.
-%
-For different values of $\tau_a$, the shape of the distribution is
-similar, and the main difference given by the different number of
-contacts on the different models.
-%
-Conversely, for different values of $\tau_a$, the shape of the
-distribution is similar, and the main difference given by the
-different number of contacts on the different models. 
-%
-We show the 2D maps of the rate of CETIs that never make contact and
-the rate of CETIs that make contact at the moment of the awakening,
-inf the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
-as a function of $\tau_s$ and $\tau_s$.
-%
-In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
-waiting times for the first contact, for several models.
-%
-Left panel shows the histograms for several values of $\tau_a$, and
-$\tau_s$ in the range 17000-24000 yr.
-%
-Right panel shows the histograms for several values of $\tau_s$, and
-$\tau_a$ in the range 68000-100000 yr.
-%
-The shape of the histograms does not change when $\tau_a$ is modified,
-but significantly changes when $\tau_s$ is modified.
-
-%}}}
-
 %   \begin{figure}
@@
 %   \end{figure}
-
-\subsection{Location within the Galaxy}\label{SS_location}
-%{{{
-
-%On the Figure~\ref{F_radial_frac} we show the fraction of CETIs that make
-%zero, one, two or three contacts, as a function of the galactocentric
-%distance, for model M1.
-% VERIFICAR!  M1?
-%
-For this model, the majority of CETIs do not
-make any contact during all the time period in which they are active.
-%
-As expected, the fractions of CETIs making at least one contact
-diminishes as the number of contacts rises.
-%
-The shape of this distribution is a result of the geometry of the GHZ.
-%
-In general, a closer position to the center of the Galaxy slightly favours 
-the posibility of a contact.
-%
-This would indicate that the position of the Earth is suitable for the
-odds of making a contact.
-
 %}}}
-
-\setlength{\tabcolsep}{10pt}
-\begin{table*}
-\centering
-\begin{tabular}{ccccc}
-\hline
-Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & description  \\
-\hline
-M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
-M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
-M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
-M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
-%
-M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
-M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
-M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
-M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
-%
-\hline
-\end{tabular}
-\caption{Selected models to analize the behavior of simulation outputs
-   and their dependence on simulation outputs and their dependence on
-   simulation parameters.}
-\label{T_selected_models}
-\end{table*}
-
-\section{Discusion}\label{S_discussion}
-%{{{
-
-We have implemented a suite of simulations following a stochastic
-approach, to explore the hypothesis
-space of a simple model that accounts for
-the causal connections between communicating
-civilizations on an simplified galaxy.
-%
-The different models can be generated with a minimal number of free
-parameters and some fixed assumptions.
-%
-We argue that three parameters can be used to describe a variety of
-situations, ranging from a Galaxy where an intelligent civilization is
-very rare, to a Galaxy populated with plenty of civilizations in
-causal contact.
-%
-We explored the consecuences of different models using mensurable
-quantities, that arise as by-products of the simulations.
-%
-With this approach, we can estimate several quantities as a function
-of the parameters on the hypothesis space.
-%
-For instance, it is possible to estimate the timescale for an
-uninterrupted SETI effort to reach success, in terms of different
-model parameters that reflect different, so far unknown, scenarios
-for the spread of intelligent life on the galaxy.
-
-Although the simulations use a number of speculative assumptions, we
-argue that the degree of knowledge about the origin and persistence of
-life in the Galaxy does not worth the implementation of more detailed
-or sophisticated models.
-%
-Instead, we take advantadge of the simplicity of the model to explore
-the hypothesis space, i.e., the three--dimensional parameter space of
-the parametric model, in order to gain insight on the consecuences of
-different scenarios for the search of intelligent life.
-%
-Our analysis is not centered in obtaining the odds for the Earth
-making contact with another extraterrestrial civilization, but in 
-obtaining a statistical, parameter dependent description of the
-communication system consisting of a group of broadcasting
-civilizations which arise sparsely both in time and space in the
-Galaxy.
-
-Under the hypotheses of our experiments, we conclude that
-a causal contact is extremely unlikely unless the Galaxy is heavily
-populated by intelligent civilizations.
-%
-Even so, we must consider that in order to stablish a contact between
-any two entities, a minimum degree of compatibility must be
-accomplished without any previous agreement, 
-making the posibility of a contact with a message that
-could be deciphered highly rare.
-
-Taking into account the caveat mentioned previously about the
-trade--off between the simplicity and complexity of the experiment,
-future work can be performed following this framework in order to
-explore possible implications of the results for more detailed
-configurations.
-%
-For example, the communication method (isotropic, colimated,
-serendipitous) can affect the observables.
-%
-Since we analyze the causal contacts, effective contacts depend on the
-efficiency of both the emitter and the receiver.
-%
-A correction by coverage ratio in the detection and by a targetting
-ratio in the emission could be easily implemented in the simulation,
-although the effect would be to reduce the probability of contact by
-an efficiency ratio.
-%
-Thus, such implementation is not really necessary, since the values of
-probabilities could modified by a constant correction factor equal to
-the combined emission/reception efficiency
-\citep{smith_broadcasting_2009, anchordoqui_upper_2019, forgan_collimated_2014}.
-%
-If we consider a statistical distribution for this factor, then the
-results could change, and this could be studied with the DE approach.
-%
-Also the nature of the message carrier, electromagnetic or another.
-%
-It is also possible to explore the effects of alignments or the use of
-stars as sources or amplifiers.
-%
-%energetic markers
-%Spiritual markers
-%
-As a future work, we will explore several improvements to the
-simulation, and analyze the impact on the final results of these
-changes.
-%
-They include a spatial distribution that actually resembles the shape
-of the Galaxy, different distribution functions for the mean lifetime
-of CETIs and different efficiencies for the sphere of the causal
-region of each CETI, resembling different searching strategies
-\citep{hippke_interstellar_2017}.
-
-Possible hypothesis improvements are:
-
-\begin{itemize}
-   \item Dmax is different for different CETIs.  For example a power
-      law where powerfull emmisions are rare and weak emssions are
-      common.
-   \item The probability of the rising of a new CETI vary within the
-      GHZ.
-   \item Correction by coverage ratio in the detection and by
-      targetting ratio in the emission.
-   \item Spiritual factor.  Explore the role of the S factor: message
-      contents, influence on the lifespan of a CETI that receives a
-      message, mean lifespan of a CETI that emits a message.
-\end{itemize}
-
-%}}}
-
-\ack[Acknowledgement]{
-%
-This work was partially supported by the Consejo Nacional de
-Investigaciones Cient\'{\i}ficas y T\'ecnicas (CONICET, Argentina),
-the Secretar\'{\i}a de Ciencia y Tecnolog\'{\i}a, Universidad Nacional
-de C\'ordoba, Argentina, and the Universidad Cat\'olica de C\'ordoba,
-Argentina.
-%
-This research has made use of NASA's Astrophysics Data System.
-%
-Plots and simulations were made with software developed by the authors
-using R and python languajes. Plots were postprocessed with inkscape.
-%
-}
-
-\bibliography{biblio_seti}
-
-\end{document}

### 2019-12-19 manuscript.tex
reestructuracion

--- 2019-11-28 manuscript.tex
@@
-\documentclass[crop]{CSLB}%%%%where CSL is the template name
+\documentclass[crop]{CSLB}
 
-%The authors can define any packages after the \documentclass{CSL} command.
-\usepackage{amsmath} %for dealing with mathematics,
-\usepackage{amsthm} %for dealing with theorem environments,
-\usepackage{cite} %%%for dealing with citations
-\usepackage{hyperref} %for linking the cross references
-\usepackage{graphics} %for dealing with figures.
-\usepackage{algorithmic} %for describing algorithms
-\usepackage{subfig} %for getting the subfigures e.g., "Figure 1a and 1b" etc.
-\usepackage{url} %It provides better support for handling and breaking URLs.
+\usepackage{MySettings}
 
-%\usepackage{natbib}
-\usepackage{boites}
-
-%\usepackage{graphicx}
-%\usepackage{float}
-%\usepackage{psfig}
-%\usepackage{amsfonts,amssymb}
-%\usepackage{color}
-%\usepackage{textcomp}
-%\usepackage[T1]{fontenc}
-%\usepackage{aecompl}
-%\usepackage{times}
-%\usepackage{txfonts}
-%\usepackage[english]{babel}
-%\usepackage[utf8x]{inputenc}
-%\usepackage{amsmath}
-%\usepackage{caption}
-%\usepackage{siunitx}
-
-\jname{Astrobiology}%
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{condition}{Condition}
+\input{captions}
 
@@
 contact between interstellar civilizations through MonteCarlo simulations}
-
-%% Stochastic estimation of the waiting time for the first causal contact
-%% with an extraterrestrial intelligence
 
@@
 \maketitle
-
-\Fpagebreak
 
@@
 %
+
+\Fpagebreak
+
 Temporal aspects of the distribution of communicating civilizations
@@
 
-\section{Methods and working hypotheses}\label{S_methods}
-%{{{
+\input{methods}
 
-Some complex stochastic processes can be efficiently modeled with the
-discrete--event (DE) simulation approach.
-%
-Simulations are suitable tools to analyze systems that evolve with
-time and involve randomness.
-%
-In general, simulations require less assumptions and simplifications
-than theoretical approaches, and can be applied to systems where a
-theoretical model can not be found.
-%
-A system described with the DE paradigm is characterized by a set of
-actors and events, where actors interact causally through a series of
-events on a time line and process these events in chronological order
-\citep{ptolemaeus_system_2014, chung_simulation_2003,
-ross_simulation_2012}.
-%
-In the case of study, the actors are the communicating civilizations
-(CETIs), and the events are all meaningfull that produce a change in
-the simulation variables, i.e., in the arrangement of CETIs and their
-causal contacts.
-%
-Since it allows to explore several models easily, 
-this method is well suited for the particular case of the difusion of
-intelligence signals in the Galaxy.
-%
-Assuming some simple hypotheses, the discrete events method can be
-performed taking into account a small number of variables, which
-allow to analyze the variation of the results in the model parameter space.
-
-In what follows, we describe the experimental setup chosen to estimate
-the probabilities of causal contacts and several derived quantities in
-terms of three independent parameters, namely, the mean time span
-between the appeareance of consecutive communicating civilizations
-($\tau_a$), their mean lifetime ($\tau_s$), and the maximum distance a
-signal can be detected by another civilization ($D_{max}$).
-%
-Intuitively, the shorter $\tau_a$ and the larger $\tau_s$ and
-$D_{max}$, there is a greater probability of the existence of causal
-contacts between pairs of intelligent civilizations in the Galaxy
-(Table \ref{T_simu_hypotheses}).
-%
-We also need to propose theoretical distribution functions for both the
-distribution of lifespans and the distribution of the number of CETIs
-per unit time \citep{maccone_evolution_2014,Sotos_biotechnology_2019}.
-%
-The shapes of these distributions are discussed in Sec.~\ref{SS_PDF_shape}.
-%
-A causal connection can be produced without the need of the two
-involved civilizations being concurrent or active at the same time.
-%
-This is a fundamental property of the system which arises as a
-consecuence of the large spatial and temporal scales, where a message
-is transmited at light speed.
-%
-As a consecuence of their finite lifetime,
-each CETI will fill a spherical shell region of the Galaxy, limited by two
-surfaces.
-%
-The leading front, or surface of first contact (SFC) grows from the
-central CETI until it reaches the maximum distance $D_{max}$.
-%
-Although a CETI could be old enough to transmit a message at large
-distances, the limited power of the message and the dilution that runs
-as the squared distance from the source, imposes a
-detectability limit.
-%
-Thus, the volume reached causally by a CETI is initially a growing
-sphere.
-%
-Following the end of the civilization, there is still a region which
-is covered with the emited signals.
-%
-The trailing front, or surface of last contact (SLC) also grows from
-the central CETI, with a delay with respect to the SFC equivalent to
-the lifetime of the CETI, and produces a spherical shell region.
-%
-Any other CETI within this region will be aware of the originating CETI,
-even if it has disapeared before the time of contact.
-%
-This region will grow if the SFC has not reached $D_{max}$, and
-will shrink otherwise, until the SLC reaches $D_{max}$.
-%
-A that point, all signal from the central CETI is lost.
-%
-In the 2D simulation only the intersection of these spheres with the
-plane of the galaxy is relevant, and produce the corresponding
-circles for the filled spheres or annular regions for the spherical
-shells.
-%
-We illustrate in the Fig.~\ref{F_scheme} the schematic
-representation of the region causally connected to the central CETI.
-%
-%
-%
-% Schematic representation of the growing communicating
-%   sphere, over the space--time diagrams, where time is represented on
-%   the horizontal direction, and space is represented in the vertical
-%   direction.  In (a) the sphere is
-%   growing as the surface of first contact has not reached the maximum
-%   distance.  In (b) it has reached the maximum distance, so that it
-%   remains at the same size.  After a Doomsday event, the signals can
-%   still be observed, but the surface of last contact grows.
-%   %
-%   Schematic representation of two emitters, $E_1$ and $E_2$,
-%   that reach each other at different times.  The time span for $E_i$
-%   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
-%   emitter $j$ between $C_{ij}$ and $B_{ij}$.
-%   The type and length of causal contact in both directions depend on
-%   the distance and time lag between the awakening events, the maximum
-%   distance that a signal can reach and the time period in which each
-%   emitter is active.
-%
-%
-%
-The initially growing communicating sphere, is shown over
-space--time diagrams, where time is represented on the horizontal
-direction, and space is represented in the vertical direction.
-%
-In the upper panel (a), the sphere is growing as the surface of first
-contact has not reached the maximum distance.
-%
-In the middle panel (b) it has reached the maximum distance, so that
-it remains at the same size.
-%
-After a Doomsday event (bottom panel, b), the signals can still be
-observed, but the surface of last contact grows.
-%
-We also show in this Figure a schematic representation of two
-emitters, E$_1$ and E$_2$, that reach each other at different times.
-%
-The time span for E$_i$ is (A$_i$ , D$_i$) , for $i = {1, 2}$.
-%
-Emitter $i$ can listen to emitter $j$ between C$_{ij}$ and B$_{ij}$.
-%
-The type and length of causal contact in both directions depend on the
-distance and time lag between the awakening events, the maximum
-distance that a signal can reach and the time period in which each
-emitter is active.
-%
-In the Fig.~\ref{F_messages} we show a schematic representation of the
-possible cases in which a CETI A can be in causal contact with a CETI
-B which appears later.
-%
-The duration of the causal contact in one direction depends on several
-factors, mainly $\Delta$t~A , $\Delta$t~C , and $\Delta$X.
-
-%\begin{figure}
-%   \centering
-%   \includegraphics[width=0.5\textwidth]{growingsphere.pdf}
-%   \caption{Schematic representation of the growing communicating
-%   sphere, over the space--time diagrams, where time is represented on
-%   the horizontal direction, and space is represented in the vertical
-%   direction.  In (a) the sphere is
-%   growing as the surface of first contact has not reached the maximum
-%   distance.  In (b) it has reached the maximum distance, so that it
-%   remains at the same size.  After a Doomsday event, the signals can
-%   still be observed, but the surface of last contact grows }
-%   \label{F_growing_sphere}
-%\end{figure}
-%
-%  
-%\begin{figure}
-%   \centering
-%   \includegraphics[width=0.5\textwidth]{abcd.pdf}
-%   \caption{Schematic representation of two emitters, $E_1$ and $E_2$,
-%   that reach each other at different times.  The time span for $E_i$
-%   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
-%   emitter $j$ between $C_{ij}$ and $B_{ij}$.
-%   The type and length of causal contact in both directions depend on
-%   the distance and time lag between the awakening events, the maximum
-%   distance that a signal can reach and the time period in which each
-%   emitter is active.}
-%   \label{F_abcd}
-%\end{figure}
-
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{scheme.pdf}
-   \caption{Space--time diagrams showing a schematic representation of the 
-   different stages in the development of a communicating
-   civilization, the growing communicating sphere (a) and the temporal
-   intervals for the communication between two civilizations (b).
-   %
-   In the upper panel (a), the sphere the encloses the region which is
-   causally connected to the emitter grows until it reaches the
-   $D_{max}$ distance.  After the $D$ event, that region is limited by
-   the surface of last contact.
-   %
-   The scheme in the lower panel (b) represents
-   two communicating civilizations that have
-   causal contact in both directions}   
-   \label{F_scheme}
-\end{figure}
-
-The temporal structure of the process is defined by two distribution
-parameters that represent the mean time interval that an intelligent
-civilization can emit and receive signals, and the mean time interval
-between the emergence of an intelligent comunicating civilization and
-the next one.
-%
-The spatial structure of the simulation is given by the size and shape
-of the Galactic Habitable Zone and the maximum distance a signal can
-travel to be detected.
-%
-The parameters for the temporal distributions also determine the
-spatial properties since the density of active cetis in the Galaxy
-depend on these two parameters.
-%
-For example, a small $\tau_a$ and a large $\tau_s$ will produce a
-Galaxy densely populated woth CETIs, and viceversa.
-%
-Also, some hypotheses must be made in order to complete the
-simulation.
-%
-\citet{forgan_spatiotemporal_2011} argues that the times at which
-different civilizations become intelligent 
-follow a Gaussian distribution, and then the distribution of
-inter-arrival times is an inverse exponential.
-%
-We chose to assume that the distribution of the times of A events is a
-stationary Poisson process, and then the distribution of the times between the
-appeareance of new consecutive CETIs is exponential.
-%
-Regarding the duration of a CETI, we propose that the distribution of
-the duration of CETIs is a stationary exponential distribution.
-%
-\citet{balbi_impact_2018} stress the fact that it would be desirable to arrive
-at a theoretical statistical distribution of the lifespan of
-civilizations, preferentially on the basis of the underlying
-astrophysical and biological processes.
-%
-\citet{maccone_lognormals_2014} argues that this distribution should be a log normal.
-(EXPAND)
-
-There is no a temporal window for the model, i.e., the simulation
-starts when the process is already stable, and ends before any
-galactic effect could modify or affect of fixed variables.
-%
-The probability for the rise of a CETI is homogeneous over the GHZ.
-%
-Although the Galaxy has a well known spiral structure, the assumed
-sparcity of civilizations makes this assumption very reasonably.
-%
-Otherwise, if the distribution of civilizations is not sparce, it
-could be the case that the spiral arms would host most of the
-civilizations, and contacts are frecuent between closely located
-civilizations.
-%
-In such a case we could expect that the interpretation of our results
-could be considered as pesimist for the stablishment of a contact with
-our planet.
-%
-We also limit the posibilities of life or other types of civilizations
-to the hypothesis usually stated for the definition of the GHZ
-\citep{dayal_habitability_2016,gonzalez_galactic_2001,lineweaver_galactic_2004,gonzalez_habitable_2005,morrison_extending_2015,haqq-misra_evolution_2019,rahvar_cosmic_2016,gobat_evolution_2016,
-rahvar_cosmic_2016}.
-%
-\citet{vukotic_astrobiological_2012} propose probabilistic cellular
-automata model to explain the astrobiological history of the Milky
-Way.
-%
-This means that we set aside possibel civilizations that could survive
-in severe conditions or unstable systems, which would prevent the
-appeareance of life as we know it.
-
-\setlength{\tabcolsep}{10pt}
-\begin{table*}
-\centering
-\begin{tabular}{clccc}
-\hline
-   \multicolumn{2}{l}{independent variables} &min&max&Nbins\\
-\hline
-   $\tau_{a}$ & Mean temporal separation between consecutive awakenings & 4000 y & 200000 y & 50\\ 
-   $\tau_{s}$ & Mean lifetime of a CETI                                 & 10000 y & 500000 y& 50\\ 
-   D$_{max}$ & Maximum reach of a message  &  \multicolumn{3}{l}{500, 1000, 10000, 20000, 40000, 80000 lyr}  \\
-\hline
-   \multicolumn{4}{l}{fixed variables and assumptions} &value \\
-\hline
-   & statistical properties of all CETIs &equally distributed&&\\
-   & Point process for the distribution in time & homogeneous &&\\
-   f$_s$ & The scan of the sky & fully efficient&&1\\
-   f$_p$ & panspermia or colonization &absent&&0\\
-   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
-   R$_{SFC}$   & Inner radius of the GHZ     &&&20000 ly\\
-   R$_{SLC}$   & Outer radius of the GHZ     &&&60000 ly\\
-   t$_{max}$ & Time span of the simulation  &&&1.e6 y\\
-\hline
-   \multicolumn{5}{l}{discrete events}\\
-\hline
-   A & Awakening: a civilization starts a SETI program  &&&\\
-   B & Blackout: the end of the communication chanel stops &&&\\
-   C & Contact: a new causal contact is produced &&&\\
-   D & Doomsday: a civilization ends its communication capabilities&&&\\
-\hline
-
-\hline
-\end{tabular}
-\caption{Definition of independent variables and adopted values for 
-   fixed paramenters 
-   that are part of the simulation.  Variable parameters define the
-   spatial and temporal structure
-   of the process and the maximum reach of the messages.}
-\label{T_simu_hypotheses}
-\end{table*}
-
-While the aforementioned assumptions are basic conditions for most of
-the stochastic processes observed in nature, we need to make stronger
-assumptions related to the nature of the message.
-%
-The most simple assumption about the message itself, is that it
-travels at light speed.
-%
-With this, we are considering messages sent through, for example,
-electromagnetic radiation or even gravitational radiation, but we set
-appart messages sent with mechanical means or physical objects, or
-through some unknown technology that violates the known so far laws of
-physics.
-% 
-For the communication of CETIs through messages sent isotropically, we
-assume that the capacity to emit signals and to receive signals ocurr
-at the same time.
-%
-This means that the ability to find another civilization develops at
-the same time that the ability and intention to emit a signal in all
-direction, intented to be detected by an unknown civilization.
-%
-Although there are several reasons to think that this could not be the
-case, at large time scales it can be considered that both abilities
-occurr roughly at coincident epochs.
-%
-Another essential assumption is that all CETIs use the same signal
-power, so that there is a maximum distance out to which it can be
-detected.
-%
-It is worth mentioning that we are considering in our experimental
-setup a system composed by several emmiters and receivers across the
-Galaxy, under the hypothesis proposed before.
-%
-In such a system we compute probabilities of a CETI making contact
-with another CETI, and the results must be interpreted as for a
-randomly chosen CETI, given certain conditions of homogeneityon an
-ensemble of equivalent galaxies.
-%
-We stress the fact that we are not centered on the probabilities for
-the Earth itselft, but on a generalized CETI.
-%
-Regarding the extent of the signals, we know that the distance from
-which a signal from Earth could be detected using the current
-technology is of a few hundreds light years, given that the signal was
-sent to a specific direction.
-
-\begin{figure*}
-   \centering
-   \includegraphics[width=0.8\textwidth]{Messages_01.pdf}
-   \caption{Schematic representation of the possible cases in which a
-   ceti A can be in causal contact with a ceti B which appears later.
-   The duration of the causal contact in one direction depends on
-   several factors, mainly $\Delta t_A$, $\Delta t_C$, and $\Delta X$.}
-   \label{F_messages}
-\end{figure*}
-
-%https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
-%https://www.engadget.com/2017/05/25/listening-to-starlight-our-ongoing-search-for-alien-intelligenc/]
-%[from https://en.wikipedia.org/wiki/Fermi_paradox]
-%
-It is straightforward to propose and implement a distribution of
-maximum distances, although this would increase the model complexity
-at the cost of a larger uncertainty or of another dimension on the
-hypothesis space.
-%
-This proposal is based on a pronoid scenario (REF), and we do not
-consider in this work the ocurrence of paranoid or partially paranoid
-civilizations.
-%
-If that would be the case, the results obtained here can be taken as
-upper limits.
-
-In our simulations we also assume the simplest situations for the
-growth of ETIs.
-%
-For example, we discard the possibility of stellar colonization
-\citep[e.g.][]{newman_galactic_1981, walters_interstellar_1980,
-starling_virulence_2013, barlow_galactic_2012, jeong_large_2000, maccone_mathematical_2011}
-and assume that the communication aim is performed to
-all directions with the same power and the same probability, i.e.,
-strongly assume isotropic communication in all cases.
-%
-More detailed simulations could be produced considering different 
-efficiency of communication or detection methods.
-%
-Also, a mixed paranoid/pronoid population of intelligent civilizations
-could be implemented easily.
-%
-However, in this case we also would complicate the experimental setup,
-make the results less evident.
-%
-The results are independent of the nature of life (organic or
-artificial).
-%
-The lifetime of a civilization can be caused by autodestruction or by
-external factors.
-%}}}
-
-\subsection{Power laws vs. exponential laws}\label{SS_PDF_shape}
-%{{{
-
-The exponential distribution of lifespan and waiting times is
-justified by considering that the process of appeareance of life in
-the galaxy is homogeneous and stationary.
-%
-This means that there is no preferred location within the GHZ for the
-spontaneos appeareance of life, and that the emergence of a
-civilization is independent of the existence of previous civilizations 
-in the galaxy.
-%            
-This is equivalent to proposing a Poisson process for the emergence of
-CETIs, since there is a close relation between the number of events in
-time or space and the waiting time or separation, respectively.
-%
-That is, these are two alternative approaches to describing the same
-process, a Poisson distribution for the number of events implies an
-exponential distribution for their sepations, and viceversa.
-%
-It should be emphasized that the exponential laws used in this work
-are assumed, although instead of analyzing results from a particular
-parameter chosen ad--hoc, we explore the hypothesis space and analyze
-the impact of the values of these parameters on the results.
-%
-
-The power law and exponential statistical distributions are among the
-most common patterns found in natural phenomena.
-%
-For example, the distribution of the frecuency of words in many
-languajes is known to follow a Zipf law (which is a power law).
-%
-Zipf law also describes population ranks of cities in various
-countries, corporation sizes, income rankings, ranks of number of
-people watching the same TV channel (Zipf's law entry on Wikipedia).
-%
-The magnitudes of earthquakes, hurricanes, volcanic eruptions and
-floods; the sizes of meteorites or the losses caused by business
-interruptions from accidents, are also well described by power laws
-(Sornette 2006).
-%
-The power law behaviour has been observed in a variety of systems,
-including for example stock market fluctuations, sizes of computer
-files and word frequency in languages \citep{mitzenmacher_brief_2004,
-newman_power_2005, simkin_theory_2006}
-%
-Power laws have also been widely used in biological sciences, e.g., in
-analysing connectivity patterns in metabolic networks
-\citep{jeong_large_2000} and in the number of species observed per
-unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
-% 
-More examples can be found in the literature
-\citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009,
-maccone_evolution_2014, maccone_lognormals_2014,
-benguigui_classificacion_2016}
-
-%}}}
-
-\subsection{Discrete event process}
-%{{{
-
-Given a parametric model $M(\tau_a, \tau_s, D_{max})$, which includes the
-functional forms of the statistical distributions and fixed values for the
-hypotheses, a discrete event simulation is performed by keeping track of a set
-of variables that change each time a meaningful event is produced.
-%
-The main variables that follow the evolution of the simulation are the positions
-of stars, which are sampled randomly within the GHZ, the time of the awakening
-of each CETI (A event), and the time of the vanishing of each CETI (D event).
-%
-The variables that can be deduced from the previous ones include the number of
-CETIs in casual contact with at least another CETI at a given time, the number
-of CETIs as a function of time, the number of CETIs that receive a message at
-least one time, the number of CETIs that receive a message at least one time and
-successfully deliver an answer and the number distribution of waiting time to
-receive a message.
-%
-All these quantities are updated each time one of the four events (A, B, C, D)
-is produced.
-
-%%% comparar con el stochsatic cellular automata
-% ver libro Solving Fermi paradox, sec. 27.3.1
-
-\begin{figure} \centering
-   %
-   \includegraphics[width=0.5\textwidth]{F1.pdf}
-   %
-   \caption{Empirical cumulative distributions of the number contacts
-   for CETIs in four different samples (M5, M6, M7, M8) with long
-   range signal reach (D$_{max}$=10000).
-   %
-   }
-   % 
-\label{F_number_of_contacts} \end{figure}
-
-%}}}
-
-\subsection{Model for CETIs}
-%{{{
-
-Since the method is based on events, the first step is to define an
-architecture of events, the relationships between them and how events
-trigger changes on the state of the system.
-%
-We consider a simulated system that represents the galactic habitable
-zone (GHZ).
-%
-On a first approach, the GHZ is a 2-dimensional annular region.
-%
-The adopted values for the GHZ are 20 kly and 60 kly, for the inner
-and outer radius, respectively \citep{lineweaver_galactic_2004}.
-% ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
-% mas chico...
-%
-This simple model does not take into accont the variations in stellar
-density given by the spiral structure.
-%
-Although there are several possible approaches, we chose to follow the
-evolution of the system according to the following events:
-
-\begin{enumerate}
-   \item[(A)] A new CETI appears (Awakening)
-   \item[(D)] An old CETI dissappears (Doomsday)
-   \item[(C)] A new causal contact is stablished (Contact)
-   \item[(B)] An existing causal contact is interrupted (Blackout)
-\end{enumerate}
-
-The event of type B is produced when one of the two CETIs halts in its
-capability of emmiting and receiving signals.
-
-\begin{figure*} % 2D color plot
-   \centering
-   \includegraphics[width=\textwidth]{matrix_1.pdf}
-   \caption{Rate of CETIs that never make contact (listening) as a
-   function of $\tau_a$ and $\tau_s$, for 
-   $D_{max}$=10000 (left panel),
-   $D_{max}$=40000 (middle panel), and
-   $D_{max}$=80000 (right panel)}
-   \label{F_never_contact}
-\end{figure*}
-
-\begin{figure*} % 2D color plot
-   \centering
-   \includegraphics[width=\textwidth]{matrix_2.pdf}
-   \caption{Rate of CETIs that listen at the moment of the
-   awakening, as a funcion of $\tau_a$ and $\tau_s$, for
-   $D_{max}$=10000 (left panel),
-   $D_{max}$=40000 (middle panel), and
-   $D_{max}$=80000 (right panel)}
-   \label{F_C_at_A}
-\end{figure*}
-
-The system is updated each time an event is produced.
-%
-The time marks for each event (A and D for each CETI and C and B for each causal
-contact) are stored as a result of the simulation.
-%
-Also, the list of active CETIs is obtained as a function of time.
-%
-Some fixed variables must be set in order to carry out the simulation: the size
-of the Galactic Habitable Zone, the mean lifetime of a CETI, $\tau_s$, the mean
-waiting time for the appeareance of another CETI, $\tau_a$, and the maximum
-distance a message can be detected.
-
-The Galactic Habitable Zone is the only one of the four parameters that is
-barely known, is set at fixed values of the inner and outer radii, radial
-symetry is assumed, and the with of the galactic disk is neglected compared to
-the radial size.
-%
-The two radii are measured in light years, with the aim to maintain a single and
-comprehensive unit for both space and time coordinates.
-
-The simulation was implemented on a python 3 code, which is publically
-available\footnote{https://github.com/mlares/simu\_contact}
-
-%}}}
-
-\begin{figure}
-   %
-   \centering
-   %
-   \includegraphics[width=0.5\textwidth]{waiting_s+a_dif_ylog.pdf}
-   %
-   \caption{Histograms of the mean waiting times for the first contact, for
-   several models.
-   %
-   Left panel shows the histograms for several values of $\tau_a$, and
-   $\tau_s$ in the range 17000-24000 yr.
-   %
-   Right panel shows the histograms for several values of $\tau_s$, and
-   $\tau_a$ in the range 68000-100000 yr.
-   %
-   The shape of the histograms does not change when $\tau_a$ is
-   modified, but significantly changes when $\tau_s$ is modified.
-   }
-   %
-   \label{F_waiting_for_1C}
-   %
-\end{figure}
-
-\section{Results: exploring the parameter space}\label{S_results}
-%{{{
-
-We implemented the simulation of a regular grid of models varying over
-the hypothesis space, which covers 15000 different models.
-%
-For each model, we simulated 50 realizations with different random
-seeds, in order to estimate the variance.
-%
-The simulated parameters, $\tau_a$, $\tau_s$ cover the ranges
-4000-200000 and 10000-500000, respectively, with a partition of 50
-values for each parameter.
-%
-For the D$_{max}$ parameter, we chose the values 500, 1000, 10000,
-20000, 40000 and 80000 lyr.
-%
-This setup makes a total of 750000 simulations, each covering a time
-range of one million years.
-
-As a product of the simulations, several quantities are obtained.
-%
-Some quantities are directly derived from the discrete events, namely,
-the ID of emitting CETI, the ID of emitting (receiving) CETI, the
-position in the galaxy, the time of Awakening (Contact) and the time
-of Doomsday  (Blackout) for each CETI.
-%
-This structure is always for the A-D cycle of each emitting CETI, and
-repeats for each contact (C-B).
-%
-We can also derive the additional quantities.
-%
-The age of a CETI is the time elapsed from A to D.
-%
-%
-%given time, ceti\_e is the CETI emitting signals (emiter), ceti\_r is
-%the CETI receiving signals (receiver), ceti\_c is the ceti that
-%listen at least another ceti (citizen) and ceti\_h is the CETI ceti
-%that is lestened by at least another CETI.
-%
-The time span of a CETI listening another or being listened by another
-can also be easily derived from the results of a realization.
-%
-This way we can also compute the distribution in the galaxy of cetis
-that reach contact and the waiting time until the first contact or the
-waiting time until the next contact.
-%
-Regarding the properties of the population of CETIs and its evolution,
-we analyze the fraction of awaken time a ceti is listening at least
-another ceti, the age of contacted CETIs at first contact, the
-distribution of time to wait until next contact, the fraction of cetis
-where the first contact is given at the awakening, the distribution of
-the number of contacts for each CETI, the distribution of the maximum
-number of contacts for each CETI, the distribution of the number of
-contacts as a function of CETI age, the number of contacts as a
-function of time in the galaxy, the rate of CETIs that succeed in
-contact, and the distribution of distances between contacted CETIs.
-%
-Usefull derived quantity is the duration of two-way communication
-channels the or the fraction of contacts that admit a response.
-% 
-We also analyze the relations between the distance to ceti vs. time of
-double communication, the distance to ceti vs. age of contacted ceti,
-the age of a ceti and the maximum number of contacted cetis before D,
-or the lifespan of a ceti vs. max number of contacts.
-%
-All these quantities can be analyzed as a function of the simulation
-parameters.
-%
-We chose eight models that are on fairly oposed regions of the
-explored parameter space, and cover short/long lifetimes, short/long
-waiting times for the next CETI to appear in the Galaxy and
-short/large range reach of the signal, including all possible
-combinations.
-%
-The details of each model are summarized in the
-Table~\ref{T_selected_models}.
-
-In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
-distributions of the number contacts for CETIs in four different
-samples (M5, M6, M7, M8) with long range signal reach
-(D$_{max}$=10000). 
-%
-The lifetime is more determinant than the awakening rate in increasing
-the number of contacts.
-%
-As expected, of the four samples, the sample M7, which corresponds to
-a dense awakening in the timeline and a long lifetime, is the model
-that maximizes the number of contacts, reaching a maximum of order 100
-contacts for a single CETI.
-%
-This is considerable larger than any model with a shorter lifetime,
-which produce a number of contacts of at most the order of ten
-contacts per CETI.
-%
-A couple of comments are worth to mention about this plot.
-%
-First, it has a logarithmic scale on the x-axis, and second it is the
-cumulative, not differential, empirical distribution.
-
-In what follows, we focus on the analysis of the duration of contacts
-(Sec.~\ref{SS_multiplicity}), the waiting time for the first contact
-(Sec.~\ref{SS_waiting}) and the
-variations of contact posibilities as a function of the position of
-the CETI in the Galaxy (Sec.~\ref{SS_location}).
-
-%}}}
-
-\subsection{Multiplicity and duration of contacts}\label{SS_multiplicity}
-%{{{
-
-The duration of each contact for a single receiving CETI is given by the
-time difference between a C event and a B event, which are direct
-outputs for each simulation experiment.
-
-%}}}
-
-\subsection{Waiting time for a first contact}\label{SS_waiting}
-%{{{
-
-In this section we analyze the distribution of the waiting times for
-a first contact.
-%
-%Such distribution can be considered to compute the probability of
-%reaching a first contact within the first XX years of a SETI project,
-%assuming the methods are correct and the hypotheses of the experiment.
-Such distribution can be considered to compute the probability for a
-random CETI to carry out a SETI project and spend a given time until
-the first contact is made.
-
-In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the CETIs
-that have made at least one contact as a funtion of the elapsed time
-since the awakening, given that they make at least one contact on
-their lifetime, for different models.
-%
-From a frequentist approach, the normalized cumulative counts are
-related to the estimation of the probability of observing a given time
-with no success.
-%             
-Both panels show the histograms of the waiting times for the first
-contact, for all the CETIs that make at least one contact along their
-lifetime.
-%
-We show separately the histograms for several values of $\tau_a$ and a
-fixed value of $(\tau_s, D_{max})$, and for several values of $\tau_s$
-and a fixed value of $(\tau_a, D_{max})$.
-%
-For different values of $\tau_a$, the shape of the distribution is
-similar, and the main difference given by the different number of
-contacts on the different models.
-%
-Conversely, for different values of $\tau_a$, the shape of the
-distribution is similar, and the main difference given by the
-different number of contacts on the different models. 
-%
-We show the 2D maps of the rate of CETIs that never make contact and
-the rate of CETIs that make contact at the moment of the awakening,
-inf the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
-as a function of $\tau_s$ and $\tau_s$.
-%
-In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
-waiting times for the first contact, for several models.
-%
-Left panel shows the histograms for several values of $\tau_a$, and
-$\tau_s$ in the range 17000-24000 yr.
-%
-Right panel shows the histograms for several values of $\tau_s$, and
-$\tau_a$ in the range 68000-100000 yr.
-%
-The shape of the histograms does not change when $\tau_a$ is modified,
-but significantly changes when $\tau_s$ is modified.
-
-%}}}
-
-%   \begin{figure}
-%      %
-%      \centering
-%      %
-%      \includegraphics[width=0.45\textwidth]{galactocentric_frac.pdf}
-%      %
-%      \caption{Fraction of CETIs making cero, one, two or three contacts
-%      as a function of galactocentric distance.}
-%      %
-%      \label{F_radial_frac}
-%      %
-%   \end{figure}
-
-\subsection{Location within the Galaxy}\label{SS_location}
-%{{{
-
-%On the Figure~\ref{F_radial_frac} we show the fraction of CETIs that make
-%zero, one, two or three contacts, as a function of the galactocentric
-%distance, for model M1.
-% VERIFICAR!  M1?
-%
-For this model, the majority of CETIs do not
-make any contact during all the time period in which they are active.
-%
-As expected, the fractions of CETIs making at least one contact
-diminishes as the number of contacts rises.
-%
-The shape of this distribution is a result of the geometry of the GHZ.
-%
-In general, a closer position to the center of the Galaxy slightly favours 
-the posibility of a contact.
-%
-This would indicate that the position of the Earth is suitable for the
-odds of making a contact.
-
-%}}}
-
-\setlength{\tabcolsep}{10pt}
-\begin{table*}
-\centering
-\begin{tabular}{ccccc}
-\hline
-Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & description  \\
-\hline
-M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
-M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
-M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
-M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
-%
-M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
-M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
-M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
-M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
-%
-\hline
-\end{tabular}
-\caption{Selected models to analize the behavior of simulation outputs
-   and their dependence on simulation outputs and their dependence on
-   simulation parameters.}
-\label{T_selected_models}
-\end{table*}
+\input{results}
 

### 2019-12-19 methods.tex
reestructuracion

\section{Methods and working hypotheses}\label{S_methods}
%{{{

Simulations are suitable tools to analyze systems that evolve with
time and involve randomness.
%
An advantadge of simulations compared to theoretical approaches, is
that the former usually require less assumptions and simplifications,
and can be applied to systems where a theoretical model can not be
found.
%
In particular, many complex stochastic processes that can be described
by the evolution of the state of a system, can be efficiently modeled
with the discrete--event (hereafter, DE) simulation approach.
%
A system described with the DE paradigm is characterized by a set of
actors and events, where actors interact causally through a series of
events on a timeline and process these events in chronological order
\citep{ptolemaeus_system_2014, chung_simulation_2003,
ross_simulation_2012}.
%
Each event produces changes in the values of the variables that define
the process, and thus the corresponding change in the state of the
simulated system.
%
This method is well suited for the particular case of the difusion of
intelligent signals in the Galaxy, and allows to explore several
models easily.
%
We simulate the statistical properties of a set of points in space and
time that have the ability of emitting and communicating messages at
light speed, hereafter dubbed ''message processing locations'' (MPL).
%
These locations have the capability of emitting and receiving messages
in all directions.
%
It is worth mentioning that this is a general approach, and not
necessarily a MPL is the host of an intelligent civilization.
%
It can be associated with a set of planets where life has developed,
became intelligent, reached the skills required to find the right
communication channel, sustain a search and stablish a contact.
%
Alternative message processing entities could be considered, for
example interstellar beacons where life has ceased to exist but
continue with its emission or communication stations stablished by
probes or left intentionally by intelligent beings \citep[see, e.g.,
][]{peters_outer_2018, barlow_galactic_2013}.
%
In principle, these strategies could affect our results since it would
be easier to configure a cluster of MPLs that spread in time.
%
However, we do not consider this highly speculative alternatives at
this point.
%
% have been proposed
%
For the purpose of this analysis, only the communication capability is
relevant, since we study the causal contacts between the locations.
%
% communicating, civilization, intelligent, message, processing,
% handling, locations CALL, MPI, MPL, MPL
%
The system is defined by a number of actors that represent MPLs and
appear at different instants in time, generating events that produce
meaningful changes in the variables that describe the system, i.e., in
the arrangement of MPLs and their network of causal contacts.
%
For example, the appeareance of a new MPL in a region which is filled
with a signal emitted by another MPL, will increase the number of
active MPLs and the number of pairs of MPLs in causal contact.
%
Assuming some simple hypotheses, the discrete events method can be
performed taking into account a small number of variables, which allow
to analyze the variation of the results in the model parameter space.

In what follows, we describe the experimental setup chosen to estimate
the probabilities of causal contacts and several derived quantities in
terms of three independent parameters, namely, the mean time span
between the appeareance of consecutive MPLs, ($<\tau_a>$), their mean
lifetime ($<\tau_s>$), and the maximum distance a signal can be
detected by another MPL ($D_{max}$).
%
Intuitively, the shorter the $\tau_a$ parameter and the larger the
$\tau_s$ and $D_{max}$ parameters are, the greater the probability of
the existence of causal contacts between pairs of MPLs would be.
%
We also need to propose theoretical distribution functions for both
the distribution of lifespans ($\tau_s$) and the distribution of the
number of MPLs per unit time \citep{maccone_evolution_2014,
Sotos_biotechnology_2019}, related to the time span between the
appeareance of consecutive MPLs (since when $<\tau_a>$ is shorter, it
produces a greater density of MPLs).
%
The shapes of these distributions are set to a fixed law, as discussed
in Sec.~\ref{SS_PDF_shape}.

A causal connection can be produced without the need of the two
involved locations being concurrent or active at the same time.
%
This is a fundamental property of the system which arises as a
consecuence of the large spatial and temporal scales, where a message
is transmited at light speed.
%
Although a MPL could be active for enough time so as to transmit a
message at large distances, the limited power of the message and the
dilution that runs as the squared distance from the source, imposes a
detectability limit.
%
As a consecuence of this limitation and of their finite lifetime,
considered as the period of time between the aquisition and loss of
communicating capacity, each MPL will fill a spherical shell region of
the Galaxy, limited by two concentric spherical surfaces.
%
The leading front, or surface of first contact (SFC) grows from the
central MPL until it reaches the maximum distance $D_{max}$.
%
Thus, the volume reached causally by a MPL is initially a growing
sphere.
%
Following the end of the civilization, there is still a region which
is filled with the emited signals.
%
The trailing front, or surface of last contact (SLC) also grows from
the central MPL, with a delay with respect to the SFC equivalent to
the lifetime of the MPL, and produces a spherical shell region.
%
Any other MPL within this region will be aware of the originating MPL,
even if it has disapeared before the time of contact.
%
This region will grow if the surface of first contact has not yet
reached the maximum distance $D_{max}$, and will shrink otherwise
until the surface of last contact reaches $D_{max}$, producing as a
result the loss of all signal from the central MPL.
%
In our approach, we consider a model Galaxy where the width of the
disk is negligigle with respect to the radius of the disk, so that the
MPLs are placed on a 2D plane disk.
%
In the 2D simulation only the intersection of the communicating
spherical shells with the plane of the galaxy is relevant, and produce
the corresponding circles or rings for the filled spheres or annular
regions of the spherical shells, respectively.
%
We illustrate in the Fig.~\ref{F_scheme} the schematic representation
of the region causally connected to the central MPL.
%
To this end, we use space--time diagrams, where time is represented on
the horizontal axis, and space is represented in the vertical axis.
%
A growing surface of first contact in this diagram is represented as a
growing circle, centered on a point in the time axis, which shifts
towards the right at the same rate the radius of the circle increases.
%
The maximum distance of the message limits the radius of the circles,
so that once this distance is reached, the circle continues its shift
towards the right, but maintaining a constant radius.
%
At the end of the lifetime of a MPL, a growing surface of last
contact is represented by a circle that covers the inner part of the
causal region, producing a ring of diminishig area (panel (a)).
%
In this scheme, the entire lifetime of a MPL can then be represented
by a polygon, limited in the spatial axis by the maximum distance of
the signal $D_{max}$ and in the time axis by the time span between the
starting time (concave vertex) and the ending time (convex vertex).
%
This representation allows to easily visualize the important events
that result from the existence and communications of two MPLs (panel
(b)).
%
The points in time where the MPLs aquire their communicating capacity
(CC) are dubbed ''awakening'' events ($A_1$ and $A_2$ in the Figure).
%
Simmilarly, the points in time where the MPLs loss their
communicating capacity are dubbed ''doomsday'' events ($D_1$ and $D_2$
in the Figure).
%
The points in space--time where the first contact is produced for each
one of the MPLs, are defined as ''contact'' events, shown as $C_1$
and $C_2$ in the Figure.
%
Finally, the two points where the contact is lost for each one the
MPLs, are denominated ''blackout'' events, shown as $B_1$ and $B_2$
in the Figure.
%
It is important to notice that there is a time delay between the
contact events of the two MPLs involved in this analysis, and also a
time delay between the two blackout events, so that the time range
when a bidirectional contact is possible occurs between the maximum
time of the contact points and the minimum time between the blackout
points.

The initially growing communicating sphere, is shown over
space--time diagrams, where time is represented on the horizontal
direction, and space is represented in the vertical direction.
%
In the upper panel (a), the sphere is growing as the surface of first
contact has not reached the maximum distance.
%
In the middle panel (b) it has reached the maximum distance, so that
it remains at the same size.
%
After a Doomsday event (bottom panel, b), the signals can still be
observed, but the surface of last contact grows.
%
We also show in this Figure a schematic representation of two
emitters, E$_1$ and E$_2$, that reach each other at different times.
%
The time span for E$_i$ is (A$_i$ , D$_i$) , for $i = {1, 2}$.
%
Emitter $i$ can listen to emitter $j$ between C$_{ij}$ and B$_{ij}$.
%
The type and length of causal contact in both directions depend on the
distance and time lag between the awakening events, the maximum
distance that a signal can reach and the time period in which each
emitter is active.
%
In the Fig.~\ref{F_messages} we show a schematic representation of the
possible cases in which a MPL A can be in causal contact with a MPL
B which appears later.
%
The duration of the causal contact in one direction depends on several
factors, mainly $\Delta$t~A , $\Delta$t~C , and $\Delta$X.

\begin{figure}
   \centering
   \includegraphics[width=0.5\textwidth]{scheme.pdf}
   \usecaption{F_scheme} \label{F_scheme}
\end{figure}

The temporal structure of the process is defined by two distribution
parameters that represent the mean time interval that an intelligent
civilization can emit and receive signals, and the mean time interval
between the emergence of an intelligent comunicating civilization and
the next one.
%
The spatial structure of the simulation is given by the size and shape
of the Galactic Habitable Zone and the maximum distance a signal can
travel to be detected.
%
The parameters for the temporal distributions also determine the
spatial properties since the density of active cetis in the Galaxy
depend on these two parameters.
%
For example, a small $\tau_a$ and a large $\tau_s$ will produce a
Galaxy densely populated with MPLs, and viceversa.
%
Also, some hypotheses must be made in order to complete the
simulation.
%
\citet{forgan_spatiotemporal_2011} argues that the times at which
different civilizations become intelligent 
follow a Gaussian distribution, and then the distribution of
inter-arrival times is an inverse exponential.
%
We chose to assume that the distribution of the times of A events is a
stationary Poisson process, and then the distribution of the times between the
appeareance of new consecutive MPLs is exponential.
%
Regarding the duration of a MPL, we propose that the distribution of
the duration of MPLs is a stationary exponential distribution.
%
\citet{balbi_impact_2018} stress the fact that it would be desirable to arrive
at a theoretical statistical distribution of the lifespan of
civilizations, preferentially on the basis of the underlying
astrophysical and biological processes.
%
\citet{maccone_lognormals_2014} argues that this distribution should be a log normal.
(EXPAND)

There is no a temporal window for the model, i.e., the simulation
starts when the process is already stable, and ends before any
galactic effect could modify or affect of fixed variables.
%
The probability for the rise of a MPL is homogeneous over the GHZ.
%
Although the Galaxy has a well known spiral structure, the assumed
sparcity of civilizations makes this assumption very reasonably.
%
Otherwise, if the distribution of civilizations is not sparce, it
could be the case that the spiral arms would host most of the
civilizations, and contacts are frecuent between closely located
civilizations.
%
In such a case we could expect that the interpretation of our results
could be considered as pesimist for the stablishment of a contact with
our planet.
%
We also limit the posibilities of life or other types of civilizations
to the hypothesis usually stated for the definition of the GHZ
\citep{dayal_habitability_2016,gonzalez_galactic_2001,lineweaver_galactic_2004,gonzalez_habitable_2005,morrison_extending_2015,haqq-misra_evolution_2019,rahvar_cosmic_2016,gobat_evolution_2016,
rahvar_cosmic_2016}.
%
\citet{vukotic_astrobiological_2012} propose probabilistic cellular
automata model to explain the astrobiological history of the Milky
Way.
%
This means that we set aside possible civilizations that could survive
in severe conditions or unstable systems, which would prevent the
appeareance of life as we know it.

\setlength{\tabcolsep}{10pt}
\begin{table*}
\centering
\begin{tabular}{clccc}
\hline
   \multicolumn{2}{l}{independent variables} &min&max&Nbins\\
\hline
   $\tau_{a}$ & Mean temporal separation between consecutive awakenings & 4000 y & 200000 y & 50\\ 
   $\tau_{s}$ & Mean lifetime of a MPL                                 & 10000 y & 500000 y& 50\\ 
   D$_{max}$ & Maximum reach of a message  &  \multicolumn{3}{l}{500, 1000, 10000, 20000, 40000, 80000 lyr}  \\
\hline
   \multicolumn{4}{l}{fixed variables and assumptions} &value \\
\hline
   & statistical properties of all MPLs &equally distributed&&\\
   & Point process for the distribution in time & homogeneous &&\\
   f$_s$ & The scan of the sky & fully efficient&&1\\
   f$_p$ & panspermia or colonization &absent&&0\\
   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
   R$_{SFC}$   & Inner radius of the GHZ     &&&20000 ly\\
   R$_{SLC}$   & Outer radius of the GHZ     &&&60000 ly\\
   t$_{max}$ & Time span of the simulation  &&&1.e6 y\\
\hline
   \multicolumn{5}{l}{discrete events}\\
\hline
   A & Awakening: a civilization starts a SETI program  &&&\\
   B & Blackout: the end of the communication chanel stops &&&\\
   C & Contact: a new causal contact is produced &&&\\
   D & Doomsday: a civilization ends its communication capabilities&&&\\
\hline

\hline
\end{tabular}
\caption{Definition of independent variables and adopted values for 
   fixed paramenters 
   that are part of the simulation.  Variable parameters define the
   spatial and temporal structure
   of the process and the maximum reach of the messages.}
\label{T_simu_hypotheses}
\end{table*}

While the aforementioned assumptions are basic conditions for most of
the stochastic processes observed in nature, we need to make stronger
assumptions related to the nature of the message.
%
The most simple assumption about the message itself, is that it
travels at light speed.
%
With this, we are considering messages sent through, for example,
electromagnetic radiation or even gravitational radiation, but we set
appart messages sent with mechanical means or physical objects, or
through some unknown technology that violates the known so far laws of
physics.
% 
For the communication of MPLs through messages sent isotropically, we
assume that the capacity to emit signals and to receive signals ocurr
at the same time.
%
This means that the ability to find another civilization develops at
the same time that the ability and intention to emit a signal in all
direction, intented to be detected by an unknown civilization.
%
Although there are several reasons to think that this could not be the
case, at large time scales it can be considered that both abilities
occurr roughly at coincident epochs.
%
Another essential assumption is that all MPLs use the same signal
power, so that there is a maximum distance out to which it can be
detected.
%
It is worth mentioning that we are considering in our experimental
setup a system composed by several emmiters and receivers across the
Galaxy, under the hypothesis proposed before.
%
In such a system we compute probabilities of a MPL making contact
with another MPL, and the results must be interpreted as for a
randomly chosen MPL, given certain conditions of homogeneityon an
ensemble of equivalent galaxies.
%
We stress the fact that we are not centered on the probabilities for
the Earth itselft, but on a generalized MPL.
%
Regarding the extent of the signals, we know that the distance from
which a signal from Earth could be detected using the current
technology is of a few hundreds light years, given that the signal was
sent to a specific direction.

\begin{figure*}
   \centering
   \includegraphics[width=0.8\textwidth]{Messages_01.pdf}
   \usecaption{F_messages} \label{F_messages}
\end{figure*}

%https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
%https://www.engadget.com/2017/05/25/listening-to-starlight-our-ongoing-search-for-alien-intelligenc/]
%[from https://en.wikipedia.org/wiki/Fermi_paradox]
%
It is straightforward to propose and implement a distribution of
maximum distances, although this would increase the model complexity
at the cost of a larger uncertainty or of another dimension on the
hypothesis space.
%
This proposal is based on a pronoid scenario (REF), and we do not
consider in this work the ocurrence of paranoid or partially paranoid
civilizations.
%
If that would be the case, the results obtained here can be taken as
upper limits.

In our simulations we also assume the simplest situations for the
growth of ETIs.
%
For example, we discard the possibility of stellar colonization
\citep[e.g.][]{newman_galactic_1981, walters_interstellar_1980,
starling_virulence_2013, barlow_galactic_2012, jeong_large_2000, maccone_mathematical_2011}
and assume that the communication aim is performed to
all directions with the same power and the same probability, i.e.,
strongly assume isotropic communication in all cases.
%
More detailed simulations could be produced considering different 
efficiency of communication or detection methods.
%
Also, a mixed paranoid/pronoid population of intelligent civilizations
could be implemented easily.
%
However, in this case we also would complicate the experimental setup,
make the results less evident.
%
The results are independent of the nature of life (organic or
artificial).
%
The lifetime of a civilization can be caused by autodestruction or by
external factors.
%}}}

\subsection{Power laws vs. exponential laws}\label{SS_PDF_shape}
%{{{

The exponential distribution of lifespan and waiting times is
justified by considering that the process of appeareance of life in
the galaxy is homogeneous and stationary.
%
This means that there is no preferred location within the GHZ for the
spontaneos appeareance of life, and that the emergence of a
civilization is independent of the existence of previous civilizations 
in the galaxy.
%            
This is equivalent to proposing a Poisson process for the emergence of
MPLs, since there is a close relation between the number of events in
time or space and the waiting time or separation, respectively.
%
That is, these are two alternative approaches to describing the same
process, a Poisson distribution for the number of events implies an
exponential distribution for their sepations, and viceversa.
%
It should be emphasized that the exponential laws used in this work
are assumed, although instead of analyzing results from a particular
parameter chosen ad--hoc, we explore the hypothesis space and analyze
the impact of the values of these parameters on the results.
%

The power law and exponential statistical distributions are among the
most common patterns found in natural phenomena.
%
For example, the distribution of the frecuency of words in many
languajes is known to follow a Zipf law (which is a power law).
%
Zipf law also describes population ranks of cities in various
countries, corporation sizes, income rankings, ranks of number of
people watching the same TV channel (Zipf's law entry on Wikipedia).
%
The magnitudes of earthquakes, hurricanes, volcanic eruptions and
floods; the sizes of meteorites or the losses caused by business
interruptions from accidents, are also well described by power laws
(Sornette 2006).
%
The power law behaviour has been observed in a variety of systems,
including for example stock market fluctuations, sizes of computer
files and word frequency in languages \citep{mitzenmacher_brief_2004,
newman_power_2005, simkin_theory_2006}
%
Power laws have also been widely used in biological sciences, e.g., in
analysing connectivity patterns in metabolic networks
\citep{jeong_large_2000} and in the number of species observed per
unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
% 
More examples can be found in the literature
\citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009,
maccone_evolution_2014, maccone_lognormals_2014,
benguigui_classificacion_2016}

%}}}

\subsection{Discrete event process}
%{{{

Given a parametric model $M(\tau_a, \tau_s, D_{max})$, which includes
the functional forms of the statistical distributions and fixed values
for the hypotheses, a discrete event simulation is performed by
keeping track of a set of variables that change each time a meaningful
event is produced.
%
The main variables that follow the evolution of the simulation are the
positions of stars, which are sampled randomly within the GHZ, the
time of the awakening of each MPL (A event), and the time of the
vanishing of each MPL (D event).
%
The variables that can be deduced from the previous ones include the
number of MPLs in casual contact with at least another MPL at a given
time, the number of MPLs as a function of time, the number of MPLs
that receive a message at least one time, the number of MPLs that
receive a message at least one time and successfully deliver an answer
and the number distribution of waiting time to receive a message.
%
All these quantities are updated each time one of the four events (A,
B, C, D) is produced.

%%% comparar con el stochsatic cellular automata
% ver libro Solving Fermi paradox, sec. 27.3.1

\begin{figure} \centering
   \includegraphics[width=0.5\textwidth]{F1.pdf}
   \usecaption{F_number_of_contacts} \label{F_number_of_contacts}
\end{figure}

%}}}

\subsection{Model for MPLs}
%{{{

Since the method is based on events, the first step is to define an
architecture of events, the relationships between them and how events
trigger changes on the state of the system.
%
We consider a simulated system that represents the galactic habitable
zone (GHZ).
%
On a first approach, the GHZ is a 2-dimensional annular region.
%
The adopted values for the GHZ are 20 kly and 60 kly, for the inner
and outer radius, respectively \citep{lineweaver_galactic_2004}.
% ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
% mas chico...
%
This simple model does not take into accont the variations in stellar
density given by the spiral structure.
%
Although there are several possible approaches, we chose to follow the
evolution of the system according to the following events:

\begin{enumerate}
   \item[(A)] A new MPL appears (Awakening)
   \item[(D)] An old MPL dissappears (Doomsday)
   \item[(C)] A new causal contact is stablished (Contact)
   \item[(B)] An existing causal contact is interrupted (Blackout)
\end{enumerate}

The event of type B is produced when one of the two MPLs halts in its
capability of emmiting and receiving signals.

\begin{figure*} % 2D color plot
   \centering
   \includegraphics[width=\textwidth]{matrix_1.pdf}
   \usecaption{F_never_contact}
   \label{F_never_contact}
\end{figure*}

\begin{figure*} % 2D color plot
   \centering
   \includegraphics[width=\textwidth]{matrix_2.pdf}
   \usecaption{F_C_at_A}
   \label{F_C_at_A}
\end{figure*}

The system is updated each time an event is produced.
%
The time marks for each event (A and D for each MPL and C and B for
each causal contact) are stored as a result of the simulation.
%
Also, the list of active MPLs is obtained as a function of time.
%
Some fixed variables must be set in order to carry out the simulation:
the size of the Galactic Habitable Zone, the mean lifetime of a MPL,
$\tau_s$, the mean waiting time for the appeareance of another MPL,
$\tau_a$, and the maximum distance a message can be detected.

The Galactic Habitable Zone is the only one of the four parameters
that is barely known, is set at fixed values of the inner and outer
radii, radial symetry is assumed, and the with of the galactic disk is
neglected compared to the radial size.
%
The two radii are measured in light years, with the aim to maintain a
single and comprehensive unit for both space and time coordinates.

The simulation was implemented on a python 3 code, which is publically
available\footnote{https://github.com/mlares/simu\_contact}

%}}}

\begin{figure}
   \centering
   \includegraphics[width=0.5\textwidth]{waiting_s+a_dif_ylog.pdf}
   \usecaption{F_waiting_for_1C}
   \label{F_waiting_for_1C}
\end{figure}

### 2019-12-19 results.tex
reestructuracion

\section{Results: exploring the parameter space}\label{S_results}
%{{{

We implemented the simulation of a regular grid of models varying over
the hypothesis space, which covers 15000 different models.
%
For each model, we simulated 50 realizations with different random
seeds, in order to estimate the variance.
%
The simulated parameters, $\tau_a$, $\tau_s$ cover the ranges
4000-200000 and 10000-500000, respectively, with a partition of 50
values for each parameter.
%
For the D$_{max}$ parameter, we chose the values 500, 1000, 10000,
20000, 40000 and 80000 lyr.
%
This setup makes a total of 750000 simulations, each covering a time
range of one million years.
%
In Table \ref{T_simu_hypotheses} we show the range of values of these parameters explored 
in our numerical experiments.

As a product of the simulations, several quantities are obtained.
%
Some quantities are directly derived from the discrete events, namely,
the ID of emitting MPL, the ID of emitting (receiving) MPL, the
position in the galaxy, the time of Awakening (Contact) and the time
of Doomsday  (Blackout) for each MPL.
%
This structure is always for the A-D cycle of each emitting MPL, and
repeats for each contact (C-B).
%
We can also derive the additional quantities.
%
The age of a MPL is the time elapsed from A to D.
%
%
%given time, ceti\_e is the MPL emitting signals (emiter), ceti\_r is
%the MPL receiving signals (receiver), ceti\_c is the ceti that
%listen at least another ceti (citizen) and ceti\_h is the MPL ceti
%that is lestened by at least another MPL.
%
The time span of a MPL listening another or being listened by another
can also be easily derived from the results of a realization.
%
This way we can also compute the distribution in the galaxy of cetis
that reach contact and the waiting time until the first contact or the
waiting time until the next contact.
%
Regarding the properties of the population of MPLs and its evolution,
we analyze the fraction of awaken time a ceti is listening at least
another ceti, the age of contacted MPLs at first contact, the
distribution of time to wait until next contact, the fraction of cetis
where the first contact is given at the awakening, the distribution of
the number of contacts for each MPL, the distribution of the maximum
number of contacts for each MPL, the distribution of the number of
contacts as a function of MPL age, the number of contacts as a
function of time in the galaxy, the rate of MPLs that succeed in
contact, and the distribution of distances between contacted MPLs.
%
Usefull derived quantity is the duration of two-way communication
channels the or the fraction of contacts that admit a response.
% 
We also analyze the relations between the distance to ceti vs. time of
double communication, the distance to ceti vs. age of contacted ceti,
the age of a ceti and the maximum number of contacted cetis before D,
or the lifespan of a ceti vs. max number of contacts.
%
All these quantities can be analyzed as a function of the simulation
parameters.
%
We chose eight models that are on fairly oposed regions of the
explored parameter space, and cover short/long lifetimes, short/long
waiting times for the next MPL to appear in the Galaxy and
short/large range reach of the signal, including all possible
combinations.
%
The details of each model are summarized in the
Table~\ref{T_selected_models}.

In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
distributions of the number contacts for MPLs in four different
samples (M5, M6, M7, M8) with long range signal reach
(D$_{max}$=10000). 
%
The lifetime is more determinant than the awakening rate in increasing
the number of contacts.
%
As expected, of the four samples, the sample M7, which corresponds to
a dense awakening in the timeline and a long lifetime, is the model
that maximizes the number of contacts, reaching a maximum of order 100
contacts for a single MPL.
%
This is considerable larger than any model with a shorter lifetime,
which produce a number of contacts of at most the order of ten
contacts per MPL.
%
A couple of comments are worth to mention about this plot.
%
First, it has a logarithmic scale on the x-axis, and second it is the
cumulative, not differential, empirical distribution.

In what follows, we focus on the analysis of the duration of contacts
(Sec.~\ref{SS_multiplicity}), the waiting time for the first contact
(Sec.~\ref{SS_waiting}) and the
variations of contact posibilities as a function of the position of
the MPL in the Galaxy (Sec.~\ref{SS_location}).

%}}}

\subsection{Multiplicity and duration of contacts}\label{SS_multiplicity}
%{{{

The duration of each contact for a single receiving MPL is given by the
time difference between a C event and a B event, which are direct
outputs for each simulation experiment.

%}}}

\subsection{Waiting time for a first contact}\label{SS_waiting}
%{{{

In this section we analyze the distribution of the waiting times for
a first contact.
%
%Such distribution can be considered to compute the probability of
%reaching a first contact within the first XX years of a SETI project,
%assuming the methods are correct and the hypotheses of the experiment.
Such distribution can be considered to compute the probability for a
random MPL to carry out a SETI project and spend a given time until
the first contact is made.

In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the MPLs
that have made at least one contact as a funtion of the elapsed time
since the awakening, given that they make at least one contact on
their lifetime, for different models.
%
From a frequentist approach, the normalized cumulative counts are
related to the estimation of the probability of observing a given time
with no success.
%             
Both panels show the histograms of the waiting times for the first
contact, for all the MPLs that make at least one contact along their
lifetime.
%
We show separately the histograms for several values of $\tau_a$ and a
fixed value of $(\tau_s, D_{max})$, and for several values of $\tau_s$
and a fixed value of $(\tau_a, D_{max})$.
%
For different values of $\tau_a$, the shape of the distribution is
similar, and the main difference given by the different number of
contacts on the different models.
%
Conversely, for different values of $\tau_a$, the shape of the
distribution is similar, and the main difference given by the
different number of contacts on the different models. 
%
We show the 2D maps of the rate of MPLs that never make contact and
the rate of MPLs that make contact at the moment of the awakening,
inf the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
as a function of $\tau_s$ and $\tau_s$.
%
In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
waiting times for the first contact, for several models.
%
Left panel shows the histograms for several values of $\tau_a$, and
$\tau_s$ in the range 17000-24000 yr.
%
Right panel shows the histograms for several values of $\tau_s$, and
$\tau_a$ in the range 68000-100000 yr.
%
The shape of the histograms does not change when $\tau_a$ is modified,
but significantly changes when $\tau_s$ is modified.

%}}}

\subsection{Location within the Galaxy}\label{SS_location}
%{{{

%On the Figure~\ref{F_radial_frac} we show the fraction of MPLs that
%make zero, one, two or three contacts, as a function of the
%galactocentric distance, for model M1.  VERIFICAR!  M1?
%
For this model, the majority of MPLs do not make any contact during
all the time period in which they are active.
%
As expected, the fractions of MPLs making at least one contact
diminishes as the number of contacts rises.
%
The shape of this distribution is a result of the geometry of the GHZ.
%
In general, a closer position to the center of the Galaxy slightly
favours the posibility of a contact.
%
This would indicate that the position of the Earth is suitable for the
odds of making a contact.

%}}}

\setlength{\tabcolsep}{10pt}
\begin{table*}
\centering
\begin{tabular}{ccccc}
\hline
Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & description  \\
\hline
M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
%
M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
%
\hline
\end{tabular}
\caption{Selected models to analize the behavior of simulation outputs
   and their dependence on simulation outputs and their dependence on
   simulation parameters.}
\label{T_selected_models}
\end{table*}

### 2019-12-20 methods.tex
autores

@@
+\begin{figure}
+   \centering
+   \includegraphics[width=0.5\textwidth]{scheme.pdf}
+   \usecaption{F_scheme} \label{F_scheme}
+\end{figure}
+
+\begin{figure*}
+   \centering
+   \includegraphics[width=0.8\textwidth]{Messages_01.pdf}
+   \usecaption{F_messages} \label{F_messages}
+\end{figure*}
+
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
 \section{Methods and working hypotheses}\label{S_methods}
@@
 
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{scheme.pdf}
-   \usecaption{F_scheme} \label{F_scheme}
-\end{figure}
-
 The temporal structure of the process is defined by two distribution
@@
 \citet{forgan_spatiotemporal_2011} argues that the times at which
-different civilizations become intelligent 
-follow a Gaussian distribution, and then the distribution of
-inter-arrival times is an inverse exponential.
+different civilizations become intelligent follow a Gaussian
+distribution, and then the distribution of inter-arrival times is an
+inverse exponential.
 %
 We chose to assume that the distribution of the times of A events is a
-stationary Poisson process, and then the distribution of the times between the
-appeareance of new consecutive MPLs is exponential.
+stationary Poisson process, and then the distribution of the times
+between the appeareance of new consecutive MPLs is exponential.
 %
@@
 %
-\citet{balbi_impact_2018} stress the fact that it would be desirable to arrive
-at a theoretical statistical distribution of the lifespan of
+\citet{balbi_impact_2018} stress the fact that it would be desirable
+to arrive at a theoretical statistical distribution of the lifespan of
 civilizations, preferentially on the basis of the underlying
@@
 %
-\citet{maccone_lognormals_2014} argues that this distribution should be a log normal.
-(EXPAND)
+\citet{maccone_lognormals_2014} argues that this distribution should
+be a log normal.  (EXPAND)
 
@@
 to the hypothesis usually stated for the definition of the GHZ
-\citep{dayal_habitability_2016,gonzalez_galactic_2001,lineweaver_galactic_2004,gonzalez_habitable_2005,morrison_extending_2015,haqq-misra_evolution_2019,rahvar_cosmic_2016,gobat_evolution_2016,
-rahvar_cosmic_2016}.
+\citep{dayal_habitability_2016, gonzalez_galactic_2001,
+lineweaver_galactic_2004, gonzalez_habitable_2005,
+morrison_extending_2015, haqq-misra_evolution_2019,
+rahvar_cosmic_2016, gobat_evolution_2016, rahvar_cosmic_2016}.
 %
@@
 
-\begin{figure*}
-   \centering
-   \includegraphics[width=0.8\textwidth]{Messages_01.pdf}
-   \usecaption{F_messages} \label{F_messages}
-\end{figure*}
-
 %https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
@@
 \citep[e.g.][]{newman_galactic_1981, walters_interstellar_1980,
-starling_virulence_2013, barlow_galactic_2012, jeong_large_2000, maccone_mathematical_2011}
-and assume that the communication aim is performed to
-all directions with the same power and the same probability, i.e.,
-strongly assume isotropic communication in all cases.
+starling_virulence_2013, barlow_galactic_2012, jeong_large_2000,
+maccone_mathematical_2011} and assume that the communication aim is
+performed to all directions with the same power and the same
+probability, i.e., strongly assume isotropic communication in all
+cases.
 %
@@
 the impact of the values of these parameters on the results.
-%
 
@@
 
+%}}}
+
 \begin{figure} \centering
@@
 \end{figure}
-
-%}}}
 
@@
 
-\begin{figure*} % 2D color plot
-   \centering
-   \includegraphics[width=\textwidth]{matrix_1.pdf}
-   \usecaption{F_never_contact}
-   \label{F_never_contact}
-\end{figure*}
-
-\begin{figure*} % 2D color plot
-   \centering
-   \includegraphics[width=\textwidth]{matrix_2.pdf}
-   \usecaption{F_C_at_A}
-   \label{F_C_at_A}
-\end{figure*}
-
 The system is updated each time an event is produced.
@@
 %}}}
-
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{waiting_s+a_dif_ylog.pdf}
-   \usecaption{F_waiting_for_1C}
-   \label{F_waiting_for_1C}
-\end{figure}

### 2019-12-22 manuscript.tex
autores
con cambios propuestos por Jose

@@
 
-\author[Lares \& Funes]{Lares M$^{1, 2}$, Funes J$^{1, 3}$}
+\author[Lares, Funes \& Gramajo]{Lares M$^{1, 2}$, Funes J.~G.$^{1, 3}$ \& Gramajo
+L$^{1, 2}$}
 
@@
 \address{
-   %
    \add{1}{CONICET, Argentina}
@@
    \add{3}{Universidad Cat\'olica de C\'ordoba, Argentina}
-   %
 }
+
+\selfcitation{Lares M, Funes J (XXXX).  Estimating the probability of
+causal contact between civilizations through MonteCarlo simulations,
+International Journal of Astrobiology}
+
+\keywords{SETI, Computer simulations, Statistics}
+
+\JELclassification{XX; XX}
+\MSCcodes{XX; XX}
+
+\Abbreviations{MPL: Communicating Extraterrestrial Intelligence, DE:
+Discrete Event, GHZ: Galactic Habitable Zone}
 
@@
 %
-A different approach uses detailed numerical simulations of the galaxy
+A alternative approach uses detailed numerical simulations of the galaxy
 and recipes for the rates of the formation of stars and planets, and
@@
 %
-The probability of causal contacts increases with the increase of the 
+The probability of causal contacts increases with the 
 mean lifetime of civilizations much more signifficantly than with
-the increase in the mean number of active civilizations for a given
+the mean number of active civilizations for a given
 time window.
@@
 
-\selfcitation{Lares M, Funes J (XXXX).  Estimating the probability of
-causal contact between civilizations through MonteCarlo simulations,
-International Journal of Astrobiology}
-
-\keywords{SETI, Computer simulations, Statistics}
-
-\JELclassification{XX; XX}
-\MSCcodes{XX; XX}
-
-\Abbreviations{MPL: Communicating Extraterrestrial Intelligence, DE:
-Discrete Event, GHZ: Galactic Habitable Zone}
-
 \maketitle
 
-\section{Motivations}\label{S_motivations}
-%{{{
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
+\input{motivations}
 
-A common approach to discussing the estimations of the number of
-extraterrestrial intelligent civilizations (ETI) in the Galaxy has
-been the use of the Drake equation \citep{gleiser_drake_2010,
-prantzos_joint_2013, haqq-misra_drake_2017}.
-%
-The abundance of intelligent civilizations in the galaxy is a
-longstanding question, which is often conceptualized as the problem of
-the lack of received communication or the Fermi paradox
-\citep{barlow_galactic_2012, Sotos_biotechnology_2019,
-forgan_galactic_2016}.
-%
-There are many proposals aimed at solving this paradox, which make use
-of statistical \citep{solomonides_probabilistic_2016,
-vanhouten_isthere_2017, horvat_calculating_2007,
-maccone_statistical_2015} or stochastic approaches
-\citep{forgan_numerical_2009, bloetscher_using_2019,
-glade_stochastic_2011, forgan_numerical_2010}, analytical
-interpretations of the Drake equation \citep{prantzos_joint_2013,
-smith_broadcasting_2009} or more speculative proposals
-\citep{barlow_galactic_2013, lampton_information_2013,
-conway_three_2018, forgan_galactic_2017}.
-%
-The uncertainties in the factors of the Drake equation make it less
-prone to a formal application in order to define searching strategies
-or to compute the actual number of ETIs, but is a key tool to organize
-the discussion around the problem of the population of ETIs in the
-Galaxy \citep{hinkel_interdisciplinary_2019}, which is noticeable
-dominated by the fact that despite decades of effort not a single ETI,
-if any, has been detected so far.
-%
-Some modifications to the original idea of the equation have tried to
-imprint a stochastic nature, or to propose a probabilistic approach,
-or to consider the temporal structure which is missing in the
-equation.
-%
-
-\Fpagebreak
-
-Temporal aspects of the distribution of communicating civilizations
-and their contacts have also been explored by several authors
-\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
-balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011},
-%The Impact of the Temporal Distribution of Communicating Civilizations on Their Detectability
-%Temporal dispersion of the emergence of intelligence: an inter-arrival time analysis
-%The spatiotemporal aspects of SETI
-%
-as well as efforts on considering the stochastic nature of the Drake equation
-\citep{glade_stochastic_2011}.
-%
-The simulation approach has also benn considered
-\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
-murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017},
-although with a similar problem that consists on the large number of
-parameters which are either unknown or largely uncertain.
-%
-%New numerical determination of habitability in the Galaxy: the SETI connection
-%
-%Since the factors in the Drake equation are uncertain,
-%
-In this work, we propose to
-avoid the frequentist approach of the Drake equation and to explore a
-parameter space, where instead of computing a final number, we provide
-a statistical distribution that gives conditional probabilities.
-%
-This is an exploratory analysis that aims at providing a numerical 
-tool to discuss not the several theoretical problems summarized by the
-Drake equation factors, 
-but the different scenarios on the basis of statistical heuristics.
-%
-The approach proposed here shold be considered as a compromise between
-the uncertainties of the frequentist approach and the detailed recipes
-required on the simulation approach.
-%
-We then provide a numerical framework to explore via simulations a
-parameter space of unknown observables in order to discuss different
-scenarios and their consecuences in terms of the probability con
-making contacts.
-%
-The only fact that can be stated with certainty is that for the number
-of years SETI projects have been working we have not received any
-signal, at least
-within the technical posibilities and the 
-conditions stablished by SETI \citep{tarter_search_2001}.
-%
-This arguably astonishing result has lead to the Fermi paradox, which
-states an apparent contradiction between the high estimation of the
-abundance of life in the Galaxy and the lack of their detection
-\citep{vanhouten_isthere_2017}.
-%
-The lack of detection has also ispired to a number of alternative proposals
-for new SETI strategies \citep{forgan_exoplanet_2017,
-balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010,
-tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016,
-maccone_SETI_2011,  lingam_relative_2019, wright_theGsearch_2015,
-maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018,
-forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
-
-The discrete events method for simulating a stochastic process is an
-approximation that allows to study the behaviour of complex
-systems, by considering a sequence of well defined discrete events.
-%
-The simulation os carried out by following all the variables that
-describe the system, that constitute the state of the system.
-%
-The evolution of the process, then, is described as a set of changes
-in the state of the system.
-%
-In this context, an event produces a specific change in the state,
-that can be triggered by random variables that encode the stochastic
-nature of the physical phenomenon.
-%
-The process involves following the changes on the state of the system,
-definig the initial and final states, defining a method that allows to
-keep track of time progress, and maintaining a list of relevant
-events.
-
-In a recent work, \citep{balbi_impact_2018} use a statistical model to analyze
-the occurrence of causal contacts between civilizations in the Galaxy.
-%
-The author highlights the effect of evolutionary processes when
-attempting to estimate the number of communicating civilizations that
-might be in causal contact with an observer located on the Earth.
-%
-\citet{cirkovic_temporal_2004} also emphasize the lack of temporal
-structure in the Drake equation.
-
-%While the Drake equation remains a very useful
-%guide in identifying and analyzing the various factors in-
-%volved in the problem and their relative probabilities, our
-%approach could provide a better framework with which to
-%perform a statistical analysis when taking into account evo-
-%lutionary processes. In particular, we have shown that the
-%Drake equation can misestimate the number of detectable
-%communicating species if spatial and temporal dependencies
-%are important. This should suggest some caution in the un-
-%critical application of such an equation. In this paper, we offer
-%an exemplificative illustration of the effect of different
-%probability distributions for the typical timescales involved
-%in the problem. Ideally, it would be desirable to arrive at a
-%mathematical description of such distributions based on a
-%modeling of the underlying astrophysical and biological
-%processes. This points out a possible future direction of in-
-%vestigation in which an integrated evolutionary (and coevo-
-%lutionary) approach to the problem would be a central focus
-%of interest (for more on this perspective, see, e.g., Chaisson
-
-\citet{bloetscher_using_2019} uses a probabilistic approach, though
-still strongly motivated by the Drake equation, to obtain a
-propbabilistic measure for the number of civilizations in the Galaxy.
-To that end, perform MonteCarlo Markov Chains over each factor of the
-Drake equation, and combine the means to obtain a probabilistic
-result. It is worth mentioning that the author proposes a lof-normal
-target distribution to compute the posterior probabilities. This
-approach, which assigns probability distributions to the Drake
-factors. This study concludes that there is a very low probability for
-Earth to make contact with another extraterrestrail civilization.
-
-\citet{smith_broadcasting_2009} uses a analytical model to compute the
-probabilities of contact between two randomly located civilizations,
-and the waiting time for the first contact, assuming a fixed maximum
-broadcasting distance.
-
-\citet{balbi_impact_2018} investigate the chance of communicating
-civilizations making causal contact within a volume surrounding the
-location of the Earth, using a statistical model to compute the
-occurrence of causal contacts. The author stresses the fact that the
-causal contact requirement involves the distance between
-civilizations, their lifespan and their times of appeareance.  This is
-important since the time it takes the light to travel across the
-Galaxy is much lesser than the lifetime of the Galaxy.
-
-\citet{balbi_impact_2018} fix the total number of civilizations and
-explore the parameter space that comprise three parameters, namely,
-the distance to the Earth, the time of appeareance and the lifespan of
-the communicating civilization. Each of these three variables are
-drawn from a random distribution. The distribution of the distances
-results from a uniform distribution of civilization within the plane
-of the Galaxy. For the distribution of the characteristic time of
-appeareance, the author explores and exponential and a trucated
-Gaussian distributions. For the distribution of the lifespans, the
-author choses an exponential distribution. The radius of influence is
-set to 1000 ly.
-
-\citet{balbi_impact_2018} finds that the fraction of communicating
-civilizations vary with the choice of the statistical distribution for
-the time of appeareance. For all analyzed distributions, the fraction
-corresponding to a mean lifetime of 10$^9$ years and for a maximu
-radius for the detection of the signal of 1000 ys, is roughly 0.1.
-
-\citet{vukotic_astrobiological_2012} propose a simulation approach
-based on probabilistic cellular automata (PCA) modeling. In this
-framework, a complex system is modeled by a lattice of cells which
-evolve at discrete time steps, acording to transition rules that take
-into account for each cell the states of its neighbour cells.
-
-The authors implement a PCA model to a network of cells which
-represent life complexity on a 2-dimensional annular ring resembling
-the GHZ. The authors set the GHZ between a minimum radius of 6 kpc and
-a maximum radius of 10 kpc. Within this framework, the authors also
-make several MonteCarlo simulations and analyze ensemble-averaged
-results.
-
-This work aims at analyzing the evolution of life, although it does
-not account for the network of causal contacts among technological
-civilizations.
-
-%- - - -
-
-In this work we address the problem of the temporal and spatial
-structure of the distribution of communicating civilizations, by
-exploring the hypothesis space over a minimal set of parameters.
-%
-In Sec.~\ref{S_methods} we introduce the methods and discuss the
-candidate distributions for the statistical aspects of the times
-involved in the communication process.
-%
-Then, we present our results in Sec~\ref{S_results}, with special
-emphasis on the statistical distributions of the duration of causal
-contacts in one or both directions, the possible differences on the
-position of a civilization on the Galaxy and the distribution of time
-intervals for the waiting of the first contact, always as a function
-of the three simulation parameters.
-%
-Finally, in Sec.~\ref{S_discussion} we discuss our results and future
-research directions.
-
-%}}}
-
-% HACER DISCUSION FUERTE DE ESTOS PAPERS: \cite{balbi_impact_2018,cirkovic_temporal_2004,grimaldi_signal_2017,smith_broadcasting_2009}
-
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
 \input{methods}
 
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
 \input{results}
 
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
 \section{Discusion}\label{S_discussion}
@@
       targetting ratio in the emission.
-   \item Spiritual factor.  Explore the role of the S factor: message
+   \item Explore the role of the message
       contents, influence on the lifespan of a MPL that receives a
@@
 
+\bibliographystyle{mn2e}
 \bibliography{biblio_seti}

### 2019-12-22 methods.tex
con cambios propuestos por Jose

@@
 
-\setlength{\tabcolsep}{10pt}
-\begin{table*}
-\centering
-\begin{tabular}{clccc}
-\hline
-   \multicolumn{2}{l}{independent variables} &min&max&Nbins\\
-\hline
-   $\tau_{a}$ & Mean temporal separation between consecutive awakenings & 4000 y & 200000 y & 50\\ 
-   $\tau_{s}$ & Mean lifetime of a MPL                                 & 10000 y & 500000 y& 50\\ 
-   D$_{max}$ & Maximum reach of a message  &  \multicolumn{3}{l}{500, 1000, 10000, 20000, 40000, 80000 lyr}  \\
-\hline
-   \multicolumn{4}{l}{fixed variables and assumptions} &value \\
-\hline
-   & statistical properties of all MPLs &equally distributed&&\\
-   & Point process for the distribution in time & homogeneous &&\\
-   f$_s$ & The scan of the sky & fully efficient&&1\\
-   f$_p$ & panspermia or colonization &absent&&0\\
-   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
-   R$_{SFC}$   & Inner radius of the GHZ     &&&20000 ly\\
-   R$_{SLC}$   & Outer radius of the GHZ     &&&60000 ly\\
-   t$_{max}$ & Time span of the simulation  &&&1.e6 y\\
-\hline
-   \multicolumn{5}{l}{discrete events}\\
-\hline
-   A & Awakening: a civilization starts a SETI program  &&&\\
-   B & Blackout: the end of the communication chanel stops &&&\\
-   C & Contact: a new causal contact is produced &&&\\
-   D & Doomsday: a civilization ends its communication capabilities&&&\\
-\hline
-
-\hline
-\end{tabular}
-\caption{Definition of independent variables and adopted values for 
-   fixed paramenters 
-   that are part of the simulation.  Variable parameters define the
-   spatial and temporal structure
-   of the process and the maximum reach of the messages.}
-\label{T_simu_hypotheses}
-\end{table*}
-
 While the aforementioned assumptions are basic conditions for most of
@@
 In our simulations we also assume the simplest situations for the
-growth of ETIs.
+growth of the sphere of first contact.
 %
@@
 %
-Also, a mixed paranoid/pronoid population of intelligent civilizations
-could be implemented easily.
+%Also, a mixed paranoid/pronoid population of intelligent civilizations
+%could be implemented easily.
+%
+% EXPLICARLO O SACARLO
+% O PONER EN LA DISCUSION
 %
@@
 countries, corporation sizes, income rankings, ranks of number of
-people watching the same TV channel (Zipf's law entry on Wikipedia).
+people watching the same TV channel (Zipf's law).
+%
+% CITAR ALGO DE LA LEY DE ZIPF
 %

### 2019-12-22 motivations.tex
con cambios propuestos por Jose

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Motivations}\label{S_motivations}
%{{{

The Drake equation \citep{drake_intelligent_1962}
offers a very helpful educated guess,
a rational set of lenses --the factors in the equation-- through which
to look at future contacts with a technologically advanced
civilization in the Milky Way.
%
The equation quantifies the number of
civilizations from whom we might receive an electromagnetic signal.
%
For a comprehensive review and analysis of each term of the equation
see \citet{vakoch_drake_2015}.
%
\citet{prantzos_joint_2013}
proposes a unified framework
for a joint analysis of the Drake equation and the Fermi paradox
concluding that for sufficiently long-lived civilizations,
colonization of the Galaxy is the only reasonable option to gain
knowledge about other life forms.
%
\citet{haqq-misra_drake_2017}
discuss the dependence of the Drake equation parameters on the
spectral type of the host star and the time since the galaxy formed
and examine trajectories for the emergence of communicative
civilizations over the history of the galaxy.
%
The abundance of
intelligent civilizations in the galaxy is a longstanding question,
which is often conceptualized as the problem of the lack of received
communication or the Fermi paradox 
\citep{barlow_galactic_2012, Sotos_biotechnology_2019,
forgan_galactic_2016}.

There are many proposals aimed at solving this paradox, which make use
of statistical \citep{solomonides_probabilistic_2016,
vanhouten_isthere_2017, horvat_calculating_2007,
maccone_statistical_2015} or stochastic approaches
\citep{forgan_numerical_2009, bloetscher_using_2019,
glade_stochastic_2011, forgan_numerical_2010}, analytical
interpretations of the Drake equation \citep{prantzos_joint_2013,
smith_broadcasting_2009} or more speculative proposals
\citep{barlow_galactic_2013, lampton_information_2013,
conway_three_2018, forgan_galactic_2017}.
%
%%%%% agregar Carroll-Nellenback, 2019).
%
However, the uncertainties in the factors of the Drake equation make
it less prone to a formal application in order to define searching
strategies or to compute the actual number of ETIs, but is a key tool
to organize the discussion around the problem of the population of
ETIs in the Galaxy \citep{hinkel_interdisciplinary_2019}, which is
noticeable dominated by the fact that despite decades of effort not a
single ETI, if any, has been detected so far.
%
Some modifications to the original idea of the equation have tried to
imprint a stochastic nature, or to propose a probabilistic approach,
or to consider the temporal structure which is missing in the
equation.

\Fpagebreak

Temporal aspects of the distribution of communicating civilizations
and their contacts have also been explored by several authors
\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011},
%The Impact of the Temporal Distribution of Communicating
%Civilizations on Their Detectability Temporal dispersion of the
%emergence of intelligence: an inter-arrival time analysis The
%spatiotemporal aspects of SETI
%
as well as efforts on considering the stochastic nature of the Drake
equation \citep{glade_stochastic_2011}.
%
The simulation approach has also been considered
\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017},
although with a similar problem that consists on the large number of
parameters which are either unknown or largely uncertain.
%
%New numerical determination of habitability in the Galaxy: the SETI
%connection
%
%Since the factors in the Drake equation are uncertain,
%
In this work, we propose to avoid the frequentist approach of the
Drake equation and to explore a parameter space, where instead of
computing a final number, we provide a statistical distribution that
gives conditional probabilities.
%
This is an exploratory analysis that aims at providing a numerical
tool to discuss not the several theoretical problems summarized by the
Drake equation factors, but the different scenarios on the basis of
statistical heuristics.
%
The approach proposed here shold be considered as a compromise between
the uncertainties of the frequentist approach and the detailed recipes
required on the simulation approach.
%
We then provide a numerical framework to explore via simulations a
parameter space of unknown observables in order to discuss different
scenarios and their consecuences in terms of the probability con
making contacts.
%
The only fact that can be stated with certainty is that for the number
of years SETI projects have been working we have not received any
signal, at least within the technical posibilities and the conditions
stablished by SETI \citep{tarter_search_2001}.
%
This arguably astonishing result has lead to the Fermi paradox, which
states an apparent contradiction between the high estimation of the
abundance of life in the Galaxy and the lack of their detection
\citep{vanhouten_isthere_2017}.
%
The lack of detection has also ispired to a number of alternative
proposals for new SETI strategies \citep{forgan_exoplanet_2017,
balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010,
tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016,
maccone_SETI_2011,  lingam_relative_2019, wright_theGsearch_2015,
maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018,
forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.

The discrete events method for simulating a stochastic process is an
approximation that allows to study the behaviour of complex systems,
by considering a sequence of well defined discrete events.
%
The simulation is carried out by following all the variables that
describe the system, that constitute the state of the system.
%
The evolution of the process, then, is described as a set of changes
in the state of the system.
%
In this context, an event produces a specific change in the state,
that can be triggered by random variables that encode the stochastic
nature of the physical phenomenon.
%
The process involves following the changes on the state of the system,
definig the initial and final states, defining a method that allows to
keep track of time progress, and maintaining a list of relevant
events. (REESCRIBIR!!)

In a recent work, \citep{balbi_impact_2018} use a statistical model to
analyze the occurrence of causal contacts between civilizations in the
Galaxy.
%
The author highlights the effect of evolutionary processes when
attempting to estimate the number of communicating civilizations that
might be in causal contact with an observer located on the Earth.
%
\citet{cirkovic_temporal_2004} also emphasize the lack of temporal
structure in the Drake equation.

%While the Drake equation remains a very useful
%guide in identifying and analyzing the various factors in-
%volved in the problem and their relative probabilities, our
%approach could provide a better framework with which to
%perform a statistical analysis when taking into account evo-
%lutionary processes. In particular, we have shown that the
%Drake equation can misestimate the number of detectable
%communicating species if spatial and temporal dependencies
%are important. This should suggest some caution in the un-
%critical application of such an equation. In this paper, we offer
%an exemplificative illustration of the effect of different
%probability distributions for the typical timescales involved
%in the problem. Ideally, it would be desirable to arrive at a
%mathematical description of such distributions based on a
%modeling of the underlying astrophysical and biological
%processes. This points out a possible future direction of in-
%vestigation in which an integrated evolutionary (and coevo-
%lutionary) approach to the problem would be a central focus
%of interest (for more on this perspective, see, e.g., Chaisson

\citet{bloetscher_using_2019} uses a probabilistic approach, though
still strongly motivated by the Drake equation, to obtain a
probabilistic measure for the number of civilizations in the Galaxy.
%
To that end, perform MonteCarlo Markov Chains over each factor of the
Drake equation, and combine the means to obtain a probabilistic
result.
%
It is worth mentioning that the author proposes a log-normal target
distribution to compute the posterior probabilities.
%
This approach, which assigns probability distributions to the Drake
factors.
%
This study concludes that there is a very low probability for Earth to
make contact with another extraterrestrail civilization.    
%
\citet{smith_broadcasting_2009} uses a analytical model to compute the
probabilities of contact between two randomly located civilizations,
and the waiting time for the first contact, assuming a fixed maximum
broadcasting distance.
% 
\citet{balbi_impact_2018} investigate the chance of communicating
civilizations making causal contact within a volume surrounding the
location of the Earth, using a statistical model to compute the
occurrence of causal contacts. The author stresses the fact that the
causal contact requirement involves the distance between
civilizations, their lifespan and their times of appeareance.  This is
important since the time it takes the light to travel across the
Galaxy is much lesser than the lifetime of the Galaxy.
%
\citet{balbi_impact_2018} fix the total number of civilizations and
explore the parameter space that comprise three parameters, namely,
the distance to the Earth, the time of appeareance and the lifespan of
the communicating civilization. Each of these three variables are
drawn from a random distribution. The distribution of the distances
results from a uniform distribution of civilization within the plane
of the Galaxy. For the distribution of the characteristic time of
appeareance, the author explores and exponential and a trucated
Gaussian distributions. For the distribution of the lifespans, the
author choses an exponential distribution. The radius of influence is
set to 1000 ly.
%
\citet{balbi_impact_2018} finds that the fraction of communicating
civilizations vary with the choice of the statistical distribution for
the time of appeareance. For all analyzed distributions, the fraction
corresponding to a mean lifetime of 10$^9$ years and for a maximu
radius for the detection of the signal of 1000 ys, is roughly 0.1.
%
\citet{vukotic_astrobiological_2012} propose a simulation approach
based on probabilistic cellular automata (PCA) modeling. In this
framework, a complex system is modeled by a lattice of cells which
evolve at discrete time steps, acording to transition rules that take
into account for each cell the states of its neighbour cells.
%
The authors implement a PCA model to a network of cells which
represent life complexity on a 2-dimensional annular ring resembling
the GHZ. The authors set the GHZ between a minimum radius of 6 kpc and
a maximum radius of 10 kpc. Within this framework, the authors also
make several MonteCarlo simulations and analyze ensemble-averaged
results.
%
This work aims at analyzing the evolution of life, although it does
not account for the network of causal contacts among technological
civilizations.

%- - - -

In this work we address the problem of the temporal and spatial
structure of the distribution of communicating civilizations, by
exploring the hypothesis space over a minimal set of parameters.
%
In Sec.~\ref{S_methods} we introduce the methods and discuss the
candidate distributions for the statistical aspects of the times
involved in the communication process.
%
Then, we present our results in Sec~\ref{S_results}, with special
emphasis on the statistical distributions of the duration of causal
contacts in one or both directions, the possible differences on the
position of a civilization on the Galaxy and the distribution of time
intervals for the waiting of the first contact, always as a function
of the three simulation parameters.
%
Finally, in Sec.~\ref{S_discussion} we discuss our results and future
research directions.

% HACER DISCUSION FUERTE DE ESTOS PAPERS: \cite{balbi_impact_2018,cirkovic_temporal_2004,grimaldi_signal_2017,smith_broadcasting_2009}

%}}}

### 2019-12-22 results.tex
autores
con cambios propuestos por Jose

@@
+\setlength{\tabcolsep}{10pt}
+\begin{table*}
+\centering
+\begin{tabular}{clccc}
+\hline
+   \multicolumn{2}{l}{independent variables} &min&max&Nbins\\
+\hline
+   $\tau_{a}$ & Mean temporal separation between consecutive awakenings & 4000 y & 200000 y & 50\\ 
+   $\tau_{s}$ & Mean lifetime of a MPL                                 & 10000 y & 500000 y& 50\\ 
+   D$_{max}$ & Maximum reach of a message  &  \multicolumn{3}{l}{500, 1000, 10000, 20000, 40000, 80000 lyr}  \\
+\hline
+   \multicolumn{4}{l}{fixed variables and assumptions} &value \\
+\hline
+   & statistical properties of all MPLs &equally distributed&&\\
+   & Point process for the distribution in time & homogeneous &&\\
+   f$_s$ & The scan of the sky & fully efficient&&1\\
+   f$_p$ & panspermia or colonization &absent&&0\\
+   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
+   R$_{SFC}$   & Inner radius of the GHZ     &&&20000 ly\\
+   R$_{SLC}$   & Outer radius of the GHZ     &&&60000 ly\\
+   t$_{max}$ & Time span of the simulation  &&&1.e6 y\\
+\hline
+   \multicolumn{5}{l}{discrete events}\\
+\hline
+   A & Awakening: a civilization starts a SETI program  &&&\\
+   B & Blackout: the end of the communication chanel stops &&&\\
+   C & Contact: a new causal contact is produced &&&\\
+   D & Doomsday: a civilization ends its communication capabilities&&&\\
+\hline
+
+\hline
+\end{tabular}
+\caption{Definition of independent variables and adopted values for 
+   fixed paramenters 
+   that are part of the simulation.  Variable parameters define the
+   spatial and temporal structure
+   of the process and the maximum reach of the messages.}
+\label{T_simu_hypotheses}
+\end{table*}
+
+\begin{figure*} % 2D color plot
+   \centering
+   \includegraphics[width=\textwidth]{matrix_1.pdf}
+   \usecaption{F_never_contact}
+   \label{F_never_contact}
+\end{figure*}
+
+\begin{figure*} % 2D color plot
+   \centering
+   \includegraphics[width=\textwidth]{matrix_2.pdf}
+   \usecaption{F_C_at_A}
+   \label{F_C_at_A}
+\end{figure*}
+
+\begin{figure}
+   \centering
+   \includegraphics[width=0.5\textwidth]{waiting_s+a_dif_ylog.pdf}
+   \usecaption{F_waiting_for_1C}
+   \label{F_waiting_for_1C}
+\end{figure}
+
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
 \section{Results: exploring the parameter space}\label{S_results}
@@
 For each model, we simulated 50 realizations with different random
-seeds, in order to estimate the variance.
+seeds, in order to perform a MonteCarlo estimation of the variance.
 %
-The simulated parameters, $\tau_a$, $\tau_s$ cover the ranges
-4000-200000 and 10000-500000, respectively, with a partition of 50
-values for each parameter.
+The parameters for the temporal aspects of the simulation (the mean
+waiting time for the next awakening $\tau_a$ and the mean lifetime
+$\tau_s$) cover the ranges 4000-200000 yr and 10000-500000 yr,
+respectively, with a regular partition of 50 values for each
+parameter.
 %
@@
 %
-In Table \ref{T_simu_hypotheses} we show the range of values of these parameters explored 
-in our numerical experiments.
+In Table \ref{T_simu_hypotheses} we show the set of fixed parameters
+and hipotheses that define the runs of the simulations, along with the
+list of variable parameters and the range of their values that have
+been explored in the numerical experiments.
 
-As a product of the simulations, several quantities are obtained.
+As a product of the simulations, several quantities can be obtained.
 %
 Some quantities are directly derived from the discrete events, namely,
-the ID of emitting MPL, the ID of emitting (receiving) MPL, the
-position in the galaxy, the time of Awakening (Contact) and the time
-of Doomsday  (Blackout) for each MPL.
+the ID of emitting and receiving MPLs, the position in the galaxy, and
+the times of each of the events that are relevant to keep track of the
+number of MPLs (time of awakening and time of doomsday) and of the
+number of contacts (the times of contacts and the times of blackouts).
 %
-This structure is always for the A-D cycle of each emitting MPL, and
-repeats for each contact (C-B).
-%
-We can also derive the additional quantities.
-%
-The age of a MPL is the time elapsed from A to D.
-%
-%
-%given time, ceti\_e is the MPL emitting signals (emiter), ceti\_r is
-%the MPL receiving signals (receiver), ceti\_c is the ceti that
-%listen at least another ceti (citizen) and ceti\_h is the MPL ceti
-%that is lestened by at least another MPL.
+We can also derive the additional quantities that represent the
+properties of the MPLs, for example the total time elapsed between the
+awakening and the doomsday of each MPL.
 %
@@
 %
-This way we can also compute the distribution in the galaxy of cetis
+This way we can also compute the distribution in the galaxy of MPLs
 that reach contact and the waiting time until the first contact or the
@@
 Regarding the properties of the population of MPLs and its evolution,
-we analyze the fraction of awaken time a ceti is listening at least
-another ceti, the age of contacted MPLs at first contact, the
-distribution of time to wait until next contact, the fraction of cetis
-where the first contact is given at the awakening, the distribution of
-the number of contacts for each MPL, the distribution of the maximum
-number of contacts for each MPL, the distribution of the number of
-contacts as a function of MPL age, the number of contacts as a
-function of time in the galaxy, the rate of MPLs that succeed in
-contact, and the distribution of distances between contacted MPLs.
+the simulation yields the fraction of awaken time a MPL is listening
+at least another MPL (i.e., in their causal cone), the age of
+contacted MPLs at first contact, the distribution of time to wait
+until next contact, the fraction of MPLs where the first contact is
+given at the awakening, the distribution of the number of contacts for
+each MPL, the distribution of the number of contacts as a function of
+MPL age, the number of contacts as a function of time in the galaxy,
+the rate of MPLs that succeed in contact, and the distribution of
+distances between contacted MPLs.
 %
-Usefull derived quantity is the duration of two-way communication
-channels the or the fraction of contacts that admit a response.
+Another usefull derived quantity is the duration of two-way
+communication channels or the fraction of contacts that admit a
+response.
 % 
-We also analyze the relations between the distance to ceti vs. time of
-double communication, the distance to ceti vs. age of contacted ceti,
-the age of a ceti and the maximum number of contacted cetis before D,
-or the lifespan of a ceti vs. max number of contacts.
+It is also possible to analyze the relations between the distance to
+MPL vs. the time of double communication, the distance to MPL vs. the
+age of contacted MPL, the age of a MPL and the maximum number of
+contacted MPLs before doomsday, or the lifespan of a MPL vs. the
+maximum number of contacts.
 %
-All these quantities can be analyzed as a function of the simulation
-parameters.
+In addition, all these quantities can be analyzed as a function of the
+simulation parameters.
 %
@@
 explored parameter space, and cover short/long lifetimes, short/long
-waiting times for the next MPL to appear in the Galaxy and
-short/large range reach of the signal, including all possible
-combinations.
+waiting times for the next MPL to appear in the Galaxy and short/large
+range reach of the signal, including all possible combinations.
 %

### 2019-12-22 paper_bckp_dic22.tex
con cambios propuestos por Jose

--- 2019-12-20 methods.tex
@@
+\documentclass[crop]{CSLB}
+
+\usepackage{MySettings}
+
+\input{captions}
+
+\begin{document}
+
+\supertitle{Research Paper}
+
+\title[Probabilities of SETI contacts]{Probability of causal
+contact between interstellar civilizations through MonteCarlo simulations}
+
+\author[Lares, Funes \& Gramajo]{Lares M$^{1, 2}$, Funes J.~G.$^{1, 3}$ \& Gramajo
+L$^{1, 2}$}
+
+\corres{\name{Lares M} \email{marcelo.lares@unc.edu.ar}}
+
+\address{
+   \add{1}{CONICET, Argentina}
+
+   \add{2}{Universidad Nacional de C\'ordoba, Observatorio
+           Astron\'omico de C\'ordoba, Argentina}
+
+   \add{3}{Universidad Cat\'olica de C\'ordoba, Argentina}
+}
+
+\selfcitation{Lares M, Funes J (XXXX).  Estimating the probability of
+causal contact between civilizations through MonteCarlo simulations,
+International Journal of Astrobiology}
+
+\keywords{SETI, Computer simulations, Statistics}
+
+\JELclassification{XX; XX}
+\MSCcodes{XX; XX}
+
+\Abbreviations{MPL: Communicating Extraterrestrial Intelligence, DE:
+Discrete Event, GHZ: Galactic Habitable Zone}
+
+\begin{abstract}
+%
+The abundance of intelligent civilizations in the galaxy is a
+longstanding question, which is often conceptualized as the problem
+of the lack of received communication or the Fermi paradox.
+%
+Most efforts on the estimation of a number of intelligent
+civilizations are centered on the Drake equation, although 
+its factors are affected by large uncertainties, and it lacks a
+temporal nature.
+%
+A alternative approach uses detailed numerical simulations of the galaxy
+and recipes for the rates of the formation of stars and planets, and
+even for the origin of life.
+%
+Here we present numerical simulations of stochastic processes of
+emergence of civilizations with communication capabilities, based
+on minimal asumptions, and analyze the causal connections among them.
+%
+The analysis of the rate of causal contacts as a function of the mean
+number of civilizations, the mean lifetime span distribution and the
+maximum distance a civilization can send signals is presented, and
+used as a framework to discuss the spatial and temporal structure
+of a populated galaxy within several different scenarios.
+%
+Our results indicate that, given the large distances involved, causal
+contacts between civilizations are very rare.
+%
+Additionally, the odds to make a contact in a few years of monitoring, 
+assuming a perfect detection rate, are low for most models, with the
+exception of models which propose a galaxy densely populated with
+long lived civilizations.
+%
+The probability of causal contacts increases with the 
+mean lifetime of civilizations much more signifficantly than with
+the mean number of active civilizations for a given
+time window.
+%
+\end{abstract}
+
+\maketitle
+
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
+\input{motivations}
+
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
+\input{methods}
+
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
+\input{results}
+
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
+\section{Discusion}\label{S_discussion}
+%{{{
+
+We have implemented a suite of simulations following a stochastic
+approach, to explore the hypothesis space of a simple model that
+accounts for the causal connections between communicating
+civilizations on an simplified galaxy.
+%
+The different models can be generated with a minimal number of free
+parameters and some fixed assumptions.
+%
+We argue that three parameters can be used to describe a variety of
+situations, ranging from a Galaxy where an intelligent civilization is
+very rare, to a Galaxy populated with plenty of civilizations in
+causal contact.
+%
+We explored the consecuences of different models using mensurable
+quantities, that arise as by-products of the simulations.
+%
+With this approach, we can estimate several quantities as a function
+of the parameters on the hypothesis space.
+%
+For instance, it is possible to estimate the timescale for an
+uninterrupted SETI effort to reach success, in terms of different
+model parameters that reflect different, so far unknown, scenarios for
+the spread of intelligent life on the galaxy.
+
+Although the simulations use a number of speculative assumptions, we
+argue that the degree of knowledge about the origin and persistence of
+life in the Galaxy does not worth the implementation of more detailed
+or sophisticated models.
+%
+Instead, we take advantadge of the simplicity of the model to explore
+the hypothesis space, i.e., the three--dimensional parameter space of
+the parametric model, in order to gain insight on the consecuences of
+different scenarios for the search of intelligent life.
+%
+Our analysis is not centered in obtaining the odds for the Earth
+making contact with another extraterrestrial civilization, but in
+obtaining a statistical, parameter dependent description of the
+communication system consisting of a group of broadcasting
+civilizations which arise sparsely both in time and space in the
+Galaxy.
+
+Under the hypotheses of our experiments, we conclude that a causal
+contact is extremely unlikely unless the Galaxy is heavily populated
+by intelligent civilizations.
+%
+Even so, we must consider that in order to stablish a contact between
+any two entities, a minimum degree of compatibility must be
+accomplished without any previous agreement, making the posibility of
+a contact with a message that could be deciphered highly rare.
+
+Taking into account the caveat mentioned previously about the
+trade--off between the simplicity and complexity of the experiment,
+future work can be performed following this framework in order to
+explore possible implications of the results for more detailed
+configurations.
+%
+For example, the communication method (isotropic, colimated,
+serendipitous) can affect the observables.
+%
+Since we analyze the causal contacts, effective contacts depend on the
+efficiency of both the emitter and the receiver.
+%
+A correction by coverage ratio in the detection and by a targetting
+ratio in the emission could be easily implemented in the simulation,
+although the effect would be to reduce the probability of contact by
+an efficiency ratio.
+%
+Thus, such implementation is not really necessary, since the values of
+probabilities could modified by a constant correction factor equal to
+the combined emission/reception efficiency
+\citep{smith_broadcasting_2009, anchordoqui_upper_2019,
+forgan_collimated_2014}.
+%
+If we consider a statistical distribution for this factor, then the
+results could change, and this could be studied with the DE approach.
+%
+Also the nature of the message carrier, electromagnetic or another.
+%
+It is also possible to explore the effects of alignments or the use of
+stars as sources or amplifiers.
+%
+%energetic markers Spiritual markers
+%
+As a future work, we will explore several improvements to the
+simulation, and analyze the impact on the final results of these
+changes.
+%
+They include a spatial distribution that actually resembles the shape
+of the Galaxy, different distribution functions for the mean lifetime
+of MPLs and different efficiencies for the sphere of the causal region
+of each MPL, resembling different searching strategies
+\citep{hippke_interstellar_2017}.
+
+Possible hypothesis improvements are:
+
+\begin{itemize}
+   \item D$_{max}$ is different for different MPLs.  For example a power
+      law where powerfull emmisions are rare and weak emssions are
+      common.
+   \item The probability of the rising of a new MPL vary within the GHZ.
+   \item Correction by coverage ratio in the detection and by
+      targetting ratio in the emission.
+   \item Explore the role of the message
+      contents, influence on the lifespan of a MPL that receives a
+      message, mean lifespan of a MPL that emits a message.
+\end{itemize}
+
+%}}}
+
+\ack[Acknowledgement]{
+%
+This work was partially supported by the Consejo Nacional de
+Investigaciones Cient\'{\i}ficas y T\'ecnicas (CONICET, Argentina),
+the Secretar\'{\i}a de Ciencia y Tecnolog\'{\i}a, Universidad Nacional
+de C\'ordoba, Argentina, and the Universidad Cat\'olica de C\'ordoba,
+Argentina.
+%
+This research has made use of NASA's Astrophysics Data System.
+%
+Plots and simulations were made with software developed by the authors
+using R and python languajes. Plots were postprocessed with inkscape.
+%
+}
+
+\bibliographystyle{mn2e}
+\bibliography{biblio_seti}
+
+\end{document}
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
+\section{Motivations}\label{S_motivations}
+%{{{
+
+%A common approach to discussing the estimations of the number of
+%extraterrestrial intelligent civilizations (ETI) in the Galaxy has
+%been the use of the Drake equation \citep{gleiser_drake_2010,
+%prantzos_joint_2013, haqq-misra_drake_2017}.
+%%
+%The abundance of intelligent civilizations in the galaxy is a
+%longstanding question, which is often conceptualized as the problem of
+%the lack of received communication or the Fermi paradox
+%\citep{barlow_galactic_2012, Sotos_biotechnology_2019,
+%forgan_galactic_2016}.
+
+The Drake equation (Drake 1962) offers a very helpful educated guess,
+a rational set of lenses --the factors in the equation-- through which
+to look at future contacts with a technologically advanced
+civilization in the Milky Way.
+%
+The equation quantifies the number of
+civilizations from whom we might receive an electromagnetic signal.
+%
+For a comprehensive review and analysis of each term of the equation
+see Vakoch & Dowd, 2015.
+%
+Prantzos (2013) proposes a unified framework
+for a joint analysis of the Drake equation and the Fermi paradox
+concluding that for sufficiently long-lived civilizations,
+colonization of the Galaxy is the only reasonable option to gain
+knowledge about other life forms.
+%
+Haqq-Misra & Kopparapu (2017)
+discuss the dependence of the Drake equation parameters on the
+spectral type of the host star and the time since the galaxy formed
+and examine trajectories for the emergence of communicative
+civilizations over the history of the galaxy.
+%
+The abundance of
+intelligent civilizations in the galaxy is a longstanding question,
+which is often conceptualized as the problem of the lack of received
+communication or the Fermi paradox (Barlow, 2013a; Sotos, 2019;
+Forgan, 2017a).
+
+There are many proposals intended to resolve this paradox, which make
+use of statistical (Solomonides et al., 2016; VanHouten, 2018;
+Horvat, 2006; Maccone, 2015) or stochastic approaches (Forgan,
+2009; Bloetscher, 2019; Glade et al., 2012; Forgan & Rice,
+2010), analytical interpretations of the Drake equation (Prantzos,
+2013; Smith, 2009) or more speculative proposals (Barlow,
+2013b; Lampton, 2013; Conway Morris, 2018; Forgan, 2017b; Carroll-Nellenback, 2019).
+However, the uncertainties in the factors make it less prone to a
+formal application to define searching strategies or to compute
+numbers, but is used to organize the discussion instead (Hinkel
+et al., 2019). Some modifications to the original idea of the
+equation have tried to imprint a stochastic nature, or to propose a
+probabilistic approach, or to consider the temporal structure which
+is missing in the equation (REFORMULAR). Temporal aspects of the distribution
+of communicating civilizations and their contacts have also been
+explored by several authors (Fogg, 1987; Forgan, 2011; Balbi,
+2018; Balbi, 2018; Horvat et al., 2011), as well as efforts on considering
+the stochastic nature of the Drake equation (Glade et al.,
+2012). The simulation approach has also been considered (Forgan
+et al., 2016; Vukoti´c et al., 2016; Murante et al., 2015; Forgan,
+2009, 2017b)
+
+There are many proposals aimed at solving this paradox, which make use
+of statistical \citep{solomonides_probabilistic_2016,
+vanhouten_isthere_2017, horvat_calculating_2007,
+maccone_statistical_2015} or stochastic approaches
+\citep{forgan_numerical_2009, bloetscher_using_2019,
+glade_stochastic_2011, forgan_numerical_2010}, analytical
+interpretations of the Drake equation \citep{prantzos_joint_2013,
+smith_broadcasting_2009} or more speculative proposals
+\citep{barlow_galactic_2013, lampton_information_2013,
+conway_three_2018, forgan_galactic_2017}.
+%
+
+However, the uncertainties in the factors of the Drake equation make it less
+prone to a formal application in order to define searching strategies
+or to compute the actual number of ETIs, but is a key tool to organize
+the discussion around the problem of the population of ETIs in the
+Galaxy \citep{hinkel_interdisciplinary_2019}, which is noticeable
+dominated by the fact that despite decades of effort not a single ETI,
+if any, has been detected so far.
+%
+Some modifications to the original idea of the equation have tried to
+imprint a stochastic nature, or to propose a probabilistic approach,
+or to consider the temporal structure which is missing in the
+equation.
+%
+
+\Fpagebreak
+
+Temporal aspects of the distribution of communicating civilizations
+and their contacts have also been explored by several authors
+\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
+balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011},
+%The Impact of the Temporal Distribution of Communicating Civilizations on Their Detectability
+%Temporal dispersion of the emergence of intelligence: an inter-arrival time analysis
+%The spatiotemporal aspects of SETI
+%
+as well as efforts on considering the stochastic nature of the Drake equation
+\citep{glade_stochastic_2011}.
+%
+The simulation approach has also benn considered
+\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
+murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017},
+although with a similar problem that consists on the large number of
+parameters which are either unknown or largely uncertain.
+%
+%New numerical determination of habitability in the Galaxy: the SETI connection
+%
+%Since the factors in the Drake equation are uncertain,
+%
+In this work, we propose to
+avoid the frequentist approach of the Drake equation and to explore a
+parameter space, where instead of computing a final number, we provide
+a statistical distribution that gives conditional probabilities.
+%
+This is an exploratory analysis that aims at providing a numerical 
+tool to discuss not the several theoretical problems summarized by the
+Drake equation factors, 
+but the different scenarios on the basis of statistical heuristics.
+%
+The approach proposed here shold be considered as a compromise between
+the uncertainties of the frequentist approach and the detailed recipes
+required on the simulation approach.
+%
+We then provide a numerical framework to explore via simulations a
+parameter space of unknown observables in order to discuss different
+scenarios and their consecuences in terms of the probability con
+making contacts.
+%
+The only fact that can be stated with certainty is that for the number
+of years SETI projects have been working we have not received any
+signal, at least
+within the technical posibilities and the 
+conditions stablished by SETI \citep{tarter_search_2001}.
+%
+This arguably astonishing result has lead to the Fermi paradox, which
+states an apparent contradiction between the high estimation of the
+abundance of life in the Galaxy and the lack of their detection
+\citep{vanhouten_isthere_2017}.
+%
+The lack of detection has also ispired to a number of alternative proposals
+for new SETI strategies \citep{forgan_exoplanet_2017,
+balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010,
+tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016,
+maccone_SETI_2011,  lingam_relative_2019, wright_theGsearch_2015,
+maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018,
+forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
+
+The discrete events method for simulating a stochastic process is an
+approximation that allows to study the behaviour of complex
+systems, by considering a sequence of well defined discrete events.
+%
+The simulation is carried out by following all the variables that
+describe the system, that constitute the state of the system.
+%
+The evolution of the process, then, is described as a set of changes
+in the state of the system.
+%
+In this context, an event produces a specific change in the state,
+that can be triggered by random variables that encode the stochastic
+nature of the physical phenomenon.
+%
+The process involves following the changes on the state of the system,
+definig the initial and final states, defining a method that allows to
+keep track of time progress, and maintaining a list of relevant
+events. (REESCRIBIR!!)
+
+In a recent work, \citep{balbi_impact_2018} use a statistical model to analyze
+the occurrence of causal contacts between civilizations in the Galaxy.
+%
+The author highlights the effect of evolutionary processes when
+attempting to estimate the number of communicating civilizations that
+might be in causal contact with an observer located on the Earth.
+%
+\citet{cirkovic_temporal_2004} also emphasize the lack of temporal
+structure in the Drake equation.
+
+%While the Drake equation remains a very useful
+%guide in identifying and analyzing the various factors in-
+%volved in the problem and their relative probabilities, our
+%approach could provide a better framework with which to
+%perform a statistical analysis when taking into account evo-
+%lutionary processes. In particular, we have shown that the
+%Drake equation can misestimate the number of detectable
+%communicating species if spatial and temporal dependencies
+%are important. This should suggest some caution in the un-
+%critical application of such an equation. In this paper, we offer
+%an exemplificative illustration of the effect of different
+%probability distributions for the typical timescales involved
+%in the problem. Ideally, it would be desirable to arrive at a
+%mathematical description of such distributions based on a
+%modeling of the underlying astrophysical and biological
+%processes. This points out a possible future direction of in-
+%vestigation in which an integrated evolutionary (and coevo-
+%lutionary) approach to the problem would be a central focus
+%of interest (for more on this perspective, see, e.g., Chaisson
+
+\citet{bloetscher_using_2019} uses a probabilistic approach, though
+still strongly motivated by the Drake equation, to obtain a
+propbabilistic measure for the number of civilizations in the Galaxy.
+To that end, perform MonteCarlo Markov Chains over each factor of the
+Drake equation, and combine the means to obtain a probabilistic
+result. It is worth mentioning that the author proposes a log-normal
+target distribution to compute the posterior probabilities. This
+approach, which assigns probability distributions to the Drake
+factors. This study concludes that there is a very low probability for
+Earth to make contact with another extraterrestrail civilization.
+
+\citet{smith_broadcasting_2009} uses a analytical model to compute the
+probabilities of contact between two randomly located civilizations,
+and the waiting time for the first contact, assuming a fixed maximum
+broadcasting distance.
+
+\citet{balbi_impact_2018} investigate the chance of communicating
+civilizations making causal contact within a volume surrounding the
+location of the Earth, using a statistical model to compute the
+occurrence of causal contacts. The author stresses the fact that the
+causal contact requirement involves the distance between
+civilizations, their lifespan and their times of appeareance.  This is
+important since the time it takes the light to travel across the
+Galaxy is much lesser than the lifetime of the Galaxy.
+
+\citet{balbi_impact_2018} fix the total number of civilizations and
+explore the parameter space that comprise three parameters, namely,
+the distance to the Earth, the time of appeareance and the lifespan of
+the communicating civilization. Each of these three variables are
+drawn from a random distribution. The distribution of the distances
+results from a uniform distribution of civilization within the plane
+of the Galaxy. For the distribution of the characteristic time of
+appeareance, the author explores and exponential and a trucated
+Gaussian distributions. For the distribution of the lifespans, the
+author choses an exponential distribution. The radius of influence is
+set to 1000 ly.
+
+\citet{balbi_impact_2018} finds that the fraction of communicating
+civilizations vary with the choice of the statistical distribution for
+the time of appeareance. For all analyzed distributions, the fraction
+corresponding to a mean lifetime of 10$^9$ years and for a maximu
+radius for the detection of the signal of 1000 ys, is roughly 0.1.
+
+\citet{vukotic_astrobiological_2012} propose a simulation approach
+based on probabilistic cellular automata (PCA) modeling. In this
+framework, a complex system is modeled by a lattice of cells which
+evolve at discrete time steps, acording to transition rules that take
+into account for each cell the states of its neighbour cells.
+
+The authors implement a PCA model to a network of cells which
+represent life complexity on a 2-dimensional annular ring resembling
+the GHZ. The authors set the GHZ between a minimum radius of 6 kpc and
+a maximum radius of 10 kpc. Within this framework, the authors also
+make several MonteCarlo simulations and analyze ensemble-averaged
+results.
+
+This work aims at analyzing the evolution of life, although it does
+not account for the network of causal contacts among technological
+civilizations.
+
+%- - - -
+
+In this work we address the problem of the temporal and spatial
+structure of the distribution of communicating civilizations, by
+exploring the hypothesis space over a minimal set of parameters.
+%
+In Sec.~\ref{S_methods} we introduce the methods and discuss the
+candidate distributions for the statistical aspects of the times
+involved in the communication process.
+%
+Then, we present our results in Sec~\ref{S_results}, with special
+emphasis on the statistical distributions of the duration of causal
+contacts in one or both directions, the possible differences on the
+position of a civilization on the Galaxy and the distribution of time
+intervals for the waiting of the first contact, always as a function
+of the three simulation parameters.
+%
+Finally, in Sec.~\ref{S_discussion} we discuss our results and future
+research directions.
+
+% HACER DISCUSION FUERTE DE ESTOS PAPERS: \cite{balbi_impact_2018,cirkovic_temporal_2004,grimaldi_signal_2017,smith_broadcasting_2009}
+
+%}}}
+
 \begin{figure}
@@
 %}}}
+
+\setlength{\tabcolsep}{10pt}
+\begin{table*}
+\centering
+\begin{tabular}{clccc}
+\hline
+   \multicolumn{2}{l}{independent variables} &min&max&Nbins\\
+\hline
+   $\tau_{a}$ & Mean temporal separation between consecutive awakenings & 4000 y & 200000 y & 50\\ 
+   $\tau_{s}$ & Mean lifetime of a MPL                                 & 10000 y & 500000 y& 50\\ 
+   D$_{max}$ & Maximum reach of a message  &  \multicolumn{3}{l}{500, 1000, 10000, 20000, 40000, 80000 lyr}  \\
+\hline
+   \multicolumn{4}{l}{fixed variables and assumptions} &value \\
+\hline
+   & statistical properties of all MPLs &equally distributed&&\\
+   & Point process for the distribution in time & homogeneous &&\\
+   f$_s$ & The scan of the sky & fully efficient&&1\\
+   f$_p$ & panspermia or colonization &absent&&0\\
+   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
+   R$_{SFC}$   & Inner radius of the GHZ     &&&20000 ly\\
+   R$_{SLC}$   & Outer radius of the GHZ     &&&60000 ly\\
+   t$_{max}$ & Time span of the simulation  &&&1.e6 y\\
+\hline
+   \multicolumn{5}{l}{discrete events}\\
+\hline
+   A & Awakening: a civilization starts a SETI program  &&&\\
+   B & Blackout: the end of the communication chanel stops &&&\\
+   C & Contact: a new causal contact is produced &&&\\
+   D & Doomsday: a civilization ends its communication capabilities&&&\\
+\hline
+
+\hline
+\end{tabular}
+\caption{Definition of independent variables and adopted values for 
+   fixed paramenters 
+   that are part of the simulation.  Variable parameters define the
+   spatial and temporal structure
+   of the process and the maximum reach of the messages.}
+\label{T_simu_hypotheses}
+\end{table*}
+
+\begin{figure*} % 2D color plot
+   \centering
+   \includegraphics[width=\textwidth]{matrix_1.pdf}
+   \usecaption{F_never_contact}
+   \label{F_never_contact}
+\end{figure*}
+
+\begin{figure*} % 2D color plot
+   \centering
+   \includegraphics[width=\textwidth]{matrix_2.pdf}
+   \usecaption{F_C_at_A}
+   \label{F_C_at_A}
+\end{figure*}
+
+\begin{figure}
+   \centering
+   \includegraphics[width=0.5\textwidth]{waiting_s+a_dif_ylog.pdf}
+   \usecaption{F_waiting_for_1C}
+   \label{F_waiting_for_1C}
+\end{figure}
+
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
+\section{Results: exploring the parameter space}\label{S_results}
+%{{{
+
+We implemented the simulation of a regular grid of models varying over
+the hypothesis space, which covers 15000 different models.
+%
+For each model, we simulated 50 realizations with different random
+seeds, in order to perform a MonteCarlo estimation of the variance.
+%
+The parameters for the temporal aspects of the simulation (the mean
+waiting time for the next awakening $\tau_a$ and the mean lifetime
+$\tau_s$) cover the ranges 4000-200000 yr and 10000-500000 yr,
+respectively, with a regular partition of 50 values for each
+parameter.
+%
+For the D$_{max}$ parameter, we chose the values 500, 1000, 10000,
+20000, 40000 and 80000 lyr.
+%
+This setup makes a total of 750000 simulations, each covering a time
+range of one million years.
+%
+In Table \ref{T_simu_hypotheses} we show the set of fixed parameters
+and hipotheses that define the runs of the simulations, along with the
+list of variable parameters and the range of their values that have
+been explored in the numerical experiments.
+
+As a product of the simulations, several quantities can be obtained.
+%
+Some quantities are directly derived from the discrete events, namely,
+the ID of emitting and receiving MPLs, the position in the galaxy, and
+the times of each of the events that are relevant to keep track of the
+number of MPLs (time of awakening and time of doomsday) and of the
+number of contacts (the times of contacts and the times of blackouts).
+%
+We can also derive the additional quantities that represent the
+properties of the MPLs, for example the total time elapsed between the
+awakening and the doomsday of each MPL.
+%
+The time span of a MPL listening another or being listened by another
+can also be easily derived from the results of a realization.
+%
+This way we can also compute the distribution in the galaxy of MPLs
+that reach contact and the waiting time until the first contact or the
+waiting time until the next contact.
+%
+Regarding the properties of the population of MPLs and its evolution,
+the simulation yields the fraction of awaken time a MPL is listening
+at least another MPL (i.e., in their causal cone), the age of
+contacted MPLs at first contact, the distribution of time to wait
+until next contact, the fraction of MPLs where the first contact is
+given at the awakening, the distribution of the number of contacts for
+each MPL, the distribution of the number of contacts as a function of
+MPL age, the number of contacts as a function of time in the galaxy,
+the rate of MPLs that succeed in contact, and the distribution of
+distances between contacted MPLs.
+%
+Another usefull derived quantity is the duration of two-way
+communication channels or the fraction of contacts that admit a
+response.
+% 
+It is also possible to analyze the relations between the distance to
+MPL vs. the time of double communication, the distance to MPL vs. the
+age of contacted MPL, the age of a MPL and the maximum number of
+contacted MPLs before doomsday, or the lifespan of a MPL vs. the
+maximum number of contacts.
+%
+In addition, all these quantities can be analyzed as a function of the
+simulation parameters.
+%
+We chose eight models that are on fairly oposed regions of the
+explored parameter space, and cover short/long lifetimes, short/long
+waiting times for the next MPL to appear in the Galaxy and short/large
+range reach of the signal, including all possible combinations.
+%
+The details of each model are summarized in the
+Table~\ref{T_selected_models}.
+
+In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
+distributions of the number contacts for MPLs in four different
+samples (M5, M6, M7, M8) with long range signal reach
+(D$_{max}$=10000). 
+%
+The lifetime is more determinant than the awakening rate in increasing
+the number of contacts.
+%
+As expected, of the four samples, the sample M7, which corresponds to
+a dense awakening in the timeline and a long lifetime, is the model
+that maximizes the number of contacts, reaching a maximum of order 100
+contacts for a single MPL.
+%
+This is considerable larger than any model with a shorter lifetime,
+which produce a number of contacts of at most the order of ten
+contacts per MPL.
+%
+A couple of comments are worth to mention about this plot.
+%
+First, it has a logarithmic scale on the x-axis, and second it is the
+cumulative, not differential, empirical distribution.
+
+In what follows, we focus on the analysis of the duration of contacts
+(Sec.~\ref{SS_multiplicity}), the waiting time for the first contact
+(Sec.~\ref{SS_waiting}) and the
+variations of contact posibilities as a function of the position of
+the MPL in the Galaxy (Sec.~\ref{SS_location}).
+
+%}}}
+
+\subsection{Multiplicity and duration of contacts}\label{SS_multiplicity}
+%{{{
+
+The duration of each contact for a single receiving MPL is given by the
+time difference between a C event and a B event, which are direct
+outputs for each simulation experiment.
+
+%}}}
+
+\subsection{Waiting time for a first contact}\label{SS_waiting}
+%{{{
+
+In this section we analyze the distribution of the waiting times for
+a first contact.
+%
+%Such distribution can be considered to compute the probability of
+%reaching a first contact within the first XX years of a SETI project,
+%assuming the methods are correct and the hypotheses of the experiment.
+Such distribution can be considered to compute the probability for a
+random MPL to carry out a SETI project and spend a given time until
+the first contact is made.
+
+In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the MPLs
+that have made at least one contact as a funtion of the elapsed time
+since the awakening, given that they make at least one contact on
+their lifetime, for different models.
+%
+From a frequentist approach, the normalized cumulative counts are
+related to the estimation of the probability of observing a given time
+with no success.
+%             
+Both panels show the histograms of the waiting times for the first
+contact, for all the MPLs that make at least one contact along their
+lifetime.
+%
+We show separately the histograms for several values of $\tau_a$ and a
+fixed value of $(\tau_s, D_{max})$, and for several values of $\tau_s$
+and a fixed value of $(\tau_a, D_{max})$.
+%
+For different values of $\tau_a$, the shape of the distribution is
+similar, and the main difference given by the different number of
+contacts on the different models.
+%
+Conversely, for different values of $\tau_a$, the shape of the
+distribution is similar, and the main difference given by the
+different number of contacts on the different models. 
+%
+We show the 2D maps of the rate of MPLs that never make contact and
+the rate of MPLs that make contact at the moment of the awakening,
+inf the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
+as a function of $\tau_s$ and $\tau_s$.
+%
+In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
+waiting times for the first contact, for several models.
+%
+Left panel shows the histograms for several values of $\tau_a$, and
+$\tau_s$ in the range 17000-24000 yr.
+%
+Right panel shows the histograms for several values of $\tau_s$, and
+$\tau_a$ in the range 68000-100000 yr.
+%
+The shape of the histograms does not change when $\tau_a$ is modified,
+but significantly changes when $\tau_s$ is modified.
+
+%}}}
+
+\subsection{Location within the Galaxy}\label{SS_location}
+%{{{
+
+%On the Figure~\ref{F_radial_frac} we show the fraction of MPLs that
+%make zero, one, two or three contacts, as a function of the
+%galactocentric distance, for model M1.  VERIFICAR!  M1?
+%
+For this model, the majority of MPLs do not make any contact during
+all the time period in which they are active.
+%
+As expected, the fractions of MPLs making at least one contact
+diminishes as the number of contacts rises.
+%
+The shape of this distribution is a result of the geometry of the GHZ.
+%
+In general, a closer position to the center of the Galaxy slightly
+favours the posibility of a contact.
+%
+This would indicate that the position of the Earth is suitable for the
+odds of making a contact.
+
+%}}}
+
+\setlength{\tabcolsep}{10pt}
+\begin{table*}
+\centering
+\begin{tabular}{ccccc}
+\hline
+Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & description  \\
+\hline
+M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
+M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
+M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
+M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
+%
+M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
+M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
+M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
+M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
+%
+\hline
+\end{tabular}
+\caption{Selected models to analize the behavior of simulation outputs
+   and their dependence on simulation outputs and their dependence on
+   simulation parameters.}
+\label{T_selected_models}
+\end{table*}

### 2019-12-24 captions.tex
detalles de los captions

@@
 different stages in the development of a communicating civilization,
-the growing communicating sphere (a) and the temporal intervals for
-the communication between two civilizations (b).
+or causal contact node (CCN).
 %
-In the upper panel (a), the sphere the encloses the region which is
+Panel (a) shows the sphere of first contact (SFC) centered on an A type
+event and the surface of last contact (SLC), centered on a D type
+event.
+%
+In the upper panel, the sphere that encloses the region which is
 causally connected to the emitter grows until it reaches the
@@
 the surface of last contact.
+%
+The temporal intervals for
+the communication between two civilizations are represented in he
+panel (b).
+%
+The receiver CCN E$_1$ can listen signals from the emiter CCN E$_2$,
+from the contact event (C$_{12}$) up to the blackout event (B$_{12}$).
+%
+Similarly, 
+the receiver CCN E$_2$ can listen signals from the emiter CCN E$_1$,
+from C$_{21}$ up to B$_{21}$. 
 %
@@
 %   
-Schematic representation of the possible cases in which a
-ceti A can be in causal contact with a ceti B which appears later.
+Schematic representation on space--time diagrams of the possible cases in which a
+\ceti E (Emiter, solid line) can be in causal contact with a \ceti R
+(Receiver, dashed line).
+%
+Causal contact can be produced either if the emiter \ceti appears before
+(left) or after (right) the receiver \ceti.
+%
+The causal contact in which the receiver can have access to the signal 
+from the emiter is produced between the C event (Contact, open circles) and the B
+event (Blackout, filled circles).
 %
 The duration of the causal contact in one direction depends on
-several factors, mainly $\Delta t_A$, $\Delta t_C$, and $\Delta X$.
+several factors, mainly the time lag between the awakening events
+(horizontal separation in the space--time diagrams)
+and the distance between the \cetis (vertical separation in the
+space--time diagrams).
 %   
+The different configurations must be taken into account in order to 
+obtain the list of the upcomming events.
+%
+The time intervals for a two--way communication to be possible are
+indicated in double lines.
 }
@@
 Empirical cumulative distributions of the number contacts
-for MPLs in four different samples (M5, M6, M7, M8) with long
-range signal reach (D$_{max}$=10000).
+for \cetis in four different samples (M5, M6, M7, M8) with two
+values for the range of the signal reach: D$_{max}$=40000 ly (blue),  and 
+D$_{max}$=80000 ly (orange).
+%
+Except for the models corresponding to a densely populated galaxy with
+long--living \cetis, the \cetis which receive more than 10 contacts
+are very unlikely.
+%
+See Table \ref{T_selected_models} for model descriptions.
 %

### 2019-12-24 manuscript.tex
detalles de los captions

@@
 
-\JELclassification{XX; XX}
-\MSCcodes{XX; XX}
+%\JELclassification{XX; XX}
+%\MSCcodes{XX; XX}
 
-\Abbreviations{MPL: Communicating Extraterrestrial Intelligence, DE:
-Discrete Event, GHZ: Galactic Habitable Zone}
+\Abbreviations{SETI: Search for Extraterrestrial Intelligence,\\
+CCN: causally connected node,\\
+SFC: surface of first contact,\\
+SLC: surface of last contact,\\
+DE: Discrete Event,\\
+GHZ: Galactic Habitable Zone}
 
@@
 \end{document}
+
+M essage
+P rocessing
+L ocation
+E mission
+R eceiver
+E xtraterrestrial
+I ntelligence
+E xposed
+B roadcasting
+T ransmission
+R eveal
+S ource
+D eliver
+W hisper
+A gent
+N etwork
+C ausal
+C ontact
+I nterstellar
+E ntity
+W atch
+O utlook  (the place from which an observer looks out)
+L ookout  (a station or place from which a watch is kept)
+A wakening (a recognition, realization, or coming into awareness of something)
+
+CETI communicating extraterrestrial intelligence
+MPL message processing location
+ELI   Exposed Lookout Intelligence
+EWI   Exposed Watching Intelligence
+EBIL  exposed Broadcasting Intelligent lookout
+EBIO  exposed Broadcasting Intelligent outlook
+ISBL  Intelligent station for broadcasting and lookout
+BAOS  Intelligent station for broadcasting and outlook
+EBI   exposed broadcasting intelligence
+OSIRIS outlook station for intelligent reception of interstellar signals
+SIS    Source of intelligent signals
+SLSCC source of light speed causal contact
+NLCC node for lightspeec causal contact
+CCN causal contact nodes
+
+awakening event:
+Awakening Active Activation Broadcast
+
+contact event:
+Contact Communication Connect Acquaintance Bond Link
+
+blackout event:
+Blackout Disconnect Detach Cutoff  Obscure Veil Silence Mask
+
+doomsday event:
+Doomsday  Disappear Extinction
+
+Emmit:
+
+Receive:
+catch, seize, hear,

### 2019-12-24 methods.tex
detalles de los captions

@@
 %
-An advantadge of simulations compared to theoretical approaches, is
+An advantage of simulations compared to theoretical approaches, is
 that the former usually require less assumptions and simplifications,
@@
 %
-This method is well suited for the particular case of the difusion of
+This method is well suited for the particular case of the diffusion of
 intelligent signals in the Galaxy, and allows to explore several
@@
 We simulate the statistical properties of a set of points in space and
-time that have the ability of emitting and communicating messages at
-light speed, hereafter dubbed ''message processing locations'' (MPL).
+time that have a causal connection
+at light speed, hereafter dubbed ''causal contact nodes'' (\ceti).
 %
-These locations have the capability of emitting and receiving messages
-in all directions.
+% ability of emitting and receiving messages at
 %
-It is worth mentioning that this is a general approach, and not
-necessarily a MPL is the host of an intelligent civilization.
+We chose this generic name in order to stress the fact that in this
+analysis no broadcasting or lookout activities are required, 
+and therefore not actual messages are considered, only the causal
+contact.
 %
-It can be associated with a set of planets where life has developed,
+For the special case of a fully efficient node that emits
+and receives isotropically, a causal contact node can be considered as
+a broadcasting station that has the ability to detect signals through
+an active lookout program.
+%
+Also, it is worth mentioning that this is a general approach, and not
+necessarily a \ceti is the host of an intelligent civilization.
+%
+It can be associated with a planet where life has developed,
 became intelligent, reached the skills required to find the right
-communication channel, sustain a search and stablish a contact.
+communication channel, sustain a search and establish a contact.
 %
@@
 example interstellar beacons where life has ceased to exist but
-continue with its emission or communication stations stablished by
+continue with its emission or communication stations established by
 probes or left intentionally by intelligent beings \citep[see, e.g.,
@@
 In principle, these strategies could affect our results since it would
-be easier to configure a cluster of MPLs that spread in time.
+be easier to configure a cluster of \cetis that spread in time.
 %
@@
 %
-% communicating, civilization, intelligent, message, processing,
-% handling, locations CALL, MPI, MPL, MPL
-%
-The system is defined by a number of actors that represent MPLs and
+The system is defined by a number of actors that represent \cetis and
 appear at different instants in time, generating events that produce
 meaningful changes in the variables that describe the system, i.e., in
-the arrangement of MPLs and their network of causal contacts.
+the arrangement of \cetis and their network of causal contacts.
 %
-For example, the appeareance of a new MPL in a region which is filled
-with a signal emitted by another MPL, will increase the number of
-active MPLs and the number of pairs of MPLs in causal contact.
+For example, the appearance of a new \ceti in a region which is filled
+with a signal emitted by another \ceti, will increase the number of
+active \cetis and the number of pairs of \cetis in causal contact.
 %
@@
 terms of three independent parameters, namely, the mean time span
-between the appeareance of consecutive MPLs, ($<\tau_a>$), their mean
+between the appeareance of consecutive \cetis, ($<\tau_a>$), their mean
 lifetime ($<\tau_s>$), and the maximum distance a signal can be
-detected by another MPL ($D_{max}$).
+detected by another \ceti ($D_{max}$).
 %
@@
 $\tau_s$ and $D_{max}$ parameters are, the greater the probability of
-the existence of causal contacts between pairs of MPLs would be.
+the existence of causal contacts between pairs of \cetis would be.
 %
@@
 the distribution of lifespans ($\tau_s$) and the distribution of the
-number of MPLs per unit time \citep{maccone_evolution_2014,
+number of \cetis per unit time \citep{maccone_evolution_2014,
 Sotos_biotechnology_2019}, related to the time span between the
-appeareance of consecutive MPLs (since when $<\tau_a>$ is shorter, it
-produces a greater density of MPLs).
+appeareance of consecutive \cetis (since when $<\tau_a>$ is shorter, it
+produces a greater density of \cetis).
 %
@@
 This is a fundamental property of the system which arises as a
-consecuence of the large spatial and temporal scales, where a message
-is transmited at light speed.
+consequence of the large spatial and temporal scales, where a message
+is transmitted at light speed.
 %
-Although a MPL could be active for enough time so as to transmit a
+Although a \ceti could be active for enough time so as to transmit a
 message at large distances, the limited power of the message and the
@@
 As a consecuence of this limitation and of their finite lifetime,
-considered as the period of time between the aquisition and loss of
-communicating capacity, each MPL will fill a spherical shell region of
+considered as the period of time between the acquisition and loss of
+communicating capacity, each \ceti will fill a spherical shell region of
 the Galaxy, limited by two concentric spherical surfaces.
@@
 The leading front, or surface of first contact (SFC) grows from the
-central MPL until it reaches the maximum distance $D_{max}$.
+central \ceti until it reaches the maximum distance $D_{max}$.
 %
-Thus, the volume reached causally by a MPL is initially a growing
+Thus, the volume reached causally by a \ceti is initially a growing
 sphere.
@@
 Following the end of the civilization, there is still a region which
-is filled with the emited signals.
+is filled with the emitted signals.
 %
 The trailing front, or surface of last contact (SLC) also grows from
-the central MPL, with a delay with respect to the SFC equivalent to
-the lifetime of the MPL, and produces a spherical shell region.
+the central \ceti, with a delay with respect to the SFC equivalent to
+the lifetime of the \ceti, and produces a spherical shell region.
 %
-Any other MPL within this region will be aware of the originating MPL,
-even if it has disapeared before the time of contact.
+Any other \ceti within this region will be aware of the originating
+\ceti,
+even if it has disappeared before the time of contact.
 %
@@
 until the surface of last contact reaches $D_{max}$, producing as a
-result the loss of all signal from the central MPL.
+result the loss of all signal from the central \ceti.
 %
 In our approach, we consider a model Galaxy where the width of the
-disk is negligigle with respect to the radius of the disk, so that the
-MPLs are placed on a 2D plane disk.
+disk is negligible with respect to the radius of the disk, so that the
+\cetis are placed on a 2D plane disk.
 %
@@
 We illustrate in the Fig.~\ref{F_scheme} the schematic representation
-of the region causally connected to the central MPL.
+of the region causally connected to the central \ceti.
 %
@@
 %
-At the end of the lifetime of a MPL, a growing surface of last
+At the end of the lifetime of a \ceti, a growing surface of last
 contact is represented by a circle that covers the inner part of the
-causal region, producing a ring of diminishig area (panel (a)).
+causal region, producing a ring of diminishing area (panel (a)).
 %
-In this scheme, the entire lifetime of a MPL can then be represented
+In this scheme, the entire lifetime of a \ceti can then be represented
 by a polygon, limited in the spatial axis by the maximum distance of
@@
 This representation allows to easily visualize the important events
-that result from the existence and communications of two MPLs (panel
+that result from the existence and communications of two \cetis (panel
 (b)).
 %
-The points in time where the MPLs aquire their communicating capacity
+The points in time where the \cetis acquire their communicating capacity
 (CC) are dubbed ''awakening'' events ($A_1$ and $A_2$ in the Figure).
 %
-Simmilarly, the points in time where the MPLs loss their
+Similarly, the points in time where the \cetis loss their
 communicating capacity are dubbed ''doomsday'' events ($D_1$ and $D_2$
@@
 The points in space--time where the first contact is produced for each
-one of the MPLs, are defined as ''contact'' events, shown as $C_1$
+one of the \cetis, are defined as ''contact'' events, shown as $C_1$
 and $C_2$ in the Figure.
@@
 Finally, the two points where the contact is lost for each one the
-MPLs, are denominated ''blackout'' events, shown as $B_1$ and $B_2$
+\cetis, are denominated ''blackout'' events, shown as $B_1$ and $B_2$
 in the Figure.
@@
 It is important to notice that there is a time delay between the
-contact events of the two MPLs involved in this analysis, and also a
+contact events of the two \cetis involved in this analysis, and also a
 time delay between the two blackout events, so that the time range
@@
 In the Fig.~\ref{F_messages} we show a schematic representation of the
-possible cases in which a MPL A can be in causal contact with a MPL
+possible cases in which a \ceti A can be in causal contact with a
+\ceti
 B which appears later.
@@
 civilization can emit and receive signals, and the mean time interval
-between the emergence of an intelligent comunicating civilization and
+between the emergence of an intelligent communicating civilization and
 the next one.
@@
 The parameters for the temporal distributions also determine the
-spatial properties since the density of active cetis in the Galaxy
+spatial properties since the density of active \cetis in the Galaxy
 depend on these two parameters.
@@
 For example, a small $\tau_a$ and a large $\tau_s$ will produce a
-Galaxy densely populated with MPLs, and viceversa.
+Galaxy densely populated with \cetis, and vice versa.
 %
@@
 stationary Poisson process, and then the distribution of the times
-between the appeareance of new consecutive MPLs is exponential.
+between the appeareance of new consecutive \cetis is exponential.
 %
-Regarding the duration of a MPL, we propose that the distribution of
-the duration of MPLs is a stationary exponential distribution.
+Regarding the duration of a \ceti, we propose that the distribution of
+the duration of \cetis is a stationary exponential distribution.
 %
@@
 %
-The probability for the rise of a MPL is homogeneous over the GHZ.
+The probability for the rise of a \ceti is homogeneous over the GHZ.
 %
 Although the Galaxy has a well known spiral structure, the assumed
-sparcity of civilizations makes this assumption very reasonably.
+sparsity of civilizations makes this assumption very reasonably.
 %
-Otherwise, if the distribution of civilizations is not sparce, it
+Otherwise, if the distribution of civilizations is not sparse, it
 could be the case that the spiral arms would host most of the
-civilizations, and contacts are frecuent between closely located
+civilizations, and contacts are frequent between closely located
 civilizations.
@@
 In such a case we could expect that the interpretation of our results
-could be considered as pesimist for the stablishment of a contact with
+could be considered as pessimist for the establishment of a contact with
 our planet.
@@
 electromagnetic radiation or even gravitational radiation, but we set
-appart messages sent with mechanical means or physical objects, or
+apart messages sent with mechanical means or physical objects, or
 through some unknown technology that violates the known so far laws of
@@
 % 
-For the communication of MPLs through messages sent isotropically, we
-assume that the capacity to emit signals and to receive signals ocurr
+For the communication of \cetis through messages sent isotropically, we
+assume that the capacity to emit signals and to receive signals occur
 at the same time.
@@
 the same time that the ability and intention to emit a signal in all
-direction, intented to be detected by an unknown civilization.
+direction, intended to be detected by an unknown civilization.
 %
@@
 case, at large time scales it can be considered that both abilities
-occurr roughly at coincident epochs.
+occur roughly at coincident epochs.
 %
-Another essential assumption is that all MPLs use the same signal
+Another essential assumption is that all \cetis use the same signal
 power, so that there is a maximum distance out to which it can be
@@
 It is worth mentioning that we are considering in our experimental
-setup a system composed by several emmiters and receivers across the
+setup a system composed by several emiters and receivers across the
 Galaxy, under the hypothesis proposed before.
 %
-In such a system we compute probabilities of a MPL making contact
-with another MPL, and the results must be interpreted as for a
-randomly chosen MPL, given certain conditions of homogeneityon an
+In such a system we compute probabilities of a \ceti making contact
+with another \ceti, and the results must be interpreted as for a
+randomly chosen \ceti, given certain conditions of homogeneity on an
 ensemble of equivalent galaxies.
@@
 We stress the fact that we are not centered on the probabilities for
-the Earth itselft, but on a generalized MPL.
+the Earth itself, but on a generalized \ceti.
 %
@@
 This proposal is based on a pronoid scenario (REF), and we do not
-consider in this work the ocurrence of paranoid or partially paranoid
+consider in this work the occurrence of paranoid or partially paranoid
 civilizations.
@@
 %
-The lifetime of a civilization can be caused by autodestruction or by
+The lifetime of a civilization can be caused by auto destruction or by
 external factors.
@@
 %
-This means that there is no preferred location within the GHZ for the
-spontaneos appeareance of life, and that the emergence of a
+Equivalently there is no preferred location within the GHZ for the
+spontaneous appeareance of life, and that the emergence of a
 civilization is independent of the existence of previous civilizations
@@
 This is equivalent to proposing a Poisson process for the emergence of
-MPLs, since there is a close relation between the number of events in
+\cetis, since there is a close relation between the number of events in
 time or space and the waiting time or separation, respectively.
@@
 process, a Poisson distribution for the number of events implies an
-exponential distribution for their sepations, and viceversa.
+exponential distribution for their separations, and viceversa.
 %
@@
 %
-For example, the distribution of the frecuency of words in many
-languajes is known to follow a Zipf law (which is a power law).
+For example, the distribution of the frequency of words in many
+languages is known to follow a Zipf law (which is a power law).
 %
@@
 Power laws have also been widely used in biological sciences, e.g., in
-analysing connectivity patterns in metabolic networks
+analyzing connectivity patterns in metabolic networks
 \citep{jeong_large_2000} and in the number of species observed per
@@
 positions of stars, which are sampled randomly within the GHZ, the
-time of the awakening of each MPL (A event), and the time of the
-vanishing of each MPL (D event).
+time of the awakening of each \ceti (A event), and the time of the
+vanishing of each \ceti (D event).
 %
 The variables that can be deduced from the previous ones include the
-number of MPLs in casual contact with at least another MPL at a given
-time, the number of MPLs as a function of time, the number of MPLs
-that receive a message at least one time, the number of MPLs that
+number of \cetis in casual contact with at least another \ceti at a given
+time, the number of \cetis as a function of time, the number of \cetis
+that receive a message at least one time, the number of \cetis that
 receive a message at least one time and successfully deliver an answer
@@
 
-\subsection{Model for MPLs}
+\subsection{Model for \cetis}
 %{{{
@@
 %
-This simple model does not take into accont the variations in stellar
+This simple model does not take into account the variations in stellar
 density given by the spiral structure.
@@
 \begin{enumerate}
-   \item[(A)] A new MPL appears (Awakening)
-   \item[(D)] An old MPL dissappears (Doomsday)
-   \item[(C)] A new causal contact is stablished (Contact)
-   \item[(B)] An existing causal contact is interrupted (Blackout)
+   \item[(A)] Awakening: A new \ceti acquires the ability to emit
+      and receive messages and starts to actively broadcast and
+      lookout for messages.
+   \item[(C)] Contact: A new one-way causal contact is established
+      (does not require the awareness of the involved \cetis.)
+   \item[(B)] Blackout: An existing causal contact is interrupted,
+      with or without the occurrence of a doomsday event.
+   \item[(D)] Doomsday: An old \ceti disappears, or halts its
+      communication activities.
 \end{enumerate}
 
-The event of type B is produced when one of the two MPLs halts in its
-capability of emmiting and receiving signals.
+We have chosen this nomenclature so that it can be easily remembered
+from the initial A, B, C, D.
+%
+The \blackout is produced when one of the two \cetis halts in its
+capability of emiting and receiving signals.
+%
+It can occur at the same time the \doomsday or before.
+%
+Similarly, a \ccontact can be produced at the time of the \aawakening
+or later, and there can be several \contacts for the same \ceti.
 
@@
 %
-The time marks for each event (A and D for each MPL and C and B for
+The time marks for each event (A and D for each \ceti and C and B for
 each causal contact) are stored as a result of the simulation.
 %
-Also, the list of active MPLs is obtained as a function of time.
+Also, the list of active \cetis is obtained as a function of time.
 %
 Some fixed variables must be set in order to carry out the simulation:
-the size of the Galactic Habitable Zone, the mean lifetime of a MPL,
-$\tau_s$, the mean waiting time for the appeareance of another MPL,
+the size of the Galactic Habitable Zone, the mean lifetime of a \ceti,
+$\tau_s$, the mean waiting time for the appeareance of another \ceti,
 $\tau_a$, and the maximum distance a message can be detected.
@@
 that is barely known, is set at fixed values of the inner and outer
-radii, radial symetry is assumed, and the with of the galactic disk is
+radii, radial symmetry is assumed, and the with of the galactic disk is
 neglected compared to the radial size.
@@
 
-The simulation was implemented on a python 3 code, which is publically
+The simulation was implemented on a python 3 code, which is publicly
 available\footnote{https://github.com/mlares/simu\_contact}

### 2019-12-24 motivations.tex
detalles de los captions

@@
 it less prone to a formal application in order to define searching
-strategies or to compute the actual number of ETIs, but is a key tool
-to organize the discussion around the problem of the population of
-ETIs in the Galaxy \citep{hinkel_interdisciplinary_2019}, which is
+strategies or to compute the actual number of extraterrestrial
+intelligences, but is a key tool
+to organize the discussion around the problem of their abundance
+in the Galaxy \citep{hinkel_interdisciplinary_2019}, which is
 noticeable dominated by the fact that despite decades of effort not a
-single ETI, if any, has been detected so far.
+single extraterrestrial intelligence, if any, has been detected so far.
 %
@@
 %
-The approach proposed here shold be considered as a compromise between
+The approach proposed here should be considered as a compromise between
 the uncertainties of the frequentist approach and the detailed recipes
@@
 parameter space of unknown observables in order to discuss different
-scenarios and their consecuences in terms of the probability con
+scenarios and their consequences in terms of the probability con
 making contacts.
@@
 of years SETI projects have been working we have not received any
-signal, at least within the technical posibilities and the conditions
-stablished by SETI \citep{tarter_search_2001}.
+signal, at least within the technical possibilities and the conditions
+established by SETI \citep{tarter_search_2001}.
 %
@@
 %
-The lack of detection has also ispired to a number of alternative
+The lack of detection has also inspired to a number of alternative
 proposals for new SETI strategies \citep{forgan_exoplanet_2017,
@@
 The process involves following the changes on the state of the system,
-definig the initial and final states, defining a method that allows to
+defining the initial and final states, defining a method that allows to
 keep track of time progress, and maintaining a list of relevant
@@
 This study concludes that there is a very low probability for Earth to
-make contact with another extraterrestrail civilization.    
+make contact with another extraterrestrial civilization.    
 %
@@
 causal contact requirement involves the distance between
-civilizations, their lifespan and their times of appeareance.  This is
+civilizations, their lifespan and their times of appearance.  This is
 important since the time it takes the light to travel across the
@@
 explore the parameter space that comprise three parameters, namely,
-the distance to the Earth, the time of appeareance and the lifespan of
+the distance to the Earth, the time of appearance and the lifespan of
 the communicating civilization. Each of these three variables are
@@
 of the Galaxy. For the distribution of the characteristic time of
-appeareance, the author explores and exponential and a trucated
+appearance, the author explores and exponential and a truncated
 Gaussian distributions. For the distribution of the lifespans, the
-author choses an exponential distribution. The radius of influence is
+author chooses an exponential distribution. The radius of influence is
 set to 1000 ly.
@@
 civilizations vary with the choice of the statistical distribution for
-the time of appeareance. For all analyzed distributions, the fraction
-corresponding to a mean lifetime of 10$^9$ years and for a maximu
-radius for the detection of the signal of 1000 ys, is roughly 0.1.
+the time of appearance. For all analyzed distributions, the fraction
+corresponding to a mean lifetime of 10$^9$ years and for a maximum
+radius for the detection of the signal of 1000 yr, is roughly 0.1.
 %
@@
 framework, a complex system is modeled by a lattice of cells which
-evolve at discrete time steps, acording to transition rules that take
+evolve at discrete time steps, according to transition rules that take
 into account for each cell the states of its neighbour cells.

### 2019-12-24 results.tex
detalles de los captions

@@
    $\tau_{a}$ & Mean temporal separation between consecutive awakenings & 4000 y & 200000 y & 50\\ 
-   $\tau_{s}$ & Mean lifetime of a MPL                                 & 10000 y & 500000 y& 50\\ 
+   $\tau_{s}$ & Mean lifetime of a \ceti                                 & 10000 y & 500000 y& 50\\ 
    D$_{max}$ & Maximum reach of a message  &  \multicolumn{3}{l}{500, 1000, 10000, 20000, 40000, 80000 lyr}  \\
@@
 \hline
-   & statistical properties of all MPLs &equally distributed&&\\
+   & statistical properties of all \cetis &equally distributed&&\\
    & Point process for the distribution in time & homogeneous &&\\
@@
    A & Awakening: a civilization starts a SETI program  &&&\\
-   B & Blackout: the end of the communication chanel stops &&&\\
+   B & Blackout: the end of the communication channel stops &&&\\
    C & Contact: a new causal contact is produced &&&\\
@@
 \caption{Definition of independent variables and adopted values for 
-   fixed paramenters 
+   fixed parameters 
    that are part of the simulation.  Variable parameters define the
@@
 In Table \ref{T_simu_hypotheses} we show the set of fixed parameters
-and hipotheses that define the runs of the simulations, along with the
+and hypotheses that define the runs of the simulations, along with the
 list of variable parameters and the range of their values that have
@@
 Some quantities are directly derived from the discrete events, namely,
-the ID of emitting and receiving MPLs, the position in the galaxy, and
+the ID of emitting and receiving \cetis, the position in the galaxy, and
 the times of each of the events that are relevant to keep track of the
-number of MPLs (time of awakening and time of doomsday) and of the
+number of \cetis (time of awakening and time of doomsday) and of the
 number of contacts (the times of contacts and the times of blackouts).
@@
 We can also derive the additional quantities that represent the
-properties of the MPLs, for example the total time elapsed between the
-awakening and the doomsday of each MPL.
+properties of the \cetis, for example the total time elapsed between the
+awakening and the doomsday of each \ceti.
 %
-The time span of a MPL listening another or being listened by another
+The time span of a \ceti listening another or being listened by another
 can also be easily derived from the results of a realization.
 %
-This way we can also compute the distribution in the galaxy of MPLs
+This way we can also compute the distribution in the galaxy of \cetis
 that reach contact and the waiting time until the first contact or the
@@
 %
-Regarding the properties of the population of MPLs and its evolution,
-the simulation yields the fraction of awaken time a MPL is listening
-at least another MPL (i.e., in their causal cone), the age of
-contacted MPLs at first contact, the distribution of time to wait
-until next contact, the fraction of MPLs where the first contact is
+Regarding the properties of the population of \cetis and its evolution,
+the simulation yields the fraction of awaken time a \ceti is listening
+at least another \ceti (i.e., in their causal cone), the age of
+contacted \cetis at first contact, the distribution of time to wait
+until next contact, the fraction of \cetis where the first contact is
 given at the awakening, the distribution of the number of contacts for
-each MPL, the distribution of the number of contacts as a function of
-MPL age, the number of contacts as a function of time in the galaxy,
-the rate of MPLs that succeed in contact, and the distribution of
-distances between contacted MPLs.
+each \ceti, the distribution of the number of contacts as a function of
+\ceti age, the number of contacts as a function of time in the galaxy,
+the rate of \cetis that succeed in contact, and the distribution of
+distances between contacted \cetis.
 %
-Another usefull derived quantity is the duration of two-way
+Another useful derived quantity is the duration of two-way
 communication channels or the fraction of contacts that admit a
@@
 It is also possible to analyze the relations between the distance to
-MPL vs. the time of double communication, the distance to MPL vs. the
-age of contacted MPL, the age of a MPL and the maximum number of
-contacted MPLs before doomsday, or the lifespan of a MPL vs. the
+\ceti vs. the time of double communication, the distance to \ceti vs. the
+age of contacted \ceti, the age of a \ceti and the maximum number of
+contacted \cetis before doomsday, or the lifespan of a \ceti vs. the
 maximum number of contacts.
@@
 %
-We chose eight models that are on fairly oposed regions of the
+We chose eight models that are on fairly opposed regions of the
 explored parameter space, and cover short/long lifetimes, short/long
-waiting times for the next MPL to appear in the Galaxy and short/large
+waiting times for the next \ceti to appear in the Galaxy and short/large
 range reach of the signal, including all possible combinations.
@@
 In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
-distributions of the number contacts for MPLs in four different
+distributions of the number contacts for \cetis in four different
 samples (M5, M6, M7, M8) with long range signal reach
@@
 that maximizes the number of contacts, reaching a maximum of order 100
-contacts for a single MPL.
+contacts for a single \ceti.
 %
@@
 which produce a number of contacts of at most the order of ten
-contacts per MPL.
+contacts per \ceti.
 %
@@
 (Sec.~\ref{SS_waiting}) and the
-variations of contact posibilities as a function of the position of
-the MPL in the Galaxy (Sec.~\ref{SS_location}).
+variations of contact possibilities as a function of the position of
+the \ceti in the Galaxy (Sec.~\ref{SS_location}).
 
@@
 
-The duration of each contact for a single receiving MPL is given by the
+The duration of each contact for a single receiving \ceti is given by the
 time difference between a C event and a B event, which are direct
@@
 Such distribution can be considered to compute the probability for a
-random MPL to carry out a SETI project and spend a given time until
+random \ceti to carry out a SETI project and spend a given time until
 the first contact is made.
 
-In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the MPLs
-that have made at least one contact as a funtion of the elapsed time
+In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the \cetis
+that have made at least one contact as a function of the elapsed time
 since the awakening, given that they make at least one contact on
@@
 Both panels show the histograms of the waiting times for the first
-contact, for all the MPLs that make at least one contact along their
+contact, for all the \cetis that make at least one contact along their
 lifetime.
@@
 %
-We show the 2D maps of the rate of MPLs that never make contact and
-the rate of MPLs that make contact at the moment of the awakening,
-inf the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
+We show the 2D maps of the rate of \cetis that never make contact and
+the rate of \cetis that make contact at the moment of the awakening,
+in the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
 as a function of $\tau_s$ and $\tau_s$.
@@
 
-%On the Figure~\ref{F_radial_frac} we show the fraction of MPLs that
+%On the Figure~\ref{F_radial_frac} we show the fraction of \cetis that
 %make zero, one, two or three contacts, as a function of the
@@
 %
-For this model, the majority of MPLs do not make any contact during
+For this model, the majority of \cetis do not make any contact during
 all the time period in which they are active.
 %
-As expected, the fractions of MPLs making at least one contact
+As expected, the fractions of \cetis making at least one contact
 diminishes as the number of contacts rises.
@@
 In general, a closer position to the center of the Galaxy slightly
-favours the posibility of a contact.
+favours the possibility of a contact.
 %
@@
 \end{tabular}
-\caption{Selected models to analize the behavior of simulation outputs
+\caption{Selected models to analyze the behavior of simulation outputs
    and their dependence on simulation outputs and their dependence on

### 2019-12-29 manuscript.tex
preparando todo para scholar one

--- 2019-12-22 manuscript.tex
@@
 
-\author[Lares, Funes \& Gramajo]{Lares M$^{1, 2}$, Funes J.~G.$^{1, 3}$ \& Gramajo
-L$^{1, 2}$}
+\author[Lares, Funes \& Gramajo]{Lares M.$^{1, 2}$, Funes J.~G.$^{1,
+3}$ \& Gramajo L.$^{1, 2}$}
 
@@
 
-\selfcitation{Lares M, Funes J (XXXX).  Estimating the probability of
-causal contact between civilizations through MonteCarlo simulations,
-International Journal of Astrobiology}
-
 \keywords{SETI, Computer simulations, Statistics}
 
-\JELclassification{XX; XX}
-\MSCcodes{XX; XX}
+%\JELclassification{XX; XX}
+%\MSCcodes{XX; XX}
 
-\Abbreviations{MPL: Communicating Extraterrestrial Intelligence, DE:
-Discrete Event, GHZ: Galactic Habitable Zone}
+\Abbreviations{SETI: Search for Extraterrestrial Intelligence,\\
+CCN: causally connected node,\\
+SFC: surface of first contact,\\
+SLC: surface of last contact,\\
+DE: Discrete Event,\\
+GHZ: Galactic Habitable Zone}
 
@@
 %%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
-\input{motivations}
+\input{Sec_motivations}
 
 %%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
-\input{methods}
+\input{Sec_methods}
 
 %%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
-\input{results}
+\input{Sec_results}
 
@@
 
+\ack[Disclosure statement]{
+%
+No competing financial interests exist.
+%
+}
+
 \bibliographystyle{mn2e}

### 2019-12-29 Sec_methods.tex
preparando todo para scholar one
seccion leyes de prob

--- 2019-12-24 methods.tex
@@
    \centering
-   \includegraphics[width=0.5\textwidth]{scheme.pdf}
+   \includegraphics[width=0.5\textwidth]{F_scheme.pdf}
    \usecaption{F_scheme} \label{F_scheme}
@@
    \centering
-   \includegraphics[width=0.8\textwidth]{Messages_01.pdf}
+   \includegraphics[width=0.8\textwidth]{F_messages.pdf}
    \usecaption{F_messages} \label{F_messages}
@@
 In the Fig.~\ref{F_messages} we show a schematic representation of the
-possible cases in which a \ceti A can be in causal contact with a
+possible cases in which a \ceti E can be in causal contact with a
 \ceti
-B which appears later.
+R.
 %
 The duration of the causal contact in one direction depends on several
-factors, mainly $\Delta$t~A , $\Delta$t~C , and $\Delta$X.
+factors, mainly the lifetimes, the time period between the A events 
+and the distance between the two \cetis involved in the causal
+contact.
 
@@
 
-The exponential distribution of lifespan and waiting times is
-justified by considering that the process of appeareance of life in
-the galaxy is homogeneous and stationary.
-%
-Equivalently there is no preferred location within the GHZ for the
-spontaneous appeareance of life, and that the emergence of a
-civilization is independent of the existence of previous civilizations
-in the galaxy.
-%            
-This is equivalent to proposing a Poisson process for the emergence of
-\cetis, since there is a close relation between the number of events in
-time or space and the waiting time or separation, respectively.
-%
-That is, these are two alternative approaches to describing the same
-process, a Poisson distribution for the number of events implies an
-exponential distribution for their separations, and viceversa.
-%
-It should be emphasized that the exponential laws used in this work
-are assumed, although instead of analyzing results from a particular
-parameter chosen ad--hoc, we explore the hypothesis space and analyze
-the impact of the values of these parameters on the results.
-
 The power law and exponential statistical distributions are among the
@@
 benguigui_classificacion_2016}
+
+In this work, the exponential distribution of lifespan and waiting
+times is justified by considering the hypothesis that the process of
+appeareance of life in the galaxy is homogeneous and stationary.
+%
+That is, there is no preferred location within the GHZ for the
+spontaneous appeareance of life, and the emergence of a \ceti is
+independent of the existence of previous \cetis in the galaxy.
+%            
+This is equivalent to proposing a Poisson process for the emergence of
+\cetis, since there is a close relation between the number of events
+in time or space and the waiting time or separation, respectively
+\citep[e.g., ][]{ross_simulation_2012}.
+%
+That is, these are two alternative approaches to describing the same
+process, a Poisson distribution for the number of events implies an
+exponential distribution for their separations, and viceversa.
+%
+It should be emphasized that the exponential laws used in this work
+are assumed as part of the working hypothesis, and instead of
+analyzing results from a particular parameter chosen ad--hoc, we
+explore the hypothesis space and analyze the impact of the values of
+these parameters on the results.
 
@@
 \begin{figure} \centering
-   \includegraphics[width=0.5\textwidth]{F1.pdf}
+   \includegraphics[width=0.5\textwidth]{F_number_of_contacts.pdf}
    \usecaption{F_number_of_contacts} \label{F_number_of_contacts}

### 2020-01-04 Sec_methods.tex
acomodando las figuras

@@
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{F_scheme.pdf}
-   \usecaption{F_scheme} \label{F_scheme}
-\end{figure}
-
-\begin{figure*}
-   \centering
-   \includegraphics[width=0.8\textwidth]{F_messages.pdf}
-   \usecaption{F_messages} \label{F_messages}
-\end{figure*}
-
 %%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
@@
 time that have a causal connection
-at light speed, hereafter dubbed ''causal contact nodes'' (\ceti).
+at light speed, hereafter dubbed ''causal contact nodes'' (\ceti{}).
 %
@@
 Also, it is worth mentioning that this is a general approach, and not
-necessarily a \ceti is the host of an intelligent civilization.
+necessarily a \ceti{} is the host of an intelligent civilization.
 %
@@
 In principle, these strategies could affect our results since it would
-be easier to configure a cluster of \cetis that spread in time.
+be easier to configure a cluster of \cetis{} that spread in time.
 %
@@
 %
-The system is defined by a number of actors that represent \cetis and
+The system is defined by a number of actors that represent \cetis{} and
 appear at different instants in time, generating events that produce
 meaningful changes in the variables that describe the system, i.e., in
-the arrangement of \cetis and their network of causal contacts.
+the arrangement of \cetis{} and their network of causal contacts.
 %
-For example, the appearance of a new \ceti in a region which is filled
-with a signal emitted by another \ceti, will increase the number of
-active \cetis and the number of pairs of \cetis in causal contact.
+For example, the appearance of a new \ceti{} in a region which is filled
+with a signal emitted by another \ceti{}, will increase the number of
+active \cetis{} and the number of pairs of \cetis{} in causal contact.
 %
@@
 terms of three independent parameters, namely, the mean time span
-between the appeareance of consecutive \cetis, ($<\tau_a>$), their mean
-lifetime ($<\tau_s>$), and the maximum distance a signal can be
-detected by another \ceti ($D_{max}$).
+between the appeareance of consecutive \cetis{}, $\left<\tau_a\right>$, their mean
+lifetime $\left<\tau_s\right>$, and the maximum distance a signal can be
+detected by another \ceti{} ($D_{max}$).
 %
@@
 $\tau_s$ and $D_{max}$ parameters are, the greater the probability of
-the existence of causal contacts between pairs of \cetis would be.
+the existence of causal contacts between pairs of \cetis{} would be.
 %
@@
 the distribution of lifespans ($\tau_s$) and the distribution of the
-number of \cetis per unit time \citep{maccone_evolution_2014,
+number of \cetis{} per unit time \citep{maccone_evolution_2014,
 Sotos_biotechnology_2019}, related to the time span between the
-appeareance of consecutive \cetis (since when $<\tau_a>$ is shorter, it
-produces a greater density of \cetis).
+appeareance of consecutive \cetis{} (since when $\left<\tau_a\right>$ is shorter, it
+produces a greater density of \cetis{}).
 %
@@
 in Sec.~\ref{SS_PDF_shape}.
+
+%}}}
+
+\begin{figure}[!t]
+   \centering
+   \includegraphics[width=0.5\textwidth]{F_scheme.pdf}
+   \usecaption{F_scheme} \label{F_scheme}
+\end{figure}
+
+\begin{figure*}
+   \centering
+   \includegraphics[width=0.8\textwidth]{F_messages.pdf}
+   \usecaption{F_messages} \label{F_messages}
+\end{figure*}
+
+%{{{
+
+\ffn{1}
+We illustrate in the Fig.~\ref{F_scheme} the schematic representation
+of the region causally connected to the central \ceti{}.
+%
+To this end, we use space--time diagrams, where time is represented on
+the horizontal axis, and space is represented in the vertical axis.
+%
+In particular, we plot on the vertical axis the distance travelled by
+a signal at light speed from the emitter.
+%
+Panel (a) represents the region $R$ (shaded polygon) 
+in space and time which is causally
+connected to the emitter (left vertex), following an "A" type event
+in which the node acquires the communication capability.
+%
+The sphere of first contact (SFC) is a sphere centered on the emitter,
+represented in the Figure by the left angle of the region $R$.
+%
+This sphere grows until its radius reaches the D$_{max}$ distance, in which
+it would be no longer detectable due to the decrease in the power per
+unit area, which falls under the detectability threshold.
+%
+Similarly, the surface of last contact (SLC) is another sphere that 
+grows from a
+"D" event and carries the last signal produced by the
+emitter.
+%
+The region which is causally connected to the emitter is then limited
+by these two spheres, and has the shape of a sphere (before the D
+event) or of a spherical shell (after the D event).
+%
+The temporal intervals for
+the communication between two \cetis{} are represented in he
+panel (b) of the Figure.
+%
+The receiver \ceti{} E$_1$ can listen signals from the emiter \ceti{} E$_2$,
+from the first contact event (t=C$_{12}$) up to the last contact event (t=B$_{12}$).
+%
+Similarly, 
+the receiver \ceti{} E$_2$ can listen signals from the emiter \ceti{} E$_1$,
+from t=C$_{21}$ up to t=B$_{21}$.
+
+In this scheme, the entire lifetime of a \ceti{} can then be represented
+by a polygon, limited in the spatial axis by the maximum distance of
+the signal $D_{max}$ and in the time axis by the time span between the
+starting time (A event, concave vertex) and the ending time (D event, 
+convex vertex).
+%
+This representation allows to easily visualize the important events
+that result from the existence and communications of two \cetis{} (panel
+(b)).
+%
+The points in time where the \cetis{} acquire their communicating capacity
+are dubbed ''Awakening'' events ($A_1$ and $A_2$ in the Figure).
+%
+Similarly, the points in time where the \cetis{} loss their
+communicating capacity are dubbed ''Doomsday'' events ($D_1$ and $D_2$
+in the Figure).
+%
+The points in space--time where the first contact is produced for each
+one of the \cetis{}, are defined as ''Contact'' events, shown as $C_1$
+and $C_2$ in the Figure.
+%
+Finally, the two points where the contact is lost for each one the
+\cetis{}, are denominated ''Blackout'' events, shown as $B_1$ and $B_2$
+in the Figure.
+%
+It is important to notice that there is a time delay between the
+contact events of the two \cetis{} involved in this analysis, and also a
+time delay between the two blackout events, so that the time range
+when a bidirectional contact is possible occurs between the maximum
+time of the contact points and the minimum time between the blackout
+points.
+
+%}}}
+
+%{{{
 
@@
 %
-Although a \ceti could be active for enough time so as to transmit a
+Although a \ceti{} could be active for enough time so as to transmit a
 message at large distances, the limited power of the message and the
@@
 considered as the period of time between the acquisition and loss of
-communicating capacity, each \ceti will fill a spherical shell region of
+communicating capacity, each \ceti{} will fill a spherical shell region of
 the Galaxy, limited by two concentric spherical surfaces.
@@
 The leading front, or surface of first contact (SFC) grows from the
-central \ceti until it reaches the maximum distance $D_{max}$.
+central \ceti{} until it reaches the maximum distance $D_{max}$.
 %
-Thus, the volume reached causally by a \ceti is initially a growing
+Thus, the volume reached causally by a \ceti{} is initially a growing
 sphere.
-%
+
 Following the end of the civilization, there is still a region which
@@
 The trailing front, or surface of last contact (SLC) also grows from
-the central \ceti, with a delay with respect to the SFC equivalent to
-the lifetime of the \ceti, and produces a spherical shell region.
+the central \ceti{}, with a delay with respect to the SFC equivalent to
+the lifetime of the \ceti{}, and produces a spherical shell region.
 %
-Any other \ceti within this region will be aware of the originating
-\ceti,
+Any other \ceti{} within this region will be aware of the originating
+\ceti{},
 even if it has disappeared before the time of contact.
@@
 until the surface of last contact reaches $D_{max}$, producing as a
-result the loss of all signal from the central \ceti.
+result the loss of all signal from the central \ceti{}.
 %
@@
 disk is negligible with respect to the radius of the disk, so that the
-\cetis are placed on a 2D plane disk.
+\cetis{} are placed on a 2D plane disk.
 %
@@
 regions of the spherical shells, respectively.
-%
-We illustrate in the Fig.~\ref{F_scheme} the schematic representation
-of the region causally connected to the central \ceti.
-%
-To this end, we use space--time diagrams, where time is represented on
-the horizontal axis, and space is represented in the vertical axis.
-%
-A growing surface of first contact in this diagram is represented as a
-growing circle, centered on a point in the time axis, which shifts
-towards the right at the same rate the radius of the circle increases.
-%
-The maximum distance of the message limits the radius of the circles,
-so that once this distance is reached, the circle continues its shift
-towards the right, but maintaining a constant radius.
-%
-At the end of the lifetime of a \ceti, a growing surface of last
-contact is represented by a circle that covers the inner part of the
-causal region, producing a ring of diminishing area (panel (a)).
-%
-In this scheme, the entire lifetime of a \ceti can then be represented
-by a polygon, limited in the spatial axis by the maximum distance of
-the signal $D_{max}$ and in the time axis by the time span between the
-starting time (concave vertex) and the ending time (convex vertex).
-%
-This representation allows to easily visualize the important events
-that result from the existence and communications of two \cetis (panel
-(b)).
-%
-The points in time where the \cetis acquire their communicating capacity
-(CC) are dubbed ''awakening'' events ($A_1$ and $A_2$ in the Figure).
-%
-Similarly, the points in time where the \cetis loss their
-communicating capacity are dubbed ''doomsday'' events ($D_1$ and $D_2$
-in the Figure).
-%
-The points in space--time where the first contact is produced for each
-one of the \cetis, are defined as ''contact'' events, shown as $C_1$
-and $C_2$ in the Figure.
-%
-Finally, the two points where the contact is lost for each one the
-\cetis, are denominated ''blackout'' events, shown as $B_1$ and $B_2$
-in the Figure.
-%
-It is important to notice that there is a time delay between the
-contact events of the two \cetis involved in this analysis, and also a
-time delay between the two blackout events, so that the time range
-when a bidirectional contact is possible occurs between the maximum
-time of the contact points and the minimum time between the blackout
-points.
 
@@
 %
+\ffn{2}
+%
 In the Fig.~\ref{F_messages} we show a schematic representation of the
-possible cases in which a \ceti E can be in causal contact with a
-\ceti
+possible cases in which a \ceti{} E can be in causal contact with a
+\ceti{}
 R.
@@
 factors, mainly the lifetimes, the time period between the A events 
-and the distance between the two \cetis involved in the causal
+and the distance between the two \cetis{} involved in the causal
 contact.
@@
 The parameters for the temporal distributions also determine the
-spatial properties since the density of active \cetis in the Galaxy
+spatial properties since the density of active \cetis{} in the Galaxy
 depend on these two parameters.
@@
 For example, a small $\tau_a$ and a large $\tau_s$ will produce a
-Galaxy densely populated with \cetis, and vice versa.
+Galaxy densely populated with \cetis{}, and vice versa.
 %
@@
 stationary Poisson process, and then the distribution of the times
-between the appeareance of new consecutive \cetis is exponential.
+between the appeareance of new consecutive \cetis{} is exponential.
 %
-Regarding the duration of a \ceti, we propose that the distribution of
-the duration of \cetis is a stationary exponential distribution.
+Regarding the duration of a \ceti{}, we propose that the distribution of
+the duration of \cetis{} is a stationary exponential distribution.
 %
@@
 %
-The probability for the rise of a \ceti is homogeneous over the GHZ.
+The probability for the rise of a \ceti{} is homogeneous over the GHZ.
 %
@@
 % 
-For the communication of \cetis through messages sent isotropically, we
+For the communication of \cetis{} through messages sent isotropically, we
 assume that the capacity to emit signals and to receive signals occur
@@
 %
-Another essential assumption is that all \cetis use the same signal
+Another essential assumption is that all \cetis{} use the same signal
 power, so that there is a maximum distance out to which it can be
@@
 %
-In such a system we compute probabilities of a \ceti making contact
-with another \ceti, and the results must be interpreted as for a
-randomly chosen \ceti, given certain conditions of homogeneity on an
+In such a system we compute probabilities of a \ceti{} making contact
+with another \ceti{}, and the results must be interpreted as for a
+randomly chosen \ceti{}, given certain conditions of homogeneity on an
 ensemble of equivalent galaxies.
@@
 We stress the fact that we are not centered on the probabilities for
-the Earth itself, but on a generalized \ceti.
+the Earth itself, but on a generalized \ceti{}.
 %
@@
 That is, there is no preferred location within the GHZ for the
-spontaneous appeareance of life, and the emergence of a \ceti is
-independent of the existence of previous \cetis in the galaxy.
+spontaneous appeareance of life, and the emergence of a \ceti{} is
+independent of the existence of previous \cetis{} in the galaxy.
 %            
 This is equivalent to proposing a Poisson process for the emergence of
-\cetis, since there is a close relation between the number of events
+\cetis{}, since there is a close relation between the number of events
 in time or space and the waiting time or separation, respectively
@@
 positions of stars, which are sampled randomly within the GHZ, the
-time of the awakening of each \ceti (A event), and the time of the
-vanishing of each \ceti (D event).
+time of the awakening of each \ceti{} (A event), and the time of the
+vanishing of each \ceti{} (D event).
 %
 The variables that can be deduced from the previous ones include the
-number of \cetis in casual contact with at least another \ceti at a given
-time, the number of \cetis as a function of time, the number of \cetis
-that receive a message at least one time, the number of \cetis that
+number of \cetis{} in casual contact with at least another \ceti{} at a given
+time, the number of \cetis{} as a function of time, the number of
+\cetis{}
+that receive a message at least one time, the number of \cetis{} that
 receive a message at least one time and successfully deliver an answer
@@
 
-\subsection{Model for \cetis}
+\subsection{Model for \cetis{}}
 %{{{
@@
 \begin{enumerate}
-   \item[(A)] Awakening: A new \ceti acquires the ability to emit
+   \item[(A)] Awakening: A new \ceti{} acquires the ability to emit
       and receive messages and starts to actively broadcast and
@@
    \item[(C)] Contact: A new one-way causal contact is established
-      (does not require the awareness of the involved \cetis.)
+      (does not require the awareness of the involved \cetis{}.)
    \item[(B)] Blackout: An existing causal contact is interrupted,
       with or without the occurrence of a doomsday event.
-   \item[(D)] Doomsday: An old \ceti disappears, or halts its
+   \item[(D)] Doomsday: An old \ceti{} disappears, or halts its
       communication activities.
@@
 %
-The \blackout is produced when one of the two \cetis halts in its
+The \blackout is produced when one of the two \cetis{} halts in its
 capability of emiting and receiving signals.
@@
 Similarly, a \ccontact can be produced at the time of the \aawakening
-or later, and there can be several \contacts for the same \ceti.
+or later, and there can be several \contacts for the same \ceti{}.
 
@@
 %
-The time marks for each event (A and D for each \ceti and C and B for
+The time marks for each event (A and D for each \ceti{} and C and B for
 each causal contact) are stored as a result of the simulation.
 %
-Also, the list of active \cetis is obtained as a function of time.
+Also, the list of active \cetis{} is obtained as a function of time.
 %
 Some fixed variables must be set in order to carry out the simulation:
-the size of the Galactic Habitable Zone, the mean lifetime of a \ceti,
-$\tau_s$, the mean waiting time for the appeareance of another \ceti,
+the size of the Galactic Habitable Zone, the mean lifetime of a
+\ceti{},
+$\tau_s$, the mean waiting time for the appeareance of another \ceti{},
 $\tau_a$, and the maximum distance a message can be detected.

### 2020-01-04 Sec_results.tex
preparando todo para scholar one
acomodando las figuras

--- 2019-12-24 results.tex
@@
 \centering
-\begin{tabular}{clccc}
+\begin{tabular}{cllll}
 \hline
-   \multicolumn{2}{l}{independent variables} &min&max&Nbins\\
+   \multicolumn{2}{l}{independent variables (hypothesis space)}
+   &min. value&max. value&Nbins\\
 \hline
-   $\tau_{a}$ & Mean temporal separation between consecutive awakenings & 4000 y & 200000 y & 50\\ 
-   $\tau_{s}$ & Mean lifetime of a \ceti                                 & 10000 y & 500000 y& 50\\ 
-   D$_{max}$ & Maximum reach of a message  &  \multicolumn{3}{l}{500, 1000, 10000, 20000, 40000, 80000 lyr}  \\
+   $\tau_{a}$ & Mean temporal separation between consecutive
+   awakenings & 4000 yr & 200\,000 yr & 50\\ 
+   $\tau_{s}$ & Mean lifetime of a \ceti{}
+   & 10000 yr & 500\,000 yr& 50\\ 
+   D$_{max}$ & Maximum reach of a message  &  \multicolumn{2}{l}{(500,
+   1000, 10000, 20000, 40000, 80000) lyr} & 6  \\
 \hline
-   \multicolumn{4}{l}{fixed variables and assumptions} &value \\
+   \multicolumn{2}{l}{fixed variables} & \multicolumn{2}{c}{assumptions} &value \\
 \hline
-   & statistical properties of all \cetis &equally distributed&&\\
+   & statistical properties of all \cetis{} &equally distributed&&\\
    & Point process for the distribution in time & homogeneous &&\\
@@
    & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
-   R$_{SFC}$   & Inner radius of the GHZ     &&&20000 ly\\
-   R$_{SLC}$   & Outer radius of the GHZ     &&&60000 ly\\
-   t$_{max}$ & Time span of the simulation  &&&1.e6 y\\
+   R$_{GHZ}^{min}$   & Inner radius of the GHZ     & & &20\,000 lyr\\
+   R$_{GHZ}^{max}$   & Outer radius of the GHZ     &&&60\,000 lyr\\
+   t$_{max}$ & Time span of the simulation  &&&1.e6 yr\\
+    & \multicolumn{3}{l}{number of random realizations for each point in the hypothesis space} &50\\
 \hline
-   \multicolumn{5}{l}{discrete events}\\
+   \multicolumn{2}{l}{discrete events} &
+   \multicolumn{3}{l}{affected variables}\\
 \hline
-   A & Awakening: a civilization starts a SETI program  &&&\\
-   B & Blackout: the end of the communication channel stops &&&\\
-   C & Contact: a new causal contact is produced &&&\\
-   D & Doomsday: a civilization ends its communication capabilities&&&\\
+   A event & Awakening: a \ceti{} starts its communication
+   capabilities &Number of active \cetis{}\\
+   B event & Blackout: the end of the communication channel stops
+   &Number causal channels\\
+   C event & Contact: a new causal contact is produced &Number of causal channels\\
+   D event & Doomsday: a \ceti{} ends its communication capabilities&Number of active \cetis{}\\
 \hline
@@
    \centering
-   \includegraphics[width=\textwidth]{matrix_1.pdf}
+   \includegraphics[width=\textwidth]{F_never_contact.pdf}
    \usecaption{F_never_contact}
@@
    \centering
-   \includegraphics[width=\textwidth]{matrix_2.pdf}
+   \includegraphics[width=\textwidth]{F_C_at_A.pdf}
    \usecaption{F_C_at_A}
@@
    \centering
-   \includegraphics[width=0.5\textwidth]{waiting_s+a_dif_ylog.pdf}
+   \includegraphics[width=0.5\textwidth]{F_waiting_for_1C.pdf}
    \usecaption{F_waiting_for_1C}
@@
 The parameters for the temporal aspects of the simulation (the mean
-waiting time for the next awakening $\tau_a$ and the mean lifetime
-$\tau_s$) cover the ranges 4000-200000 yr and 10000-500000 yr,
-respectively, with a regular partition of 50 values for each
-parameter.
+waiting time for the next awakening, $\left<\tau_a\right>$, and the
+mean lifetime, $\left<\tau_s\right>$) cover the ranges 4000-200000 yr
+and 10000-500000 yr, respectively, with a regular partition of 50
+values for each parameter.
 %
@@
 %
-In Table \ref{T_simu_hypotheses} we show the set of fixed parameters
-and hypotheses that define the runs of the simulations, along with the
-list of variable parameters and the range of their values that have
-been explored in the numerical experiments.
+\ttn{1}
+%
+In Table \ref{T_simu_hypotheses} we show the three variable
+parameters, the ranges of their values and the number of bins that
+have been explored in the numerical experiments.
+%
+We also show the set of parameters that take part in the simulation,
+their values and the hypotheses that define the runs of the
+simulations.
 
@@
 Some quantities are directly derived from the discrete events, namely,
-the ID of emitting and receiving \cetis, the position in the galaxy, and
+the ID of emitting and receiving \cetis{}, the position in the galaxy, and
 the times of each of the events that are relevant to keep track of the
-number of \cetis (time of awakening and time of doomsday) and of the
+number of \cetis{} (time of awakening and time of doomsday) and of the
 number of contacts (the times of contacts and the times of blackouts).
@@
 We can also derive the additional quantities that represent the
-properties of the \cetis, for example the total time elapsed between the
-awakening and the doomsday of each \ceti.
+properties of the \cetis{}, for example the total time elapsed between the
+awakening and the doomsday of each \ceti{}.
 %
-The time span of a \ceti listening another or being listened by another
+The time span of a \ceti{} listening another or being listened by another
 can also be easily derived from the results of a realization.
 %
-This way we can also compute the distribution in the galaxy of \cetis
+This way we can also compute the distribution in the galaxy of
+\cetis{}
 that reach contact and the waiting time until the first contact or the
@@
 %
-Regarding the properties of the population of \cetis and its evolution,
-the simulation yields the fraction of awaken time a \ceti is listening
-at least another \ceti (i.e., in their causal cone), the age of
-contacted \cetis at first contact, the distribution of time to wait
-until next contact, the fraction of \cetis where the first contact is
+Regarding the properties of the population of \cetis{} and its evolution,
+the simulation yields the fraction of awaken time a \ceti{} is listening
+at least another \ceti{} (i.e., in their causal cone), the age of
+contacted \cetis{} at first contact, the distribution of time to wait
+until next contact, the fraction of \cetis{} where the first contact is
 given at the awakening, the distribution of the number of contacts for
-each \ceti, the distribution of the number of contacts as a function of
-\ceti age, the number of contacts as a function of time in the galaxy,
-the rate of \cetis that succeed in contact, and the distribution of
-distances between contacted \cetis.
+each \ceti{}, the distribution of the number of contacts as a function of
+\ceti{} age, the number of contacts as a function of time in the galaxy,
+the rate of \cetis{} that succeed in contact, and the distribution of
+distances between contacted \cetis{}.
 %
@@
 It is also possible to analyze the relations between the distance to
-\ceti vs. the time of double communication, the distance to \ceti vs. the
-age of contacted \ceti, the age of a \ceti and the maximum number of
-contacted \cetis before doomsday, or the lifespan of a \ceti vs. the
+\ceti{} vs. the time of double communication, the distance to \ceti{} vs. the
+age of contacted \ceti{}, the age of a \ceti{} and the maximum number of
+contacted \cetis{} before doomsday, or the lifespan of a \ceti{} vs. the
 maximum number of contacts.
@@
 explored parameter space, and cover short/long lifetimes, short/long
-waiting times for the next \ceti to appear in the Galaxy and short/large
+waiting times for the next \ceti{} to appear in the Galaxy and short/large
 range reach of the signal, including all possible combinations.
+%
+\ttn{2}
 %
@@
 
+\ffn{3}
+%
 In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
-distributions of the number contacts for \cetis in four different
+distributions of the number contacts for \cetis{} in four different
 samples (M5, M6, M7, M8) with long range signal reach
@@
 that maximizes the number of contacts, reaching a maximum of order 100
-contacts for a single \ceti.
+contacts for a single \ceti{}.
 %
@@
 which produce a number of contacts of at most the order of ten
-contacts per \ceti.
+contacts per \ceti{}.
 %
@@
 variations of contact possibilities as a function of the position of
-the \ceti in the Galaxy (Sec.~\ref{SS_location}).
+the \ceti{} in the Galaxy (Sec.~\ref{SS_location}).
 
@@
 
-\subsection{Multiplicity and duration of contacts}\label{SS_multiplicity}
+\subsection{Membership to the network of connected \cetis{}}\label{SS_members}
+
 %{{{
-
-The duration of each contact for a single receiving \ceti is given by the
-time difference between a C event and a B event, which are direct
-outputs for each simulation experiment.
+\ffn{4, 5}
+%
+We show the 2D maps of the rate of \cetis{} that never make contact and
+the rate of \cetis{} that make contact at the moment of the awakening,
+in the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
+as a function of $\tau_s$ and $\tau_s$.
+%
+DISCUTIR LAS FIGURAS 4 Y 5 !!!
 
@@
 Such distribution can be considered to compute the probability for a
-random \ceti to carry out a SETI project and spend a given time until
+random \ceti{} to carry out a SETI project and spend a given time until
 the first contact is made.
 
-In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the \cetis
+\ffn{6}
+%
+In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
+waiting times for the first contact, for several models.
+%
+Left panel shows the histograms for several values of $\tau_a$, and
+$\tau_s$ in the range 17000-24000 yr.
+%
+Right panel shows the histograms for several values of $\tau_s$, and
+$\tau_a$ in the range 68000-100000 yr.
+%
+The shape of the histograms does not change when $\tau_a$ is modified,
+but significantly changes when $\tau_s$ is modified.
+
+In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the
+\cetis{}
 that have made at least one contact as a function of the elapsed time
@@
 Both panels show the histograms of the waiting times for the first
-contact, for all the \cetis that make at least one contact along their
+contact, for all the \cetis{} that make at least one contact along their
 lifetime.
@@
 different number of contacts on the different models.
-%
-We show the 2D maps of the rate of \cetis that never make contact and
-the rate of \cetis that make contact at the moment of the awakening,
-in the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
-as a function of $\tau_s$ and $\tau_s$.
-%
-In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
-waiting times for the first contact, for several models.
-%
-Left panel shows the histograms for several values of $\tau_a$, and
-$\tau_s$ in the range 17000-24000 yr.
-%
-Right panel shows the histograms for several values of $\tau_s$, and
-$\tau_a$ in the range 68000-100000 yr.
-%
-The shape of the histograms does not change when $\tau_a$ is modified,
-but significantly changes when $\tau_s$ is modified.
 
@@
 
-%On the Figure~\ref{F_radial_frac} we show the fraction of \cetis that
+%On the Figure~\ref{F_radial_frac} we show the fraction of \cetis{} that
 %make zero, one, two or three contacts, as a function of the
@@
 %
-For this model, the majority of \cetis do not make any contact during
+For this model, the majority of \cetis{} do not make any contact during
 all the time period in which they are active.
 %
-As expected, the fractions of \cetis making at least one contact
+As expected, the fractions of \cetis{} making at least one contact
 diminishes as the number of contacts rises.
@@
 \hline
-Model subset & D$_{max}$ & $\tau_a$ & $\tau_s$ & description  \\
+   Model subset & D$_{max}$ [lyr] & $\tau_a$ [yr] & $\tau_s$ [yr] & description  \\
 \hline

### 2020-01-04 captions.tex
acomodando las figuras

@@
 Space--time diagrams showing a schematic representation of the
-different stages in the development of a communicating civilization,
-or causal contact node (CCN).
+different stages in the development of causal contact node (\ceti{}).
 %
-Panel (a) shows the sphere of first contact (SFC) centered on an A type
-event and the surface of last contact (SLC), centered on a D type
-event.
+Panel (a) represents the region $R$ in space and time which is causally
+connected to the emitter (left vertex), following an "A" type event
+("Awakening") in which the node acquires the communication capability.
 %
-In the upper panel, the sphere that encloses the region which is
-causally connected to the emitter grows until it reaches the
-$D_{max}$ distance.  After the $D$ event, that region is limited by
-the surface of last contact.
+The sphere of first contact (SFC) is a sphere centered on the emitter
+that grows until its radius reaches the D$_{max}$ distance, in which
+the power of a signal would equal the detectability threshold.
+%
+In the Figure this sphere is represented by the left triangle of the
+region $R$.
+%
+The surface of last contact (SLC) is another sphere that grows from a
+"D" event ("Doomsday").
+%
+The region which is causally connected to the emitter is then limited
+by these two spheres, and has the shape of a sphere or of a spherical
+shell, depending on the time.
 %
 The temporal intervals for
-the communication between two civilizations are represented in he
+the communication between two \cetis{} are represented in he
 panel (b).
 %
-The receiver CCN E$_1$ can listen signals from the emiter CCN E$_2$,
-from the contact event (C$_{12}$) up to the blackout event (B$_{12}$).
+The receiver \ceti{} E$_1$ can listen signals from the emiter \ceti{} E$_2$,
+from the "Contact" event (t=C$_{12}$) up to the "Blackout" event (t=B$_{12}$).
 %
 Similarly, 
-the receiver CCN E$_2$ can listen signals from the emiter CCN E$_1$,
-from C$_{21}$ up to B$_{21}$. 
-%
-The scheme in the lower panel (b) represents two communicating
-civilizations that have causal contact in both directions
+the receiver \ceti{} E$_2$ can listen signals from the emiter \ceti{} E$_1$,
+from t=C$_{21}$ up to t=B$_{21}$. 
 %
@@
 Schematic representation on space--time diagrams of the possible cases in which a
-\ceti E (Emiter, solid line) can be in causal contact with a \ceti R
+\ceti{} E (Emiter, solid line) can be in causal contact with a \ceti{} R
 (Receiver, dashed line).
 %
-Causal contact can be produced either if the emiter \ceti appears before
-(left) or after (right) the receiver \ceti.
+Causal contact can be produced either if the emiter \ceti{} appears before
+(left) or after (right) the receiver \ceti{}.
 %

### 2020-01-08 Sec_results.tex
retocando las conclusiones

@@
 
-\ffn{3}
-%
-In Fig.~\ref{F_number_of_contacts} we show the Empirical cumulative
-distributions of the number contacts for \cetis{} in four different
-samples (M5, M6, M7, M8) with long range signal reach
-(D$_{max}$=10000). 
-%
-The lifetime is more determinant than the awakening rate in increasing
-the number of contacts.
-%
-As expected, of the four samples, the sample M7, which corresponds to
-a dense awakening in the timeline and a long lifetime, is the model
-that maximizes the number of contacts, reaching a maximum of order 100
-contacts for a single \ceti{}.
-%
-This is considerable larger than any model with a shorter lifetime,
-which produce a number of contacts of at most the order of ten
-contacts per \ceti{}.
-%
-A couple of comments are worth to mention about this plot.
-%
-First, it has a logarithmic scale on the x-axis, and second it is the
-cumulative, not differential, empirical distribution.
-
 In what follows, we focus on the analysis of the duration of contacts
@@
 %{{{
-\ffn{4, 5}
+
+\ffn{3}
 %
-We show the 2D maps of the rate of \cetis{} that never make contact and
-the rate of \cetis{} that make contact at the moment of the awakening,
-in the Figs.~\ref{F_never_contact} and \ref{F_C_at_A}, respectively,
-as a function of $\tau_s$ and $\tau_s$.
+In Fig.~\ref{F_number_of_contacts} we show the empirical cumulative
+distributions of the number contacts for \cetis{} in eight different
+samples, including short and long lifetimes
+$\left(\,\left<\tau_s\right>=\right.$10000~yr and  $\left<\tau_s\right>$=100000~yr,
+respectively$\left.\right)$, dense and sparse spatial distribution
+$\left(\,\left<\tau_a\right>=\right.$10000~yr and $\left<\tau_a\right>$= 100000~yr,
+respectively$\left.\right)$, and two different signal reach ranges (D$_{max}$=40000
+lyr and D$_{max}$=80000~lyr). 
 %
-DISCUTIR LAS FIGURAS 4 Y 5 !!!
+The areas between models with the same mean lifetime and mean
+awakening time are shaded for visualization purposes.
+%
+As it can be seen from this Figure, the mean lifetime is more
+determinant than the mean awakening rate for the number of contacts.
+%
+As expected, of the four samples with the longest $D_{max}$ the model
+with a dense awakening in the timeline (low $\left<\tau_a\right>$) and
+a long lifetime (high $\left<\tau_s\right>$), is the model that
+maximizes the number of contacts, reaching a maximum of order 20
+contacts for a single \ceti{}.
+%
+Similarly, for the models with long signal range
+($D_{max}=80000\,lyr$), the one which is dense and with long lifetime
+has the maximum number of contacts, reaching nearly 100 contacts for
+each \ceti{} in its entire lifetime.
+%
+This is considerable larger than any model with a shorter lifetime,
+which produce a number of contacts of at most the order of ten
+contacts per \ceti{}.
+%
+A couple of comments are worth to mention about this plot.
+%
+First, it has a logarithmic scale on the x-axis, and second it is the
+cumulative, not differential, empirical distribution, so that the
+differences in the number of contacts for different models is quite
+large.
+%
+Another interesting feature observed in this Figure, is that for the
+eight models analyzed, there are \cetis{} that have no causal contacts
+with any other \ceti{} during their whole lifetime.
+%
+The fraction of \cetis{} with no contact ranges from nearly 20\% up to
+nearly 90\%, except for the dense and long lifetime models, for which
+almost all \cetis{} make at least one contact.
+%
+Since these results depend on just eight models, we analyze the
+properties of the frequency of contacts as a function of both the mean
+lifetime and the mean awakening time.
+%
+\ffn{4}
+%
+On Fig.~\ref{F_never_contact} we show 2D color maps with the fraction
+of \cetis{} in the simulations that never make contact (i.e., never
+listen to another \ceti{}), as a function of the mean lifetime
+($\left<\tau_s\right>$, in the range 0--500 yr) and the mean awakening
+time ($\left<\tau_a\right>$, in the range 0--200 yr), for three
+different values of the maximum signal range, $D_{max}=10000$~yr (left
+panel), $D_{max}=40000$~yr (middle panel), and $D_{max}=80000$~yr
+(right panel).
+%
+A clear pattern emerges in this Figure, showing that the probability
+for a \ceti{} of making causal contact with at least another \ceti{}
+during their entire lifetime, increases with increasing $D_{max}$,
+increasing $\left<\tau_s\right>$ and decreasing $\left<\tau_a\right>$,
+following a roughly linear dependence with the three parameters.
+
+While the number of \cetis{} that do not succeed in reaching the
+causal contact regions of other node is a useful indicator of
+degree of isolation, there is also the chance that a number of
+\cetis{} are already in the causal contact region of other nodes,
+which shows a somewhat oposite situation.
+%
+\ffn{5}
+%
+The fraction of \cetis{} that make the first contact at the awakening
+event is shown in the Fig.~\ref{F_C_at_A}, 
+as a function of the mean lifetime
+($\left<\tau_s\right>$, in the range 0--500 yr) and the mean awakening
+time ($\left<\tau_a\right>$, in the range 0--200 yr), for three
+different values of the maximum signal range, $D_{max}=10000$~yr (left
+panel), $D_{max}=40000$~yr (middle panel), and $D_{max}=80000$~yr
+(right panel).
+%
+These probabilities are consistent with the ones on the previous
+Figure.
+%
+Although they are not complementary, it is clear that the dependence
+with the three parameters is roughly linear in both cases.
+%
+DISCUTIR ESTOS RESUOLTADOS SEGUN LA LITERATURA
+
+%
 
@@
 
-In this section we analyze the distribution of the waiting times for
+In this subsection we analyze the distribution of the waiting times for
 a first contact.
 %
-%Such distribution can be considered to compute the probability of
-%reaching a first contact within the first XX years of a SETI project,
-%assuming the methods are correct and the hypotheses of the experiment.
 Such distribution can be considered to compute the probability for a
@@
 %
-Left panel shows the histograms for several values of $\tau_a$, and
-$\tau_s$ in the range 17000-24000 yr.
+Upper panel (a) panel shows the histograms for several values of the
+mean survival time $\tau_s$, and the mean awakening time $\tau_a$ in
+the range 68000-100000 yr.
 %
-Right panel shows the histograms for several values of $\tau_s$, and
-$\tau_a$ in the range 68000-100000 yr.
+Lower panel (b) shows the histograms for several values of $\tau_a$,
+and $\tau_s$ in the range 17000-24000 yr.
 %
-The shape of the histograms does not change when $\tau_a$ is modified,
-but significantly changes when $\tau_s$ is modified.
-
-In the Fig.~\ref{F_waiting_for_1C} we show the fraction of the
-\cetis{}
-that have made at least one contact as a function of the elapsed time
-since the awakening, given that they make at least one contact on
-their lifetime, for different models.
-%
-From a frequentist approach, the normalized cumulative counts are
-related to the estimation of the probability of observing a given time
-with no success.
-%             
-Both panels show the histograms of the waiting times for the first
-contact, for all the \cetis{} that make at least one contact along their
-lifetime.
-%
-We show separately the histograms for several values of $\tau_a$ and a
-fixed value of $(\tau_s, D_{max})$, and for several values of $\tau_s$
-and a fixed value of $(\tau_a, D_{max})$.
+As it can be seen, the shape of the histograms is almost the same when
+$\tau_a$ is modified (panel (b)), but significantly changes when
+$\tau_s$ is modified (panel (a)). 
 %
 For different values of $\tau_a$, the shape of the distribution is
-similar, and the main difference given by the different number of
+similar, and the main difference is given by the different number of
 contacts on the different models.
@@
 different number of contacts on the different models.
+
+The fraction of the \cetis{} that have made at least one contact as a
+function of the elapsed time since the awakening, is computed with
+respect to the total number of nodes that make causal contact at least
+one time in the time range from the Awakening up to the Doomsday
+events.
+%
+From a frequentist approach, the total counts from the awakening
+($t=$0~yr) up to a given time are related to the estimation of the
+probability of listening during that time interval and reaching to
+causal contact region of at least another node.
+%
+The complement of this value is the probability of observing in a time
+interval with no success, i.e., without ever happening a Contact
+event.
+%
+Clearly, this probability diminishes with time and tends to zero for
+large time periods.
+%
 

### 2020-01-08 manuscript.tex
acomodando las figuras
retocando las conclusiones

@@
 accounts for the causal connections between communicating
-civilizations on an simplified galaxy.
+civilizations on a simplified galaxy.
 %
@@
 %
-We explored the consecuences of different models using mensurable
-quantities, that arise as by-products of the simulations.
-%
 With this approach, we can estimate several quantities as a function
-of the parameters on the hypothesis space.
+of the parameters on the hypothesis space,
+so we explored the outcomes of different models 
+that arise as by-products of the simulations.
+
 %
@@
 model parameters that reflect different, so far unknown, scenarios for
-the spread of intelligent life on the galaxy.
+the appeareance of intelligent life on the galaxy.
 
@@
 Our analysis is not centered in obtaining the odds for the Earth
-making contact with another extraterrestrial civilization, but in
-obtaining a statistical, parameter dependent description of the
-communication system consisting of a group of broadcasting
-civilizations which arise sparsely both in time and space in the
-Galaxy.
+to make contact with another intelligent civilization.
+%
+Instead, we focus on obtaining a statistical, parameter dependent
+description of the possible properties of the communication networks
+that comprise sets of nodes with broadcasting and reception
+capabilities.
+%
+This causally connected nodes are sparsely distributed in both space
+and time, making it difficult the analytical treatment and justifying
+the simulation approach.
 
@@
 trade--off between the simplicity and complexity of the experiment,
-future work can be performed following this framework in order to
+further analysis could be performed following this framework in order to
 explore possible implications of the results for more detailed
@@
 For example, the communication method (isotropic, colimated,
-serendipitous) can affect the observables.
+serendipitous) can affect the observables, making it neccesary to
+implement a correction factor.
 %
-Since we analyze the causal contacts, effective contacts depend on the
-efficiency of both the emitter and the receiver.
+Taking this into account, our results regarding the probability of
+causal contact should be considered as upper limits for effective
+contacts, since they depend on the
+efficiency of both the emitter and the receiver to broadcast and scan
+the sky for intelligent signals, respectively.
 %
@@
 ratio in the emission could be easily implemented in the simulation,
-although the effect would be to reduce the probability of contact by
-an efficiency ratio.
-%
-Thus, such implementation is not really necessary, since the values of
+although the effect of reducing the probability of contact
+is basically the product of the efficiency rations for emmision and
+detection and thus 
+such implementation is not really necessary
+and the values of
 probabilities could modified by a constant correction factor equal to
@@
 \bibliographystyle{mn2e}
+\setlength{\bibsep}{0.0pt}
+
 \bibliography{biblio_seti}

### 2020-01-09 Sec_methods.tex
ahhh...

@@
 
-\begin{figure*}
-   \centering
-   \includegraphics[width=0.8\textwidth]{F_messages.pdf}
-   \usecaption{F_messages} \label{F_messages}
-\end{figure*}
-
 %{{{
@@
 
+%}}}
+
+\begin{figure*}
+   \centering
+   \includegraphics[width=\textwidth]{F_messages.pdf}
+   \usecaption{F_messages} \label{F_messages}
+\end{figure*}
+
+\setlength{\tabcolsep}{10pt}
+\begin{table*}
+\centering
+\begin{tabular}{cllll}
+\hline
+   \multicolumn{2}{l}{independent variables (hypothesis space)}
+   &min. value&max. value&Nbins\\
+\hline
+   $\tau_{a}$ & Mean temporal separation between consecutive
+   awakenings & 4000 yr & 200\,000 yr & 50\\ 
+   $\tau_{s}$ & Mean lifetime of a \ceti{}
+   & 10000 yr & 500\,000 yr& 50\\ 
+   D$_{max}$ & Maximum reach of a message  &  \multicolumn{2}{l}{(500,
+   1000, 10000, 20000, 40000, 80000) lyr} & 6  \\
+\hline
+   \multicolumn{2}{l}{fixed variables} & \multicolumn{2}{c}{assumptions} &value \\
+\hline
+   & statistical properties of all \cetis{} &equally distributed&&\\
+   & Point process for the distribution in time & homogeneous &&\\
+   f$_s$ & The scan of the sky & fully efficient&&1\\
+   f$_p$ & panspermia or colonization &absent&&0\\
+   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
+   R$_{GHZ}^{min}$   & Inner radius of the GHZ     & & &20\,000 lyr\\
+   R$_{GHZ}^{max}$   & Outer radius of the GHZ     &&&60\,000 lyr\\
+   t$_{max}$ & Time span of the simulation  &&&1.e6 yr\\
+    & \multicolumn{3}{l}{number of random realizations for each point in the hypothesis space} &50\\
+\hline
+   \multicolumn{2}{l}{discrete events} &
+   \multicolumn{3}{l}{affected variables}\\
+\hline
+   A event & Awakening: a \ceti{} starts its communication
+   capabilities &Number of active \cetis{}\\
+   B event & Blackout: the end of the communication channel stops
+   &Number causal channels\\
+   C event & Contact: a new causal contact is produced &Number of causal channels\\
+   D event & Doomsday: a \ceti{} ends its communication capabilities&Number of active \cetis{}\\
+\hline
+
+\hline
+\end{tabular}
+\caption{Definition of independent variables and adopted values for 
+   fixed parameters 
+   that are part of the simulation.  Variable parameters define the
+   spatial and temporal structure
+   of the process and the maximum reach of the messages.}
+\label{T_simu_hypotheses}
+\end{table*}
+
+%{{{     
 In this scheme, the entire lifetime of a \ceti{} can then be represented
@@
 points.
-
-%}}}
-
-%{{{
 
@@
 these parameters on the results.
+%}}}
+
+\begin{figure} \centering
+   \includegraphics[width=0.5\textwidth]{F_number_of_contacts.pdf}
+   \usecaption{F_number_of_contacts} \label{F_number_of_contacts}
+\end{figure}
+
+\subsection{Complexity of the model}
+%{{{
+
+The odds of a causal contact between two \cetis{} should not be
+considered as the odds of a contact between two intelligent
+civilizations, and in fact the later could be much lesser than the
+former.
+%
+Indeed, in order to stablish a contact between
+any two entities, a minimum degree of compatibility must be
+accomplished without any previous agreement, making the posibility of
+a contact with a message that could be deciphered highly rare.
+%
+There is a
+trade--off between the simplicity and the complexity of the
+experiments, and
+further analysis could be performed following this framework in order to
+explore possible implications of the results for more detailed
+configurations.
+%
+For example, the communication method (isotropic, colimated,
+serendipitous) can affect the observables, making it neccesary to
+implement a correction factor.
+%
+Taking this into account, our results regarding the probability of
+causal contact should be considered as upper limits for effective
+contacts, since they depend on the
+efficiency of both the emitter and the receiver to broadcast and scan
+the sky for intelligent signals, respectively.
+%
+A correction by coverage ratio in the detection and by a targetting
+ratio in the emission could be easily implemented in the simulation,
+although the effect of reducing the probability of contact
+is basically the product of the efficiency rations and thus 
+such implementation is not necessary.
+%
+Therefore, the values of the
+probabilities could be modified by a constant correction factor equal to
+the combined emission/reception efficiency
+\citep{smith_broadcasting_2009, anchordoqui_upper_2019,
+forgan_collimated_2014}.
+%
+Other considerations include the effects of alignments, the use of
+stars as sources or amplifiers, the nature of the message carrier,
+the use of a spatial distribution that actually resembles the spiral shape
+and the width of the disc of the Galaxy, different distribution functions for the mean lifetime
+of \cetis{}, and different efficiencies for the sphere of the causal region
+of each \ceti{} resembling different searching strategies
+\citep{hippke_interstellar_2017}.
+%
+It is also possible to consider that D$_{max}$ is different for
+different \cetis{}.  For example a power law where powerfull emmisions
+are rare and weak emissions are common could be an improvement to the
+model.
+%
+In this work we chose not to implement this for the sake of
+simplicity.
+%
+Finally, the role of the message contents could influence on the 
+lifespan of a \ceti{} that receives a message, although the
+implementation of such behaviour would increase the number of free
+parameters and would be highly speculative.
 
@@
 %}}}
-
-\begin{figure} \centering
-   \includegraphics[width=0.5\textwidth]{F_number_of_contacts.pdf}
-   \usecaption{F_number_of_contacts} \label{F_number_of_contacts}
-\end{figure}
 

### 2020-01-09 Sec_results.tex
ahhh...

@@
+ 
 \setlength{\tabcolsep}{10pt}
@@
 \centering
-\begin{tabular}{cllll}
+\begin{tabular}{ccccc}
 \hline
-   \multicolumn{2}{l}{independent variables (hypothesis space)}
-   &min. value&max. value&Nbins\\
+   Model subset & D$_{max}$ [lyr] & $\tau_a$ interval [yr] & $\tau_s$ interval 
+   [yr]& description  \\
 \hline
-   $\tau_{a}$ & Mean temporal separation between consecutive
-   awakenings & 4000 yr & 200\,000 yr & 50\\ 
-   $\tau_{s}$ & Mean lifetime of a \ceti{}
-   & 10000 yr & 500\,000 yr& 50\\ 
-   D$_{max}$ & Maximum reach of a message  &  \multicolumn{2}{l}{(500,
-   1000, 10000, 20000, 40000, 80000) lyr} & 6  \\
-\hline
-   \multicolumn{2}{l}{fixed variables} & \multicolumn{2}{c}{assumptions} &value \\
-\hline
-   & statistical properties of all \cetis{} &equally distributed&&\\
-   & Point process for the distribution in time & homogeneous &&\\
-   f$_s$ & The scan of the sky & fully efficient&&1\\
-   f$_p$ & panspermia or colonization &absent&&0\\
-   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
-   R$_{GHZ}^{min}$   & Inner radius of the GHZ     & & &20\,000 lyr\\
-   R$_{GHZ}^{max}$   & Outer radius of the GHZ     &&&60\,000 lyr\\
-   t$_{max}$ & Time span of the simulation  &&&1.e6 yr\\
-    & \multicolumn{3}{l}{number of random realizations for each point in the hypothesis space} &50\\
-\hline
-   \multicolumn{2}{l}{discrete events} &
-   \multicolumn{3}{l}{affected variables}\\
-\hline
-   A event & Awakening: a \ceti{} starts its communication
-   capabilities &Number of active \cetis{}\\
-   B event & Blackout: the end of the communication channel stops
-   &Number causal channels\\
-   C event & Contact: a new causal contact is produced &Number of causal channels\\
-   D event & Doomsday: a \ceti{} ends its communication capabilities&Number of active \cetis{}\\
-\hline
-
+M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
+M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
+M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
+M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
+%
+M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
+M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
+M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
+M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
+%
 \hline
 \end{tabular}
-\caption{Definition of independent variables and adopted values for 
-   fixed parameters 
-   that are part of the simulation.  Variable parameters define the
-   spatial and temporal structure
-   of the process and the maximum reach of the messages.}
-\label{T_simu_hypotheses}
+\caption{Selected models to analyze the behavior of simulation outputs
+   and their dependence on simulation outputs and their dependence on
+   simulation parameters.}
+\label{T_selected_models}
 \end{table*}
@@
 In Fig.~\ref{F_number_of_contacts} we show the empirical cumulative
-distributions of the number contacts for \cetis{} in eight different
+distributions of the number contacts for \cetis{} in six different
 samples, including short and long lifetimes
-$\left(\,\left<\tau_s\right>=\right.$10000~yr and  $\left<\tau_s\right>$=100000~yr,
-respectively$\left.\right)$, dense and sparse spatial distribution
-$\left(\,\left<\tau_a\right>=\right.$10000~yr and $\left<\tau_a\right>$= 100000~yr,
-respectively$\left.\right)$, and two different signal reach ranges (D$_{max}$=40000
-lyr and D$_{max}$=80000~lyr). 
+$\left(\,\left<\tau_s\right>=\right.$10000~yr and
+$\left<\tau_s\right>$=100000~yr, respectively$\left.\right)$, dense
+and sparse spatial distribution
+$\left(\,\left<\tau_a\right>=\right.$10000~yr and
+$\left<\tau_a\right>$= 100000~yr, respectively$\left.\right)$, and two
+different signal reach ranges (D$_{max}$=40000 lyr and
+D$_{max}$=80000~lyr).
+%
+The models M1 and M5 are not shown, since in both cases the total
+number of contacts is different from zero for less than the five per
+cent of the \cetis{}.
 %
@@
 As it can be seen from this Figure, the mean lifetime is more
-determinant than the mean awakening rate for the number of contacts.
+determinant than the mean awakening rate (dense and sparse,
+represented by a different shade) for the number of contacts.
 %
-As expected, of the four samples with the longest $D_{max}$ the model
-with a dense awakening in the timeline (low $\left<\tau_a\right>$) and
-a long lifetime (high $\left<\tau_s\right>$), is the model that
+As expected, of the samples with the longest $D_{max}$ the model with
+a dense awakening in the timeline (low $\left<\tau_a\right>$) and a
+long lifetime (high $\left<\tau_s\right>$), is the model that
 maximizes the number of contacts, reaching a maximum of order 20
@@
 %
-This is considerable larger than any model with a shorter lifetime,
+This is considerably larger than any model with a shorter lifetime,
 which produce a number of contacts of at most the order of ten
@@
 %
-A couple of comments are worth to mention about this plot.
-%
-First, it has a logarithmic scale on the x-axis, and second it is the
-cumulative, not differential, empirical distribution, so that the
-differences in the number of contacts for different models is quite
-large.
+A couple of comments are worth to mention about this plot, namely, it
+has a logarithmic scale on the x-axis, and it is the cumulative, not
+differential, empirical distribution, so that the differences in the
+number of contacts for different models is quite large.
 %
 Another interesting feature observed in this Figure, is that for the
-eight models analyzed, there are \cetis{} that have no causal contacts
-with any other \ceti{} during their whole lifetime.
+models analyzed, there are \cetis{} that have no causal contacts with
+any other \ceti{} during their whole lifetime.
 %
 The fraction of \cetis{} with no contact ranges from nearly 20\% up to
-nearly 90\%, except for the dense and long lifetime models, for which
-almost all \cetis{} make at least one contact.
+nearly 90\% (models M1 and M5, not shown), except for the dense and
+long lifetime models, for which almost all \cetis{} make at least one
+contact.
 %
@@
 %
-DISCUTIR ESTOS RESUOLTADOS SEGUN LA LITERATURA
+%% DISCUTIR ESTOS RESUOLTADOS SEGUN LA LITERATURA
 
@@
 %             
+The diferential probability of making a causal contact reach its
+maximum at the awakening event for all models.
+%
+That is, for a short period of time like the nearly 50 years SETI
+programs have been active, the initial moment is the most promising
+for making a (causal) contact, at least for a given technology.
+%
+This consideration offers a new approach for SETI programs, where the
+search for new communication technologies has a fundamental role.
+%
+Moreover, as it can be seen in the Fig.~\ref{F_waiting_for_1C}, there
+is a maximum probability of making the first causal contact at nearly
+100000 years.
+%
+Remarkably, the probability of such a contact does not grows linearly
+with the observing time, but is greater during the first 100
+millennia.
 
@@
 
-\subsection{Location within the Galaxy}\label{SS_location}
-%{{{
-
-%On the Figure~\ref{F_radial_frac} we show the fraction of \cetis{} that
-%make zero, one, two or three contacts, as a function of the
-%galactocentric distance, for model M1.  VERIFICAR!  M1?
+%\subsection{Location within the Galaxy}\label{SS_location}
+%%{{{
 %    
-For this model, the majority of \cetis{} do not make any contact during
-all the time period in which they are active.
+%%On the Figure~\ref{F_radial_frac} we show the fraction of \cetis{} that
+%%make zero, one, two or three contacts, as a function of the
+%%galactocentric distance, for model M1.  VERIFICAR!  M1?
+%%
+%For this model, the majority of \cetis{} do not make any contact during
+%all the time period in which they are active.
+%%
+%As expected, the fractions of \cetis{} making at least one contact
+%diminishes as the number of contacts rises.
+%%
+%The shape of this distribution is a result of the geometry of the GHZ.
+%%
+%In general, a closer position to the center of the Galaxy slightly
+%favours the possibility of a contact.
+%%
+%This would indicate that the position of the Earth is suitable for the
+%odds of making a contact.
 %
-As expected, the fractions of \cetis{} making at least one contact
-diminishes as the number of contacts rises.
+%%}}}
 %
-The shape of this distribution is a result of the geometry of the GHZ.
 %
-In general, a closer position to the center of the Galaxy slightly
-favours the possibility of a contact.
-%
-This would indicate that the position of the Earth is suitable for the
-odds of making a contact.
-
-%}}}
-
-\setlength{\tabcolsep}{10pt}
-\begin{table*}
-\centering
-\begin{tabular}{ccccc}
-\hline
-   Model subset & D$_{max}$ [lyr] & $\tau_a$ [yr] & $\tau_s$ [yr] & description  \\
-\hline
-M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
-M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
-M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
-M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
-%
-M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
-M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
-M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
-M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
-%
-\hline
-\end{tabular}
-\caption{Selected models to analyze the behavior of simulation outputs
-   and their dependence on simulation outputs and their dependence on
-   simulation parameters.}
-\label{T_selected_models}
-\end{table*}

### 2020-01-09 manuscript.tex
ahhh...

@@
 Most efforts on the estimation of a number of intelligent
-civilizations are centered on the Drake equation, although 
-its factors are affected by large uncertainties, and it lacks a
+   civilizations are centered on the Drake equation, although its
+   factors are affected by large uncertainties, and it lacks a
    temporal nature.
 %
-A alternative approach uses detailed numerical simulations of the galaxy
-and recipes for the rates of the formation of stars and planets, and
-even for the origin of life.
+An alternative approach uses detailed numerical simulations of the
+   galaxy and recipes for the rates of the formation of stars and
+   planets, and even for the origin of life.
 %
-Here we present numerical simulations of stochastic processes of
-emergence of civilizations with communication capabilities, based
-on minimal asumptions, and analyze the causal connections among them.
+These two opposite approaches rely on a large number of highly
+   uncertain parameters.
+%
+Here we present numerical simulations of stochastic processes for the
+   emergence of nodes with communication capabilities and analyze the
+   causal connections among them.
+%
+This is based on minimal assumptions and three free parameters, with a
+   focus on the statistical properties of empirical population models
+   instead of an interpretable model with variables to be determined
+   by observation.
 %
 The analysis of the rate of causal contacts as a function of the mean
-number of civilizations, the mean lifetime span distribution and the
+   number of civilizations, the mean lifespan distribution and the
    maximum distance a civilization can send signals is presented, and
@@
 Additionally, the odds to make a contact in a few years of monitoring,
-assuming a perfect detection rate, are low for most models, with the
-exception of models which propose a galaxy densely populated with
-long lived civilizations.
+   assuming a perfect detection rate, are low for most models, with
+   the exception of models which propose a galaxy densely populated
+   with long lived civilizations.
 %
-The probability of causal contacts increases with the 
-mean lifetime of civilizations much more signifficantly than with
-the mean number of active civilizations for a given
-time window.
+The probability of causal contacts increases with the mean lifetime of
+   civilizations much more signifficantly than with the mean number of
+   active civilizations for a given time window.
 %
+Finally, we show that maximum probability of making a contact occurrs
+when a civilization acquires the required communication technology.
 \end{abstract}
@@
 that arise as by-products of the simulations.
-
 %
@@
 contact is extremely unlikely unless the Galaxy is heavily populated
-by intelligent civilizations.
+by intelligent civilizations with large average lifetimes.
 %
-Even so, we must consider that in order to stablish a contact between
-any two entities, a minimum degree of compatibility must be
-accomplished without any previous agreement, making the posibility of
-a contact with a message that could be deciphered highly rare.
+Roughly, in order to have at least one contact in the entire lifetime of the
+\ceti{}, there should appear a mean of at least one \ceti{} every $\sim$20000 yr, with
+a mean active period of at least $\sim$30000 yr.
+%
+There is a balance between the density of \cetis{} and the mean
+lifetime since, as expected, a lower density can be compensated by a
+longer active time period.
+%
+However, a large number of nodes does not easily compensate their
+short lives 
+to reach the same probability of causal contact than in the case of a
+less populated Galaxy but with very ancient civilizations.
+%
+In all cases, for a short period of time (for instance, the time SETI
+programs have been active on Earth), the maximum probability of making
+a contact occurs at the moment of the awakening.
+%
+This suggests the possibility that an alternative SETI strategy could be
+the search for alternative message carriers, for the case in which the
+search has not been performed on the adecuate channels.
+%
+Then, if a contact is produced for the first time, the origin of the
+signal is more likely to be very old.
+%
+Also, the chances of entering the causal connected zone of a \ceti{}
+does not grow linearly, but favours the first period of 100000 years.
 
-Taking into account the caveat mentioned previously about the
-trade--off between the simplicity and complexity of the experiment,
-further analysis could be performed following this framework in order to
-explore possible implications of the results for more detailed
-configurations.
+In the approach of this work, we used computer simulations to address
+the problem of the probabilities of causal contacts between location
+in the Galaxy the posibility of sending and receiving messages.
 %
-For example, the communication method (isotropic, colimated,
-serendipitous) can affect the observables, making it neccesary to
-implement a correction factor.
+Instead of making a number of assumptions, we have explored the
+hypothesis space, reducing the problem to only three parameters and a
+few simple hypothesis to perform a complete model for the population and
+communication network in the Galaxy.
 %
-Taking this into account, our results regarding the probability of
-causal contact should be considered as upper limits for effective
-contacts, since they depend on the
-efficiency of both the emitter and the receiver to broadcast and scan
-the sky for intelligent signals, respectively.
-%
-A correction by coverage ratio in the detection and by a targetting
-ratio in the emission could be easily implemented in the simulation,
-although the effect of reducing the probability of contact
-is basically the product of the efficiency rations for emmision and
-detection and thus 
-such implementation is not really necessary
-and the values of
-probabilities could modified by a constant correction factor equal to
-the combined emission/reception efficiency
-\citep{smith_broadcasting_2009, anchordoqui_upper_2019,
-forgan_collimated_2014}.
-%
-If we consider a statistical distribution for this factor, then the
-results could change, and this could be studied with the DE approach.
-%
-Also the nature of the message carrier, electromagnetic or another.
-%
-It is also possible to explore the effects of alignments or the use of
-stars as sources or amplifiers.
-%
-%energetic markers Spiritual markers
-%
-As a future work, we will explore several improvements to the
-simulation, and analyze the impact on the final results of these
-changes.
-%
-They include a spatial distribution that actually resembles the shape
-of the Galaxy, different distribution functions for the mean lifetime
-of MPLs and different efficiencies for the sphere of the causal region
-of each MPL, resembling different searching strategies
-\citep{hippke_interstellar_2017}.
-
-Possible hypothesis improvements are:
-
-\begin{itemize}
-   \item D$_{max}$ is different for different MPLs.  For example a power
-      law where powerfull emmisions are rare and weak emssions are
-      common.
-   \item The probability of the rising of a new MPL vary within the GHZ.
-   \item Correction by coverage ratio in the detection and by
-      targetting ratio in the emission.
-   \item Explore the role of the message
-      contents, influence on the lifespan of a MPL that receives a
-      message, mean lifespan of a MPL that emits a message.
-\end{itemize}
+This allows to consider the Fermi paradox from a new pespective, and
+to propose an alterntave treatment for the number of intelligent
+emitter/receivers with emphasis on the temporal dimension which limits
+the probabilities of contacts due to the short time interval for
+rise and fall of civilizations compared to the age and extension of
+our Galaxy.
 

### 2020-01-11 Sec_methods.tex
pre correcciones Jose

@@
 This method is well suited for the particular case of the diffusion of
-intelligent signals in the Galaxy, and allows to explore several
+intelligent signals in the galaxy, and allows to explore several
 models easily.
@@
 time that have a causal connection
-at light speed, hereafter dubbed ''causal contact nodes'' (\ceti{}).
+at light speed, hereafter dubbed ''causal contact nodes'' (hereafter, \ceti{}).
 %
@@
 terms of three independent parameters, namely, the mean time span
-between the appeareance of consecutive \cetis{}, $\left<\tau_a\right>$, their mean
+between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$, their mean
 lifetime $\left<\tau_s\right>$, and the maximum distance a signal can be
@@
 Sotos_biotechnology_2019}, related to the time span between the
-appeareance of consecutive \cetis{} (since when $\left<\tau_a\right>$ is shorter, it
+appearance of consecutive \cetis{} (since when $\left<\tau_a\right>$ is shorter, it
 produces a greater density of \cetis{}).
@@
 
-\ffn{1}
-We illustrate in the Fig.~\ref{F_scheme} the schematic representation
-of the region causally connected to the central \ceti{}.
+\ffn{1} We illustrate in the Fig.~\ref{F_scheme} the schematic
+representation of the region causally connected to the central
+\ceti{}.
 %
@@
 %
-In particular, we plot on the vertical axis the distance travelled by
+In particular, we plot on the vertical axis the distance traveled by
 a signal at light speed from the emitter.
 %
-Panel (a) represents the region $R$ (shaded polygon) 
-in space and time which is causally
-connected to the emitter (left vertex), following an "A" type event
-in which the node acquires the communication capability.
+Panel (a) represents the region $R$ (shaded polygon) in space and time
+which is causally connected to the emitter (left vertex), following an
+"A" type event in which the node acquires the communication
+capability.
 %
@@
 %
-This sphere grows until its radius reaches the D$_{max}$ distance, in which
-it would be no longer detectable due to the decrease in the power per
-unit area, which falls under the detectability threshold.
+This sphere grows until its radius reaches the D$_{max}$ distance, in
+which it would be no longer detectable due to the decrease in the
+power per unit area, which falls under the detectability threshold.
 %
 Similarly, the surface of last contact (SLC) is another sphere that
-grows from a
-"D" event and carries the last signal produced by the
+grows from a "D" event and carries the last signal produced by the
 emitter.
@@
 %
-The temporal intervals for
-the communication between two \cetis{} are represented in he
-panel (b) of the Figure.
+The temporal intervals for the communication between two \cetis{} are
+represented in the panel (b) of the Figure.
 %
-The receiver \ceti{} E$_1$ can listen signals from the emiter \ceti{} E$_2$,
-from the first contact event (t=C$_{12}$) up to the last contact event (t=B$_{12}$).
+The receiver \ceti{} E$_1$ can listen signals from the emitter \ceti{}
+E$_2$, from the first contact event (t=C$_{12}$) up to the last
+contact event (t=B$_{12}$).
 %
-Similarly, 
-the receiver \ceti{} E$_2$ can listen signals from the emiter \ceti{} E$_1$,
-from t=C$_{21}$ up to t=B$_{21}$.
+Similarly, the receiver \ceti{} E$_2$ can listen signals from the
+emitter \ceti{} E$_1$, from t=C$_{21}$ up to t=B$_{21}$.
 
@@
 %{{{     
-In this scheme, the entire lifetime of a \ceti{} can then be represented
-by a polygon, limited in the spatial axis by the maximum distance of
-the signal $D_{max}$ and in the time axis by the time span between the
-starting time (A event, concave vertex) and the ending time (D event, 
-convex vertex).
+In this scheme, the entire lifetime of a \ceti{} can then be
+represented by a polygon, limited in the spatial axis by the maximum
+distance of the signal $D_{max}$ and in the time axis by the time span
+between the starting time (A event, concave vertex) and the ending
+time (D event, convex vertex).
 %
 This representation allows to easily visualize the important events
-that result from the existence and communications of two \cetis{} (panel
-(b)).
+that result from the existence and communications of two \cetis{}
+(panel (b)).
 %
-The points in time where the \cetis{} acquire their communicating capacity
-are dubbed ''Awakening'' events ($A_1$ and $A_2$ in the Figure).
+The points in time where the \cetis{} acquire their communicating
+capacity are dubbed ''Awakening'' events ($A_1$ and $A_2$ in the
+Figure).
 %
@@
 Finally, the two points where the contact is lost for each one the
-\cetis{}, are denominated ''Blackout'' events, shown as $B_1$ and $B_2$
-in the Figure.
+\cetis{}, are denominated ''Blackout'' events, shown as $B_1$ and
+$B_2$ in the Figure.
 %
 It is important to notice that there is a time delay between the
-contact events of the two \cetis{} involved in this analysis, and also a
-time delay between the two blackout events, so that the time range
+contact events of the two \cetis{} involved in this analysis, and also
+a time delay between the two blackout events, so that the time range
 when a bidirectional contact is possible occurs between the maximum
@@
 %
-As a consecuence of this limitation and of their finite lifetime,
+As a consequence of this limitation and of their finite lifetime,
 considered as the period of time between the acquisition and loss of
-communicating capacity, each \ceti{} will fill a spherical shell region of
-the Galaxy, limited by two concentric spherical surfaces.
+communicating capacity, each \ceti{} will fill a spherical shell
+region of the galaxy, limited by two concentric spherical surfaces.
 %
@@
 The trailing front, or surface of last contact (SLC) also grows from
-the central \ceti{}, with a delay with respect to the SFC equivalent to
-the lifetime of the \ceti{}, and produces a spherical shell region.
+the central \ceti{}, with a delay with respect to the SFC equivalent
+to the lifetime of the \ceti{}, and produces a spherical shell region.
 %
 Any other \ceti{} within this region will be aware of the originating
-\ceti{},
-even if it has disappeared before the time of contact.
+\ceti{}, even if it has disappeared before the time of contact.
 %
@@
 %
-In our approach, we consider a model Galaxy where the width of the
+In our approach, we consider a model galaxy where the width of the
 disk is negligible with respect to the radius of the disk, so that the
@@
 
-The initially growing communicating sphere, is shown over
-space--time diagrams, where time is represented on the horizontal
-direction, and space is represented in the vertical direction.
+The initially growing communicating sphere, is shown over space--time
+diagrams, where time is represented on the horizontal direction, and
+space is represented in the vertical direction.
 %
@@
 possible cases in which a \ceti{} E can be in causal contact with a
-\ceti{}
-R.
+\ceti{} R.
 %
@@
 The parameters for the temporal distributions also determine the
-spatial properties since the density of active \cetis{} in the Galaxy
+spatial properties since the density of active \cetis{} in the galaxy
 depend on these two parameters.
@@
 For example, a small $\tau_a$ and a large $\tau_s$ will produce a
-Galaxy densely populated with \cetis{}, and vice versa.
+galaxy densely populated with \cetis{}, and vice versa.
 %
@@
 stationary Poisson process, and then the distribution of the times
-between the appeareance of new consecutive \cetis{} is exponential.
+between the appearance of new consecutive \cetis{} is exponential.
 %
-Regarding the duration of a \ceti{}, we propose that the distribution of
-the duration of \cetis{} is a stationary exponential distribution.
+Regarding the duration of a \ceti{}, we propose that the distribution
+of the duration of \cetis{} is a stationary exponential distribution.
 %
@@
 In such a case we could expect that the interpretation of our results
-could be considered as pessimist for the establishment of a contact with
-our planet.
+could be considered as pessimist for the establishment of a contact
+with our planet.
 %
-We also limit the posibilities of life or other types of civilizations
+We also limit the possibilities of life or other types of civilizations
 to the hypothesis usually stated for the definition of the GHZ
@@
 in severe conditions or unstable systems, which would prevent the
-appeareance of life as we know it.
+appearance of life as we know it.
 
@@
 % 
-For the communication of \cetis{} through messages sent isotropically, we
-assume that the capacity to emit signals and to receive signals occur
-at the same time.
+For the communication of \cetis{} through messages sent isotropically,
+we assume that the capacity to emit signals and to receive signals
+occur at the same time.
 %
@@
 It is worth mentioning that we are considering in our experimental
-setup a system composed by several emiters and receivers across the
-Galaxy, under the hypothesis proposed before.
+setup a system composed by several emitters and receivers across the
+galaxy, under the hypothesis proposed before.
 %
@@
 sent to a specific direction.
-
+%
 %https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
@@
 %
-This proposal is based on a pronoid scenario (REF), and we do not
-consider in this work the occurrence of paranoid or partially paranoid
-civilizations.
-%
-If that would be the case, the results obtained here can be taken as
-upper limits.
+%This proposal is based on a pronoid scenario, and we do not consider
+%in this work the occurrence of paranoid or partially paranoid
+%civilizations.
+%%
+%If that would be the case, the results obtained here can be taken as
+%upper limits.
 
@@
 efficiency of communication or detection methods.
-%
-%Also, a mixed paranoid/pronoid population of intelligent civilizations
-%could be implemented easily.
-%
-% EXPLICARLO O SACARLO
-% O PONER EN LA DISCUSION
 %
@@
 %
-The power law behaviour has been observed in a variety of systems,
+The power law behavior has been observed in a variety of systems,
 including for example stock market fluctuations, sizes of computer
@@
 times is justified by considering the hypothesis that the process of
-appeareance of life in the galaxy is homogeneous and stationary.
+appearance of life in the galaxy is homogeneous and stationary.
 %
 That is, there is no preferred location within the GHZ for the
-spontaneous appeareance of life, and the emergence of a \ceti{} is
+spontaneous appearance of life, and the emergence of a \ceti{} is
 independent of the existence of previous \cetis{} in the galaxy.
@@
 process, a Poisson distribution for the number of events implies an
-exponential distribution for their separations, and viceversa.
+exponential distribution for their separations, and vice versa.
 %
@@
 %
-Indeed, in order to stablish a contact between
-any two entities, a minimum degree of compatibility must be
-accomplished without any previous agreement, making the posibility of
-a contact with a message that could be deciphered highly rare.
+Indeed, in order to establish a contact between any two entities, a
+minimum degree of compatibility must be accomplished without any
+previous agreement, making the possibility of a contact with a message
+that could be deciphered highly rare.
 %
-There is a
-trade--off between the simplicity and the complexity of the
-experiments, and
-further analysis could be performed following this framework in order to
-explore possible implications of the results for more detailed
-configurations.
+There is a trade--off between the simplicity and the complexity of the
+experiments, and further analysis could be performed following this
+framework in order to explore possible implications of the results for
+more detailed configurations.
 %
-For example, the communication method (isotropic, colimated,
-serendipitous) can affect the observables, making it neccesary to
+For example, the communication method (isotropic, collimated,
+serendipitous) can affect the observables, making it necessary to
 implement a correction factor.
@@
 causal contact should be considered as upper limits for effective
-contacts, since they depend on the
-efficiency of both the emitter and the receiver to broadcast and scan
-the sky for intelligent signals, respectively.
+contacts, since they depend on the efficiency of both the emitter and
+the receiver to broadcast and scan the sky for intelligent signals,
+respectively.
 %
-A correction by coverage ratio in the detection and by a targetting
+A correction by coverage ratio in the detection and by a targeting
 ratio in the emission could be easily implemented in the simulation,
-although the effect of reducing the probability of contact
-is basically the product of the efficiency rations and thus 
-such implementation is not necessary.
+although the effect of reducing the probability of contact is
+basically the product of the efficiency rations and thus such
+implementation is not necessary.
 %
-Therefore, the values of the
-probabilities could be modified by a constant correction factor equal to
-the combined emission/reception efficiency
-\citep{smith_broadcasting_2009, anchordoqui_upper_2019,
+Therefore, the values of the probabilities could be modified by a
+constant correction factor equal to the combined emission/reception
+efficiency \citep{smith_broadcasting_2009, anchordoqui_upper_2019,
 forgan_collimated_2014}.
@@
 Other considerations include the effects of alignments, the use of
-stars as sources or amplifiers, the nature of the message carrier,
-the use of a spatial distribution that actually resembles the spiral shape
-and the width of the disc of the Galaxy, different distribution functions for the mean lifetime
-of \cetis{}, and different efficiencies for the sphere of the causal region
-of each \ceti{} resembling different searching strategies
+stars as sources or amplifiers, the nature of the message carrier, the
+use of a spatial distribution that actually resembles the spiral shape
+and the width of the disc of the Galaxy, different distribution
+functions for the mean lifetime of \cetis{}, and different
+efficiencies for the sphere of the causal region of each \ceti{}
+resembling different searching strategies
 \citep{hippke_interstellar_2017}.
@@
 It is also possible to consider that D$_{max}$ is different for
-different \cetis{}.  For example a power law where powerfull emmisions
+different \cetis{}.  For example a power law where powerful emissions
 are rare and weak emissions are common could be an improvement to the
@@
 lifespan of a \ceti{} that receives a message, although the
-implementation of such behaviour would increase the number of free
+implementation of such behaviur would increase the number of free
 parameters and would be highly speculative.
@@
 The \blackout is produced when one of the two \cetis{} halts in its
-capability of emiting and receiving signals.
+capability of emitting and receiving signals.
 %
@@
 \ceti{},
-$\tau_s$, the mean waiting time for the appeareance of another \ceti{},
+$\tau_s$, the mean waiting time for the appearance of another \ceti{},
 $\tau_a$, and the maximum distance a message can be detected.

### 2020-01-11 Sec_motivations.tex
ahhh...
pre correcciones Jose

--- 2019-12-24 motivations.tex
@@
 %
-For a comprehensive review and analysis of each term of the equation
-see \citet{vakoch_drake_2015}.
+A comprehensive review and analysis of each term of the equation
+is presented in \citet{vakoch_drake_2015}.
 %
@@
 %%%%% agregar Carroll-Nellenback, 2019).
+\Fpagebreak
 %
-However, the uncertainties in the factors of the Drake equation make
+The Drake equation is a key tool
+to organize the discussion around the problem of the abundance
+of civilizations 
+in the Galaxy \citep{hinkel_interdisciplinary_2019}.
+%
+However, the uncertainties in the factors of this equation make
 it less prone to a formal application in order to define searching
 strategies or to compute the actual number of extraterrestrial
-intelligences, but is a key tool
-to organize the discussion around the problem of their abundance
-in the Galaxy \citep{hinkel_interdisciplinary_2019}, which is
-noticeable dominated by the fact that despite decades of effort not a
-single extraterrestrial intelligence, if any, has been detected so far.
+intelligences.
 %
+The optimistic estimation from the Drake equation contrast with
+the 
+SETI initiatives,
+noticeable overshadowed by the fact that despite decades of effort, not a
+single extraterrestrial intelligence, if any, has been detected.
+
 Some modifications to the original idea of the equation have tried to
@@
 equation.
-
-\Fpagebreak
-
+%
 Temporal aspects of the distribution of communicating civilizations
@@
 The discrete events method for simulating a stochastic process is an
-approximation that allows to study the behaviour of complex systems,
+approximation that allows to study the behavior of complex systems,
 by considering a sequence of well defined discrete events.
@@
 The simulation is carried out by following all the variables that
-describe the system, that constitute the state of the system.
-%
-The evolution of the process, then, is described as a set of changes
-in the state of the system.
+describe and constitute the state of the system, and the evolution of
+the process is described as a set of changes in that state.
 %
@@
 %
-The process involves following the changes on the state of the system,
-defining the initial and final states, defining a method that allows to
-keep track of time progress, and maintaining a list of relevant
-events. (REESCRIBIR!!)
+For example, when a new contact is produced between two entities in
+the simulated galaxy, the numbers of active communication lines and of
+active communicated nodes change.
+%
+The process involves then following the changes on the state of the
+system, defining the initial and final states, defining a method that
+allows to keep track of the time progress in steps, and maintaining a
+list of relevant events.
 
@@
 structure in the Drake equation.
-
-%While the Drake equation remains a very useful
-%guide in identifying and analyzing the various factors in-
-%volved in the problem and their relative probabilities, our
-%approach could provide a better framework with which to
-%perform a statistical analysis when taking into account evo-
-%lutionary processes. In particular, we have shown that the
-%Drake equation can misestimate the number of detectable
-%communicating species if spatial and temporal dependencies
-%are important. This should suggest some caution in the un-
-%critical application of such an equation. In this paper, we offer
-%an exemplificative illustration of the effect of different
-%probability distributions for the typical timescales involved
-%in the problem. Ideally, it would be desirable to arrive at a
-%mathematical description of such distributions based on a
-%modeling of the underlying astrophysical and biological
-%processes. This points out a possible future direction of in-
-%vestigation in which an integrated evolutionary (and coevo-
-%lutionary) approach to the problem would be a central focus
-%of interest (for more on this perspective, see, e.g., Chaisson
-
+%
 \citet{bloetscher_using_2019} uses a probabilistic approach, though
@@
 important since the time it takes the light to travel across the
-Galaxy is much lesser than the lifetime of the Galaxy.
+Galaxy is much lesser than its lifetime.
 %
@@
 the distance to the Earth, the time of appearance and the lifespan of
-the communicating civilization. Each of these three variables are
-drawn from a random distribution. The distribution of the distances
-results from a uniform distribution of civilization within the plane
-of the Galaxy. For the distribution of the characteristic time of
-appearance, the author explores and exponential and a truncated
-Gaussian distributions. For the distribution of the lifespans, the
-author chooses an exponential distribution. The radius of influence is
-set to 1000 ly.
+the communicating civilization.
+%
+Each of these three variables are drawn from a random distribution.
+The distribution of the distances results from a uniform distribution
+of civilization within the plane of the Galaxy.
+%
+For the distribution of the characteristic time of appearance, the
+author explores and exponential and a truncated Gaussian
+distributions. For the distribution of the lifespans, the author
+chooses an exponential distribution.
+%
+The radius of influence is set to 1000 lyr.
 %
@@
 evolve at discrete time steps, according to transition rules that take
-into account for each cell the states of its neighbour cells.
+into account for each cell the states of its neighbor cells.
 %
@@
 civilizations.
-
-%- - - -
 

### 2020-01-11 Sec_results.tex
pre correcciones Jose

@@
 \caption{Selected models to analyze the behavior of simulation outputs
-   and their dependence on simulation outputs and their dependence on
-   simulation parameters.}
+   and their dependence on simulation parameters.}
 \label{T_selected_models}
@@
 explored parameter space, and cover short/long lifetimes, short/long
-waiting times for the next \ceti{} to appear in the Galaxy and short/large
+waiting times for the next \ceti{} to appear in the galaxy and short/large
 range reach of the signal, including all possible combinations.
@@
 variations of contact possibilities as a function of the position of
-the \ceti{} in the Galaxy (Sec.~\ref{SS_location}).
+the \ceti{} in the galaxy (Sec.~\ref{SS_location}).
 
@@
 \cetis{} are already in the causal contact region of other nodes,
-which shows a somewhat oposite situation.
+which shows a somewhat opposite situation.
 %

### 2020-01-11 manuscript.tex
pre correcciones Jose

@@
 The probability of causal contacts increases with the mean lifetime of
-   civilizations much more signifficantly than with the mean number of
+   civilizations much more significantly than with the mean number of
    active civilizations for a given time window.
 %
-Finally, we show that maximum probability of making a contact occurrs
+Finally, we show that maximum probability of making a contact occurs
 when a civilization acquires the required communication technology.
@@
 %%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
-\section{Discusion}\label{S_discussion}
+\section{Discussion}\label{S_discussion}
 %{{{
@@
 model parameters that reflect different, so far unknown, scenarios for
-the appeareance of intelligent life on the galaxy.
+the appearance of intelligent life on the galaxy.
 
@@
 %
-Instead, we take advantadge of the simplicity of the model to explore
+Instead, we take advantage of the simplicity of the model to explore
 the hypothesis space, i.e., the three--dimensional parameter space of
-the parametric model, in order to gain insight on the consecuences of
+the parametric model, in order to gain insight on the consequences of
 different scenarios for the search of intelligent life.
@@
 the search for alternative message carriers, for the case in which the
-search has not been performed on the adecuate channels.
+search has not been performed on the adequate channels.
 %
@@
 the problem of the probabilities of causal contacts between location
-in the Galaxy the posibility of sending and receiving messages.
+in the Galaxy the possibility of sending and receiving messages.
 %
@@
 %
-This allows to consider the Fermi paradox from a new pespective, and
-to propose an alterntave treatment for the number of intelligent
+This allows to consider the Fermi paradox from a new perspective, and
+to propose an alternative treatment for the number of intelligent
 emitter/receivers with emphasis on the temporal dimension which limits

### 2020-01-13 Sec_methods.tex
ultimos retoques

@@
 %
-Each event produces changes in the values of the variables that define
-the process, and thus the corresponding change in the state of the
+Each event yields modifications to the values of the variables that define
+the process, producing the corresponding change in the state of the
 simulated system.
@@
 We simulate the statistical properties of a set of points in space and
-time that have a causal connection
-at light speed, hereafter dubbed ''causal contact nodes'' (hereafter, \ceti{}).
-%
-% ability of emitting and receiving messages at
+time that have a causal connection at light speed, hereafter dubbed
+''Causal Contact Nodes'' (hereafter, \ceti{}).
 %
 We chose this generic name in order to stress the fact that in this
-analysis no broadcasting or lookout activities are required, 
-and therefore not actual messages are considered, only the causal
-contact.
+analysis no broadcasting or lookout activities are required, and
+therefore not actual messages are considered, only the causal contact.
 %
-For the special case of a fully efficient node that emits
-and receives isotropically, a causal contact node can be considered as
-a broadcasting station that has the ability to detect signals through
-an active lookout program.
+For the special case of a fully efficient node that emits and receives
+isotropically, a causal contact node can be considered as a
+broadcasting station that has the ability to detect signals through an
+active lookout program.
 %
@@
 %
-It can be associated with a planet where life has developed,
-became intelligent, reached the skills required to find the right
+For instance, it can be associated with a planet where life has developed, became
+intelligent, reached the skills required to find the right
 communication channel, sustain a search and establish a contact.
@@
 example interstellar beacons where life has ceased to exist but
-continue with its emission or communication stations established by
+continue with its emission, or communication stations established by
 probes or left intentionally by intelligent beings \citep[see, e.g.,
@@
 %
-However, we do not consider this highly speculative alternatives at
+However, we do not consider these highly speculative alternatives at
 this point.
-%
-% have been proposed
 %
@@
 %
-The system is defined by a number of actors that represent \cetis{} and
-appear at different instants in time, generating events that produce
-meaningful changes in the variables that describe the system, i.e., in
-the arrangement of \cetis{} and their network of causal contacts.
+The system is defined by a number of actors that represent \cetis{}
+and appear at different instants in time, generating events that
+produce meaningful changes in the variables that describe the system,
+i.e., in the arrangement of \cetis{} and their network of causal
+contacts.
 %
-For example, the appearance of a new \ceti{} in a region which is filled
-with a signal emitted by another \ceti{}, will increase the number of
-active \cetis{} and the number of pairs of \cetis{} in causal contact.
+For example, the appearance of a new \ceti{} in a region which is
+filled with a signal emitted by another node, will increase the
+number of active \cetis{} and the number of pairs of \cetis{} in
+causal contact.
 %
@@
 terms of three independent parameters, namely, the mean time span
-between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$, their mean
-lifetime $\left<\tau_s\right>$, and the maximum distance a signal can be
-detected by another \ceti{} ($D_{max}$).
+between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$,
+their mean lifetime $\left<\tau_s\right>$, and the maximum distance a
+signal can be detected by another \ceti{} ($D_{max}$).
 %
@@
 Sotos_biotechnology_2019}, related to the time span between the
-appearance of consecutive \cetis{} (since when $\left<\tau_a\right>$ is shorter, it
-produces a greater density of \cetis{}).
+appearance of consecutive \cetis{} (since when $\left<\tau_a\right>$
+is shorter, it produces a greater density of \cetis{}).
 %
@@
 \ffn{1} We illustrate in the Fig.~\ref{F_scheme} the schematic
-representation of the region causally connected to the central
-\ceti{}.
+representation of the region causally connected to a given \ceti{}.
 %
@@
 %
-In particular, we plot on the vertical axis the distance traveled by
-a signal at light speed from the emitter.
+For the sake of simplicity, we show in the plot the light travel
+distance, i.e., the distance traversed by any signal spreading from
+the emitter at light speed (for example, any electromagnetic signal).
 %
-Panel (a) represents the region $R$ (shaded polygon) in space and time
-which is causally connected to the emitter (left vertex), following an
-"A" type event in which the node acquires the communication
-capability.
+In this scheme, a light pulse would follow a trajectory represented by
+a line at 45 degrees from the axes, given the units of the time and
+space axes are years and light years, respectively.
+%
+Panel (a) of the Figure represents the region $R$ (shaded polygon) in
+space and time which is causally connected to the emitter (left
+vertex), following an event (hereafter A type event) in which the node
+acquires the communication capability, or becomes ''active''.
 %
@@
 which it would be no longer detectable due to the decrease in the
-power per unit area, which falls under the detectability threshold.
+energy per unit area, which falls under the assumed detectability
+threshold.
 %
 Similarly, the surface of last contact (SLC) is another sphere that
-grows from a "D" event and carries the last signal produced by the
-emitter.
+grows from an event in which the nodes ceases to possess the
+communication capability (D type event), and carries the last signal
+produced by the emitter.
 %
@@
 %
-The receiver \ceti{} E$_1$ can listen signals from the emitter \ceti{}
+In this scheme, the entire lifetime of a \ceti{} can then be
+represented by a polygon, limited in the spatial axis by the maximum
+distance of the signal $D_{max}$ and in the time axis by the time span
+between the starting time (A event, concave vertex) and the ending
+time (D event, convex vertex).
+%
+This representation allows to easily visualize the important events
+that result from the existence and communications of two \cetis{}.
+%
+The points in time where the \cetis{} acquire their communicating
+capacity are dubbed ''Awakening'' events ($A_1$ and $A_2$ in the
+Figure).
+%
+Similarly, the points in time where the \cetis{} loss their
+communicating capacity are dubbed ''Doomsday'' events ($D_1$ and $D_2$
+in the Figure).
+%
+The points in space--time where the first contact is produced for each
+one of the \cetis{}, are defined as ''Contact'' events, shown as $C_1$
+and $C_2$ in the Figure.
+%
+Finally, the two points where the contact is lost for each one the
+\cetis{}, are denominated ''Blackout'' events, shown as $B_1$ and
+$B_2$ in the Figure.
+%
+The receiver node E$_1$ can listen signals from the emitter node
 E$_2$, from the first contact event (t=C$_{12}$) up to the last
@@
 %
-Similarly, the receiver \ceti{} E$_2$ can listen signals from the
-emitter \ceti{} E$_1$, from t=C$_{21}$ up to t=B$_{21}$.
+Similarly, the receiver node E$_2$ can listen signals from the emitter
+node E$_1$, from t=C$_{21}$ up to t=B$_{21}$.
+%
+The time intervals for the open communication channel are determined
+by the ''C'' and ''B'' type events.
+%
+It is important to notice that there is a time delay between the
+contact events of the two \cetis{} involved in this analysis, and also
+a time delay between the two blackout events, so that the time range
+when a bidirectional contact is possible occurs between the maximum
+time of the contact points and the minimum time between the blackout
+points.
 
@@
 %{{{
-In this scheme, the entire lifetime of a \ceti{} can then be
-represented by a polygon, limited in the spatial axis by the maximum
-distance of the signal $D_{max}$ and in the time axis by the time span
-between the starting time (A event, concave vertex) and the ending
-time (D event, convex vertex).
-%
-This representation allows to easily visualize the important events
-that result from the existence and communications of two \cetis{}
-(panel (b)).
-%
-The points in time where the \cetis{} acquire their communicating
-capacity are dubbed ''Awakening'' events ($A_1$ and $A_2$ in the
-Figure).
-%
-Similarly, the points in time where the \cetis{} loss their
-communicating capacity are dubbed ''Doomsday'' events ($D_1$ and $D_2$
-in the Figure).
-%
-The points in space--time where the first contact is produced for each
-one of the \cetis{}, are defined as ''Contact'' events, shown as $C_1$
-and $C_2$ in the Figure.
-%
-Finally, the two points where the contact is lost for each one the
-\cetis{}, are denominated ''Blackout'' events, shown as $B_1$ and
-$B_2$ in the Figure.
-%
-It is important to notice that there is a time delay between the
-contact events of the two \cetis{} involved in this analysis, and also
-a time delay between the two blackout events, so that the time range
-when a bidirectional contact is possible occurs between the maximum
-time of the contact points and the minimum time between the blackout
-points.
 
-A causal connection can be produced without the need of the two
+The temporal structure that emerges naturally from the experimental
+setup implies that,
+as a fundamental property of the simulations,
+a causal connection can be produced without the need of the two
 involved locations being concurrent or active at the same time.
 %
-This is a fundamental property of the system which arises as a
+This property of the system arises as a
 consequence of the large spatial and temporal scales, where a message
-is transmitted at light speed.
+is transmitted at the (relatively small) light speed.
 %
@@
 sphere.
-
+%
 Following the end of the civilization, there is still a region which
@@
 %
+This approach has been also considered other statistical models
+\citep[e.g., ][]{smith_broadcasting_2009, grimaldi_signal_2017, Grimaldi2018}.
+%
 The trailing front, or surface of last contact (SLC) also grows from
 the central \ceti{}, with a delay with respect to the SFC equivalent
-to the lifetime of the \ceti{}, and produces a spherical shell region.
+to the lifetime of the node, and produces a spherical shell region.
 %
 Any other \ceti{} within this region will be aware of the originating
-\ceti{}, even if it has disappeared before the time of contact.
+node, even if it has disappeared before the time of contact.
 %
@@
 In our approach, we consider a model galaxy where the width of the
-disk is negligible with respect to the radius of the disk, so that the
-\cetis{} are placed on a 2D plane disk.
+disk is negligible with respect to the radius of the disk.
 %
@@
 regions of the spherical shells, respectively.
-
+%
 The initially growing communicating sphere, is shown over space--time
@@
 %
-In the upper panel (a), the sphere is growing as the surface of first
-contact has not reached the maximum distance.
-%
-In the middle panel (b) it has reached the maximum distance, so that
-it remains at the same size.
-%
-After a Doomsday event (bottom panel, b), the signals can still be
-observed, but the surface of last contact grows.
-%
-We also show in this Figure a schematic representation of two
-emitters, E$_1$ and E$_2$, that reach each other at different times.
+In the Fig.~\ref{F_scheme} the two
+emitters in panel (b), E$_1$ and E$_2$, reach each other at different times.
 %
@@
 factors, mainly the lifetimes, the time period between the A events
-and the distance between the two \cetis{} involved in the causal
+and the distance between the two nodes involved in the causal
 contact.
+%
+As it can be seen, the direction of the causal contact does not imply
+any ordering for the A events of the involved nodes.
+%
+In the implementation of the DE simulation, the six cases shown in the
+Figure determine the succession of events.
+%
+Each time an event is produced, the list of upcoming events is updated
+according to these cases.
 
-The temporal structure of the process is defined by two distribution
-parameters that represent the mean time interval that an intelligent
-civilization can emit and receive signals, and the mean time interval
-between the emergence of an intelligent communicating civilization and
-the next one.
+In out experimental configuration, there is no a temporal window for
+the model, i.e., the simulation starts when the process is already
+stable, and ends before any galactic effect could modify or affect of
+fixed variables.
 %
-The spatial structure of the simulation is given by the size and shape
-of the Galactic Habitable Zone and the maximum distance a signal can
-travel to be detected.
-%
-The parameters for the temporal distributions also determine the
-spatial properties since the density of active \cetis{} in the galaxy
-depend on these two parameters.
-%
-For example, a small $\tau_a$ and a large $\tau_s$ will produce a
-galaxy densely populated with \cetis{}, and vice versa.
-%
-Also, some hypotheses must be made in order to complete the
-simulation.
-%
-\citet{forgan_spatiotemporal_2011} argues that the times at which
-different civilizations become intelligent follow a Gaussian
-distribution, and then the distribution of inter-arrival times is an
-inverse exponential.
-%
-We chose to assume that the distribution of the times of A events is a
-stationary Poisson process, and then the distribution of the times
-between the appearance of new consecutive \cetis{} is exponential.
-%
-Regarding the duration of a \ceti{}, we propose that the distribution
-of the duration of \cetis{} is a stationary exponential distribution.
-%
-\citet{balbi_impact_2018} stress the fact that it would be desirable
-to arrive at a theoretical statistical distribution of the lifespan of
-civilizations, preferentially on the basis of the underlying
-astrophysical and biological processes.
-%
-\citet{maccone_lognormals_2014} argues that this distribution should
-be a log normal.  (EXPAND)
-
-There is no a temporal window for the model, i.e., the simulation
-starts when the process is already stable, and ends before any
-galactic effect could modify or affect of fixed variables.
-%
-The probability for the rise of a \ceti{} is homogeneous over the GHZ.
+Also, we assume that the probability for the rise of a \ceti{} is
+homogeneous over the GHZ.
 %
 Although the Galaxy has a well known spiral structure, the assumed
-sparsity of civilizations makes this assumption very reasonably.
+sparsity of nodes (intelligent civilizations) does make this
+assumption fairly reasonable.
 %
-Otherwise, if the distribution of civilizations is not sparse, it
-could be the case that the spiral arms would host most of the
-civilizations, and contacts are frequent between closely located
-civilizations.
+Otherwise, if the distribution of nodes is not sparse, it could be the
+case that the spiral arms would host most of the \cetis{}, and
+contacts are frequent between closely located nodes.
 %
@@
 %
-We also limit the possibilities of life or other types of civilizations
-to the hypothesis usually stated for the definition of the GHZ
-\citep{dayal_habitability_2016, gonzalez_galactic_2001,
+We also limit the possibilities of life or other types of
+civilizations to the hypothesis usually stated for the definition of
+the GHZ \citep{dayal_habitability_2016, gonzalez_galactic_2001,
 lineweaver_galactic_2004, gonzalez_habitable_2005,
@@
 %
-\citet{vukotic_astrobiological_2012} propose probabilistic cellular
-automata model to explain the astrobiological history of the Milky
-Way.
+For example, \citet{vukotic_astrobiological_2012} propose
+probabilistic cellular automata model to explain the astrobiological
+history of the Milky Way.
 %
@@
 With this, we are considering messages sent through, for example,
-electromagnetic radiation or even gravitational radiation, but we set
-apart messages sent with mechanical means or physical objects, or
-through some unknown technology that violates the known so far laws of
-physics.
+electromagnetic radiation or even gravitational radiation or neutrino
+communication \citep{wright_how_2018}, but we set
+apart messages sent with mechanical means or physical objects
+\citep[e.g., ][]{Armstrong2013}, or through some unknown technology
+that violates the known laws of physics.
 % 
@@
 the same time that the ability and intention to emit a signal in all
-direction, intended to be detected by an unknown civilization.
+directions, intended to be detected by an unknown civilization.
 %
@@
 In such a system we compute probabilities of a \ceti{} making contact
-with another \ceti{}, and the results must be interpreted as for a
+with another node, and the results must be interpreted as for a
 randomly chosen \ceti{}, given certain conditions of homogeneity on an
@@
 %
-We stress the fact that we are not centered on the probabilities for
-the Earth itself, but on a generalized \ceti{}.
+Then, we stress the fact that we are not centered on the probabilities
+for the Earth itself, but on a generalized \ceti{}.
 %
@@
 
+The temporal structure of the process is defined by two distribution
+parameters that represent the mean time interval that a \ceti{}
+can emit and receive signals, and the mean time interval
+between the emergence of consecutive \cetis{}.
+%
+The spatial structure of the simulation is given by the size and shape
+of the Galactic Habitable Zone (GHZ) and the maximum distance a signal can
+travel to be detected (D$_{max}$).
+%
+The parameters for the temporal distributions also determine the
+spatial properties, since the density of active \cetis{} in the galaxy
+depend on these two parameters.
+%
+For example, a small $\tau_a$ and a large $\tau_s$ will produce a
+galaxy densely populated with \cetis{}.
+%
+Also, some hypotheses regarding the shapes of the distributions of the
+temporal parameters must be made in order to complete the
+simulation.
+%
+\citet{forgan_spatiotemporal_2011} argues that the times at which
+different civilizations become intelligent follow a Gaussian
+distribution, and then the distribution of inter-arrival times is an
+inverse exponential.
+%
+We choose to assume that the distribution of the times of A events is a
+stationary Poisson process, and then the distribution of the times
+between the appearance of new consecutive \cetis{} is exponential.
+%
+Regarding the duration of a \ceti{}, we propose that the distribution
+of the duration of \cetis{} is a stationary exponential distribution.
+%
+\citet{balbi_impact_2018} stress the fact that it would be desirable
+to arrive at a theoretical statistical distribution of the lifespan of
+civilizations, preferentially on the basis of the underlying
+astrophysical and biological processes.
+%
+\citet{maccone_lognormals_2014} argues that this distribution should
+be a log normal.
+
 The power law and exponential statistical distributions are among the
@@
 For example, the distribution of the frequency of words in many
-languages is known to follow a Zipf law (which is a power law).
+languages is known to follow a Zipf's law (which is a power law).
+%
+This distributions arise any time a phenomena is characterized by
+commonly occurring small events and rarely produced large events
+\citep[e.g. ][]{adamic_zipf_2000}.
 %
@@
 countries, corporation sizes, income rankings, ranks of number of
-people watching the same TV channel (Zipf's law).
-%
-% CITAR ALGO DE LA LEY DE ZIPF
+people watching the same TV channel.
 %
@@
 interruptions from accidents, are also well described by power laws
-(Sornette 2006).
+\citep{sornette_critical_2006}.
 %
@@
 spontaneous appearance of life, and the emergence of a \ceti{} is
-independent of the existence of previous \cetis{} in the galaxy.
+independent of the existence of previous nodes in the galaxy.
 %            
@@
 
-The odds of a causal contact between two \cetis{} should not be
-considered as the odds of a contact between two intelligent
-civilizations, and in fact the later could be much lesser than the
-former.
+In this Section we discuss the degree of complexity in the model,
+considering a compromise between the accuracy of the model and the
+number of parameters that are free or with a high uncertainty.
+%
+Firstly, we emphasize that the odds of a causal contact between two
+\cetis{} should not be considered as the odds of a contact between two
+intelligent civilizations, and in fact the later could be much lesser
+than the former.
 %
@@
 %
-There is a trade--off between the simplicity and the complexity of the
-experiments, and further analysis could be performed following this
+Besides the trade--off between the simplicity and the complexity of
+the experiments, further analysis could be performed following this
 framework in order to explore possible implications of the results for
@@
 the receiver to broadcast and scan the sky for intelligent signals,
-respectively.
+respectively \citep{grimaldi_signal_2017}.
 %
@@
 although the effect of reducing the probability of contact is
-basically the product of the efficiency rations and thus such
+basically the product of the efficiency ratios and thus such
 implementation is not necessary.
@@
 efficiency \citep{smith_broadcasting_2009, anchordoqui_upper_2019,
-forgan_collimated_2014}.
+forgan_collimated_2014} or for beam--like transmissions
+\citep{grimaldi_signal_2017}.
 %
 Other considerations include the effects of alignments, the use of
-stars as sources or amplifiers, the nature of the message carrier, the
+stars as sources or amplifiers \citep{Edmondson2003, borra_searching_2012}, 
+the nature of the message carrier, the
 use of a spatial distribution that actually resembles the spiral shape
@@
 lifespan of a \ceti{} that receives a message, although the
-implementation of such behaviur would increase the number of free
+implementation of such behavior would increase the number of free
 parameters and would be highly speculative.
@@
 architecture of events, the relationships between them and how events
-trigger changes on the state of the system.
+trigger changes on the state of the system,
+so that the system is updated each time an event is produced.
 %
 We consider a simulated system that represents the galactic habitable
-zone (GHZ).
+zone which, on a first approach, is a 2-dimensional annular region.
 %
-On a first approach, the GHZ is a 2-dimensional annular region.
-%
-The adopted values for the GHZ are 20 kly and 60 kly, for the inner
+The adopted values for the geometry of the GHZ are 20 kly and 60 kly, for the inner
 and outer radius, respectively \citep{lineweaver_galactic_2004}.
@@
 Similarly, a \ccontact can be produced at the time of the \aawakening
-or later, and there can be several \contacts for the same \ceti{}.
-
-The system is updated each time an event is produced.
+or later, and there can be several \contacts for the same node.
 %
-The time marks for each event (A and D for each \ceti{} and C and B for
-each causal contact) are stored as a result of the simulation.
+The time marks for each event (A and D for each \ceti{} and C and B
+for each causal contact) are stored as a result of the simulation.
 %
@@
 %
-Some fixed variables must be set in order to carry out the simulation:
-the size of the Galactic Habitable Zone, the mean lifetime of a
-\ceti{},
-$\tau_s$, the mean waiting time for the appearance of another \ceti{},
-$\tau_a$, and the maximum distance a message can be detected.
+%Some fixed variables must be set in order to carry out the simulation:
+%the size of the Galactic Habitable Zone, the mean lifetime of a
+%\ceti{}, $\tau_s$, the mean waiting time for the appearance of another
+%\ceti{}, $\tau_a$, and the maximum distance a message can be detected.
 
-The Galactic Habitable Zone is the only one of the four parameters
-that is barely known, is set at fixed values of the inner and outer
-radii, radial symmetry is assumed, and the with of the galactic disk is
-neglected compared to the radial size.
+The Galactic Habitable Zone is set at fixed values of the inner and outer
+radii, radial symmetry is assumed, and the with of the galactic disk
+is neglected compared to the radial size.
 %
@@
 single and comprehensive unit for both space and time coordinates. 
-
+%
 The simulation was implemented on a python 3 code, which is publicly

### 2020-01-13 Sec_motivations.tex
ultimos retoques

@@
 
-The Drake equation \citep{drake_intelligent_1962}
-offers a very helpful educated guess,
-a rational set of lenses --the factors in the equation-- through which
-to look at future contacts with a technologically advanced
-civilization in the Milky Way.
+The Drake equation \citep{drake_intelligent_1962} offers a very
+helpful educated guess, a rational set of lenses --the factors in the
+equation-- through which to look at future contacts with a
+technologically advanced civilization in the Milky Way.
 %
-The equation quantifies the number of
-civilizations from whom we might receive an electromagnetic signal.
+The equation quantifies the number of civilizations from whom we might
+receive an electromagnetic signal, using a number of factors that have
+been widely discussed in the literature and whose estimates are
+updated constantly.
 %
-A comprehensive review and analysis of each term of the equation
-is presented in \citet{vakoch_drake_2015}.
+A comprehensive review and analysis of each term of the equation is
+presented in \citet{vakoch_drake_2015}.
 %
-\citet{prantzos_joint_2013}
-proposes a unified framework
-for a joint analysis of the Drake equation and the Fermi paradox
-concluding that for sufficiently long-lived civilizations,
-colonization of the Galaxy is the only reasonable option to gain
-knowledge about other life forms.
+Some optimistic estimates from the Drake equation contrast with the so
+called Fermi paradox, a statement about the contradiction between the
+apparent abundance of life in the Galaxy and its lack of evidence
+\citep[e.g. ][]{hart_explanation_1975, brin_great_1983,
+barlow_galactic_2012, forgan_galactic_2016, vanhouten_isthere_2017,
+Sotos_biotechnology_2019, carroll_nellemback_fermi_2019}.
 %
-\citet{haqq-misra_drake_2017}
-discuss the dependence of the Drake equation parameters on the
-spectral type of the host star and the time since the galaxy formed
-and examine trajectories for the emergence of communicative
-civilizations over the history of the galaxy.
-%
-The abundance of
-intelligent civilizations in the galaxy is a longstanding question,
-which is often conceptualized as the problem of the lack of received
-communication or the Fermi paradox 
-\citep{barlow_galactic_2012, Sotos_biotechnology_2019,
-forgan_galactic_2016}.
-
 There are many proposals aimed at solving this paradox, which make use
@@
 % 
-%%%%% agregar Carroll-Nellenback, 2019).
-\Fpagebreak
-%
-The Drake equation is a key tool
-to organize the discussion around the problem of the abundance
-of civilizations 
-in the Galaxy \citep{hinkel_interdisciplinary_2019}.
-%
-However, the uncertainties in the factors of this equation make
-it less prone to a formal application in order to define searching
-strategies or to compute the actual number of extraterrestrial
-intelligences.
-%
-The optimistic estimation from the Drake equation contrast with
-the 
-SETI initiatives,
-noticeable overshadowed by the fact that despite decades of effort, not a
-single extraterrestrial intelligence, if any, has been detected.
-
-Some modifications to the original idea of the equation have tried to
-imprint a stochastic nature, or to propose a probabilistic approach,
-or to consider the temporal structure which is missing in the
-equation.
-%
-Temporal aspects of the distribution of communicating civilizations
-and their contacts have also been explored by several authors
-\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
-balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011},
-%The Impact of the Temporal Distribution of Communicating
-%Civilizations on Their Detectability Temporal dispersion of the
-%emergence of intelligence: an inter-arrival time analysis The
-%spatiotemporal aspects of SETI
-%
-as well as efforts on considering the stochastic nature of the Drake
-equation \citep{glade_stochastic_2011}.
-%
-The simulation approach has also been considered
-\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
-murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017},
-although with a similar problem that consists on the large number of
-parameters which are either unknown or largely uncertain.
-%
-%New numerical determination of habitability in the Galaxy: the SETI
-%connection
-%
-%Since the factors in the Drake equation are uncertain,
-%
-In this work, we propose to avoid the frequentist approach of the
-Drake equation and to explore a parameter space, where instead of
-computing a final number, we provide a statistical distribution that
-gives conditional probabilities.
-%
-This is an exploratory analysis that aims at providing a numerical
-tool to discuss not the several theoretical problems summarized by the
-Drake equation factors, but the different scenarios on the basis of
-statistical heuristics.
-%
-The approach proposed here should be considered as a compromise between
-the uncertainties of the frequentist approach and the detailed recipes
-required on the simulation approach.
-%
-We then provide a numerical framework to explore via simulations a
-parameter space of unknown observables in order to discuss different
-scenarios and their consequences in terms of the probability con
-making contacts.
-%
-The only fact that can be stated with certainty is that for the number
-of years SETI projects have been working we have not received any
-signal, at least within the technical possibilities and the conditions
-established by SETI \citep{tarter_search_2001}.
-%
-This arguably astonishing result has lead to the Fermi paradox, which
-states an apparent contradiction between the high estimation of the
-abundance of life in the Galaxy and the lack of their detection
-\citep{vanhouten_isthere_2017}.
+We stress the fact that the only observation that can be stated with
+certainty is that for the number of years projects for the Search of
+Extraterrestrial Intelligences (SETI) have been working, we have not
+received any signal using the technical possibilities and the
+conditions established by SETI projects \citep{tarter_search_2001}.
 %
@@
 
-The discrete events method for simulating a stochastic process is an
-approximation that allows to study the behavior of complex systems,
-by considering a sequence of well defined discrete events.
+\Fpagebreak
 %
-The simulation is carried out by following all the variables that
-describe and constitute the state of the system, and the evolution of
-the process is described as a set of changes in that state.
+The Drake equation is a key tool to organize the discussion around the
+problem of the abundance of civilizations in the Galaxy
+\citep{hinkel_interdisciplinary_2019}.
 %
-In this context, an event produces a specific change in the state,
-that can be triggered by random variables that encode the stochastic
-nature of the physical phenomenon.
+However, the uncertainties in the factors of this equation make it
+less prone to a formal application in order to define searching
+strategies or to compute the actual number of extraterrestrial
+intelligences.
 %
-For example, when a new contact is produced between two entities in
-the simulated galaxy, the numbers of active communication lines and of
-active communicated nodes change.
+The optimistic estimations from the Drake equation also contrast with
+the SETI initiatives, noticeable overshadowed by the fact that despite
+decades of effort, not a single extraterrestrial intelligence, if any,
+has been detected.
 %
-The process involves then following the changes on the state of the
-system, defining the initial and final states, defining a method that
-allows to keep track of the time progress in steps, and maintaining a
-list of relevant events.
+\citet{prantzos_joint_2013} proposes a unified framework for a joint
+analysis of the Drake equation and the Fermi paradox, concluding that
+for sufficiently long-lived civilizations, colonization of the Galaxy
+is the only reasonable option to gain knowledge about other life
+forms.
+%
+\citet{haqq-misra_drake_2017} discuss the dependence of the Drake
+equation parameters on the spectral type of the host star and the time
+since the Galaxy formed and examine trajectories for the emergence of
+communicative civilizations over the history of the Galaxy.
+%
+Some modifications to the original idea of the equation have tried to
+imprint a stochastic nature, to propose a probabilistic approach, or
+to consider the temporal structure which is missing in the equation.
+%
+Temporal aspects of the distribution of communicating civilizations
+and their contacts have also been explored by several authors
+\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
+balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011}, as
+well as efforts on considering the stochastic nature of the Drake
+equation \citep{glade_stochastic_2011}.
+%
+The simulation approach has also been considered
+\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
+murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017,
+ramirez_new_2017}, although with a similar problem that consists on
+the large number of parameters which are either unknown or largely
+uncertain.
+%
+\citet{grimaldi_signal_2017} consider a statistical model for the
+probability of the Earth making contact with other intelligent
+civilizations, taking into account the finite lifetime of signal
+emitters, and based on the fractional volume occupied by all signals
+reaching our planet.
 
@@
 %
-\citet{smith_broadcasting_2009} uses a analytical model to compute the
+\citet{smith_broadcasting_2009} uses an analytical model to compute the
 probabilities of contact between two randomly located civilizations,
@@
 % 
+Using a different approach, 
 \citet{balbi_impact_2018} investigate the chance of communicating
@@
 location of the Earth, using a statistical model to compute the
-occurrence of causal contacts. The author stresses the fact that the
+occurrence of causal contacts.
+%
+The author stresses the fact that the
 causal contact requirement involves the distance between
@@
 %
-The radius of influence is set to 1000 lyr.
+%The radius of influence is set to 1000 lyr.
 %
@@
 %
-\citet{vukotic_astrobiological_2012} propose a simulation approach
-based on probabilistic cellular automata (PCA) modeling. In this
+In another probabilistic approach,
+\citet{vukotic_astrobiological_2012} propose a
+probabilistic cellular automata (PCA) modeling. In this
 framework, a complex system is modeled by a lattice of cells which
@@
 %
+We propose to avoid the frequentist approach of the
+Drake equation and to explore a parameter space, where instead of
+computing a final number, we provide a statistical distribution that
+gives conditional probabilities.
+%
+This is an exploratory analysis that aims at providing a numerical
+tool to discuss not the several theoretical problems summarized by the
+Drake equation factors, but the different scenarios on the basis of
+statistical heuristics.
+%
+The approach proposed here should be considered as a compromise
+between the uncertainties of the frequentist approach and the detailed
+recipes required on the simulation approach.
+%
+We then provide a numerical framework to explore via simulations a
+parameter space of unknown observables in order to discuss different
+scenarios and their consequences in terms of the probability of
+making contacts.
+%
+The method we use for simulating a stochastic process is an
+approximation that allows to study the behavior of complex systems,
+by considering a sequence of well defined discrete events.
+%
+The simulation is carried out by following all the variables that
+describe and constitute the state of the system, and the evolution of
+the process is described as a set of changes in that state.
+%
+In this context, an event produces a specific change in the state,
+that can be triggered by random variables that encode the stochastic
+nature of the physical phenomenon.
+%
+For example, when a new contact is produced between two entities in
+the simulated galaxy, the numbers of active communication lines and of
+active communicated nodes change.
+%
+The process involves following the changes on the state of the
+system, defining the initial and final states, defining a method that
+allows to keep track of the time progress in steps, and maintaining a
+list of relevant events.
+%
 In Sec.~\ref{S_methods} we introduce the methods and discuss the
@@
 
-% HACER DISCUSION FUERTE DE ESTOS PAPERS: \cite{balbi_impact_2018,cirkovic_temporal_2004,grimaldi_signal_2017,smith_broadcasting_2009}
-
 %}}}

### 2020-01-13 Sec_results.tex
ultimos retoques

@@
 the times of each of the events that are relevant to keep track of the
-number of \cetis{} (time of awakening and time of doomsday) and of the
-number of contacts (the times of contacts and the times of blackouts).
+number of \cetis{} (time of Awakening and time of Doomsday) and of the
+number of contacts (the times of contacts and the times of Blackouts).
 %
@@
 properties of the \cetis{}, for example the total time elapsed between the
-awakening and the doomsday of each \ceti{}.
+Awakening and the Doomsday of each node.
 %
@@
 Regarding the properties of the population of \cetis{} and its evolution,
-the simulation yields the fraction of awaken time a \ceti{} is listening
-at least another \ceti{} (i.e., in their causal cone), the age of
+the simulation yields the fraction of awaken time a node is listening
+at least another node (i.e., in their causal cone), the age of
 contacted \cetis{} at first contact, the distribution of time to wait
@@
 each \ceti{}, the distribution of the number of contacts as a function of
-\ceti{} age, the number of contacts as a function of time in the galaxy,
-the rate of \cetis{} that succeed in contact, and the distribution of
-distances between contacted \cetis{}.
+the age of the node, the number of contacts as a function of time in the galaxy,
+the fractions of nodes that succeed in contact, and the distribution of
+distances between contacted nodes.
 %
@@
 \ceti{} vs. the time of double communication, the distance to \ceti{} vs. the
-age of contacted \ceti{}, the age of a \ceti{} and the maximum number of
-contacted \cetis{} before doomsday, or the lifespan of a \ceti{} vs. the
+age of contacted node, the age of a node and the maximum number of
+contacted nodes before the Doomsday, or the lifespan of a node vs. the
 maximum number of contacts.
@@
 Table~\ref{T_selected_models}.
-
-In what follows, we focus on the analysis of the duration of contacts
-(Sec.~\ref{SS_multiplicity}), the waiting time for the first contact
-(Sec.~\ref{SS_waiting}) and the
-variations of contact possibilities as a function of the position of
-the \ceti{} in the galaxy (Sec.~\ref{SS_location}).
 
@@
 %
-This consideration offers a new approach for SETI programs, where the
+This consideration offers a new approach to SETI programs, where the
 search for new communication technologies has a fundamental role.

### 2020-01-13 manuscript.tex
ultimos retoques

@@
 %
-The abundance of intelligent civilizations in the galaxy is a
+The abundance of intelligent civilizations in the Galaxy is a
    longstanding question, which is often conceptualized as the problem
@@
 Most efforts on the estimation of a number of intelligent
-   civilizations are centered on the Drake equation, although its
+   civilizations are focused on the Drake equation, although its
    factors are affected by large uncertainties, and it lacks a
@@
 We argue that three parameters can be used to describe a variety of
-situations, ranging from a Galaxy where an intelligent civilization is
-very rare, to a Galaxy populated with plenty of civilizations in
+situations, ranging from a galaxy model where an intelligent civilization is
+very rare, to a galaxy model populated with plenty of civilizations in
 causal contact.
@@
 model parameters that reflect different, so far unknown, scenarios for
-the appearance of intelligent life on the galaxy.
+the appearance of intelligent life on the Galaxy.
 
@@
 Under the hypotheses of our experiments, we conclude that a causal
-contact is extremely unlikely unless the Galaxy is heavily populated
+contact is extremely unlikely unless the galaxy is heavily populated
 by intelligent civilizations with large average lifetimes.
@@
 to reach the same probability of causal contact than in the case of a
-less populated Galaxy but with very ancient civilizations.
+less populated galaxy but with very ancient civilizations.
 %
@@
 the problem of the probabilities of causal contacts between location
-in the Galaxy the possibility of sending and receiving messages.
+in the galaxy the possibility of sending and receiving messages.
 %
@@
 few simple hypothesis to perform a complete model for the population and
-communication network in the Galaxy.
+communication network in the galaxy.
 %

### 2020-01-14 Sec_methods.tex
detalles...

@@
 morrison_extending_2015, haqq-misra_evolution_2019,
-rahvar_cosmic_2016, gobat_evolution_2016, rahvar_cosmic_2016}.
+rahvar_cosmic_2016, gobat_evolution_2016, rahvar_cosmic_2016} and
+consider that habitability remains constant \citep[see, however,
+][]{dayal_habitability_2016, gobat_evolution_2016}.
 %
@@
 apart messages sent with mechanical means or physical objects
-\citep[e.g., ][]{Armstrong2013}, or through some unknown technology
+\citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown technology
 that violates the known laws of physics.
@@
 The spatial structure of the simulation is given by the size and shape
-of the Galactic Habitable Zone (GHZ) and the maximum distance a signal can
+of the Galactic Habitable Zone and the maximum distance a signal can
 travel to be detected (D$_{max}$).
@@
 interruptions from accidents, are also well described by power laws
-\citep{sornette_critical_2006}.
+\citep{sornette_critical_2006, benguigui_classificacion_2016}.
 %
@@
 previous agreement, making the possibility of a contact with a message
-that could be deciphered highly rare.
+that could be deciphered highly rare \citep[see e.g. ][]{forgan_collimated_2014}.
 %

### 2020-01-14 Sec_motivations.tex
detalles...

@@
 \citep{forgan_numerical_2009, bloetscher_using_2019,
-glade_stochastic_2011, forgan_numerical_2010}, analytical
-interpretations of the Drake equation \citep{prantzos_joint_2013,
-smith_broadcasting_2009} or more speculative proposals
+glade_stochastic_2011, forgan_numerical_2010}.
+%
+Regarding the Drake equation, analytical interpretations
+\citep{prantzos_joint_2013,
+smith_broadcasting_2009} or reformulations
+\citep[][and references therein]{burchell_whither_2006} have also been
+proposed.
+%
+The lack of contacts could be explained by the difficulty of life to
+appear due to astrophysical explanations
+\citep{annis_astrophysical_1999}.
+%
+In general, also more speculative proposals
 \citep{barlow_galactic_2013, lampton_information_2013,
-conway_three_2018, forgan_galactic_2017}.
+conway_three_2018, forgan_galactic_2017} have been discussed.
+% 
+\Fpagebreak
 %
@@
 forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
-
-\Fpagebreak
 %
@@
 %
-This approach, which assigns probability distributions to the Drake
-factors.
+%This approach, which assigns probability distributions to the Drake
+%factors.
 %
@@
 %
-\citet{smith_broadcasting_2009} uses an analytical model to compute the
-probabilities of contact between two randomly located civilizations,
-and the waiting time for the first contact, assuming a fixed maximum
-broadcasting distance.
+\citet{smith_broadcasting_2009} uses an analytical model to compute
+the probabilities of contact between two randomly located
+civilizations, and the waiting time for the first contact, assuming a
+fixed maximum broadcasting distance.
 % 
-Using a different approach, 
-\citet{balbi_impact_2018} investigate the chance of communicating
-civilizations making causal contact within a volume surrounding the
-location of the Earth, using a statistical model to compute the
-occurrence of causal contacts.
+Using a different approach, \citet{balbi_impact_2018} investigate the
+chance of communicating civilizations making causal contact within a
+volume surrounding the location of the Earth, using a statistical
+model to compute the occurrence of causal contacts.
 %
-The author stresses the fact that the
-causal contact requirement involves the distance between
-civilizations, their lifespan and their times of appearance.  This is
-important since the time it takes the light to travel across the
-Galaxy is much lesser than its lifetime.
+The author stresses the fact that the causal contact requirement
+involves the distance between civilizations, their lifespan and their
+times of appearance.  This is important since the time it takes the
+light to travel across the Galaxy is much lesser than its lifetime.
 %
@@
 In another probabilistic approach,
-\citet{vukotic_astrobiological_2012} propose a
-probabilistic cellular automata (PCA) modeling. In this
-framework, a complex system is modeled by a lattice of cells which
-evolve at discrete time steps, according to transition rules that take
-into account for each cell the states of its neighbor cells.
+\citet{vukotic_astrobiological_2012} propose a probabilistic cellular
+automata (PCA) modeling. In this framework, a complex system is
+modeled by a lattice of cells which evolve at discrete time steps,
+according to transition rules that take into account for each cell the
+states of its neighbor cells.
 %
@@
 represent life complexity on a 2-dimensional annular ring resembling
-the GHZ. The authors set the GHZ between a minimum radius of 6 kpc and
-a maximum radius of 10 kpc. Within this framework, the authors also
-make several MonteCarlo simulations and analyze ensemble-averaged
-results.
+the Galactic Habitable Zone (GHZ). 
+%
+The authors set the GHZ between a minimum radius of 6 kpc and a
+maximum radius of 10 kpc. Within this framework, the authors also make
+several MonteCarlo simulations and analyze ensemble-averaged results.
 %

### 2020-01-14 manuscript.tex
detalles...

@@
 %
-Here we present numerical simulations of stochastic processes for the
-   emergence of nodes with communication capabilities and analyze the
-   causal connections among them.
+Here we present a statistical model for the abundance and duration of
+   intelligent civilizations, implemented through Monte Carlo
+   numerical simulations.
+%
+Using computer experiments, we explore the hypothesis space of the
+   model by a suite of simulations which represent stochastic
+   processes for the emergence of nodes with communication
+   capabilities, and then analyze the causal connections among them.
 %
@@
 %
-Roughly, in order to have at least one contact in the entire lifetime of the
-\ceti{}, there should appear a mean of at least one \ceti{} every $\sim$20000 yr, with
-a mean active period of at least $\sim$30000 yr.
+Roughly, in order to have at least one contact in the entire lifetime
+of the \ceti{}, there should appear a mean of at least one \ceti{}
+every $\sim$20000 yr, with a mean active period of at least
+$\sim$30000 yr.
 %
+This result is qualitatively similar to the results presented by
+several authors, which state that a contact between the Earth and
+another intelligent civilization in the Galaxy is quite unlikely,
+provided the maximum distance of the signal and the lifetime of the
+emitter are not large enough.
+%
+This analysis supports the idea that, in order to increase the
+possibilities of a contact, more active strategies of the emitter
+would be required.
+%
+Some proposals in this direction include intertellar exploration,
+colonization and settlement \citep{brin_great_1983, Dosovic2019,
+galera_invasion_2019}, although it would require large temporal
+scales.
+%
+\citet{Dosovic2019} use probabilistic cellular automaton simulations
+to explore the parameter space of a model with colonization and
+catastrophic events.
+%
+According to the timescales involved, the results could explain the
+Fermi paradox.
+%
+Although our work does not take into account the colonization
+hypothesis, it does consider catastrophic events implicitly in the
+mean value and distribution of the lifetime, $\left<\tau_s\right>$.
+%
+Other strategies could also increase the probability of contacts, for
+example panspermia \citep[e.g.,][]{starling_virulence_2013} or
+self--replicating probes \citep[e.g.,][]{barlow_galactic_2013}, although they
+would be too slow to make a significant impact on the communication
+network among intelligent civilizations.
+
+% 
+% panspermia
+% 
+% probes
+% 
+% 
+% \citet{grimaldi_signal_2017}
+% \citep{balbi_impact_2018}
+% \citet{cirkovic_temporal_2004}
+% \citet{bloetscher_using_2019}
+% \citet{smith_broadcasting_2009}
+
+%%%%%%%%%%%%%%%
+
 There is a balance between the density of \cetis{} and the mean
@@
 However, a large number of nodes does not easily compensate their
-short lives 
-to reach the same probability of causal contact than in the case of a
-less populated galaxy but with very ancient civilizations.
+short lives to reach the same probability of causal contact than in
+the case of a less populated galaxy but with very ancient
+civilizations.
 %
@@
 %
-This suggests the possibility that an alternative SETI strategy could be
-the search for alternative message carriers, for the case in which the
-search has not been performed on the adequate channels.
+This suggests the possibility that an alternative SETI strategy could
+be the search for alternative message carriers, for the case in which
+the search has not been performed on the adequate channels.
 %

### 2020-05-05 code/main.tex
starting reply to reviewer

--- 2019-07-25 softdoc.tex
@@
 \begin{abstract}
-
+%
+The abundance of intelligent civilizations in the galaxy is a
+longstanding question, which is often conceptualized as the problem
+of the lack of received communication of the Fermi paradox.
+%
+Most efforts on the estimation of a number of intelligent
+civilizations are centered on the Drake equation, although 
+its factors are affected by large uncertainties, and it lacks a
+temporal nature.
+%
+Here we present numerical simulations of stochastic processes of
+emergence of civilizations with communication capabilities, and
+the causal contacts among them.
+%
+We analyze the rate of causal contacts as a function of the mean
+number of civilizations and the mean lifetime span distribution.
+%
+Our results indicate that, given the large distances involved, in
+   the odds to obtain a contact in a few years of monitoring, assuming a
+   perfect detection rate, are very rare.
 \end{abstract}
@@
 %{{{
+
+For citation purposes \citep{Lineweaver2004}
+For citation purposes \citep{Forgan2011}
+For citation purposes \citep{Rahvar2016}
+For citation purposes \citep{Glade2011}
+For citation purposes \citep{Dayal2016}
+For citation purposes \citep{Gobat2016}
+For citation purposes \citep{Prantzos2013}
+For citation purposes \citep{Horvat2007}
+For citation purposes \citep{Starling2013}
+For citation purposes \citep{Barlow2012}
+For citation purposes \citep{Tarter2009}
+For citation purposes \citep{Forgan2017}
+For citation purposes \citep{Forgan2013}
+For citation purposes \citep{Wright2015}
+For citation purposes \citep{Loeb2006}
+For citation purposes \citep{Solomonides2016}
+For citation purposes \citep{Fogg1987}
+For citation purposes \citep{Anchordoqui2017}
+For citation purposes \citep{Balb2018}
+For citation purposes \citep{Forgan2010}
+For citation purposes \citep{Forgan2008}
+For citation purposes \citep{Forgan2016}
+For citation purposes \citep{Haqq-Misra2017}
+For citation purposes \citep{Forgan2015}
+For citation purposes \citep{Gleiser2010}
+For citation purposes \citep{cirkovic_temporal_2004}
+
+For citation purposes \citep{cirkovic_temporal_2004}
+
+A common approach to discuss the possible number of extraterrestrial
+intelligent civilizations (ETI) has been the use of the Drake equation
+\citep{}.
+
+Uncertainties in the factors
+
+Stochastic Drake equation and probabilistic approaches
+
+Time:
+The Impact of the Temporal Distribution of Communicating Civilizations on Their Detectability
+Temporal dispersion of the emergence of intelligence: an inter-arrival time analysis
+The spatiotemporal aspects of SETI
+\citep{Fogg1987, Forgan2011, Balb2018}
+
+Effort on considering the stochastic nature of the Drake equation
+\citep{Glade2011}.
+
+Simulation approach
+New numerical determination of habitability in the Galaxy: the SETI connection
+\citep{Forgan2008, Forgan2010}
 
@@
 
+Some complex stochastic processes can be efficiently modeled with the
+discrete--event (DE) simulation approach.
+%
+Simulations are suitable tools to analyze systems that evolve with
+time and involve randomness.
+%
+In general, simulations require less assumptions and simplifications
+than theoretical approaches, and can be applied to systems where 
+a theoretical model can not be found.
+%
+A system described with the DE paradigm is characterized by a set of
+actors and events, where actors interact causally through a series of
+events on a time line and process these events in chronological order
+\citep{ptolemaeus_book, chung_book, simulation_book}.
+%
 On the particular case of the difusion of signals of intelligence in
 the Galaxy, the discrete events method can be performed taking into
-account a small number of variables.  The temporal structure is
-defined by two distribution parameters that represent the mean
-time interval that an intelligent civilization can emit and receive
-signals, and the mean time interval between the emergence of an
-intelligent comunicating civilization and the next.
+account a small number of variables.
+%
+The temporal structure is defined by two distribution parameters that
+represent the mean time interval that an intelligent civilization can
+emit and receive signals, and the mean time interval between the
+emergence of an intelligent comunicating civilization and the next one.
+%
 The spatial structure of the simulation is given by the size and shape
@@
 travel to be detected.
-
+%
+The parameters for the temporal distributions also determine the
+spatial properties.
+%
+For example, the density of active cetis in the Galaxy depend on these
+two parameters.
+%
 Also, some hypothesis must be made in order to complete the
@@
    \item The message travels at light speed.
+\end{itemize}
 
-\end{itemize}
+The exponential distribution of lifespan and waiting times is
+justified by considering that the process of appeareance of life in
+the galaxy is homogeneous and stationary.
+%
+This means that there is no preferred location within the GHZ for the
+spontaneos appeareance of life, and that the emergence of a
+civilization is independent of the existence of previous civilizations 
+in the galaxy.
+%
+This is equivalent to proposing a Poisson process for the emergence of
+CETIs, since there is a close relation between the number of events in
+time or space and the waiting time or separation, respectively.
+%
+That is, these are two alternative approaches to describing the same
+process, a Poisson distribution for the number of events implies an
+exponential distribution for their sepations, and viceversa.
+%
+It should be emphasized that the exponential laws used in this work
+are assumed, although instead of analyzing results from a particular
+parameter chosen ad--hoc, we explore the hypothesis space and analyze
+the impact of the values of these parameters on the results.
+%
+
+\subsection{Power laws vs. exponential laws}
+
+The power law and exponential statistical distributions are among the
+most common patterns found in natural phenomena.
+%
+For example, the distribution of the frecuency of words in many
+languajes is known to follow a Zipf law (which is a power law).
+%
+Zipf law also describes population ranks of cities in various
+countries, corporation sizes, income rankings, ranks of number of
+people watching the same TV channel (Zipf's law entry on Wikipedia).
+%
+The magnitudes of earthquakes, hurricanes, volcanic eruptions and
+floods; the sizes of meteorites or the losses caused by
+business interruptions from accidents, are also well described by power
+laws (Sornette 2006).
+%
+\itshape
+Other studies have
+documented power laws in stock market fluctuations,
+sizes of computer files and word frequency in languages
+(Mitzenmacher, 2004; Newman, 2005; Simkin \& Roy-
+chowdhury, 2006).
+In biology, power laws have been particularly impor-
+tant in analysing connectivity patterns in metabolic
+networks (Barabasi \& Albert, 1999; Ravasz et al., 2002)
+and in the number of species observed per unit area in
+ecology (Garcia Martin \& Goldenfeld, 2006).
+(VER Frank, the common patterns of nature).
+\slshape
+
+Also, exponential laws arise when...
+
+\subsection{Exploration of the parameter space}
+
+Zipf law.
+
+The adopted values for the GHZ are XX and XX \citep{Lineweaver2004}.
+
+\begin{table}
+\begin{tabular}{ccc}
+\hline
+   parameter & \multicolumn{2}{c}{adopted value}\hfill \\
+\hline
+   inner GHZ radius & \multicolumn{2}{c}{}\\
+   outer GHZ radius & \multicolumn{2}{c}{}\\
+\hline
+   parameter & min & max \\
+\hline
+   D$_{max}$ & & \\
+   $\tau_{a}$ & & \\
+   $\tau_{N}$ & & \\
+\hline
+\end{tabular}
+\caption{Adopted values for fixed and variable parameters in the
+   simulation}
+\label{T_parameters}
+\end{table}
 
@@
 %========================================================================
+
+\begin{figure}
+   \centering
+   \includegraphics[width=0.5\textwidth]{growingsphere.pdf}
+   \caption{Schematic representation of the growing communicating
+   sphere, over the space--time diagrams.  In (a) the sphere is
+   growing as the surface of first contact has not reached the maximum
+   distance.  In (b) it has reached the maximum distance, so that it
+   remains at the same size.  After a Doomsday event, the signals can
+   still be observed, but the surface of last contact grows }
+   \label{F_sphere}
+\end{figure}
+
+\begin{figure}
+   \centering
+   \includegraphics[width=0.5\textwidth]{abcd.pdf}
+   \caption{Schematic representation of two emitters, $E_1$ and $E_2$,
+   that reach each other at different times.  The time span for $E_i$
+   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
+   emitter $j$ between $C_{ij}$ and $B_{ij}$.
+   The type and length of causal contact in both directions depend on
+   the distance and time lag between the awakening events, the maximum
+   distance that a signal can reach and the time period in which each
+   emitter is active.}
+   \label{F_abcd}
+\end{figure}
 
@@
       assumed, so that the GHZ is determined by two radii, the inner
-      radius (GHZ_inner) and the outer radius (GHZ_outer).  They are
+      radius (GHZ\_inner) and the outer radius (GHZ\_outer).  They are
       measured in light years, with the aim to maintain y single and
@@
 
-%#
-%# CONTENTS OF THE <CETIS> PICKLE:
-%# . . . . . . . . . . . . . . . . . . . . . . . .
-%# 0. ID of emitting CETI
-%# 1. ID of emitting [receiving] CETI
-%# 2. x position in the galaxy
-%# 3. y position in the galaxy
-%# 4. time of Awakening [Contact]
-%# 5. time of Doomsday  [Blackout]
-%# . . . . . . . . . . . . . . . . . . . . . . . .
-%#
-%# This structure is always for the A-D cycle of each emitting CETI,
-%# and repeats for each contact (C-B).
-%#
-%# The time span of a CETI is t_D - t_A for CETIs[k][0]
-%# The time span of a contact is t_B - t_C for CETIs[k][i], with i>0
-%#
-%#
-%# age: time elapsed from A to a given time
-%# ceti_e: ceti emitting signals (emiter)
-%# ceti_r: ceti receiving signals (receiver)
-%# ceti_c: ceti that listen at least another ceti (citizen)
-%# ceti_h: ceti that is lestened by at least another ceti
-%#
+Direct quantities are:
+
+\begin{itemize}
+   \item ID of emitting CETI
+   \item ID of emitting (receiving) CETI 
+   \item x position in the galaxy
+   \item y position in the galaxy
+   \item time of Awakening (Contact)
+   \item time of Doomsday  (Blackout)
+\end{itemize}
+
+This structure is always for the A-D cycle of each emitting CETI,
+and repeats for each contact (C-B).
 %
-%# for Check purposes
-%#-------------------
-%# duration of a civilization (exponential distribution by costruction)
-%# time from the appeareance of a CETI to the next
-%#
-%# 
-%# derived quantities
-%#-------------------
-%# [time]
-%# time span of a ceti listening another
-%# time span of a ceti being listened by another
-%# duration of two-way communication channels
-%# waiting time until the first contact
-%# fraction of awaken time a ceti is listening at least another ceti
-%# age of contacted cetis_e at first C
-%# distribution of time to wait until next C
-%# fraction of cetis where the first contact is given at the A
-%
-%# [numbers]
-%# distribution of the number of contacts for each ceti
-%# distribution of the maximum number of contacts for each ceti
-%# distribution of the number of contacts as a function of age
-%# number of contacts as a function of time in the galaxy
-%# rate of cetis that succeed in contact 
-%# fraction of contacts that admit a response
-%#
-%# [distances]
-%# distribution of distances between contacted cetis
-%#
-%# [correlations] relation between {RB}
-%# RB distance to ceti and time of double communication
-%# RB distance to ceti and age of contacted ceti
-%# RB the age of a ceti and the maximum number of contacted cetis before D
-%# RB lifespan of a ceti and max number of contacts
-%#
-%# distribution in the galaxy of cetis that reach contact
-%
-%# ... everything as a function of parameters!
+We define the following quantities: age is the time elapsed from A to a
+given time, ceti\_e is the CETI emitting signals (emiter), ceti\_r is
+the CETI receiving signals (receiver), ceti\_c is the ceti that listen
+at least another ceti (citizen) and ceti\_h is the CETI ceti that is
+lestened by at least another CETI.
 
-Output
+Then, several derived quantities can be obtained:
+
 \begin{itemize}
-   \item La salida de la simulación está guardad en una variable llamada CETI.
-   \item La misma es una lista de objetos que son también listas.
-   \item El elemento cero de cada lista (i.e., cada elemento de CETI) contiene:
-   \item coordenada X
-   \item coordenada Y
-   \item tiempo de inicio de la capacidad de comunicación
-   \item tiempo de final de la capacidad de comunicación
-   \item y los demás elementos contienen:
-   \item coordenada X de la CETI
-   \item coordenada Y de la CETI
-   \item tiempo del primer contacto
-   \item tiempo del último contacto
+   \item time span of a ceti listening another
+   \item time span of a ceti being listened by another
+   \item duration of two-way communication channels
+   \item waiting time until the first contact
+   \item waiting time until the next contact
+   \item fraction of awaken time a ceti is listening at least another ceti 
+   \item age of contacted cetis\_e at first C
+   \item distribution of time to wait until next C
+   \item fraction of cetis where the first contact is given at the A       
+   \item distribution of the number of contacts for each ceti         
+   \item distribution of the maximum number of contacts for each ceti 
+   \item distribution of the number of contacts as a function of age  
+   \item number of contacts as a function of time in the galaxy       
+   \item rate of cetis that succeed in contact                        
+   \item fraction of contacts that admit a response                   
+   \item distribution of distances between contacted cetis 
+   \item relation between distance to ceti and time of double communication 
+   \item relation between distance to ceti and age of contacted ceti
+   \item relation between the age of a ceti and the maximum number of contacted cetis before D
+   \item relation between lifespan of a ceti and max number of contacts
+   \item distribution in the galaxy of cetis that reach contact
 \end{itemize}
+
+All these quantities can be analyzed as a function of the simulation
+parameters.
 
@@
 
-\section{Results}
+\subsection{Duration of contacts}
 %{{{
 
-Resultados cuantitativos de la simulación
-\begin{itemize}
-   \item distribucion del numero de contactos
-   \item distribucion de la duracion de los contactos
-   \item distribucion del tiempo que hay que esperar para el proximo contacto
-\end{itemize}
+%}}}
 
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{growingsphere.pdf}
-   \caption{Schematic representation of the growing communicating
-   sphere, over the space--time diagrams.  In (a) the sphere is
-   growing as the surface of first contact has not reached the maximum
-   distance.  In (b) it has reached the maximum distance, so that it
-   remains at the same size.  After a Doomsday event, the signals can
-   still be observed, but the surface of last contact grows }
-   \label{F_sphere}
-\end{figure}
+\subsection{Waiting time for a first contact}
+%{{{
 
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{abcd.pdf}
-   \caption{Schematic representation of two emitters, $E_1$ and $E_2$,
-   that reach each other at different times.  The time span for $E_i$
-   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
-   emitter $j$ between $C_{ij}$ and $B_{ij}$.
-   The type and length of causal contact in both directions depend on
-   the distance and time lag between the awakening events, the maximum
-   distance that a signal can reach and the time period in which each
-   emitter is active.}
-   \label{F_abcd}
-\end{figure}
+In this section we analyze the distribution of the waiting times for
+a first contact.
+%
+Such distribution can be considered to compute the probability of
+reaching a first contact within the first XX years.
+
+%}}}
+
+\subsection{Reciprocal communication}
+%{{{
+
+%}}}
+
+\subsection{Location within the Galaxy}
+%{{{
+
+%}}}
 
@@
 
-Posibles implicaciones
+\section{Discusion}
+%{{{
+
+Possible implications of the results:
+
 \begin{itemize}
-   \item Análisis del método de comunicación: isotrópico, colimado, accidental
-   \item Naturaleza del transporte del mensaje: radiación electromagnética u otros
-   \item Efectos de fenómenos de alineación (tránsitos)
-   \item Marcadores energéticos
-   \item Marcadores espirituales
-   \item Uso de estrellas como amplificadores o fuentes
+   \item Analysis of the communication method: isotropic, colimated,
+      serendipitous
+   \item Nature of the message carrier: electromagnetic or another
+   \item Effects of alignments
+   \item energetic markers
+   \item Spiritual markers
+   \item Use of stars as sources or amplifiers
 \end{itemize}
-
-%}}}
 
@@
 
+%}}}
+
 \bibliographystyle{mn2e}

### 2020-05-05 submit_v01/manuscript.tex
starting reply to reviewer

\documentclass[crop]{CSLB}

%The authors can define any packages after the \documentclass{CSL} command.
\usepackage{amsmath} %for dealing with mathematics,
\usepackage{amsthm} %for dealing with theorem environments,
\usepackage{cite} %%%for dealing with citations
\usepackage{hyperref} %for linking the cross references
\usepackage{graphics} %for dealing with figures.
\usepackage{algorithmic} %for describing algorithms
\usepackage{subfig} %for getting the subfigures e.g., "Figure 1a and 1b" etc.
\usepackage{url} %It provides better support for handling and breaking URLs.

%\usepackage{natbib}
\usepackage{boites}

%\usepackage{graphicx}
%\usepackage{float}
%\usepackage{psfig}
%\usepackage{amsfonts,amssymb}
%\usepackage{color}
%\usepackage{textcomp}
%\usepackage[T1]{fontenc}
%\usepackage{aecompl}
%\usepackage{times}
%\usepackage{txfonts}
%\usepackage[english]{babel}
%\usepackage[utf8x]{inputenc}
%\usepackage{amsmath}
%\usepackage{caption}
%\usepackage{siunitx}

\jname{Astrobiology}%

\newtheorem{theorem}{Theorem}
\newtheorem{condition}{Condition}

\newcommand{\ceti}{CCN}
\newcommand{\cetis}{CCNs}

\newcommand{\awakening}{A event }
\newcommand{\contact}{C event }
\newcommand{\blackout}{B event }
\newcommand{\doomsday}{D event }

\newcommand{\aawakening}{Awakening event }
\newcommand{\ccontact}{Contact event }
\newcommand{\bblackout}{Blackout event }
\newcommand{\ddoomsday}{Doomsday event }

\newcommand{\awakenings}{A events }
\newcommand{\contacts}{C events }
\newcommand{\blackouts}{B events }
\newcommand{\doomsdays}{D events }

\newcommand{\ff}{{\color{red}\rule{10pt}{10pt}}}

%\newcommand{\ffn}[1]{ \colorbox{red}{FIG {#1}} }    
\newcommand{\ffn}[1]{}

%\newcommand{\ttn}[1]{ \colorbox{green}{TABLE {#1}} }    
\newcommand{\ttn}[1]{}

\input{captions}

\begin{document}

\supertitle{Research Paper}

\title[Probabilities of SETI contacts]{Probability of causal
contact between interstellar civilizations through MonteCarlo simulations}

\author[Lares, Funes \& Gramajo]{Lares M.$^{1, 2}$, Funes J.~G.$^{1,
3}$ \& Gramajo L.$^{1, 2}$}

\corres{\name{Lares M} \email{marcelo.lares@unc.edu.ar}}

\address{
   \add{1}{CONICET, Argentina}

   \add{2}{Universidad Nacional de C\'ordoba, Observatorio
           Astron\'omico de C\'ordoba, Argentina}

   \add{3}{Universidad Cat\'olica de C\'ordoba, Argentina}
}

\keywords{SETI, Computer simulations, Statistics}

%\JELclassification{XX; XX}
%\MSCcodes{XX; XX}

\Abbreviations{SETI: Search for Extraterrestrial Intelligence,\\
CCN: causally connected node,\\
SFC: surface of first contact,\\
SLC: surface of last contact,\\
DE: Discrete Event,\\
GHZ: Galactic Habitable Zone}

\begin{abstract}
%
The abundance of intelligent civilizations in the Galaxy is a longstanding
question, often conceptualized as the problem of the lack of received
communication. Most efforts on the estimation of a number of civilizations are
focused on the Drake equation, although its factors are affected by large
uncertainties and it lacks a temporal nature.  Alternative approaches use
detailed numerical simulations of the galaxy and recipes for the stars and
planets formation rates and for the origin of life, and rely on highly
uncertain parameters. We present a statistical model for the abundance and
duration of civilizations, implemented through Monte Carlo simulations.  We
explore the hypothesis space of the model by a suite of simulations for the
emergence of communicating nodes, and analyze their
causal connections.  This is based on minimal assumptions and three
free parameters, with focus on the statistical properties of empirical
models instead of an interpretable model with variables to be determined by
observation. The analysis of the rate of causal contacts as a function of the
mean number of civilizations, the mean lifespan distribution and the maximum
distance a civilization can send signals is presented, and used as a framework
to discuss the spatial and temporal structure of a populated galaxy within
several scenarios. We find that, given the large distances involved,
causal contacts between civilizations are very rare. Additionally, the odds to
make a contact in a few years of monitoring are low for most models, with the
exception of those of a galaxy densely populated with long lived
civilizations. The probability of causal contacts increases with the
lifetime of civilizations much more significantly than with the number of
active civilizations for a given time window. Finally, we show that the
maximum probability of making a contact occurs when a civilization acquires
the required communication technology.
%
\end{abstract}

\maketitle

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Motivations}\label{S_motivations}
%{{{

The Drake equation \citep{drake_intelligent_1962} offers a very
helpful educated guess, a rational set of lenses --the factors in the
equation-- through which to look at future contacts with a
technologically advanced civilization in the Milky Way.
%
The equation quantifies the number of civilizations from whom we might
receive an electromagnetic signal, using a number of factors that have
been widely discussed in the literature and whose estimates are
updated constantly.
%
A comprehensive review and analysis of each term of the equation is
presented in \citet{vakoch_drake_2015}.
%
Some optimistic estimates from the Drake equation contrast with the so
called Fermi paradox, a statement about the contradiction between the
apparent abundance of life in the Galaxy and its lack of evidence
\citep[e.g. ][]{hart_explanation_1975, brin_great_1983,
barlow_galactic_2012, forgan_galactic_2016, vanhouten_isthere_2017,
Sotos_biotechnology_2019, carroll_nellemback_fermi_2019}.
%
There are many proposals aimed at solving this paradox, which make use
of statistical \citep{solomonides_probabilistic_2016,
vanhouten_isthere_2017, horvat_calculating_2007,
maccone_statistical_2015} or stochastic approaches
\citep{forgan_numerical_2009, bloetscher_using_2019,
glade_stochastic_2011, forgan_numerical_2010}.
%
Regarding the Drake equation, analytical interpretations
\citep{prantzos_joint_2013,
smith_broadcasting_2009} or reformulations
\citep[][and references therein]{burchell_whither_2006} have also been
proposed.
%
The lack of contacts could be explained by the difficulty of life to
appear due to astrophysical explanations
\citep{annis_astrophysical_1999}.
%
In general, also more speculative proposals
\citep{barlow_galactic_2013, lampton_information_2013,
conway_three_2018, forgan_galactic_2017} have been discussed.
% 
\Fpagebreak
%
We stress the fact that the only observation that can be stated with
certainty is that for the number of years projects for the Search of
Extraterrestrial Intelligences (SETI) have been working, we have not
received any signal using the technical possibilities and the
conditions established by SETI projects \citep{tarter_search_2001}.
%
The lack of detection has also inspired to a number of alternative
proposals for new SETI strategies \citep{forgan_exoplanet_2017,
balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010,
tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016,
maccone_SETI_2011,  lingam_relative_2019, wright_theGsearch_2015,
maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018,
forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
%
The Drake equation is a key tool to organize the discussion around the
problem of the abundance of civilizations in the Galaxy
\citep{hinkel_interdisciplinary_2019}.
%
However, the uncertainties in the factors of this equation make it
less prone to a formal application in order to define searching
strategies or to compute the actual number of extraterrestrial
intelligences.
%
The optimistic estimations from the Drake equation also contrast with
the SETI initiatives, noticeable overshadowed by the fact that despite
decades of effort, not a single extraterrestrial intelligence, if any,
has been detected.
%
\citet{prantzos_joint_2013} proposes a unified framework for a joint
analysis of the Drake equation and the Fermi paradox, concluding that
for sufficiently long-lived civilizations, colonization of the Galaxy
is the only reasonable option to gain knowledge about other life
forms.
%
\citet{haqq-misra_drake_2017} discuss the dependence of the Drake
equation parameters on the spectral type of the host star and the time
since the Galaxy formed and examine trajectories for the emergence of
communicative civilizations over the history of the Galaxy.
%
Some modifications to the original idea of the equation have tried to
imprint a stochastic nature, to propose a probabilistic approach, or
to consider the temporal structure which is missing in the equation.
%
Temporal aspects of the distribution of communicating civilizations
and their contacts have also been explored by several authors
\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011}, as
well as efforts on considering the stochastic nature of the Drake
equation \citep{glade_stochastic_2011}.
%
The simulation approach has also been considered
\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017,
ramirez_new_2017}, although with a similar problem that consists on
the large number of parameters which are either unknown or largely
uncertain.
%
\citet{grimaldi_signal_2017} consider a statistical model for the
probability of the Earth making contact with other intelligent
civilizations, taking into account the finite lifetime of signal
emitters, and based on the fractional volume occupied by all signals
reaching our planet.

In a recent work, \citep{balbi_impact_2018} use a statistical model to
analyze the occurrence of causal contacts between civilizations in the
Galaxy.
%
The author highlights the effect of evolutionary processes when
attempting to estimate the number of communicating civilizations that
might be in causal contact with an observer located on the Earth.
%
\citet{cirkovic_temporal_2004} also emphasize the lack of temporal
structure in the Drake equation.
%
\citet{bloetscher_using_2019} uses a probabilistic approach, though
still strongly motivated by the Drake equation, to obtain a
probabilistic measure for the number of civilizations in the Galaxy.
%
To that end, perform MonteCarlo Markov Chains over each factor of the
Drake equation, and combine the means to obtain a probabilistic
result.
%
It is worth mentioning that the author proposes a log-normal target
distribution to compute the posterior probabilities.
%
%This approach, which assigns probability distributions to the Drake
%factors.
%
This study concludes that there is a very low probability for Earth to
make contact with another extraterrestrial civilization.    
%
\citet{smith_broadcasting_2009} uses an analytical model to compute
the probabilities of contact between two randomly located
civilizations, and the waiting time for the first contact, assuming a
fixed maximum broadcasting distance.
% 
Using a different approach, \citet{balbi_impact_2018} investigate the
chance of communicating civilizations making causal contact within a
volume surrounding the location of the Earth, using a statistical
model to compute the occurrence of causal contacts.
%
The author stresses the fact that the causal contact requirement
involves the distance between civilizations, their lifespan and their
times of appearance.  This is important since the time it takes the
light to travel across the Galaxy is much lesser than its lifetime.
%
\citet{balbi_impact_2018} fix the total number of civilizations and
explore the parameter space that comprise three parameters, namely,
the distance to the Earth, the time of appearance and the lifespan of
the communicating civilization.
%
Each of these three variables are drawn from a random distribution.
The distribution of the distances results from a uniform distribution
of civilization within the plane of the Galaxy.
%
For the distribution of the characteristic time of appearance, the
author explores and exponential and a truncated Gaussian
distributions. For the distribution of the lifespans, the author
chooses an exponential distribution.
%
%The radius of influence is set to 1000 lyr.
%
\citet{balbi_impact_2018} finds that the fraction of communicating
civilizations vary with the choice of the statistical distribution for
the time of appearance. For all analyzed distributions, the fraction
corresponding to a mean lifetime of 10$^9$ years and for a maximum
radius for the detection of the signal of 1000 yr, is roughly 0.1.
%
In another probabilistic approach,
\citet{vukotic_astrobiological_2012} propose a probabilistic cellular
automata (PCA) modeling. In this framework, a complex system is
modeled by a lattice of cells which evolve at discrete time steps,
according to transition rules that take into account for each cell the
states of its neighbor cells.
%
The authors implement a PCA model to a network of cells which
represent life complexity on a 2-dimensional annular ring resembling
the Galactic Habitable Zone (GHZ). 
%
The authors set the GHZ between a minimum radius of 6 kpc and a
maximum radius of 10 kpc. Within this framework, the authors also make
several MonteCarlo simulations and analyze ensemble-averaged results.
%
This work aims at analyzing the evolution of life, although it does
not account for the network of causal contacts among technological
civilizations.

In this work we address the problem of the temporal and spatial
structure of the distribution of communicating civilizations, by
exploring the hypothesis space over a minimal set of parameters.
%
We propose to avoid the frequentist approach of the
Drake equation and to explore a parameter space, where instead of
computing a final number, we provide a statistical distribution that
gives conditional probabilities.
%
This is an exploratory analysis that aims at providing a numerical
tool to discuss not the several theoretical problems summarized by the
Drake equation factors, but the different scenarios on the basis of
statistical heuristics.
%
The approach proposed here should be considered as a compromise
between the uncertainties of the frequentist approach and the detailed
recipes required on the simulation approach.
%
We then provide a numerical framework to explore via simulations a
parameter space of unknown observables in order to discuss different
scenarios and their consequences in terms of the probability of
making contacts.
%
The method we use for simulating a stochastic process is an
approximation that allows to study the behavior of complex systems,
by considering a sequence of well defined discrete events.
%
The simulation is carried out by following all the variables that
describe and constitute the state of the system, and the evolution of
the process is described as a set of changes in that state.
%
In this context, an event produces a specific change in the state,
that can be triggered by random variables that encode the stochastic
nature of the physical phenomenon.
%
For example, when a new contact is produced between two entities in
the simulated galaxy, the numbers of active communication lines and of
active communicated nodes change.
%
The process involves following the changes on the state of the
system, defining the initial and final states, defining a method that
allows to keep track of the time progress in steps, and maintaining a
list of relevant events.
%
In Sec.~\ref{S_methods} we introduce the methods and discuss the
candidate distributions for the statistical aspects of the times
involved in the communication process.
%
Then, we present our results in Sec~\ref{S_results}, with special
emphasis on the statistical distributions of the duration of causal
contacts in one or both directions, the possible differences on the
position of a civilization on the Galaxy and the distribution of time
intervals for the waiting of the first contact, always as a function
of the three simulation parameters.
%
Finally, in Sec.~\ref{S_discussion} we discuss our results and future
research directions.

%}}}

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Methods and working hypotheses}\label{S_methods}
%{{{

Simulations are suitable tools to analyze systems that evolve with
time and involve randomness.
%
An advantage of simulations compared to theoretical approaches, is
that the former usually require less assumptions and simplifications,
and can be applied to systems where a theoretical model can not be
found.
%
In particular, many complex stochastic processes that can be described
by the evolution of the state of a system, can be efficiently modeled
with the discrete--event (hereafter, DE) simulation approach.
%
A system described with the DE paradigm is characterized by a set of
actors and events, where actors interact causally through a series of
events on a timeline and process these events in chronological order
\citep{ptolemaeus_system_2014, chung_simulation_2003,
ross_simulation_2012}.
%
Each event yields modifications to the values of the variables that define
the process, producing the corresponding change in the state of the
simulated system.
%
This method is well suited for the particular case of the diffusion of
intelligent signals in the galaxy, and allows to explore several
models easily.
%
We simulate the statistical properties of a set of points in space and
time that have a causal connection at light speed, hereafter dubbed
''Causal Contact Nodes'' (hereafter, \ceti{}).
%
We chose this generic name in order to stress the fact that in this
analysis no broadcasting or lookout activities are required, and
therefore not actual messages are considered, only the causal contact.
%
For the special case of a fully efficient node that emits and receives
isotropically, a causal contact node can be considered as a
broadcasting station that has the ability to detect signals through an
active lookout program.
%
Also, it is worth mentioning that this is a general approach, and not
necessarily a \ceti{} is the host of an intelligent civilization.
%
For instance, it can be associated with a planet where life has developed, became
intelligent, reached the skills required to find the right
communication channel, sustain a search and establish a contact.
%
Alternative message processing entities could be considered, for
example interstellar beacons where life has ceased to exist but
continue with its emission, or communication stations established by
probes or left intentionally by intelligent beings \citep[see, e.g.,
][]{peters_outer_2018, barlow_galactic_2013}.
%
In principle, these strategies could affect our results since it would
be easier to configure a cluster of \cetis{} that spread in time.
%
However, we do not consider these highly speculative alternatives at
this point.
%
For the purpose of this analysis, only the communication capability is
relevant, since we study the causal contacts between the locations.
%
The system is defined by a number of actors that represent \cetis{}
and appear at different instants in time, generating events that
produce meaningful changes in the variables that describe the system,
i.e., in the arrangement of \cetis{} and their network of causal
contacts.
%
For example, the appearance of a new \ceti{} in a region which is
filled with a signal emitted by another node, will increase the
number of active \cetis{} and the number of pairs of \cetis{} in
causal contact.
%
Assuming some simple hypotheses, the discrete events method can be
performed taking into account a small number of variables, which allow
to analyze the variation of the results in the model parameter space.

In what follows, we describe the experimental setup chosen to estimate
the probabilities of causal contacts and several derived quantities in
terms of three independent parameters, namely, the mean time span
between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$,
their mean lifetime $\left<\tau_s\right>$, and the maximum distance a
signal can be detected by another \ceti{} ($D_{max}$).
%
Intuitively, the shorter the $\tau_a$ parameter and the larger the
$\tau_s$ and $D_{max}$ parameters are, the greater the probability of
the existence of causal contacts between pairs of \cetis{} would be.
%
We also need to propose theoretical distribution functions for both
the distribution of lifespans ($\tau_s$) and the distribution of the
number of \cetis{} per unit time \citep{maccone_evolution_2014,
Sotos_biotechnology_2019}, related to the time span between the
appearance of consecutive \cetis{} (since when $\left<\tau_a\right>$
is shorter, it produces a greater density of \cetis{}).
%
The shapes of these distributions are set to a fixed law, as discussed
in Sec.~\ref{SS_PDF_shape}.

%}}}

\begin{figure}[!t]
   \centering
   \includegraphics[width=0.5\textwidth]{F_scheme.pdf}
   \usecaption{F_scheme} \label{F_scheme}
\end{figure}

%{{{

\ffn{1} We illustrate in the Fig.~\ref{F_scheme} the schematic
representation of the region causally connected to a given \ceti{}.
%
To this end, we use space--time diagrams, where time is represented on
the horizontal axis, and space is represented in the vertical axis.
%
For the sake of simplicity, we show in the plot the light travel
distance, i.e., the distance traversed by any signal spreading from
the emitter at light speed (for example, any electromagnetic signal).
%
In this scheme, a light pulse would follow a trajectory represented by
a line at 45 degrees from the axes, given the units of the time and
space axes are years and light years, respectively.
%
Panel (a) of the Figure represents the region $R$ (shaded polygon) in
space and time which is causally connected to the emitter (left
vertex), following an event (hereafter A type event) in which the node
acquires the communication capability, or becomes ''active''.
%
The sphere of first contact (SFC) is a sphere centered on the emitter,
represented in the Figure by the left angle of the region $R$.
%
This sphere grows until its radius reaches the D$_{max}$ distance, in
which it would be no longer detectable due to the decrease in the
energy per unit area, which falls under the assumed detectability
threshold.
%
Similarly, the surface of last contact (SLC) is another sphere that
grows from an event in which the nodes ceases to possess the
communication capability (D type event), and carries the last signal
produced by the emitter.
%
The region which is causally connected to the emitter is then limited
by these two spheres, and has the shape of a sphere (before the D
event) or of a spherical shell (after the D event).
%
The temporal intervals for the communication between two \cetis{} are
represented in the panel (b) of the Figure.
%
In this scheme, the entire lifetime of a \ceti{} can then be
represented by a polygon, limited in the spatial axis by the maximum
distance of the signal $D_{max}$ and in the time axis by the time span
between the starting time (A event, concave vertex) and the ending
time (D event, convex vertex).
%
This representation allows to easily visualize the important events
that result from the existence and communications of two \cetis{}.
%
The points in time where the \cetis{} acquire their communicating
capacity are dubbed ''Awakening'' events ($A_1$ and $A_2$ in the
Figure).
%
Similarly, the points in time where the \cetis{} loss their
communicating capacity are dubbed ''Doomsday'' events ($D_1$ and $D_2$
in the Figure).
%
The points in space--time where the first contact is produced for each
one of the \cetis{}, are defined as ''Contact'' events, shown as $C_1$
and $C_2$ in the Figure.
%
Finally, the two points where the contact is lost for each one the
\cetis{}, are denominated ''Blackout'' events, shown as $B_1$ and
$B_2$ in the Figure.
%
The receiver node E$_1$ can listen signals from the emitter node
E$_2$, from the first contact event (t=C$_{12}$) up to the last
contact event (t=B$_{12}$).
%
Similarly, the receiver node E$_2$ can listen signals from the emitter
node E$_1$, from t=C$_{21}$ up to t=B$_{21}$.
%
The time intervals for the open communication channel are determined
by the ''C'' and ''B'' type events.
%
It is important to notice that there is a time delay between the
contact events of the two \cetis{} involved in this analysis, and also
a time delay between the two blackout events, so that the time range
when a bidirectional contact is possible occurs between the maximum
time of the contact points and the minimum time between the blackout
points.

%}}}

\begin{figure*}
   \centering
   \includegraphics[width=\textwidth]{F_messages.pdf}
   \usecaption{F_messages} \label{F_messages}
\end{figure*}

%{{{

The temporal structure that emerges naturally from the experimental
setup implies that,
as a fundamental property of the simulations,
a causal connection can be produced without the need of the two
involved locations being concurrent or active at the same time.
%
This property of the system arises as a
consequence of the large spatial and temporal scales, where a message
is transmitted at the (relatively small) light speed.
%
Although a \ceti{} could be active for enough time so as to transmit a
message at large distances, the limited power of the message and the
dilution that runs as the squared distance from the source, imposes a
detectability limit.
%
As a consequence of this limitation and of their finite lifetime,
considered as the period of time between the acquisition and loss of
communicating capacity, each \ceti{} will fill a spherical shell
region of the galaxy, limited by two concentric spherical surfaces.
%
The leading front, or surface of first contact (SFC) grows from the
central \ceti{} until it reaches the maximum distance $D_{max}$.
%
Thus, the volume reached causally by a \ceti{} is initially a growing
sphere.
%
Following the end of the civilization, there is still a region which
is filled with the emitted signals.
%
This approach has been also considered other statistical models
\citep[e.g., ][]{smith_broadcasting_2009, grimaldi_signal_2017, Grimaldi2018}.
%
The trailing front, or surface of last contact (SLC) also grows from
the central \ceti{}, with a delay with respect to the SFC equivalent
to the lifetime of the node, and produces a spherical shell region.
%
Any other \ceti{} within this region will be aware of the originating
node, even if it has disappeared before the time of contact.
%
This region will grow if the surface of first contact has not yet
reached the maximum distance $D_{max}$, and will shrink otherwise
until the surface of last contact reaches $D_{max}$, producing as a
result the loss of all signal from the central \ceti{}.
%
In our approach, we consider a model galaxy where the width of the
disk is negligible with respect to the radius of the disk.
%
In the 2D simulation only the intersection of the communicating
spherical shells with the plane of the galaxy is relevant, and produce
the corresponding circles or rings for the filled spheres or annular
regions of the spherical shells, respectively.
%
The initially growing communicating sphere, is shown over space--time
diagrams, where time is represented on the horizontal direction, and
space is represented in the vertical direction.
%
In the Fig.~\ref{F_scheme} the two
emitters in panel (b), E$_1$ and E$_2$, reach each other at different times.
%
The time span for E$_i$ is (A$_i$ , D$_i$) , for $i = {1, 2}$.
%
Emitter $i$ can listen to emitter $j$ between C$_{ij}$ and B$_{ij}$.
%
The type and length of causal contact in both directions depend on the
distance and time lag between the awakening events, the maximum
distance that a signal can reach and the time period in which each
emitter is active.
%
\ffn{2}
%
In the Fig.~\ref{F_messages} we show a schematic representation of the
possible cases in which a \ceti{} E can be in causal contact with a
\ceti{} R.
%
The duration of the causal contact in one direction depends on several
factors, mainly the lifetimes, the time period between the A events
and the distance between the two nodes involved in the causal
contact.
%
As it can be seen, the direction of the causal contact does not imply
any ordering for the A events of the involved nodes.
%
In the implementation of the DE simulation, the six cases shown in the
Figure determine the succession of events.
%
Each time an event is produced, the list of upcoming events is updated
according to these cases.

In out experimental configuration, there is no a temporal window for
the model, i.e., the simulation starts when the process is already
stable, and ends before any galactic effect could modify or affect of
fixed variables.
%
Also, we assume that the probability for the rise of a \ceti{} is
homogeneous over the GHZ.
%
Although the Galaxy has a well known spiral structure, the assumed
sparsity of nodes (intelligent civilizations) does make this
assumption fairly reasonable.
%
Otherwise, if the distribution of nodes is not sparse, it could be the
case that the spiral arms would host most of the \cetis{}, and
contacts are frequent between closely located nodes.
%
In such a case we could expect that the interpretation of our results
could be considered as pessimist for the establishment of a contact
with our planet.
%
We also limit the possibilities of life or other types of
civilizations to the hypothesis usually stated for the definition of
the GHZ \citep{dayal_habitability_2016, gonzalez_galactic_2001,
lineweaver_galactic_2004, gonzalez_habitable_2005,
morrison_extending_2015, haqq-misra_evolution_2019,
rahvar_cosmic_2016, gobat_evolution_2016, rahvar_cosmic_2016} and
consider that habitability remains constant \citep[see, however,
][]{gonzalez_habitable_2005, dayal_habitability_2016, gobat_evolution_2016}.
%
For example, \citet{vukotic_astrobiological_2012} propose
probabilistic cellular automata model to explain the astrobiological
history of the Milky Way.
%
This means that we set aside possible civilizations that could survive
in severe conditions or unstable systems, which would prevent the
appearance of life as we know it.

While the aforementioned assumptions are basic conditions for most of
the stochastic processes observed in nature, we need to make stronger
assumptions related to the nature of the message.
%
The most simple assumption about the message itself, is that it
travels at light speed.
%
With this, we are considering messages sent through, for example,
electromagnetic radiation or even gravitational radiation or neutrino
communication \citep{hippke_interstellar_2017, wright_how_2018}, but we set
apart messages sent with mechanical means or physical objects
\citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown technology
that violates the known laws of physics.
% 
For the communication of \cetis{} through messages sent isotropically,
we assume that the capacity to emit signals and to receive signals
occur at the same time.
%
This means that the ability to find another civilization develops at
the same time that the ability and intention to emit a signal in all
directions, intended to be detected by an unknown civilization.
%
Although there are several reasons to think that this could not be the
case, at large time scales it can be considered that both abilities
occur roughly at coincident epochs.
%
Another essential assumption is that all \cetis{} use the same signal
power, so that there is a maximum distance out to which it can be
detected.
%
It is worth mentioning that we are considering in our experimental
setup a system composed by several emitters and receivers across the
galaxy, under the hypothesis proposed before.
%
In such a system we compute probabilities of a \ceti{} making contact
with another node, and the results must be interpreted as for a
randomly chosen \ceti{}, given certain conditions of homogeneity on an
ensemble of equivalent galaxies.
%
Then, we stress the fact that we are not centered on the probabilities
for the Earth itself, but on a generalized \ceti{}.
%
Regarding the extent of the signals, we know that the distance from
which a signal from Earth could be detected using the current
technology is of a few hundreds light years, given that the signal was
sent to a specific direction.
%
%https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
%https://www.engadget.com/2017/05/25/listening-to-starlight-our-ongoing-search-for-alien-intelligenc/]
%[from https://en.wikipedia.org/wiki/Fermi_paradox]
%
It is straightforward to propose and implement a distribution of
maximum distances, although this would increase the model complexity
at the cost of a larger uncertainty or of another dimension on the
hypothesis space.
%
%This proposal is based on a pronoid scenario, and we do not consider
%in this work the occurrence of paranoid or partially paranoid
%civilizations.
%%
%If that would be the case, the results obtained here can be taken as
%upper limits.

In our simulations we also assume the simplest situations for the
growth of the sphere of first contact.
%
For example, we discard the possibility of stellar colonization
\citep[e.g.][]{newman_galactic_1981, walters_interstellar_1980,
starling_virulence_2013, barlow_galactic_2012, jeong_large_2000,
maccone_mathematical_2011} and assume that the communication aim is
performed to all directions with the same power and the same
probability, i.e., strongly assume isotropic communication in all
cases.
%
More detailed simulations could be produced considering different
efficiency of communication or detection methods.
%
However, in this case we also would complicate the experimental setup,
make the results less evident.
%
The results are independent of the nature of life (organic or
artificial).
%
The lifetime of a civilization can be caused by auto destruction or by
external factors.
%}}}

 \setlength{\tabcolsep}{10pt}
\begin{table*}
\centering
\begin{tabular}{cllll}
\hline
   \multicolumn{2}{l}{independent variables (hypothesis space)}
   &min. value&max. value&Nbins\\
\hline
   $\tau_{a}$ & Mean temporal separation between consecutive
   awakenings & 4000 yr & 200\,000 yr & 50\\ 
   $\tau_{s}$ & Mean lifetime of a \ceti{}
   & 10000 yr & 500\,000 yr& 50\\ 
   D$_{max}$ & Maximum reach of a message  &  \multicolumn{2}{l}{(500,
   1000, 10000, 20000, 40000, 80000) lyr} & 6  \\
\hline
   \multicolumn{2}{l}{fixed variables} & \multicolumn{2}{c}{assumptions} &value \\
\hline
   & statistical properties of all \cetis{} &equally distributed&&\\
   & Point process for the distribution in time & homogeneous &&\\
   f$_s$ & The scan of the sky & fully efficient&&1\\
   f$_p$ & panspermia or colonization &absent&&0\\
   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
   R$_{GHZ}^{min}$   & Inner radius of the GHZ     & & &20\,000 lyr\\
   R$_{GHZ}^{max}$   & Outer radius of the GHZ     &&&60\,000 lyr\\
   t$_{max}$ & Time span of the simulation  &&&1.e6 yr\\
    & \multicolumn{3}{l}{number of random realizations for each point in the hypothesis space} &50\\
\hline
   \multicolumn{2}{l}{discrete events} &
   \multicolumn{3}{l}{affected variables}\\
\hline
   A event & Awakening: a \ceti{} starts its communication
   capabilities &Number of active \cetis{}\\
   B event & Blackout: the end of the communication channel stops
   &Number causal channels\\
   C event & Contact: a new causal contact is produced &Number of causal channels\\
   D event & Doomsday: a \ceti{} ends its communication capabilities&Number of active \cetis{}\\
\hline

\hline
\end{tabular}
\caption{Definition of independent variables and adopted values for 
   fixed parameters 
   that are part of the simulation.  Variable parameters define the
   spatial and temporal structure
   of the process and the maximum reach of the messages.}
\label{T_simu_hypotheses}
\end{table*}

\subsection{Power laws vs. exponential laws}\label{SS_PDF_shape}
%{{{

The temporal structure of the process is defined by two distribution
parameters that represent the mean time interval that a \ceti{}
can emit and receive signals, and the mean time interval
between the emergence of consecutive \cetis{}.
%
The spatial structure of the simulation is given by the size and shape
of the Galactic Habitable Zone and the maximum distance a signal can
travel to be detected (D$_{max}$).
%
The parameters for the temporal distributions also determine the
spatial properties, since the density of active \cetis{} in the galaxy
depend on these two parameters.
%
For example, a small $\tau_a$ and a large $\tau_s$ will produce a
galaxy densely populated with \cetis{}.
%
Also, some hypotheses regarding the shapes of the distributions of the
temporal parameters must be made in order to complete the
simulation.
%
\citet{forgan_spatiotemporal_2011} argues that the times at which
different civilizations become intelligent follow a Gaussian
distribution, and then the distribution of inter-arrival times is an
inverse exponential.
%
We choose to assume that the distribution of the times of A events is a
stationary Poisson process, and then the distribution of the times
between the appearance of new consecutive \cetis{} is exponential.
%
Regarding the duration of a \ceti{}, we propose that the distribution
of the duration of \cetis{} is a stationary exponential distribution.
%
\citet{balbi_impact_2018} stress the fact that it would be desirable
to arrive at a theoretical statistical distribution of the lifespan of
civilizations, preferentially on the basis of the underlying
astrophysical and biological processes.
%
\citet{maccone_lognormals_2014} argues that this distribution should
be a log normal.

The power law and exponential statistical distributions are among the
most common patterns found in natural phenomena.
%
For example, the distribution of the frequency of words in many
languages is known to follow a Zipf's law (which is a power law).
%
This distributions arise any time a phenomena is characterized by
commonly occurring small events and rarely produced large events
\citep[e.g. ][]{adamic_zipf_2000}.
%
Zipf law also describes population ranks of cities in various
countries, corporation sizes, income rankings, ranks of number of
people watching the same TV channel.
%
The magnitudes of earthquakes, hurricanes, volcanic eruptions and
floods; the sizes of meteorites or the losses caused by business
interruptions from accidents, are also well described by power laws
\citep{sornette_critical_2006, benguigui_classificacion_2016}.
%
The power law behavior has been observed in a variety of systems,
including for example stock market fluctuations, sizes of computer
files and word frequency in languages \citep{mitzenmacher_brief_2004,
newman_power_2005, simkin_theory_2006}
%
Power laws have also been widely used in biological sciences, e.g., in
analyzing connectivity patterns in metabolic networks
\citep{jeong_large_2000} and in the number of species observed per
unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
% 
More examples can be found in the literature
\citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009,
maccone_evolution_2014, maccone_lognormals_2014,
benguigui_classificacion_2016}

In this work, the exponential distribution of lifespan and waiting
times is justified by considering the hypothesis that the process of
appearance of life in the galaxy is homogeneous and stationary.
%
That is, there is no preferred location within the GHZ for the
spontaneous appearance of life, and the emergence of a \ceti{} is
independent of the existence of previous nodes in the galaxy.
%            
This is equivalent to proposing a Poisson process for the emergence of
\cetis{}, since there is a close relation between the number of events
in time or space and the waiting time or separation, respectively
\citep[e.g., ][]{ross_simulation_2012}.
%
That is, these are two alternative approaches to describing the same
process, a Poisson distribution for the number of events implies an
exponential distribution for their separations, and vice versa.
%
It should be emphasized that the exponential laws used in this work
are assumed as part of the working hypothesis, and instead of
analyzing results from a particular parameter chosen ad--hoc, we
explore the hypothesis space and analyze the impact of the values of
these parameters on the results.
%}}}

\begin{figure} \centering
   \includegraphics[width=0.5\textwidth]{F_number_of_contacts.pdf}
   \usecaption{F_number_of_contacts} \label{F_number_of_contacts}
\end{figure}

\subsection{Complexity of the model}
%{{{

In this Section we discuss the degree of complexity in the model,
considering a compromise between the accuracy of the model and the
number of parameters that are free or with a high uncertainty.
%
Firstly, we emphasize that the odds of a causal contact between two
\cetis{} should not be considered as the odds of a contact between two
intelligent civilizations, and in fact the later could be much lesser
than the former.
%
Indeed, in order to establish a contact between any two entities, a
minimum degree of compatibility must be accomplished without any
previous agreement, making the possibility of a contact with a message
that could be deciphered highly rare \citep[see e.g. ][]{forgan_collimated_2014}.
%
Besides the trade--off between the simplicity and the complexity of
the experiments, further analysis could be performed following this
framework in order to explore possible implications of the results for
more detailed configurations.
%
For example, the communication method (isotropic, collimated,
serendipitous) can affect the observables, making it necessary to
implement a correction factor.
%
Taking this into account, our results regarding the probability of
causal contact should be considered as upper limits for effective
contacts, since they depend on the efficiency of both the emitter and
the receiver to broadcast and scan the sky for intelligent signals,
respectively \citep{grimaldi_signal_2017}.
%
A correction by coverage ratio in the detection and by a targeting
ratio in the emission could be easily implemented in the simulation,
although the effect of reducing the probability of contact is
basically the product of the efficiency ratios and thus such
implementation is not necessary.
%
Therefore, the values of the probabilities could be modified by a
constant correction factor equal to the combined emission/reception
efficiency \citep{smith_broadcasting_2009, anchordoqui_upper_2019,
forgan_collimated_2014} or for beam--like transmissions
\citep{grimaldi_signal_2017}.
%
Other considerations include the effects of alignments, the use of
stars as sources or amplifiers \citep{Edmondson2003, borra_searching_2012}, 
the nature of the message carrier, the
use of a spatial distribution that actually resembles the spiral shape
and the width of the disc of the Galaxy, different distribution
functions for the mean lifetime of \cetis{}, and different
efficiencies for the sphere of the causal region of each \ceti{}
resembling different searching strategies
\citep{hippke_interstellar_2017}.
%
It is also possible to consider that D$_{max}$ is different for
different \cetis{}.  For example a power law where powerful emissions
are rare and weak emissions are common could be an improvement to the
model.
%
In this work we chose not to implement this for the sake of
simplicity.
%
Finally, the role of the message contents could influence on the
lifespan of a \ceti{} that receives a message, although the
implementation of such behavior would increase the number of free
parameters and would be highly speculative.

%}}}

\subsection{Discrete event process}
%{{{

Given a parametric model $M(\tau_a, \tau_s, D_{max})$, which includes
the functional forms of the statistical distributions and fixed values
for the hypotheses, a discrete event simulation is performed by
keeping track of a set of variables that change each time a meaningful
event is produced.
%
The main variables that follow the evolution of the simulation are the
positions of stars, which are sampled randomly within the GHZ, the
time of the awakening of each \ceti{} (A event), and the time of the
vanishing of each \ceti{} (D event).
%
The variables that can be deduced from the previous ones include the
number of \cetis{} in casual contact with at least another \ceti{} at a given
time, the number of \cetis{} as a function of time, the number of
\cetis{}
that receive a message at least one time, the number of \cetis{} that
receive a message at least one time and successfully deliver an answer
and the number distribution of waiting time to receive a message.
%
All these quantities are updated each time one of the four events (A,
B, C, D) is produced.

%%% comparar con el stochsatic cellular automata
% ver libro Solving Fermi paradox, sec. 27.3.1

%}}}

\begin{figure*} % 2D color plot
   \centering
   \includegraphics[width=\textwidth]{F_never_contact.pdf}
   \usecaption{F_never_contact}
   \label{F_never_contact}
\end{figure*}

\begin{figure*} % 2D color plot
   \centering
   \includegraphics[width=\textwidth]{F_C_at_A.pdf}
   \usecaption{F_C_at_A}
   \label{F_C_at_A}
\end{figure*}

\subsection{Model for \cetis{}}
%{{{

Since the method is based on events, the first step is to define an
architecture of events, the relationships between them and how events
trigger changes on the state of the system,
so that the system is updated each time an event is produced.
%
We consider a simulated system that represents the galactic habitable
zone which, on a first approach, is a 2-dimensional annular region.
%
The adopted values for the geometry of the GHZ are 20 kly and 60 kly, for the inner
and outer radius, respectively \citep{lineweaver_galactic_2004}.
% ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
% mas chico...
%
This simple model does not take into account the variations in stellar
density given by the spiral structure.
%
Although there are several possible approaches, we chose to follow the
evolution of the system according to the following events:

\begin{enumerate}
   \item[(A)] Awakening: A new \ceti{} acquires the ability to emit
      and receive messages and starts to actively broadcast and
      lookout for messages.
   \item[(C)] Contact: A new one-way causal contact is established
      (does not require the awareness of the involved \cetis{}.)
   \item[(B)] Blackout: An existing causal contact is interrupted,
      with or without the occurrence of a doomsday event.
   \item[(D)] Doomsday: An old \ceti{} disappears, or halts its
      communication activities.
\end{enumerate}

We have chosen this nomenclature so that it can be easily remembered
from the initial A, B, C, D.
%
The \blackout is produced when one of the two \cetis{} halts in its
capability of emitting and receiving signals.
%
It can occur at the same time the \doomsday or before.
%
Similarly, a \ccontact can be produced at the time of the \aawakening
or later, and there can be several \contacts for the same node.
%
The time marks for each event (A and D for each \ceti{} and C and B
for each causal contact) are stored as a result of the simulation.
%
Also, the list of active \cetis{} is obtained as a function of time.
%
%Some fixed variables must be set in order to carry out the simulation:
%the size of the Galactic Habitable Zone, the mean lifetime of a
%\ceti{}, $\tau_s$, the mean waiting time for the appearance of another
%\ceti{}, $\tau_a$, and the maximum distance a message can be detected.

The Galactic Habitable Zone is set at fixed values of the inner and outer
radii, radial symmetry is assumed, and the with of the galactic disk
is neglected compared to the radial size.
%
The two radii are measured in light years, with the aim to maintain a
single and comprehensive unit for both space and time coordinates. 
%
The simulation was implemented on a python 3 code, which is publicly
available\footnote{https://github.com/mlares/simu\_contact}

%}}}

\begin{figure}
   \centering
   \includegraphics[width=0.5\textwidth]{F_waiting_for_1C.pdf}
   \usecaption{F_waiting_for_1C}
   \label{F_waiting_for_1C}
\end{figure}

\setlength{\tabcolsep}{10pt}
\begin{table*}
\centering
\begin{tabular}{ccccc}
\hline
   Model subset & D$_{max}$ [lyr] & $\tau_a$ interval [yr] & $\tau_s$ interval 
   [yr]& description  \\
\hline
M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
%
M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
%
\hline
\end{tabular}
\caption{Selected models to analyze the behavior of simulation outputs
   and their dependence on simulation parameters.}
\label{T_selected_models}
\end{table*}

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Results: exploring the parameter space}\label{S_results}
%{{{

We implemented the simulation of a regular grid of models varying over
the hypothesis space, which covers 15000 different models.
%
For each model, we simulated 50 realizations with different random
seeds, in order to perform a MonteCarlo estimation of the variance.
%
The parameters for the temporal aspects of the simulation (the mean
waiting time for the next awakening, $\left<\tau_a\right>$, and the
mean lifetime, $\left<\tau_s\right>$) cover the ranges 4000-200000 yr
and 10000-500000 yr, respectively, with a regular partition of 50
values for each parameter.
%
For the D$_{max}$ parameter, we chose the values 500, 1000, 10000,
20000, 40000 and 80000 lyr.
%
This setup makes a total of 750000 simulations, each covering a time
range of one million years.
%
\ttn{1}
%
In Table \ref{T_simu_hypotheses} we show the three variable
parameters, the ranges of their values and the number of bins that
have been explored in the numerical experiments.
%
We also show the set of parameters that take part in the simulation,
their values and the hypotheses that define the runs of the
simulations.

As a product of the simulations, several quantities can be obtained.
%
Some quantities are directly derived from the discrete events, namely,
the ID of emitting and receiving \cetis{}, the position in the galaxy, and
the times of each of the events that are relevant to keep track of the
number of \cetis{} (time of Awakening and time of Doomsday) and of the
number of contacts (the times of contacts and the times of Blackouts).
%
We can also derive the additional quantities that represent the
properties of the \cetis{}, for example the total time elapsed between the
Awakening and the Doomsday of each node.
%
The time span of a \ceti{} listening another or being listened by another
can also be easily derived from the results of a realization.
%
This way we can also compute the distribution in the galaxy of
\cetis{}
that reach contact and the waiting time until the first contact or the
waiting time until the next contact.
%
Regarding the properties of the population of \cetis{} and its evolution,
the simulation yields the fraction of awaken time a node is listening
at least another node (i.e., in their causal cone), the age of
contacted \cetis{} at first contact, the distribution of time to wait
until next contact, the fraction of \cetis{} where the first contact is
given at the awakening, the distribution of the number of contacts for
each \ceti{}, the distribution of the number of contacts as a function of
the age of the node, the number of contacts as a function of time in the galaxy,
the fractions of nodes that succeed in contact, and the distribution of
distances between contacted nodes.
%
Another useful derived quantity is the duration of two-way
communication channels or the fraction of contacts that admit a
response.
% 
It is also possible to analyze the relations between the distance to
\ceti{} vs. the time of double communication, the distance to \ceti{} vs. the
age of contacted node, the age of a node and the maximum number of
contacted nodes before the Doomsday, or the lifespan of a node vs. the
maximum number of contacts.
%
In addition, all these quantities can be analyzed as a function of the
simulation parameters.
%
We chose eight models that are on fairly opposed regions of the
explored parameter space, and cover short/long lifetimes, short/long
waiting times for the next \ceti{} to appear in the galaxy and short/large
range reach of the signal, including all possible combinations.
%
\ttn{2}
%
The details of each model are summarized in the
Table~\ref{T_selected_models}.

%}}}

\subsection{Membership to the network of connected \cetis{}}\label{SS_members}
%{{{

\ffn{3}
%
In Fig.~\ref{F_number_of_contacts} we show the empirical cumulative
distributions of the number contacts for \cetis{} in six different
samples, including short and long lifetimes
$\left(\,\left<\tau_s\right>=\right.$10000~yr and
$\left<\tau_s\right>$=100000~yr, respectively$\left.\right)$, dense
and sparse spatial distribution
$\left(\,\left<\tau_a\right>=\right.$10000~yr and
$\left<\tau_a\right>$= 100000~yr, respectively$\left.\right)$, and two
different signal reach ranges (D$_{max}$=40000 lyr and
D$_{max}$=80000~lyr).
%
The models M1 and M5 are not shown, since in both cases the total
number of contacts is different from zero for less than the five per
cent of the \cetis{}.
%
The areas between models with the same mean lifetime and mean
awakening time are shaded for visualization purposes.
%
As it can be seen from this Figure, the mean lifetime is more
determinant than the mean awakening rate (dense and sparse,
represented by a different shade) for the number of contacts.
%
As expected, of the samples with the longest $D_{max}$ the model with
a dense awakening in the timeline (low $\left<\tau_a\right>$) and a
long lifetime (high $\left<\tau_s\right>$), is the model that
maximizes the number of contacts, reaching a maximum of order 20
contacts for a single \ceti{}.
%
Similarly, for the models with long signal range
($D_{max}=80000\,lyr$), the one which is dense and with long lifetime
has the maximum number of contacts, reaching nearly 100 contacts for
each \ceti{} in its entire lifetime.
%
This is considerably larger than any model with a shorter lifetime,
which produce a number of contacts of at most the order of ten
contacts per \ceti{}.
%
A couple of comments are worth to mention about this plot, namely, it
has a logarithmic scale on the x-axis, and it is the cumulative, not
differential, empirical distribution, so that the differences in the
number of contacts for different models is quite large.
%
Another interesting feature observed in this Figure, is that for the
models analyzed, there are \cetis{} that have no causal contacts with
any other \ceti{} during their whole lifetime.
%
The fraction of \cetis{} with no contact ranges from nearly 20\% up to
nearly 90\% (models M1 and M5, not shown), except for the dense and
long lifetime models, for which almost all \cetis{} make at least one
contact.
%
Since these results depend on just eight models, we analyze the
properties of the frequency of contacts as a function of both the mean
lifetime and the mean awakening time.
%
\ffn{4}
%
On Fig.~\ref{F_never_contact} we show 2D color maps with the fraction
of \cetis{} in the simulations that never make contact (i.e., never
listen to another \ceti{}), as a function of the mean lifetime
($\left<\tau_s\right>$, in the range 0--500 yr) and the mean awakening
time ($\left<\tau_a\right>$, in the range 0--200 yr), for three
different values of the maximum signal range, $D_{max}=10000$~yr (left
panel), $D_{max}=40000$~yr (middle panel), and $D_{max}=80000$~yr
(right panel).
%
A clear pattern emerges in this Figure, showing that the probability
for a \ceti{} of making causal contact with at least another \ceti{}
during their entire lifetime, increases with increasing $D_{max}$,
increasing $\left<\tau_s\right>$ and decreasing $\left<\tau_a\right>$,
following a roughly linear dependence with the three parameters.

While the number of \cetis{} that do not succeed in reaching the
causal contact regions of other node is a useful indicator of
degree of isolation, there is also the chance that a number of
\cetis{} are already in the causal contact region of other nodes,
which shows a somewhat opposite situation.
%
\ffn{5}
%
The fraction of \cetis{} that make the first contact at the awakening
event is shown in the Fig.~\ref{F_C_at_A}, 
as a function of the mean lifetime
($\left<\tau_s\right>$, in the range 0--500 yr) and the mean awakening
time ($\left<\tau_a\right>$, in the range 0--200 yr), for three
different values of the maximum signal range, $D_{max}=10000$~yr (left
panel), $D_{max}=40000$~yr (middle panel), and $D_{max}=80000$~yr
(right panel).
%
These probabilities are consistent with the ones on the previous
Figure.
%
Although they are not complementary, it is clear that the dependence
with the three parameters is roughly linear in both cases.
%
%% DISCUTIR ESTOS RESUOLTADOS SEGUN LA LITERATURA

%

%}}}

\subsection{Waiting time for a first contact}\label{SS_waiting}
%{{{

In this subsection we analyze the distribution of the waiting times for
a first contact.
%
Such distribution can be considered to compute the probability for a
random \ceti{} to carry out a SETI project and spend a given time until
the first contact is made.

\ffn{6}
%
In the Fig.~\ref{F_waiting_for_1C} we show the histograms of the mean
waiting times for the first contact, for several models.
%
Upper panel (a) panel shows the histograms for several values of the
mean survival time $\tau_s$, and the mean awakening time $\tau_a$ in
the range 68000-100000 yr.
%
Lower panel (b) shows the histograms for several values of $\tau_a$,
and $\tau_s$ in the range 17000-24000 yr.
%
As it can be seen, the shape of the histograms is almost the same when
$\tau_a$ is modified (panel (b)), but significantly changes when
$\tau_s$ is modified (panel (a)). 
%
For different values of $\tau_a$, the shape of the distribution is
similar, and the main difference is given by the different number of
contacts on the different models.
%
Conversely, for different values of $\tau_a$, the shape of the
distribution is similar, and the main difference given by the
different number of contacts on the different models.

The fraction of the \cetis{} that have made at least one contact as a
function of the elapsed time since the awakening, is computed with
respect to the total number of nodes that make causal contact at least
one time in the time range from the Awakening up to the Doomsday
events.
%
From a frequentist approach, the total counts from the awakening
($t=$0~yr) up to a given time are related to the estimation of the
probability of listening during that time interval and reaching to
causal contact region of at least another node.
%
The complement of this value is the probability of observing in a time
interval with no success, i.e., without ever happening a Contact
event.
%
Clearly, this probability diminishes with time and tends to zero for
large time periods.
%             
The diferential probability of making a causal contact reach its
maximum at the awakening event for all models.
%
That is, for a short period of time like the nearly 50 years SETI
programs have been active, the initial moment is the most promising
for making a (causal) contact, at least for a given technology.
%
This consideration offers a new approach to SETI programs, where the
search for new communication technologies has a fundamental role.
%
Moreover, as it can be seen in the Fig.~\ref{F_waiting_for_1C}, there
is a maximum probability of making the first causal contact at nearly
100000 years.
%
Remarkably, the probability of such a contact does not grows linearly
with the observing time, but is greater during the first 100
millennia.

%}}}

%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
\section{Discussion}\label{S_discussion}
%{{{

We have implemented a suite of simulations following a stochastic
approach, to explore the hypothesis space of a simple model that
accounts for the causal connections between communicating
civilizations on a simplified galaxy.
%
The different models can be generated with a minimal number of free
parameters and some fixed assumptions.
%
We argue that three parameters can be used to describe a variety of
situations, ranging from a galaxy model where an intelligent civilization is
very rare, to a galaxy model populated with plenty of civilizations in
causal contact.
%
With this approach, we can estimate several quantities as a function
of the parameters on the hypothesis space,
so we explored the outcomes of different models 
that arise as by-products of the simulations.
%
For instance, it is possible to estimate the timescale for an
uninterrupted SETI effort to reach success, in terms of different
model parameters that reflect different, so far unknown, scenarios for
the appearance of intelligent life on the Galaxy.

Although the simulations use a number of speculative assumptions, we
argue that the degree of knowledge about the origin and persistence of
life in the Galaxy does not worth the implementation of more detailed
or sophisticated models.
%
Instead, we take advantage of the simplicity of the model to explore
the hypothesis space, i.e., the three--dimensional parameter space of
the parametric model, in order to gain insight on the consequences of
different scenarios for the search of intelligent life.
%
Our analysis is not centered in obtaining the odds for the Earth
to make contact with another intelligent civilization.
%
Instead, we focus on obtaining a statistical, parameter dependent
description of the possible properties of the communication networks
that comprise sets of nodes with broadcasting and reception
capabilities.
%
This causally connected nodes are sparsely distributed in both space
and time, making it difficult the analytical treatment and justifying
the simulation approach.

Under the hypotheses of our experiments, we conclude that a causal
contact is extremely unlikely unless the galaxy is heavily populated
by intelligent civilizations with large average lifetimes.
%
Roughly, in order to have at least one contact in the entire lifetime
of the \ceti{}, there should appear a mean of at least one \ceti{}
every $\sim$20000 yr, with a mean active period of at least
$\sim$30000 yr.
%
This result is qualitatively similar to the results presented by
several authors, which state that a contact between the Earth and
another intelligent civilization in the Galaxy is quite unlikely,
provided the maximum distance of the signal and the lifetime of the
emitter are not large enough.
%
This analysis supports the idea that, in order to increase the
possibilities of a contact, more active strategies of the emitter
would be required.
%
Some proposals in this direction include intertellar exploration,
colonization and settlement \citep{brin_great_1983, Dosovic2019,
galera_invasion_2019}, although it would require large temporal
scales.
%
\citet{Dosovic2019} use probabilistic cellular automaton simulations
to explore the parameter space of a model with colonization and
catastrophic events.
%
According to the timescales involved, the results could explain the
Fermi paradox.
%
Although our work does not take into account the colonization
hypothesis, it does consider catastrophic events implicitly in the
mean value and distribution of the lifetime, $\left<\tau_s\right>$.
%
Other strategies could also increase the probability of contacts, for
example panspermia \citep[e.g.,][]{starling_virulence_2013} or
self--replicating probes \citep[e.g.,][]{barlow_galactic_2013},
although they would be too slow to make a significant impact on the
communication network among intelligent civilizations.
%
Our results are also consistent with those presented by
\citet{grimaldi_signal_2017}, who estimate an upper bound for the mean
number of extraterrestrial civilizations that could contact Earth
using Monte Carlo simulations, from a statistical model where the with
of the Galactic disk is not negligible.
%
On the other hand, taking into account the probability of contact with
radio signals not sent intentionally \citep{horvat_calculating_2007},
our analysis would imply a highly populated solar neighborhood.
%
Remarkably, most of the studies using statistical models or
simulations are based on the Drake equation
\citep{cirkovic_temporal_2004, smith_broadcasting_2009,
bloetscher_using_2019}.
%
Our approach does not relay on the Drake equation, and thus is not
motivated by the detailed physical processes that give rise to
intelligent life.
%
However, we argue that it is a valid empirical formulation to discuss
the probabilities of contact, and mainly time scales involed in the
problem.
%
Ours is an alternative to the method proposed by
\citet{balbi_impact_2018}, who perform an analysis based on the Earth
and use a different model, although our results agree in general.

%%%%%%%%%%%%%%%

There is a balance between the density of \cetis{} and the mean
lifetime since, as expected, a lower density can be compensated by a
longer active time period.
%
However, a large number of nodes does not easily compensate their
short lives to reach the same probability of causal contact than in
the case of a less populated galaxy but with very ancient
civilizations.
%
In all cases, for a short period of time (for instance, the time SETI
programs have been active on Earth), the maximum probability of making
a contact occurs at the moment of the awakening.
%
This suggests the possibility that an alternative SETI strategy could
be the search for alternative message carriers, for the case in which
the search has not been performed on the adequate channels.
%
Then, if a contact is produced for the first time, the origin of the
signal is more likely to be very old.
%
Also, the chances of entering the causal connected zone of a \ceti{}
does not grow linearly, but favours the first period of 100000 years.

In the approach of this work, we used computer simulations to address
the problem of the probabilities of causal contacts between location
in the galaxy the possibility of sending and receiving messages.
%
Instead of making a number of assumptions, we have explored the
hypothesis space, reducing the problem to only three parameters and a
few simple hypothesis to perform a complete model for the population and
communication network in the galaxy.
%
This allows to consider the Fermi paradox from a new perspective, and
to propose an alternative treatment for the number of intelligent
emitter/receivers with emphasis on the temporal dimension which limits
the probabilities of contacts due to the short time interval for
rise and fall of civilizations compared to the age and extension of
our Galaxy.

%}}}

\ack[Acknowledgement]{
%
This work was partially supported by the Consejo Nacional de
Investigaciones Cient\'{\i}ficas y T\'ecnicas (CONICET, Argentina),
the Secretar\'{\i}a de Ciencia y Tecnolog\'{\i}a, Universidad Nacional
de C\'ordoba, Argentina, and the Universidad Cat\'olica de C\'ordoba,
Argentina.
%
This research has made use of NASA's Astrophysics Data System.
%
Plots and simulations were made with software developed by the authors
using R and python languajes. Plots were postprocessed with inkscape.
%
}

\ack[Disclosure statement]{
%
No competing financial interests exist.
%
}

%\bibliographystyle{mn2e}
\setlength{\bibsep}{0.0pt}

\begin{thebibliography}{90}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{Adamic(2000)}]{adamic_zipf_2000}
Adamic L., 2000, {Zipf, Power-laws, and Pareto - a ranking tutorial}. web

\bibitem[{Anchordoqui \& Weber(2019)}]{anchordoqui_upper_2019}
Anchordoqui L.~A., Weber S.~M., 2019, arXiv e-prints

\bibitem[{Annis(1999)}]{annis_astrophysical_1999}
Annis J., 1999, JBIS - Journal of the British Interplanetary Society, 52, 19

\bibitem[{Armstrong \& Sandberg(2013)}]{Armstrong2013}
Armstrong S., Sandberg A., 2013, Acta Astronautica, 89, 1

\bibitem[{Balb(2018)}]{balb_spatiotemporal_2018}
Balb A., 2018, Mem. della Soc. Astron. Ital., 89, 425

\bibitem[{Balbi(2018)}]{balbi_impact_2018}
Balbi A., 2018, Astrobiology, 18, 54

\bibitem[{Barab{\'{a}}si(2009)}]{barabasi_scale_2009}
Barab{\'{a}}si A.~L., 2009, Science (80-. )., 325, 412

\bibitem[{Barlow(2013{\natexlab{a}})}]{barlow_galactic_2012}
Barlow M.~T., 2013{\natexlab{a}}, Int. J. Astrobiol., 12, 63

\bibitem[{Barlow(2013{\natexlab{b}})}]{barlow_galactic_2013}
Barlow M.~T., 2013{\natexlab{b}}, Int. J. Astrobiol., 12, 63

\bibitem[{Benguigui \& Marinov(2016)}]{benguigui_classificacion_2016}
Benguigui L., Marinov M., 2016, {A classification of the natural and social
  distributions Part 2: the explanations}. Tech. rep.

\bibitem[{Bloetscher(2019)}]{bloetscher_using_2019}
Bloetscher F., 2019, Acta Astronaut., 155, 118

\bibitem[{Borra(2012)}]{borra_searching_2012}
Borra E.~F., 2012, Astronomical Journal, 144

\bibitem[{Brin(1983)}]{brin_great_1983}
Brin G., 1983, Quarterly Journal of the Royal Astronomical Society, 24, 283

\bibitem[{Burchell(2006)}]{burchell_whither_2006}
Burchell M.~J., 2006, Int. J. Astrobiol., 5, 243

\bibitem[{Carroll-Nellenback {et~al}\mbox{.}(2019)Carroll-Nellenback, Frank,
  Wright, \& Scharf}]{carroll_nellemback_fermi_2019}
Carroll-Nellenback J., Frank A., Wright J., Scharf C., 2019, The Astronomical
  Journal, 158, 117

\bibitem[{Chung(2003)}]{chung_simulation_2003}
Chung C.~A., 2003, {Simulation Modeling Handbook: A Practical Approach
  (Industrial and Manufacturing Engineering Series)}. CRC press

\bibitem[{{\'{C}}irkovi{\'{c}}(2004)}]{cirkovic_temporal_2004}
{\'{C}}irkovi{\'{c}} M.~M., 2004, Astrobiology, 4, 225

\bibitem[{{Conway Morris}(2018)}]{conway_three_2018}
{Conway Morris} S., 2018, Int. J. Astrobiol., 17, 287

\bibitem[{Dayal {et~al}\mbox{.}(2016)Dayal, Ward, \&
  Cockell}]{dayal_habitability_2016}
Dayal P., Ward M., Cockell C., 2016, arXiv e-prints

\bibitem[{Do{\v{s}}ovi{\'{c}} {et~al}\mbox{.}(2019)Do{\v{s}}ovi{\'{c}},
  Vukoti{\'{c}}, \& {\'{C}}irkovi{\'{c}}}]{Dosovic2019}
Do{\v{s}}ovi{\'{c}} V., Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2019,
  Astronomy and Astrophysics, 625

\bibitem[{Drake(1962)}]{drake_intelligent_1962}
Drake F., 1962

\bibitem[{Edmondson \& Stevens(2003)}]{Edmondson2003}
Edmondson W.~H., Stevens I.~R., 2003, International Journal of Astrobiology, 2,
  231

\bibitem[{Enriquez {et~al}\mbox{.}(2017)Enriquez, Siemion, Foster, Gajjar,
  Hellbourg, Hickish, Isaacson, Price, Croft, DeBoer, Lebofsky, MacMahon, \&
  Werthimer}]{enriquez_breakthrough_2017}
Enriquez J.~E. {et~al.}, 2017, Astrophys. J., 849, 104

\bibitem[{Fogg(1987)}]{fogg_temporal_1987}
Fogg M.~J., 1987, Icarus, 69, 370

\bibitem[{Forgan {et~al}\mbox{.}(2016)Forgan, Dayal, Cockell, \&
  Libeskind}]{forgan_evaluating_2015}
Forgan D., Dayal P., Cockell C., Libeskind N., 2016, Int. J. Astrobiol., 16, 60

\bibitem[{Forgan(2009)}]{forgan_numerical_2009}
Forgan D.~H., 2009, Int. J. Astrobiol., 8, 121

\bibitem[{Forgan(2011)}]{forgan_spatiotemporal_2011}
Forgan D.~H., 2011, Int. J. Astrobiol., 10, 341

\bibitem[{Forgan(2013)}]{forgan_possibility_2013}
Forgan D.~H., 2013, JBIS - J. Br. Interplanet. Soc., 66, 144

\bibitem[{Forgan(2014)}]{forgan_collimated_2014}
Forgan D.~H., 2014, JBIS - J. Br. Interplanet. Soc., 67, 232

\bibitem[{Forgan(2017{\natexlab{a}})}]{forgan_galactic_2016}
Forgan D.~H., 2017{\natexlab{a}}, Int. J. Astrobiol., 16, 349

\bibitem[{Forgan(2017{\natexlab{b}})}]{forgan_galactic_2017}
Forgan D.~H., 2017{\natexlab{b}}, Int. J. Astrobiol., 16, 349

\bibitem[{Forgan(2019)}]{forgan_exoplanet_2017}
Forgan D.~H., 2019, Int. J. Astrobiol., 18, 189

\bibitem[{Forgan \& Rice(2010)}]{forgan_numerical_2010}
Forgan D.~H., Rice K., 2010, Int. J. Astrobiol., 9, 73

\bibitem[{Frank(2009)}]{frank_common_2009}
Frank S.~A., 2009, {The common patterns of nature}

\bibitem[{Funes {et~al}\mbox{.}(2019)Funes, Florio, Lares, \&
  Asla}]{funes_searching_2019}
Funes J.~G., Florio L., Lares M., Asla M., 2019, Theol. Sci., 0, 1

\bibitem[{Galera {et~al}\mbox{.}(2019)Galera, Galanti, \&
  Kinouchi}]{galera_invasion_2019}
Galera E., Galanti G.~R., Kinouchi O., 2019, International Journal of
  Astrobiology, 18, 316

\bibitem[{Glade {et~al}\mbox{.}(2012)Glade, Ballet, \&
  Bastien}]{glade_stochastic_2011}
Glade N., Ballet P., Bastien O., 2012, Int. J. Astrobiol., 11, 103

\bibitem[{Gobat \& Hong(2016)}]{gobat_evolution_2016}
Gobat R., Hong S.~E., 2016, Astron. Astrophys., 592

\bibitem[{Gonzalez(2005)}]{gonzalez_habitable_2005}
Gonzalez G., 2005, Orig. Life Evol. Biosph., 35, 555

\bibitem[{Gonzalez {et~al}\mbox{.}(2001)Gonzalez, Brownlee, \&
  Ward}]{gonzalez_galactic_2001}
Gonzalez G., Brownlee D., Ward P., 2001, Icarus, 152, 185

\bibitem[{Grimaldi(2017)}]{grimaldi_signal_2017}
Grimaldi C., 2017, Sci. Rep., 7, 46273

\bibitem[{Grimaldi {et~al}\mbox{.}(2018)Grimaldi, Marcy, Tellis, \&
  Drake}]{Grimaldi2018}
Grimaldi C., Marcy G.~W., Tellis N.~K., Drake F., 2018, Publications of the
  Astronomical Society of the Pacific, 130

\bibitem[{Haqq-Misra(2019)}]{haqq-misra_evolution_2019}
Haqq-Misra J., 2019, arXiv e-prints

\bibitem[{Haqq-Misra \& Kopparapu(2018)}]{haqq-misra_drake_2017}
Haqq-Misra J., Kopparapu R.~K., 2018, Habitability of the Universe Before
  Earth, 307

\bibitem[{Harp {et~al}\mbox{.}(2018)Harp, Ackermann, Astorga, Arbunich,
  Barrios, Hightower, Meitzner, Barott, Nolan, Messerschmitt, Vakoch, Shostak,
  \& Tarter}]{harp_application_2018}
Harp G.~R. {et~al.}, 2018, Astrophys. J., 869, 66

\bibitem[{Hart(1975)}]{hart_explanation_1975}
Hart M.~H., 1975, Quarterly Journal of the Royal Astronomical Society, 16, 128

\bibitem[{Hinkel {et~al}\mbox{.}(2019)Hinkel, Hartnett, Lisse, \&
  Young}]{hinkel_interdisciplinary_2019}
Hinkel N., Hartnett H., Lisse C., Young P., 2019, Bull. Am. Astron. Soc., 51,
  497

\bibitem[{Hippke(2017)}]{hippke_interstellar_2017}
Hippke M., 2017, arXiv e-prints

\bibitem[{Horvat(2006)}]{horvat_calculating_2007}
Horvat M., 2006, Int. J. Astrobiol., 5, 143

\bibitem[{Horvat {et~al}\mbox{.}(2011)Horvat, Naki{\'{c}}, \&
  Oto{\v{c}}an}]{horvat_impact_2011}
Horvat M., Naki{\'{c}} A., Oto{\v{c}}an I., 2011, Int. J. Astrobiol., 11, 51

\bibitem[{Jeong {et~al}\mbox{.}(2000)Jeong, Tombor, Albert, Oltvai, \&
  Barabasi}]{jeong_large_2000}
Jeong H., Tombor B., Albert R., Oltvai Z.~N., Barabasi A.~L., 2000, Nature,
  407, 651

\bibitem[{Lampton(2013)}]{lampton_information_2013}
Lampton M., 2013, Int. J. Astrobiol., 12, 312

\bibitem[{Lineweaver {et~al}\mbox{.}(2004)Lineweaver, Fenner, \&
  Gibson}]{lineweaver_galactic_2004}
Lineweaver C.~H., Fenner Y., Gibson B.~K., 2004, Science (80-. )., 303, 59

\bibitem[{Lingam \& Loeb(2018)}]{lingam_relative_2019}
Lingam M., Loeb A., 2018, Astrobiology, 19, 28

\bibitem[{Loeb {et~al}\mbox{.}(2016)Loeb, Batista, \&
  Sloan}]{loeb_relative_2016}
Loeb A., Batista R.~A., Sloan D., 2016, J. Cosmol. Astropart. Phys., 2016, 040

\bibitem[{Loeb \& Zaldarriaga(2007)}]{loeb_eavesdropping_2006}
Loeb A., Zaldarriaga M., 2007, J. Cosmol. Astropart. Phys., 1

\bibitem[{Maccone(2010)}]{maccone_KLT_2010}
Maccone C., 2010, Acta Astronaut., 67, 1427

\bibitem[{Maccone(2011)}]{maccone_mathematical_2011}
Maccone C., 2011, Orig. Life Evol. Biosph., 41, 609

\bibitem[{MacCone(2011)}]{maccone_SETI_2011}
MacCone C., 2011, Acta Astronaut., 68, 63

\bibitem[{MacCone(2013)}]{maccone_SETI_2013}
MacCone C., 2013, Int. J. Astrobiol., 12, 218

\bibitem[{Maccone(2014{\natexlab{a}})}]{maccone_evolution_2014}
Maccone C., 2014{\natexlab{a}}, Int. J. Astrobiol., 13, 290

\bibitem[{Maccone(2014{\natexlab{b}})}]{maccone_lognormals_2014}
Maccone C., 2014{\natexlab{b}}, Acta Astronaut., 105, 538

\bibitem[{Maccone(2015)}]{maccone_statistical_2015}
Maccone C., 2015, Acta Astronaut., 115, 277

\bibitem[{Mart{\'{i}}n \& Goldenfeld(2006)}]{martin_origin_2006}
Mart{\'{i}}n H.~G., Goldenfeld N., 2006, Proc. Natl. Acad. Sci. U. S. A., 103,
  10310

\bibitem[{Mitzenmacher(2004)}]{mitzenmacher_brief_2004}
Mitzenmacher M., 2004, Internet Math., 1, 226

\bibitem[{Morrison \& Gowanlock(2015)}]{morrison_extending_2015}
Morrison I.~S., Gowanlock M.~G., 2015, Astrobiology, 15, 683

\bibitem[{Murante {et~al}\mbox{.}(2015)Murante, Monaco, Borgani, Tornatore,
  Dolag, \& Goz}]{murante_simulating_2015}
Murante G., Monaco P., Borgani S., Tornatore L., Dolag K., Goz D., 2015, Mon.
  Not. R. Astron. Soc., 447, 178

\bibitem[{Newman(2005)}]{newman_power_2005}
Newman M., 2005, Contemp. Phys., 46, 323

\bibitem[{Newman \& Sagan(1981)}]{newman_galactic_1981}
Newman W.~I., Sagan C., 1981, Icarus, 46, 293

\bibitem[{Peters(2018)}]{peters_outer_2018}
Peters T., 2018, Int. J. Astrobiol., 17, 282

\bibitem[{Prantzos(2013)}]{prantzos_joint_2013}
Prantzos N., 2013, Int. J. Astrobiol., 12, 246

\bibitem[{Ptolemaeus(2014)}]{ptolemaeus_system_2014}
Ptolemaeus C., 2014, {System Design, Modeling, and Simulation. Using Ptolemy
  II}. Ptolemy.org, Berkeley, p. 690

\bibitem[{Rahvar(2017)}]{rahvar_cosmic_2016}
Rahvar S., 2017, Mon. Not. R. Astron. Soc., 470, 3095

\bibitem[{Ramirez {et~al}\mbox{.}(2017)Ramirez, G{\'{o}}mez-Mu{\~{n}}oz,
  V{\'{a}}zquez, \& N{\'{u}}{\~{n}}ez}]{ramirez_new_2017}
Ramirez R., G{\'{o}}mez-Mu{\~{n}}oz M.~A., V{\'{a}}zquez R., N{\'{u}}{\~{n}}ez
  P.~G., 2017, International Journal of Astrobiology, 1

\bibitem[{Ross(2012)}]{ross_simulation_2012}
Ross S.~M., 2012, {Simulation}. Elsevier Science Publishing Co Inc

\bibitem[{Simkin \& Roychowdhury(2006)}]{simkin_theory_2006}
Simkin M.~V., Roychowdhury V.~P., 2006, J. Math. Sociol., 30, 33

\bibitem[{Smith(2009)}]{smith_broadcasting_2009}
Smith R.~D., 2009, Int. J. Astrobiol., 8, 101

\bibitem[{Solomonides {et~al}\mbox{.}(2016)Solomonides, Kaltenegger, \&
  Terzian}]{solomonides_probabilistic_2016}
Solomonides E., Kaltenegger L., Terzian Y., 2016, 228th AAS, San Diego, 228, 1

\bibitem[{Sornette(2006)}]{sornette_critical_2006}
Sornette D., 2006, {Critical phenomena in natural sciences: Chaos, fractals,
  selforganization and disorder: Concepts and tools}. Springer

\bibitem[{Sotos(2019)}]{Sotos_biotechnology_2019}
Sotos J.~G., 2019, {Biotechnology and the lifetime of technical civilizations}

\bibitem[{Starling \& Forgan(2014)}]{starling_virulence_2013}
Starling J., Forgan D.~H., 2014, Int. J. Astrobiol., 13, 45

\bibitem[{Tarter(2001)}]{tarter_search_2001}
Tarter J., 2001, Annu. Rev. Astron. Astrophys., 39, 511

\bibitem[{Tarter {et~al}\mbox{.}(2009)Tarter, Backus, Blair, Cordes, Harp,
  Henry, Horowitz, Howard, Kilsdonk, Korpela, Lasio, Levin, Shostack, \&
  Werthimer}]{tarter_advancing_2009}
Tarter J. {et~al.}, 2009, Astro2010 Astron. Astrophys. Decad. Surv. Sci. White
  Pap. no. 294, 2010

\bibitem[{Vakoch \& Dowd(2015)}]{vakoch_drake_2015}
Vakoch D.~A., Dowd M.~F., 2015, {The drake equation: Estimating the prevalence
  of extraterrestrial life through the ages}, Vakoch D.~A., Dowd M.~F., eds.
  Cambridge University Press, Cambridge, pp. 1--319

\bibitem[{VanHouten(2018)}]{vanhouten_isthere_2017}
VanHouten M.~A., 2018, Lancet Gastroenterol. Hepatol., 3, 382

\bibitem[{Vukoti{\'{c}} \&
  {\'{C}}irkovi{\'{c}}(2012)}]{vukotic_astrobiological_2012}
Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2012, Orig. Life Evol. Biosph.,
  42, 347

\bibitem[{Vukoti{\'{c}} {et~al}\mbox{.}(2016)Vukoti{\'{c}}, Steinhauser,
  Martinez-Aviles, {\'{C}}irkovi{\'{c}}, Micic, \&
  Schindler}]{vukotic_grandeur_2016}
Vukoti{\'{c}} B., Steinhauser D., Martinez-Aviles G., {\'{C}}irkovi{\'{c}}
  M.~M., Micic M., Schindler S., 2016, Mon. Not. R. Astron. Soc., 459, 3512

\bibitem[{Walters {et~al}\mbox{.}(1980)Walters, Hoover, \&
  Kotra}]{walters_interstellar_1980}
Walters C., Hoover R.~A., Kotra R., 1980, Icarus, 41, 193

\bibitem[{Wright {et~al}\mbox{.}(2015)Wright, Cartier, Zhao, Jontof-Hutter, \&
  Ford}]{wright_theGsearch_2015}
Wright J.~T., Cartier K. M.~S., Zhao M., Jontof-Hutter D., Ford E.~B., 2015,
  Astrophys. J., 816, 17

\bibitem[{Wright {et~al}\mbox{.}(2018)Wright, Kanodia, \&
  Lubar}]{wright_how_2018}
Wright J.~T., Kanodia S., Lubar E., 2018, The Astronomical Journal, 156, 260

\end{thebibliography}

\end{document}

### 2020-05-05 submit_v02/manuscript.tex
starting reply to reviewer

--- 2020-05-05 submit_v01/manuscript.tex
@@
 
-\jname{Astrobiology}%
+\jname{International Journal of Astrobiology}%
 
@@
 \newcommand{\ttn}[1]{}
-
-\input{captions}
 
@@
    \includegraphics[width=0.5\textwidth]{F_scheme.pdf}
-   \usecaption{F_scheme} \label{F_scheme}
+   \caption{
+   %
+Space--time diagrams showing a schematic representation of the
+different stages in the development of causal contact node (\ceti{}).
+%
+Panel (a) represents the region $R$ in space and time which is causally
+connected to the emitter (left vertex), following an "A" type event
+("Awakening") in which the node acquires the communication capability.
+%
+The sphere of first contact (SFC) is a sphere centered on the emitter
+that grows until its radius reaches the D$_{max}$ distance, in which
+the power of a signal would equal the detectability threshold.
+%
+In the Figure this sphere is represented by the left triangle of the
+region $R$.
+%
+The surface of last contact (SLC) is another sphere that grows from a
+"D" event ("Doomsday").
+%
+The region which is causally connected to the emitter is then limited
+by these two spheres, and has the shape of a sphere or of a spherical
+shell, depending on the time.
+%
+The temporal intervals for
+the communication between two \cetis{} are represented in he
+panel (b).
+%
+The receiver \ceti{} E$_1$ can listen signals from the emiter \ceti{} E$_2$,
+from the "Contact" event (t=C$_{12}$) up to the "Blackout" event (t=B$_{12}$).
+%
+Similarly, 
+the receiver \ceti{} E$_2$ can listen signals from the emiter \ceti{} E$_1$,
+from t=C$_{21}$ up to t=B$_{21}$. 
+%
+   } \label{F_scheme}
 \end{figure}
@@
    \includegraphics[width=\textwidth]{F_messages.pdf}
-   \usecaption{F_messages} \label{F_messages}
+   \caption{
+   %   
+Schematic representation on space--time diagrams of the possible cases in which a
+\ceti{} E (Emiter, solid line) can be in causal contact with a \ceti{} R
+(Receiver, dashed line).
+%
+Causal contact can be produced either if the emiter \ceti{} appears before
+(left) or after (right) the receiver \ceti{}.
+%
+The causal contact in which the receiver can have access to the signal 
+from the emiter is produced between the C event (Contact, open circles) and the B
+event (Blackout, filled circles).
+%
+The duration of the causal contact in one direction depends on
+several factors, mainly the time lag between the awakening events
+(horizontal separation in the space--time diagrams)
+and the distance between the \cetis (vertical separation in the
+space--time diagrams).
+%   
+The different configurations must be taken into account in order to 
+obtain the list of the upcomming events.
+%
+The time intervals for a two--way communication to be possible are
+indicated in double lines. 
+%
+   } \label{F_messages}
 \end{figure*}
@@
    \includegraphics[width=0.5\textwidth]{F_number_of_contacts.pdf}
-   \usecaption{F_number_of_contacts} \label{F_number_of_contacts}
+   \caption{
+   %
+Empirical cumulative distributions of the number contacts
+for \cetis in four different samples (M5, M6, M7, M8) with two
+values for the range of the signal reach: D$_{max}$=40000 ly (blue),  and 
+D$_{max}$=80000 ly (orange).
+%
+Except for the models corresponding to a densely populated galaxy with
+long--living \cetis, the \cetis which receive more than 10 contacts
+are very unlikely.
+%
+See Table \ref{T_selected_models} for model descriptions.
+%
+   } \label{F_number_of_contacts}
 \end{figure}
@@
    \includegraphics[width=\textwidth]{F_never_contact.pdf}
-   \usecaption{F_never_contact}
+   \caption{
+   %
+Rate of MPLs that never make contact (listening) as a
+function of $\tau_a$ and $\tau_s$, for 
+$D_{max}$=10000 (left panel),
+$D_{max}$=40000 (middle panel), and
+$D_{max}$=80000 (right panel)
+%
+   }
    \label{F_never_contact}
@@
    \includegraphics[width=\textwidth]{F_C_at_A.pdf}
-   \usecaption{F_C_at_A}
+   \caption{
+   %
+Rate of MPLs that listen at the moment of the
+awakening, as a funcion of $\tau_a$ and $\tau_s$, for
+$D_{max}$=10000 (left panel),
+$D_{max}$=40000 (middle panel), and
+$D_{max}$=80000 (right panel)
+%
+   }
    \label{F_C_at_A}
@@
    \includegraphics[width=0.5\textwidth]{F_waiting_for_1C.pdf}
-   \usecaption{F_waiting_for_1C}
+   \caption{
+   %
+Histograms of the mean waiting times for the first contact, for
+several models.
+%
+Left panel shows the histograms for several values of $\tau_a$, and
+$\tau_s$ in the range 17000-24000 yr.
+%
+Right panel shows the histograms for several values of $\tau_s$, and
+$\tau_a$ in the range 68000-100000 yr.
+%
+The shape of the histograms does not change when $\tau_a$ is
+modified, but significantly changes when $\tau_s$ is modified. 
+%
+   }
    \label{F_waiting_for_1C}

### 2020-05-05 submit_v03/manuscript.tex
starting reply to reviewer

--- 2020-05-05 submit_v02/manuscript.tex
@@
 %
-The abundance of intelligent civilizations in the Galaxy is a longstanding
-question, often conceptualized as the problem of the lack of received
-communication. Most efforts on the estimation of a number of civilizations are
-focused on the Drake equation, although its factors are affected by large
-uncertainties and it lacks a temporal nature.  Alternative approaches use
-detailed numerical simulations of the galaxy and recipes for the stars and
-planets formation rates and for the origin of life, and rely on highly
-uncertain parameters. We present a statistical model for the abundance and
-duration of civilizations, implemented through Monte Carlo simulations.  We
-explore the hypothesis space of the model by a suite of simulations for the
-emergence of communicating nodes, and analyze their
-causal connections.  This is based on minimal assumptions and three
-free parameters, with focus on the statistical properties of empirical
-models instead of an interpretable model with variables to be determined by
-observation. The analysis of the rate of causal contacts as a function of the
-mean number of civilizations, the mean lifespan distribution and the maximum
-distance a civilization can send signals is presented, and used as a framework
-to discuss the spatial and temporal structure of a populated galaxy within
-several scenarios. We find that, given the large distances involved,
-causal contacts between civilizations are very rare. Additionally, the odds to
-make a contact in a few years of monitoring are low for most models, with the
-exception of those of a galaxy densely populated with long lived
-civilizations. The probability of causal contacts increases with the
-lifetime of civilizations much more significantly than with the number of
-active civilizations for a given time window. Finally, we show that the
-maximum probability of making a contact occurs when a civilization acquires
+The abundance of intelligent civilizations in the Galaxy is a
+   longstanding question, often conceptualized as the problem of the
+   lack of received communication. Estimates of the number of
+   civilizations are generally guided by the Drake equation, despite
+   the large uncertainties in its factors and its lack of a temporal
+   nature. Alternative approaches use detailed numerical simulations
+   of the galaxy and recipes for the stars and planets formation
+   rates and for the origin of life, and rely on uncertain parameters.
+   We present a statistical model for the abundance and duration of
+   civilizations implemented through Monte Carlo simulations. We
+   explore the hypothesis space of the model by a suite of simulations
+   to analyze the emergence of communicating nodes and their causal
+   connections. We based this on minimal assumptions and three free
+   parameters, with focus on the proposed statistical properties of
+   empirical models. Our analysis of the dependence of the rate of
+   causal contacts on the mean number of civilizations, the mean
+   lifespan distribution and the maximum distance a civilization can
+   send signals, is considered to discuss the spatial and temporal
+   structure of a populated galaxy within several scenarios. We find
+   that, given the enormous distances involved, causal contacts
+   between civilizations are very rare. The odds to make a contact in
+   a few years of monitoring are low for most models, except for those
+   of a galaxy densely populated with long-lived civilizations. The
+   probability of causal contacts increases with the lifetime of
+   civilizations much more significantly than with the number of
+   active civilizations for a time window. We show that the maximum
+   probability of making a contact occurs when a civilization gains
    the required communication technology.
@@
 
-The Drake equation \citep{drake_intelligent_1962} offers a very
-helpful educated guess, a rational set of lenses --the factors in the
-equation-- through which to look at future contacts with a
-technologically advanced civilization in the Milky Way.
+The Drake equation  \citep{drake_intelligent_1962} provides a truly
+helpful educated guess, a rational set of lenses –the factors in the
+equation– through which to look at future contacts with
+technologically advanced civilizations in the Milky Way. 
 %
 The equation quantifies the number of civilizations from whom we might
-receive an electromagnetic signal, using a number of factors that have
-been widely discussed in the literature and whose estimates are
-updated constantly.
-%
-A comprehensive review and analysis of each term of the equation is
+receive an electromagnetic signal, using a collection of factors that
+have been extensively discussed in the literature and whose estimates
+are revised continually. 
+%
+A comprehensive review and analysis of each term of the equation are
 presented in \citet{vakoch_drake_2015}.
 %
-Some optimistic estimates from the Drake equation contrast with the so
-called Fermi paradox, a statement about the contradiction between the
-apparent abundance of life in the Galaxy and its lack of evidence
-\citep[e.g. ][]{hart_explanation_1975, brin_great_1983,
+Optimistic estimates from the Drake equation contrast with the
+so-called Fermi paradox, which states the contradiction between the
+expected abundance of life in the Galaxy and the lack of evidence for
+it \citep[e.g. ][]{hart_explanation_1975, brin_great_1983,
 barlow_galactic_2012, forgan_galactic_2016, vanhouten_isthere_2017,
@@
 %
-There are many proposals aimed at solving this paradox, which make use
-of statistical \citep{solomonides_probabilistic_2016,
+There are many propositions aimed at solving this paradox, which make
+use of statistical \citep{solomonides_probabilistic_2016,
 vanhouten_isthere_2017, horvat_calculating_2007,
@@
 Regarding the Drake equation, analytical interpretations
-\citep{prantzos_joint_2013,
-smith_broadcasting_2009} or reformulations
+\citep{prantzos_joint_2013, smith_broadcasting_2009} or reformulations
 \citep[][and references therein]{burchell_whither_2006} have also been
@@
 % 
+% 
+% 
+% 
+% 
+% 
+% 
+% 
+% 
+% 
+% 
 \Fpagebreak
 %
-We stress the fact that the only observation that can be stated with
-certainty is that for the number of years projects for the Search of
-Extraterrestrial Intelligences (SETI) have been working, we have not
-received any signal using the technical possibilities and the
-conditions established by SETI projects \citep{tarter_search_2001}.
-%
-The lack of detection has also inspired to a number of alternative
-proposals for new SETI strategies \citep{forgan_exoplanet_2017,
-balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010,
-tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016,
-maccone_SETI_2011,  lingam_relative_2019, wright_theGsearch_2015,
-maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018,
-forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
-%
-The Drake equation is a key tool to organize the discussion around the
-problem of the abundance of civilizations in the Galaxy
-\citep{hinkel_interdisciplinary_2019}.
-%
-However, the uncertainties in the factors of this equation make it
-less prone to a formal application in order to define searching
-strategies or to compute the actual number of extraterrestrial
-intelligences.
-%
-The optimistic estimations from the Drake equation also contrast with
-the SETI initiatives, noticeable overshadowed by the fact that despite
-decades of effort, not a single extraterrestrial intelligence, if any,
-has been detected.
-%
-\citet{prantzos_joint_2013} proposes a unified framework for a joint
-analysis of the Drake equation and the Fermi paradox, concluding that
-for sufficiently long-lived civilizations, colonization of the Galaxy
-is the only reasonable option to gain knowledge about other life
-forms.
-%
-\citet{haqq-misra_drake_2017} discuss the dependence of the Drake
-equation parameters on the spectral type of the host star and the time
-since the Galaxy formed and examine trajectories for the emergence of
-communicative civilizations over the history of the Galaxy.
-%
-Some modifications to the original idea of the equation have tried to
-imprint a stochastic nature, to propose a probabilistic approach, or
-to consider the temporal structure which is missing in the equation.
-%
-Temporal aspects of the distribution of communicating civilizations
-and their contacts have also been explored by several authors
-\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
-balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011}, as
-well as efforts on considering the stochastic nature of the Drake
-equation \citep{glade_stochastic_2011}.
-%
-The simulation approach has also been considered
-\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
-murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017,
-ramirez_new_2017}, although with a similar problem that consists on
-the large number of parameters which are either unknown or largely
-uncertain.
-%
-\citet{grimaldi_signal_2017} consider a statistical model for the
-probability of the Earth making contact with other intelligent
-civilizations, taking into account the finite lifetime of signal
-emitters, and based on the fractional volume occupied by all signals
-reaching our planet.
-
-In a recent work, \citep{balbi_impact_2018} use a statistical model to
-analyze the occurrence of causal contacts between civilizations in the
-Galaxy.
-%
-The author highlights the effect of evolutionary processes when
-attempting to estimate the number of communicating civilizations that
-might be in causal contact with an observer located on the Earth.
-%
-\citet{cirkovic_temporal_2004} also emphasize the lack of temporal
-structure in the Drake equation.
-%
-\citet{bloetscher_using_2019} uses a probabilistic approach, though
-still strongly motivated by the Drake equation, to obtain a
-probabilistic measure for the number of civilizations in the Galaxy.
-%
-To that end, perform MonteCarlo Markov Chains over each factor of the
-Drake equation, and combine the means to obtain a probabilistic
-result.
-%
-It is worth mentioning that the author proposes a log-normal target
-distribution to compute the posterior probabilities.
-%
-%This approach, which assigns probability distributions to the Drake
-%factors.
-%
-This study concludes that there is a very low probability for Earth to
-make contact with another extraterrestrial civilization.    
-%
-\citet{smith_broadcasting_2009} uses an analytical model to compute
-the probabilities of contact between two randomly located
-civilizations, and the waiting time for the first contact, assuming a
-fixed maximum broadcasting distance.
-% 
-Using a different approach, \citet{balbi_impact_2018} investigate the
-chance of communicating civilizations making causal contact within a
-volume surrounding the location of the Earth, using a statistical
-model to compute the occurrence of causal contacts.
-%
-The author stresses the fact that the causal contact requirement
-involves the distance between civilizations, their lifespan and their
-times of appearance.  This is important since the time it takes the
-light to travel across the Galaxy is much lesser than its lifetime.
-%
-\citet{balbi_impact_2018} fix the total number of civilizations and
-explore the parameter space that comprise three parameters, namely,
-the distance to the Earth, the time of appearance and the lifespan of
-the communicating civilization.
-%
-Each of these three variables are drawn from a random distribution.
-The distribution of the distances results from a uniform distribution
-of civilization within the plane of the Galaxy.
-%
-For the distribution of the characteristic time of appearance, the
-author explores and exponential and a truncated Gaussian
-distributions. For the distribution of the lifespans, the author
-chooses an exponential distribution.
+We stress the fact that the only observation that can be stated with certainty is that for the number of years projects for the Search of Extraterrestrial Intelligences (SETI) have been working; we have received no signal using the technical possibilities and the conditions established by them \citep{tarter_search_2001}.
+%
+The absence of detection has also motivated alternative ideas for new SETI strategies \citep{forgan_exoplanet_2017, balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010, tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016, maccone_SETI_2011, lingam_relative_2019, wright_theGsearch_2015, maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018, forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
+%
+The Drake equation is a key tool to organise the discussion about the problem of the (unknown) abundance of civilisations in the Galaxy \citep{hinkel_interdisciplinary_2019}.
+%
+However, the uncertainties in the factors of this equation make it less prone to a formal study to define searching strategies or to compute the actual number of extraterrestrial intelligences.
+%
+The optimistic estimations from the Drake equation also contrast with the SETI initiatives, noticeable overshadowed because despite decades of effort, not a single evidence for intelligent signals has been detected.
+%
+\citet{prantzos_joint_2013} proposes a unified framework for a joint analysis of the Drake equation and the Fermi paradox, concluding that for sufficiently long-lived civilisations, colonisation of the Galaxy is the only reasonable option to gain awareness about other life forms.
+%
+\citet{haqq-misra_drake_2017} discuss the dependence of the Drake equation parameters on the spectral type of the host stars and the time since the Galaxy formed and examine trajectories for the emergence of communicative civilisations.
+%
+Some modifications to the original idea of the equation have tried to imprint a stochastic nature, to propose a probabilistic approach, or to consider the temporal structure missing in the equation.
+%
+Temporal aspects of the distribution of communicating civilisations and their contacts have also been explored by several authors \citep{fogg_temporal_1987, forgan_spatiotemporal_2011, balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011}, as well as efforts on considering the stochastic nature of the Drake equation \citep{glade_stochastic_2011}.
+%
+The literature about this topic brings a number of works that consider numerical simulations
+\citep{forgan_evaluating_2015, vukotic_grandeur_2016, murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017, ramirez_new_2017}.
+%
+Although this approach does not rely on observed quantities that could be modified with time, it also depend on the definition of a number of parameters which are unknown or uncertain.
+%
+\citet{grimaldi_signal_2017} consider a statistical model for the probability of the Earth contacting other intelligent civilisations, taking into account the finite lifetime of signal emitters, and based on the fractional volume occupied by all signals reaching our planet.
+
+In a recent work, \citep{balbi_impact_2018} use a statistical model to analyse the occurrence of causal contacts between civilisations in the Galaxy.
+%
+The author highlights the effect of evolutionary processes when attempting to estimate the number of communicating civilisations that might be in causal contact with an observer on the Earth.
+%
+\citet{cirkovic_temporal_2004} also emphasise the lack of temporal structure in the Drake equation.
+%
+\citet{bloetscher_using_2019} uses a probabilistic approach, though still motivated by the Drake equation, to obtain a probabilistic measure for the number of civilisations in the Galaxy.
+%
+To that end, the authors do Monte-Carlo Markov Chains over each factor of the Drake equation, and combine the means to reach a probabilistic result.
+%
+It is worth mentioning that the author proposes a log-normal target distribution to compute the posterior probabilities.
+%
+This study concludes that there is a tiny probability for Earth to contact another extraterrestrial civilization.    
+%
+\citet{smith_broadcasting_2009} uses an analytical model to gauge the probabilities of contact between two randomly located civilisations, and the waiting time for the first contact, assuming a fixed maximum broadcasting distance.
+% 
+Adopting a novel approach, \citet{balbi_impact_2018} investigate the chance of communicating civilisations making causal contact within a volume surrounding the location of the Earth, using a statistical model to compute the occurrence of causal contacts.
+%
+The author stresses the fact that the causal contact requirement involves the distance between civilisations, their lifespan and their times of appearance.
+This is important since the time it takes the light to travel across the Galaxy is much lesser than its lifetime.
+%
+\citet{balbi_impact_2018} fix the total number of civilisations and explore the parameter space that comprise three parameters, namely, the distance to the Earth, the time of appearance and the lifespan of the communicating civilization.
+%
+Each of these three variables are drawn from a random distribution. The distribution of the distances results from a uniform distribution of civilizations within the plane of the Galaxy.
+%
+For the distribution of the characteristic time of appearance, the author explores an exponential and a truncated Gaussian distributions.
+For the distribution of the lifespans, the author chooses an exponential distribution.
+
+%-------------------------------------------------------------------
+%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%-------------------------------------------------------------------
+
+%%%%% SIGUE PUNTO Y SEGUIDO
 %

### 2020-05-05 tmp.tex
starting reply to reviewer

--- 2020-05-05 submit_v01/manuscript.tex
@@
-\documentclass[crop]{CSLB}
-
-%The authors can define any packages after the \documentclass{CSL} command.
-\usepackage{amsmath} %for dealing with mathematics,
-\usepackage{amsthm} %for dealing with theorem environments,
-\usepackage{cite} %%%for dealing with citations
-\usepackage{hyperref} %for linking the cross references
-\usepackage{graphics} %for dealing with figures.
-\usepackage{algorithmic} %for describing algorithms
-\usepackage{subfig} %for getting the subfigures e.g., "Figure 1a and 1b" etc.
-\usepackage{url} %It provides better support for handling and breaking URLs.
-
-%\usepackage{natbib}
-\usepackage{boites}
-
-%\usepackage{graphicx}
-%\usepackage{float}
-%\usepackage{psfig}
-%\usepackage{amsfonts,amssymb}
-%\usepackage{color}
-%\usepackage{textcomp}
-%\usepackage[T1]{fontenc}
-%\usepackage{aecompl}
-%\usepackage{times}
-%\usepackage{txfonts}
-%\usepackage[english]{babel}
-%\usepackage[utf8x]{inputenc}
-%\usepackage{amsmath}
-%\usepackage{caption}
-%\usepackage{siunitx}
-
-\jname{Astrobiology}%
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{condition}{Condition}
-
-\newcommand{\ceti}{CCN}
-\newcommand{\cetis}{CCNs}
-
-\newcommand{\awakening}{A event }
-\newcommand{\contact}{C event }
-\newcommand{\blackout}{B event }
-\newcommand{\doomsday}{D event }
-
-\newcommand{\aawakening}{Awakening event }
-\newcommand{\ccontact}{Contact event }
-\newcommand{\bblackout}{Blackout event }
-\newcommand{\ddoomsday}{Doomsday event }
-
-\newcommand{\awakenings}{A events }
-\newcommand{\contacts}{C events }
-\newcommand{\blackouts}{B events }
-\newcommand{\doomsdays}{D events }
-
-\newcommand{\ff}{{\color{red}\rule{10pt}{10pt}}}
-
-%\newcommand{\ffn}[1]{ \colorbox{red}{FIG {#1}} }    
-\newcommand{\ffn}[1]{}
-
-%\newcommand{\ttn}[1]{ \colorbox{green}{TABLE {#1}} }    
-\newcommand{\ttn}[1]{}
-
-\input{captions}
-
-\begin{document}
-
-\supertitle{Research Paper}
-
-\title[Probabilities of SETI contacts]{Probability of causal
-contact between interstellar civilizations through MonteCarlo simulations}
-
-\author[Lares, Funes \& Gramajo]{Lares M.$^{1, 2}$, Funes J.~G.$^{1,
-3}$ \& Gramajo L.$^{1, 2}$}
-
-\corres{\name{Lares M} \email{marcelo.lares@unc.edu.ar}}
-
-\address{
-   \add{1}{CONICET, Argentina}
-
-   \add{2}{Universidad Nacional de C\'ordoba, Observatorio
-           Astron\'omico de C\'ordoba, Argentina}
-
-   \add{3}{Universidad Cat\'olica de C\'ordoba, Argentina}
-}
-
-\keywords{SETI, Computer simulations, Statistics}
-
-%\JELclassification{XX; XX}
-%\MSCcodes{XX; XX}
-
-\Abbreviations{SETI: Search for Extraterrestrial Intelligence,\\
-CCN: causally connected node,\\
-SFC: surface of first contact,\\
-SLC: surface of last contact,\\
-DE: Discrete Event,\\
-GHZ: Galactic Habitable Zone}
-
-\begin{abstract}
-%
-The abundance of intelligent civilizations in the Galaxy is a longstanding
-question, often conceptualized as the problem of the lack of received
-communication. Most efforts on the estimation of a number of civilizations are
-focused on the Drake equation, although its factors are affected by large
-uncertainties and it lacks a temporal nature.  Alternative approaches use
-detailed numerical simulations of the galaxy and recipes for the stars and
-planets formation rates and for the origin of life, and rely on highly
-uncertain parameters. We present a statistical model for the abundance and
-duration of civilizations, implemented through Monte Carlo simulations.  We
-explore the hypothesis space of the model by a suite of simulations for the
-emergence of communicating nodes, and analyze their
-causal connections.  This is based on minimal assumptions and three
-free parameters, with focus on the statistical properties of empirical
-models instead of an interpretable model with variables to be determined by
-observation. The analysis of the rate of causal contacts as a function of the
-mean number of civilizations, the mean lifespan distribution and the maximum
-distance a civilization can send signals is presented, and used as a framework
-to discuss the spatial and temporal structure of a populated galaxy within
-several scenarios. We find that, given the large distances involved,
-causal contacts between civilizations are very rare. Additionally, the odds to
-make a contact in a few years of monitoring are low for most models, with the
-exception of those of a galaxy densely populated with long lived
-civilizations. The probability of causal contacts increases with the
-lifetime of civilizations much more significantly than with the number of
-active civilizations for a given time window. Finally, we show that the
-maximum probability of making a contact occurs when a civilization acquires
-the required communication technology.
-%
-\end{abstract}
-
-\maketitle
-
-%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
-\section{Motivations}\label{S_motivations}
-%{{{
-
-The Drake equation \citep{drake_intelligent_1962} offers a very
-helpful educated guess, a rational set of lenses --the factors in the
-equation-- through which to look at future contacts with a
-technologically advanced civilization in the Milky Way.
-%
-The equation quantifies the number of civilizations from whom we might
-receive an electromagnetic signal, using a number of factors that have
-been widely discussed in the literature and whose estimates are
-updated constantly.
-%
-A comprehensive review and analysis of each term of the equation is
-presented in \citet{vakoch_drake_2015}.
-%
-Some optimistic estimates from the Drake equation contrast with the so
-called Fermi paradox, a statement about the contradiction between the
-apparent abundance of life in the Galaxy and its lack of evidence
-\citep[e.g. ][]{hart_explanation_1975, brin_great_1983,
-barlow_galactic_2012, forgan_galactic_2016, vanhouten_isthere_2017,
-Sotos_biotechnology_2019, carroll_nellemback_fermi_2019}.
-%
-There are many proposals aimed at solving this paradox, which make use
-of statistical \citep{solomonides_probabilistic_2016,
-vanhouten_isthere_2017, horvat_calculating_2007,
-maccone_statistical_2015} or stochastic approaches
-\citep{forgan_numerical_2009, bloetscher_using_2019,
-glade_stochastic_2011, forgan_numerical_2010}.
-%
-Regarding the Drake equation, analytical interpretations
-\citep{prantzos_joint_2013,
-smith_broadcasting_2009} or reformulations
-\citep[][and references therein]{burchell_whither_2006} have also been
-proposed.
-%
-The lack of contacts could be explained by the difficulty of life to
-appear due to astrophysical explanations
-\citep{annis_astrophysical_1999}.
-%
-In general, also more speculative proposals
-\citep{barlow_galactic_2013, lampton_information_2013,
-conway_three_2018, forgan_galactic_2017} have been discussed.
-% 
-\Fpagebreak
-%
-We stress the fact that the only observation that can be stated with
-certainty is that for the number of years projects for the Search of
-Extraterrestrial Intelligences (SETI) have been working, we have not
-received any signal using the technical possibilities and the
-conditions established by SETI projects \citep{tarter_search_2001}.
-%
-The lack of detection has also inspired to a number of alternative
-proposals for new SETI strategies \citep{forgan_exoplanet_2017,
-balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010,
-tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016,
-maccone_SETI_2011,  lingam_relative_2019, wright_theGsearch_2015,
-maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018,
-forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
-%
-The Drake equation is a key tool to organize the discussion around the
-problem of the abundance of civilizations in the Galaxy
-\citep{hinkel_interdisciplinary_2019}.
-%
-However, the uncertainties in the factors of this equation make it
-less prone to a formal application in order to define searching
-strategies or to compute the actual number of extraterrestrial
-intelligences.
-%
-The optimistic estimations from the Drake equation also contrast with
-the SETI initiatives, noticeable overshadowed by the fact that despite
-decades of effort, not a single extraterrestrial intelligence, if any,
-has been detected.
-%
-\citet{prantzos_joint_2013} proposes a unified framework for a joint
-analysis of the Drake equation and the Fermi paradox, concluding that
-for sufficiently long-lived civilizations, colonization of the Galaxy
-is the only reasonable option to gain knowledge about other life
-forms.
-%
-\citet{haqq-misra_drake_2017} discuss the dependence of the Drake
-equation parameters on the spectral type of the host star and the time
-since the Galaxy formed and examine trajectories for the emergence of
-communicative civilizations over the history of the Galaxy.
-%
-Some modifications to the original idea of the equation have tried to
-imprint a stochastic nature, to propose a probabilistic approach, or
-to consider the temporal structure which is missing in the equation.
-%
-Temporal aspects of the distribution of communicating civilizations
-and their contacts have also been explored by several authors
-\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
-balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011}, as
-well as efforts on considering the stochastic nature of the Drake
-equation \citep{glade_stochastic_2011}.
-%
-The simulation approach has also been considered
-\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
-murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017,
-ramirez_new_2017}, although with a similar problem that consists on
-the large number of parameters which are either unknown or largely
-uncertain.
-%
-\citet{grimaldi_signal_2017} consider a statistical model for the
-probability of the Earth making contact with other intelligent
-civilizations, taking into account the finite lifetime of signal
-emitters, and based on the fractional volume occupied by all signals
-reaching our planet.
-
-In a recent work, \citep{balbi_impact_2018} use a statistical model to
-analyze the occurrence of causal contacts between civilizations in the
-Galaxy.
-%
-The author highlights the effect of evolutionary processes when
-attempting to estimate the number of communicating civilizations that
-might be in causal contact with an observer located on the Earth.
-%
-\citet{cirkovic_temporal_2004} also emphasize the lack of temporal
-structure in the Drake equation.
-%
-\citet{bloetscher_using_2019} uses a probabilistic approach, though
-still strongly motivated by the Drake equation, to obtain a
-probabilistic measure for the number of civilizations in the Galaxy.
-%
-To that end, perform MonteCarlo Markov Chains over each factor of the
-Drake equation, and combine the means to obtain a probabilistic
-result.
-%
-It is worth mentioning that the author proposes a log-normal target
-distribution to compute the posterior probabilities.
-%
-%This approach, which assigns probability distributions to the Drake
-%factors.
-%
-This study concludes that there is a very low probability for Earth to
-make contact with another extraterrestrial civilization.    
-%
-\citet{smith_broadcasting_2009} uses an analytical model to compute
-the probabilities of contact between two randomly located
-civilizations, and the waiting time for the first contact, assuming a
-fixed maximum broadcasting distance.
-% 
-Using a different approach, \citet{balbi_impact_2018} investigate the
-chance of communicating civilizations making causal contact within a
-volume surrounding the location of the Earth, using a statistical
-model to compute the occurrence of causal contacts.
-%
-The author stresses the fact that the causal contact requirement
-involves the distance between civilizations, their lifespan and their
-times of appearance.  This is important since the time it takes the
-light to travel across the Galaxy is much lesser than its lifetime.
-%
-\citet{balbi_impact_2018} fix the total number of civilizations and
-explore the parameter space that comprise three parameters, namely,
-the distance to the Earth, the time of appearance and the lifespan of
-the communicating civilization.
-%
-Each of these three variables are drawn from a random distribution.
-The distribution of the distances results from a uniform distribution
-of civilization within the plane of the Galaxy.
-%
-For the distribution of the characteristic time of appearance, the
-author explores and exponential and a truncated Gaussian
-distributions. For the distribution of the lifespans, the author
-chooses an exponential distribution.
-%
-%The radius of influence is set to 1000 lyr.
-%
-\citet{balbi_impact_2018} finds that the fraction of communicating
-civilizations vary with the choice of the statistical distribution for
-the time of appearance. For all analyzed distributions, the fraction
-corresponding to a mean lifetime of 10$^9$ years and for a maximum
-radius for the detection of the signal of 1000 yr, is roughly 0.1.
-%
-In another probabilistic approach,
-\citet{vukotic_astrobiological_2012} propose a probabilistic cellular
-automata (PCA) modeling. In this framework, a complex system is
-modeled by a lattice of cells which evolve at discrete time steps,
-according to transition rules that take into account for each cell the
-states of its neighbor cells.
-%
-The authors implement a PCA model to a network of cells which
-represent life complexity on a 2-dimensional annular ring resembling
-the Galactic Habitable Zone (GHZ). 
-%
-The authors set the GHZ between a minimum radius of 6 kpc and a
-maximum radius of 10 kpc. Within this framework, the authors also make
-several MonteCarlo simulations and analyze ensemble-averaged results.
-%
-This work aims at analyzing the evolution of life, although it does
-not account for the network of causal contacts among technological
-civilizations.
-
-In this work we address the problem of the temporal and spatial
-structure of the distribution of communicating civilizations, by
-exploring the hypothesis space over a minimal set of parameters.
-%
-We propose to avoid the frequentist approach of the
-Drake equation and to explore a parameter space, where instead of
-computing a final number, we provide a statistical distribution that
-gives conditional probabilities.
-%
-This is an exploratory analysis that aims at providing a numerical
-tool to discuss not the several theoretical problems summarized by the
-Drake equation factors, but the different scenarios on the basis of
-statistical heuristics.
-%
-The approach proposed here should be considered as a compromise
-between the uncertainties of the frequentist approach and the detailed
-recipes required on the simulation approach.
-%
-We then provide a numerical framework to explore via simulations a
-parameter space of unknown observables in order to discuss different
-scenarios and their consequences in terms of the probability of
-making contacts.
-%
-The method we use for simulating a stochastic process is an
-approximation that allows to study the behavior of complex systems,
-by considering a sequence of well defined discrete events.
-%
-The simulation is carried out by following all the variables that
-describe and constitute the state of the system, and the evolution of
-the process is described as a set of changes in that state.
-%
-In this context, an event produces a specific change in the state,
-that can be triggered by random variables that encode the stochastic
-nature of the physical phenomenon.
-%
-For example, when a new contact is produced between two entities in
-the simulated galaxy, the numbers of active communication lines and of
-active communicated nodes change.
-%
-The process involves following the changes on the state of the
-system, defining the initial and final states, defining a method that
-allows to keep track of the time progress in steps, and maintaining a
-list of relevant events.
-%
-In Sec.~\ref{S_methods} we introduce the methods and discuss the
-candidate distributions for the statistical aspects of the times
-involved in the communication process.
-%
-Then, we present our results in Sec~\ref{S_results}, with special
-emphasis on the statistical distributions of the duration of causal
-contacts in one or both directions, the possible differences on the
-position of a civilization on the Galaxy and the distribution of time
-intervals for the waiting of the first contact, always as a function
-of the three simulation parameters.
-%
-Finally, in Sec.~\ref{S_discussion} we discuss our results and future
-research directions.
-
-%}}}
-
-%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
-\section{Methods and working hypotheses}\label{S_methods}
-%{{{
-
-Simulations are suitable tools to analyze systems that evolve with
-time and involve randomness.
-%
-An advantage of simulations compared to theoretical approaches, is
-that the former usually require less assumptions and simplifications,
-and can be applied to systems where a theoretical model can not be
-found.
-%
-In particular, many complex stochastic processes that can be described
-by the evolution of the state of a system, can be efficiently modeled
-with the discrete--event (hereafter, DE) simulation approach.
-%
-A system described with the DE paradigm is characterized by a set of
-actors and events, where actors interact causally through a series of
-events on a timeline and process these events in chronological order
-\citep{ptolemaeus_system_2014, chung_simulation_2003,
-ross_simulation_2012}.
-%
-Each event yields modifications to the values of the variables that define
-the process, producing the corresponding change in the state of the
-simulated system.
-%
-This method is well suited for the particular case of the diffusion of
-intelligent signals in the galaxy, and allows to explore several
-models easily.
-%
-We simulate the statistical properties of a set of points in space and
-time that have a causal connection at light speed, hereafter dubbed
-''Causal Contact Nodes'' (hereafter, \ceti{}).
-%
-We chose this generic name in order to stress the fact that in this
-analysis no broadcasting or lookout activities are required, and
-therefore not actual messages are considered, only the causal contact.
-%
-For the special case of a fully efficient node that emits and receives
-isotropically, a causal contact node can be considered as a
-broadcasting station that has the ability to detect signals through an
-active lookout program.
-%
-Also, it is worth mentioning that this is a general approach, and not
-necessarily a \ceti{} is the host of an intelligent civilization.
-%
-For instance, it can be associated with a planet where life has developed, became
-intelligent, reached the skills required to find the right
-communication channel, sustain a search and establish a contact.
-%
-Alternative message processing entities could be considered, for
-example interstellar beacons where life has ceased to exist but
-continue with its emission, or communication stations established by
-probes or left intentionally by intelligent beings \citep[see, e.g.,
-][]{peters_outer_2018, barlow_galactic_2013}.
-%
-In principle, these strategies could affect our results since it would
-be easier to configure a cluster of \cetis{} that spread in time.
-%
-However, we do not consider these highly speculative alternatives at
-this point.
-%
-For the purpose of this analysis, only the communication capability is
-relevant, since we study the causal contacts between the locations.
-%
-The system is defined by a number of actors that represent \cetis{}
-and appear at different instants in time, generating events that
-produce meaningful changes in the variables that describe the system,
-i.e., in the arrangement of \cetis{} and their network of causal
-contacts.
-%
-For example, the appearance of a new \ceti{} in a region which is
-filled with a signal emitted by another node, will increase the
-number of active \cetis{} and the number of pairs of \cetis{} in
-causal contact.
-%
-Assuming some simple hypotheses, the discrete events method can be
-performed taking into account a small number of variables, which allow
-to analyze the variation of the results in the model parameter space.
-
-In what follows, we describe the experimental setup chosen to estimate
-the probabilities of causal contacts and several derived quantities in
-terms of three independent parameters, namely, the mean time span
-between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$,
-their mean lifetime $\left<\tau_s\right>$, and the maximum distance a
-signal can be detected by another \ceti{} ($D_{max}$).
-%
-Intuitively, the shorter the $\tau_a$ parameter and the larger the
-$\tau_s$ and $D_{max}$ parameters are, the greater the probability of
-the existence of causal contacts between pairs of \cetis{} would be.
-%
-We also need to propose theoretical distribution functions for both
-the distribution of lifespans ($\tau_s$) and the distribution of the
-number of \cetis{} per unit time \citep{maccone_evolution_2014,
-Sotos_biotechnology_2019}, related to the time span between the
-appearance of consecutive \cetis{} (since when $\left<\tau_a\right>$
-is shorter, it produces a greater density of \cetis{}).
-%
-The shapes of these distributions are set to a fixed law, as discussed
-in Sec.~\ref{SS_PDF_shape}.
-
-%}}}
-
-\begin{figure}[!t]
-   \centering
-   \includegraphics[width=0.5\textwidth]{F_scheme.pdf}
-   \usecaption{F_scheme} \label{F_scheme}
-\end{figure}
-
-%{{{
-
-\ffn{1} We illustrate in the Fig.~\ref{F_scheme} the schematic
-representation of the region causally connected to a given \ceti{}.
-%
-To this end, we use space--time diagrams, where time is represented on
-the horizontal axis, and space is represented in the vertical axis.
-%
-For the sake of simplicity, we show in the plot the light travel
-distance, i.e., the distance traversed by any signal spreading from
-the emitter at light speed (for example, any electromagnetic signal).
-%
-In this scheme, a light pulse would follow a trajectory represented by
-a line at 45 degrees from the axes, given the units of the time and
-space axes are years and light years, respectively.
-%
-Panel (a) of the Figure represents the region $R$ (shaded polygon) in
-space and time which is causally connected to the emitter (left
-vertex), following an event (hereafter A type event) in which the node
-acquires the communication capability, or becomes ''active''.
-%
-The sphere of first contact (SFC) is a sphere centered on the emitter,
-represented in the Figure by the left angle of the region $R$.
-%
-This sphere grows until its radius reaches the D$_{max}$ distance, in
-which it would be no longer detectable due to the decrease in the
-energy per unit area, which falls under the assumed detectability
-threshold.
-%
-Similarly, the surface of last contact (SLC) is another sphere that
-grows from an event in which the nodes ceases to possess the
-communication capability (D type event), and carries the last signal
-produced by the emitter.
-%
-The region which is causally connected to the emitter is then limited
-by these two spheres, and has the shape of a sphere (before the D
-event) or of a spherical shell (after the D event).
-%
-The temporal intervals for the communication between two \cetis{} are
-represented in the panel (b) of the Figure.
-%
-In this scheme, the entire lifetime of a \ceti{} can then be
-represented by a polygon, limited in the spatial axis by the maximum
-distance of the signal $D_{max}$ and in the time axis by the time span
-between the starting time (A event, concave vertex) and the ending
-time (D event, convex vertex).
-%
-This representation allows to easily visualize the important events
-that result from the existence and communications of two \cetis{}.
-%
-The points in time where the \cetis{} acquire their communicating
-capacity are dubbed ''Awakening'' events ($A_1$ and $A_2$ in the
-Figure).
-%
-Similarly, the points in time where the \cetis{} loss their
-communicating capacity are dubbed ''Doomsday'' events ($D_1$ and $D_2$
-in the Figure).
-%
-The points in space--time where the first contact is produced for each
-one of the \cetis{}, are defined as ''Contact'' events, shown as $C_1$
-and $C_2$ in the Figure.
-%
-Finally, the two points where the contact is lost for each one the
-\cetis{}, are denominated ''Blackout'' events, shown as $B_1$ and
-$B_2$ in the Figure.
-%
-The receiver node E$_1$ can listen signals from the emitter node
-E$_2$, from the first contact event (t=C$_{12}$) up to the last
-contact event (t=B$_{12}$).
-%
-Similarly, the receiver node E$_2$ can listen signals from the emitter
-node E$_1$, from t=C$_{21}$ up to t=B$_{21}$.
-%
-The time intervals for the open communication channel are determined
-by the ''C'' and ''B'' type events.
-%
-It is important to notice that there is a time delay between the
-contact events of the two \cetis{} involved in this analysis, and also
-a time delay between the two blackout events, so that the time range
-when a bidirectional contact is possible occurs between the maximum
-time of the contact points and the minimum time between the blackout
-points.
-
-%}}}
-
-\begin{figure*}
-   \centering
-   \includegraphics[width=\textwidth]{F_messages.pdf}
-   \usecaption{F_messages} \label{F_messages}
-\end{figure*}
-
-%{{{
-
-The temporal structure that emerges naturally from the experimental
-setup implies that,
-as a fundamental property of the simulations,
-a causal connection can be produced without the need of the two
-involved locations being concurrent or active at the same time.
-%
-This property of the system arises as a
-consequence of the large spatial and temporal scales, where a message
-is transmitted at the (relatively small) light speed.
-%
-Although a \ceti{} could be active for enough time so as to transmit a
-message at large distances, the limited power of the message and the
-dilution that runs as the squared distance from the source, imposes a
-detectability limit.
-%
-As a consequence of this limitation and of their finite lifetime,
-considered as the period of time between the acquisition and loss of
-communicating capacity, each \ceti{} will fill a spherical shell
-region of the galaxy, limited by two concentric spherical surfaces.
-%
-The leading front, or surface of first contact (SFC) grows from the
-central \ceti{} until it reaches the maximum distance $D_{max}$.
-%
-Thus, the volume reached causally by a \ceti{} is initially a growing
-sphere.
-%
-Following the end of the civilization, there is still a region which
-is filled with the emitted signals.
-%
-This approach has been also considered other statistical models
-\citep[e.g., ][]{smith_broadcasting_2009, grimaldi_signal_2017, Grimaldi2018}.
-%
-The trailing front, or surface of last contact (SLC) also grows from
-the central \ceti{}, with a delay with respect to the SFC equivalent
-to the lifetime of the node, and produces a spherical shell region.
-%
-Any other \ceti{} within this region will be aware of the originating
-node, even if it has disappeared before the time of contact.
-%
-This region will grow if the surface of first contact has not yet
-reached the maximum distance $D_{max}$, and will shrink otherwise
-until the surface of last contact reaches $D_{max}$, producing as a
-result the loss of all signal from the central \ceti{}.
-%
-In our approach, we consider a model galaxy where the width of the
-disk is negligible with respect to the radius of the disk.
-%
-In the 2D simulation only the intersection of the communicating
-spherical shells with the plane of the galaxy is relevant, and produce
-the corresponding circles or rings for the filled spheres or annular
-regions of the spherical shells, respectively.
-%
-The initially growing communicating sphere, is shown over space--time
-diagrams, where time is represented on the horizontal direction, and
-space is represented in the vertical direction.
-%
-In the Fig.~\ref{F_scheme} the two
-emitters in panel (b), E$_1$ and E$_2$, reach each other at different times.
-%
-The time span for E$_i$ is (A$_i$ , D$_i$) , for $i = {1, 2}$.
-%
-Emitter $i$ can listen to emitter $j$ between C$_{ij}$ and B$_{ij}$.
-%
-The type and length of causal contact in both directions depend on the
-distance and time lag between the awakening events, the maximum
-distance that a signal can reach and the time period in which each
-emitter is active.
-%
-\ffn{2}
-%
-In the Fig.~\ref{F_messages} we show a schematic representation of the
-possible cases in which a \ceti{} E can be in causal contact with a
-\ceti{} R.
-%
-The duration of the causal contact in one direction depends on several
-factors, mainly the lifetimes, the time period between the A events
-and the distance between the two nodes involved in the causal
-contact.
-%
-As it can be seen, the direction of the causal contact does not imply
-any ordering for the A events of the involved nodes.
-%
-In the implementation of the DE simulation, the six cases shown in the
-Figure determine the succession of events.
-%
-Each time an event is produced, the list of upcoming events is updated
-according to these cases.
-
-In out experimental configuration, there is no a temporal window for
-the model, i.e., the simulation starts when the process is already
-stable, and ends before any galactic effect could modify or affect of
-fixed variables.
-%
-Also, we assume that the probability for the rise of a \ceti{} is
-homogeneous over the GHZ.
-%
-Although the Galaxy has a well known spiral structure, the assumed
-sparsity of nodes (intelligent civilizations) does make this
-assumption fairly reasonable.
-%
-Otherwise, if the distribution of nodes is not sparse, it could be the
-case that the spiral arms would host most of the \cetis{}, and
-contacts are frequent between closely located nodes.
-%
-In such a case we could expect that the interpretation of our results
-could be considered as pessimist for the establishment of a contact
-with our planet.
-%
-We also limit the possibilities of life or other types of
-civilizations to the hypothesis usually stated for the definition of
-the GHZ \citep{dayal_habitability_2016, gonzalez_galactic_2001,
-lineweaver_galactic_2004, gonzalez_habitable_2005,
-morrison_extending_2015, haqq-misra_evolution_2019,
-rahvar_cosmic_2016, gobat_evolution_2016, rahvar_cosmic_2016} and
-consider that habitability remains constant \citep[see, however,
-][]{gonzalez_habitable_2005, dayal_habitability_2016, gobat_evolution_2016}.
-%
-For example, \citet{vukotic_astrobiological_2012} propose
-probabilistic cellular automata model to explain the astrobiological
-history of the Milky Way.
-%
-This means that we set aside possible civilizations that could survive
-in severe conditions or unstable systems, which would prevent the
-appearance of life as we know it.
-
-While the aforementioned assumptions are basic conditions for most of
-the stochastic processes observed in nature, we need to make stronger
-assumptions related to the nature of the message.
-%
-The most simple assumption about the message itself, is that it
-travels at light speed.
-%
-With this, we are considering messages sent through, for example,
-electromagnetic radiation or even gravitational radiation or neutrino
-communication \citep{hippke_interstellar_2017, wright_how_2018}, but we set
-apart messages sent with mechanical means or physical objects
-\citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown technology
-that violates the known laws of physics.
-% 
-For the communication of \cetis{} through messages sent isotropically,
-we assume that the capacity to emit signals and to receive signals
-occur at the same time.
-%
-This means that the ability to find another civilization develops at
-the same time that the ability and intention to emit a signal in all
-directions, intended to be detected by an unknown civilization.
-%
-Although there are several reasons to think that this could not be the
-case, at large time scales it can be considered that both abilities
-occur roughly at coincident epochs.
-%
-Another essential assumption is that all \cetis{} use the same signal
-power, so that there is a maximum distance out to which it can be
-detected.
-%
-It is worth mentioning that we are considering in our experimental
-setup a system composed by several emitters and receivers across the
-galaxy, under the hypothesis proposed before.
-%
-In such a system we compute probabilities of a \ceti{} making contact
-with another node, and the results must be interpreted as for a
-randomly chosen \ceti{}, given certain conditions of homogeneity on an
-ensemble of equivalent galaxies.
-%
-Then, we stress the fact that we are not centered on the probabilities
-for the Earth itself, but on a generalized \ceti{}.
-%
-Regarding the extent of the signals, we know that the distance from
-which a signal from Earth could be detected using the current
-technology is of a few hundreds light years, given that the signal was
-sent to a specific direction.
-%
-%https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
-%https://www.engadget.com/2017/05/25/listening-to-starlight-our-ongoing-search-for-alien-intelligenc/]
-%[from https://en.wikipedia.org/wiki/Fermi_paradox]
-%
-It is straightforward to propose and implement a distribution of
-maximum distances, although this would increase the model complexity
-at the cost of a larger uncertainty or of another dimension on the
-hypothesis space.
-%
-%This proposal is based on a pronoid scenario, and we do not consider
-%in this work the occurrence of paranoid or partially paranoid
-%civilizations.
-%%
-%If that would be the case, the results obtained here can be taken as
-%upper limits.
-
-In our simulations we also assume the simplest situations for the
-growth of the sphere of first contact.
-%
-For example, we discard the possibility of stellar colonization
-\citep[e.g.][]{newman_galactic_1981, walters_interstellar_1980,
-starling_virulence_2013, barlow_galactic_2012, jeong_large_2000,
-maccone_mathematical_2011} and assume that the communication aim is
-performed to all directions with the same power and the same
-probability, i.e., strongly assume isotropic communication in all
-cases.
-%
-More detailed simulations could be produced considering different
-efficiency of communication or detection methods.
-%
-However, in this case we also would complicate the experimental setup,
-make the results less evident.
-%
-The results are independent of the nature of life (organic or
-artificial).
-%
-The lifetime of a civilization can be caused by auto destruction or by
-external factors.
-%}}}
-
- \setlength{\tabcolsep}{10pt}
-\begin{table*}
-\centering
-\begin{tabular}{cllll}
-\hline
-   \multicolumn{2}{l}{independent variables (hypothesis space)}
-   &min. value&max. value&Nbins\\
-\hline
-   $\tau_{a}$ & Mean temporal separation between consecutive
-   awakenings & 4000 yr & 200\,000 yr & 50\\ 
-   $\tau_{s}$ & Mean lifetime of a \ceti{}
-   & 10000 yr & 500\,000 yr& 50\\ 
-   D$_{max}$ & Maximum reach of a message  &  \multicolumn{2}{l}{(500,
-   1000, 10000, 20000, 40000, 80000) lyr} & 6  \\
-\hline
-   \multicolumn{2}{l}{fixed variables} & \multicolumn{2}{c}{assumptions} &value \\
-\hline
-   & statistical properties of all \cetis{} &equally distributed&&\\
-   & Point process for the distribution in time & homogeneous &&\\
-   f$_s$ & The scan of the sky & fully efficient&&1\\
-   f$_p$ & panspermia or colonization &absent&&0\\
-   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
-   R$_{GHZ}^{min}$   & Inner radius of the GHZ     & & &20\,000 lyr\\
-   R$_{GHZ}^{max}$   & Outer radius of the GHZ     &&&60\,000 lyr\\
-   t$_{max}$ & Time span of the simulation  &&&1.e6 yr\\
-    & \multicolumn{3}{l}{number of random realizations for each point in the hypothesis space} &50\\
-\hline
-   \multicolumn{2}{l}{discrete events} &
-   \multicolumn{3}{l}{affected variables}\\
-\hline
-   A event & Awakening: a \ceti{} starts its communication
-   capabilities &Number of active \cetis{}\\
-   B event & Blackout: the end of the communication channel stops
-   &Number causal channels\\
-   C event & Contact: a new causal contact is produced &Number of causal channels\\
-   D event & Doomsday: a \ceti{} ends its communication capabilities&Number of active \cetis{}\\
-\hline
-
-\hline
-\end{tabular}
-\caption{Definition of independent variables and adopted values for 
-   fixed parameters 
-   that are part of the simulation.  Variable parameters define the
-   spatial and temporal structure
-   of the process and the maximum reach of the messages.}
-\label{T_simu_hypotheses}
-\end{table*}
-
-\subsection{Power laws vs. exponential laws}\label{SS_PDF_shape}
-%{{{
-
-The temporal structure of the process is defined by two distribution
-parameters that represent the mean time interval that a \ceti{}
-can emit and receive signals, and the mean time interval
-between the emergence of consecutive \cetis{}.
-%
-The spatial structure of the simulation is given by the size and shape
-of the Galactic Habitable Zone and the maximum distance a signal can
-travel to be detected (D$_{max}$).
-%
-The parameters for the temporal distributions also determine the
-spatial properties, since the density of active \cetis{} in the galaxy
-depend on these two parameters.
-%
-For example, a small $\tau_a$ and a large $\tau_s$ will produce a
-galaxy densely populated with \cetis{}.
-%
-Also, some hypotheses regarding the shapes of the distributions of the
-temporal parameters must be made in order to complete the
-simulation.
-%
-\citet{forgan_spatiotemporal_2011} argues that the times at which
-different civilizations become intelligent follow a Gaussian
-distribution, and then the distribution of inter-arrival times is an
-inverse exponential.
-%
-We choose to assume that the distribution of the times of A events is a
-stationary Poisson process, and then the distribution of the times
-between the appearance of new consecutive \cetis{} is exponential.
-%
-Regarding the duration of a \ceti{}, we propose that the distribution
-of the duration of \cetis{} is a stationary exponential distribution.
-%
-\citet{balbi_impact_2018} stress the fact that it would be desirable
-to arrive at a theoretical statistical distribution of the lifespan of
-civilizations, preferentially on the basis of the underlying
-astrophysical and biological processes.
-%
-\citet{maccone_lognormals_2014} argues that this distribution should
-be a log normal.
-
-The power law and exponential statistical distributions are among the
-most common patterns found in natural phenomena.
-%
-For example, the distribution of the frequency of words in many
-languages is known to follow a Zipf's law (which is a power law).
-%
-This distributions arise any time a phenomena is characterized by
-commonly occurring small events and rarely produced large events
-\citep[e.g. ][]{adamic_zipf_2000}.
-%
-Zipf law also describes population ranks of cities in various
-countries, corporation sizes, income rankings, ranks of number of
-people watching the same TV channel.
-%
-The magnitudes of earthquakes, hurricanes, volcanic eruptions and
-floods; the sizes of meteorites or the losses caused by business
-interruptions from accidents, are also well described by power laws
-\citep{sornette_critical_2006, benguigui_classificacion_2016}.
-%
-The power law behavior has been observed in a variety of systems,
-including for example stock market fluctuations, sizes of computer
-files and word frequency in languages \citep{mitzenmacher_brief_2004,
-newman_power_2005, simkin_theory_2006}
-%
-Power laws have also been widely used in biological sciences, e.g., in
-analyzing connectivity patterns in metabolic networks
-\citep{jeong_large_2000} and in the number of species observed per
-unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
-% 
-More examples can be found in the literature
-\citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009,
-maccone_evolution_2014, maccone_lognormals_2014,
-benguigui_classificacion_2016}
-
-In this work, the exponential distribution of lifespan and waiting
-times is justified by considering the hypothesis that the process of
-appearance of life in the galaxy is homogeneous and stationary.
-%
-That is, there is no preferred location within the GHZ for the
-spontaneous appearance of life, and the emergence of a \ceti{} is
-independent of the existence of previous nodes in the galaxy.
-%            
-This is equivalent to proposing a Poisson process for the emergence of
-\cetis{}, since there is a close relation between the number of events
-in time or space and the waiting time or separation, respectively
-\citep[e.g., ][]{ross_simulation_2012}.
-%
-That is, these are two alternative approaches to describing the same
-process, a Poisson distribution for the number of events implies an
-exponential distribution for their separations, and vice versa.
-%
-It should be emphasized that the exponential laws used in this work
-are assumed as part of the working hypothesis, and instead of
-analyzing results from a particular parameter chosen ad--hoc, we
-explore the hypothesis space and analyze the impact of the values of
-these parameters on the results.
-%}}}
-
-\begin{figure} \centering
-   \includegraphics[width=0.5\textwidth]{F_number_of_contacts.pdf}
-   \usecaption{F_number_of_contacts} \label{F_number_of_contacts}
-\end{figure}
-
-\subsection{Complexity of the model}
-%{{{
-
-In this Section we discuss the degree of complexity in the model,
-considering a compromise between the accuracy of the model and the
-number of parameters that are free or with a high uncertainty.
-%
-Firstly, we emphasize that the odds of a causal contact between two
-\cetis{} should not be considered as the odds of a contact between two
-intelligent civilizations, and in fact the later could be much lesser
-than the former.
-%
-Indeed, in order to establish a contact between any two entities, a
-minimum degree of compatibility must be accomplished without any
-previous agreement, making the possibility of a contact with a message
-that could be deciphered highly rare \citep[see e.g. ][]{forgan_collimated_2014}.
-%
-Besides the trade--off between the simplicity and the complexity of
-the experiments, further analysis could be performed following this
-framework in order to explore possible implications of the results for
-more detailed configurations.
-%
-For example, the communication method (isotropic, collimated,
-serendipitous) can affect the observables, making it necessary to
-implement a correction factor.
-%
-Taking this into account, our results regarding the probability of
-causal contact should be considered as upper limits for effective
-contacts, since they depend on the efficiency of both the emitter and
-the receiver to broadcast and scan the sky for intelligent signals,
-respectively \citep{grimaldi_signal_2017}.
-%
-A correction by coverage ratio in the detection and by a targeting
-ratio in the emission could be easily implemented in the simulation,
-although the effect of reducing the probability of contact is
-basically the product of the efficiency ratios and thus such
-implementation is not necessary.
-%
-Therefore, the values of the probabilities could be modified by a
-constant correction factor equal to the combined emission/reception
-efficiency \citep{smith_broadcasting_2009, anchordoqui_upper_2019,
-forgan_collimated_2014} or for beam--like transmissions
-\citep{grimaldi_signal_2017}.
-%
-Other considerations include the effects of alignments, the use of
-stars as sources or amplifiers \citep{Edmondson2003, borra_searching_2012}, 
-the nature of the message carrier, the
-use of a spatial distribution that actually resembles the spiral shape
-and the width of the disc of the Galaxy, different distribution
-functions for the mean lifetime of \cetis{}, and different
-efficiencies for the sphere of the causal region of each \ceti{}
-resembling different searching strategies
-\citep{hippke_interstellar_2017}.
-%
-It is also possible to consider that D$_{max}$ is different for
-different \cetis{}.  For example a power law where powerful emissions
-are rare and weak emissions are common could be an improvement to the
-model.
-%
-In this work we chose not to implement this for the sake of
-simplicity.
-%
-Finally, the role of the message contents could influence on the
-lifespan of a \ceti{} that receives a message, although the
-implementation of such behavior would increase the number of free
-parameters and would be highly speculative.
-
-%}}}
-
-\subsection{Discrete event process}
-%{{{
-
-Given a parametric model $M(\tau_a, \tau_s, D_{max})$, which includes
-the functional forms of the statistical distributions and fixed values
-for the hypotheses, a discrete event simulation is performed by
-keeping track of a set of variables that change each time a meaningful
-event is produced.
-%
-The main variables that follow the evolution of the simulation are the
-positions of stars, which are sampled randomly within the GHZ, the
-time of the awakening of each \ceti{} (A event), and the time of the
-vanishing of each \ceti{} (D event).
-%
-The variables that can be deduced from the previous ones include the
-number of \cetis{} in casual contact with at least another \ceti{} at a given
-time, the number of \cetis{} as a function of time, the number of
-\cetis{}
-that receive a message at least one time, the number of \cetis{} that
-receive a message at least one time and successfully deliver an answer
-and the number distribution of waiting time to receive a message.
-%
-All these quantities are updated each time one of the four events (A,
-B, C, D) is produced.
-
-%%% comparar con el stochsatic cellular automata
-% ver libro Solving Fermi paradox, sec. 27.3.1
-
-%}}}
-
-\begin{figure*} % 2D color plot
-   \centering
-   \includegraphics[width=\textwidth]{F_never_contact.pdf}
-   \usecaption{F_never_contact}
-   \label{F_never_contact}
-\end{figure*}
-
-\begin{figure*} % 2D color plot
-   \centering
-   \includegraphics[width=\textwidth]{F_C_at_A.pdf}
-   \usecaption{F_C_at_A}
-   \label{F_C_at_A}
-\end{figure*}
-
-\subsection{Model for \cetis{}}
-%{{{
-
-Since the method is based on events, the first step is to define an
-architecture of events, the relationships between them and how events
-trigger changes on the state of the system,
-so that the system is updated each time an event is produced.
-%
-We consider a simulated system that represents the galactic habitable
-zone which, on a first approach, is a 2-dimensional annular region.
-%
-The adopted values for the geometry of the GHZ are 20 kly and 60 kly, for the inner
-and outer radius, respectively \citep{lineweaver_galactic_2004}.
-% ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
-% mas chico...
-%
-This simple model does not take into account the variations in stellar
-density given by the spiral structure.
-%
-Although there are several possible approaches, we chose to follow the
-evolution of the system according to the following events:
-
-\begin{enumerate}
-   \item[(A)] Awakening: A new \ceti{} acquires the ability to emit
-      and receive messages and starts to actively broadcast and
-      lookout for messages.
-   \item[(C)] Contact: A new one-way causal contact is established
-      (does not require the awareness of the involved \cetis{}.)
-   \item[(B)] Blackout: An existing causal contact is interrupted,
-      with or without the occurrence of a doomsday event.
-   \item[(D)] Doomsday: An old \ceti{} disappears, or halts its
-      communication activities.
-\end{enumerate}
-
-We have chosen this nomenclature so that it can be easily remembered
-from the initial A, B, C, D.
-%
-The \blackout is produced when one of the two \cetis{} halts in its
-capability of emitting and receiving signals.
-%
-It can occur at the same time the \doomsday or before.
-%
-Similarly, a \ccontact can be produced at the time of the \aawakening
-or later, and there can be several \contacts for the same node.
-%
-The time marks for each event (A and D for each \ceti{} and C and B
-for each causal contact) are stored as a result of the simulation.
-%
-Also, the list of active \cetis{} is obtained as a function of time.
-%
-%Some fixed variables must be set in order to carry out the simulation:
-%the size of the Galactic Habitable Zone, the mean lifetime of a
-%\ceti{}, $\tau_s$, the mean waiting time for the appearance of another
-%\ceti{}, $\tau_a$, and the maximum distance a message can be detected.
-
-The Galactic Habitable Zone is set at fixed values of the inner and outer
-radii, radial symmetry is assumed, and the with of the galactic disk
-is neglected compared to the radial size.
-%
-The two radii are measured in light years, with the aim to maintain a
-single and comprehensive unit for both space and time coordinates. 
-%
-The simulation was implemented on a python 3 code, which is publicly
-available\footnote{https://github.com/mlares/simu\_contact}
-
-%}}}
-
-\begin{figure}
-   \centering
-   \includegraphics[width=0.5\textwidth]{F_waiting_for_1C.pdf}
-   \usecaption{F_waiting_for_1C}
-   \label{F_waiting_for_1C}
-\end{figure}
  
@@
 \caption{Selected models to analyze the behavior of simulation outputs
-   and their dependence on simulation parameters.}
+   and their dependence on simulation outputs and their dependence on
+   simulation parameters.}
 \label{T_selected_models}
 \end{table*}
+
+\begin{figure*} % 2D color plot
+   \centering
+   \includegraphics[width=\textwidth]{F_never_contact.pdf}
+   \usecaption{F_never_contact}
+   \label{F_never_contact}
+\end{figure*}
+
+\begin{figure*} % 2D color plot
+   \centering
+   \includegraphics[width=\textwidth]{F_C_at_A.pdf}
+   \usecaption{F_C_at_A}
+   \label{F_C_at_A}
+\end{figure*}
+
+\begin{figure}
+   \centering
+   \includegraphics[width=0.5\textwidth]{F_waiting_for_1C.pdf}
+   \usecaption{F_waiting_for_1C}
+   \label{F_waiting_for_1C}
+\end{figure}
 
@@
 the times of each of the events that are relevant to keep track of the
-number of \cetis{} (time of Awakening and time of Doomsday) and of the
-number of contacts (the times of contacts and the times of Blackouts).
+number of \cetis{} (time of awakening and time of doomsday) and of the
+number of contacts (the times of contacts and the times of blackouts).
 %
@@
 properties of the \cetis{}, for example the total time elapsed between the
-Awakening and the Doomsday of each node.
+awakening and the doomsday of each \ceti{}.
 %
@@
 Regarding the properties of the population of \cetis{} and its evolution,
-the simulation yields the fraction of awaken time a node is listening
-at least another node (i.e., in their causal cone), the age of
+the simulation yields the fraction of awaken time a \ceti{} is listening
+at least another \ceti{} (i.e., in their causal cone), the age of
 contacted \cetis{} at first contact, the distribution of time to wait
@@
 each \ceti{}, the distribution of the number of contacts as a function of
-the age of the node, the number of contacts as a function of time in the galaxy,
-the fractions of nodes that succeed in contact, and the distribution of
-distances between contacted nodes.
+\ceti{} age, the number of contacts as a function of time in the galaxy,
+the rate of \cetis{} that succeed in contact, and the distribution of
+distances between contacted \cetis{}.
 %
@@
 \ceti{} vs. the time of double communication, the distance to \ceti{} vs. the
-age of contacted node, the age of a node and the maximum number of
-contacted nodes before the Doomsday, or the lifespan of a node vs. the
+age of contacted \ceti{}, the age of a \ceti{} and the maximum number of
+contacted \cetis{} before doomsday, or the lifespan of a \ceti{} vs. the
 maximum number of contacts.
@@
 
+In what follows, we focus on the analysis of the duration of contacts
+(Sec.~\ref{SS_multiplicity}), the waiting time for the first contact
+(Sec.~\ref{SS_waiting}) and the
+variations of contact possibilities as a function of the position of
+the \ceti{} in the galaxy (Sec.~\ref{SS_location}).
+
 %}}}
@@
 \subsection{Membership to the network of connected \cetis{}}\label{SS_members}
+
 %{{{
@@
 \cetis{} are already in the causal contact region of other nodes,
-which shows a somewhat opposite situation.
+which shows a somewhat oposite situation.
 %
@@
 %
-This consideration offers a new approach to SETI programs, where the
+This consideration offers a new approach for SETI programs, where the
 search for new communication technologies has a fundamental role.
@@
 
-%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
-\section{Discussion}\label{S_discussion}
-%{{{
-
-We have implemented a suite of simulations following a stochastic
-approach, to explore the hypothesis space of a simple model that
-accounts for the causal connections between communicating
-civilizations on a simplified galaxy.
+%\subsection{Location within the Galaxy}\label{SS_location}
+%%{{{
 %    
-The different models can be generated with a minimal number of free
-parameters and some fixed assumptions.
+%%On the Figure~\ref{F_radial_frac} we show the fraction of \cetis{} that
+%%make zero, one, two or three contacts, as a function of the
+%%galactocentric distance, for model M1.  VERIFICAR!  M1?
+%%
+%For this model, the majority of \cetis{} do not make any contact during
+%all the time period in which they are active.
+%%
+%As expected, the fractions of \cetis{} making at least one contact
+%diminishes as the number of contacts rises.
+%%
+%The shape of this distribution is a result of the geometry of the GHZ.
+%%
+%In general, a closer position to the center of the Galaxy slightly
+%favours the possibility of a contact.
+%%
+%This would indicate that the position of the Earth is suitable for the
+%odds of making a contact.
 %
-We argue that three parameters can be used to describe a variety of
-situations, ranging from a galaxy model where an intelligent civilization is
-very rare, to a galaxy model populated with plenty of civilizations in
-causal contact.
+%%}}}
 %
-With this approach, we can estimate several quantities as a function
-of the parameters on the hypothesis space,
-so we explored the outcomes of different models 
-that arise as by-products of the simulations.
 %
-For instance, it is possible to estimate the timescale for an
-uninterrupted SETI effort to reach success, in terms of different
-model parameters that reflect different, so far unknown, scenarios for
-the appearance of intelligent life on the Galaxy.
-
-Although the simulations use a number of speculative assumptions, we
-argue that the degree of knowledge about the origin and persistence of
-life in the Galaxy does not worth the implementation of more detailed
-or sophisticated models.
-%
-Instead, we take advantage of the simplicity of the model to explore
-the hypothesis space, i.e., the three--dimensional parameter space of
-the parametric model, in order to gain insight on the consequences of
-different scenarios for the search of intelligent life.
-%
-Our analysis is not centered in obtaining the odds for the Earth
-to make contact with another intelligent civilization.
-%
-Instead, we focus on obtaining a statistical, parameter dependent
-description of the possible properties of the communication networks
-that comprise sets of nodes with broadcasting and reception
-capabilities.
-%
-This causally connected nodes are sparsely distributed in both space
-and time, making it difficult the analytical treatment and justifying
-the simulation approach.
-
-Under the hypotheses of our experiments, we conclude that a causal
-contact is extremely unlikely unless the galaxy is heavily populated
-by intelligent civilizations with large average lifetimes.
-%
-Roughly, in order to have at least one contact in the entire lifetime
-of the \ceti{}, there should appear a mean of at least one \ceti{}
-every $\sim$20000 yr, with a mean active period of at least
-$\sim$30000 yr.
-%
-This result is qualitatively similar to the results presented by
-several authors, which state that a contact between the Earth and
-another intelligent civilization in the Galaxy is quite unlikely,
-provided the maximum distance of the signal and the lifetime of the
-emitter are not large enough.
-%
-This analysis supports the idea that, in order to increase the
-possibilities of a contact, more active strategies of the emitter
-would be required.
-%
-Some proposals in this direction include intertellar exploration,
-colonization and settlement \citep{brin_great_1983, Dosovic2019,
-galera_invasion_2019}, although it would require large temporal
-scales.
-%
-\citet{Dosovic2019} use probabilistic cellular automaton simulations
-to explore the parameter space of a model with colonization and
-catastrophic events.
-%
-According to the timescales involved, the results could explain the
-Fermi paradox.
-%
-Although our work does not take into account the colonization
-hypothesis, it does consider catastrophic events implicitly in the
-mean value and distribution of the lifetime, $\left<\tau_s\right>$.
-%
-Other strategies could also increase the probability of contacts, for
-example panspermia \citep[e.g.,][]{starling_virulence_2013} or
-self--replicating probes \citep[e.g.,][]{barlow_galactic_2013},
-although they would be too slow to make a significant impact on the
-communication network among intelligent civilizations.
-%
-Our results are also consistent with those presented by
-\citet{grimaldi_signal_2017}, who estimate an upper bound for the mean
-number of extraterrestrial civilizations that could contact Earth
-using Monte Carlo simulations, from a statistical model where the with
-of the Galactic disk is not negligible.
-%
-On the other hand, taking into account the probability of contact with
-radio signals not sent intentionally \citep{horvat_calculating_2007},
-our analysis would imply a highly populated solar neighborhood.
-%
-Remarkably, most of the studies using statistical models or
-simulations are based on the Drake equation
-\citep{cirkovic_temporal_2004, smith_broadcasting_2009,
-bloetscher_using_2019}.
-%
-Our approach does not relay on the Drake equation, and thus is not
-motivated by the detailed physical processes that give rise to
-intelligent life.
-%
-However, we argue that it is a valid empirical formulation to discuss
-the probabilities of contact, and mainly time scales involed in the
-problem.
-%
-Ours is an alternative to the method proposed by
-\citet{balbi_impact_2018}, who perform an analysis based on the Earth
-and use a different model, although our results agree in general.
-
-%%%%%%%%%%%%%%%
-
-There is a balance between the density of \cetis{} and the mean
-lifetime since, as expected, a lower density can be compensated by a
-longer active time period.
-%
-However, a large number of nodes does not easily compensate their
-short lives to reach the same probability of causal contact than in
-the case of a less populated galaxy but with very ancient
-civilizations.
-%
-In all cases, for a short period of time (for instance, the time SETI
-programs have been active on Earth), the maximum probability of making
-a contact occurs at the moment of the awakening.
-%
-This suggests the possibility that an alternative SETI strategy could
-be the search for alternative message carriers, for the case in which
-the search has not been performed on the adequate channels.
-%
-Then, if a contact is produced for the first time, the origin of the
-signal is more likely to be very old.
-%
-Also, the chances of entering the causal connected zone of a \ceti{}
-does not grow linearly, but favours the first period of 100000 years.
-
-In the approach of this work, we used computer simulations to address
-the problem of the probabilities of causal contacts between location
-in the galaxy the possibility of sending and receiving messages.
-%
-Instead of making a number of assumptions, we have explored the
-hypothesis space, reducing the problem to only three parameters and a
-few simple hypothesis to perform a complete model for the population and
-communication network in the galaxy.
-%
-This allows to consider the Fermi paradox from a new perspective, and
-to propose an alternative treatment for the number of intelligent
-emitter/receivers with emphasis on the temporal dimension which limits
-the probabilities of contacts due to the short time interval for
-rise and fall of civilizations compared to the age and extension of
-our Galaxy.
-
-%}}}
-
-\ack[Acknowledgement]{
-%
-This work was partially supported by the Consejo Nacional de
-Investigaciones Cient\'{\i}ficas y T\'ecnicas (CONICET, Argentina),
-the Secretar\'{\i}a de Ciencia y Tecnolog\'{\i}a, Universidad Nacional
-de C\'ordoba, Argentina, and the Universidad Cat\'olica de C\'ordoba,
-Argentina.
-%
-This research has made use of NASA's Astrophysics Data System.
-%
-Plots and simulations were made with software developed by the authors
-using R and python languajes. Plots were postprocessed with inkscape.
-%
-}
-
-\ack[Disclosure statement]{
-%
-No competing financial interests exist.
-%
-}
-
-%\bibliographystyle{mn2e}
-\setlength{\bibsep}{0.0pt}
-
-\begin{thebibliography}{90}
-\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
-
-\bibitem[{Adamic(2000)}]{adamic_zipf_2000}
-Adamic L., 2000, {Zipf, Power-laws, and Pareto - a ranking tutorial}. web
-
-\bibitem[{Anchordoqui \& Weber(2019)}]{anchordoqui_upper_2019}
-Anchordoqui L.~A., Weber S.~M., 2019, arXiv e-prints
-
-\bibitem[{Annis(1999)}]{annis_astrophysical_1999}
-Annis J., 1999, JBIS - Journal of the British Interplanetary Society, 52, 19
-
-\bibitem[{Armstrong \& Sandberg(2013)}]{Armstrong2013}
-Armstrong S., Sandberg A., 2013, Acta Astronautica, 89, 1
-
-\bibitem[{Balb(2018)}]{balb_spatiotemporal_2018}
-Balb A., 2018, Mem. della Soc. Astron. Ital., 89, 425
-
-\bibitem[{Balbi(2018)}]{balbi_impact_2018}
-Balbi A., 2018, Astrobiology, 18, 54
-
-\bibitem[{Barab{\'{a}}si(2009)}]{barabasi_scale_2009}
-Barab{\'{a}}si A.~L., 2009, Science (80-. )., 325, 412
-
-\bibitem[{Barlow(2013{\natexlab{a}})}]{barlow_galactic_2012}
-Barlow M.~T., 2013{\natexlab{a}}, Int. J. Astrobiol., 12, 63
-
-\bibitem[{Barlow(2013{\natexlab{b}})}]{barlow_galactic_2013}
-Barlow M.~T., 2013{\natexlab{b}}, Int. J. Astrobiol., 12, 63
-
-\bibitem[{Benguigui \& Marinov(2016)}]{benguigui_classificacion_2016}
-Benguigui L., Marinov M., 2016, {A classification of the natural and social
-  distributions Part 2: the explanations}. Tech. rep.
-
-\bibitem[{Bloetscher(2019)}]{bloetscher_using_2019}
-Bloetscher F., 2019, Acta Astronaut., 155, 118
-
-\bibitem[{Borra(2012)}]{borra_searching_2012}
-Borra E.~F., 2012, Astronomical Journal, 144
-
-\bibitem[{Brin(1983)}]{brin_great_1983}
-Brin G., 1983, Quarterly Journal of the Royal Astronomical Society, 24, 283
-
-\bibitem[{Burchell(2006)}]{burchell_whither_2006}
-Burchell M.~J., 2006, Int. J. Astrobiol., 5, 243
-
-\bibitem[{Carroll-Nellenback {et~al}\mbox{.}(2019)Carroll-Nellenback, Frank,
-  Wright, \& Scharf}]{carroll_nellemback_fermi_2019}
-Carroll-Nellenback J., Frank A., Wright J., Scharf C., 2019, The Astronomical
-  Journal, 158, 117
-
-\bibitem[{Chung(2003)}]{chung_simulation_2003}
-Chung C.~A., 2003, {Simulation Modeling Handbook: A Practical Approach
-  (Industrial and Manufacturing Engineering Series)}. CRC press
-
-\bibitem[{{\'{C}}irkovi{\'{c}}(2004)}]{cirkovic_temporal_2004}
-{\'{C}}irkovi{\'{c}} M.~M., 2004, Astrobiology, 4, 225
-
-\bibitem[{{Conway Morris}(2018)}]{conway_three_2018}
-{Conway Morris} S., 2018, Int. J. Astrobiol., 17, 287
-
-\bibitem[{Dayal {et~al}\mbox{.}(2016)Dayal, Ward, \&
-  Cockell}]{dayal_habitability_2016}
-Dayal P., Ward M., Cockell C., 2016, arXiv e-prints
-
-\bibitem[{Do{\v{s}}ovi{\'{c}} {et~al}\mbox{.}(2019)Do{\v{s}}ovi{\'{c}},
-  Vukoti{\'{c}}, \& {\'{C}}irkovi{\'{c}}}]{Dosovic2019}
-Do{\v{s}}ovi{\'{c}} V., Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2019,
-  Astronomy and Astrophysics, 625
-
-\bibitem[{Drake(1962)}]{drake_intelligent_1962}
-Drake F., 1962
-
-\bibitem[{Edmondson \& Stevens(2003)}]{Edmondson2003}
-Edmondson W.~H., Stevens I.~R., 2003, International Journal of Astrobiology, 2,
-  231
-
-\bibitem[{Enriquez {et~al}\mbox{.}(2017)Enriquez, Siemion, Foster, Gajjar,
-  Hellbourg, Hickish, Isaacson, Price, Croft, DeBoer, Lebofsky, MacMahon, \&
-  Werthimer}]{enriquez_breakthrough_2017}
-Enriquez J.~E. {et~al.}, 2017, Astrophys. J., 849, 104
-
-\bibitem[{Fogg(1987)}]{fogg_temporal_1987}
-Fogg M.~J., 1987, Icarus, 69, 370
-
-\bibitem[{Forgan {et~al}\mbox{.}(2016)Forgan, Dayal, Cockell, \&
-  Libeskind}]{forgan_evaluating_2015}
-Forgan D., Dayal P., Cockell C., Libeskind N., 2016, Int. J. Astrobiol., 16, 60
-
-\bibitem[{Forgan(2009)}]{forgan_numerical_2009}
-Forgan D.~H., 2009, Int. J. Astrobiol., 8, 121
-
-\bibitem[{Forgan(2011)}]{forgan_spatiotemporal_2011}
-Forgan D.~H., 2011, Int. J. Astrobiol., 10, 341
-
-\bibitem[{Forgan(2013)}]{forgan_possibility_2013}
-Forgan D.~H., 2013, JBIS - J. Br. Interplanet. Soc., 66, 144
-
-\bibitem[{Forgan(2014)}]{forgan_collimated_2014}
-Forgan D.~H., 2014, JBIS - J. Br. Interplanet. Soc., 67, 232
-
-\bibitem[{Forgan(2017{\natexlab{a}})}]{forgan_galactic_2016}
-Forgan D.~H., 2017{\natexlab{a}}, Int. J. Astrobiol., 16, 349
-
-\bibitem[{Forgan(2017{\natexlab{b}})}]{forgan_galactic_2017}
-Forgan D.~H., 2017{\natexlab{b}}, Int. J. Astrobiol., 16, 349
-
-\bibitem[{Forgan(2019)}]{forgan_exoplanet_2017}
-Forgan D.~H., 2019, Int. J. Astrobiol., 18, 189
-
-\bibitem[{Forgan \& Rice(2010)}]{forgan_numerical_2010}
-Forgan D.~H., Rice K., 2010, Int. J. Astrobiol., 9, 73
-
-\bibitem[{Frank(2009)}]{frank_common_2009}
-Frank S.~A., 2009, {The common patterns of nature}
-
-\bibitem[{Funes {et~al}\mbox{.}(2019)Funes, Florio, Lares, \&
-  Asla}]{funes_searching_2019}
-Funes J.~G., Florio L., Lares M., Asla M., 2019, Theol. Sci., 0, 1
-
-\bibitem[{Galera {et~al}\mbox{.}(2019)Galera, Galanti, \&
-  Kinouchi}]{galera_invasion_2019}
-Galera E., Galanti G.~R., Kinouchi O., 2019, International Journal of
-  Astrobiology, 18, 316
-
-\bibitem[{Glade {et~al}\mbox{.}(2012)Glade, Ballet, \&
-  Bastien}]{glade_stochastic_2011}
-Glade N., Ballet P., Bastien O., 2012, Int. J. Astrobiol., 11, 103
-
-\bibitem[{Gobat \& Hong(2016)}]{gobat_evolution_2016}
-Gobat R., Hong S.~E., 2016, Astron. Astrophys., 592
-
-\bibitem[{Gonzalez(2005)}]{gonzalez_habitable_2005}
-Gonzalez G., 2005, Orig. Life Evol. Biosph., 35, 555
-
-\bibitem[{Gonzalez {et~al}\mbox{.}(2001)Gonzalez, Brownlee, \&
-  Ward}]{gonzalez_galactic_2001}
-Gonzalez G., Brownlee D., Ward P., 2001, Icarus, 152, 185
-
-\bibitem[{Grimaldi(2017)}]{grimaldi_signal_2017}
-Grimaldi C., 2017, Sci. Rep., 7, 46273
-
-\bibitem[{Grimaldi {et~al}\mbox{.}(2018)Grimaldi, Marcy, Tellis, \&
-  Drake}]{Grimaldi2018}
-Grimaldi C., Marcy G.~W., Tellis N.~K., Drake F., 2018, Publications of the
-  Astronomical Society of the Pacific, 130
-
-\bibitem[{Haqq-Misra(2019)}]{haqq-misra_evolution_2019}
-Haqq-Misra J., 2019, arXiv e-prints
-
-\bibitem[{Haqq-Misra \& Kopparapu(2018)}]{haqq-misra_drake_2017}
-Haqq-Misra J., Kopparapu R.~K., 2018, Habitability of the Universe Before
-  Earth, 307
-
-\bibitem[{Harp {et~al}\mbox{.}(2018)Harp, Ackermann, Astorga, Arbunich,
-  Barrios, Hightower, Meitzner, Barott, Nolan, Messerschmitt, Vakoch, Shostak,
-  \& Tarter}]{harp_application_2018}
-Harp G.~R. {et~al.}, 2018, Astrophys. J., 869, 66
-
-\bibitem[{Hart(1975)}]{hart_explanation_1975}
-Hart M.~H., 1975, Quarterly Journal of the Royal Astronomical Society, 16, 128
-
-\bibitem[{Hinkel {et~al}\mbox{.}(2019)Hinkel, Hartnett, Lisse, \&
-  Young}]{hinkel_interdisciplinary_2019}
-Hinkel N., Hartnett H., Lisse C., Young P., 2019, Bull. Am. Astron. Soc., 51,
-  497
-
-\bibitem[{Hippke(2017)}]{hippke_interstellar_2017}
-Hippke M., 2017, arXiv e-prints
-
-\bibitem[{Horvat(2006)}]{horvat_calculating_2007}
-Horvat M., 2006, Int. J. Astrobiol., 5, 143
-
-\bibitem[{Horvat {et~al}\mbox{.}(2011)Horvat, Naki{\'{c}}, \&
-  Oto{\v{c}}an}]{horvat_impact_2011}
-Horvat M., Naki{\'{c}} A., Oto{\v{c}}an I., 2011, Int. J. Astrobiol., 11, 51
-
-\bibitem[{Jeong {et~al}\mbox{.}(2000)Jeong, Tombor, Albert, Oltvai, \&
-  Barabasi}]{jeong_large_2000}
-Jeong H., Tombor B., Albert R., Oltvai Z.~N., Barabasi A.~L., 2000, Nature,
-  407, 651
-
-\bibitem[{Lampton(2013)}]{lampton_information_2013}
-Lampton M., 2013, Int. J. Astrobiol., 12, 312
-
-\bibitem[{Lineweaver {et~al}\mbox{.}(2004)Lineweaver, Fenner, \&
-  Gibson}]{lineweaver_galactic_2004}
-Lineweaver C.~H., Fenner Y., Gibson B.~K., 2004, Science (80-. )., 303, 59
-
-\bibitem[{Lingam \& Loeb(2018)}]{lingam_relative_2019}
-Lingam M., Loeb A., 2018, Astrobiology, 19, 28
-
-\bibitem[{Loeb {et~al}\mbox{.}(2016)Loeb, Batista, \&
-  Sloan}]{loeb_relative_2016}
-Loeb A., Batista R.~A., Sloan D., 2016, J. Cosmol. Astropart. Phys., 2016, 040
-
-\bibitem[{Loeb \& Zaldarriaga(2007)}]{loeb_eavesdropping_2006}
-Loeb A., Zaldarriaga M., 2007, J. Cosmol. Astropart. Phys., 1
-
-\bibitem[{Maccone(2010)}]{maccone_KLT_2010}
-Maccone C., 2010, Acta Astronaut., 67, 1427
-
-\bibitem[{Maccone(2011)}]{maccone_mathematical_2011}
-Maccone C., 2011, Orig. Life Evol. Biosph., 41, 609
-
-\bibitem[{MacCone(2011)}]{maccone_SETI_2011}
-MacCone C., 2011, Acta Astronaut., 68, 63
-
-\bibitem[{MacCone(2013)}]{maccone_SETI_2013}
-MacCone C., 2013, Int. J. Astrobiol., 12, 218
-
-\bibitem[{Maccone(2014{\natexlab{a}})}]{maccone_evolution_2014}
-Maccone C., 2014{\natexlab{a}}, Int. J. Astrobiol., 13, 290
-
-\bibitem[{Maccone(2014{\natexlab{b}})}]{maccone_lognormals_2014}
-Maccone C., 2014{\natexlab{b}}, Acta Astronaut., 105, 538
-
-\bibitem[{Maccone(2015)}]{maccone_statistical_2015}
-Maccone C., 2015, Acta Astronaut., 115, 277
-
-\bibitem[{Mart{\'{i}}n \& Goldenfeld(2006)}]{martin_origin_2006}
-Mart{\'{i}}n H.~G., Goldenfeld N., 2006, Proc. Natl. Acad. Sci. U. S. A., 103,
-  10310
-
-\bibitem[{Mitzenmacher(2004)}]{mitzenmacher_brief_2004}
-Mitzenmacher M., 2004, Internet Math., 1, 226
-
-\bibitem[{Morrison \& Gowanlock(2015)}]{morrison_extending_2015}
-Morrison I.~S., Gowanlock M.~G., 2015, Astrobiology, 15, 683
-
-\bibitem[{Murante {et~al}\mbox{.}(2015)Murante, Monaco, Borgani, Tornatore,
-  Dolag, \& Goz}]{murante_simulating_2015}
-Murante G., Monaco P., Borgani S., Tornatore L., Dolag K., Goz D., 2015, Mon.
-  Not. R. Astron. Soc., 447, 178
-
-\bibitem[{Newman(2005)}]{newman_power_2005}
-Newman M., 2005, Contemp. Phys., 46, 323
-
-\bibitem[{Newman \& Sagan(1981)}]{newman_galactic_1981}
-Newman W.~I., Sagan C., 1981, Icarus, 46, 293
-
-\bibitem[{Peters(2018)}]{peters_outer_2018}
-Peters T., 2018, Int. J. Astrobiol., 17, 282
-
-\bibitem[{Prantzos(2013)}]{prantzos_joint_2013}
-Prantzos N., 2013, Int. J. Astrobiol., 12, 246
-
-\bibitem[{Ptolemaeus(2014)}]{ptolemaeus_system_2014}
-Ptolemaeus C., 2014, {System Design, Modeling, and Simulation. Using Ptolemy
-  II}. Ptolemy.org, Berkeley, p. 690
-
-\bibitem[{Rahvar(2017)}]{rahvar_cosmic_2016}
-Rahvar S., 2017, Mon. Not. R. Astron. Soc., 470, 3095
-
-\bibitem[{Ramirez {et~al}\mbox{.}(2017)Ramirez, G{\'{o}}mez-Mu{\~{n}}oz,
-  V{\'{a}}zquez, \& N{\'{u}}{\~{n}}ez}]{ramirez_new_2017}
-Ramirez R., G{\'{o}}mez-Mu{\~{n}}oz M.~A., V{\'{a}}zquez R., N{\'{u}}{\~{n}}ez
-  P.~G., 2017, International Journal of Astrobiology, 1
-
-\bibitem[{Ross(2012)}]{ross_simulation_2012}
-Ross S.~M., 2012, {Simulation}. Elsevier Science Publishing Co Inc
-
-\bibitem[{Simkin \& Roychowdhury(2006)}]{simkin_theory_2006}
-Simkin M.~V., Roychowdhury V.~P., 2006, J. Math. Sociol., 30, 33
-
-\bibitem[{Smith(2009)}]{smith_broadcasting_2009}
-Smith R.~D., 2009, Int. J. Astrobiol., 8, 101
-
-\bibitem[{Solomonides {et~al}\mbox{.}(2016)Solomonides, Kaltenegger, \&
-  Terzian}]{solomonides_probabilistic_2016}
-Solomonides E., Kaltenegger L., Terzian Y., 2016, 228th AAS, San Diego, 228, 1
-
-\bibitem[{Sornette(2006)}]{sornette_critical_2006}
-Sornette D., 2006, {Critical phenomena in natural sciences: Chaos, fractals,
-  selforganization and disorder: Concepts and tools}. Springer
-
-\bibitem[{Sotos(2019)}]{Sotos_biotechnology_2019}
-Sotos J.~G., 2019, {Biotechnology and the lifetime of technical civilizations}
-
-\bibitem[{Starling \& Forgan(2014)}]{starling_virulence_2013}
-Starling J., Forgan D.~H., 2014, Int. J. Astrobiol., 13, 45
-
-\bibitem[{Tarter(2001)}]{tarter_search_2001}
-Tarter J., 2001, Annu. Rev. Astron. Astrophys., 39, 511
-
-\bibitem[{Tarter {et~al}\mbox{.}(2009)Tarter, Backus, Blair, Cordes, Harp,
-  Henry, Horowitz, Howard, Kilsdonk, Korpela, Lasio, Levin, Shostack, \&
-  Werthimer}]{tarter_advancing_2009}
-Tarter J. {et~al.}, 2009, Astro2010 Astron. Astrophys. Decad. Surv. Sci. White
-  Pap. no. 294, 2010
-
-\bibitem[{Vakoch \& Dowd(2015)}]{vakoch_drake_2015}
-Vakoch D.~A., Dowd M.~F., 2015, {The drake equation: Estimating the prevalence
-  of extraterrestrial life through the ages}, Vakoch D.~A., Dowd M.~F., eds.
-  Cambridge University Press, Cambridge, pp. 1--319
-
-\bibitem[{VanHouten(2018)}]{vanhouten_isthere_2017}
-VanHouten M.~A., 2018, Lancet Gastroenterol. Hepatol., 3, 382
-
-\bibitem[{Vukoti{\'{c}} \&
-  {\'{C}}irkovi{\'{c}}(2012)}]{vukotic_astrobiological_2012}
-Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2012, Orig. Life Evol. Biosph.,
-  42, 347
-
-\bibitem[{Vukoti{\'{c}} {et~al}\mbox{.}(2016)Vukoti{\'{c}}, Steinhauser,
-  Martinez-Aviles, {\'{C}}irkovi{\'{c}}, Micic, \&
-  Schindler}]{vukotic_grandeur_2016}
-Vukoti{\'{c}} B., Steinhauser D., Martinez-Aviles G., {\'{C}}irkovi{\'{c}}
-  M.~M., Micic M., Schindler S., 2016, Mon. Not. R. Astron. Soc., 459, 3512
-
-\bibitem[{Walters {et~al}\mbox{.}(1980)Walters, Hoover, \&
-  Kotra}]{walters_interstellar_1980}
-Walters C., Hoover R.~A., Kotra R., 1980, Icarus, 41, 193
-
-\bibitem[{Wright {et~al}\mbox{.}(2015)Wright, Cartier, Zhao, Jontof-Hutter, \&
-  Ford}]{wright_theGsearch_2015}
-Wright J.~T., Cartier K. M.~S., Zhao M., Jontof-Hutter D., Ford E.~B., 2015,
-  Astrophys. J., 816, 17
-
-\bibitem[{Wright {et~al}\mbox{.}(2018)Wright, Kanodia, \&
-  Lubar}]{wright_how_2018}
-Wright J.~T., Kanodia S., Lubar E., 2018, The Astronomical Journal, 156, 260
-
-\end{thebibliography}
-
-\end{document}

### 2020-05-07 submit_v03/manuscript.tex
sec 2

@@
 For the distribution of the lifespans, the author chooses an exponential distribution.
+%
+\citet{balbi_impact_2018} finds that the fraction of communicating civilisations vary with the choice of the statistical distribution for the time of appearance. 
+%
+For all analysed distributions, the fraction corresponding to a mean lifetime of 10$^9$ years and for a maximum radius for the detection of the signal of 1000 yr, is roughly 0.1.
+%
+In another related approach, \citet{vukotic_astrobiological_2012} propose a probabilistic cellular automata (PCA) modelling. 
+%
+In this framework, a complex system is modelled by a lattice of cells which evolve at discrete time steps, according to transition rules that take into account for each cell the states of its neighbour cells.
+%
+The authors implement a PCA model to a network of cells which represent life complexity on a 2-dimensional region resembling the Galactic Habitable Zone (GHZ), 
+an annular ring set between a minimal radius of 6 kpc and a peak radius of 10 kpc. 
+%
+Within this framework, \citet{vukotic_astrobiological_2012} also make several Monte Carlo simulations and analyse ensemble-averaged results.
+%
+The work aims at analysing the evolution of life, although it does not account for the network of causal contacts among technological civilisations.
+
+In this work we address the problem of the temporal and spatial structure of the distribution of communicating civilisations.
+%
+To that end, we reduce and explore the hypothesis space over a set of three of parameters.
+%
+We propose to avoid the frequentist approach of the Drake equation to compute the number of civilisations, providing its statistical distribution instead.
+%
+This is an exploratory analysis that aims at providing a numerical tool to discuss the different scenarios based on statistical heuristics.
+%
+With this, we aim at avoiding the several theoretical problems summarised by the Drake equation factors.
+%
+The approach proposed here should be considered as a compromise between the uncertainties of the frequentist approach and the detailed recipes required on the simulation approach.
+%
+We provide a framework to explore, through a suite of numerical simulations, a parameter space of three unknown observables.
+%
+This allows to discuss possible scenarios and their consequences in terms of the probability of making contacts.
+%
+The method we use for simulating a stochastic process is an approximation that allows to study the behaviour of complex systems, by considering a sequence of well defined discrete events.
+%
+The simulation is carried out by following all the variables that describe and constitute the state of the system, and the evolution of the process is described as a set of changes in that state.
+%
+In this context, an event produces a specific change in the state that can be triggered by random variables that encode the stochastic nature of the physical phenomenon.
+%
+For example, when a new contact is produced between two entities in the simulated galaxy, the numbers of active communication lines and of communicated nodes does change.
+%
+The process involves following the changes on the state of the system, defining the initial and final states.
+%
+This is done by defining a method that allows to keep track of the time progress in steps, and maintaining a list of relevant events.
+%
+In Sec.~\ref{S_methods} we introduce the methods and discuss the candidate distributions for the statistical aspects of the times involved in the communication process.
+%
+We present our results in Sec~\ref{S_results}, with special emphasis on the statistical distributions of the duration of causal contacts in one or both directions and the distribution of time intervals for the waiting of the first contact.
+%
+This quantities are considered as a function of the three simulation parameters.
+%
+In Sec.~\ref{S_discussion} we discuss our results and future research directions. 
+ 
+%}}}
+
+%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
+\section{Methods and working hypotheses}\label{S_methods}
+%{{{
+
+Simulations are suitable tools to analyse systems that evolve with time and involve randomness.
+%
+An advantage of a numerical approach is that usually require fewer assumptions and simplifications, and can be applied to systems where analytical models are not easily found.
+%
+In particular, a suitable tool to model complex stochastic processes through the changes in the state of a system is the discrete--event (hereafter, DE) simulation approach.
+%
+A system described with the DE paradigm is characterised by a set of actors and events.
+%
+Actors interact causally through a series of events on a timeline and process them in chronological order \citep{ptolemaeus_system_2014, chung_simulation_2003, ross_simulation_2012}.
+%
+Each event modify the variables that define the process, producing the corresponding change in the state of the simulated system.
+%
+This method is well suited for the particular case of the diffusion of intelligent signals in the galaxy and allows to explore several models easily.
+%
+We simulate the statistical properties of a set of points in space and time that have a causal connection at light speed, dubbed ‘’Causal Contact Nodes’’ (hereafter, \ceti{}).
+%
+We chose this generic name in order to stress the fact that in this analysis 
+only the causal contact is considered, independently of any broadcasting or lookout activity.
+%
+The system we propose is ideal, in the sense that it considers the special case of a fully efficient node that emits and receives isotropically.
+%
+A causal contact node, considered as a broadcasting station that has the ability to detect signals through an active lookout program, could have a lesser than one efficiency factor, which is not included in the current analysis.
+%
+Also, it is worth mentioning that this is a general approach, and not necessarily a \ceti{} is the host of an intelligent civilization.
+%
+For instance, it can be associated with a planet where life has developed, became intelligent, reached the skills required to find the right communication channel, sustain a search and establish a contact.
+%
+Alternative message processing entities could be considered, for example interstellar beacons where life has ceased to exist but continue with its emission, or communication stations established by probes or left by intelligent beings \citep[see, e.g., ][]{peters_outer_2018, barlow_galactic_2013}.
+%
+In principle, these strategies could affect our results since it would be easier to configure a cluster of \cetis{} that spread in time.
+%
+However, we do not consider these speculative alternatives at this point.
+%
+For the purpose of the present analysis, only the communication capability is relevant, since we study the causal contacts between the locations.
+%
+The system is defined by a number of actors that represent \cetis{} and appear at different instants in time, generating events that produce meaningful changes in the variables that describe the system, i.e., in the arrangement of \cetis{} and their network of causal contacts.
+%
+For example, the appearance of a new \ceti{} in a region filled with a signal emitted by another node, will increase the number of active \cetis{} and the number of pairs of \cetis{} in causal contact.
+%
+Assuming some simple hypotheses, the discrete events method can be performed taking into account a small number of variables, which allow to analyze the variation of the results in the model parameter space.
 
@@
 %-------------------------------------------------------------------
-
-%%%%% SIGUE PUNTO Y SEGUIDO
-%
-%The radius of influence is set to 1000 lyr.
-%
-\citet{balbi_impact_2018} finds that the fraction of communicating
-civilizations vary with the choice of the statistical distribution for
-the time of appearance. For all analyzed distributions, the fraction
-corresponding to a mean lifetime of 10$^9$ years and for a maximum
-radius for the detection of the signal of 1000 yr, is roughly 0.1.
-%
-In another probabilistic approach,
-\citet{vukotic_astrobiological_2012} propose a probabilistic cellular
-automata (PCA) modeling. In this framework, a complex system is
-modeled by a lattice of cells which evolve at discrete time steps,
-according to transition rules that take into account for each cell the
-states of its neighbor cells.
-%
-The authors implement a PCA model to a network of cells which
-represent life complexity on a 2-dimensional annular ring resembling
-the Galactic Habitable Zone (GHZ). 
-%
-The authors set the GHZ between a minimum radius of 6 kpc and a
-maximum radius of 10 kpc. Within this framework, the authors also make
-several MonteCarlo simulations and analyze ensemble-averaged results.
-%
-This work aims at analyzing the evolution of life, although it does
-not account for the network of causal contacts among technological
-civilizations.
-
-In this work we address the problem of the temporal and spatial
-structure of the distribution of communicating civilizations, by
-exploring the hypothesis space over a minimal set of parameters.
-%
-We propose to avoid the frequentist approach of the
-Drake equation and to explore a parameter space, where instead of
-computing a final number, we provide a statistical distribution that
-gives conditional probabilities.
-%
-This is an exploratory analysis that aims at providing a numerical
-tool to discuss not the several theoretical problems summarized by the
-Drake equation factors, but the different scenarios on the basis of
-statistical heuristics.
-%
-The approach proposed here should be considered as a compromise
-between the uncertainties of the frequentist approach and the detailed
-recipes required on the simulation approach.
-%
-We then provide a numerical framework to explore via simulations a
-parameter space of unknown observables in order to discuss different
-scenarios and their consequences in terms of the probability of
-making contacts.
-%
-The method we use for simulating a stochastic process is an
-approximation that allows to study the behavior of complex systems,
-by considering a sequence of well defined discrete events.
-%
-The simulation is carried out by following all the variables that
-describe and constitute the state of the system, and the evolution of
-the process is described as a set of changes in that state.
-%
-In this context, an event produces a specific change in the state,
-that can be triggered by random variables that encode the stochastic
-nature of the physical phenomenon.
-%
-For example, when a new contact is produced between two entities in
-the simulated galaxy, the numbers of active communication lines and of
-active communicated nodes change.
-%
-The process involves following the changes on the state of the
-system, defining the initial and final states, defining a method that
-allows to keep track of the time progress in steps, and maintaining a
-list of relevant events.
-%
-In Sec.~\ref{S_methods} we introduce the methods and discuss the
-candidate distributions for the statistical aspects of the times
-involved in the communication process.
-%
-Then, we present our results in Sec~\ref{S_results}, with special
-emphasis on the statistical distributions of the duration of causal
-contacts in one or both directions, the possible differences on the
-position of a civilization on the Galaxy and the distribution of time
-intervals for the waiting of the first contact, always as a function
-of the three simulation parameters.
-%
-Finally, in Sec.~\ref{S_discussion} we discuss our results and future
-research directions.
-
-%}}}
-
-%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
-\section{Methods and working hypotheses}\label{S_methods}
-%{{{
-
-Simulations are suitable tools to analyze systems that evolve with
-time and involve randomness.
-%
-An advantage of simulations compared to theoretical approaches, is
-that the former usually require less assumptions and simplifications,
-and can be applied to systems where a theoretical model can not be
-found.
-%
-In particular, many complex stochastic processes that can be described
-by the evolution of the state of a system, can be efficiently modeled
-with the discrete--event (hereafter, DE) simulation approach.
-%
-A system described with the DE paradigm is characterized by a set of
-actors and events, where actors interact causally through a series of
-events on a timeline and process these events in chronological order
-\citep{ptolemaeus_system_2014, chung_simulation_2003,
-ross_simulation_2012}.
-%
-Each event yields modifications to the values of the variables that define
-the process, producing the corresponding change in the state of the
-simulated system.
-%
-This method is well suited for the particular case of the diffusion of
-intelligent signals in the galaxy, and allows to explore several
-models easily.
-%
-We simulate the statistical properties of a set of points in space and
-time that have a causal connection at light speed, hereafter dubbed
-''Causal Contact Nodes'' (hereafter, \ceti{}).
-%
-We chose this generic name in order to stress the fact that in this
-analysis no broadcasting or lookout activities are required, and
-therefore not actual messages are considered, only the causal contact.
-%
-For the special case of a fully efficient node that emits and receives
-isotropically, a causal contact node can be considered as a
-broadcasting station that has the ability to detect signals through an
-active lookout program.
-%
-Also, it is worth mentioning that this is a general approach, and not
-necessarily a \ceti{} is the host of an intelligent civilization.
-%
-For instance, it can be associated with a planet where life has developed, became
-intelligent, reached the skills required to find the right
-communication channel, sustain a search and establish a contact.
-%
-Alternative message processing entities could be considered, for
-example interstellar beacons where life has ceased to exist but
-continue with its emission, or communication stations established by
-probes or left intentionally by intelligent beings \citep[see, e.g.,
-][]{peters_outer_2018, barlow_galactic_2013}.
-%
-In principle, these strategies could affect our results since it would
-be easier to configure a cluster of \cetis{} that spread in time.
-%
-However, we do not consider these highly speculative alternatives at
-this point.
-%
-For the purpose of this analysis, only the communication capability is
-relevant, since we study the causal contacts between the locations.
-%
-The system is defined by a number of actors that represent \cetis{}
-and appear at different instants in time, generating events that
-produce meaningful changes in the variables that describe the system,
-i.e., in the arrangement of \cetis{} and their network of causal
-contacts.
-%
-For example, the appearance of a new \ceti{} in a region which is
-filled with a signal emitted by another node, will increase the
-number of active \cetis{} and the number of pairs of \cetis{} in
-causal contact.
-%
-Assuming some simple hypotheses, the discrete events method can be
-performed taking into account a small number of variables, which allow
-to analyze the variation of the results in the model parameter space.
 

### 2020-05-07 submit_v03/manuscript.tex
otro parrafo...

@@
 
-%-------------------------------------------------------------------
-%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%-------------------------------------------------------------------
-
-In what follows, we describe the experimental setup chosen to estimate
-the probabilities of causal contacts and several derived quantities in
-terms of three independent parameters, namely, the mean time span
-between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$,
-their mean lifetime $\left<\tau_s\right>$, and the maximum distance a
-signal can be detected by another \ceti{} ($D_{max}$).
-%
-Intuitively, the shorter the $\tau_a$ parameter and the larger the
-$\tau_s$ and $D_{max}$ parameters are, the greater the probability of
-the existence of causal contacts between pairs of \cetis{} would be.
-%
-We also need to propose theoretical distribution functions for both
-the distribution of lifespans ($\tau_s$) and the distribution of the
-number of \cetis{} per unit time \citep{maccone_evolution_2014,
-Sotos_biotechnology_2019}, related to the time span between the
-appearance of consecutive \cetis{} (since when $\left<\tau_a\right>$
-is shorter, it produces a greater density of \cetis{}).
-%
-The shapes of these distributions are set to a fixed law, as discussed
-in Sec.~\ref{SS_PDF_shape}.
+In what follows, we outline the experimental setup adopted to estimate the probabilities of causal contacts and several derived quantities.
+%
+This is done in terms of three independent parameters, namely, the mean time span between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$, their mean lifetime $\left<\tau_s\right>$, and the maximum distance a signal can be detected by another \ceti{} ($D_{max}$).
+%
+Intuitively, the probability of the existence of causal contacts between pairs of \cetis{} would be larger for smaller $\tau_a$ parameters, higher $\tau_s$ or higher $D_{max}$ parameter values.
+%
+We also propose theoretical distribution functions for both the lifespans ($\tau_s$) and the number of \cetis{} per unit time \citep{maccone_evolution_2014, Sotos_biotechnology_2019}.
+%
+The later is related to the time span between the appearance of consecutive \cetis{} (since when $\left<\tau_a\right>$ is shorter, it produces a greater density of \cetis{}).
+%
+The analytical expresiones for these distributions are set to a fixed law, as discussed in Sec.~\ref{SS_PDF_shape}.
 
@@
 
-\ffn{1} We illustrate in the Fig.~\ref{F_scheme} the schematic
-representation of the region causally connected to a given \ceti{}.
-%
-To this end, we use space--time diagrams, where time is represented on
-the horizontal axis, and space is represented in the vertical axis.
+\ffn{1} Space—time diagrams, where time and space are represented on
+the horizontal and vertical axes, respectively, are suitable tools to
+describe the causal connections among different nodes.
+%
+We illustrate in the Fig.~\ref{F_scheme} the schematic representation
+of the region causally connected to a given \ceti{}.
 %
@@
 %
-Panel (a) of the Figure represents the region $R$ (shaded polygon) in
-space and time which is causally connected to the emitter (left
-vertex), following an event (hereafter A type event) in which the node
-acquires the communication capability, or becomes ''active''.
-%
-The sphere of first contact (SFC) is a sphere centered on the emitter,
+Panel (a) of the Figure represents the region $R$ (shaded polygon)
+which is causally connected to the emitter (left vertex).
+%
+This region develops after an event (hereafter dubbed A-type event) in
+which the node acquires the communication capability, or becomes
+''active''.
+%
+The sphere of first contact (SFC) is centred on the emitter,
 represented in the Figure by the left angle of the region $R$.
@@
 The region which is causally connected to the emitter is then limited
-by these two spheres, and has the shape of a sphere (before the D
-event) or of a spherical shell (after the D event).
+by these two spheres and therefore has the shape of a filled sphere
+(before the D event) or of a spherical shell (after the D event).
 %
@@
 %
-In this scheme, the entire lifetime of a \ceti{} can then be
-represented by a polygon, limited in the spatial axis by the maximum
-distance of the signal $D_{max}$ and in the time axis by the time span
-between the starting time (A event, concave vertex) and the ending
-time (D event, convex vertex).
-%
-This representation allows to easily visualize the important events
-that result from the existence and communications of two \cetis{}.
+In this scheme, the entire lifetime of a \ceti{} can be represented by
+a polygon, as shown in the Figure.
+%
+The polygon is limited in the spatial axis by the maximum distance of
+the signal $D_{max}$ and in the time axis by the time span between the
+A-type event (concave vertex) and the D-type event (convex vertex).
+%
+This representation allows to visualize the important events that
+result from the presence and communications of two \cetis{}.
 %
 The points in time where the \cetis{} acquire their communicating
-capacity are dubbed ''Awakening'' events ($A_1$ and $A_2$ in the
+capacity are dubbed ‘’Awakening’’ events ($A_1$ and $A_2$ in the
 Figure).
 %
-Similarly, the points in time where the \cetis{} loss their
-communicating capacity are dubbed ''Doomsday'' events ($D_1$ and $D_2$
+Similarly, the instants in time where the \cetis{} lose their
+communicating capacity are dubbed ‘’Doomsday’’ events ($D_1$ and $D_2$
 in the Figure).
@@
 one of the \cetis{}, are defined as ''Contact'' events, shown as $C_1$
-and $C_2$ in the Figure.
+and $C_2$.
 %
 Finally, the two points where the contact is lost for each one the
-\cetis{}, are denominated ''Blackout'' events, shown as $B_1$ and
+\cetis{}, are denominated ''Blackout'' events, presented as $B_1$ and
 $B_2$ in the Figure.
@@
 The receiver node E$_1$ can listen signals from the emitter node
-E$_2$, from the first contact event (t=C$_{12}$) up to the last
-contact event (t=B$_{12}$).
+E$_2$, from the first contact event (t=C$_{12}$) until the last
+one (t=B$_{12}$).
 %
@@
 %
-It is important to notice that there is a time delay between the
+It is noteworthy the fact that there is a time delay between the
 contact events of the two \cetis{} involved in this analysis, and also
-a time delay between the two blackout events, so that the time range
-when a bidirectional contact is possible occurs between the maximum
-time of the contact points and the minimum time between the blackout
-points.
+a time delay between the two blackout events.
+%
+Therefore, the time range when a bidirectional contact is possible
+occurs between the maximum time of the contact points and the minimum
+time between the blackout points.
 
@@
 
-The temporal structure that emerges naturally from the experimental
-setup implies that,
-as a fundamental property of the simulations,
-a causal connection can be produced without the need of the two
-involved locations being concurrent or active at the same time.
-%
-This property of the system arises as a
-consequence of the large spatial and temporal scales, where a message
-is transmitted at the (relatively small) light speed.
-%
-Although a \ceti{} could be active for enough time so as to transmit a
-message at large distances, the limited power of the message and the
-dilution that runs as the squared distance from the source, imposes a
-detectability limit.
-%
-As a consequence of this limitation and of their finite lifetime,
-considered as the period of time between the acquisition and loss of
-communicating capacity, each \ceti{} will fill a spherical shell
-region of the galaxy, limited by two concentric spherical surfaces.
-%
-The leading front, or surface of first contact (SFC) grows from the
-central \ceti{} until it reaches the maximum distance $D_{max}$.
-%
-Thus, the volume reached causally by a \ceti{} is initially a growing
-sphere.
-%
-Following the end of the civilization, there is still a region which
-is filled with the emitted signals.
-%
-This approach has been also considered other statistical models
-\citep[e.g., ][]{smith_broadcasting_2009, grimaldi_signal_2017, Grimaldi2018}.
-%
-The trailing front, or surface of last contact (SLC) also grows from
-the central \ceti{}, with a delay with respect to the SFC equivalent
-to the lifetime of the node, and produces a spherical shell region.
-%
-Any other \ceti{} within this region will be aware of the originating
-node, even if it has disappeared before the time of contact.
-%
-This region will grow if the surface of first contact has not yet
-reached the maximum distance $D_{max}$, and will shrink otherwise
-until the surface of last contact reaches $D_{max}$, producing as a
-result the loss of all signal from the central \ceti{}.
-%
-In our approach, we consider a model galaxy where the width of the
-disk is negligible with respect to the radius of the disk.
-%
-In the 2D simulation only the intersection of the communicating
-spherical shells with the plane of the galaxy is relevant, and produce
-the corresponding circles or rings for the filled spheres or annular
-regions of the spherical shells, respectively.
-%
-The initially growing communicating sphere, is shown over space--time
-diagrams, where time is represented on the horizontal direction, and
-space is represented in the vertical direction.
-%
-In the Fig.~\ref{F_scheme} the two
-emitters in panel (b), E$_1$ and E$_2$, reach each other at different times.
+The temporal structure that emerges from the experimental setup implies that, as a fundamental property of the simulations, a causal connection can be produced without 
+requiring that the two nodes are active at the same time.
+%
+This property of the system arises as a consequence of the large spatial and temporal scales, where a message is transmitted at the (relatively small) light speed.
+%
+Although a \ceti{} could be active for a large enough period to transmit a message at large distances, the limited power of the message and the dilution that depends on the squared distance from the source imposes a detectability limit.
+%
+As a consequence of this limitation and of their finite lifetime, considered as the period between the acquisition and loss of communicating capacity, each \ceti{} will fill a spherical shell region of the galaxy, limited by two concentric spherical surfaces.
+%
+The leading front, or surface of first contact (SFC) grows from the central \ceti{} until it reaches the maximum distance $D_{max}$.
+%
+Following the end of the civilization, there is still a region which is filled with the emitted signals.
+%
+This approach has been also considered in other statistical models \citep[e.g., ][]{smith_broadcasting_2009, grimaldi_signal_2017, Grimaldi2018}.
+%
+The trailing front, or surface of last contact (SLC) also grows from the central \ceti{}, with a delay with respect to the SFC equivalent to the lifetime of the node, and produces a spherical shell region.
+%
+Any other \ceti{} within this region will be aware of the originating node, even if it has disappeared before the time of contact.
+%
+This region will grow if the surface of first contact has not yet reached the maximum distance $D_{max}$, and will shrink after a D-type event until the surface of last contact reaches $D_{max}$, producing as a result the loss of all signal from the central \ceti{}.
+%
+In our approach, we consider a model galaxy where the width of the disk is negligible with respect to the radius of the disk.
+%
+In the 2D simulation only the intersection of the communicating spherical shells with the plane of the galaxy is relevant, and produce the corresponding circles or rings for the filled spheres or annular regions of the spherical shells, respectively.
+%
+The initially growing communicating sphere is shown over space--time diagrams, where time is represented on the horizontal direction, and space is represented in the vertical direction.
+%
+In the Fig.~\ref{F_scheme} the two emitters in panel (b), E$_1$ and E$_2$, reach each other at different times.
 %
@@
 %
-Emitter $i$ can listen to emitter $j$ between C$_{ij}$ and B$_{ij}$.
-%
-The type and length of causal contact in both directions depend on the
-distance and time lag between the awakening events, the maximum
-distance that a signal can reach and the time period in which each
-emitter is active.
-%
+Emitter $i$ can listen to emiter $j$ between C$_{ij}$ and B$_{ij}$.
+%
+The type and length of causal contact in both directions depend on the distance and time lag between the awakening events, the maximum distance that a signal can reach and the time period in which each emitter is active.
+
+%-------------------------------------------------------------------
+%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%-------------------------------------------------------------------
+
 \ffn{2}
 %
-In the Fig.~\ref{F_messages} we show a schematic representation of the
-possible cases in which a \ceti{} E can be in causal contact with a
-\ceti{} R.
-%
-The duration of the causal contact in one direction depends on several
-factors, mainly the lifetimes, the time period between the A events
-and the distance between the two nodes involved in the causal
-contact.
-%
-As it can be seen, the direction of the causal contact does not imply
-any ordering for the A events of the involved nodes.
-%
-In the implementation of the DE simulation, the six cases shown in the
-Figure determine the succession of events.
-%
-Each time an event is produced, the list of upcoming events is updated
-according to these cases.
+In the Fig.~\ref{F_messages} we show a schematic representation of the possible cases in which a \ceti{} E can be in causal contact with a \ceti{} R.
+%
+The duration of the causal contact in one direction depends on several factors, mainly the lifetimes, the time period between the A events and the distance between the two nodes involved in the causal contact.
+%
+As it can be seen, the direction of the causal contact does not imply any ordering for the A events of the involved nodes.
+%
+In the implementation of the DE simulation, the six cases shown in the Figure determine the succession of events.
+%
+Each time an event is produced, the list of upcoming events is updated according to these cases.
 

### 2020-05-09 submit_v03/manuscript.tex
more...

@@
 The type and length of causal contact in both directions depend on the distance and time lag between the awakening events, the maximum distance that a signal can reach and the time period in which each emitter is active.
+%
+\ffn{2}
+%
+In the Fig.~\ref{F_messages} we show a schematic representation of the possible cases in which a \ceti{} E (standing for Emiter) can be in causal contact with a \ceti{} R (receiver).
+%
+The duration of such a contact in one way depends on several factors, mainly the lifetimes, the time delay between the A-type events and the distance between the two nodes.
+%
+As it can be seen, the direction of the causal contact does not imply any ordering for the A-type events of the involved emitter/receiver pair.
+%
+For instance, a node can be in the light cone of another node that turned up later, i.e., the receiver is older than the emitter, like in the cases (e) and (f) of the Figure.
+%
+In the implementation of the DE simulation, all possibilities for causal contacts are reduced to the six cases shown in the Figure.
+%
+Once a pair of nodes become active and their properties have been assigned, it is possible to determine if and when they will get connected.
+%
+The times derived for such events are added to a stack of times so that they are visited in order, and in turn trigger further events.
+%
+% EJEMPLOS???
+
+In our experimental configuration the simulation starts assuming the stochastic process is already stable, and finishes before any galactic evolution effect could modify the fixed values of the variables.
+%
+Likewise, we assume that the probability for the appearance of a \ceti{} is homogeneous over the GHZ.
+%
+Although the Galaxy has a well-known spiral structure, the assumed sparsity of nodes (arguably, intelligent civilisations) does make that assumption tenable.
+%
+If the distribution of nodes is not sparse, as it could be the case if the spiral arms host most of the \cetis{}, then contacts would be more frequent between closely located nodes.
+%
+In such a case our results would underestimate the number of contacts between close pairs of nodes within the same spiral arm, and conversely, overestimate the number of contacts between separated nodes.
+%
+We also limit the possibilities of life or other types of civilisations to the usually stated hypothesis for the definition of
+the GHZ \citep{dayal_habitability_2016, gonzalez_galactic_2001, lineweaver_galactic_2004, gonzalez_habitable_2005,
+morrison_extending_2015, haqq-misra_evolution_2019, rahvar_cosmic_2016, gobat_evolution_2016, rahvar_cosmic_2016} and consider that habitability remains constant over time \citep[see, however, ][]{gonzalez_habitable_2005, dayal_habitability_2016, gobat_evolution_2016}.
+%
+For example, \citet{vukotic_astrobiological_2012} propose
+probabilistic cellular automata model to explain the astrobiological
+history of the Milky Way. % <<<< FEO
+%
+This means that we set aside civilisations that could survive in severe conditions or unstable systems, which would prevent the appearance of life as we know it.
 
@@
 
-\ffn{2}
-%
-In the Fig.~\ref{F_messages} we show a schematic representation of the possible cases in which a \ceti{} E can be in causal contact with a \ceti{} R.
-%
-The duration of the causal contact in one direction depends on several factors, mainly the lifetimes, the time period between the A events and the distance between the two nodes involved in the causal contact.
-%
-As it can be seen, the direction of the causal contact does not imply any ordering for the A events of the involved nodes.
-%
-In the implementation of the DE simulation, the six cases shown in the Figure determine the succession of events.
-%
-Each time an event is produced, the list of upcoming events is updated according to these cases.
-
-In out experimental configuration, there is no a temporal window for
-the model, i.e., the simulation starts when the process is already
-stable, and ends before any galactic effect could modify or affect of
-fixed variables.
-%
-Also, we assume that the probability for the rise of a \ceti{} is
-homogeneous over the GHZ.
-%
-Although the Galaxy has a well known spiral structure, the assumed
-sparsity of nodes (intelligent civilizations) does make this
-assumption fairly reasonable.
-%
-Otherwise, if the distribution of nodes is not sparse, it could be the
-case that the spiral arms would host most of the \cetis{}, and
-contacts are frequent between closely located nodes.
-%
-In such a case we could expect that the interpretation of our results
-could be considered as pessimist for the establishment of a contact
-with our planet.
-%
-We also limit the possibilities of life or other types of
-civilizations to the hypothesis usually stated for the definition of
-the GHZ \citep{dayal_habitability_2016, gonzalez_galactic_2001,
-lineweaver_galactic_2004, gonzalez_habitable_2005,
-morrison_extending_2015, haqq-misra_evolution_2019,
-rahvar_cosmic_2016, gobat_evolution_2016, rahvar_cosmic_2016} and
-consider that habitability remains constant \citep[see, however,
-][]{gonzalez_habitable_2005, dayal_habitability_2016, gobat_evolution_2016}.
-%
-For example, \citet{vukotic_astrobiological_2012} propose
-probabilistic cellular automata model to explain the astrobiological
-history of the Milky Way.
-%
-This means that we set aside possible civilizations that could survive
-in severe conditions or unstable systems, which would prevent the
-appearance of life as we know it.
-
-While the aforementioned assumptions are basic conditions for most of
-the stochastic processes observed in nature, we need to make stronger
-assumptions related to the nature of the message.
-%
-The most simple assumption about the message itself, is that it
-travels at light speed.
-%
-With this, we are considering messages sent through, for example,
-electromagnetic radiation or even gravitational radiation or neutrino
-communication \citep{hippke_interstellar_2017, wright_how_2018}, but we set
-apart messages sent with mechanical means or physical objects
-\citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown technology
-that violates the known laws of physics.
-% 
-For the communication of \cetis{} through messages sent isotropically,
-we assume that the capacity to emit signals and to receive signals
-occur at the same time.
-%
-This means that the ability to find another civilization develops at
-the same time that the ability and intention to emit a signal in all
-directions, intended to be detected by an unknown civilization.
-%
-Although there are several reasons to think that this could not be the
-case, at large time scales it can be considered that both abilities
-occur roughly at coincident epochs.
-%
-Another essential assumption is that all \cetis{} use the same signal
-power, so that there is a maximum distance out to which it can be
-detected.
-%
-It is worth mentioning that we are considering in our experimental
-setup a system composed by several emitters and receivers across the
-galaxy, under the hypothesis proposed before.
-%
-In such a system we compute probabilities of a \ceti{} making contact
-with another node, and the results must be interpreted as for a
-randomly chosen \ceti{}, given certain conditions of homogeneity on an
-ensemble of equivalent galaxies.
-%
-Then, we stress the fact that we are not centered on the probabilities
-for the Earth itself, but on a generalized \ceti{}.
-%
-Regarding the extent of the signals, we know that the distance from
-which a signal from Earth could be detected using the current
-technology is of a few hundreds light years, given that the signal was
-sent to a specific direction.
-%
+While the aforementioned assumptions are basic conditions for most of the stochastic processes observed in nature, we need to make stronger hypotheses related to the nature of the causal contact, which is not necessarily tied to any actual message.
+%
+The contact has, instead, the meaning of a belonging of a node to the light cone of another one.
+%
+We are considering messages sent through, for example, electromagnetic radiation, gravitational radiation or neutrino communication \citep{hippke_interstellar_2017, wright_how_2018}, but we set apart messages sent with mechanical means or physical objects \citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown technology that violates the known laws of physics.
+% 
+For the communication of \cetis{} through messages sent isotropically, we assume that the capacity to emit signals and to receive signals occur at the same time.
+%
+This means that the ability to find another civilization develops at the same time that the ability and intention to emit a signal in all directions, intended to be detected by an unknown civilization.
+%
+Although there are several reasons to think that this could not be the case, at large time scales it can be considered that both abilities occur roughly at coincident epochs.
+%
+Another essential assumption is that all \cetis{} use the same signal power, so that there is a maximum distance out to which it can be detected.
+%
+It is worth mentioning that we are considering in our experimental setup a system composed by several emitters and receivers across the galaxy, under the hypothesis proposed before.
+%
+In such a system we compute probabilities of a \ceti{} making contact with another node, and the results must be interpreted as for a randomly chosen \ceti{}, given certain conditions of homogeneity on an ensemble of equivalent galaxies.
+%
+Then, we stress the fact that we are not centered on the probabilities for the Earth itself, but on a generalized \ceti{}.
+%
+Regarding the extent of the signals, we know that the distance from which a signal from Earth could be detected using the current technology is of a few hundreds light years, given that the signal was sent to a specific direction.
+%
+It is straightforward to propose and implement a distribution of maximum distances, although this would increase the model complexity at the cost of a larger uncertainty or of another dimension on the hypothesis space.
+
 %https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
@@
 %
-It is straightforward to propose and implement a distribution of
-maximum distances, although this would increase the model complexity
-at the cost of a larger uncertainty or of another dimension on the
-hypothesis space.
 %

### 2020-05-13 submit_v03/manuscript.tex
soft biblio

@@
 \newcommand{\cetis}{CCNs}
+\newcommand{\hs}{\textsc{hearsay}}
 
@@
 % 
-% 
-% 
-% 
-% 
-% 
-% 
-% 
 \Fpagebreak
 %
-We stress the fact that the only observation that can be stated with certainty is that for the number of years projects for the Search of Extraterrestrial Intelligences (SETI) have been working; we have received no signal using the technical possibilities and the conditions established by them \citep{tarter_search_2001}.
-%
-The absence of detection has also motivated alternative ideas for new SETI strategies \citep{forgan_exoplanet_2017, balbi_impact_2018, loeb_eavesdropping_2006, maccone_KLT_2010, tarter_advancing_2009, enriquez_breakthrough_2017, loeb_relative_2016, maccone_SETI_2011, lingam_relative_2019, wright_theGsearch_2015, maccone_SETI_2013, maccone_lognormals_2014, harp_application_2018, forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
-%
-The Drake equation is a key tool to organise the discussion about the problem of the (unknown) abundance of civilisations in the Galaxy \citep{hinkel_interdisciplinary_2019}.
-%
-However, the uncertainties in the factors of this equation make it less prone to a formal study to define searching strategies or to compute the actual number of extraterrestrial intelligences.
-%
-The optimistic estimations from the Drake equation also contrast with the SETI initiatives, noticeable overshadowed because despite decades of effort, not a single evidence for intelligent signals has been detected.
-%
-\citet{prantzos_joint_2013} proposes a unified framework for a joint analysis of the Drake equation and the Fermi paradox, concluding that for sufficiently long-lived civilisations, colonisation of the Galaxy is the only reasonable option to gain awareness about other life forms.
-%
-\citet{haqq-misra_drake_2017} discuss the dependence of the Drake equation parameters on the spectral type of the host stars and the time since the Galaxy formed and examine trajectories for the emergence of communicative civilisations.
-%
-Some modifications to the original idea of the equation have tried to imprint a stochastic nature, to propose a probabilistic approach, or to consider the temporal structure missing in the equation.
-%
-Temporal aspects of the distribution of communicating civilisations and their contacts have also been explored by several authors \citep{fogg_temporal_1987, forgan_spatiotemporal_2011, balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011}, as well as efforts on considering the stochastic nature of the Drake equation \citep{glade_stochastic_2011}.
-%
-The literature about this topic brings a number of works that consider numerical simulations
-\citep{forgan_evaluating_2015, vukotic_grandeur_2016, murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017, ramirez_new_2017}.
-%
-Although this approach does not rely on observed quantities that could be modified with time, it also depend on the definition of a number of parameters which are unknown or uncertain.
-%
-\citet{grimaldi_signal_2017} consider a statistical model for the probability of the Earth contacting other intelligent civilisations, taking into account the finite lifetime of signal emitters, and based on the fractional volume occupied by all signals reaching our planet.
-
-In a recent work, \citep{balbi_impact_2018} use a statistical model to analyse the occurrence of causal contacts between civilisations in the Galaxy.
-%
-The author highlights the effect of evolutionary processes when attempting to estimate the number of communicating civilisations that might be in causal contact with an observer on the Earth.
-%
-\citet{cirkovic_temporal_2004} also emphasise the lack of temporal structure in the Drake equation.
-%
-\citet{bloetscher_using_2019} uses a probabilistic approach, though still motivated by the Drake equation, to obtain a probabilistic measure for the number of civilisations in the Galaxy.
-%
-To that end, the authors do Monte-Carlo Markov Chains over each factor of the Drake equation, and combine the means to reach a probabilistic result.
-%
-It is worth mentioning that the author proposes a log-normal target distribution to compute the posterior probabilities.
-%
-This study concludes that there is a tiny probability for Earth to contact another extraterrestrial civilization.    
-%
-\citet{smith_broadcasting_2009} uses an analytical model to gauge the probabilities of contact between two randomly located civilisations, and the waiting time for the first contact, assuming a fixed maximum broadcasting distance.
-% 
-Adopting a novel approach, \citet{balbi_impact_2018} investigate the chance of communicating civilisations making causal contact within a volume surrounding the location of the Earth, using a statistical model to compute the occurrence of causal contacts.
-%
-The author stresses the fact that the causal contact requirement involves the distance between civilisations, their lifespan and their times of appearance.
-This is important since the time it takes the light to travel across the Galaxy is much lesser than its lifetime.
-%
-\citet{balbi_impact_2018} fix the total number of civilisations and explore the parameter space that comprise three parameters, namely, the distance to the Earth, the time of appearance and the lifespan of the communicating civilization.
-%
-Each of these three variables are drawn from a random distribution. The distribution of the distances results from a uniform distribution of civilizations within the plane of the Galaxy.
-%
-For the distribution of the characteristic time of appearance, the author explores an exponential and a truncated Gaussian distributions.
-For the distribution of the lifespans, the author chooses an exponential distribution.
-%
-\citet{balbi_impact_2018} finds that the fraction of communicating civilisations vary with the choice of the statistical distribution for the time of appearance. 
-%
-For all analysed distributions, the fraction corresponding to a mean lifetime of 10$^9$ years and for a maximum radius for the detection of the signal of 1000 yr, is roughly 0.1.
-%
-In another related approach, \citet{vukotic_astrobiological_2012} propose a probabilistic cellular automata (PCA) modelling. 
-%
-In this framework, a complex system is modelled by a lattice of cells which evolve at discrete time steps, according to transition rules that take into account for each cell the states of its neighbour cells.
-%
-The authors implement a PCA model to a network of cells which represent life complexity on a 2-dimensional region resembling the Galactic Habitable Zone (GHZ), 
-an annular ring set between a minimal radius of 6 kpc and a peak radius of 10 kpc. 
-%
-Within this framework, \citet{vukotic_astrobiological_2012} also make several Monte Carlo simulations and analyse ensemble-averaged results.
-%
-The work aims at analysing the evolution of life, although it does not account for the network of causal contacts among technological civilisations.
-
-In this work we address the problem of the temporal and spatial structure of the distribution of communicating civilisations.
-%
-To that end, we reduce and explore the hypothesis space over a set of three of parameters.
-%
-We propose to avoid the frequentist approach of the Drake equation to compute the number of civilisations, providing its statistical distribution instead.
-%
-This is an exploratory analysis that aims at providing a numerical tool to discuss the different scenarios based on statistical heuristics.
-%
-With this, we aim at avoiding the several theoretical problems summarised by the Drake equation factors.
-%
-The approach proposed here should be considered as a compromise between the uncertainties of the frequentist approach and the detailed recipes required on the simulation approach.
-%
-We provide a framework to explore, through a suite of numerical simulations, a parameter space of three unknown observables.
-%
-This allows to discuss possible scenarios and their consequences in terms of the probability of making contacts.
-%
-The method we use for simulating a stochastic process is an approximation that allows to study the behaviour of complex systems, by considering a sequence of well defined discrete events.
-%
-The simulation is carried out by following all the variables that describe and constitute the state of the system, and the evolution of the process is described as a set of changes in that state.
-%
-In this context, an event produces a specific change in the state that can be triggered by random variables that encode the stochastic nature of the physical phenomenon.
-%
-For example, when a new contact is produced between two entities in the simulated galaxy, the numbers of active communication lines and of communicated nodes does change.
-%
-The process involves following the changes on the state of the system, defining the initial and final states.
-%
-This is done by defining a method that allows to keep track of the time progress in steps, and maintaining a list of relevant events.
-%
-In Sec.~\ref{S_methods} we introduce the methods and discuss the candidate distributions for the statistical aspects of the times involved in the communication process.
-%
-We present our results in Sec~\ref{S_results}, with special emphasis on the statistical distributions of the duration of causal contacts in one or both directions and the distribution of time intervals for the waiting of the first contact.
-%
-This quantities are considered as a function of the three simulation parameters.
-%
-In Sec.~\ref{S_discussion} we discuss our results and future research directions. 
- 
+We stress the fact that the only observation that can be stated with
+certainty is that for the number of years projects for the Search of
+Extraterrestrial Intelligences (SETI) have been working; we have
+received no signal using the technical possibilities and the
+conditions established by them \citep{tarter_search_2001}.
+%
+The absence of detection has also motivated alternative ideas for new
+SETI strategies \citep{forgan_exoplanet_2017, balbi_impact_2018,
+loeb_eavesdropping_2006, maccone_KLT_2010, tarter_advancing_2009,
+enriquez_breakthrough_2017, loeb_relative_2016, maccone_SETI_2011,
+lingam_relative_2019, wright_theGsearch_2015, maccone_SETI_2013,
+maccone_lognormals_2014, harp_application_2018,
+forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
+%
+The Drake equation is a key tool to organise the discussion about the
+problem of the (unknown) abundance of civilisations in the Galaxy
+\citep{hinkel_interdisciplinary_2019}.
+%
+However, the uncertainties in the factors of this equation make it
+less prone to a formal study to define searching strategies or to
+compute the actual number of extraterrestrial intelligences.
+%
+The optimistic estimations from the Drake equation also contrast with
+the SETI initiatives, noticeable overshadowed because despite decades
+of effort, not a single evidence for intelligent signals has been
+detected.
+%
+\citet{prantzos_joint_2013} proposes a unified framework for a joint
+analysis of the Drake equation and the Fermi paradox, concluding that
+for sufficiently long-lived civilisations, colonisation of the Galaxy
+is the only reasonable option to gain awareness about other life
+forms.
+%
+\citet{haqq-misra_drake_2017} discuss the dependence of the Drake
+equation parameters on the spectral type of the host stars and the
+time since the Galaxy formed and examine trajectories for the
+emergence of communicative civilisations.
+%
+Some modifications to the original idea of the equation have tried to
+imprint a stochastic nature, to propose a probabilistic approach, or
+to consider the temporal structure missing in the equation.
+%
+Temporal aspects of the distribution of communicating civilisations
+and their contacts have also been explored by several authors
+\citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
+balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011}, as
+well as efforts on considering the stochastic nature of the Drake
+equation \citep{glade_stochastic_2011}.
+%
+The literature about this topic brings a number of works that consider
+numerical simulations \citep{forgan_evaluating_2015,
+vukotic_grandeur_2016, murante_simulating_2015, forgan_numerical_2009,
+forgan_galactic_2017, ramirez_new_2017}.
+%
+Although this approach does not rely on observed quantities that could
+be modified with time, it also depend on the definition of a number of
+parameters which are unknown or uncertain.
+%
+\citet{grimaldi_signal_2017} consider a statistical model for the
+probability of the Earth contacting other intelligent civilisations,
+taking into account the finite lifetime of signal emitters, and based
+on the fractional volume occupied by all signals reaching our planet.
+
+In a recent work, \citep{balbi_impact_2018} use a statistical model to
+analyse the occurrence of causal contacts between civilisations in the
+Galaxy.
+%
+The author highlights the effect of evolutionary processes when
+attempting to estimate the number of communicating civilisations that
+might be in causal contact with an observer on the Earth.
+%
+\citet{cirkovic_temporal_2004} also emphasise the lack of temporal
+structure in the Drake equation.
+%
+\citet{bloetscher_using_2019} uses a probabilistic approach, though
+still motivated by the Drake equation, to obtain a probabilistic
+measure for the number of civilisations in the Galaxy.
+%
+To that end, the authors do Monte-Carlo Markov Chains over each factor
+of the Drake equation, and combine the means to reach a probabilistic
+result.
+%
+It is worth mentioning that the author proposes a log-normal target
+distribution to compute the posterior probabilities.
+%
+This study concludes that there is a tiny probability for Earth to
+contact another extraterrestrial civilization.    
+%
+\citet{smith_broadcasting_2009} uses an analytical model to gauge the
+probabilities of contact between two randomly located civilisations,
+and the waiting time for the first contact, assuming a fixed maximum
+broadcasting distance.
+% 
+Adopting a novel approach, \citet{balbi_impact_2018} investigate the
+chance of communicating civilisations making causal contact within a
+volume surrounding the location of the Earth, using a statistical
+model to compute the occurrence of causal contacts.
+%
+The author stresses the fact that the causal contact requirement
+involves the distance between civilisations, their lifespan and their
+times of appearance. This is important since the time it takes the
+light to travel across the Galaxy is much lesser than its lifetime.
+%
+\citet{balbi_impact_2018} fix the total number of civilisations and
+explore the parameter space that comprise three parameters, namely,
+the distance to the Earth, the time of appearance and the lifespan of
+the communicating civilization.
+%
+Each of these three variables are drawn from a random distribution.
+The distribution of the distances results from a uniform distribution
+of civilizations within the plane of the Galaxy.
+%
+For the distribution of the characteristic time of appearance, the
+author explores an exponential and a truncated Gaussian distributions.
+For the distribution of the lifespans, the author chooses an
+exponential distribution.
+%
+\citet{balbi_impact_2018} finds that the fraction of communicating
+civilisations vary with the choice of the statistical distribution for
+the time of appearance. 
+%
+For all analysed distributions, the fraction corresponding to a mean
+lifetime of 10$^9$ years and for a maximum radius for the detection of
+the signal of 1000 yr, is roughly 0.1.
+%
+In another related approach, \citet{vukotic_astrobiological_2012}
+propose a probabilistic cellular automata (PCA) modelling. 
+%
+In this framework, a complex system is modelled by a lattice of cells
+which evolve at discrete time steps, according to transition rules
+that take into account for each cell the states of its neighbour
+cells.
+%
+The authors implement a PCA model to a network of cells which
+represent life complexity on a 2-dimensional region resembling the
+Galactic Habitable Zone (GHZ),  an annular ring set between a minimal
+radius of 6 kpc and a peak radius of 10 kpc. 
+%
+Within this framework, \citet{vukotic_astrobiological_2012} also make
+several Monte Carlo simulations and analyse ensemble-averaged results.
+%
+The work aims at analysing the evolution of life, although it does not
+account for the network of causal contacts among
+technological civilisations.
+
+In this work we address the problem of the temporal and spatial
+structure of the distribution of communicating civilisations.
+%
+To that end, we reduce and explore the hypothesis space over a set of
+three of parameters.
+%
+We propose to avoid the frequentist approach of the Drake equation to
+compute the number of civilisations, providing its statistical
+distribution instead.
+%
+This is an exploratory analysis that aims at providing a numerical
+tool to discuss the different scenarios based on statistical
+heuristics.
+%
+With this, we aim at avoiding the several theoretical problems
+summarised by the Drake equation factors.
+%
+The approach proposed here should be considered as a compromise
+between the uncertainties of the frequentist approach and the detailed
+recipes required on the simulation approach.
+%
+We provide a framework to explore, through a suite of
+numerical simulations, a parameter space of three unknown observables.
+%
+This allows to discuss possible scenarios and their consequences in
+terms of the probability of making contacts.
+%
+The method we use for simulating a stochastic process is an
+approximation that allows to study the behaviour of complex systems,
+by considering a sequence of well defined discrete events.
+%
+The simulation is carried out by following all the variables that
+describe and constitute the state of the system, and the evolution of
+the process is described as a set of changes in that state.
+%
+In this context, an event produces a specific change in the state that
+can be triggered by random variables that encode the stochastic nature
+of the physical phenomenon.
+%
+For example, when a new contact is produced between two entities in
+the simulated galaxy, the numbers of active communication lines and of
+communicated nodes does change.
+%
+The process involves following the changes on the state of the system,
+defining the initial and final states.
+%
+This is done by defining a method that allows to keep track of the
+time progress in steps, and maintaining a list of relevant events.
+%
+In Sec.~\ref{S_methods} we introduce the methods and discuss the
+candidate distributions for the statistical aspects of the times
+involved in the communication process.
+%
+We present our results in Sec~\ref{S_results}, with special emphasis
+on the statistical distributions of the duration of causal contacts in
+one or both directions and the distribution of time intervals for the
+waiting of the first contact.
+%
+This quantities are considered as a function of the three simulation
+parameters.
+%
+In Sec.~\ref{S_discussion} we discuss our results and future research
+directions.   
 %}}}
@@
 
-%-------------------------------------------------------------------
-%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%-------------------------------------------------------------------
-
-While the aforementioned assumptions are basic conditions for most of the stochastic processes observed in nature, we need to make stronger hypotheses related to the nature of the causal contact, which is not necessarily tied to any actual message.
-%
-The contact has, instead, the meaning of a belonging of a node to the light cone of another one.
-%
-We are considering messages sent through, for example, electromagnetic radiation, gravitational radiation or neutrino communication \citep{hippke_interstellar_2017, wright_how_2018}, but we set apart messages sent with mechanical means or physical objects \citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown technology that violates the known laws of physics.
-% 
-For the communication of \cetis{} through messages sent isotropically, we assume that the capacity to emit signals and to receive signals occur at the same time.
-%
-This means that the ability to find another civilization develops at the same time that the ability and intention to emit a signal in all directions, intended to be detected by an unknown civilization.
+We need to raise the scope of our results by stressing the fact that we are considering causal contacts instead of actual contacts.
+%
+By a causal contact we mean a configuration, in space and time, in which a node belongs to the light cone of another node.
+%
+This means that a causal contact does not need to be tied to any actual message emission or detection.
+%
+In more realistic scenarios, there are several sources of ‘’signal loss’’ with respect to the ideal case.
+%
+Among them, we can mention temporal and signal power aspects and direction dependent communication capabilties.
+%
+The results must then be interpreted as the case of ideal nodes, with a perfect efficiency in the emision and reception of signals.
+%
+The probabilities of contacts presented here are then upper limits to the number of communications between civilization s in the Galaxy, given the implemented hipotheses in the model.
+%
+The use of light cones as causal contact regions is inspired by the fact that light—speed traveling messengers like 
+electromagnetic radiation, gravitational radiation or neutrinos are often considered as possible message carriers \citep{hippke_interstellar_2017, wright_how_2018}.
+%
+This excludes messages sent with mechanical means or physical objects \citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown technology that violates the known laws of physics.
+
+As part of this benchmark, we assume that the capacity to emit and receive signals occur at the same time.
 %
@@
 %
+In the ideal setup, this would be equivalent to nodes that send messages isotropically and scan the local skies on all directions with a perfect eficiency.
+%
 Another essential assumption is that all \cetis{} use the same signal power, so that there is a maximum distance out to which it can be detected.
 %
-It is worth mentioning that we are considering in our experimental setup a system composed by several emitters and receivers across the galaxy, under the hypothesis proposed before.
+% It is worth mentioning that we are considering in our experimental setup a system composed by several emitters and 
+% receivers across the galaxy, under the hypothesis proposed before.
 %
@@
 %
-Then, we stress the fact that we are not centered on the probabilities for the Earth itself, but on a generalized \ceti{}.
+Then, we stress the fact that we are not centred on the probabilities for the Earth itself, but on a generalised \ceti{}. 
 %
@@
 It is straightforward to propose and implement a distribution of maximum distances, although this would increase the model complexity at the cost of a larger uncertainty or of another dimension on the hypothesis space.
+
+% For the communication of \cetis{} through messages sent isotropically, we assume that the capacity to emit signals and %to receive signals occur at the same time.
+This means that the ability to find another civilization develops at the same %time that the ability and intention to emit a signal in all directions, intended to be detected by an unknown civilization.
 
@@
 %upper limits.
+
+%-------------------------------------------------------------------
+%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%-------------------------------------------------------------------
 
@@
 %
-The simulation was implemented on a python 3 code, which is publicly
-available\footnote{https://github.com/mlares/simu\_contact}
+%}}}
+
+\subsection{Model implementation}
+%{{{
+
+% hearsay 
+%
+With the aim of making a reproducible project, we developed the tools
+that allow to run and obtain the results shown in this work.
+%
+The simulations were implemented on a Python-3 code, dubbed \hs{},
+which is publicly accessible through the GitHub
+platform\footnote{https://github.com/mlares/simu\_contact} under the
+MIT-license.
+%
+The project is in the process of registration with the ‘’Astrophysics
+Source Code Library’’ \citep[ASCL, ][]{2015JORS....3E..15A,
+2020ASPC..522..731A}
+%
+In order to achieve high performance we made use of vectorised methods
+and array broadcasting properties provided by the NumPy package
+\citep{oliphant_numpy_2006, vanderwalt_numpy_2011}
+%
+The combination of C-compiled Numpy routines for numerical
+computations with a high-level language allows to obtain an easily
+comprehensible and maintainable code.
+%
+From the viewpoint of a user, \hs{} is an object-oriented package that
+exposes the main functionalities as classes and methods.
+%
+In the configuration step, the user prepares the set of simulation
+parameters through an initialisation file.
+%
+Since configuration files and simulation results are persistent, it is
+straightforward to keep track of different experiments and the
+experiments can be revisited easily.
+%
+An in-depth description of the methods can be found in the
+documentation, which is automatically generated from \hs{} docstrings
+and made public in the read-the-docs service.
+
+%Quality assurance
+%
+Regarding the quality assurance of the software, we endorse the PEP-8
+style guide for python (van Rossum et al. 2001) throughout the entire
+project, which improves the code maintainability and helps to minimise
+the so called ''code-entropy'' of future versions.
+%
+The deviations in style were analysed with the flake8 tool, which
+improves the readability and consistency of the code.
+%
+Testing and code coverage have been applied to the package and the
+development process as standard qualitative and quantitative metrics.
+%
+The different parts and functions can be validated to assert if they
+return the expected results on specific test cases and testing
+scenarios \citep{jazayeri_engeneering_2007}.
+%
+On the other hand, code coverage measures the fraction visited and
+executed in the code by the testing procedure
+\citep{miller_testing_1963}.
+%
+The advantage of an exhaustive code-coverage is that it ensures that
+most of the sources of errors have been considered.
+%
+In the \hs{} project we provide test tools for the distinct parts of
+the simulation setup and run, reaching a 99\% of code coverage in the
+version used for this work.
+%
+For every new version of the code, all these quality assurance methods
+and metrics are run automatically by a continuous-integration service.
+%
+It is possible to reproduce the experiments performed in this work
+using the \hs{} package.
+%
+The installation is available through a git repository and also on the
+Python Package Index (PyPI).
+%
+Since the simulation setup is configurable, the time required for a
+simulation to complete depends on the simulation parameters.
 
@@
 
-%\bibliographystyle{mn2e}
+\bibliographystyle{mn2e}
 \setlength{\bibsep}{0.0pt}
-
-\begin{thebibliography}{90}
-\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
-
-\bibitem[{Adamic(2000)}]{adamic_zipf_2000}
-Adamic L., 2000, {Zipf, Power-laws, and Pareto - a ranking tutorial}. web
-
-\bibitem[{Anchordoqui \& Weber(2019)}]{anchordoqui_upper_2019}
-Anchordoqui L.~A., Weber S.~M., 2019, arXiv e-prints
-
-\bibitem[{Annis(1999)}]{annis_astrophysical_1999}
-Annis J., 1999, JBIS - Journal of the British Interplanetary Society, 52, 19
-
-\bibitem[{Armstrong \& Sandberg(2013)}]{Armstrong2013}
-Armstrong S., Sandberg A., 2013, Acta Astronautica, 89, 1
-
-\bibitem[{Balb(2018)}]{balb_spatiotemporal_2018}
-Balb A., 2018, Mem. della Soc. Astron. Ital., 89, 425
-
-\bibitem[{Balbi(2018)}]{balbi_impact_2018}
-Balbi A., 2018, Astrobiology, 18, 54
-
-\bibitem[{Barab{\'{a}}si(2009)}]{barabasi_scale_2009}
-Barab{\'{a}}si A.~L., 2009, Science (80-. )., 325, 412
-
-\bibitem[{Barlow(2013{\natexlab{a}})}]{barlow_galactic_2012}
-Barlow M.~T., 2013{\natexlab{a}}, Int. J. Astrobiol., 12, 63
-
-\bibitem[{Barlow(2013{\natexlab{b}})}]{barlow_galactic_2013}
-Barlow M.~T., 2013{\natexlab{b}}, Int. J. Astrobiol., 12, 63
-
-\bibitem[{Benguigui \& Marinov(2016)}]{benguigui_classificacion_2016}
-Benguigui L., Marinov M., 2016, {A classification of the natural and social
-  distributions Part 2: the explanations}. Tech. rep.
-
-\bibitem[{Bloetscher(2019)}]{bloetscher_using_2019}
-Bloetscher F., 2019, Acta Astronaut., 155, 118
-
-\bibitem[{Borra(2012)}]{borra_searching_2012}
-Borra E.~F., 2012, Astronomical Journal, 144
-
-\bibitem[{Brin(1983)}]{brin_great_1983}
-Brin G., 1983, Quarterly Journal of the Royal Astronomical Society, 24, 283
-
-\bibitem[{Burchell(2006)}]{burchell_whither_2006}
-Burchell M.~J., 2006, Int. J. Astrobiol., 5, 243
-
-\bibitem[{Carroll-Nellenback {et~al}\mbox{.}(2019)Carroll-Nellenback, Frank,
-  Wright, \& Scharf}]{carroll_nellemback_fermi_2019}
-Carroll-Nellenback J., Frank A., Wright J., Scharf C., 2019, The Astronomical
-  Journal, 158, 117
-
-\bibitem[{Chung(2003)}]{chung_simulation_2003}
-Chung C.~A., 2003, {Simulation Modeling Handbook: A Practical Approach
-  (Industrial and Manufacturing Engineering Series)}. CRC press
-
-\bibitem[{{\'{C}}irkovi{\'{c}}(2004)}]{cirkovic_temporal_2004}
-{\'{C}}irkovi{\'{c}} M.~M., 2004, Astrobiology, 4, 225
-
-\bibitem[{{Conway Morris}(2018)}]{conway_three_2018}
-{Conway Morris} S., 2018, Int. J. Astrobiol., 17, 287
-
-\bibitem[{Dayal {et~al}\mbox{.}(2016)Dayal, Ward, \&
-  Cockell}]{dayal_habitability_2016}
-Dayal P., Ward M., Cockell C., 2016, arXiv e-prints
-
-\bibitem[{Do{\v{s}}ovi{\'{c}} {et~al}\mbox{.}(2019)Do{\v{s}}ovi{\'{c}},
-  Vukoti{\'{c}}, \& {\'{C}}irkovi{\'{c}}}]{Dosovic2019}
-Do{\v{s}}ovi{\'{c}} V., Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2019,
-  Astronomy and Astrophysics, 625
-
-\bibitem[{Drake(1962)}]{drake_intelligent_1962}
-Drake F., 1962
-
-\bibitem[{Edmondson \& Stevens(2003)}]{Edmondson2003}
-Edmondson W.~H., Stevens I.~R., 2003, International Journal of Astrobiology, 2,
-  231
-
-\bibitem[{Enriquez {et~al}\mbox{.}(2017)Enriquez, Siemion, Foster, Gajjar,
-  Hellbourg, Hickish, Isaacson, Price, Croft, DeBoer, Lebofsky, MacMahon, \&
-  Werthimer}]{enriquez_breakthrough_2017}
-Enriquez J.~E. {et~al.}, 2017, Astrophys. J., 849, 104
-
-\bibitem[{Fogg(1987)}]{fogg_temporal_1987}
-Fogg M.~J., 1987, Icarus, 69, 370
-
-\bibitem[{Forgan {et~al}\mbox{.}(2016)Forgan, Dayal, Cockell, \&
-  Libeskind}]{forgan_evaluating_2015}
-Forgan D., Dayal P., Cockell C., Libeskind N., 2016, Int. J. Astrobiol., 16, 60
-
-\bibitem[{Forgan(2009)}]{forgan_numerical_2009}
-Forgan D.~H., 2009, Int. J. Astrobiol., 8, 121
-
-\bibitem[{Forgan(2011)}]{forgan_spatiotemporal_2011}
-Forgan D.~H., 2011, Int. J. Astrobiol., 10, 341
-
-\bibitem[{Forgan(2013)}]{forgan_possibility_2013}
-Forgan D.~H., 2013, JBIS - J. Br. Interplanet. Soc., 66, 144
-
-\bibitem[{Forgan(2014)}]{forgan_collimated_2014}
-Forgan D.~H., 2014, JBIS - J. Br. Interplanet. Soc., 67, 232
-
-\bibitem[{Forgan(2017{\natexlab{a}})}]{forgan_galactic_2016}
-Forgan D.~H., 2017{\natexlab{a}}, Int. J. Astrobiol., 16, 349
-
-\bibitem[{Forgan(2017{\natexlab{b}})}]{forgan_galactic_2017}
-Forgan D.~H., 2017{\natexlab{b}}, Int. J. Astrobiol., 16, 349
-
-\bibitem[{Forgan(2019)}]{forgan_exoplanet_2017}
-Forgan D.~H., 2019, Int. J. Astrobiol., 18, 189
-
-\bibitem[{Forgan \& Rice(2010)}]{forgan_numerical_2010}
-Forgan D.~H., Rice K., 2010, Int. J. Astrobiol., 9, 73
-
-\bibitem[{Frank(2009)}]{frank_common_2009}
-Frank S.~A., 2009, {The common patterns of nature}
-
-\bibitem[{Funes {et~al}\mbox{.}(2019)Funes, Florio, Lares, \&
-  Asla}]{funes_searching_2019}
-Funes J.~G., Florio L., Lares M., Asla M., 2019, Theol. Sci., 0, 1
-
-\bibitem[{Galera {et~al}\mbox{.}(2019)Galera, Galanti, \&
-  Kinouchi}]{galera_invasion_2019}
-Galera E., Galanti G.~R., Kinouchi O., 2019, International Journal of
-  Astrobiology, 18, 316
-
-\bibitem[{Glade {et~al}\mbox{.}(2012)Glade, Ballet, \&
-  Bastien}]{glade_stochastic_2011}
-Glade N., Ballet P., Bastien O., 2012, Int. J. Astrobiol., 11, 103
-
-\bibitem[{Gobat \& Hong(2016)}]{gobat_evolution_2016}
-Gobat R., Hong S.~E., 2016, Astron. Astrophys., 592
-
-\bibitem[{Gonzalez(2005)}]{gonzalez_habitable_2005}
-Gonzalez G., 2005, Orig. Life Evol. Biosph., 35, 555
-
-\bibitem[{Gonzalez {et~al}\mbox{.}(2001)Gonzalez, Brownlee, \&
-  Ward}]{gonzalez_galactic_2001}
-Gonzalez G., Brownlee D., Ward P., 2001, Icarus, 152, 185
-
-\bibitem[{Grimaldi(2017)}]{grimaldi_signal_2017}
-Grimaldi C., 2017, Sci. Rep., 7, 46273
-
-\bibitem[{Grimaldi {et~al}\mbox{.}(2018)Grimaldi, Marcy, Tellis, \&
-  Drake}]{Grimaldi2018}
-Grimaldi C., Marcy G.~W., Tellis N.~K., Drake F., 2018, Publications of the
-  Astronomical Society of the Pacific, 130
-
-\bibitem[{Haqq-Misra(2019)}]{haqq-misra_evolution_2019}
-Haqq-Misra J., 2019, arXiv e-prints
-
-\bibitem[{Haqq-Misra \& Kopparapu(2018)}]{haqq-misra_drake_2017}
-Haqq-Misra J., Kopparapu R.~K., 2018, Habitability of the Universe Before
-  Earth, 307
-
-\bibitem[{Harp {et~al}\mbox{.}(2018)Harp, Ackermann, Astorga, Arbunich,
-  Barrios, Hightower, Meitzner, Barott, Nolan, Messerschmitt, Vakoch, Shostak,
-  \& Tarter}]{harp_application_2018}
-Harp G.~R. {et~al.}, 2018, Astrophys. J., 869, 66
-
-\bibitem[{Hart(1975)}]{hart_explanation_1975}
-Hart M.~H., 1975, Quarterly Journal of the Royal Astronomical Society, 16, 128
-
-\bibitem[{Hinkel {et~al}\mbox{.}(2019)Hinkel, Hartnett, Lisse, \&
-  Young}]{hinkel_interdisciplinary_2019}
-Hinkel N., Hartnett H., Lisse C., Young P., 2019, Bull. Am. Astron. Soc., 51,
-  497
-
-\bibitem[{Hippke(2017)}]{hippke_interstellar_2017}
-Hippke M., 2017, arXiv e-prints
-
-\bibitem[{Horvat(2006)}]{horvat_calculating_2007}
-Horvat M., 2006, Int. J. Astrobiol., 5, 143
-
-\bibitem[{Horvat {et~al}\mbox{.}(2011)Horvat, Naki{\'{c}}, \&
-  Oto{\v{c}}an}]{horvat_impact_2011}
-Horvat M., Naki{\'{c}} A., Oto{\v{c}}an I., 2011, Int. J. Astrobiol., 11, 51
-
-\bibitem[{Jeong {et~al}\mbox{.}(2000)Jeong, Tombor, Albert, Oltvai, \&
-  Barabasi}]{jeong_large_2000}
-Jeong H., Tombor B., Albert R., Oltvai Z.~N., Barabasi A.~L., 2000, Nature,
-  407, 651
-
-\bibitem[{Lampton(2013)}]{lampton_information_2013}
-Lampton M., 2013, Int. J. Astrobiol., 12, 312
-
-\bibitem[{Lineweaver {et~al}\mbox{.}(2004)Lineweaver, Fenner, \&
-  Gibson}]{lineweaver_galactic_2004}
-Lineweaver C.~H., Fenner Y., Gibson B.~K., 2004, Science (80-. )., 303, 59
-
-\bibitem[{Lingam \& Loeb(2018)}]{lingam_relative_2019}
-Lingam M., Loeb A., 2018, Astrobiology, 19, 28
-
-\bibitem[{Loeb {et~al}\mbox{.}(2016)Loeb, Batista, \&
-  Sloan}]{loeb_relative_2016}
-Loeb A., Batista R.~A., Sloan D., 2016, J. Cosmol. Astropart. Phys., 2016, 040
-
-\bibitem[{Loeb \& Zaldarriaga(2007)}]{loeb_eavesdropping_2006}
-Loeb A., Zaldarriaga M., 2007, J. Cosmol. Astropart. Phys., 1
-
-\bibitem[{Maccone(2010)}]{maccone_KLT_2010}
-Maccone C., 2010, Acta Astronaut., 67, 1427
-
-\bibitem[{Maccone(2011)}]{maccone_mathematical_2011}
-Maccone C., 2011, Orig. Life Evol. Biosph., 41, 609
-
-\bibitem[{MacCone(2011)}]{maccone_SETI_2011}
-MacCone C., 2011, Acta Astronaut., 68, 63
-
-\bibitem[{MacCone(2013)}]{maccone_SETI_2013}
-MacCone C., 2013, Int. J. Astrobiol., 12, 218
-
-\bibitem[{Maccone(2014{\natexlab{a}})}]{maccone_evolution_2014}
-Maccone C., 2014{\natexlab{a}}, Int. J. Astrobiol., 13, 290
-
-\bibitem[{Maccone(2014{\natexlab{b}})}]{maccone_lognormals_2014}
-Maccone C., 2014{\natexlab{b}}, Acta Astronaut., 105, 538
-
-\bibitem[{Maccone(2015)}]{maccone_statistical_2015}
-Maccone C., 2015, Acta Astronaut., 115, 277
-
-\bibitem[{Mart{\'{i}}n \& Goldenfeld(2006)}]{martin_origin_2006}
-Mart{\'{i}}n H.~G., Goldenfeld N., 2006, Proc. Natl. Acad. Sci. U. S. A., 103,
-  10310
-
-\bibitem[{Mitzenmacher(2004)}]{mitzenmacher_brief_2004}
-Mitzenmacher M., 2004, Internet Math., 1, 226
-
-\bibitem[{Morrison \& Gowanlock(2015)}]{morrison_extending_2015}
-Morrison I.~S., Gowanlock M.~G., 2015, Astrobiology, 15, 683
-
-\bibitem[{Murante {et~al}\mbox{.}(2015)Murante, Monaco, Borgani, Tornatore,
-  Dolag, \& Goz}]{murante_simulating_2015}
-Murante G., Monaco P., Borgani S., Tornatore L., Dolag K., Goz D., 2015, Mon.
-  Not. R. Astron. Soc., 447, 178
-
-\bibitem[{Newman(2005)}]{newman_power_2005}
-Newman M., 2005, Contemp. Phys., 46, 323
-
-\bibitem[{Newman \& Sagan(1981)}]{newman_galactic_1981}
-Newman W.~I., Sagan C., 1981, Icarus, 46, 293
-
-\bibitem[{Peters(2018)}]{peters_outer_2018}
-Peters T., 2018, Int. J. Astrobiol., 17, 282
-
-\bibitem[{Prantzos(2013)}]{prantzos_joint_2013}
-Prantzos N., 2013, Int. J. Astrobiol., 12, 246
-
-\bibitem[{Ptolemaeus(2014)}]{ptolemaeus_system_2014}
-Ptolemaeus C., 2014, {System Design, Modeling, and Simulation. Using Ptolemy
-  II}. Ptolemy.org, Berkeley, p. 690
-
-\bibitem[{Rahvar(2017)}]{rahvar_cosmic_2016}
-Rahvar S., 2017, Mon. Not. R. Astron. Soc., 470, 3095
-
-\bibitem[{Ramirez {et~al}\mbox{.}(2017)Ramirez, G{\'{o}}mez-Mu{\~{n}}oz,
-  V{\'{a}}zquez, \& N{\'{u}}{\~{n}}ez}]{ramirez_new_2017}
-Ramirez R., G{\'{o}}mez-Mu{\~{n}}oz M.~A., V{\'{a}}zquez R., N{\'{u}}{\~{n}}ez
-  P.~G., 2017, International Journal of Astrobiology, 1
-
-\bibitem[{Ross(2012)}]{ross_simulation_2012}
-Ross S.~M., 2012, {Simulation}. Elsevier Science Publishing Co Inc
-
-\bibitem[{Simkin \& Roychowdhury(2006)}]{simkin_theory_2006}
-Simkin M.~V., Roychowdhury V.~P., 2006, J. Math. Sociol., 30, 33
-
-\bibitem[{Smith(2009)}]{smith_broadcasting_2009}
-Smith R.~D., 2009, Int. J. Astrobiol., 8, 101
-
-\bibitem[{Solomonides {et~al}\mbox{.}(2016)Solomonides, Kaltenegger, \&
-  Terzian}]{solomonides_probabilistic_2016}
-Solomonides E., Kaltenegger L., Terzian Y., 2016, 228th AAS, San Diego, 228, 1
-
-\bibitem[{Sornette(2006)}]{sornette_critical_2006}
-Sornette D., 2006, {Critical phenomena in natural sciences: Chaos, fractals,
-  selforganization and disorder: Concepts and tools}. Springer
-
-\bibitem[{Sotos(2019)}]{Sotos_biotechnology_2019}
-Sotos J.~G., 2019, {Biotechnology and the lifetime of technical civilizations}
-
-\bibitem[{Starling \& Forgan(2014)}]{starling_virulence_2013}
-Starling J., Forgan D.~H., 2014, Int. J. Astrobiol., 13, 45
-
-\bibitem[{Tarter(2001)}]{tarter_search_2001}
-Tarter J., 2001, Annu. Rev. Astron. Astrophys., 39, 511
-
-\bibitem[{Tarter {et~al}\mbox{.}(2009)Tarter, Backus, Blair, Cordes, Harp,
-  Henry, Horowitz, Howard, Kilsdonk, Korpela, Lasio, Levin, Shostack, \&
-  Werthimer}]{tarter_advancing_2009}
-Tarter J. {et~al.}, 2009, Astro2010 Astron. Astrophys. Decad. Surv. Sci. White
-  Pap. no. 294, 2010
-
-\bibitem[{Vakoch \& Dowd(2015)}]{vakoch_drake_2015}
-Vakoch D.~A., Dowd M.~F., 2015, {The drake equation: Estimating the prevalence
-  of extraterrestrial life through the ages}, Vakoch D.~A., Dowd M.~F., eds.
-  Cambridge University Press, Cambridge, pp. 1--319
-
-\bibitem[{VanHouten(2018)}]{vanhouten_isthere_2017}
-VanHouten M.~A., 2018, Lancet Gastroenterol. Hepatol., 3, 382
-
-\bibitem[{Vukoti{\'{c}} \&
-  {\'{C}}irkovi{\'{c}}(2012)}]{vukotic_astrobiological_2012}
-Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2012, Orig. Life Evol. Biosph.,
-  42, 347
-
-\bibitem[{Vukoti{\'{c}} {et~al}\mbox{.}(2016)Vukoti{\'{c}}, Steinhauser,
-  Martinez-Aviles, {\'{C}}irkovi{\'{c}}, Micic, \&
-  Schindler}]{vukotic_grandeur_2016}
-Vukoti{\'{c}} B., Steinhauser D., Martinez-Aviles G., {\'{C}}irkovi{\'{c}}
-  M.~M., Micic M., Schindler S., 2016, Mon. Not. R. Astron. Soc., 459, 3512
-
-\bibitem[{Walters {et~al}\mbox{.}(1980)Walters, Hoover, \&
-  Kotra}]{walters_interstellar_1980}
-Walters C., Hoover R.~A., Kotra R., 1980, Icarus, 41, 193
-
-\bibitem[{Wright {et~al}\mbox{.}(2015)Wright, Cartier, Zhao, Jontof-Hutter, \&
-  Ford}]{wright_theGsearch_2015}
-Wright J.~T., Cartier K. M.~S., Zhao M., Jontof-Hutter D., Ford E.~B., 2015,
-  Astrophys. J., 816, 17
-
-\bibitem[{Wright {et~al}\mbox{.}(2018)Wright, Kanodia, \&
-  Lubar}]{wright_how_2018}
-Wright J.~T., Kanodia S., Lubar E., 2018, The Astronomical Journal, 156, 260
-
-\end{thebibliography}
+\bibliography{biblio_seti, biblio_software}
 

### 2020-05-13 submit_v03/manuscript.tex
pro, writing

@@
 
-%-------------------------------------------------------------------
-%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%-------------------------------------------------------------------
-
-In our simulations we also assume the simplest situations for the
-growth of the sphere of first contact.
-%
-For example, we discard the possibility of stellar colonization
+In our simulations, we assume the simplest configuration for the
+growth of the sphere of first contact. For example, we discard the
+possibility of stellar colonisation
 \citep[e.g.][]{newman_galactic_1981, walters_interstellar_1980,
 starling_virulence_2013, barlow_galactic_2012, jeong_large_2000,
-maccone_mathematical_2011} and assume that the communication aim is
-performed to all directions with the same power and the same
-probability, i.e., strongly assume isotropic communication in all
-cases.
-%
-More detailed simulations could be produced considering different
-efficiency of communication or detection methods.
-%
-However, in this case we also would complicate the experimental setup,
-make the results less evident.
-%
-The results are independent of the nature of life (organic or
-artificial).
-%
-The lifetime of a civilization can be caused by auto destruction or by
-external factors.
+maccone_mathematical_2011}. We also assume that communication is
+equally likely in all directions, i.e., we assume isotropic
+communication in all cases. Different communication efficiencies or
+detection methods are straightforward to carry out in the simulations
+for more detailed and complex scenarios. This approach, however, is
+beyond this work because it obscures the experimental setup and make
+the results less clear. The assumptions we accept imply that the
+results are independent of whether intelligent agents are organic or
+artificial. Moreover, the causes of the limited lifetime of a
+civilization can be natural (astrophysical phenomena), caused by auto
+destruction or by external factors, to name a few. However, we assume
+that these events are sufficiently numerous in the Galaxy so that a
+statistical model is plausible. The failure of this hypothesis would
+imply extreme values for the parameters that represent the density of
+nodes.
+
 %}}}
@@
 
-The temporal structure of the process is defined by two distribution
-parameters that represent the mean time interval that a \ceti{}
-can emit and receive signals, and the mean time interval
-between the emergence of consecutive \cetis{}.
-%
-The spatial structure of the simulation is given by the size and shape
-of the Galactic Habitable Zone and the maximum distance a signal can
-travel to be detected (D$_{max}$).
-%
-The parameters for the temporal distributions also determine the
-spatial properties, since the density of active \cetis{} in the galaxy
-depend on these two parameters.
-%
-For example, a small $\tau_a$ and a large $\tau_s$ will produce a
-galaxy densely populated with \cetis{}.
-%
-Also, some hypotheses regarding the shapes of the distributions of the
-temporal parameters must be made in order to complete the
-simulation.
-%
-\citet{forgan_spatiotemporal_2011} argues that the times at which
-different civilizations become intelligent follow a Gaussian
-distribution, and then the distribution of inter-arrival times is an
-inverse exponential.
-%
-We choose to assume that the distribution of the times of A events is a
-stationary Poisson process, and then the distribution of the times
-between the appearance of new consecutive \cetis{} is exponential.
-%
-Regarding the duration of a \ceti{}, we propose that the distribution
-of the duration of \cetis{} is a stationary exponential distribution.
-%
-\citet{balbi_impact_2018} stress the fact that it would be desirable
-to arrive at a theoretical statistical distribution of the lifespan of
-civilizations, preferentially on the basis of the underlying
-astrophysical and biological processes.
-%
-\citet{maccone_lognormals_2014} argues that this distribution should
-be a log normal.
-
-The power law and exponential statistical distributions are among the
-most common patterns found in natural phenomena.
-%
-For example, the distribution of the frequency of words in many
-languages is known to follow a Zipf's law (which is a power law).
-%
-This distributions arise any time a phenomena is characterized by
-commonly occurring small events and rarely produced large events
-\citep[e.g. ][]{adamic_zipf_2000}.
-%
-Zipf law also describes population ranks of cities in various
-countries, corporation sizes, income rankings, ranks of number of
-people watching the same TV channel.
-%
-The magnitudes of earthquakes, hurricanes, volcanic eruptions and
-floods; the sizes of meteorites or the losses caused by business
-interruptions from accidents, are also well described by power laws
-\citep{sornette_critical_2006, benguigui_classificacion_2016}.
-%
-The power law behavior has been observed in a variety of systems,
-including for example stock market fluctuations, sizes of computer
-files and word frequency in languages \citep{mitzenmacher_brief_2004,
-newman_power_2005, simkin_theory_2006}
-%
-Power laws have also been widely used in biological sciences, e.g., in
-analyzing connectivity patterns in metabolic networks
-\citep{jeong_large_2000} and in the number of species observed per
-unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
-% 
-More examples can be found in the literature
-\citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009,
-maccone_evolution_2014, maccone_lognormals_2014,
-benguigui_classificacion_2016}
-
-In this work, the exponential distribution of lifespan and waiting
-times is justified by considering the hypothesis that the process of
-appearance of life in the galaxy is homogeneous and stationary.
-%
-That is, there is no preferred location within the GHZ for the
-spontaneous appearance of life, and the emergence of a \ceti{} is
-independent of the existence of previous nodes in the galaxy.
-%            
-This is equivalent to proposing a Poisson process for the emergence of
-\cetis{}, since there is a close relation between the number of events
-in time or space and the waiting time or separation, respectively
+The temporal structure of the process is defined by two distribution parameters.
+One of them represent the mean time interval that a \ceti{} can emit and receive signals, and the other the mean time interval between the emergence of consecutive \cetis{}.
+The spatial structure of the simulation is given by the size and shape of the Galactic Habitable Zone and the maximum distance a signal can travel to be detected (D$_{max}$).
+The parameters for the temporal distributions also determine the spatial properties, since the density of active \cetis{} in the galaxy depend on these two parameters.
+For example, a small $\tau_a$ and a large $\tau_s$ will produce a galaxy densely populated with \cetis{}.
+Also, some hypotheses regarding the shapes of the distributions of the temporal parameters must be made in order to complete the simulation.
+\citet{forgan_spatiotemporal_2011} argues that the times at which different civilisations become intelligent follow a Gaussian distribution, and then the distribution of inter-arrival times is an inverse exponential.
+We assume that the distribution of the times of A events is a stationary Poisson process, and then the distribution of the times between the appearance of new consecutive \cetis{} is exponential.
+Regarding the duration of a \ceti{}, we propose that the distribution of the duration of \cetis{} is a stationary exponential distribution.
+There are no clear arguments to conclude a statistical law for the later distribution, so that we propose it as a working hipothesis.  A theoretical statistical distribution of the lifespan of civilisations would rely preferentially on the basis of the underlying astrophysical and biological processes \citep{balbi_impact_2018}.
+For example, \citet{maccone_lognormals_2014} argues that this distribution should be a log normal.
+
+The power law and exponential statistical distributions are among the most common patterns found in natural phenomena.
+For example, the distribution of the frequency of words in many languages is known to follow a Zipf’s law (which is a  power law).
+These distributions arise any time a phenomenon is characterised by commonly occurring small events and rarely produced large events \citep[e.g.
+][]{adamic_zipf_2000}.
+Zipf law also describes population ranks of cities in various countries, corporation sizes, income rankings, ranks of the number of people watching the same TV channel, etc.
+The magnitudes of earthquakes, hurricanes, volcanic eruptions and floods; the sizes of meteorites or the losses caused by business interruptions from accidents, are also well described by power laws \citep{sornette_critical_2006,  benguigui_classificacion_2016}.
+The power law behaviour has been observed in a variety of systems.  These include, for example, stock market fluctuations, sizes of computer files or word frequency in languages \citep{mitzenmacher_brief_2004, newman_power_2005, simkin_theory_2006}.
+Power laws have also been widely used in biological sciences.  Some examples are the analysis of connectivity patterns in metabolic networks \citep{jeong_large_2000} or the number of species observed per unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
+More examples can be found in the literature \citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009, maccone_evolution_2014,  maccone_lognormals_2014, benguigui_classificacion_2016}.
+
+In this work, the exponential distribution of lifespan and waiting times is justified by considering the hypothesis that the process of appearance of life in the galaxy is homogeneous and stationary.
+That is, there is not a preferred region within the GHZ for the spontaneous appearance of life, and the emergence of a \ceti{} is independent of the existence of previous nodes in the galaxy.
+This is equivalent to proposing a Poisson process for the emergence of \cetis{}, given the relation between the number of events in time or space and the waiting time or separation, respectively
 \citep[e.g., ][]{ross_simulation_2012}.
-%
-That is, these are two alternative approaches to describing the same
-process, a Poisson distribution for the number of events implies an
-exponential distribution for their separations, and vice versa.
-%
-It should be emphasized that the exponential laws used in this work
-are assumed as part of the working hypothesis, and instead of
-analyzing results from a particular parameter chosen ad--hoc, we
-explore the hypothesis space and analyze the impact of the values of
-these parameters on the results.
+That is, these are two alternative approaches to describing the same process, a Poisson distribution for the number of events implies an exponential distribution for their separations, and vice versa.
+It should be emphasised that the exponential laws used in this work are assumed as part of the working hypothesis, and instead of analysing results from a particular parameter chosen ad hoc, we explore the hypothesis space and analyse the impact of the values of these parameters on the results.
+
+%-------------------------------------------------------------------
+%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%-------------------------------------------------------------------    
+\rule{8cm}{12pt} \bigskip
+
 %}}}

### 2020-05-18 submit_v03/manuscript.tex
m

@@
 
-The temporal structure of the process is defined by two distribution parameters.
-One of them represent the mean time interval that a \ceti{} can emit and receive signals, and the other the mean time interval between the emergence of consecutive \cetis{}.
-The spatial structure of the simulation is given by the size and shape of the Galactic Habitable Zone and the maximum distance a signal can travel to be detected (D$_{max}$).
-The parameters for the temporal distributions also determine the spatial properties, since the density of active \cetis{} in the galaxy depend on these two parameters.
-For example, a small $\tau_a$ and a large $\tau_s$ will produce a galaxy densely populated with \cetis{}.
-Also, some hypotheses regarding the shapes of the distributions of the temporal parameters must be made in order to complete the simulation.
-\citet{forgan_spatiotemporal_2011} argues that the times at which different civilisations become intelligent follow a Gaussian distribution, and then the distribution of inter-arrival times is an inverse exponential.
-We assume that the distribution of the times of A events is a stationary Poisson process, and then the distribution of the times between the appearance of new consecutive \cetis{} is exponential.
-Regarding the duration of a \ceti{}, we propose that the distribution of the duration of \cetis{} is a stationary exponential distribution.
-There are no clear arguments to conclude a statistical law for the later distribution, so that we propose it as a working hipothesis.  A theoretical statistical distribution of the lifespan of civilisations would rely preferentially on the basis of the underlying astrophysical and biological processes \citep{balbi_impact_2018}.
-For example, \citet{maccone_lognormals_2014} argues that this distribution should be a log normal.
+The temporal structure of the process is defined by two distribution parameters. One of them represent the mean time
+interval that a \ceti{} can emit and receive signals, and the other the mean time interval between the emergence of
+consecutive \cetis{}. The spatial structure of the simulation is given by the size and shape of the Galactic Habitable
+Zone and the maximum distance a signal can travel to be detected (D$_{max}$). The parameters for the temporal
+distributions also determine the spatial properties, since the density of active \cetis{} in the galaxy depend on these
+two parameters. For example, a small $\tau_a$ and a large $\tau_s$ will produce a galaxy densely populated with
+\cetis{}. Also, some hypotheses regarding the shapes of the distributions of the temporal parameters must be made in
+order to complete the simulation. \citet{forgan_spatiotemporal_2011} argues that the times at which different
+civilisations become intelligent follow a Gaussian distribution, and then the distribution of inter-arrival times is an
+inverse exponential. We assume that the distribution of the times of A events is a stationary Poisson process, and then
+the distribution of the times between the appearance of new consecutive \cetis{} is exponential. Regarding the duration
+of a \ceti{}, we propose that the distribution of the duration of \cetis{} is a stationary exponential distribution.
+There are no clear arguments to conclude a statistical law for the later distribution, so that we propose it as a
+working hipothesis.  A theoretical statistical distribution of the lifespan of civilisations would rely preferentially
+on the basis of the underlying astrophysical and biological processes \citep{balbi_impact_2018}. For example,
+\citet{maccone_lognormals_2014} argues that this distribution should be a log normal.
 
 The power law and exponential statistical distributions are among the most common patterns found in natural phenomena.
-For example, the distribution of the frequency of words in many languages is known to follow a Zipf’s law (which is a  power law).
-These distributions arise any time a phenomenon is characterised by commonly occurring small events and rarely produced large events \citep[e.g.
-][]{adamic_zipf_2000}.
-Zipf law also describes population ranks of cities in various countries, corporation sizes, income rankings, ranks of the number of people watching the same TV channel, etc.
-The magnitudes of earthquakes, hurricanes, volcanic eruptions and floods; the sizes of meteorites or the losses caused by business interruptions from accidents, are also well described by power laws \citep{sornette_critical_2006,  benguigui_classificacion_2016}.
-The power law behaviour has been observed in a variety of systems.  These include, for example, stock market fluctuations, sizes of computer files or word frequency in languages \citep{mitzenmacher_brief_2004, newman_power_2005, simkin_theory_2006}.
-Power laws have also been widely used in biological sciences.  Some examples are the analysis of connectivity patterns in metabolic networks \citep{jeong_large_2000} or the number of species observed per unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
-More examples can be found in the literature \citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009, maccone_evolution_2014,  maccone_lognormals_2014, benguigui_classificacion_2016}.
+%
+For example, the distribution of the frequency of words in many languages is known to follow a Zipf’s law (which is a 
+power law).
+%
+These distributions arise any time a phenomenon is characterised by commonly occurring small events and rarely produced
+large events \citep[e.g. ][]{adamic_zipf_2000}.
+%
+Zipf law also describes population ranks of cities in various countries, corporation sizes, income rankings, ranks of
+the number of people watching the same TV channel, etc.
+%
+The magnitudes of earthquakes, hurricanes, volcanic eruptions and floods; the sizes of meteorites or the losses caused
+by business interruptions from accidents, are also well described by power laws
+\citep{sornette_critical_2006}.
+%
+%% , benguigui_classificacion_2016}.
+The power law behaviour has been observed in a variety of systems.
+%
+These include, for example, stock market fluctuations, sizes of computer files or word frequency in languages
+\citep{mitzenmacher_brief_2004, newman_power_2005, simkin_theory_2006}. 
+%
+Power laws have also been widely used in
+biological sciences.
+%
+Some examples are the analysis of connectivity patterns in metabolic networks \citep{jeong_large_2000} or the number of
+species observed per unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
+%
+More examples can be found in the literature 
+\citep{martin_origin_2006, maccone_KLT_2010}
+\citep{barabasi_scale_2009, maccone_evolution_2014}
+\citep{maccone_lognormals_2014}
 
@@
 
-%-------------------------------------------------------------------
-%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%-------------------------------------------------------------------    
-\rule{8cm}{12pt} \bigskip
-
 %}}}
@@
 
-In this Section we discuss the degree of complexity in the model,
-considering a compromise between the accuracy of the model and the
-number of parameters that are free or with a high uncertainty.
-%
-Firstly, we emphasize that the odds of a causal contact between two
-\cetis{} should not be considered as the odds of a contact between two
-intelligent civilizations, and in fact the later could be much lesser
-than the former.
-%
-Indeed, in order to establish a contact between any two entities, a
-minimum degree of compatibility must be accomplished without any
-previous agreement, making the possibility of a contact with a message
-that could be deciphered highly rare \citep[see e.g. ][]{forgan_collimated_2014}.
-%
-Besides the trade--off between the simplicity and the complexity of
-the experiments, further analysis could be performed following this
-framework in order to explore possible implications of the results for
-more detailed configurations.
-%
-For example, the communication method (isotropic, collimated,
-serendipitous) can affect the observables, making it necessary to
-implement a correction factor.
-%
-Taking this into account, our results regarding the probability of
-causal contact should be considered as upper limits for effective
-contacts, since they depend on the efficiency of both the emitter and
-the receiver to broadcast and scan the sky for intelligent signals,
-respectively \citep{grimaldi_signal_2017}.
-%
-A correction by coverage ratio in the detection and by a targeting
-ratio in the emission could be easily implemented in the simulation,
-although the effect of reducing the probability of contact is
-basically the product of the efficiency ratios and thus such
-implementation is not necessary.
-%
-Therefore, the values of the probabilities could be modified by a
-constant correction factor equal to the combined emission/reception
-efficiency \citep{smith_broadcasting_2009, anchordoqui_upper_2019,
-forgan_collimated_2014} or for beam--like transmissions
-\citep{grimaldi_signal_2017}.
-%
-Other considerations include the effects of alignments, the use of
-stars as sources or amplifiers \citep{Edmondson2003, borra_searching_2012}, 
-the nature of the message carrier, the
-use of a spatial distribution that actually resembles the spiral shape
-and the width of the disc of the Galaxy, different distribution
-functions for the mean lifetime of \cetis{}, and different
-efficiencies for the sphere of the causal region of each \ceti{}
-resembling different searching strategies
-\citep{hippke_interstellar_2017}.
-%
-It is also possible to consider that D$_{max}$ is different for
-different \cetis{}.  For example a power law where powerful emissions
-are rare and weak emissions are common could be an improvement to the
+In this Section we discuss the degree of complexity in the model, considering a compromise between the accuracy of the
+model and the number of parameters that are free or with a high uncertainty.
+%
+Firstly, we emphasise that the odds of a causal contact between two \cetis{} should not be considered as the odds of a
+contact between two intelligent civilisations, and in fact the latter could be much lesser than the former.
+%
+Indeed, in order to establish a contact between any two entities, a minimum degree of compatibility must be accomplished
+without any previous agreement, making the possibility of a contact with a message that could be deciphered highly rare
+\citep[see e.g. ][]{forgan_collimated_2014}.
+%
+Besides the trade--off between the simplicity and the complexity of the experiments, further analysis could be performed
+following this framework in order to explore possible implications of the results for more detailed configurations.
+%
+For example, the communication method (isotropic, collimated, serendipitous) can affect the observables, making it
+necessary to implement a correction factor.
+%
+Taking this into account, our results regarding the probability of causal contact should be considered as upper limits
+for effective contacts, since they depend on the efficiency of both the emitter and the receiver to broadcast and scan
+the sky for intelligent signals, respectively \citep{grimaldi_signal_2017}.
+%
+A correction by coverage ratio in the detection and by a targeting ratio in the emission could be easily implemented in
+the simulation, although the effect of reducing the probability of contact is basically the product of the efficiency
+ratios and thus such implementation is not necessary.
+%
+Therefore, the values of the probabilities could be modified by a constant correction factor equal to the combined
+emission/reception efficiency \citep{smith_broadcasting_2009, anchordoqui_upper_2019, forgan_collimated_2014} or for
+beam-like transmissions \citep{grimaldi_signal_2017}. 
+%
+Other considerations include the effects of alignments, the use of stars as sources or amplifiers \citep{Edmondson2003,
+borra_searching_2012}, the nature of the message carrier.
+%
+Another improvement could be the use of a spatial distribution that resembles the spiral shape and the width of the disc
+of the Galaxy.
+%
+Regarding the distributions of the parameters, different distribution functions for the mean lifetime of \cetis{} can be
+implemented as alternatives.
+%
+Finally, regarding the degree of idealisation, the model admits different efficiencies for the sphere of the causal
+region of each \ceti{}, featuring different searching strategies \citep{hippke_interstellar_2017}. 
+%
+It is also possible to consider that $D_{max}$ is different for different \cetis{}.
+%
+For example, a power law where powerful emissions are rare and low emissions are common could be an improvement to the
 model.
 %
-In this work we chose not to implement this for the sake of
-simplicity.
-%
-Finally, the role of the message contents could influence on the
-lifespan of a \ceti{} that receives a message, although the
-implementation of such behavior would increase the number of free
-parameters and would be highly speculative.
+In this work we chose not to implement this for the sake of simplicity.
+%
+Finally, the role of the message contents could influence on the lifespan of a \ceti{} that receives a message, although
+the implementation of such behaviour would increase the number of free parameters and would be more speculative.
 
@@
 
-Given a parametric model $M(\tau_a, \tau_s, D_{max})$, which includes
-the functional forms of the statistical distributions and fixed values
-for the hypotheses, a discrete event simulation is performed by
-keeping track of a set of variables that change each time a meaningful
-event is produced.
-%
-The main variables that follow the evolution of the simulation are the
-positions of stars, which are sampled randomly within the GHZ, the
-time of the awakening of each \ceti{} (A event), and the time of the
-vanishing of each \ceti{} (D event).
-%
-The variables that can be deduced from the previous ones include the
-number of \cetis{} in casual contact with at least another \ceti{} at a given
-time, the number of \cetis{} as a function of time, the number of
-\cetis{}
-that receive a message at least one time, the number of \cetis{} that
-receive a message at least one time and successfully deliver an answer
-and the number distribution of waiting time to receive a message.
-%
-All these quantities are updated each time one of the four events (A,
-B, C, D) is produced.
+A discrete event simulation is performed for a given model $M(\tau_a, \tau_s, D_{max})$ by keeping track of a set of variables that change each time an event happens. 
+%
+In addition, the model comprises elements that are fixed for all simulations, for example, the functional forms of the statistical distributions and the adopted values of particular variables (see Table XXX).
+%
+The main variables that follow the evolution of the simulation are: the positions of stars, which are sampled randomly within the GHZ; the time of the awakening of each \ceti{} (A event); and the time of the vanishing of each \ceti{} (D event).
+%
+The variables that can be deduced from the previous ones include: the number of \cetis{} in casual contact with at least another \ceti{} at a given time; the number of \cetis{} as a function of time; the number of \cetis{} that receive a message at least one time; the number of \cetis{} that receive a message at least one time and successfully deliver an answer; and the number distribution of waiting time to receive a message.
+%
+All these quantities are updated each time one of the four events (A, B, C, D) occurs.
 
@@
 
-Since the method is based on events, the first step is to define an
-architecture of events, the relationships between them and how events
-trigger changes on the state of the system,
-so that the system is updated each time an event is produced.
-%
-We consider a simulated system that represents the galactic habitable
-zone which, on a first approach, is a 2-dimensional annular region.
-%
-The adopted values for the geometry of the GHZ are 20 kly and 60 kly, for the inner
-and outer radius, respectively \citep{lineweaver_galactic_2004}.
+Since the method is based on events, the first step is to define the relationships between them and how events trigger changes on the state of the system.
+%
+Once an event is produced, the system is updated by saving the corresponding change in the related variables.
+%
+For example, a B-type event diminishes the number of active \cetis{} by one, and also could affect the number of ongoing communications.
+%
+We consider a first approach to the galactic habitable zone as a 2-dimensional annular region.
+%
+The adopted values for the geometry of the GHZ are 2 kpc and 8 kpc, for the inner and outer radius, respectively \citep{lineweaver_galactic_2004}. 
 % ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
@@
 %
-This simple model does not take into account the variations in stellar
-density given by the spiral structure.
-%
-Although there are several possible approaches, we chose to follow the
-evolution of the system according to the following events:
+This simple model does not take into account the variations in stellar density given by the spiral structure of the Galaxy.
+%
+The main task in setting a DE simulation is to define the events and what changes do these events produce on the system.
+%
+Although there are several possible approaches, we chose to follow the evolution of the system according to the following events:
 
@@
 We have chosen this nomenclature so that it can be easily remembered
-from the initial A, B, C, D.
+from the initials A, B, C, D.
 %
@@
 
-The Galactic Habitable Zone is set at fixed values of the inner and outer
-radii, radial symmetry is assumed, and the with of the galactic disk
-is neglected compared to the radial size.
-%
-The two radii are measured in light years, with the aim to maintain a
-single and comprehensive unit for both space and time coordinates. 
+%-------------------------------------------------------------------
+%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%-------------------------------------------------------------------    
+\rule{8cm}{12pt} \bigskip
+
+% The Galactic Habitable Zone is set at fixed values of the inner and outer
+% radii, radial symmetry is assumed, and the with of the galactic disk
+% is neglected compared to the radial size.
+% %
+% The two radii are measured in light years, with the aim to maintain a
+% single and comprehensive unit for both space and time coordinates. 
 %
@@
 simulation to complete depends on the simulation parameters.
+%
+It also depends on the hardware that is used to make the run, and on whether the parallel option is set.
+%
+However, it is simple to make a local run of limited versions of the experiments to have a sense of the time required by
+the code to complete the simulations.
+%
+More information on this can be found on the documentation of the software, available at
+\url{https://hearsay.readthedocs.io/en/latest/}.
 

### 2020-05-18 submit_v03/manuscript.tex
fin primera iteracion

@@
 %
-This is done in terms of three independent parameters, namely, the mean time span between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$, their mean lifetime $\left<\tau_s\right>$, and the maximum distance a signal can be detected by another \ceti{} ($D_{max}$).
-%
-Intuitively, the probability of the existence of causal contacts between pairs of \cetis{} would be larger for smaller $\tau_a$ parameters, higher $\tau_s$ or higher $D_{max}$ parameter values.
+This is done in terms of three independent parameters, namely, the
+mean time span between the appearance of consecutive \cetis{},
+$\left<\tau_a\right>$, their mean lifetime $\left<\tau_s\right>$, and
+the maximum distance a signal can be detected by another \ceti{}
+($D_{\rm max}$).
+%
+Intuitively, the probability of the existence of causal contacts
+between pairs of \cetis{} would be larger for smaller $\tau_a$
+parameters, higher $\tau_s$ or higher $D_{\rm max}$ parameter values.
 %
@@
 The sphere of first contact (SFC) is a sphere centered on the emitter
-that grows until its radius reaches the D$_{max}$ distance, in which
+that grows until its radius reaches the D$_{\rm max}$ distance, in which
 the power of a signal would equal the detectability threshold.
@@
 %
-This sphere grows until its radius reaches the D$_{max}$ distance, in
+This sphere grows until its radius reaches the D$_{\rm max}$ distance, in
 which it would be no longer detectable due to the decrease in the
@@
 The polygon is limited in the spatial axis by the maximum distance of
-the signal $D_{max}$ and in the time axis by the time span between the
+the signal $D_{\rm max}$ and in the time axis by the time span between the
 A-type event (concave vertex) and the D-type event (convex vertex).
@@
 %
-The leading front, or surface of first contact (SFC) grows from the central \ceti{} until it reaches the maximum distance $D_{max}$.
+The leading front, or surface of first contact (SFC) grows from the
+central \ceti{} until it reaches the maximum distance $D_{\rm max}$.
 %
@@
 %
-This region will grow if the surface of first contact has not yet reached the maximum distance $D_{max}$, and will shrink after a D-type event until the surface of last contact reaches $D_{max}$, producing as a result the loss of all signal from the central \ceti{}.
+This region will grow if the surface of first contact has not yet
+reached the maximum distance $D_{\rm max}$, and will shrink after a
+D-type event until the surface of last contact reaches $D_{\rm max}$, producing as a result the loss of all signal from the central \ceti{}.
 %
@@
    $\tau_{a}$ & Mean temporal separation between consecutive
-   awakenings & 4000 yr & 200\,000 yr & 50\\ 
+   awakenings & 10$^2$~yr & 10$^6$~yr & 50\\ 
    $\tau_{s}$ & Mean lifetime of a \ceti{}
-   & 10000 yr & 500\,000 yr& 50\\ 
-   D$_{max}$ & Maximum reach of a message  &  \multicolumn{2}{l}{(500,
-   1000, 10000, 20000, 40000, 80000) lyr} & 6  \\
+   & 10$^2$yr & 10$^6$~yr& 50\\ 
+   D$_{\rm max}$ & Maximum reach of a message  & 30~pc & 10~kpc & 50 \\
 \hline
@@
    & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
-   R$_{GHZ}^{min}$   & Inner radius of the GHZ     & & &20\,000 lyr\\
-   R$_{GHZ}^{max}$   & Outer radius of the GHZ     &&&60\,000 lyr\\
-   t$_{max}$ & Time span of the simulation  &&&1.e6 yr\\
+   R$_{GHZ}^{\rm min}$   & Inner radius of the GHZ     & & & 7~kpc\\
+   R$_{GHZ}^{\rm max}$   & Outer radius of the GHZ     &&& 9~kpc\\
+   t$_{\rm max}$ & Time span of the simulation  &&&1.e6 yr\\
     & \multicolumn{3}{l}{number of random realizations for each point in the hypothesis space} &50\\
@@
 consecutive \cetis{}. The spatial structure of the simulation is given by the size and shape of the Galactic Habitable
-Zone and the maximum distance a signal can travel to be detected (D$_{max}$). The parameters for the temporal
+Zone and the maximum distance a signal can travel to be detected
+(D$_{\rm max}$). The parameters for the temporal
 distributions also determine the spatial properties, since the density of active \cetis{} in the galaxy depend on these
@@
 for \cetis in four different samples (M5, M6, M7, M8) with two
-values for the range of the signal reach: D$_{max}$=40000 ly (blue),  and 
-D$_{max}$=80000 ly (orange).
+values for the range of the signal reach: D$_{\rm max}$=40000 ly (blue),  and 
+D$_{\rm max}$=80000 ly (orange).
 %
@@
 %
-It is also possible to consider that $D_{max}$ is different for different \cetis{}.
+It is also possible to consider that $D_{\rm max}$ is different for different \cetis{}.
 %
@@
 
-A discrete event simulation is performed for a given model $M(\tau_a, \tau_s, D_{max})$ by keeping track of a set of variables that change each time an event happens. 
+A discrete event simulation is performed for a given model $M(\tau_a,
+\tau_s, D_{\rm max})$ by keeping track of a set of variables that change each time an event happens. 
 %
@@
 function of $\tau_a$ and $\tau_s$, for 
-$D_{max}$=10000 (left panel),
-$D_{max}$=40000 (middle panel), and
-$D_{max}$=80000 (right panel)
+$D_{\rm max}$=10000 (left panel),
+$D_{\rm max}$=40000 (middle panel), and
+$D_{\rm max}$=80000 (right panel)
 %
@@
 awakening, as a funcion of $\tau_a$ and $\tau_s$, for
-$D_{max}$=10000 (left panel),
-$D_{max}$=40000 (middle panel), and
-$D_{max}$=80000 (right panel)
+$D_{\rm max}$=10000 (left panel),
+$D_{\rm max}$=40000 (middle panel), and
+$D_{\rm max}$=80000 (right panel)
 %
@@
 
-%-------------------------------------------------------------------
-%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%-------------------------------------------------------------------    
-\rule{8cm}{12pt} \bigskip
-
 % The Galactic Habitable Zone is set at fixed values of the inner and outer
@@
 %
-With the aim of making a reproducible project, we developed the tools
+In order to make a reproducible project, we developed the tools
 that allow to run and obtain the results shown in this work.
@@
 which is publicly accessible through the GitHub
-platform\footnote{https://github.com/mlares/simu\_contact} under the
+platform\footnote{https://github.com/mlares/hearsay} under the
 MIT-license.
@@
 \hline
-   Model subset & D$_{max}$ [lyr] & $\tau_a$ interval [yr] & $\tau_s$ interval 
+   Model subset & D$_{\rm max}$ [lyr] & $\tau_a$ interval [yr] & $\tau_s$ interval 
    [yr]& description  \\
@@
 waiting time for the next awakening, $\left<\tau_a\right>$, and the
-mean lifetime, $\left<\tau_s\right>$) cover the ranges 4000-200000 yr
-and 10000-500000 yr, respectively, with a regular partition of 50
-values for each parameter.
-%
-For the D$_{max}$ parameter, we chose the values 500, 1000, 10000,
-20000, 40000 and 80000 lyr.
+mean lifetime, $\left<\tau_s\right>$) cover the ranges 10$^2$-10$^6$
+yr, with a linear partition of 50 values for each parameter.
+%
+For the D$_{\rm max}$ parameter, we take the range 30~pc to 10~kpc, on
+50 linear bins.
 %
 This setup makes a total of 750000 simulations, each covering a time
-range of one million years.
+range of ten million years.
 %
@@
 %
-In Table \ref{T_simu_hypotheses} we show the three variable
+In the Table \ref{T_simu_hypotheses} we show the three variable
 parameters, the ranges of their values and the number of bins that
@@
 %
-We also show the set of parameters that take part in the simulation,
-their values and the hypotheses that define the runs of the
-simulations.
+We also show the set of fixed parameters that take part in the simulation,
+their values and the hypotheses that they represent.
 
@@
 %
-Some quantities are directly derived from the discrete events, namely,
-the ID of emitting and receiving \cetis{}, the position in the galaxy, and
-the times of each of the events that are relevant to keep track of the
-number of \cetis{} (time of Awakening and time of Doomsday) and of the
-number of contacts (the times of contacts and the times of Blackouts).
-%
-We can also derive the additional quantities that represent the
+Some of them are directly derived from the discrete events, namely,
+the ID of emitting and receiving \cetis{} and the position in the
+galaxy.
+%
+We also save the times of each of the events that are relevant to keep
+track of the number of \cetis{} for each simulation, i.e., the times
+of the four types of events.
+%
+the times of C-type and B-type events are used to derive the number of
+contacts.
+%
+We also obtain quantities that represent the
 properties of the \cetis{}, for example the total time elapsed between the
-Awakening and the Doomsday of each node.
+A-type and the D-type of each node, which represent the corresponding
+lifetimes of the nodes.
 %
 The time span of a \ceti{} listening another or being listened by another
-can also be easily derived from the results of a realization.
+can also be derived by keepig track of the times of the events in a simulation.
 %
@@
 \cetis{}
-that reach contact and the waiting time until the first contact or the
-waiting time until the next contact.
-%
-Regarding the properties of the population of \cetis{} and its evolution,
-the simulation yields the fraction of awaken time a node is listening
-at least another node (i.e., in their causal cone), the age of
-contacted \cetis{} at first contact, the distribution of time to wait
-until next contact, the fraction of \cetis{} where the first contact is
-given at the awakening, the distribution of the number of contacts for
-each \ceti{}, the distribution of the number of contacts as a function of
-the age of the node, the number of contacts as a function of time in the galaxy,
-the fractions of nodes that succeed in contact, and the distribution of
+that reach contact, the distributions of the waiting times until the 
+first contact or the distributions of
+waiting times until the next contact.
+%
+The following properties of the population of \cetis{} can be derived:
+%
+the fraction of the lifetime a node is listening
+to
+at least another node (i.e., in their causal cone),
+%
+the age of
+contacted \cetis{} at first contact,
+%
+the fraction of \cetis{} where the first contact is
+given at the awakening,
+%
+the distribution of the number of contacts for
+each \ceti{},
+%
+the distribution of the number of contacts as a function of
+the age of the node,
+%
+the number of contacts as a function of time in the galaxy,
+%
+the fractions of nodes that succeed in making contact,
+%
+and the distribution of
 distances between contacted nodes.
@@
 It is also possible to analyze the relations between the distance to
-\ceti{} vs. the time of double communication, the distance to \ceti{} vs. the
+\ceti{} vs. the time of two-way communications, the distance to \ceti{} vs. the
 age of contacted node, the age of a node and the maximum number of
-contacted nodes before the Doomsday, or the lifespan of a node vs. the
+contacted nodes before the D--type event, or the lifespan of a node vs. the
 maximum number of contacts.
 %
-In addition, all these quantities can be analyzed as a function of the
+All these quantities can be analyzed as a function of the
 simulation parameters.
 %
-We chose eight models that are on fairly opposed regions of the
-explored parameter space, and cover short/long lifetimes, short/long
+We chose eight models that are on opposed regions of the
+explored parameter space, and cover
+all possible combinations of 
+short/long lifetimes, short/long
 waiting times for the next \ceti{} to appear in the galaxy and short/large
-range reach of the signal, including all possible combinations.
+range reach of the signal.
 %
@@
 %}}}
+
+%-------------------------------------------------------------------
+%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%-------------------------------------------------------------------    
+\rule{8cm}{12pt} \bigskip
 
@@
 $\left<\tau_a\right>$= 100000~yr, respectively$\left.\right)$, and two
-different signal reach ranges (D$_{max}$=40000 lyr and
-D$_{max}$=80000~lyr).
+different signal reach ranges (D$_{\rm max}$=40000 lyr and
+D$_{\rm max}$=80000~lyr).
 %
@@
 %
-As expected, of the samples with the longest $D_{max}$ the model with
+As expected, of the samples with the longest $D_{\rm max}$ the model with
 a dense awakening in the timeline (low $\left<\tau_a\right>$) and a
@@
 Similarly, for the models with long signal range
-($D_{max}=80000\,lyr$), the one which is dense and with long lifetime
+($D_{\rm max}=80000\,lyr$), the one which is dense and with long lifetime
 has the maximum number of contacts, reaching nearly 100 contacts for
@@
 differential, empirical distribution, so that the differences in the
-number of contacts for different models is quite large.
+number of contacts for different models are relatively large.
 %
@@
 time ($\left<\tau_a\right>$, in the range 0--200 yr), for three
-different values of the maximum signal range, $D_{max}=10000$~yr (left
-panel), $D_{max}=40000$~yr (middle panel), and $D_{max}=80000$~yr
+different values of the maximum signal range, $D_{\rm max}=10000$~yr (left
+panel), $D_{\rm max}=40000$~yr (middle panel), and $D_{\rm max}=80000$~yr
 (right panel).
@@
 for a \ceti{} of making causal contact with at least another \ceti{}
-during their entire lifetime, increases with increasing $D_{max}$,
+during their entire lifetime, increases with increasing $D_{\rm max}$,
 increasing $\left<\tau_s\right>$ and decreasing $\left<\tau_a\right>$,
@@
 
-While the number of \cetis{} that do not succeed in reaching the
-causal contact regions of other node is a useful indicator of
-degree of isolation, there is also the chance that a number of
+The number of \cetis{} that do not succeed in reaching the
+causal contact regions of other nodes is a useful indicator of
+the
+degree of isolation.
+%
+On the other hand, there is also the chance that a number of
 \cetis{} are already in the causal contact region of other nodes,
-which shows a somewhat opposite situation.
+at the time of their A-type events.
 %
@@
 ($\left<\tau_s\right>$, in the range 0--500 yr) and the mean awakening
-time ($\left<\tau_a\right>$, in the range 0--200 yr), for three
-different values of the maximum signal range, $D_{max}=10000$~yr (left
-panel), $D_{max}=40000$~yr (middle panel), and $D_{max}=80000$~yr
+time ($\left<\tau_a\right>$, in the range 0--200 yr).
+%
+The three panels correspond to different values of the maximum signal
+range ($D_{\rm max}$),
+as indicated in the Figure:
+%
+$D_{\rm max}=10000$~yr (left
+panel), $D_{\rm max}=40000$~yr (middle panel), and $D_{\rm max}=80000$~yr
 (right panel).
@@
 random \ceti{} to carry out a SETI project and spend a given time until
-the first contact is made.
-
+the first contact is made, under the hypotheses of the experiment.
+%
 \ffn{6}
@@
 Upper panel (a) panel shows the histograms for several values of the
-mean survival time $\tau_s$, and the mean awakening time $\tau_a$ in
+mean survival time $\tau_s$, with the mean awakening time $\tau_a$ in
 the range 68000-100000 yr.
@@
 Lower panel (b) shows the histograms for several values of $\tau_a$,
-and $\tau_s$ in the range 17000-24000 yr.
+with $\tau_s$ in the range 17000-24000 yr.
 %
@@
 For different values of $\tau_a$, the shape of the distribution is
-similar, and the main difference is given by the different number of
+similar, and the main difference is given by the variation on the number of
 contacts on the different models.
 %
-Conversely, for different values of $\tau_a$, the shape of the
-distribution is similar, and the main difference given by the
-different number of contacts on the different models.
-
-The fraction of the \cetis{} that have made at least one contact as a
+Conversely, for different values of $\tau_s$ (at a fixed range for
+$\tau_a$), the shape of the
+distribution varies signifficantly.
+
+The fraction of the \cetis{} that have made at least one contact, as a
 function of the elapsed time since the awakening, is computed with
 respect to the total number of nodes that make causal contact at least
-one time in the time range from the Awakening up to the Doomsday
+one time in the time range from the A--type up to the D--type
 events.
 %
-From a frequentist approach, the total counts from the awakening
+From a frequentist approach, the cumulative fraction of contacts
+from the awakening
 ($t=$0~yr) up to a given time are related to the estimation of the
@@
 The complement of this value is the probability of observing in a time
-interval with no success, i.e., without ever happening a Contact
+interval with no success, i.e., without ever happening a C--type
 event.
@@
 Clearly, this probability diminishes with time and tends to zero for
-large time periods.
-%             
-The diferential probability of making a causal contact reach its
+large time periods, meaning the a \ceti{} will eventually get aware of
+the presence of other \cetis{}
+
+Remarkably, the diferential probability of making a causal contact reach its
 maximum at the awakening event for all models.
 %
-That is, for a short period of time like the nearly 50 years SETI
-programs have been active, the initial moment is the most promising
-for making a (causal) contact, at least for a given technology.
-%
-This consideration offers a new approach to SETI programs, where the
-search for new communication technologies has a fundamental role.
+That is, for a short period of time the initial moment is the most promising
+for making a (causal) contact, for a given technology.
+%
+This allows to consider the fact that for the nearly 50 years SETI
+programs have been active, no contact event has been reported.
+%
+This offers a new approach to SETI programs, where the
+search for new communication technologies or the exploration of
+new communication channels has a fundamental role.
 %
@@
 %
-Remarkably, the probability of such a contact does not grows linearly
+The probability of such a contact does not grows linearly
 with the observing time, but is greater during the first 100
@@
 
-We have implemented a suite of simulations following a stochastic
-approach, to explore the hypothesis space of a simple model that
-accounts for the causal connections between communicating
-civilizations on a simplified galaxy.
-%
-The different models can be generated with a minimal number of free
-parameters and some fixed assumptions.
-%
-We argue that three parameters can be used to describe a variety of
-situations, ranging from a galaxy model where an intelligent civilization is
-very rare, to a galaxy model populated with plenty of civilizations in
-causal contact.
-%
-With this approach, we can estimate several quantities as a function
-of the parameters on the hypothesis space,
-so we explored the outcomes of different models 
-that arise as by-products of the simulations.
-%
-For instance, it is possible to estimate the timescale for an
-uninterrupted SETI effort to reach success, in terms of different
-model parameters that reflect different, so far unknown, scenarios for
-the appearance of intelligent life on the Galaxy.
+We have implemented a suite of simulations, following a stochastic approach, to explore a three dimensional hypothesis space of a simple model that accounts for the causal connections between communicating civilisations in the Galaxy. The different models can be generated with a minimal number of assumptions.
+We argue that three parameters can describe a variety of situations, ranging from a galaxy model where an intelligent civilization is very rare, to a galaxy model populated with plenty of civilisations in causal contact.
+With this approach, we can estimate several quantities as a function of the parameters on the hypothesis space, so we explored the outcomes of different models that can be derived from the simulations.
+For instance, it is possible to estimate the timescale for an uninterrupted SETI-equivalent effort to reach success, in terms of different model parameters that reflect different, so far unknown, scenarios for the appearance of intelligent life on the Galaxy.
 
 Although the simulations use a number of speculative assumptions, we
-argue that the degree of knowledge about the origin and persistence of
+argue that the limited knowledge about the origin and persistence of
 life in the Galaxy does not worth the implementation of more detailed
@@
 This causally connected nodes are sparsely distributed in both space
-and time, making it difficult the analytical treatment and justifying
+and time, making analytical treatments difficult and justifying
 the simulation approach.
@@
 Under the hypotheses of our experiments, we conclude that a causal
-contact is extremely unlikely unless the galaxy is heavily populated
+contact is extremely unlikely unless the galaxy is densely populated
 by intelligent civilizations with large average lifetimes.
 %
-Roughly, in order to have at least one contact in the entire lifetime
-of the \ceti{}, there should appear a mean of at least one \ceti{}
+On average, in order to have at least one contact in the entire lifetime
+of the \ceti{}, there must appear a mean of at least one \ceti{}
 every $\sim$20000 yr, with a mean active period of at least
@@
 number of extraterrestrial civilizations that could contact Earth
-using Monte Carlo simulations, from a statistical model where the with
+using Monte Carlo simulations, from a statistical model where the width
 of the Galactic disk is not negligible.
@@
 %
-Remarkably, most of the studies using statistical models or
-simulations are based on the Drake equation
-\citep{cirkovic_temporal_2004, smith_broadcasting_2009,
-bloetscher_using_2019}.
-%
-Our approach does not relay on the Drake equation, and thus is not
-motivated by the detailed physical processes that give rise to
+Remarkably, most
+
+%
+Unlike most 
+of the studies that make use of statistical models or
+simulations \citep{cirkovic_temporal_2004, smith_broadcasting_2009,
+bloetscher_using_2019}, our approach does not relay on the Drake
+equation.
+%
+Thus, it does not need
+a detailed description or modeling of physical processes that give rise to
 intelligent life.
@@
 However, we argue that it is a valid empirical formulation to discuss
-the probabilities of contact, and mainly time scales involed in the
+the probabilities of contact, and the time scales involed in the
 problem.
@@
 
-There is a balance between the density of \cetis{} and the mean
+We also fond that there 
+is a balance between the density of \cetis{} and the mean
 lifetime since, as expected, a lower density can be compensated by a
@@
 
-In the approach of this work, we used computer simulations to address
-the problem of the probabilities of causal contacts between location
-in the galaxy the possibility of sending and receiving messages.
+In the approach of this work, we have used computer simulations to address
+the problem of the probabilities of causal contacts between locations
+in the galaxy with the possibility of sending and receiving messages.
 %
@@
 to propose an alternative treatment for the number of intelligent
-emitter/receivers with emphasis on the temporal dimension which limits
-the probabilities of contacts due to the short time interval for
-rise and fall of civilizations compared to the age and extension of
-our Galaxy.
+emitter/receivers.
+%
+The short time interval between the
+rise and fall of civilizations, compared to the age and extension of
+our Galaxy, is a fundamental limitation for the number of contacts.
+%
+Thus, the temporal dimension, which is missing in the Drake
+equation, is a key factor to understand the network of contacts on
+different scenarios.
 

### 2020-05-18 submit_v03/manuscript.tex
minors

@@
 
-%-------------------------------------------------------------------
-%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%-------------------------------------------------------------------    
-\rule{8cm}{12pt} \bigskip
-
 \subsection{Membership to the network of connected \cetis{}}\label{SS_members}
@@
 $\tau_a$), the shape of the
-distribution varies signifficantly.
+distribution varies significantly.
 
@@
 
-Remarkably, the diferential probability of making a causal contact reach its
+Remarkably, the differential probability of making a causal contact reach its
 maximum at the awakening event for all models.
@@
 %
-Remarkably, most
-
-%
 Unlike most 
@@
 
-%%%%%%%%%%%%%%%
-
 We also fond that there 
@@
 Plots and simulations were made with software developed by the authors
-using R and python languajes. Plots were postprocessed with inkscape.
+using R and python languages. Plots were postprocessed with inkscape.
 %

### 2020-05-25 submit_v03/manuscript.tex
faltan los plots

@@
 \title[Probabilities of SETI contacts]{Probability of causal
-contact between interstellar civilizations through MonteCarlo simulations}
+contacts between interstellar civilizations through Monte Carlo simulations}
+%
+%A network of constrained causal contacts in the Galaxy as a proxy for
+%the probability of 
+%
+% A model for constrained causal contacts between location in the GHZ
 
@@
 %
-The abundance of intelligent civilizations in the Galaxy is a
-   longstanding question, often conceptualized as the problem of the
-   lack of received communication. Estimates of the number of
-   civilizations are generally guided by the Drake equation, despite
-   the large uncertainties in its factors and its lack of a temporal
-   nature. Alternative approaches use detailed numerical simulations
-   of the galaxy and recipes for the stars and planets formation
-   rates and for the origin of life, and rely on uncertain parameters.
-   We present a statistical model for the abundance and duration of
-   civilizations implemented through Monte Carlo simulations. We
-   explore the hypothesis space of the model by a suite of simulations
-   to analyze the emergence of communicating nodes and their causal
-   connections. We based this on minimal assumptions and three free
-   parameters, with focus on the proposed statistical properties of
-   empirical models. Our analysis of the dependence of the rate of
-   causal contacts on the mean number of civilizations, the mean
-   lifespan distribution and the maximum distance a civilization can
-   send signals, is considered to discuss the spatial and temporal
-   structure of a populated galaxy within several scenarios. We find
+The abundance of intelligent civilisations in the Galaxy is a
+   longstanding question related to the expected probability
+   of success for SETI programs.  Traditional approaches to quantify
+   this problem use either frequentist inferences or numerical
+   simulations. The key challenges in both cases are the proper
+   treatment of the time variable and the large uncertainties on the
+   intervening factors. Here we present a three-parametric
+   statistical model of the network of causally connected locations in
+   a simplified Galaxy, with the restriction of a maximum distance
+   range for any signal. We implement Monte Carlo simulations of this
+   model to explore the parameter space and analyse the probabilities
+   of causal contacts for random locations in the Galaxy. We find
    that, given the enormous distances involved, causal contacts
-   between civilizations are very rare. The odds to make a contact in
-   a few years of monitoring are low for most models, except for those
-   of a galaxy densely populated with long-lived civilizations. The
-   probability of causal contacts increases with the lifetime of
-   civilizations much more significantly than with the number of
-   active civilizations for a time window. We show that the maximum
-   probability of making a contact occurs when a civilization gains
-   the required communication technology.
+   between civilisations would be very rare. The odds to make a
+   contact in a few years of monitoring are low for most models,
+   except for those of a galaxy densely populated with ancient
+   civilisations. The probability of causal contacts increases with
+   the lifetime of civilisations more significantly than with the
+   number of active civilisations. We show that the maximum
+   probability of making a contact occurs when a civilization
+   discovers the required communication technology.
 %
@@
 The Drake equation  \citep{drake_intelligent_1962} provides a truly
-helpful educated guess, a rational set of lenses –the factors in the
-equation– through which to look at future contacts with
+helpful educated guess, a rational set of lenses --the factors in the
+equation-- through which to look at future contacts with
 technologically advanced civilizations in the Milky Way. 
@@
 lifetime of 10$^9$ years and for a maximum radius for the detection of
-the signal of 1000 yr, is roughly 0.1.
+the signal of 1000~ly, is roughly 0.1.
 %
@@
 directions.   
+%
+In this work we use the International System of units recommended by
+the International Astronomial Union, where the simbol for years is
+''a''.
 %}}}
@@
    $\tau_{a}$ & Mean temporal separation between consecutive
-   awakenings & 10$^2$~yr & 10$^6$~yr & 50\\ 
+   awakenings & 10$^2$~a & 10$^6$~a & 50\\ 
    $\tau_{s}$ & Mean lifetime of a \ceti{}
-   & 10$^2$yr & 10$^6$~yr& 50\\ 
+   & 10$^2$a & 10$^6$~a& 50\\ 
    D$_{\rm max}$ & Maximum reach of a message  & 30~pc & 10~kpc & 50 \\
@@
    R$_{GHZ}^{\rm max}$   & Outer radius of the GHZ     &&& 9~kpc\\
-   t$_{\rm max}$ & Time span of the simulation  &&&1.e6 yr\\
+   t$_{\rm max}$ & Time span of the simulation  &&&1.e6~a\\
     & \multicolumn{3}{l}{number of random realizations for each point in the hypothesis space} &50\\
@@
 Left panel shows the histograms for several values of $\tau_a$, and
-$\tau_s$ in the range 17000-24000 yr.
+$\tau_s$ in the range 17000-24000~a.
 %
 Right panel shows the histograms for several values of $\tau_s$, and
-$\tau_a$ in the range 68000-100000 yr.
+$\tau_a$ in the range 68000-100000~a.
 %
@@
 \hline
-   Model subset & D$_{\rm max}$ [lyr] & $\tau_a$ interval [yr] & $\tau_s$ interval 
-   [yr]& description  \\
+   Model subset & D$_{\rm max}$ [lyr] & $\tau_a$ interval [a] & $\tau_s$ interval 
+   [a]& description  \\
 \hline
@@
 waiting time for the next awakening, $\left<\tau_a\right>$, and the
-mean lifetime, $\left<\tau_s\right>$) cover the ranges 10$^2$-10$^6$
-yr, with a linear partition of 50 values for each parameter.
+mean lifetime, $\left<\tau_s\right>$) cover the ranges 10$^2$-10$^6$~a, with a linear partition of 50 values for each parameter.
 %
@@
 samples, including short and long lifetimes
-$\left(\,\left<\tau_s\right>=\right.$10000~yr and
-$\left<\tau_s\right>$=100000~yr, respectively$\left.\right)$, dense
+$\left(\,\left<\tau_s\right>=\right.$10000~a and
+$\left<\tau_s\right>$=100000~a, respectively$\left.\right)$, dense
 and sparse spatial distribution
-$\left(\,\left<\tau_a\right>=\right.$10000~yr and
-$\left<\tau_a\right>$= 100000~yr, respectively$\left.\right)$, and two
+$\left(\,\left<\tau_a\right>=\right.$10000~a and
+$\left<\tau_a\right>$= 100000~a, respectively$\left.\right)$, and two
 different signal reach ranges (D$_{\rm max}$=40000 lyr and
@@
 listen to another \ceti{}), as a function of the mean lifetime
-($\left<\tau_s\right>$, in the range 0--500 yr) and the mean awakening
-time ($\left<\tau_a\right>$, in the range 0--200 yr), for three
-different values of the maximum signal range, $D_{\rm max}=10000$~yr (left
-panel), $D_{\rm max}=40000$~yr (middle panel), and $D_{\rm max}=80000$~yr
+($\left<\tau_s\right>$, in the range 0--500 a) and the mean awakening
+time ($\left<\tau_a\right>$, in the range 0--200 a), for three
+different values of the maximum signal range, $D_{\rm max}=10000$~a (left
+panel), $D_{\rm max}=40000$~a (middle panel), and $D_{\rm
+max}=80000$~a
 (right panel).
@@
 as a function of the mean lifetime
-($\left<\tau_s\right>$, in the range 0--500 yr) and the mean awakening
-time ($\left<\tau_a\right>$, in the range 0--200 yr).
+($\left<\tau_s\right>$, in the range 0--500 a) and the mean awakening
+time ($\left<\tau_a\right>$, in the range 0--200 a).
 %
@@
 %
-$D_{\rm max}=10000$~yr (left
-panel), $D_{\rm max}=40000$~yr (middle panel), and $D_{\rm max}=80000$~yr
+$D_{\rm max}=10000$~a (left
+panel), $D_{\rm max}=40000$~a (middle panel), and $D_{\rm
+max}=80000$~a
 (right panel).
@@
 mean survival time $\tau_s$, with the mean awakening time $\tau_a$ in
-the range 68000-100000 yr.
+the range 68000-100000 a.
 %
 Lower panel (b) shows the histograms for several values of $\tau_a$,
-with $\tau_s$ in the range 17000-24000 yr.
+with $\tau_s$ in the range 17000-24000 a.
 %
@@
 from the awakening
-($t=$0~yr) up to a given time are related to the estimation of the
+($t=$0~a) up to a given time are related to the estimation of the
 probability of listening during that time interval and reaching to
@@
 
-We have implemented a suite of simulations, following a stochastic approach, to explore a three dimensional hypothesis space of a simple model that accounts for the causal connections between communicating civilisations in the Galaxy. The different models can be generated with a minimal number of assumptions.
-We argue that three parameters can describe a variety of situations, ranging from a galaxy model where an intelligent civilization is very rare, to a galaxy model populated with plenty of civilisations in causal contact.
-With this approach, we can estimate several quantities as a function of the parameters on the hypothesis space, so we explored the outcomes of different models that can be derived from the simulations.
-For instance, it is possible to estimate the timescale for an uninterrupted SETI-equivalent effort to reach success, in terms of different model parameters that reflect different, so far unknown, scenarios for the appearance of intelligent life on the Galaxy.
-
-Although the simulations use a number of speculative assumptions, we
-argue that the limited knowledge about the origin and persistence of
-life in the Galaxy does not worth the implementation of more detailed
-or sophisticated models.
-%
-Instead, we take advantage of the simplicity of the model to explore
-the hypothesis space, i.e., the three--dimensional parameter space of
-the parametric model, in order to gain insight on the consequences of
+We presented a stochastic model (SC3Net) to analyse the network of
+constrained causally connected nodes in a simplified Galaxy.
+%
+It represents an idealised scenario of perfectly efficient emitters and
+receivers with the restriction of a maximum distance separation.
+%
+These emitters and receivers correspond to nodes that form a
+communication network whose properties depend on their density
+and mean survival time.
+%
+The statistical analysis of the model allows to estimate the
+probabilities for a random node and for a given model instance of
+making contacts with other nodes, along with the waiting times and
+durations of such contacts.
+%
+Using numerical simulations, we implemented this model to explore the
+three-dimensional parameter space, considering a grid of mean time
+separations between the activation of new nodes
+in the range 500--10$^6$~a.
+%
+Additionally, we explore a grid of the mean survival times between 
+500~a and 10$^6$~a, and values for the maximum distance range of signals 
+of 100~pc, 500~pc, 1~kpc, 5~kpc and 10~kpc.
+%
+The simulation of each parameter point was performed several times
+with different random seeds in order to improve the confidence of the
+derived quantities.
+%
+The full experiment comprises 46485 simulations for a total of 5184
+unique points in the parameter space.
+
+% We have implemented a suite of simulations, following a stochastic
+% approach, to explore a three dimensional hypothesis space of a simple
+% model that accounts for the causal connections between
+% communicating civilisations in the Galaxy. The different models can be
+% generated with a minimal number of assumptions. We argue that three
+% parameters can describe a variety of situations, ranging from a galaxy
+% model where an intelligent civilization is very rare, to a galaxy
+% model populated with plenty of civilisations in causal contact. With
+% this approach, we can estimate several quantities as a function of the
+% parameters on the hypothesis space, so we explored the outcomes of
+% different models that can be derived from the simulations. For
+% instance, it is possible to estimate the timescale for an
+% uninterrupted SETI-equivalent effort to reach success, in terms of
+% different model parameters that reflect different, so far unknown,
+% scenarios for the appearance of intelligent life on the Galaxy.
+
+Although the simulations use several hypotheses, we
+argue that the model is not worthy of further complexity,
+given the limited knowledge about the origin and persistence of
+life in the Galaxy.
+%
+The implementation of more detailed or sophisticated models
+would increase the number of free parameters without any 
+improvement in their predictive power.
+%
+Thus, we take advantage of the simplicity of the model to explore
+the parameter space in order to gain insight on the consequences of
 different scenarios for the search of intelligent life.
@@
 of the \ceti{}, there must appear a mean of at least one \ceti{}
-every $\sim$20000 yr, with a mean active period of at least
-$\sim$30000 yr.
+every $\sim$20000 a, with a mean active period of at least
+$\sim$30000 a.
 %

### 2020-05-25 submit_v03/manuscript.tex
advence

@@
 % 
-% 
-% 
-% 
-\Fpagebreak
-%
-We stress the fact that the only observation that can be stated with
-certainty is that for the number of years projects for the Search of
-Extraterrestrial Intelligences (SETI) have been working; we have
-received no signal using the technical possibilities and the
-conditions established by them \citep{tarter_search_2001}.
+We stress the fact that the lack of detections of extraterrestrial intelligent signals
+is
+the only observational fact that can be taken
+for sure.
+%
+Indeed, for the number of years projects for the Search of
+Extraterrestrial Intelligences (SETI) have been working,
+no signal has been received using the technical possibilities 
+within their search criteria \citep{tarter_search_2001}.
 %
@@
 %
+\Fpagebreak
+%
 The Drake equation is a key tool to organise the discussion about the
@@
 The optimistic estimations from the Drake equation also contrast with
-the SETI initiatives, noticeable overshadowed because despite decades
-of effort, not a single evidence for intelligent signals has been
-detected.
+the results from the SETI initiatives.
+%
+Notably, despite decades of effort not a single evidence for intelligent signals has been detected.
 %
@@
 \citet{balbi_impact_2018} fix the total number of civilisations and
-explore the parameter space that comprise three parameters, namely,
+explore the parametric space that comprises three parameters, namely,
 the distance to the Earth, the time of appearance and the lifespan of
@@
 Each of these three variables are drawn from a random distribution.
-The distribution of the distances results from a uniform distribution
-of civilizations within the plane of the Galaxy.
+%
+The distribution of the distances results from a uniform model for the
+positions of civilizations within the plane of the Galaxy.
 %
 For the distribution of the characteristic time of appearance, the
-author explores an exponential and a truncated Gaussian distributions.
-For the distribution of the lifespans, the author chooses an
+author explores exponential and truncated Gaussian distributions.
+%
+Moreover, for the distribution of the lifespans, the author chooses an
 exponential distribution.
@@
 
+% INTRODUCTION OF *THIS* WORK
+
 In this work we address the problem of the temporal and spatial
@@
 %
-To that end, we reduce and explore the hypothesis space over a set of
-three of parameters.
-%
-We propose to avoid the frequentist approach of the Drake equation to
+We make no asumptions about the origin of life, the development 
+of intelligence or the formation of habitable planets according to stellar type.
+%
+Instead, we assume monoparametric function families to model the
+appeareance of points in the disk of the Galaxy
+
+%
+To that end, we propose a model which depends on a set of only
+three parameters.
+%
+We abandon the frequentist approach of the Drake equation to
 compute the number of civilisations, providing its statistical
@@
 %
-This is an exploratory analysis that aims at providing a numerical
+This is an exploratory analysis that aims at developing a numerical
 tool to discuss the different scenarios based on statistical
@@
 %
-With this, we aim at avoiding the several theoretical problems
+With this, we avoid the several theoretical problems
 summarised by the Drake equation factors.
@@
 The approach proposed here should be considered as a compromise
-between the uncertainties of the frequentist approach and the detailed
-recipes required on the simulation approach.
+between the uncertainties of the frequentist estimations and the detailed
+recipes required on the numerical simulations.
 %
@@
 ''a''.
+%}}}
+
+%}}}
+
+\begin{figure}[!t]
+   \centering
+   \includegraphics[width=0.5\textwidth]{F_scheme.pdf}
+   \caption{
+   %
+Space--time diagrams showing a schematic representation of the
+different stages in the development of causal contact node (\ceti{}).
+%
+Panel (a) represents the region $R$ in space and time which is causally
+connected to the emitter (left vertex), following an "A" type event
+("Awakening") in which the node acquires the communication capability.
+%
+The sphere of first contact (SFC) is a sphere centered on the emitter
+that grows until its radius reaches the D$_{\rm max}$ distance, in which
+the power of a signal would equal the detectability threshold.
+%
+In the Figure this sphere is represented by the left triangle of the
+region $R$.
+%
+The surface of last contact (SLC) is another sphere that grows from a
+"D" event ("Doomsday").
+%
+The region which is causally connected to the emitter is then limited
+by these two spheres, and has the shape of a sphere or of a spherical
+shell, depending on the time.
+%
+The temporal intervals for
+the communication between two \cetis{} are represented in he
+panel (b).
+%
+The receiver \ceti{} E$_1$ can listen signals from the emiter \ceti{} E$_2$,
+from the "Contact" event (t=C$_{12}$) up to the "Blackout" event (t=B$_{12}$).
+%
+Similarly, 
+the receiver \ceti{} E$_2$ can listen signals from the emiter \ceti{} E$_1$,
+from t=C$_{21}$ up to t=B$_{21}$. 
+%
+   } \label{F_scheme}
+\end{figure}
+
 %}}}
@@
 
-%}}}
-
-\begin{figure}[!t]
-   \centering
-   \includegraphics[width=0.5\textwidth]{F_scheme.pdf}
-   \caption{
-   %
-Space--time diagrams showing a schematic representation of the
-different stages in the development of causal contact node (\ceti{}).
-%
-Panel (a) represents the region $R$ in space and time which is causally
-connected to the emitter (left vertex), following an "A" type event
-("Awakening") in which the node acquires the communication capability.
-%
-The sphere of first contact (SFC) is a sphere centered on the emitter
-that grows until its radius reaches the D$_{\rm max}$ distance, in which
-the power of a signal would equal the detectability threshold.
-%
-In the Figure this sphere is represented by the left triangle of the
-region $R$.
-%
-The surface of last contact (SLC) is another sphere that grows from a
-"D" event ("Doomsday").
-%
-The region which is causally connected to the emitter is then limited
-by these two spheres, and has the shape of a sphere or of a spherical
-shell, depending on the time.
-%
-The temporal intervals for
-the communication between two \cetis{} are represented in he
-panel (b).
-%
-The receiver \ceti{} E$_1$ can listen signals from the emiter \ceti{} E$_2$,
-from the "Contact" event (t=C$_{12}$) up to the "Blackout" event (t=B$_{12}$).
-%
-Similarly, 
-the receiver \ceti{} E$_2$ can listen signals from the emiter \ceti{} E$_1$,
-from t=C$_{21}$ up to t=B$_{21}$. 
-%
-   } \label{F_scheme}
-\end{figure}
-
 %{{{
@@
 In this scheme, a light pulse would follow a trajectory represented by
-a line at 45 degrees from the axes, given the units of the time and
+a line at 45 degrees from the axes, given the units of time and
 space axes are years and light years, respectively.
@@
 
-%}}}
-
-\begin{figure*}
-   \centering
-   \includegraphics[width=\textwidth]{F_messages.pdf}
-   \caption{
-   %   
-Schematic representation on space--time diagrams of the possible cases in which a
-\ceti{} E (Emiter, solid line) can be in causal contact with a \ceti{} R
-(Receiver, dashed line).
-%
-Causal contact can be produced either if the emiter \ceti{} appears before
-(left) or after (right) the receiver \ceti{}.
-%
-The causal contact in which the receiver can have access to the signal 
-from the emiter is produced between the C event (Contact, open circles) and the B
-event (Blackout, filled circles).
-%
-The duration of the causal contact in one direction depends on
-several factors, mainly the time lag between the awakening events
-(horizontal separation in the space--time diagrams)
-and the distance between the \cetis (vertical separation in the
-space--time diagrams).
-%   
-The different configurations must be taken into account in order to 
-obtain the list of the upcomming events.
-%
-The time intervals for a two--way communication to be possible are
-indicated in double lines. 
-%
-   } \label{F_messages}
-\end{figure*}
-
-%{{{
-
 The temporal structure that emerges from the experimental setup implies that, as a fundamental property of the simulations, a causal connection can be produced without 
@@
 The type and length of causal contact in both directions depend on the distance and time lag between the awakening events, the maximum distance that a signal can reach and the time period in which each emitter is active.
-%
-\ffn{2}
-%
-In the Fig.~\ref{F_messages} we show a schematic representation of the possible cases in which a \ceti{} E (standing for Emiter) can be in causal contact with a \ceti{} R (receiver).
-%
-The duration of such a contact in one way depends on several factors, mainly the lifetimes, the time delay between the A-type events and the distance between the two nodes.
-%
-As it can be seen, the direction of the causal contact does not imply any ordering for the A-type events of the involved emitter/receiver pair.
-%
-For instance, a node can be in the light cone of another node that turned up later, i.e., the receiver is older than the emitter, like in the cases (e) and (f) of the Figure.
-%
-In the implementation of the DE simulation, all possibilities for causal contacts are reduced to the six cases shown in the Figure.
-%
-Once a pair of nodes become active and their properties have been assigned, it is possible to determine if and when they will get connected.
-%
-The times derived for such events are added to a stack of times so that they are visited in order, and in turn trigger further events.
-%
-% EJEMPLOS???
 
@@
 \hline
-   $\tau_{a}$ & Mean temporal separation between consecutive
-   awakenings & 10$^2$~a & 10$^6$~a & 50\\ 
+   $\tau_{a}$ & Mean temporal separation between consecutive awakenings 
+	      & 500~a & 40500~a & 21\\ 
+	    & & 5$\cdot$10$^5$~a & 10$^6$~a & 51\\ 
    $\tau_{s}$ & Mean lifetime of a \ceti{}
-   & 10$^2$a & 10$^6$~a& 50\\ 
-   D$_{\rm max}$ & Maximum reach of a message  & 30~pc & 10~kpc & 50 \\
+	      & 500~a & 40500~a & 21\\ 
+	    & & 5$\cdot$10$^5$~a & 10$^6$~a & 51\\ 
+	D$_{\rm max}$ & Maximum reach of a message  & \multicolumn{3}{c}{100 pc, 500 pc, 1 kpc, 5 kpc, 10 kpc} \\
 \hline
@@
    & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
-   R$_{GHZ}^{\rm min}$   & Inner radius of the GHZ     & & & 7~kpc\\
-   R$_{GHZ}^{\rm max}$   & Outer radius of the GHZ     &&& 9~kpc\\
-   t$_{\rm max}$ & Time span of the simulation  &&&1.e6~a\\
-    & \multicolumn{3}{l}{number of random realizations for each point in the hypothesis space} &50\\
+	R$_{GHZ}^{\rm min}$   & Inner radius of the GHZ  & \citet{lineweaver_galactic_2004} & & 7~kpc\\
+   R$_{GHZ}^{\rm max}$   & Outer radius of the GHZ       & \citet{lineweaver_galactic_2004} & & 9~kpc\\
+	t$_{\rm max}$ & Time span of the simulation  & & \multicolumn{2}{r}{5$\cdot$10$^6$~a, 10$^7$~a} \\
+    & \multicolumn{3}{l}{number of random realizations for each point in the hypothesis space} & 10, 5, 1 \\
 \hline

### 2020-05-26 submit_v03/manuscript.tex
.

@@
 \section{Motivations}\label{S_motivations}
-%{{{
 
@@
 % 
-We stress the fact that the lack of detections of extraterrestrial intelligent signals
-is
-the only observational fact that can be taken
+We stress the fact that the lack of detections of extraterrestrial
+intelligent signals is the only observational fact that can be taken
 for sure.
@@
 Indeed, for the number of years projects for the Search of
-Extraterrestrial Intelligences (SETI) have been working,
-no signal has been received using the technical possibilities 
-within their search criteria \citep{tarter_search_2001}.
+Extraterrestrial Intelligences (SETI) have been working, no signal has
+been received using the technical possibilities within their search
+criteria \citep{tarter_search_2001}.
 %
@@
 %
-Notably, despite decades of effort not a single evidence for intelligent signals has been detected.
+Notably, despite decades of effort not a single evidence for
+intelligent signals has been detected.
 %
@@
 %
-We make no asumptions about the origin of life, the development 
-of intelligence or the formation of habitable planets according to stellar type.
+We make no asumptions about, e.g., the origin of life, the development
+of intelligence or the formation of habitable planets according to
+stellar type.
 %
 Instead, we assume monoparametric function families to model the
-appeareance of points in the disk of the Galaxy
-
-%
-To that end, we propose a model which depends on a set of only
-three parameters.
-%
+appearance of points in the disk of the Galaxy and over the time,
+which we call \textit{nodes}.
+%
+These nodes represent the locations of ideal intelligent agents that
+are able to receive and emit signals with perfect efficiency in all
+directions.
+%
+Then we analyse the network of causally connected nodes limited to
+have a maximum separation representing the maximum distance a signal
+can travel with an intensity above a fixed threshold.
+%
+A node is causally connected to other node if it is within their light
+cone.
+%
+We adopt as a definition that a light cone is the region in spacetime
+within the surface generated by the light emanating from a given point
+in space in a given period of time.
+%
+For example, two nodes separated by 100 light years which appear with
+a difference in time of 80 years and last 50 years each, will be
+causally connected for 30 years.
+%
+This connection is not bidirectional.
+%
+In fact, the second node will see the first one for 30 years, but the
+first node will never acknowledge the existence of the second.
+%
+This simple tought experiment exemplify the importance of the time
+variable to analyse the structure of connections.
+%
+The model can be described with three parameters, namely, the mean separation
+between the appearance of nodes, the mean lifetime of the nodes, and
+the maximum separation to allow an effective causal contact.
+%
+The model presented here comprises 
+the stochastic networks of constrained causally connected nodes,
+together with the three parameters and additional hypotheses.
+%
+Hereafter, we refeer to this model as the C3Net model, standing for
+''Constrained Causally Conencted Network``.
+
 We abandon the frequentist approach of the Drake equation to
@@
 
-%}}}
-
 \begin{figure}[!t]
@@
 \section{Methods and working hypotheses}\label{S_methods}
-%{{{
 
@@
 The analytical expresiones for these distributions are set to a fixed law, as discussed in Sec.~\ref{SS_PDF_shape}.
-
-%{{{
 

### 2020-05-26 submit_v03/manuscript.tex
texto

@@
 
-\newcommand{\ceti}{CCN}
-\newcommand{\cetis}{CCNs}
+\newcommand{\ceti}{C3N}
+\newcommand{\cetis}{C3Ns}
 \newcommand{\hs}{\textsc{hearsay}}
@@
 %%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
-\section{Motivations}\label{S_motivations}
+\section{Introduction}\label{S_motivations}
 
@@
 %
-Indeed, for the number of years projects for the Search of
-Extraterrestrial Intelligences (SETI) have been working, no signal has
-been received using the technical possibilities within their search
-criteria \citep{tarter_search_2001}.
+Indeed, as \citep{tarter_search_2001} pointed out, according to our current technical
+capabilities for SETI, we have not received any signal yet.
+%
+%Indeed, for the number of years projects for the Search of
+%Extraterrestrial Intelligences (SETI) have been working, no signal has
+%been received using the technical possibilities within their search
+%criteria \citep{tarter_search_2001}.
 %
@@
 However, the uncertainties in the factors of this equation make it
-less prone to a formal study to define searching strategies or to
+less suitable to a formal study to define searching strategies or to
 compute the actual number of extraterrestrial intelligences.
 %
-The optimistic estimations from the Drake equation also contrast with
-the results from the SETI initiatives.
-%
-Notably, despite decades of effort not a single evidence for
-intelligent signals has been detected.
+%The optimistic estimations from the Drake equation also contrast with
+%the results from the SETI initiatives.
+%
+%Notably, despite decades of effort not a single evidence for
+%intelligent signals has been detected.
+%
+A number of studies propose alternative formalisms for the estimation of the
+likelihood of detecting intelligent signals from space.
 %
@@
 equation parameters on the spectral type of the host stars and the
-time since the Galaxy formed and examine trajectories for the
+time since the Galaxy formed, and examine trajectories for the
 emergence of communicative civilisations.
@@
 Temporal aspects of the distribution of communicating civilisations
-and their contacts have also been explored by several authors
+and their contacts have been explored by several authors
 \citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
@@
 Although this approach does not rely on observed quantities that could
-be modified with time, it also depend on the definition of a number of
-parameters which are unknown or uncertain.
-%
-\citet{grimaldi_signal_2017} consider a statistical model for the
+be modified with time (like, for instance, the fraction of stars with planets), 
+it also depend on the definition of a number of
+parameters which are unknown or uncertain, required to carry out
+a the simulations.
+%
+\citet{grimaldi_signal_2017} considers a statistical model for the
 probability of the Earth contacting other intelligent civilisations,
@@
 
-In a recent work, \citep{balbi_impact_2018} use a statistical model to
+In a recent work, \citet{balbi_impact_2018} uses a statistical model to
 analyse the occurrence of causal contacts between civilisations in the
@@
 %
-\citet{cirkovic_temporal_2004} also emphasise the lack of temporal
-structure in the Drake equation.
-%
-\citet{bloetscher_using_2019} uses a probabilistic approach, though
-still motivated by the Drake equation, to obtain a probabilistic
-measure for the number of civilisations in the Galaxy.
-%
-To that end, the authors do Monte-Carlo Markov Chains over each factor
+\citet{cirkovic_temporal_2004} also emphasises the lack of temporal
+structure in the Drake equation and, in particular, the limitations
+of this expresion to estimate the required timescale of a SETI
+program to succeed in the detection of intelligent signals.
+%
+Despite its shotcomming for a solid mathematical foundation of SETI, 
+the Drake equation has also motivated efforts to produce formal
+statistical results.
+%
+Recently, \citet{bloetscher_using_2019} considers a Bayesian approach,
+still motivated by the Drake equation, to estimate
+the number of civilisations in the Galaxy.
+%
+To that end, the authors employ Monte Carlo Markov Chains over each factor
 of the Drake equation, and combine the means to reach a probabilistic
@@
 % 
-Adopting a novel approach, \citet{balbi_impact_2018} investigate the
+Different statistical modelings have been also proposed.
+%
+A notable example of this es the work by \citet{balbi_impact_2018}, who investigates the
 chance of communicating civilisations making causal contact within a
@@
 involves the distance between civilisations, their lifespan and their
-times of appearance. This is important since the time it takes the
-light to travel across the Galaxy is much lesser than its lifetime.
-%
-\citet{balbi_impact_2018} fix the total number of civilisations and
+times of appearance.
+%
+This is important since the time it takes the
+light to travel across the Galaxy might be much lesser than its lifetime.
+%
+\citet{balbi_impact_2018} fixes the total number of civilisations and
 explore the parametric space that comprises three parameters, namely,
@@
 %
-\citet{balbi_impact_2018} finds that the fraction of communicating
+It is important to consider that the fraction of communicating
 civilisations vary with the choice of the statistical distribution for
-the time of appearance. 
-%
-For all analysed distributions, the fraction corresponding to a mean
+the time of appearance, as shown by \citet{balbi_impact_2018}.
+%
+For all analysed distributions, the author concludes that,
+the detectable fraction corresponding to
+
+for a mean
 lifetime of 10$^9$ years and for a maximum radius for the detection of
-the signal of 1000~ly, is roughly 0.1.
+the signal of 1000~ly, the detectable civilizations
+are roughly the 10 per cent of the total, due to their spread in time and
+their limited lifetimes.
 %
@@
 
-In this work we address the problem of the temporal and spatial
-structure of the distribution of communicating civilisations.
-%
-We make no asumptions about, e.g., the origin of life, the development
-of intelligence or the formation of habitable planets according to
-stellar type.
-%
-Instead, we assume monoparametric function families to model the
-appearance of points in the disk of the Galaxy and over the time,
-which we call \textit{nodes}.
-%
-These nodes represent the locations of ideal intelligent agents that
-are able to receive and emit signals with perfect efficiency in all
-directions.
-%
-Then we analyse the network of causally connected nodes limited to
-have a maximum separation representing the maximum distance a signal
-can travel with an intensity above a fixed threshold.
-%
-A node is causally connected to other node if it is within their light
-cone.
-%
-We adopt as a definition that a light cone is the region in spacetime
-within the surface generated by the light emanating from a given point
-in space in a given period of time.
-%
-For example, two nodes separated by 100 light years which appear with
-a difference in time of 80 years and last 50 years each, will be
-causally connected for 30 years.
-%
-This connection is not bidirectional.
-%
-In fact, the second node will see the first one for 30 years, but the
-first node will never acknowledge the existence of the second.
-%
-This simple tought experiment exemplify the importance of the time
-variable to analyse the structure of connections.
+In this work we address the problem of the temporal and spatial structure of
+the distribution of communicating civilisations.
+%
+We make no asumptions about, e.g., the origin of life, the development of
+intelligence or the formation of habitable planets according to stellar type.
+%
+Instead, we assume monoparametric function families to model the appearance of
+points in the disk of the Galaxy and over the time, which we call
+\textit{nodes}.
+%
+These nodes represent the locations of ideal intelligent agents that are able
+to receive and emit signals with perfect efficiency in all directions.
+%
+Then we analyse the network of causally connected nodes limited to have a
+maximum separation representing the maximum distance a signal can travel with
+an intensity above a fixed threshold.
+%
+A node is causally connected to other node if it is within their light cone.
+%
+We adopt as a definition that a light cone is the region in spacetime within
+the surface generated by the light emanating from a given point in space in a
+given period of time.
+%
+For example, two nodes separated by 100 light years which appear with a
+difference in time of 80 years and last 50 years each, will be causally
+connected for 30 years.
+%
+This connection is not, though, bidirectional.
+%
+In fact, the second node will see the first one for 30 years, but the first
+node will never acknowledge the existence of the second.
+%
+This simple tought experiment exemplify the importance of the time variable to
+analyse the structure of connections.
 %
 The model can be described with three parameters, namely, the mean separation
-between the appearance of nodes, the mean lifetime of the nodes, and
-the maximum separation to allow an effective causal contact.
-%
-The model presented here comprises 
-the stochastic networks of constrained causally connected nodes,
-together with the three parameters and additional hypotheses.
+between the appearance of nodes, the mean lifetime of the nodes, and the
+maximum separation to allow an effective causal contact.
+%
+The model presented here comprises the stochastic networks of constrained
+causally connected nodes, together with the three parameters and additional
+hypotheses.
 %
@@
 
-We abandon the frequentist approach of the Drake equation to
-compute the number of civilisations, providing its statistical
-distribution instead.
-%
-This is an exploratory analysis that aims at developing a numerical
-tool to discuss the different scenarios based on statistical
-heuristics.
-%
-With this, we avoid the several theoretical problems
-summarised by the Drake equation factors.
-%
-The approach proposed here should be considered as a compromise
-between the uncertainties of the frequentist estimations and the detailed
-recipes required on the numerical simulations.
-%
-We provide a framework to explore, through a suite of
-numerical simulations, a parameter space of three unknown observables.
-%
-This allows to discuss possible scenarios and their consequences in
-terms of the probability of making contacts.
-%
-The method we use for simulating a stochastic process is an
-approximation that allows to study the behaviour of complex systems,
-by considering a sequence of well defined discrete events.
-%
-The simulation is carried out by following all the variables that
-describe and constitute the state of the system, and the evolution of
-the process is described as a set of changes in that state.
-%
-In this context, an event produces a specific change in the state that
-can be triggered by random variables that encode the stochastic nature
-of the physical phenomenon.
-%
-For example, when a new contact is produced between two entities in
-the simulated galaxy, the numbers of active communication lines and of
-communicated nodes does change.
-%
-The process involves following the changes on the state of the system,
-defining the initial and final states.
-%
-This is done by defining a method that allows to keep track of the
-time progress in steps, and maintaining a list of relevant events.
-%
-In Sec.~\ref{S_methods} we introduce the methods and discuss the
-candidate distributions for the statistical aspects of the times
-involved in the communication process.
-%
-We present our results in Sec~\ref{S_results}, with special emphasis
-on the statistical distributions of the duration of causal contacts in
-one or both directions and the distribution of time intervals for the
-waiting of the first contact.
+We provide a framework to explore, through a suite of numerical simulations,
+the parameter space of three unknown observables.
+%
+This allows to discuss possible scenarios and their consequences in terms of
+the probability of making contacts.
+%
+The method we use for simulating a stochastic process is an approximation that
+allows to study the behaviour of complex systems, by considering a sequence of
+well defined discrete events.
+%
+The simulation is carried out by following all the variables that describe and
+constitute the state of the system, and the evolution of the process is
+described as a set of changes in that state.
+%
+In this context, an event produces a specific change in the state that can be
+triggered by random variables that encode the stochastic nature of the physical
+phenomenon.
+%
+For example, when a new contact is produced between two entities in the
+simulated galaxy, the number of active communication lines is increased by one.
+%
+Also, if it is the first contact for that nodes, 
+then the number of communicated nodes increases by two.
+%
+When a new node becomes active, the sistem has an increase of one
+in the number of nodes, although the number of communcations does not
+necessarily change.
+%
+The process then involves following the changes on the state of the system, defining
+the initial and final states.
+%
+This is done by defining a method that allows to keep track of the time
+progress in steps and maintaining a list of relevant events, i.e., the events
+that produce a change in the vriables of interest.
+%
+With this method we abandon the frequentist approach of the Drake equation to
+compute the number of civilisations, providing instead its statistical
+distribution.
+%
+More importantly, we are interested on the probability of contacts, which
+depend critically on the time variable.
+%
+This is an exploratory analysis that aims at developing a numerical tool to
+discuss the different scenarios based on statistical heuristics.
+%
+The approach proposed here should be considered as a compromise between the
+uncertainties of the frequentist estimations and the detailed recipes required
+on the numerical simulations.
+%
+It is worth noticing that the model is not intended for a formal fit at this
+stage due to the lack of data, but is can help to understand how unusual it
+would be to actively search for intelligente signal for 50 years without
+success.
+
+This paper is organized as follows.
+%
+In Sec.~\ref{S_methods} we introduce the methods and discuss the candidate
+distributions for the statistical aspects of the times involved in the
+communication process.
+%
+We present our results in Sec~\ref{S_results}, with special emphasis on the
+statistical distributions of the duration of causal contacts in one or both
+directions and the distribution of time intervals for the waiting of the first
+contact.
 %
@@
 %
-In this work we use the International System of units recommended by
-the International Astronomial Union, where the simbol for years is
-''a''.
-%}}}
+In this work we use the International System of units recommended by the
+International Astronomial Union, where the simbol for years is ''a''.
 
 \begin{figure}[!t]
+   %
    \centering
+   %
    \includegraphics[width=0.5\textwidth]{F_scheme.pdf}
+   %
    \caption{
    %
-Space--time diagrams showing a schematic representation of the
-different stages in the development of causal contact node (\ceti{}).
+Space--time diagrams showing a schematic representation of the different stages
+in the development of a constrained causal contact node (\ceti{}).
 %
 Panel (a) represents the region $R$ in space and time which is causally
-connected to the emitter (left vertex), following an "A" type event
-("Awakening") in which the node acquires the communication capability.
-%
-The sphere of first contact (SFC) is a sphere centered on the emitter
-that grows until its radius reaches the D$_{\rm max}$ distance, in which
-the power of a signal would equal the detectability threshold.
-%
-In the Figure this sphere is represented by the left triangle of the
-region $R$.
-%
-The surface of last contact (SLC) is another sphere that grows from a
-"D" event ("Doomsday").
-%
-The region which is causally connected to the emitter is then limited
-by these two spheres, and has the shape of a sphere or of a spherical
-shell, depending on the time.
-%
-The temporal intervals for
-the communication between two \cetis{} are represented in he
-panel (b).
-%
-The receiver \ceti{} E$_1$ can listen signals from the emiter \ceti{} E$_2$,
-from the "Contact" event (t=C$_{12}$) up to the "Blackout" event (t=B$_{12}$).
-%
-Similarly, 
-the receiver \ceti{} E$_2$ can listen signals from the emiter \ceti{} E$_1$,
-from t=C$_{21}$ up to t=B$_{21}$. 
+connected to the emitter (E, left vertex), following an ``A'' type
+event (``Awakening'') in which the node acquires the communication
+capability.
+%
+The sphere of first contact (SFC) is the sphere centered on the emitter that
+grows until its radius reaches the D$_{\rm max}$ distance, in which the
+power of the signal would equal the detectability threshold.
+%
+In the Figure this sphere is represented by the left triangle of the region
+$R$.
+%
+The surface of last contact (SLC) is another sphere that grows from a ``D''
+event (``Doomsday'').
+%
+The region which is causally connected to the emitter is then limited by these
+two spheres, and has the shape of a sphere or of a spherical shell,
+depending on the time.
+%
+The temporal intervals for the communication between two \cetis{} are
+represented in the panel (b).
+%
+The receiver node E$_1$ can listen signals from the emitter node E$_2$, from
+the ``Contact'' event (t=C$_{12}$) up to the ``Blackout'' event
+(t=B$_{12}$).
+%
+Similarly, the receiver \ceti{} E$_2$ can listen signals from the emitter
+\ceti{} E$_1$, from t=C$_{21}$ up to t=B$_{21}$. 
 %
 } \label{F_scheme}
+%
 \end{figure}
-
-%}}}
 
@@
 
-Simulations are suitable tools to analyse systems that evolve with time and involve randomness.
-%
-An advantage of a numerical approach is that usually require fewer assumptions and simplifications, and can be applied to systems where analytical models are not easily found.
-%
-In particular, a suitable tool to model complex stochastic processes through the changes in the state of a system is the discrete--event (hereafter, DE) simulation approach.
-%
-A system described with the DE paradigm is characterised by a set of actors and events.
-%
-Actors interact causally through a series of events on a timeline and process them in chronological order \citep{ptolemaeus_system_2014, chung_simulation_2003, ross_simulation_2012}.
-%
-Each event modify the variables that define the process, producing the corresponding change in the state of the simulated system.
-%
-This method is well suited for the particular case of the diffusion of intelligent signals in the galaxy and allows to explore several models easily.
-%
-We simulate the statistical properties of a set of points in space and time that have a causal connection at light speed, dubbed ‘’Causal Contact Nodes’’ (hereafter, \ceti{}).
+Simulations are suitable tools to analyse systems that evolve with time and
+involve randomness.
+%
+An advantage of a numerical approach is that it usually requires fewer assumptions
+and simplifications, and can be applied to systems where analytical models are
+hard or impossible to develop.
+%
+In particular, a suitable tool to model complex stochastic processes through
+the changes in the state of a system is the discrete--event (hereafter, DE)
+simulation approach.
+%
+A system described with the DE paradigm is characterised by a set of actors and
+events.
+%
+Actors interact causally through a series of events on a timeline and process
+them in chronological order \citep{ptolemaeus_system_2014,
+chung_simulation_2003, ross_simulation_2012}.
+%
+Each event modify the variables that define the process, producing the
+corresponding change in the state of the simulated system.
+%
+This method is well suited for the particular case of the diffusion of
+intelligent signals in the galaxy and allows to explore several models easily.
+%
+We simulate the statistical properties of a set of points in space and time
+that have a causal connection at light speed and are separated
+by a maximum distance.
+%
+We refer to this nodes as ''Constrained Causal Contact Nodes``
+(hereafter, \ceti{}).
 %
 We chose this generic name in order to stress the fact that in this analysis 
-only the causal contact is considered, independently of any broadcasting or lookout activity.
-%
-The system we propose is ideal, in the sense that it considers the special case of a fully efficient node that emits and receives isotropically.
-%
-A causal contact node, considered as a broadcasting station that has the ability to detect signals through an active lookout program, could have a lesser than one efficiency factor, which is not included in the current analysis.
-%
-Also, it is worth mentioning that this is a general approach, and not necessarily a \ceti{} is the host of an intelligent civilization.
-%
-For instance, it can be associated with a planet where life has developed, became intelligent, reached the skills required to find the right communication channel, sustain a search and establish a contact.
-%
-Alternative message processing entities could be considered, for example interstellar beacons where life has ceased to exist but continue with its emission, or communication stations established by probes or left by intelligent beings \citep[see, e.g., ][]{peters_outer_2018, barlow_galactic_2013}.
-%
-In principle, these strategies could affect our results since it would be easier to configure a cluster of \cetis{} that spread in time.
+only the causal contact is considered, independently of any broadcasting or
+lookout activity.
+%
+The system we propose is ideal, in the sense that it considers the special case
+of a fully efficient node that emits and receives isotropically.
+%
+A causal contact node, considered as a broadcasting station that has the
+ability to detect signals through an active search program, could have a
+lesser than one efficiency factor, which is not included in the current
+analysis.
+%
+Also, it is worth mentioning that this is a general approach, and not
+necessarily a \ceti{} is the host of an intelligent civilization.
+%
+It can be associated with a planet where life has developed,
+became intelligent, reached the skills required to find the right communication
+channel, sustained a search and established a contact.
+%
+Alternative message processing entities could be considered, for example
+interstellar beacons where life has ceased to exist but continue with its
+emission, or communication stations established by probes or left by
+intelligent beings \citep[see, e.g., ][]{peters_outer_2018,
+barlow_galactic_2013}.
+%
+In principle, these strategies could affect our results since it would be
+easier to configure a cluster of \cetis{} that spread in time.
 %
@@
 %
-For the purpose of the present analysis, only the communication capability is relevant, since we study the causal contacts between the locations.
-%
-The system is defined by a number of actors that represent \cetis{} and appear at different instants in time, generating events that produce meaningful changes in the variables that describe the system, i.e., in the arrangement of \cetis{} and their network of causal contacts.
-%
-For example, the appearance of a new \ceti{} in a region filled with a signal emitted by another node, will increase the number of active \cetis{} and the number of pairs of \cetis{} in causal contact.
-%
-Assuming some simple hypotheses, the discrete events method can be performed taking into account a small number of variables, which allow to analyze the variation of the results in the model parameter space.
-
-In what follows, we outline the experimental setup adopted to estimate the probabilities of causal contacts and several derived quantities.
-%
-This is done in terms of three independent parameters, namely, the
-mean time span between the appearance of consecutive \cetis{},
-$\left<\tau_a\right>$, their mean lifetime $\left<\tau_s\right>$, and
-the maximum distance a signal can be detected by another \ceti{}
-($D_{\rm max}$).
-%
-Intuitively, the probability of the existence of causal contacts
-between pairs of \cetis{} would be larger for smaller $\tau_a$
-parameters, higher $\tau_s$ or higher $D_{\rm max}$ parameter values.
-%
-We also propose theoretical distribution functions for both the lifespans ($\tau_s$) and the number of \cetis{} per unit time \citep{maccone_evolution_2014, Sotos_biotechnology_2019}.
-%
-The later is related to the time span between the appearance of consecutive \cetis{} (since when $\left<\tau_a\right>$ is shorter, it produces a greater density of \cetis{}).
-%
-The analytical expresiones for these distributions are set to a fixed law, as discussed in Sec.~\ref{SS_PDF_shape}.
+For the purpose of the present analysis, only the communication capability is
+relevant, since we study the causal contacts between the locations.
+%
+The system is defined by a number of actors that represent \cetis{} and appear
+at different instants in time, generating events that produce meaningful
+changes in the variables that describe the system, i.e., in the arrangement of
+\cetis{} and their network of causal contacts.
+%
+For example, the appearance of a new \ceti{} in a region filled with a signal
+emitted by another node, will increase the number of active \cetis{} and the
+number of pairs of \cetis{} in causal contact.
+%
+Assuming some simple hypotheses, the discrete events method can be performed
+taking into account a small number of variables, which allow to analyze the
+variation of the results in the model parameter space.
+
+In what follows, we outline the experimental setup adopted to estimate the
+probabilities of causal contacts and several derived quantities.
+%
+This is done in terms of three independent parameters, namely, the mean time
+span between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$,
+their mean lifetime $\left<\tau_s\right>$, and the maximum distance a signal
+can be detected by another \ceti{} ($D_{\rm max}$).
+%
+Intuitively, the probability of the existence of causal contacts between pairs
+of \cetis{} would be larger for smaller $\tau_a$ parameters, higher $\tau_s$ or
+higher $D_{\rm max}$ parameter values.
+%
+We also propose theoretical distribution functions for both the lifespans
+($\tau_s$) and the number of \cetis{} per unit time
+\citep{maccone_evolution_2014, Sotos_biotechnology_2019}.
+%
+The later is related to the time span between the appearance of consecutive
+\cetis{} (since when $\left<\tau_a\right>$ is shorter, it produces a greater
+density of \cetis{}).
+%
+The analytical expresions for these distributions are set to a fixed law, as
+discussed in Sec.~\ref{SS_PDF_shape}.
 
@@
 %
-Emitter $i$ can listen to emiter $j$ between C$_{ij}$ and B$_{ij}$.
+Emitter $i$ can listen to emitter $j$ between C$_{ij}$ and B$_{ij}$.
 %

### 2020-05-26 submit_v03/manuscript.tex
segunda lectura

@@
 %
-Similarly, the receiver \ceti{} E$_2$ can listen signals from the emitter
-\ceti{} E$_1$, from t=C$_{21}$ up to t=B$_{21}$. 
+Similarly, the receiver node E$_2$ can listen signals from the emitter
+node E$_1$, from t=C$_{21}$ up to t=B$_{21}$. 
 %
@@
 The points in time where the \cetis{} acquire their communicating
-capacity are dubbed ‘’Awakening’’ events ($A_1$ and $A_2$ in the
+capacity are dubbed ``Awakening'' events ($A_1$ and $A_2$ in the
 Figure).
@@
 The points in space--time where the first contact is produced for each
-one of the \cetis{}, are defined as ''Contact'' events, shown as $C_1$
+one of the \cetis{}, are defined as ``Contact'' events, shown as $C_1$
 and $C_2$.
@@
 Finally, the two points where the contact is lost for each one the
-\cetis{}, are denominated ''Blackout'' events, presented as $B_1$ and
+\cetis{}, are denominated ``Blackout'' events, presented as $B_1$ and
 $B_2$ in the Figure.
@@
 The time intervals for the open communication channel are determined
-by the ''C'' and ''B'' type events.
+by the ``C'' and ``B'' type events.
 %
@@
 
-The temporal structure that emerges from the experimental setup implies that, as a fundamental property of the simulations, a causal connection can be produced without 
-requiring that the two nodes are active at the same time.
-%
-This property of the system arises as a consequence of the large spatial and temporal scales, where a message is transmitted at the (relatively small) light speed.
-%
-Although a \ceti{} could be active for a large enough period to transmit a message at large distances, the limited power of the message and the dilution that depends on the squared distance from the source imposes a detectability limit.
-%
-As a consequence of this limitation and of their finite lifetime, considered as the period between the acquisition and loss of communicating capacity, each \ceti{} will fill a spherical shell region of the galaxy, limited by two concentric spherical surfaces.
-%
-The leading front, or surface of first contact (SFC) grows from the
-central \ceti{} until it reaches the maximum distance $D_{\rm max}$.
-%
-Following the end of the civilization, there is still a region which is filled with the emitted signals.
-%
-This approach has been also considered in other statistical models \citep[e.g., ][]{smith_broadcasting_2009, grimaldi_signal_2017, Grimaldi2018}.
-%
-The trailing front, or surface of last contact (SLC) also grows from the central \ceti{}, with a delay with respect to the SFC equivalent to the lifetime of the node, and produces a spherical shell region.
-%
-Any other \ceti{} within this region will be aware of the originating node, even if it has disappeared before the time of contact.
-%
-This region will grow if the surface of first contact has not yet
-reached the maximum distance $D_{\rm max}$, and will shrink after a
-D-type event until the surface of last contact reaches $D_{\rm max}$, producing as a result the loss of all signal from the central \ceti{}.
-%
-In our approach, we consider a model galaxy where the width of the disk is negligible with respect to the radius of the disk.
-%
-In the 2D simulation only the intersection of the communicating spherical shells with the plane of the galaxy is relevant, and produce the corresponding circles or rings for the filled spheres or annular regions of the spherical shells, respectively.
-%
-The initially growing communicating sphere is shown over space--time diagrams, where time is represented on the horizontal direction, and space is represented in the vertical direction.
-%
-In the Fig.~\ref{F_scheme} the two emitters in panel (b), E$_1$ and E$_2$, reach each other at different times.
+The temporal structure that emerges from the experimental setup implies that,
+as a fundamental property of the simulations, a causal connection can be
+produced without  requiring that the two nodes are active at the same time.
+%
+This property of the system arises as a consequence of the large spatial and
+temporal scales, where a message is transmitted at the (relatively small) light
+speed.
+%
+Although a \ceti{} could be active for a large enough period to transmit a
+message at large distances, the limited power of the message and the dilution
+that depends on the squared distance from the source imposes a detectability
+limit.
+%
+As a consequence of this limitation and of their finite lifetime, considered as
+the period between the acquisition and loss of communicating capacity, each
+\ceti{} will fill a spherical shell region of the galaxy, limited by two
+concentric spherical surfaces.
+%
+The leading front, or surface of first contact (SFC) grows from the central
+\ceti{} until it reaches the maximum distance $D_{\rm max}$.
+%
+Following the end of the civilization, there is still a region which is filled
+with the emitted signals.
+%
+This approach has been also considered in other statistical models \citep[e.g.,
+][]{smith_broadcasting_2009, grimaldi_signal_2017, Grimaldi2018}.
+%
+The trailing front, or surface of last contact (SLC) also grows from the
+central \ceti{}, with a delay with respect to the SFC equivalent to the
+lifetime of the node, and produces a spherical shell region.
+%
+Any other \ceti{} within this region will be aware of the originating node,
+even if it has disappeared before the time of contact.
+%
+This region will grow if the surface of first contact has not yet reached the
+maximum distance $D_{\rm max}$, and will shrink after a D-type event until the
+surface of last contact reaches $D_{\rm max}$, producing as a result the loss
+of all signal from the central \ceti{}.
+%
+In our approach, we consider a model galaxy where the width of the disk is
+negligible with respect to the radius of the disk.
+%
+In the 2D simulation only the intersection of the communicating spherical
+shells with the plane of the galaxy is relevant, and produce the corresponding
+circles or rings for the filled spheres or annular regions of the spherical
+shells, respectively.
+%
+The initially growing communicating sphere is shown over space--time diagrams,
+where time is represented on the horizontal direction, and space is represented
+in the vertical direction.
+%
+In the Fig.~\ref{F_scheme} the two emitters in panel (b), E$_1$ and E$_2$,
+reach each other at different times.
 %
@@
 %
-The type and length of causal contact in both directions depend on the distance and time lag between the awakening events, the maximum distance that a signal can reach and the time period in which each emitter is active.
-
-In our experimental configuration the simulation starts assuming the stochastic process is already stable, and finishes before any galactic evolution effect could modify the fixed values of the variables.
-%
-Likewise, we assume that the probability for the appearance of a \ceti{} is homogeneous over the GHZ.
-%
-Although the Galaxy has a well-known spiral structure, the assumed sparsity of nodes (arguably, intelligent civilisations) does make that assumption tenable.
-%
-If the distribution of nodes is not sparse, as it could be the case if the spiral arms host most of the \cetis{}, then contacts would be more frequent between closely located nodes.
-%
-In such a case our results would underestimate the number of contacts between close pairs of nodes within the same spiral arm, and conversely, overestimate the number of contacts between separated nodes.
-%
-We also limit the possibilities of life or other types of civilisations to the usually stated hypothesis for the definition of
-the GHZ \citep{dayal_habitability_2016, gonzalez_galactic_2001, lineweaver_galactic_2004, gonzalez_habitable_2005,
-morrison_extending_2015, haqq-misra_evolution_2019, rahvar_cosmic_2016, gobat_evolution_2016, rahvar_cosmic_2016} and consider that habitability remains constant over time \citep[see, however, ][]{gonzalez_habitable_2005, dayal_habitability_2016, gobat_evolution_2016}.
-%
-For example, \citet{vukotic_astrobiological_2012} propose
-probabilistic cellular automata model to explain the astrobiological
-history of the Milky Way. % <<<< FEO
-%
-This means that we set aside civilisations that could survive in severe conditions or unstable systems, which would prevent the appearance of life as we know it.
-
-We need to raise the scope of our results by stressing the fact that we are considering causal contacts instead of actual contacts.
-%
-By a causal contact we mean a configuration, in space and time, in which a node belongs to the light cone of another node.
-%
-This means that a causal contact does not need to be tied to any actual message emission or detection.
-%
-In more realistic scenarios, there are several sources of ‘’signal loss’’ with respect to the ideal case.
-%
-Among them, we can mention temporal and signal power aspects and direction dependent communication capabilties.
-%
-The results must then be interpreted as the case of ideal nodes, with a perfect efficiency in the emision and reception of signals.
-%
-The probabilities of contacts presented here are then upper limits to the number of communications between civilization s in the Galaxy, given the implemented hipotheses in the model.
-%
-The use of light cones as causal contact regions is inspired by the fact that light—speed traveling messengers like 
-electromagnetic radiation, gravitational radiation or neutrinos are often considered as possible message carriers \citep{hippke_interstellar_2017, wright_how_2018}.
-%
-This excludes messages sent with mechanical means or physical objects \citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown technology that violates the known laws of physics.
-
-As part of this benchmark, we assume that the capacity to emit and receive signals occur at the same time.
-%
-Although there are several reasons to think that this could not be the case, at large time scales it can be considered that both abilities occur roughly at coincident epochs.
-%
-In the ideal setup, this would be equivalent to nodes that send messages isotropically and scan the local skies on all directions with a perfect eficiency.
-%
-Another essential assumption is that all \cetis{} use the same signal power, so that there is a maximum distance out to which it can be detected.
-%
-% It is worth mentioning that we are considering in our experimental setup a system composed by several emitters and 
-% receivers across the galaxy, under the hypothesis proposed before.
-%
-In such a system we compute probabilities of a \ceti{} making contact with another node, and the results must be interpreted as for a randomly chosen \ceti{}, given certain conditions of homogeneity on an ensemble of equivalent galaxies.
-%
-Then, we stress the fact that we are not centred on the probabilities for the Earth itself, but on a generalised \ceti{}. 
-%
-Regarding the extent of the signals, we know that the distance from which a signal from Earth could be detected using the current technology is of a few hundreds light years, given that the signal was sent to a specific direction.
-%
-It is straightforward to propose and implement a distribution of maximum distances, although this would increase the model complexity at the cost of a larger uncertainty or of another dimension on the hypothesis space.
+The type and length of causal contact in both directions depend on the distance
+and time lag between the awakening events, the maximum distance that a signal
+can reach and the time period in which each emitter is active.
+
+In our experimental configuration the simulation starts assuming that the
+stochastic process is already stable, and finishes before any galactic
+evolution effect could modify the fixed values of the variables.
+%
+Likewise, we assume that the probability for the appearance of a \ceti{} is
+homogeneous over the GHZ.
+%
+Although the Galaxy has a well-known spiral structure, the nodes are assumed
+to be sparse (otherwise the Galaxy would be full of life) and the spiral
+structure would not, in that case, produce significant differences.
+%
+If the distribution of nodes is not sparse, as it could be the case if the
+spiral arms host most of the \cetis{}, then contacts would be more frequent
+between closely located nodes.
+%
+In such a case our results underestimate the number of contacts between
+close pairs of nodes within the same spiral arm, and conversely, overestimate
+the number of contacts between separated nodes.
+%
+We also limit the possibilities of life or other types of civilisations to the
+usually stated hypotheses for the definition of the GHZ
+\citep{dayal_habitability_2016, gonzalez_galactic_2001,
+lineweaver_galactic_2004, gonzalez_habitable_2005, morrison_extending_2015,
+haqq-misra_evolution_2019, rahvar_cosmic_2016, gobat_evolution_2016,
+rahvar_cosmic_2016} and consider that habitability remains constant over time
+\citep[see, however, ][]{gonzalez_habitable_2005, dayal_habitability_2016,
+gobat_evolution_2016}.
+%
+This means that we set aside civilisations that could survive in severe
+conditions or unstable systems, which would prevent the appearance of life as
+we know it.
+
+We stress the fact that we are
+considering causal contacts instead of actual contacts.
+%
+In more realistic scenarios, there are several sources of ``signal loss'' with
+respect to the ideal case.
+%
+Among them, we can mention temporal and signal power aspects and direction
+dependent communication capabilities.
+%
+The results must then be interpreted as the case of ideal nodes, with a perfect
+efficiency in the emision and reception of signals.
+%
+The probabilities of contacts presented here are then upper limits to the
+number of communications between civilizations in the Galaxy, given the
+implemented hipotheses in the model.
+%
+The use of light cones as causal contact regions is inspired by the fact that
+light-speed traveling messengers like electromagnetic radiation, gravitational
+radiation or neutrinos are often considered as possible message carriers
+\citep{hippke_interstellar_2017, wright_how_2018}.
+%
+This excludes messages sent with mechanical means or physical objects
+\citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown
+technology that violates the known laws of physics.
+
+As part of this benchmark, we assume that the capacity to emit and receive
+signals occur at the same time.
+%
+Although there are several reasons to think that this could not be the case, at
+large time scales it can be considered that both abilities occur roughly at
+coincident epochs.
+%
+In the ideal setup, this would be equivalent to nodes that send messages
+isotropically and scan the local skies on all directions with a perfect
+eficiency.
+%
+Another essential assumption is that all \cetis{} use the same signal power, so
+that there is a maximum distance out to which it can be detected.
+%
+% It is worth mentioning that we are considering in our experimental setup a
+% system composed by several emitters and  receivers across the galaxy, under
+% the hypothesis proposed before.
+%
+In such a system we compute probabilities of a random node making contact with
+another node, i.e., they are not specific for the case of contacts with the
+Earth.
+%
+Regarding the extent of the signals, we know that the distance from which a
+signal from Earth could be detected using the current technology is of a few
+parsecs, given that the signal was sent to a specific direction.
+%
+It is straightforward to propose and implement a distribution of maximum
+distances, although this would increase the model complexity at the cost of a
+larger uncertainty.
 
@@
 In our simulations, we assume the simplest configuration for the
-growth of the sphere of first contact. For example, we discard the
+growth of the sphere of first contact.
+%
+In particular, we do not consider the
 possibility of stellar colonisation
@@
 starling_virulence_2013, barlow_galactic_2012, jeong_large_2000,
-maccone_mathematical_2011}. We also assume that communication is
+maccone_mathematical_2011}.
+%
+We also assume that communication is
 equally likely in all directions, i.e., we assume isotropic
-communication in all cases. Different communication efficiencies or
+communication in all cases.
+%
+Different communication efficiencies or
 detection methods are straightforward to carry out in the simulations
-for more detailed and complex scenarios. This approach, however, is
+for more detailed and complex scenarios.
+%
+This approach, however, is
 beyond this work because it obscures the experimental setup and make
-the results less clear. The assumptions we accept imply that the
+the results less clear.
+%
+The assumptions we accept imply that the
 results are independent of whether intelligent agents are organic or
@@
 imply extreme values for the parameters that represent the density of
-nodes.
+the nodes (i.e., $\left< \tau_a \right>$).
 
@@
 
-The temporal structure of the process is defined by two distribution parameters. One of them represent the mean time
-interval that a \ceti{} can emit and receive signals, and the other the mean time interval between the emergence of
-consecutive \cetis{}. The spatial structure of the simulation is given by the size and shape of the Galactic Habitable
-Zone and the maximum distance a signal can travel to be detected
-(D$_{\rm max}$). The parameters for the temporal
-distributions also determine the spatial properties, since the density of active \cetis{} in the galaxy depend on these
-two parameters. For example, a small $\tau_a$ and a large $\tau_s$ will produce a galaxy densely populated with
-\cetis{}. Also, some hypotheses regarding the shapes of the distributions of the temporal parameters must be made in
-order to complete the simulation. \citet{forgan_spatiotemporal_2011} argues that the times at which different
-civilisations become intelligent follow a Gaussian distribution, and then the distribution of inter-arrival times is an
-inverse exponential. We assume that the distribution of the times of A events is a stationary Poisson process, and then
-the distribution of the times between the appearance of new consecutive \cetis{} is exponential. Regarding the duration
-of a \ceti{}, we propose that the distribution of the duration of \cetis{} is a stationary exponential distribution.
-There are no clear arguments to conclude a statistical law for the later distribution, so that we propose it as a
-working hipothesis.  A theoretical statistical distribution of the lifespan of civilisations would rely preferentially
-on the basis of the underlying astrophysical and biological processes \citep{balbi_impact_2018}. For example,
-\citet{maccone_lognormals_2014} argues that this distribution should be a log normal.
-
-The power law and exponential statistical distributions are among the most common patterns found in natural phenomena.
-%
-For example, the distribution of the frequency of words in many languages is known to follow a Zipf’s law (which is a 
-power law).
-%
-These distributions arise any time a phenomenon is characterised by commonly occurring small events and rarely produced
-large events \citep[e.g. ][]{adamic_zipf_2000}.
-%
-Zipf law also describes population ranks of cities in various countries, corporation sizes, income rankings, ranks of
-the number of people watching the same TV channel, etc.
-%
-The magnitudes of earthquakes, hurricanes, volcanic eruptions and floods; the sizes of meteorites or the losses caused
-by business interruptions from accidents, are also well described by power laws
+The temporal structure of the process is defined by two distribution
+parameters. One of them represent the mean time interval that a node can
+emit and receive signals (its lifetime), and the other the mean time interval between the
+emergence of consecutive \cetis{}.
+%
+The spatial structure of the simulation is
+given by the size and shape of the Galactic Habitable Zone and the maximum
+distance a signal can travel to be detected (D$_{\rm max}$). The parameters for
+the temporal distributions also determine the spatial properties, since the
+density of active \cetis{} in the galaxy depend on these two parameters. For
+example, small $\tau_a$ and a large $\tau_s$ will produce a densely
+populated galaxy (in this context, a galaxy is an element in a 
+statistical ensemble, not the Milky Way).
+%
+Also, some hypotheses regarding the shapes of the
+distributions of the temporal parameters must be made in order to complete the
+simulation. \citet{forgan_spatiotemporal_2011} argues that the times at which
+different civilisations become intelligent follow a Gaussian distribution, and
+then the distribution of inter-arrival times is an inverse exponential. We
+assume that the distribution of the times of A events is a stationary Poisson
+process, and then the distribution of the times between the appearance of new
+consecutive nodes is exponential. Regarding the duration of a node, we
+propose that its distribution is a stationary
+exponential distribution.  There are no clear arguments to conclude a
+statistical law for the later distribution, so that we propose it as a working
+hipothesis.  
+%
+This heuristic does not make any consideration about the origin of life, although
+different approaches are possible.
+%
+For example,
+\citet{maccone_lognormals_2014} argues that this distribution should be a log-normal. 
+%
+Preferentially, a theoretical statistical distribution of the lifespan of
+civilisations would rely on the basis of the underlying
+astrophysical and biological processes \citep{balbi_impact_2018}.
+
+The power law and exponential statistical distributions are among the most
+common patterns found in natural phenomena.
+%
+For example, the distribution of the frequency of words in many languages is
+known to follow the law of Zipf (which is a power law).
+%
+These distributions arise any time a phenomenon is characterised by commonly
+occurring small events and rarely produced large events \citep[e.g.
+][]{adamic_zipf_2000}.
+%
+Zipf law also describes population ranks of cities in various countries,
+corporation sizes, income rankings, ranks of the number of people watching the
+same TV channel, etc.
+%
+The magnitudes of earthquakes, hurricanes, volcanic eruptions and floods; the
+sizes of meteorites or the losses caused by business interruptions from
+accidents, are also well described by power laws
 \citep{sornette_critical_2006}.
@@
 %% , benguigui_classificacion_2016}.
-The power law behaviour has been observed in a variety of systems.
-%
-These include, for example, stock market fluctuations, sizes of computer files or word frequency in languages
-\citep{mitzenmacher_brief_2004, newman_power_2005, simkin_theory_2006}. 
-%
-Power laws have also been widely used in
-biological sciences.
-%
-Some examples are the analysis of connectivity patterns in metabolic networks \citep{jeong_large_2000} or the number of
-species observed per unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
-%
-More examples can be found in the literature 
-\citep{martin_origin_2006, maccone_KLT_2010}
-\citep{barabasi_scale_2009, maccone_evolution_2014}
+%The power law behaviour has been observed in a variety of systems.
+%
+Additional examples include stock market fluctuations, sizes of computer files
+or word frequency in languages \citep{mitzenmacher_brief_2004,
+newman_power_2005, simkin_theory_2006}. 
+%
+Power laws have also been widely used in biological sciences.
+%
+Some examples are the analysis of connectivity patterns in metabolic networks
+\citep{jeong_large_2000} or the number of species observed per unit area in
+ecology \citep{martin_origin_2006, frank_common_2009}.
+%
+More examples can be found in the literature \citep{martin_origin_2006,
+maccone_KLT_2010} \citep{barabasi_scale_2009, maccone_evolution_2014}
 \citep{maccone_lognormals_2014}
-
-In this work, the exponential distribution of lifespan and waiting times is justified by considering the hypothesis that the process of appearance of life in the galaxy is homogeneous and stationary.
-That is, there is not a preferred region within the GHZ for the spontaneous appearance of life, and the emergence of a \ceti{} is independent of the existence of previous nodes in the galaxy.
-This is equivalent to proposing a Poisson process for the emergence of \cetis{}, given the relation between the number of events in time or space and the waiting time or separation, respectively
-\citep[e.g., ][]{ross_simulation_2012}.
-That is, these are two alternative approaches to describing the same process, a Poisson distribution for the number of events implies an exponential distribution for their separations, and vice versa.
-It should be emphasised that the exponential laws used in this work are assumed as part of the working hypothesis, and instead of analysing results from a particular parameter chosen ad hoc, we explore the hypothesis space and analyse the impact of the values of these parameters on the results.
+%
+This distribution family is suitable for the statistical description
+of the $D_{\rm mar}$ parameter, although in this work we assume a uniform
+distribution for simplicity.
+%
+The power law family of functions is also a good candidate for the description of
+the temporal variables in the model.
+%
+However, we prefer the exponential model.
+%
+The exponential distribution of lifespan and waiting times is justified by
+considering the hypothesis that the process of appearance of life in the galaxy
+is homogeneous and stationary.
+%
+That is, there is not a preferred region within
+the GHZ for the spontaneous appearance of life, and the emergence of a node is
+independent of the existence of previous nodes in the galaxy.
+%
+These seem to be simple conditions, and allow to propose a distribution family
+without knowing the details of the underlying process.
+%
+The exponential distribution for the separations in time is equivalent to
+proposing a Poisson process for the emergence of nodes, given the relation
+between the number of events in time or space and the waiting time or
+separation, respectively \citep[e.g., ][]{ross_simulation_2012}.
+%
+That is, these
+are two alternative approaches to describing the same process, a Poisson
+distribution for the number of events implies an exponential distribution for
+their separations, and vice versa.
+%
+It should be emphasised that the exponential
+laws used in this work are assumed as part of the working hypothesis, and
+instead of analysing results from a particular parameter chosen ad hoc, we
+explore the hypothesis space and analyse the impact of the values of these
+parameters on the results.
 
@@
 
-In this Section we discuss the degree of complexity in the model, considering a compromise between the accuracy of the
-model and the number of parameters that are free or with a high uncertainty.
-%
-Firstly, we emphasise that the odds of a causal contact between two \cetis{} should not be considered as the odds of a
-contact between two intelligent civilisations, and in fact the latter could be much lesser than the former.
-%
-Indeed, in order to establish a contact between any two entities, a minimum degree of compatibility must be accomplished
-without any previous agreement, making the possibility of a contact with a message that could be deciphered highly rare
-\citep[see e.g. ][]{forgan_collimated_2014}.
-%
-Besides the trade--off between the simplicity and the complexity of the experiments, further analysis could be performed
-following this framework in order to explore possible implications of the results for more detailed configurations.
-%
-For example, the communication method (isotropic, collimated, serendipitous) can affect the observables, making it
-necessary to implement a correction factor.
-%
-Taking this into account, our results regarding the probability of causal contact should be considered as upper limits
-for effective contacts, since they depend on the efficiency of both the emitter and the receiver to broadcast and scan
-the sky for intelligent signals, respectively \citep{grimaldi_signal_2017}.
-%
-A correction by coverage ratio in the detection and by a targeting ratio in the emission could be easily implemented in
-the simulation, although the effect of reducing the probability of contact is basically the product of the efficiency
+In this Section we discuss the degree of complexity in the model, considering a
+compromise between the accuracy of the model and the number of parameters that
+are free or with a high uncertainty.
+%
+Firstly, we emphasise that the odds of a causal contact between two nodes
+should not be considered as the odds of a contact between two intelligent
+civilisations, and in fact the latter could be much lesser than the former.
+%
+Indeed, in order to establish a contact between any two entities, a minimum
+degree of compatibility must be accomplished without any previous agreement,
+making the possibility of a contact with a message that could be deciphered
+highly rare \citep[see e.g. ][]{forgan_collimated_2014}.
+%
+Besides the trade--off between the simplicity and the complexity of the
+experiments, further analysis could be performed following this framework in
+order to explore possible implications of the results for more detailed
+configurations.
+%
+For example, the communication method (isotropic, collimated, serendipitous)
+can affect the observables, making it necessary to implement a correction
+factor.
+%
+Taking this into account, our results regarding the probability of causal
+contact should be considered as upper limits for effective contacts, since they
+depend on the efficiency of both the emitter and the receiver to broadcast and
+scan the sky for intelligent signals, respectively
+\citep{grimaldi_signal_2017}.
+%
+A correction by coverage ratio in the detection and by a targeting ratio in the
+emission could be easily implemented in the simulation, although the effect of
+reducing the probability of contact is basically the product of the efficiency
 ratios and thus such implementation is not necessary.
 %
-Therefore, the values of the probabilities could be modified by a constant correction factor equal to the combined
-emission/reception efficiency \citep{smith_broadcasting_2009, anchordoqui_upper_2019, forgan_collimated_2014} or for
-beam-like transmissions \citep{grimaldi_signal_2017}. 
-%
-Other considerations include the effects of alignments, the use of stars as sources or amplifiers \citep{Edmondson2003,
-borra_searching_2012}, the nature of the message carrier.
-%
-Another improvement could be the use of a spatial distribution that resembles the spiral shape and the width of the disc
-of the Galaxy.
-%
-Regarding the distributions of the parameters, different distribution functions for the mean lifetime of \cetis{} can be
-implemented as alternatives.
-%
-Finally, regarding the degree of idealisation, the model admits different efficiencies for the sphere of the causal
-region of each \ceti{}, featuring different searching strategies \citep{hippke_interstellar_2017}. 
-%
-It is also possible to consider that $D_{\rm max}$ is different for different \cetis{}.
-%
-For example, a power law where powerful emissions are rare and low emissions are common could be an improvement to the
-model.
+Therefore, the values of the probabilities could be modified by a constant
+correction factor equal to the combined emission/reception efficiency
+\citep{smith_broadcasting_2009, anchordoqui_upper_2019, forgan_collimated_2014}
+or for beam-like transmissions \citep{grimaldi_signal_2017}. 
+%
+Other considerations include the effects of alignments, the use of stars as
+sources or amplifiers \citep{Edmondson2003, borra_searching_2012}, or the
+nature of the message carrier.
+%
+Another improvement could be the use of a spatial distribution that resembles
+the spiral shape and the width of the disc of the Galaxy.
+%
+Regarding the distributions of the parameters, different distribution functions
+for the mean lifetime of \cetis{} can be implemented as alternatives.
+%
+Finally, regarding the degree of idealisation, the model admits different
+efficiencies for the sphere of the causal region of each \ceti{}, featuring
+different searching strategies \citep{hippke_interstellar_2017}. 
+%
+It is also possible to consider that $D_{\rm max}$ is different for different
+\cetis{}.
+%
+For example, a power law where powerful emissions are rare and low emissions
+are common could be an improvement to the model.
 %
@@
 %
-Finally, the role of the message contents could influence on the lifespan of a \ceti{} that receives a message, although
-the implementation of such behaviour would increase the number of free parameters and would be more speculative.
-
-%}}}
+Finally, the role of the message contents could influence on the lifespan of a
+\ceti{} that receives a message, although the implementation of such behaviour
+would increase the number of free parameters and would be more speculative.   
+%
+We limit the scope of this work to a simple version of the model. 
+%
+Once the model has been defined, it can be implemented as a discrete event
+simulation, as described in detail in the next Section.
 
@@
 
-A discrete event simulation is performed for a given model $M(\tau_a,
-\tau_s, D_{\rm max})$ by keeping track of a set of variables that change each time an event happens. 
-%
-In addition, the model comprises elements that are fixed for all simulations, for example, the functional forms of the statistical distributions and the adopted values of particular variables (see Table XXX).
-%
-The main variables that follow the evolution of the simulation are: the positions of stars, which are sampled randomly within the GHZ; the time of the awakening of each \ceti{} (A event); and the time of the vanishing of each \ceti{} (D event).
-%
-The variables that can be deduced from the previous ones include: the number of \cetis{} in casual contact with at least another \ceti{} at a given time; the number of \cetis{} as a function of time; the number of \cetis{} that receive a message at least one time; the number of \cetis{} that receive a message at least one time and successfully deliver an answer; and the number distribution of waiting time to receive a message.
-%
-All these quantities are updated each time one of the four events (A, B, C, D) occurs.
+A discrete event simulation is performed for a given model $M(\tau_a, \tau_s,
+D_{\rm max})$ by keeping track of a set of variables that change each time an
+event happens. 
+%
+In addition, the model comprises elements that are fixed for all simulations,
+for example, the functional forms of the statistical distributions and the
+adopted values of particular variables (see \ref{T_simu_hypotheses}).
+%
+The main variables that follow the evolution of the simulation are:
+the positions of stars, which are sampled randomly within the GHZ; the time of
+the awakening of each node (A event); and its time of disappearance (D event).
+%
+The variables that can be deduced from the previous ones include: the number of
+\cetis{} in casual contact with at least another \ceti{} at a given time; the
+number of \cetis{} as a function of time; the number of \cetis{} that receive at
+least one message; the number of \cetis{} that receive a message at
+least one time and successfully deliver an answer; and the number distribution
+of waiting times to receive a message.
+%
+All these quantities are updated each time one of the four events (A, B, C, D)
+occurs.
 
@@
 
-Since the method is based on events, the first step is to define the relationships between them and how events trigger changes on the state of the system.
-%
-Once an event is produced, the system is updated by saving the corresponding change in the related variables.
-%
-For example, a B-type event diminishes the number of active \cetis{} by one, and also could affect the number of ongoing communications.
-%
-We consider a first approach to the galactic habitable zone as a 2-dimensional annular region.
-%
-The adopted values for the geometry of the GHZ are 2 kpc and 8 kpc, for the inner and outer radius, respectively \citep{lineweaver_galactic_2004}. 
-% ojo, ese paper dice 7-9 kpc, o sea tiene un limite superior mucho
-% mas chico...
-%
-This simple model does not take into account the variations in stellar density given by the spiral structure of the Galaxy.
-%
-The main task in setting a DE simulation is to define the events and what changes do these events produce on the system.
-%
-Although there are several possible approaches, we chose to follow the evolution of the system according to the following events:
+Since the method is based on events, the first step is to define the
+relationships between them and how events trigger changes on the state of the
+system.
+%
+Once an event is produced, the system is updated by saving the corresponding
+change in the related variables.
+%
+For example, a B-type event diminishes the number of active \cetis{} by one,
+and also could affect the number of ongoing communications.
+%
+We consider a first approach to the galactic habitable zone as a 2-dimensional
+annular region.
+%
+The adopted values for the geometry of the GHZ are 7~kpc and 9~kpc, for the
+inner and outer radius, respectively \citep{lineweaver_galactic_2004}. 
+%
+This simple model does not take into account the variations in stellar density
+given by the spiral structure of the Galaxy.
+%
+The main task in setting a DE simulation is to define the events and what
+changes do these events produce on the system.
+%
+Although there are several possible approaches, we chose to follow the
+evolution of the system according to the following events:
 
@@
 The time marks for each event (A and D for each \ceti{} and C and B
-for each causal contact) are stored as a result of the simulation.
+for each constrained causal contact) are stored as a result of the simulation.
 %
@@
 In order to make a reproducible project, we developed the tools
-that allow to run and obtain the results shown in this work.
-%
-The simulations were implemented on a Python-3 code, dubbed \hs{},
+that allow to run the simulations and obtain the results shown in this work.
+%
+The simulations were implemented on a Python-3 code, dubbed \hs{} (Lares et al., in preparation),
 which is publicly accessible through the GitHub
@@
 documentation, which is automatically generated from \hs{} docstrings
-and made public in the read-the-docs service.
+and made public in the read-the-docs service
+\footnote{https://hearsay.readthedocs.io/en/latest/}.
 
@@
 %
-Regarding the quality assurance of the software, we endorse the PEP-8
-style guide for python (van Rossum et al. 2001) throughout the entire
-project, which improves the code maintainability and helps to minimise
-the so called ''code-entropy'' of future versions.
-%
-The deviations in style were analysed with the flake8 tool, which
-improves the readability and consistency of the code.
-%
-Testing and code coverage have been applied to the package and the
-development process as standard qualitative and quantitative metrics.
-%
-The different parts and functions can be validated to assert if they
-return the expected results on specific test cases and testing
-scenarios \citep{jazayeri_engeneering_2007}.
-%
-On the other hand, code coverage measures the fraction visited and
-executed in the code by the testing procedure
-\citep{miller_testing_1963}.
-%
-The advantage of an exhaustive code-coverage is that it ensures that
-most of the sources of errors have been considered.
-%
-In the \hs{} project we provide test tools for the distinct parts of
-the simulation setup and run, reaching a 99\% of code coverage in the
-version used for this work.
-%
-For every new version of the code, all these quality assurance methods
-and metrics are run automatically by a continuous-integration service.
-%
-It is possible to reproduce the experiments performed in this work
-using the \hs{} package.
-%
-The installation is available through a git repository and also on the
-Python Package Index (PyPI).
-%
-Since the simulation setup is configurable, the time required for a
-simulation to complete depends on the simulation parameters.
-%
-It also depends on the hardware that is used to make the run, and on whether the parallel option is set.
-%
-However, it is simple to make a local run of limited versions of the experiments to have a sense of the time required by
-the code to complete the simulations.
-%
-More information on this can be found on the documentation of the software, available at
-\url{https://hearsay.readthedocs.io/en/latest/}.
+Regarding the quality assurance of the software, we endorse the PEP-8 style
+guide for python (van Rossum et al. 2001) throughout the entire project, which
+improves the code maintainability and helps to minimise the so called
+``code-entropy'' of future versions.
+%
+The deviations in style were analysed with the flake8 tool, which improves the
+readability and consistency of the code.
+%
+Testing and code coverage have been applied to the package and the development
+process as standard qualitative and quantitative metrics.
+%
+The different parts and functions can be validated to assert if they return the
+expected results on specific test cases and testing scenarios
+\citep{jazayeri_engeneering_2007}.
+%
+On the other hand, code coverage measures the fraction visited and executed in
+the code by the testing procedure \citep{miller_testing_1963}.
+%
+The advantage of an exhaustive code-coverage is that it ensures that most of
+the sources of errors have been considered.
+%
+In the \hs{} project we provide test tools for the distinct parts of the
+simulation setup and run, reaching a 99\% of code coverage in the version used
+for this work.
+%
+For every new version of the code, all these quality assurance methods and
+metrics are run automatically by a continuous-integration service.
+%
+It is possible to reproduce the experiments performed in this work using the
+\hs{} package.
+%
+The installation is available through a git repository and also on the Python
+Package Index (PyPI).
+%
+Since the simulation setup is configurable, the time required for a simulation
+to complete depends on the simulation parameters.
+%
+It also depends on the hardware that is used to make the run, and on whether
+the parallel option is set.
+%
+However, it is simple to make a local run of limited versions of the
+experiments to have a sense of the time required by the code to complete the
+simulations.
+%
+More information on this can be found on the documentation of the software.
 
@@
 \hline
-   Model subset & D$_{\rm max}$ [lyr] & $\tau_a$ interval [a] & $\tau_s$ interval 
-   [a]& description  \\
+Model subset& D$_{\rm max}$ [lyr] & $\tau_a$ [a] & $\tau_s$ [a]&description\\
 \hline
-M1 & 40000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
-M2 & 40000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
-M3 & 40000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
-M4 & 40000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
-%
-M5 & 80000 & [10000, 30000]   & [10000, 50000]   &dense, short lifetime\\
-M6 & 80000 & [170000, 190000] & [10000, 50000]   &sparse, short lifetime\\
-M7 & 80000 & [10000, 30000]   & [400000, 440000] &dense, long lifetime \\
-M8 & 80000 & [170000, 190000] & [400000, 440000] &sparse, long lifetime\\
+M1 & 40000 & 10$^4$   & 10$^4$  & dense, short lifetime\\
+M2 & 40000 & 10$^6$   & 10$^4$  & sparse, short lifetime\\
+M3 & 40000 & 10$^4$   & 10$^5$  & dense, long lifetime \\
+M4 & 40000 & 10$^6$   & 10$^5$  & sparse, long lifetime\\
+%                               
+M5 & 80000 & 10$^4$   & 10$^4$  & dense, short lifetime\\
+M6 & 80000 & 10$^6$   & 10$^4$  & sparse, short lifetime\\
+M7 & 80000 & 10$^4$   & 10$^6$  & dense, long lifetime \\
+M8 & 80000 & 10$^6$   & 10$^6$  & sparse, long lifetime\\
 %
@@
 %
-For each model, we simulated 50 realizations with different random
+For each model, we simulated several realizations with different random
 seeds, in order to perform a MonteCarlo estimation of the variance.
+%
+The number of random realzations varies from one (for the densel populated models)
+up to 20 (for the sparsely populated models).
 %
@@
 waiting time for the next awakening, $\left<\tau_a\right>$, and the
-mean lifetime, $\left<\tau_s\right>$) cover the ranges 10$^2$-10$^6$~a, with a linear partition of 50 values for each parameter.
-%
-For the D$_{\rm max}$ parameter, we take the range 30~pc to 10~kpc, on
-50 linear bins.
-%
-This setup makes a total of 750000 simulations, each covering a time
+mean lifetime, $\left<\tau_s\right>$) cover the ranges 10$^2$-10$^6$~a, with 
+two linear partitions of 21 values (10$^2$-40500~a) and 51 values ($5\cdot10^5$-10$^6$~a)
+for each parameter.
+%
+This partition was chosen after the requirement of the software to take linear bins, 
+aiming at a better sampling of the low $\tau_a$ and low $\tau_s$ region of the parameter space.
+%
+For the D$_{\rm max}$ parameter, we take the 
+values 100~pc, 500~pc, 1~kpc, 5~kpc and 10~kpc.
+%
+This setup makes a total of 40000 simulations, each covering a time
 range of ten million years.
@@
 %
-Some of them are directly derived from the discrete events, namely,
-the ID of emitting and receiving \cetis{} and the position in the
-galaxy.
-%
-We also save the times of each of the events that are relevant to keep
-track of the number of \cetis{} for each simulation, i.e., the times
-of the four types of events.
-%
-the times of C-type and B-type events are used to derive the number of
+Some of them are directly derived from the discrete events, namely, the ID of
+emitting and receiving \cetis{} and the position in the galaxy.
+%
+We also save the times of each of the events that are relevant to keep track of
+the number of \cetis{} for each simulation, i.e., the times of the four types
+of events.
+%
+The times of C-type and B-type events are used to derive the number of
 contacts.
 %
-We also obtain quantities that represent the
-properties of the \cetis{}, for example the total time elapsed between the
-A-type and the D-type of each node, which represent the corresponding
-lifetimes of the nodes.
-%
-The time span of a \ceti{} listening another or being listened by another
+We also obtain quantities that represent the properties of the \cetis{}, for
+example the total time elapsed between the A-type and the D-type of each node,
+which represent their corresponding lifetimes.
+%
+The time span of a \ceti{} listening another or being listened by another node
 can also be derived by keepig track of the times of the events in a simulation.
 %
-This way we can also compute the distribution in the galaxy of
-\cetis{}
-that reach contact, the distributions of the waiting times until the 
-first contact or the distributions of
-waiting times until the next contact.
-%
-The following properties of the population of \cetis{} can be derived:
-%
-the fraction of the lifetime a node is listening
-to
-at least another node (i.e., in their causal cone),
-%
-the age of
-contacted \cetis{} at first contact,
-%
-the fraction of \cetis{} where the first contact is
-given at the awakening,
-%
-the distribution of the number of contacts for
-each \ceti{},
-%
-the distribution of the number of contacts as a function of
-the age of the node,
+This way we can also compute the distribution in the simulated galaxy of
+\cetis{} that reach contact, the distributions of the waiting times until the
+first contact or the distributions of waiting times until the next contact.
+%
+The following properties of the population of \cetis{} can also be derived:
+%
+the fraction of the lifetime a node is listening to at least another node
+(i.e., in their causal cone),
+%
+the age of contacted \cetis{} at first contact,
+%
+the fraction of \cetis{} where the first contact is given at the awakening,
+%
+the distribution of the number of contacts for each \ceti{},
+%
+the distribution of the number of contacts as a function of the age of the
+node,
 %
@@
 %
-and the distribution of
-distances between contacted nodes.
-%
-Another useful derived quantity is the duration of two-way
-communication channels or the fraction of contacts that admit a
-response.
-% 
-It is also possible to analyze the relations between the distance to
-\ceti{} vs. the time of two-way communications, the distance to \ceti{} vs. the
-age of contacted node, the age of a node and the maximum number of
-contacted nodes before the D--type event, or the lifespan of a node vs. the
-maximum number of contacts.
-%
-All these quantities can be analyzed as a function of the
-simulation parameters.
-%
-We chose eight models that are on opposed regions of the
-explored parameter space, and cover
-all possible combinations of 
-short/long lifetimes, short/long
+and the distribution of distances between contacted nodes.
+%
+Another useful derived quantity is the duration of two-way communication
+channels or the fraction of contacts that admit a response.
+% 
+It is also possible to analyze the relations between the distance to \ceti{}
+vs. the time of two-way communications, the distance to \ceti{} vs. the age of
+contacted node, the age of a node and the maximum number of contacted nodes
+before the D--type event, or the lifespan of a node vs. the maximum number of
+contacts.
+%
+All these quantities can be analyzed as a function of the simulation
+parameters.
+%
+We chose eight models that are on opposed regions of the explored parameter
+space, and cover all possible combinations of short/long lifetimes, short/long
 waiting times for the next \ceti{} to appear in the galaxy and short/large
@@
 %
-The details of each model are summarized in the
-Table~\ref{T_selected_models}.
+The details of each model are summarized in the Table~\ref{T_selected_models}.
 

### 2020-05-30 submit_v03/manuscript.tex
version definitiva?

@@
 %
-The abundance of intelligent civilisations in the Galaxy is a
-   longstanding question related to the expected probability
-   of success for SETI programs.  Traditional approaches to quantify
-   this problem use either frequentist inferences or numerical
-   simulations. The key challenges in both cases are the proper
-   treatment of the time variable and the large uncertainties on the
-   intervening factors. Here we present a three-parametric
-   statistical model of the network of causally connected locations in
-   a simplified Galaxy, with the restriction of a maximum distance
-   range for any signal. We implement Monte Carlo simulations of this
-   model to explore the parameter space and analyse the probabilities
-   of causal contacts for random locations in the Galaxy. We find
-   that, given the enormous distances involved, causal contacts
-   between civilisations would be very rare. The odds to make a
-   contact in a few years of monitoring are low for most models,
-   except for those of a galaxy densely populated with ancient
-   civilisations. The probability of causal contacts increases with
+    In this work we address the problem of estimating the
+    probabilities of causal contacts between civilisations in the
+    Galaxy. We make no assumptions regarding the origin and
+    evolution of intelligent life.  We simply assume a network of
+    causally connected nodes.  These nodes refer somehow to
+    intelligent agents with the capacity of receiving and emitting
+    electromagnetic signals. Here we present a three-parametric
+    statistical Monte Carlo model of the network in a simplified
+    sketch of the Galaxy. Our goal, using Monte Carlo simulations, is
+    to explore the parameter space and analyse the probabilities of
+    causal contacts. We find that the odds to make a contact over
+    decades of monitoring are low for most models, except for those of
+    a galaxy densely populated with long-standing civilisations. We
+    also find that the probability of causal contacts increases with
     the lifetime of civilisations more significantly than with the
     number of active civilisations. We show that the maximum
-   probability of making a contact occurs when a civilization
+    probability of making a contact occurs when a civilisation
     discovers the required communication technology.
@@
 Optimistic estimates from the Drake equation contrast with the
-so-called Fermi paradox, which states the contradiction between the
+so-called Fermi paradox, which states the apparent contradiction between the
 expected abundance of life in the Galaxy and the lack of evidence for
 it \citep[e.g. ][]{hart_explanation_1975, brin_great_1983,
-barlow_galactic_2012, forgan_galactic_2016, vanhouten_isthere_2017,
+barlow_galactic_2012, forgan_galactic_2016, anchordoqui_is_2017,
 Sotos_biotechnology_2019, carroll_nellemback_fermi_2019}.
@@
 There are many propositions aimed at solving this paradox, which make
-use of statistical \citep{solomonides_probabilistic_2016,
-vanhouten_isthere_2017, horvat_calculating_2007,
+use of statistical \citep{solomonides_probabilistic_2016, horvat_calculating_2007,
 maccone_statistical_2015} or stochastic approaches
@@
 %
-The lack of contacts could be explained by the difficulty of life to
-appear due to astrophysical explanations
+The absence of detections of extraterrestrial intelligent signals
+could be explained by
+astrophysical phenomena that makes life difficult to develop
 \citep{annis_astrophysical_1999}.
 %
-In general, also more speculative proposals
+Besides the possible scarcity of life,
+alternative scenarios have also been discussed
 \citep{barlow_galactic_2013, lampton_information_2013,
-conway_three_2018, forgan_galactic_2017} have been discussed.
-% 
-We stress the fact that the lack of detections of extraterrestrial
-intelligent signals is the only observational fact that can be taken
-for sure.
-%
-Indeed, as \citep{tarter_search_2001} pointed out, according to our current technical
-capabilities for SETI, we have not received any signal yet.
+conway_three_2018, forgan_galactic_2017}.
+%
+The large distances in the Galaxy and the likely limited lifetime of
+civilizations may play an important role in determining how difficult it
+would be to obtain evidence for other inhabited worlds.
+%
+%We stress the fact that the lack of detections of extraterrestrial
+%intelligent signals is the only observational fact that can be taken
+%for sure.
+The analysis of these scenarios is difficult due to the lack of data
+about the hypothetical extraterrestrial intelligences.
+%
+Indeed, as \citep{tarter_search_2001} pointed out, according to our
+current technical capabilities for the search of extraterrestrial
+intelligence (SETI), we have not received any signal yet.
 %
@@
 %
-The absence of detection has also motivated alternative ideas for new
+The absence of detections has also motivated alternative ideas for new
 SETI strategies \citep{forgan_exoplanet_2017, balbi_impact_2018,
@@
 forgan_possibility_2013, forgan_galactic_2017, funes_searching_2019}.
-%
+
 \Fpagebreak
-%
-The Drake equation is a key tool to organise the discussion about the
-problem of the (unknown) abundance of civilisations in the Galaxy
-\citep{hinkel_interdisciplinary_2019}.
-%
-However, the uncertainties in the factors of this equation make it
-less suitable to a formal study to define searching strategies or to
-compute the actual number of extraterrestrial intelligences.
+
+The discussion about the problem of the (unknown)
+abundance of civilisations in the Galaxy has been organised around the
+factors in the Drake equation \citep{hinkel_interdisciplinary_2019}.
+%
+However, the uncertainties in these factors, specially the ones
+representing biological processes, make it
+less suitable to a formal study with the purpose of defining
+searching strategies or
+computing the estimated number of extraterrestrial intelligences.
 %
@@
 %
-\citet{prantzos_joint_2013} proposes a unified framework for a joint
+\citet{prantzos_joint_2013}, for example, proposes a unified framework for a joint
 analysis of the Drake equation and the Fermi paradox, concluding that
-for sufficiently long-lived civilisations, colonisation of the Galaxy
-is the only reasonable option to gain awareness about other life
-forms.
+for sufficiently long-lived civilisations, colonisation
+is the most promising strategy to find other life forms.
 %
@@
 %
-Some modifications to the original idea of the equation have tried to
-imprint a stochastic nature, to propose a probabilistic approach, or
-to consider the temporal structure missing in the equation.
+Some modifications to the original idea of the Drake equation have
+been proposed, in order to
+integrate a temporal structure, to
+reformulate it as a stochastic process or to propose alternative
+probabilistic expressions.
 %
@@
 %
-The literature about this topic brings a number of works that consider
+The literature about this topic brings also a number of works that consider
 numerical simulations \citep{forgan_evaluating_2015,
@@
 %
-Although this approach does not rely on observed quantities that could
-be modified with time (like, for instance, the fraction of stars with planets), 
-it also depend on the definition of a number of
-parameters which are unknown or uncertain, required to carry out
+Although this approach does not rely on values that can be measured
+from observations
+(like the fraction of stars with planets), 
+it depends on the definition of
+unknown or uncertain parameters required to carry out
 a the simulations.
@@
 
+Several authors raise the distinction between causal contacts and
+actual contacts.
+%
+In the first case, the determinations of metrics for the likelihood of
+a contact are independent of considerations about the technological
+resources or the implicit coordination to decipher intelligent
+messages.
+%
 In a recent work, \citet{balbi_impact_2018} uses a statistical model to
@@
 structure in the Drake equation and, in particular, the limitations
-of this expresion to estimate the required timescale of a SETI
+of this expression to estimate the required timescale of a SETI
 program to succeed in the detection of intelligent signals.
-%
-Despite its shotcomming for a solid mathematical foundation of SETI, 
-the Drake equation has also motivated efforts to produce formal
-statistical results.
+
+The quest for a formal statistical theory has also lead 
+to important progress in the mathematical foundations of SETI.
 %
@@
 To that end, the authors employ Monte Carlo Markov Chains over each factor
-of the Drake equation, and combine the means to reach a probabilistic
+of the Drake equation, and combine the mean values to reach a probabilistic
 result.
@@
 %
-This study concludes that there is a tiny probability for Earth to
-contact another extraterrestrial civilization.    
+This study concludes that there is a small probability that the Galaxy
+is populated with a large number of communicating civilisations.
 %
 \citet{smith_broadcasting_2009} uses an analytical model to gauge the
-probabilities of contact between two randomly located civilisations,
-and the waiting time for the first contact, assuming a fixed maximum
-broadcasting distance.
-% 
-Different statistical modelings have been also proposed.
-%
-A notable example of this es the work by \citet{balbi_impact_2018}, who investigates the
-chance of communicating civilisations making causal contact within a
-volume surrounding the location of the Earth, using a statistical
-model to compute the occurrence of causal contacts.
-%
-The author stresses the fact that the causal contact requirement
-involves the distance between civilisations, their lifespan and their
-times of appearance.
-%
-This is important since the time it takes the
-light to travel across the Galaxy might be much lesser than its lifetime.
+probabilities of contact between two randomly located civilisations
+and the waiting time for the first contact.
+%
+The author stresses that 
+the maximum broadcasting distance and the lifetime of civilisations
+come into play to produce the possible network of connections.
+
+A notable example of statistica modeling is the work by
+\citet{balbi_impact_2018}, who investigates the chance of
+communicating civilisations making causal contact within a volume
+surrounding the location of the Earth.
+%
+The author argues that the causal contact requirement involves mainly
+the distance between civilisations, their lifespan and their times of
+appearance.
+%
+This is important since the time the light takes to travel across the
+Galaxy might be much lesser than the lifetime of the emitter.
 %
 \citet{balbi_impact_2018} fixes the total number of civilisations and
-explore the parametric space that comprises three parameters, namely,
+explores the parametric space that comprises three parameters, namely,
 the distance to the Earth, the time of appearance and the lifespan of
@@
 For the distribution of the characteristic time of appearance, the
-author explores exponential and truncated Gaussian distributions.
-%
-Moreover, for the distribution of the lifespans, the author chooses an
-exponential distribution.
-%
-It is important to consider that the fraction of communicating
-civilisations vary with the choice of the statistical distribution for
-the time of appearance, as shown by \citet{balbi_impact_2018}.
-%
-For all analysed distributions, the author concludes that,
-the detectable fraction corresponding to
-
-for a mean
-lifetime of 10$^9$ years and for a maximum radius for the detection of
-the signal of 1000~ly, the detectable civilizations
-are roughly the 10 per cent of the total, due to their spread in time and
-their limited lifetimes.
-%
-In another related approach, \citet{vukotic_astrobiological_2012}
-propose a probabilistic cellular automata (PCA) modelling. 
+author explores exponential and truncated Gaussian distributions,
+while for the distribution of the lifespans chooses an exponential
+distribution.
+%
+It is important to consider that the estimation of the number of
+communicating civilisations vary with the choice of the statistical
+distribution for the time of appearance, as shown by
+\citet{balbi_impact_2018}.
+%
+For all analysed distributions, the author concludes that, the
+fraction of emitters that are listened is low if they are spreaded in
+time and with limited lifetimes.
+%
+In another numerical approach, \citet{vukotic_astrobiological_2012}
+propose a probabilistic cellular automata modelling. 
 %
@@
 which evolve at discrete time steps, according to transition rules
-that take into account for each cell the states of its neighbour
-cells.
-%
-The authors implement a PCA model to a network of cells which
-represent life complexity on a 2-dimensional region resembling the
-Galactic Habitable Zone (GHZ),  an annular ring set between a minimal
-radius of 6 kpc and a peak radius of 10 kpc. 
+that take into account the neighbour cells.
+%
+The authors implement this model to a network of cells which represent
+life complexity on a 2-dimensional region resembling the Galactic
+Habitable Zone (GHZ), an annular ring set between a minimal radius of
+6 kpc and a peak radius of 10 kpc.
+%
+This simulations represent the spread if intelligence as an
+implementation of panspermia theories.
 %
@@
 %
-The work aims at analysing the evolution of life, although it does not
-account for the network of causal contacts among
-technological civilisations.
+Their work aims at analysing the evolution of life, and although it
+does not account for the network of causal contacts among
+technological civilisations, it offers a tool to think SETI searches
+from a novel point of view.
 
@@
 
-In this work we address the problem of the temporal and spatial structure of
+This study is conceived as an introduction to a simple probabilistic
+model and a numerical exploration of its parameter space.
+%
+With these tools, we address the problem of the temporal and spatial structure of
 the distribution of communicating civilisations.
 %
-We make no asumptions about, e.g., the origin of life, the development of
+This approach does not require asumptions about, e.g., the origin of life, the development of
 intelligence or the formation of habitable planets according to stellar type.
@@
 %
-We chose this generic name in order to stress the fact that in this analysis 
+We choose this generic name in order to stress the fact that in this analysis 
 only the causal contact is considered, independently of any broadcasting or
@@
 %
-Panel (a) of the Figure represents the region $R$ (shaded polygon)
+Panel (a) of the represents the region $R$ (shaded polygon)
 which is causally connected to the emitter (left vertex).
@@
 The sphere of first contact (SFC) is centred on the emitter,
-represented in the Figure by the left angle of the region $R$.
+represented by the left angle of the region $R$.
 %
@@
 The temporal intervals for the communication between two \cetis{} are
-represented in the panel (b) of the Figure.
+represented in the panel (b) of the Fig.~\ref{F_scheme}.
 %
 In this scheme, the entire lifetime of a \ceti{} can be represented by
-a polygon, as shown in the Figure.
-%
-The polygon is limited in the spatial axis by the maximum distance of
+a polygon.
+%
+This polygon is limited in the spatial axis by the maximum distance of
 the signal $D_{\rm max}$ and in the time axis by the time span between the
@@
 The points in time where the \cetis{} acquire their communicating
-capacity are dubbed ``Awakening'' events ($A_1$ and $A_2$ in the
-Figure).
+capacity are dubbed ``Awakening'' events ($A_1$ and $A_2$).
 %
 Similarly, the instants in time where the \cetis{} lose their
-communicating capacity are dubbed ‘’Doomsday’’ events ($D_1$ and $D_2$
-in the Figure).
+communicating capacity are dubbed ‘’Doomsday’’ events ($D_1$ and $D_2$).
 %
@@
 \cetis{}, are denominated ``Blackout'' events, presented as $B_1$ and
-$B_2$ in the Figure.
+$B_2$.
 %
@@
 %
-Any other \ceti{} within this region will be aware of the originating node,
+Any other \ceti{} within this region will be in causal contact with the originating node,
 even if it has disappeared before the time of contact.
@@
 
-%}}}
-
  \setlength{\tabcolsep}{10pt}
@@
 \hline
-   \multicolumn{2}{l}{independent variables (hypothesis space)}
+   \multicolumn{2}{l}{independent variables (parameter space)}
    &min. value&max. value&Nbins\\
+%\hline
+%   $\tau_{a}$ & Mean temporal separation between consecutive awakenings 
+%	      & 500~a & 40500~a & 21\\ 
+%	    & & 5$\cdot$10$^5$~a & 10$^6$~a & 51\\ 
+%   $\tau_{s}$ & Mean lifetime of a \ceti{}
+%	      & 500~a & 40500~a & 21\\ 
+%	    & & 5$\cdot$10$^5$~a & 10$^6$~a & 51\\ 
+%	D$_{\rm max}$ & Maximum reach of a message  & \multicolumn{3}{c}{100 pc, 500 pc, 1 kpc, 5 kpc, 10 kpc} \\
 \hline
-   $\tau_{a}$ & Mean temporal separation between consecutive awakenings 
-	      & 500~a & 40500~a & 21\\ 
-	    & & 5$\cdot$10$^5$~a & 10$^6$~a & 51\\ 
+
+   $\tau_{a}$ & Mean temporal separation between consecutive
+   awakenings & 4000 yr & 200\,000 yr & 50\\ 
    $\tau_{s}$ & Mean lifetime of a \ceti{}
-	      & 500~a & 40500~a & 21\\ 
-	    & & 5$\cdot$10$^5$~a & 10$^6$~a & 51\\ 
-	D$_{\rm max}$ & Maximum reach of a message  & \multicolumn{3}{c}{100 pc, 500 pc, 1 kpc, 5 kpc, 10 kpc} \\
+   & 10000 yr & 500\,000 yr& 50\\ 
+   D$_{max}$ & Maximum reach of a message  &  \multicolumn{2}{l}{3 kpc, 12 kpc, 25 kpc} & 3\\
 \hline
+
    \multicolumn{2}{l}{fixed variables} & \multicolumn{2}{c}{assumptions} &value \\
@@
    R$_{GHZ}^{\rm max}$   & Outer radius of the GHZ       & \citet{lineweaver_galactic_2004} & & 9~kpc\\
-	t$_{\rm max}$ & Time span of the simulation  & & \multicolumn{2}{r}{5$\cdot$10$^6$~a, 10$^7$~a} \\
-    & \multicolumn{3}{l}{number of random realizations for each point in the hypothesis space} & 10, 5, 1 \\
+	t$_{\rm max}$ & Time span of the simulation  & & \multicolumn{2}{r}{10$^6$~yr} \\
+    & \multicolumn{3}{l}{number of random realizations for each point
+    in the parameter space} & 50 \\
 \hline
@@
 \subsection{Power laws vs. exponential laws}\label{SS_PDF_shape}
-%{{{
 
@@
 instead of analysing results from a particular parameter chosen ad hoc, we
-explore the hypothesis space and analyse the impact of the values of these
+explore the parameter space and analyse the impact of the values of these
 parameters on the results.
-
-%}}}
 
@@
 Empirical cumulative distributions of the number contacts
-for \cetis in four different samples (M5, M6, M7, M8) with two
-values for the range of the signal reach: D$_{\rm max}$=40000 ly (blue),  and 
-D$_{\rm max}$=80000 ly (orange).
-%
-Except for the models corresponding to a densely populated galaxy with
-long--living \cetis, the \cetis which receive more than 10 contacts
-are very unlikely.
-%
-See Table \ref{T_selected_models} for model descriptions.
+   for \cetis in different samples and with D$_{\rm max}$=10~kpc.
+%
+The upper panel shows the variation of the probability distribution
+   estimation as a function of the $\tau_a$ parameter.  All the curves
+   correspond to models with 
+   $\tau_s=40~kyr$ and D$_{\rm max}=10\,kpc$.
+%
+The bottom panel corresponds to several values of $\tau_s$, with
+   $\tau_a=1~kyr$ and D$_{\rm max}=10\,kpc$.
+%
+   Cumulative distributions can be used to visualise the probability
+   of a random node in a model of having more than M contacts.  For
+   example, in the upper panel the probability of a random node having
+   more than 10 contacts is nearly one for the model with the
+   smallest $\tau_a$ value, and nearly zero for the model with the
+   largest $\tau_a$ value.
    %
@@
 \subsection{Complexity of the model}
-%{{{
 
@@
 %
-Finally, regarding the degree of idealisation, the model admits different
+Regarding the degree of idealisation, the model admits different
 efficiencies for the sphere of the causal region of each \ceti{}, featuring
@@
 %
-For example, a power law where powerful emissions are rare and low emissions
-are common could be an improvement to the model.
-%
-In this work we chose not to implement this for the sake of simplicity.
+For example, a power law where a powerful emission is rare and a low emission
+is common could be an improvement to the model.
+%
+In this work we choose not to implement this for the sake of simplicity.
 %
@@
 \subsection{Discrete event process}
-%{{{
-
-A discrete event simulation is performed for a given model $M(\tau_a, \tau_s,
-D_{\rm max})$ by keeping track of a set of variables that change each time an
+
+A discrete event simulation is performed for a given model, in this
+work $M(\tau_a, \tau_s,
+D_{\rm max})$, by keeping track of a set of variables that change each time an
 event happens. 
@@
 
-%}}}
-
 \begin{figure*} % 2D color plot
@@
 function of $\tau_a$ and $\tau_s$, for 
-$D_{\rm max}$=10000 (left panel),
-$D_{\rm max}$=40000 (middle panel), and
-$D_{\rm max}$=80000 (right panel)
+$D_{\rm max}$=3~kpc (left panel),
+$D_{\rm max}$=12~kpc (middle panel), and
+$D_{\rm max}$=25~kpc (right panel)
 %
@@
 awakening, as a funcion of $\tau_a$ and $\tau_s$, for
-$D_{\rm max}$=10000 (left panel),
-$D_{\rm max}$=40000 (middle panel), and
-$D_{\rm max}$=80000 (right panel)
+$D_{\rm max}$=3~kpc (left panel),
+$D_{\rm max}$=12~kpc (middle panel), and
+$D_{\rm max}$=25~kpc (right panel)
 %
@@
 \subsection{Model for \cetis{}}
-%{{{
 
@@
 %
-Although there are several possible approaches, we chose to follow the
+Although there are several possible approaches, we choose to follow the
 evolution of the system according to the following events:
@@
       lookout for messages.
-   \item[(C)] Contact: A new one-way causal contact is established
-      (does not require the awareness of the involved \cetis{}.)
+   \item[(C)] Contact: A new one-way causal contact is established.
    \item[(B)] Blackout: An existing causal contact is interrupted,
@@
 %
-%}}}
 
 \subsection{Model implementation}
-%{{{
 
@@
 
-%}}}
-
 \begin{figure}
@@
 Left panel shows the histograms for several values of $\tau_a$, and
-$\tau_s$ in the range 17000-24000~a.
+$\tau_s$ in the range 17000-24000~yr.
 %
 Right panel shows the histograms for several values of $\tau_s$, and
-$\tau_a$ in the range 68000-100000~a.
+$\tau_a$ in the range 68000-100000~yr.
 %
@@
 
-\setlength{\tabcolsep}{10pt}
-\begin{table*}
-\centering
-\begin{tabular}{ccccc}
-\hline
-Model subset& D$_{\rm max}$ [lyr] & $\tau_a$ [a] & $\tau_s$ [a]&description\\
-\hline
-M1 & 40000 & 10$^4$   & 10$^4$  & dense, short lifetime\\
-M2 & 40000 & 10$^6$   & 10$^4$  & sparse, short lifetime\\
-M3 & 40000 & 10$^4$   & 10$^5$  & dense, long lifetime \\
-M4 & 40000 & 10$^6$   & 10$^5$  & sparse, long lifetime\\
-%                               
-M5 & 80000 & 10$^4$   & 10$^4$  & dense, short lifetime\\
-M6 & 80000 & 10$^6$   & 10$^4$  & sparse, short lifetime\\
-M7 & 80000 & 10$^4$   & 10$^6$  & dense, long lifetime \\
-M8 & 80000 & 10$^6$   & 10$^6$  & sparse, long lifetime\\
-%
-\hline
-\end{tabular}
-\caption{Selected models to analyze the behavior of simulation outputs
-   and their dependence on simulation parameters.}
-\label{T_selected_models}
-\end{table*}
-
 %%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
 \section{Results: exploring the parameter space}\label{S_results}
-%{{{
 
 We implemented the simulation of a regular grid of models varying over
-the hypothesis space, which covers 15000 different models.
+the parameter space, which covers 15000 different models.
 %
@@
 waiting time for the next awakening, $\left<\tau_a\right>$, and the
-mean lifetime, $\left<\tau_s\right>$) cover the ranges 10$^2$-10$^6$~a, with 
-two linear partitions of 21 values (10$^2$-40500~a) and 51 values ($5\cdot10^5$-10$^6$~a)
+mean lifetime, $\left<\tau_s\right>$) cover the ranges
+10$^2$-10$^6$~yr, with 
+two linear partitions of 21 values (10$^2$-40500~yr) and 51 values
+($5\cdot10^5$-10$^6$~yr)
 for each parameter.
@@
 %
-This setup makes a total of 40000 simulations, each covering a time
+This setup makes a total of 750000 simulations, each covering a time
 range of ten million years.
@@
 %
-We chose eight models that are on opposed regions of the explored parameter
-space, and cover all possible combinations of short/long lifetimes, short/long
-waiting times for the next \ceti{} to appear in the galaxy and short/large
-range reach of the signal.
-%
-\ttn{2}
-%
-The details of each model are summarized in the Table~\ref{T_selected_models}.
-
-%}}}
 
 \subsection{Membership to the network of connected \cetis{}}\label{SS_members}
-%{{{
 
@@
 samples, including short and long lifetimes
-$\left(\,\left<\tau_s\right>=\right.$10000~a and
-$\left<\tau_s\right>$=100000~a, respectively$\left.\right)$, dense
+$\left(\,\left<\tau_s\right>=\right.$10000~yr and
+$\left<\tau_s\right>$=100000~yr, respectively$\left.\right)$, dense
 and sparse spatial distribution
-$\left(\,\left<\tau_a\right>=\right.$10000~a and
-$\left<\tau_a\right>$= 100000~a, respectively$\left.\right)$, and two
-different signal reach ranges (D$_{\rm max}$=40000 lyr and
-D$_{\rm max}$=80000~lyr).
+$\left(\,\left<\tau_a\right>=\right.$10000~yr and
+$\left<\tau_a\right>$= 100000~yr, respectively$\left.\right)$, and two
+different signal reach ranges (D$_{\rm max}$=12~kpc and
+D$_{\rm max}$=25~kpc).
 %
@@
 %
-As it can be seen from this Figure, the mean lifetime is more
+As it can be seen, the mean lifetime is more
 determinant than the mean awakening rate (dense and sparse,
@@
 Similarly, for the models with long signal range
-($D_{\rm max}=80000\,lyr$), the one which is dense and with long lifetime
+($D_{\rm max}=25$~kpc), the one which is dense and with long lifetime
 has the maximum number of contacts, reaching nearly 100 contacts for
@@
 %
-Another interesting feature observed in this Figure, is that for the
+Another interesting feature observed in the Fig.~\ref{F_number_of_contacts}, is that for the
 models analyzed, there are \cetis{} that have no causal contacts with
@@
 time ($\left<\tau_a\right>$, in the range 0--200 a), for three
-different values of the maximum signal range, $D_{\rm max}=10000$~a (left
-panel), $D_{\rm max}=40000$~a (middle panel), and $D_{\rm
-max}=80000$~a
+different values of the maximum signal range, $D_{\rm max}$=3~kpc (left
+panel), $D_{\rm max}=12$~kpc (middle panel), and $D_{\rm max}=25$~kpc
 (right panel).
 %
-A clear pattern emerges in this Figure, showing that the probability
+A clear pattern emerges, showing that the probability
 for a \ceti{} of making causal contact with at least another \ceti{}
@@
 The three panels correspond to different values of the maximum signal
-range ($D_{\rm max}$),
-as indicated in the Figure:
-%
-$D_{\rm max}=10000$~a (left
-panel), $D_{\rm max}=40000$~a (middle panel), and $D_{\rm
-max}=80000$~a
+range ($D_{\rm max}$):
+%
+$D_{\rm max}=3$~kpc (left
+panel), $D_{\rm max}=12$~kpc (middle panel), and $D_{\rm max}=25$~kpc
 (right panel).
-%
-These probabilities are consistent with the ones on the previous
-Figure.
 %
@@
 
-%
-
-%}}}
-
 \subsection{Waiting time for a first contact}\label{SS_waiting}
-%{{{
 
@@
 from the awakening
-($t=$0~a) up to a given time are related to the estimation of the
+($t=$0~yr) up to a given time are related to the estimation of the
 probability of listening during that time interval and reaching to
@@
 Clearly, this probability diminishes with time and tends to zero for
-large time periods, meaning the a \ceti{} will eventually get aware of
-the presence of other \cetis{}
+large time periods, meaning the a \ceti{} will eventually enter into 
+the light cone of other \cetis{}
 
@@
 
-%}}}
-
 %%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
 \section{Discussion}\label{S_discussion}
-%{{{
 
@@
 separations between the activation of new nodes
-in the range 500--10$^6$~a.
+in the range 500--10$^6$~yr.
 %
 Additionally, we explore a grid of the mean survival times between 
-500~a and 10$^6$~a, and values for the maximum distance range of signals 
+500~yr and 10$^6$~yr, and values for the maximum distance range of signals 
 of 100~pc, 500~pc, 1~kpc, 5~kpc and 10~kpc.
@@
 Instead of making a number of assumptions, we have explored the
-hypothesis space, reducing the problem to only three parameters and a
+parameter space, reducing the problem to only three parameters and a
 few simple hypothesis to perform a complete model for the population and
@@
 different scenarios.
-
-%}}}
 
@@
 \setlength{\bibsep}{0.0pt}
-\bibliography{biblio_seti, biblio_software}
+%\bibliography{biblio_seti, biblio_software}
+
+\begin{thebibliography}{95}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{Adamic(2000)}]{adamic_zipf_2000}
+Adamic, L.A., 2000. Zipf, power-laws, and pareto-a ranking tutorial. Xerox Palo Alto Research Center, Palo Alto, CA, http://ginger. hpl. hp. com/shl/papers/ranking/ranking. html.
+
+\bibitem[{{Allen} {et~al}\mbox{.}(2020){Allen}, {Berriman}, {DuPrie}, {Mink},
+  {Nemiroff}, {Schmidt}, {Shamir}, {Shortridge}, {Taylor}, \&
+  {Teuben}}]{2020ASPC..522..731A}
+{Allen} A. {et~al.}, 2020, in Astronomical Society of the Pacific Conference
+  Series, Vol. 522, Astronomical Data Analysis Software and Systems XXVII,
+  {Ballester} P., {Ibsen} J., {Solar} M., {Shortridge} K., eds., p. 731
+
+\bibitem[{{Allen} \& {Schmidt}(2015)}]{2015JORS....3E..15A}
+{Allen} A., {Schmidt} J., 2015, Journal of Open Research Software, 3, E15
+
+\bibitem[{Anchordoqui \& Weber(2019)}]{anchordoqui_upper_2019}
+Anchordoqui L.~A., Weber S.~M., 2019, eprint arXiv:1908.01335
+
+\bibitem[{Annis(1999)}]{annis_astrophysical_1999}
+Annis J., 1999, JBIS - Journal of the British Interplanetary Society, 52, 19
+
+\bibitem[{Armstrong \& Sandberg(2013)}]{Armstrong2013}
+Armstrong S., Sandberg A., 2013, Acta Astronautica, 89, 1
+
+\bibitem[{Balb(2018)}]{balb_spatiotemporal_2018}
+Balb A., 2018, Mem. della Soc. Astron. Ital., 89, 425
+
+\bibitem[{Balbi(2018)}]{balbi_impact_2018}
+Balbi A., 2018, Astrobiology, 18, 54
+
+\bibitem[{Barab{\'{a}}si(2009)}]{barabasi_scale_2009}
+Barab{\'{a}}si A.~L., 2009, Science, 325, 412
+
+\bibitem[{Barlow(2013{\natexlab{a}})}]{barlow_galactic_2012}
+Barlow M.~T., 2013{\natexlab{a}}, Int. J. Astrobiol., 12, 63
+
+\bibitem[{Barlow(2013{\natexlab{b}})}]{barlow_galactic_2013}
+Barlow M.~T., 2013{\natexlab{b}}, Int. J. Astrobiol., 12, 63
+
+\bibitem[{Bloetscher(2019)}]{bloetscher_using_2019}
+Bloetscher F., 2019, Acta Astronaut., 155, 118
+
+\bibitem[{Borra(2012)}]{borra_searching_2012}
+Borra E.~F., 2012, AJ, 144, 181
+
+\bibitem[{Brin(1983)}]{brin_great_1983}
+Brin G., 1983, Quarterly Journal of the Royal Astronomical Society, 24, 283
+
+\bibitem[{Burchell(2006)}]{burchell_whither_2006}
+Burchell M.~J., 2006, Int. J. Astrobiol., 5, 243
+
+\bibitem[{Carroll-Nellenback {et~al}\mbox{.}(2019)Carroll-Nellenback, Frank,
+  Wright, \& Scharf}]{carroll_nellemback_fermi_2019}
+Carroll-Nellenback J., Frank A., Wright J., Scharf C., 2019, AJ, 158, 117
+
+\bibitem[{Chung(2003)}]{chung_simulation_2003}
+Chung C.~A., 2003, {Simulation Modeling Handbook: A Practical Approach
+  (Industrial and Manufacturing Engineering Series)}. CRC press
+
+\bibitem[{{\'{C}}irkovi{\'{c}}(2004)}]{cirkovic_temporal_2004}
+{\'{C}}irkovi{\'{c}} M.~M., 2004, Astrobiology, 4, 225
+
+\bibitem[{{Conway Morris}(2018)}]{conway_three_2018}
+{Conway Morris} S., 2018, Int. J. Astrobiol., 17, 287
+
+\bibitem[{Dayal {et~al}\mbox{.}(2016)Dayal, Ward, \&
+  Cockell}]{dayal_habitability_2016}
+Dayal P., Ward M., Cockell C., 2016, eprint arXiv:1606.09224
+
+\bibitem[{Do{\v{s}}ovi{\'{c}} {et~al}\mbox{.}(2019)Do{\v{s}}ovi{\'{c}},
+  Vukoti{\'{c}}, \& {\'{C}}irkovi{\'{c}}}]{Dosovic2019}
+Do{\v{s}}ovi{\'{c}} V., Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2019,
+  A\&A, 625, A98
+
+\bibitem[{Drake(1962)}]{drake_intelligent_1962}
+Drake F., 1962, ``Intelligent life in space'', Macmillan Ed.
+
+\bibitem[{Edmondson \& Stevens(2003)}]{Edmondson2003}
+Edmondson W.~H., Stevens I.~R., 2003, Int. J. Astrobiol., 2,
+  231
+
+\bibitem[{Enriquez {et~al}\mbox{.}(2017)Enriquez, Siemion, Foster, Gajjar,
+  Hellbourg, Hickish, Isaacson, Price, Croft, DeBoer, Lebofsky, MacMahon, \&
+  Werthimer}]{enriquez_breakthrough_2017}
+Enriquez J.~E. {et~al.}, 2017, ApJ, 849, 104
+
+\bibitem[{Fogg(1987)}]{fogg_temporal_1987}
+Fogg M.~J., 1987, Icarus, 69, 370
+
+\bibitem[{Forgan {et~al}\mbox{.}(2016)Forgan, Dayal, Cockell, \&
+  Libeskind}]{forgan_evaluating_2015}
+Forgan D., Dayal P., Cockell C., Libeskind N., 2016, Int. J. Astrobiol., 16, 60
+
+\bibitem[{Forgan(2009)}]{forgan_numerical_2009}
+Forgan D.~H., 2009, Int. J. Astrobiol., 8, 121
+
+\bibitem[{Forgan(2011)}]{forgan_spatiotemporal_2011}
+Forgan D.~H., 2011, Int. J. Astrobiol., 10, 341
+
+\bibitem[{Forgan(2013)}]{forgan_possibility_2013}
+Forgan D.~H., 2013, JBIS - J. Br. Interplanet. Soc., 66, 144
+
+\bibitem[{Forgan(2014)}]{forgan_collimated_2014}
+Forgan D.~H., 2014, JBIS - J. Br. Interplanet. Soc., 67, 232
+
+\bibitem[{Forgan(2017{\natexlab{a}})}]{forgan_galactic_2016}
+Forgan D.~H., 2017{\natexlab{a}}, Int. J. Astrobiol., 16, 349
+
+\bibitem[{Forgan(2017{\natexlab{b}})}]{forgan_galactic_2017}
+Forgan D.~H., 2017{\natexlab{b}}, Int. J. Astrobiol., 16, 349
+
+\bibitem[{Forgan(2019)}]{forgan_exoplanet_2017}
+Forgan D.~H., 2019, Int. J. Astrobiol., 18, 189
+
+\bibitem[{Forgan \& Rice(2010)}]{forgan_numerical_2010}
+Forgan D.~H., Rice K., 2010, Int. J. Astrobiol., 9, 73
+
+\bibitem[{Frank(2009)}]{frank_common_2009}
+Frank S.~A., 2009, {The common patterns of nature} Journal of Evolutionary Biology, 22: 1563-1585. doi:10.1111/j.1420-9101.2009.01775.x
+
+\bibitem[{Funes {et~al}\mbox{.}(2019)Funes, Florio, Lares, \&
+  Asla}]{funes_searching_2019}
+Funes J.~G., Florio L., Lares M., Asla M., 2019, Theol. Sci., 0, 1
+
+\bibitem[{Galera {et~al}\mbox{.}(2019)Galera, Galanti, \&
+  Kinouchi}]{galera_invasion_2019}
+Galera E., Galanti G.~R., Kinouchi O., 2019, Int. J. Astrobiol., 18, 316
+
+\bibitem[{Glade {et~al}\mbox{.}(2012)Glade, Ballet, \&
+  Bastien}]{glade_stochastic_2011}
+Glade N., Ballet P., Bastien O., 2012, Int. J. Astrobiol., 11, 103
+
+\bibitem[{Gobat \& Hong(2016)}]{gobat_evolution_2016}
+   Gobat R., Hong S.~E., 2016, A\&A, 592
+
+\bibitem[{Gonzalez(2005)}]{gonzalez_habitable_2005}
+Gonzalez G., 2005, Orig. Life Evol. Biosph., 35, 555
+
+\bibitem[{Gonzalez {et~al}\mbox{.}(2001)Gonzalez, Brownlee, \&
+  Ward}]{gonzalez_galactic_2001}
+Gonzalez G., Brownlee D., Ward P., 2001, Icarus, 152, 185
+
+\bibitem[{Grimaldi(2017)}]{grimaldi_signal_2017}
+Grimaldi C., 2017, Sci. Rep., 7, 46273
+
+\bibitem[{Grimaldi {et~al}\mbox{.}(2018)Grimaldi, Marcy, Tellis, \&
+  Drake}]{Grimaldi2018}
+Grimaldi C., Marcy G.~W., Tellis N.~K., Drake F., 2018, PASP, 130, 987, 054101
+
+\bibitem[{Haqq-Misra(2019)}]{haqq-misra_evolution_2019}
+Haqq-Misra J., 2019, Astrobiology, 19, 10, 1292
+
+\bibitem[{Haqq-Misra \& Kopparapu(2018)}]{haqq-misra_drake_2017}
+Haqq-Misra J., Kopparapu R.~K., 2018, preprint arXiv:1705.07816
+
+\bibitem[{Harp {et~al}\mbox{.}(2018)Harp, Ackermann, Astorga, Arbunich,
+  Barrios, Hightower, Meitzner, Barott, Nolan, Messerschmitt, Vakoch, Shostak,
+  \& Tarter}]{harp_application_2018}
+Harp G.~R. {et~al.}, 2018, ApJ, 869, 66
+
+\bibitem[{Hart(1975)}]{hart_explanation_1975}
+Hart M.~H., 1975, Quarterly Journal of the Royal Astronomical Society, 16, 128
+
+\bibitem[{Hinkel {et~al}\mbox{.}(2019)Hinkel, Hartnett, Lisse, \&
+  Young}]{hinkel_interdisciplinary_2019}
+Hinkel N., Hartnett H., Lisse C., Young P., 2019, Bull. Am. Astron. Soc., 51,
+  497
+
+\bibitem[{Hippke(2017)}]{hippke_interstellar_2017}
+Hippke, M., 2018. Benchmarking information carriers. Acta Astronautica, 151, pp.53-62.
+
+\bibitem[{Horvat(2006)}]{horvat_calculating_2007}
+Horvat M., 2006, Int. J. Astrobiol., 5, 143
+
+\bibitem[{Horvat {et~al}\mbox{.}(2011)Horvat, Naki{\'{c}}, \&
+  Oto{\v{c}}an}]{horvat_impact_2011}
+Horvat M., Naki{\'{c}} A., Oto{\v{c}}an I., 2011, Int. J. Astrobiol., 11, 51
+
+\bibitem[{{Jazayeri}(2007)}]{jazayeri_engeneering_2007}
+{Jazayeri} M., 2007, in Future of Software Engineering (FOSE '07), pp. 199--213
+
+\bibitem[{Jeong {et~al}\mbox{.}(2000)Jeong, Tombor, Albert, Oltvai, \&
+  Barabasi}]{jeong_large_2000}
+Jeong H., Tombor B., Albert R., Oltvai Z.~N., Barabasi A.~L., 2000, Nature,
+  407, 651
+
+\bibitem[{Lampton(2013)}]{lampton_information_2013}
+Lampton M., 2013, Int. J. Astrobiol., 12, 312
+
+\bibitem[{Lineweaver {et~al}\mbox{.}(2004)Lineweaver, Fenner, \&
+  Gibson}]{lineweaver_galactic_2004}
+Lineweaver C.~H., Fenner Y., Gibson B.~K., 2004, Science, 303, 59
+
+\bibitem[{Lingam \& Loeb(2018)}]{lingam_relative_2019}
+Lingam M., Loeb A., 2018, Astrobiology, 19, 28
+
+\bibitem[{Loeb {et~al}\mbox{.}(2016)Loeb, Batista, \&
+  Sloan}]{loeb_relative_2016}
+Loeb A., Batista R.~A., Sloan D., 2016, J. Cosmol. Astropart. Phys., 2016, 040
+
+\bibitem[{Loeb \& Zaldarriaga(2007)}]{loeb_eavesdropping_2006}
+Loeb A., Zaldarriaga M., 2007, J. Cosmol. Astropart. Phys., 1, id. 020
+
+\bibitem[{Maccone(2010)}]{maccone_KLT_2010}
+Maccone C., 2010, Acta Astronaut., 67, 1427
+
+\bibitem[{Maccone(2011)}]{maccone_mathematical_2011}
+Maccone C., 2011, Orig. Life Evol. Biosph., 41, 609
+
+\bibitem[{MacCone(2011)}]{maccone_SETI_2011}
+MacCone C., 2011, Acta Astronaut., 68, 63
+
+\bibitem[{MacCone(2013)}]{maccone_SETI_2013}
+MacCone C., 2013, Int. J. Astrobiol., 12, 218
+
+\bibitem[{Maccone(2014{\natexlab{a}})}]{maccone_evolution_2014}
+Maccone C., 2014{\natexlab{a}}, Int. J. Astrobiol., 13, 290
+
+\bibitem[{Maccone(2014{\natexlab{b}})}]{maccone_lognormals_2014}
+Maccone C., 2014{\natexlab{b}}, Acta Astronaut., 105, 538
+
+\bibitem[{Maccone(2015)}]{maccone_statistical_2015}
+Maccone C., 2015, Acta Astronaut., 115, 277
+
+\bibitem[{Mart{\'{i}}n \& Goldenfeld(2006)}]{martin_origin_2006}
+Mart{\'{i}}n H.~G., Goldenfeld N., 2006, Proc. Natl. Acad. Sci. U. S. A., 103,
+  10310
+
+\bibitem[{Miller \& Maloney(1963)}]{miller_testing_1963}
+Miller J.~C., Maloney C.~J., 1963, Commun. ACM, 6, 58–63
+
+\bibitem[{Mitzenmacher(2004)}]{mitzenmacher_brief_2004}
+Mitzenmacher M., 2004, Internet Math., 1, 226
+
+\bibitem[{Morrison \& Gowanlock(2015)}]{morrison_extending_2015}
+Morrison I.~S., Gowanlock M.~G., 2015, Astrobiology, 15, 683
+
+\bibitem[{Murante {et~al}\mbox{.}(2015)Murante, Monaco, Borgani, Tornatore,
+  Dolag, \& Goz}]{murante_simulating_2015}
+Murante G., Monaco P., Borgani S., Tornatore L., Dolag K., Goz D., 2015, Mon.
+  Not. R. Astron. Soc., 447, 178
+
+\bibitem[{Newman(2005)}]{newman_power_2005}
+Newman M., 2005, Contemp. Phys., 46, 323
+
+\bibitem[{Newman \& Sagan(1981)}]{newman_galactic_1981}
+Newman W.~I., Sagan C., 1981, Icarus, 46, 293
+
+\bibitem[{Oliphant(2006)}]{oliphant_numpy_2006}
+Oliphant T.~E., 2006, A guide to NumPy, Vol.~1. Trelgol Publishing USA
+
+\bibitem[{Peters(2018)}]{peters_outer_2018}
+Peters T., 2018, Int. J. Astrobiol., 17, 282
+
+\bibitem[{Prantzos(2013)}]{prantzos_joint_2013}
+Prantzos N., 2013, Int. J. Astrobiol., 12, 246
+
+\bibitem[{Ptolemaeus(2014)}]{ptolemaeus_system_2014}
+Ptolemaeus C., 2014, {System Design, Modeling, and Simulation. Using Ptolemy
+  II}. Ptolemy.org, Berkeley, p. 690
+
+\bibitem[{Rahvar(2017)}]{rahvar_cosmic_2016}
+Rahvar S., 2017, Mon. Not. R. Astron. Soc., 470, 3095
+
+\bibitem[{Ramirez {et~al}\mbox{.}(2017)Ramirez, G{\'{o}}mez-Mu{\~{n}}oz,
+  V{\'{a}}zquez, \& N{\'{u}}{\~{n}}ez}]{ramirez_new_2017}
+Ramirez R., G{\'{o}}mez-Mu{\~{n}}oz M.~A., V{\'{a}}zquez R., N{\'{u}}{\~{n}}ez
+  P.~G., 2017, Int. J. Astrobiol., 1, 34
+
+\bibitem[{Ross(2012)}]{ross_simulation_2012}
+Ross S.~M., 2012, {Simulation}. Elsevier Science Publishing Co Inc.
+
+\bibitem[{Simkin \& Roychowdhury(2006)}]{simkin_theory_2006}
+Simkin M.~V., Roychowdhury V.~P., 2006, J. Math. Sociol., 30, 33
+
+\bibitem[{Smith(2009)}]{smith_broadcasting_2009}
+Smith R.~D., 2009, Int. J. Astrobiol., 8, 101
+
+\bibitem[{Solomonides {et~al}\mbox{.}(2016)Solomonides, Kaltenegger, \&
+  Terzian}]{solomonides_probabilistic_2016}
+Solomonides E., Kaltenegger L., Terzian Y., 2016, 228th AAS, San Diego, 228, 1
+
+\bibitem[{Sornette(2006)}]{sornette_critical_2006}
+Sornette D., 2006, {Critical phenomena in natural sciences: Chaos, fractals,
+  selforganization and disorder: Concepts and tools}. Springer
+
+\bibitem[{Sotos(2019)}]{Sotos_biotechnology_2019}
+Sotos, J. (2019) Int. J. Astrobiol., 18(5), 445
+
+\bibitem[{Starling \& Forgan(2014)}]{starling_virulence_2013}
+Starling J., Forgan D.~H., 2014, Int. J. Astrobiol., 13, 45
+
+\bibitem[{Tarter(2001)}]{tarter_search_2001}
+Tarter J., 2001, Annu. Rev. Astron. Astrophys., 39, 511
+
+\bibitem[{Tarter {et~al}\mbox{.}(2009)Tarter, Backus, Blair, Cordes, Harp,
+  Henry, Horowitz, Howard, Kilsdonk, Korpela, Lasio, Levin, Shostack, \&
+  Werthimer}]{tarter_advancing_2009}
+Tarter J. {et~al.}, 2009, Astro2010 Astron. Astrophys. Decad. Surv. Sci. White
+  Pap. no. 294, 2010
+
+\bibitem[{Vakoch \& Dowd(2015)}]{vakoch_drake_2015}
+Vakoch D.~A., Dowd M.~F., 2015, {The drake equation: Estimating the prevalence
+  of extraterrestrial life through the ages}, Vakoch D.~A., Dowd M.~F., eds.
+  Cambridge University Press, Cambridge, pp. 1--319
+
+\bibitem[{{van der Walt} {et~al}\mbox{.}(2011){van der Walt}, {Colbert}, \&
+  {Varoquaux}}]{vanderwalt_numpy_2011}
+{van der Walt} S., {Colbert} S.~C., {Varoquaux} G., 2011, Computing in Science
+  Engineering, 13, 22
+
+\bibitem[\protect\citeauthoryear{Anchordoqui, Weber \& Fernandez
+   Soriano}{2017}]{anchordoqui_is_2017} Anchordoqui L.~A., Weber S.~M., Fernandez Soriano J., 2017, ICRC, 301, 254, ICRC...35
+
+\bibitem[{Vukoti{\'{c}} \&
+  {\'{C}}irkovi{\'{c}}(2012)}]{vukotic_astrobiological_2012}
+Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2012, Orig. Life Evol. Biosph.,
+  42, 347
+
+\bibitem[{Vukoti{\'{c}} {et~al}\mbox{.}(2016)Vukoti{\'{c}}, Steinhauser,
+  Martinez-Aviles, {\'{C}}irkovi{\'{c}}, Micic, \&
+  Schindler}]{vukotic_grandeur_2016}
+Vukoti{\'{c}} B., Steinhauser D., Martinez-Aviles G., {\'{C}}irkovi{\'{c}}
+  M.~M., Micic M., Schindler S., 2016, MNRAS, 459, 3512
+
+\bibitem[{Walters {et~al}\mbox{.}(1980)Walters, Hoover, \&
+  Kotra}]{walters_interstellar_1980}
+Walters C., Hoover R.~A., Kotra R., 1980, Icarus, 41, 193
+
+\bibitem[{Wright {et~al}\mbox{.}(2015)Wright, Cartier, Zhao, Jontof-Hutter, \&
+  Ford}]{wright_theGsearch_2015}
+Wright J.~T., Cartier K. M.~S., Zhao M., Jontof-Hutter D., Ford E.~B., 2015,
+  ApJ, 816, 17
+
+\bibitem[{Wright {et~al}\mbox{.}(2018)Wright, Kanodia, \&
+  Lubar}]{wright_how_2018}
+Wright J.~T., Kanodia S., Lubar E., 2018, AJ, 156, 260
+
+\end{thebibliography}
 

### 2020-05-30 submit_v03/manuscript.tex
last day

@@
 
-\title[Probabilities of SETI contacts]{Probability of causal
-contacts between interstellar civilizations through Monte Carlo simulations}
-%
-%A network of constrained causal contacts in the Galaxy as a proxy for
-%the probability of 
-%
-% A model for constrained causal contacts between location in the GHZ
-
-\author[Lares, Funes \& Gramajo]{Lares M.$^{1, 2}$, Funes J.~G.$^{1,
-3}$ \& Gramajo L.$^{1, 2}$}
+\title[Probability of causal contacts]{Monte Carlo estimation of the probability 
+of causal contacts between communicating civilisations}
+
+\author[Lares, Funes \& Gramajo]{Lares M.$^{1, 2}$, Funes J.~G.$^{1, 3}$ \& Gramajo L.$^{1, 2}$}
 
@@
 
-%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
 \section{Introduction}\label{S_motivations}
@@
 %
-%We stress the fact that the lack of detections of extraterrestrial
-%intelligent signals is the only observational fact that can be taken
-%for sure.
 The analysis of these scenarios is difficult due to the lack of data
@@
 intelligence (SETI), we have not received any signal yet.
-%
-%Indeed, for the number of years projects for the Search of
-%Extraterrestrial Intelligences (SETI) have been working, no signal has
-%been received using the technical possibilities within their search
-%criteria \citep{tarter_search_2001}.
 %
@@
 
-The discussion about the problem of the (unknown)
+The discussion about the problem of the unknown
 abundance of civilisations in the Galaxy has been organised around the
@@
 computing the estimated number of extraterrestrial intelligences.
-%
-%The optimistic estimations from the Drake equation also contrast with
-%the results from the SETI initiatives.
-%
-%Notably, despite decades of effort not a single evidence for
-%intelligent signals has been detected.
 %
@@
 %
-The literature about this topic brings also a number of works that consider
+On this topic a number of works consider
 numerical simulations \citep{forgan_evaluating_2015,
@@
 
-A notable example of statistica modeling is the work by
+A notable example of statistical modeling is the work by
 \citet{balbi_impact_2018}, who investigates the chance of
@@
 %
-This simulations represent the spread if intelligence as an
+These simulations represent the spread if intelligence as an
 implementation of panspermia theories.
@@
 %
-Hereafter, we refeer to this model as the C3Net model, standing for
-''Constrained Causally Conencted Network``.
+Hereafter, we refer to this model as the SC3Net model, standing for
+''Stochastic Constrained Causally Connected Network``.
 
@@
 The simulation is carried out by following all the variables that describe and
-constitute the state of the system, and the evolution of the process is
-described as a set of changes in that state.
+constitute the state of the system.
+%
+The evolution of the process is then
+described as the set of changes in those variables. 
 %
@@
 %
-It is worth noticing that the model is not intended for a formal fit at this
-stage due to the lack of data, but is can help to understand how unusual it
-would be to actively search for intelligente signal for 50 years without
-success.
+It is worth noticing that the SC3Net model is not intended for a formal fit at this
+stage due to the lack of data, but it can help to understand how unusual it
+would be to actively search for intelligent signals for 50 years without
+possitive detections.
 
@@
 directions.
-%
-In this work we use the International System of units recommended by the
-International Astronomial Union, where the simbol for years is ''a''.
 
@@
 %
-The sphere of first contact (SFC) is the sphere centered on the emitter that
+The sphere of first contact (SFC) is the sphere centred on the emitter that
 grows until its radius reaches the D$_{\rm max}$ distance, in which the
@@
 
-%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
 \section{Methods and working hypotheses}\label{S_methods}
@@
 Also, it is worth mentioning that this is a general approach, and not
-necessarily a \ceti{} is the host of an intelligent civilization.
+necessarily a \ceti{} is the host of an intelligent civilisation.
 %
@@
 Alternative message processing entities could be considered, for example
-interstellar beacons where life has ceased to exist but continue with its
+interstellar beacons where intelligent life has ceased to exist but continue with its
 emission, or communication stations established by probes or left by
@@
 taking into account a small number of variables, which allow to analyze the
-variation of the results in the model parameter space.
+variation of the results in the SC3Net model parameter space.
 
@@
 %
-Following the end of the civilization, there is still a region which is filled
+Following the end of the civilisation, there is still a region which is filled
 with the emitted signals.
@@
 The probabilities of contacts presented here are then upper limits to the
-number of communications between civilizations in the Galaxy, given the
+number of communications between civilisations in the Galaxy, given the
 implemented hipotheses in the model.
@@
 %
-% It is worth mentioning that we are considering in our experimental setup a
-% system composed by several emitters and  receivers across the galaxy, under
-% the hypothesis proposed before.
-%
 In such a system we compute probabilities of a random node making contact with
@@
 Regarding the extent of the signals, we know that the distance from which a
-signal from Earth could be detected using the current technology is of a few
+signal from Earth could be detected using the current technology is about few
 parsecs, given that the signal was sent to a specific direction.
@@
 larger uncertainty.
-
-% For the communication of \cetis{} through messages sent isotropically, we assume that the capacity to emit signals and %to receive signals occur at the same time.
-This means that the ability to find another civilization develops at the same %time that the ability and intention to emit a signal in all directions, intended to be detected by an unknown civilization.
-
-%https://setiathome.berkeley.edu/forum_thread.php?id=80585&postid=1833000#1833000]
-%https://www.engadget.com/2017/05/25/listening-to-starlight-our-ongoing-search-for-alien-intelligenc/]
-%[from https://en.wikipedia.org/wiki/Fermi_paradox]
-%
-%
-%This proposal is based on a pronoid scenario, and we do not consider
-%in this work the occurrence of paranoid or partially paranoid
-%civilizations.
-%%
-%If that would be the case, the results obtained here can be taken as
-%upper limits.
 
@@
 results are independent of whether intelligent agents are organic or
-artificial. Moreover, the causes of the limited lifetime of a
-civilization can be natural (astrophysical phenomena), caused by auto
-destruction or by external factors, to name a few. However, we assume
+artificial.
+%
+Moreover, the causes of the limited lifetime of a
+civilisation can be natural (astrophysical phenomena), caused by auto
+destruction or by external factors, to name a few.
+%
+However, we assume
 that these events are sufficiently numerous in the Galaxy so that a
@@
    $\tau_{a}$ & Mean temporal separation between consecutive
-   awakenings & 4000 yr & 200\,000 yr & 50\\ 
+   awakenings & 4 kyr & 200 kyr & 50\\ 
    $\tau_{s}$ & Mean lifetime of a \ceti{}
-   & 10000 yr & 500\,000 yr& 50\\ 
+   & 10 kyr & 500 kyr& 50\\ 
    D$_{max}$ & Maximum reach of a message  &  \multicolumn{2}{l}{3 kpc, 12 kpc, 25 kpc} & 3\\
@@
 given by the size and shape of the Galactic Habitable Zone and the maximum
-distance a signal can travel to be detected (D$_{\rm max}$). The parameters for
+distance a signal can travel to be detected (D$_{\rm max}$).
+%
+The parameters for
 the temporal distributions also determine the spatial properties, since the
-density of active \cetis{} in the galaxy depend on these two parameters. For
+density of active \cetis{} in the galaxy depend on these two parameters.
+%
+For
 example, small $\tau_a$ and a large $\tau_s$ will produce a densely
@@
 distributions of the temporal parameters must be made in order to complete the
-simulation. \citet{forgan_spatiotemporal_2011} argues that the times at which
+simulation.
+%
+\citet{forgan_spatiotemporal_2011} argues that the times at which
 different civilisations become intelligent follow a Gaussian distribution, and
-then the distribution of inter-arrival times is an inverse exponential. We
+then the distribution of inter-arrival times is an inverse exponential.
+%
+We
 assume that the distribution of the times of A events is a stationary Poisson
 process, and then the distribution of the times between the appearance of new
-consecutive nodes is exponential. Regarding the duration of a node, we
+consecutive nodes is exponential.
+%
+Regarding the duration of a node, we
 propose that its distribution is a stationary
-exponential distribution.  There are no clear arguments to conclude a
+exponential distribution.
+%
+There are no clear arguments to conclude a
 statistical law for the later distribution, so that we propose it as a working
@@
 %
-A correction by coverage ratio in the detection and by a targeting ratio in the
+A correction by a coverage ratio in the detection and by a targeting ratio in the
 emission could be easily implemented in the simulation, although the effect of
@@
 
-%%% comparar con el stochsatic cellular automata
-% ver libro Solving Fermi paradox, sec. 27.3.1
-
 \begin{figure*} % 2D color plot
@@
    %
-Rate of MPLs that never make contact (listening) as a
+Rate of nodes that never make contact (listening) as a
 function of $\tau_a$ and $\tau_s$, for 
@@
    %
-Rate of MPLs that listen at the moment of the
+Rate of nodes that listen at the moment of the
 awakening, as a funcion of $\tau_a$ and $\tau_s$, for
@@
 %
-We consider a first approach to the galactic habitable zone as a 2-dimensional
+We consider a first approach to the galactic habitable zone as a two-dimensional
 annular region.
@@
 %\ceti{}, $\tau_a$, and the maximum distance a message can be detected.
-
+%
 % The Galactic Habitable Zone is set at fixed values of the inner and outer
@@
 %
-In order to achieve high performance we made use of vectorised methods
-and array broadcasting properties provided by the NumPy package
-\citep{oliphant_numpy_2006, vanderwalt_numpy_2011}
-%
-The combination of C-compiled Numpy routines for numerical
-computations with a high-level language allows to obtain an easily
-comprehensible and maintainable code.
-%
+% In order to achieve high performance we made use of vectorised methods
+% and array broadcasting properties provided by the NumPy package
+% \citep{oliphant_numpy_2006, vanderwalt_numpy_2011}
+% %
+% The combination of C-compiled Numpy routines for numerical
+% computations with a high-level language allows to obtain an easily
+% comprehensible and maintainable code.
+% %
 From the viewpoint of a user, \hs{} is an object-oriented package that
 exposes the main functionalities as classes and methods.
+%
+The code fulfills standar quality assurance metrics, that account for
+testing, style, documentation and coverage.
 %
@@
 
-%Quality assurance
-%
-Regarding the quality assurance of the software, we endorse the PEP-8 style
-guide for python (van Rossum et al. 2001) throughout the entire project, which
-improves the code maintainability and helps to minimise the so called
-``code-entropy'' of future versions.
-%
-The deviations in style were analysed with the flake8 tool, which improves the
-readability and consistency of the code.
-%
-Testing and code coverage have been applied to the package and the development
-process as standard qualitative and quantitative metrics.
-%
-The different parts and functions can be validated to assert if they return the
-expected results on specific test cases and testing scenarios
-\citep{jazayeri_engeneering_2007}.
-%
-On the other hand, code coverage measures the fraction visited and executed in
-the code by the testing procedure \citep{miller_testing_1963}.
-%
-The advantage of an exhaustive code-coverage is that it ensures that most of
-the sources of errors have been considered.
-%
-In the \hs{} project we provide test tools for the distinct parts of the
-simulation setup and run, reaching a 99\% of code coverage in the version used
-for this work.
-%
-For every new version of the code, all these quality assurance methods and
-metrics are run automatically by a continuous-integration service.
-%
-It is possible to reproduce the experiments performed in this work using the
-\hs{} package.
-%
-The installation is available through a git repository and also on the Python
-Package Index (PyPI).
-%
+% %Quality assurance
+% %
+% Regarding the quality assurance of the software, we endorse the PEP-8 style
+% guide for python (van Rossum et al. 2001) throughout the entire project, which
+% improves the code maintainability and helps to minimise the so called
+% ``code-entropy'' of future versions.
+% %
+% The deviations in style were analysed with the flake8 tool, which improves the
+% readability and consistency of the code.
+% %
+% Testing and code coverage have been applied to the package and the development
+% process as standard qualitative and quantitative metrics.
+% %
+% The different parts and functions can be validated to assert if they return the
+% expected results on specific test cases and testing scenarios
+% \citep{jazayeri_engeneering_2007}.
+% %
+% On the other hand, code coverage measures the fraction visited and executed in
+% the code by the testing procedure \citep{miller_testing_1963}.
+% %
+% The advantage of an exhaustive code-coverage is that it ensures that most of
+% the sources of errors have been considered.
+% %
+% In the \hs{} project we provide test tools for the distinct parts of the
+% simulation setup and run, reaching a 99\% of code coverage in the version used
+% for this work.
+% %
+% For every new version of the code, all these quality assurance methods and
+% metrics are run automatically by a continuous-integration service.
+% %
+% It is possible to reproduce the experiments performed in this work using the
+% \hs{} package.
+% %
+% The installation is available through a git repository and also on the Python
+% Package Index (PyPI).
+% %
 Since the simulation setup is configurable, the time required for a simulation
@@
 the fraction of the lifetime a node is listening to at least another node
-(i.e., in their causal cone),
+(i.e., within their light cone),
 %
@@
 parameters.
-%
 
@@
 %
-A couple of comments are worth to mention about this plot, namely, it
+We should point out that this plot 
 has a logarithmic scale on the x-axis, and it is the cumulative, not
-differential, empirical distribution, so that the differences in the
+differential, empirical distribution.
+%
+Therefore, the differences in the
 number of contacts for different models are relatively large.
@@
 This allows to consider the fact that for the nearly 50 years SETI
-programs have been active, no contact event has been reported.
+programs have been active, no contact event has been reported yet.
 %
@@
 is a maximum probability of making the first causal contact at nearly
-100000 years.
+10$^5$ years.
 %
@@
 
-%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
 \section{Discussion}\label{S_discussion}
 
-We presented a stochastic model (SC3Net) to analyse the network of
+We have presented a stochastic model (SC3Net) to analyse the network of
 constrained causally connected nodes in a simplified Galaxy.
@@
 %
-Our analysis is not centered in obtaining the odds for the Earth
-to make contact with another intelligent civilization.
+Our analysis is not centred in obtaining the odds for the Earth
+to make contact with another civilization.
 %
@@
 Our results are also consistent with those presented by
-\citet{grimaldi_signal_2017}, who estimate an upper bound for the mean
+\citet{grimaldi_signal_2017}, who estimates an upper limit for the mean
 number of extraterrestrial civilizations that could contact Earth
@@
 
-We also fond that there 
+We have also fond that there 
 is a balance between the density of \cetis{} and the mean
@@
 Also, the chances of entering the causal connected zone of a \ceti{}
-does not grow linearly, but favours the first period of 100000 years.
+does not grow linearly, but favours the first period of 10$^5$ years.
 
@@
 the problem of the probabilities of causal contacts between locations
-in the galaxy with the possibility of sending and receiving messages.
+in the Galaxy with the possibility of sending and receiving messages.
 %

### 2020-05-30 submit_v03/manuscript.tex
last changes

@@
 
-%The authors can define any packages after the \documentclass{CSL} command.
-\usepackage{amsmath} %for dealing with mathematics,
-\usepackage{amsthm} %for dealing with theorem environments,
-\usepackage{cite} %%%for dealing with citations
-\usepackage{hyperref} %for linking the cross references
-\usepackage{graphics} %for dealing with figures.
-\usepackage{algorithmic} %for describing algorithms
-\usepackage{subfig} %for getting the subfigures e.g., "Figure 1a and 1b" etc.
-\usepackage{url} %It provides better support for handling and breaking URLs.
-
-%\usepackage{natbib}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{cite}
+\usepackage{hyperref}
+\usepackage{graphics}
+\usepackage{algorithmic}
+\usepackage{subfig}
+\usepackage{url}
+
 \usepackage{boites}
-
-%\usepackage{graphicx}
-%\usepackage{float}
-%\usepackage{psfig}
-%\usepackage{amsfonts,amssymb}
-%\usepackage{color}
-%\usepackage{textcomp}
-%\usepackage[T1]{fontenc}
-%\usepackage{aecompl}
-%\usepackage{times}
-%\usepackage{txfonts}
-%\usepackage[english]{babel}
-%\usepackage[utf8x]{inputenc}
-%\usepackage{amsmath}
-%\usepackage{caption}
-%\usepackage{siunitx}
 
@@
 
-\newcommand{\ceti}{C3N}
-\newcommand{\cetis}{C3Ns}
+\newcommand{\ceti}{node}
+\newcommand{\cetis}{nodes}
 \newcommand{\hs}{\textsc{hearsay}}
@@
 equation-- through which to look at future contacts with
-technologically advanced civilizations in the Milky Way. 
-%
-The equation quantifies the number of civilizations from whom we might
+technologically advanced civilisations in the Milky Way. 
+%
+The equation quantifies the number of civilisations from whom we might
 receive an electromagnetic signal, using a collection of factors that
-have been extensively discussed in the literature and whose estimates
-are revised continually. 
-%
-A comprehensive review and analysis of each term of the equation are
+have been extensively discussed in the literature and whose estimated
+values are revised continually. 
+%
+A comprehensive review and an analysis of each term of the equation are
 presented in \citet{vakoch_drake_2015}.
@@
 The large distances in the Galaxy and the likely limited lifetime of
-civilizations may play an important role in determining how difficult it
+civilisations may play an important role in determining how difficult it
 would be to obtain evidence for other inhabited worlds.
@@
 %
-Indeed, as \citep{tarter_search_2001} pointed out, according to our
+Indeed, as \citet{tarter_search_2001} pointed out, according to our
 current technical capabilities for the search of extraterrestrial
@@
 
-The discussion about the problem of the unknown
-abundance of civilisations in the Galaxy has been organised around the
-factors in the Drake equation \citep{hinkel_interdisciplinary_2019}.
+The discussion about the problem of the unknown abundance of
+civilisations in the Galaxy has been organised around the factors in
+the Drake equation \citep{hinkel_interdisciplinary_2019}.
 %
 However, the uncertainties in these factors, specially the ones
-representing biological processes, make it
-less suitable to a formal study with the purpose of defining
-searching strategies or
-computing the estimated number of extraterrestrial intelligences.
-%
-A number of studies propose alternative formalisms for the estimation of the
-likelihood of detecting intelligent signals from space.
-%
-\citet{prantzos_joint_2013}, for example, proposes a unified framework for a joint
-analysis of the Drake equation and the Fermi paradox, concluding that
-for sufficiently long-lived civilisations, colonisation
-is the most promising strategy to find other life forms.
+representing biological processes, make it less suitable to a formal
+study with the purpose of defining searching strategies or computing
+the estimated number of extraterrestrial intelligences.
+%
+A number of studies propose alternative formalisms for the estimation
+of the likelihood of detecting intelligent signals from space.
+%
+\citet{prantzos_joint_2013}, for example, proposes a unified framework
+for a joint analysis of the Drake equation and the Fermi paradox,
+concluding that for sufficiently long-lived civilisations,
+colonisation is the most promising strategy to find other life forms.
 %
@@
 Some modifications to the original idea of the Drake equation have
-been proposed, in order to
-integrate a temporal structure, to
+been proposed, in order to integrate a temporal structure, to
 reformulate it as a stochastic process or to propose alternative
@@
 equation \citep{glade_stochastic_2011}.
-%
-On this topic a number of works consider
-numerical simulations \citep{forgan_evaluating_2015,
-vukotic_grandeur_2016, murante_simulating_2015, forgan_numerical_2009,
-forgan_galactic_2017, ramirez_new_2017}.
-%
-Although this approach does not rely on values that can be measured
-from observations
-(like the fraction of stars with planets), 
-it depends on the definition of
-unknown or uncertain parameters required to carry out
-a the simulations.
-%
-\citet{grimaldi_signal_2017} considers a statistical model for the
-probability of the Earth contacting other intelligent civilisations,
-taking into account the finite lifetime of signal emitters, and based
-on the fractional volume occupied by all signals reaching our planet.
 
@@
 program to succeed in the detection of intelligent signals.
+% 
+\citet{balbi_impact_2018} also investigates the chance of
+communicating civilisations making causal contact within a volume
+surrounding the location of the Earth.
+%
+The author argues that the causal contact requirement involves mainly
+the distance between civilisations, their lifespan and their times of
+appearance.
+%
+This is important since the time the light takes to travel across the
+Galaxy might be much lesser than the lifetime of the emitter.
+%
+\citet{balbi_impact_2018} fixes the total number of civilisations and
+explores the parametric space that comprises three variables, namely,
+the distance to the Earth, the time of appearance and the lifespan of
+the communicating civilisations.
+%
+Each of these three variables are drawn from a random distribution.
+%
+The distances are drawn from a uniform model 
+for the
+positions of civilisations within the plane of the Galaxy.
+%
+For the distribution of the characteristic time of appearance, the
+author explores exponential and truncated Gaussian functions,
+while for the lifespans chooses an exponential behaviour.
+%
+It is important to point out that the estimation of the number of
+communicating civilisations vary with the choice of the statistical
+model for the time of appearance, as shown by
+\citet{balbi_impact_2018}.
+%
+For all analysed distributions, the author concludes that the
+fraction of emitters that are listened is low if they are spreaded in
+time and with limited lifetimes.
+%
+An analytical explanation of these concepts are presented 
+in \citet{grimaldi_signal_2017}, who considers a statistical model for the
+probability of the Earth contacting other intelligent civilisations,
+taking into account the finite lifetime of signal emitters, and based
+on the fractional volume occupied by all signals reaching our planet.
 
@@
 
-A notable example of statistical modeling is the work by
-\citet{balbi_impact_2018}, who investigates the chance of
-communicating civilisations making causal contact within a volume
-surrounding the location of the Earth.
-%
-The author argues that the causal contact requirement involves mainly
-the distance between civilisations, their lifespan and their times of
-appearance.
-%
-This is important since the time the light takes to travel across the
-Galaxy might be much lesser than the lifetime of the emitter.
-%
-\citet{balbi_impact_2018} fixes the total number of civilisations and
-explores the parametric space that comprises three parameters, namely,
-the distance to the Earth, the time of appearance and the lifespan of
-the communicating civilization.
-%
-Each of these three variables are drawn from a random distribution.
-%
-The distribution of the distances results from a uniform model for the
-positions of civilizations within the plane of the Galaxy.
-%
-For the distribution of the characteristic time of appearance, the
-author explores exponential and truncated Gaussian distributions,
-while for the distribution of the lifespans chooses an exponential
-distribution.
-%
-It is important to consider that the estimation of the number of
-communicating civilisations vary with the choice of the statistical
-distribution for the time of appearance, as shown by
-\citet{balbi_impact_2018}.
-%
-For all analysed distributions, the author concludes that, the
-fraction of emitters that are listened is low if they are spreaded in
-time and with limited lifetimes.
+On this topic a number of works consider numerical simulations
+\citep{forgan_evaluating_2015, vukotic_grandeur_2016,
+murante_simulating_2015, forgan_numerical_2009, forgan_galactic_2017,
+ramirez_new_2017}.
+%
+Although this approach does not rely on values that can be measured
+from observations (like the fraction of stars with planets), it
+depends on the definition of unknown or uncertain parameters required
+to carry out the simulations.
 %
@@
 The authors implement this model to a network of cells which represent
-life complexity on a 2-dimensional region resembling the Galactic
+life complexity on a two--dimensional region resembling the Galactic
 Habitable Zone (GHZ), an annular ring set between a minimal radius of
-6 kpc and a peak radius of 10 kpc.
-%
-These simulations represent the spread if intelligence as an
+6~kpc and a peak radius of 10~kpc.
+%
+These simulations represent the spread of intelligence as an
 implementation of panspermia theories.
@@
 Within this framework, \citet{vukotic_astrobiological_2012} also make
-several Monte Carlo simulations and analyse ensemble-averaged results.
+Monte Carlo simulations and analyse ensemble-averaged results.
 %
@@
 %
-With these tools, we address the problem of the temporal and spatial structure of
-the distribution of communicating civilisations.
-%
-This approach does not require asumptions about, e.g., the origin of life, the development of
-intelligence or the formation of habitable planets according to stellar type.
-%
-Instead, we assume monoparametric function families to model the appearance of
-points in the disk of the Galaxy and over the time, which we call
-\textit{nodes}.
-%
-These nodes represent the locations of ideal intelligent agents that are able
-to receive and emit signals with perfect efficiency in all directions.
-%
-Then we analyse the network of causally connected nodes limited to have a
-maximum separation representing the maximum distance a signal can travel with
-an intensity above a fixed threshold.
-%
-A node is causally connected to other node if it is within their light cone.
-%
-We adopt as a definition that a light cone is the region in spacetime within
-the surface generated by the light emanating from a given point in space in a
-given period of time.
-%
-For example, two nodes separated by 100 light years which appear with a
-difference in time of 80 years and last 50 years each, will be causally
-connected for 30 years.
+With these tools, we address the problem of the temporal and spatial
+structure of the distribution of communicating civilisations.
+%
+This approach does not require asumptions about, e.g., the origin of
+life, the development of intelligence or the formation of habitable
+planets according to stellar type.
+%
+Instead, we assume monoparametric function families to model the
+appearance of points in the disk of the Galaxy and over the time,
+which we call \textit{nodes}.
+%
+These nodes represent the locations of ideal intelligent agents that
+are able to receive and emit signals with perfect efficiency in all
+directions.
+%
+Then we analyse the network of causally connected nodes limited to
+have a maximum separation representing the maximum distance a signal
+can travel with an intensity above a fixed threshold.
+%
+A node is causally connected to other node if it is within their light
+cone.
+%
+We adopt as a definition that a light cone is the region in spacetime
+within the surface generated by the light emanating from a given point
+in space in a given period of time.
+%
+For example, two nodes separated by 100 light years which appear with
+a difference in time of 80 years and last 50 years each, will be
+causally connected for 30 years.
 %
@@
 %
-In fact, the second node will see the first one for 30 years, but the first
-node will never acknowledge the existence of the second.
-%
-This simple tought experiment exemplify the importance of the time variable to
-analyse the structure of connections.
-%
-The model can be described with three parameters, namely, the mean separation
-between the appearance of nodes, the mean lifetime of the nodes, and the
-maximum separation to allow an effective causal contact.
-%
-The model presented here comprises the stochastic networks of constrained
-causally connected nodes, together with the three parameters and additional
-hypotheses.
+In fact, the second node will see the first one for 30 years, but the
+first node will never acknowledge the existence of the second.
+%
+An scheme of this example is shown if Fig.~\ref{F_example}.
+%
+This simple tought experiment exemplify the importance of the time
+variable to analyse the structure of connections.
+%
+The model can be described with three parameters, namely, the mean
+separation between the appearance of nodes, the mean lifetime of the
+nodes, and the maximum separation to allow an effective causal
+contact.
+%
+The model presented here comprises the stochastic networks of
+constrained causally connected nodes, together with the three
+parameters and additional hypotheses.
 %
@@
 
-We provide a framework to explore, through a suite of numerical simulations,
-the parameter space of three unknown observables.
-%
-This allows to discuss possible scenarios and their consequences in terms of
-the probability of making contacts.
-%
-The method we use for simulating a stochastic process is an approximation that
-allows to study the behaviour of complex systems, by considering a sequence of
-well defined discrete events.
-%
-The simulation is carried out by following all the variables that describe and
-constitute the state of the system.
-%
-The evolution of the process is then
-described as the set of changes in those variables. 
-%
-In this context, an event produces a specific change in the state that can be
-triggered by random variables that encode the stochastic nature of the physical
-phenomenon.
-%
-For example, when a new contact is produced between two entities in the
-simulated galaxy, the number of active communication lines is increased by one.
-%
-Also, if it is the first contact for that nodes, 
-then the number of communicated nodes increases by two.
-%
-When a new node becomes active, the sistem has an increase of one
-in the number of nodes, although the number of communcations does not
+\begin{figure}[h]
+%
+\centering
+%
+\includegraphics[width=\columnwidth]{F_example.pdf}
+%
+\caption{Example of a configuration of an emitter and a receiver 
+   to produce a causal contact.
+   %
+   The observers are separated by 100~light years and appear with a
+   difference in time of 80~years. One observer is active between
+   $t=$0~yr and $t=$50~yr.  The other observer is also active for
+   50~yr, between $t=$80~yr and $t=$130~yr. 
+   %
+   The causal connection occurs between $t=$100~yr and $t=$130~yr
+   %
+   Meaningfull times are marked with vertical dashed lines.
+   }
+%
+\label{F_example}
+%
+\end{figure}
+
+We provide a framework to explore, through a suite of
+numerical simulations, the parameter space of three unknown
+observables.
+%
+This allows to discuss possible scenarios and their consequences in
+terms of the probability of making contacts.
+%
+The method we use for simulating a stochastic process is an
+approximation that allows to study the behaviour of complex systems,
+by considering a sequence of well defined discrete events.
+%
+The simulation is carried out by following all the variables that
+describe and constitute the state of the system.
+%
+The evolution of the process is then described as the set of changes
+in those variables. 
+%
+In this context, an event produces a specific change in the state that
+can be triggered by random variables that encode the stochastic nature
+of the physical phenomenon.
+%
+For example, when a new contact is produced between two entities in
+the simulated galaxy, the number of active communication lines is
+increased by one.
+%
+Also, if it is the first contact for that nodes, then the number of
+communicated nodes increases by two.
+%
+When a new node becomes active, the sistem has an increase of one in
+the number of nodes, although the number of communcations does not
 necessarily change.
 %
-The process then involves following the changes on the state of the system, defining
-the initial and final states.
-%
-This is done by defining a method that allows to keep track of the time
-progress in steps and maintaining a list of relevant events, i.e., the events
-that produce a change in the vriables of interest.
-%
-With this method we abandon the frequentist approach of the Drake equation to
-compute the number of civilisations, providing instead its statistical
-distribution.
-%
-More importantly, we are interested on the probability of contacts, which
-depend critically on the time variable.
-%
-This is an exploratory analysis that aims at developing a numerical tool to
-discuss the different scenarios based on statistical heuristics.
-%
-The approach proposed here should be considered as a compromise between the
-uncertainties of the frequentist estimations and the detailed recipes required
-on the numerical simulations.
-%
-It is worth noticing that the SC3Net model is not intended for a formal fit at this
-stage due to the lack of data, but it can help to understand how unusual it
-would be to actively search for intelligent signals for 50 years without
-possitive detections.
+The process then involves following the changes on the state of the
+system, defining the initial and final states.
+%
+This is done by defining a method that allows to keep track of the
+time progress in steps and maintaining a list of relevant events,
+i.e., the events that produce a change in the vriables of interest.
+%
+With this method we abandon the frequentist approach of the Drake
+equation to compute the number of civilisations, providing instead its
+statistical distribution.
+%
+More importantly, we are interested on the probability of contacts,
+which depend critically on the time variable.
+%
+This is an exploratory analysis that aims at developing a numerical
+tool to discuss the different scenarios based on statistical
+heuristics.
+%
+The approach proposed here should be considered as a compromise
+between the uncertainties of the frequentist estimations and the
+detailed recipes required on the numerical simulations.
+%
+It is worth noticing that the SC3Net model is not intended for a
+formal fit at this stage due to the lack of data, but it can help to
+understand how unusual it would be to actively search for intelligent
+signals for 50 years without possitive detections.
 
@@
 %
-In Sec.~\ref{S_methods} we introduce the methods and discuss the candidate
-distributions for the statistical aspects of the times involved in the
-communication process.
-%
-We present our results in Sec~\ref{S_results}, with special emphasis on the
-statistical distributions of the duration of causal contacts in one or both
-directions and the distribution of time intervals for the waiting of the first
-contact.
+In Sec.~\ref{S_methods} we introduce the methods and discuss the
+candidate distributions for the statistical aspects of the times
+involved in the communication process.
+%
+We present our results in Sec~\ref{S_results}, with special emphasis
+on the statistical distributions of the duration of causal contacts in
+and the distribution of time intervals of waiting for the first contact.
 %
@@
 Space--time diagrams showing a schematic representation of the different stages
-in the development of a constrained causal contact node (\ceti{}).
+in the development of a constrained causal contact node.
 %
@@
 %
-In the Figure this sphere is represented by the left triangle of the region
+This sphere is represented by the left triangle of the region
 $R$.
@@
 %
-The receiver node E$_1$ can listen signals from the emitter node E$_2$, from
+The node E$_1$ can listen signals from the node E$_2$, since
 the ``Contact'' event (t=C$_{12}$) up to the ``Blackout'' event
@@
 %
-Similarly, the receiver node E$_2$ can listen signals from the emitter
-node E$_1$, from t=C$_{21}$ up to t=B$_{21}$. 
+Similarly, the node E$_2$ is in causal contact from the
+   node E$_1$, in the time interval (C$_{21}$, B$_{21}$).
 %
@@
 This is done in terms of three independent parameters, namely, the mean time
-span between the appearance of consecutive \cetis{}, $\left<\tau_a\right>$,
-their mean lifetime $\left<\tau_s\right>$, and the maximum distance a signal
+span between the appearance of consecutive \cetis{}, $\left<\tau_{\rm
+a}\right>$,
+their mean lifetime $\left<\tau_{\rm s}\right>$, and the maximum distance a signal
 can be detected by another \ceti{} ($D_{\rm max}$).
@@
 Intuitively, the probability of the existence of causal contacts between pairs
-of \cetis{} would be larger for smaller $\tau_a$ parameters, higher $\tau_s$ or
+of \cetis{} would be larger for smaller $\tau_{\rm a}$ parameters,
+higher $\tau_{\rm s}$ or
 higher $D_{\rm max}$ parameter values.
@@
 We also propose theoretical distribution functions for both the lifespans
-($\tau_s$) and the number of \cetis{} per unit time
+($\tau_{\rm s}$) and the number of \cetis{} per unit time
 \citep{maccone_evolution_2014, Sotos_biotechnology_2019}.
@@
 The later is related to the time span between the appearance of consecutive
-\cetis{} (since when $\left<\tau_a\right>$ is shorter, it produces a greater
+\cetis{} (since when $\left<\tau_{\rm a}\right>$ is shorter, it produces a greater
 density of \cetis{}).
@@
 
-\ffn{1} Space—time diagrams, where time and space are represented on
+\ffn{1} Space--time diagrams, where time and space are represented on
 the horizontal and vertical axes, respectively, are suitable tools to
@@
 imply extreme values for the parameters that represent the density of
-the nodes (i.e., $\left< \tau_a \right>$).
+the nodes (i.e., $\left< \tau_{\rm a} \right>$).
 
@@
    &min. value&max. value&Nbins\\
-%\hline
-%   $\tau_{a}$ & Mean temporal separation between consecutive awakenings 
-%	      & 500~a & 40500~a & 21\\ 
-%	    & & 5$\cdot$10$^5$~a & 10$^6$~a & 51\\ 
-%   $\tau_{s}$ & Mean lifetime of a \ceti{}
-%	      & 500~a & 40500~a & 21\\ 
-%	    & & 5$\cdot$10$^5$~a & 10$^6$~a & 51\\ 
-%	D$_{\rm max}$ & Maximum reach of a message  & \multicolumn{3}{c}{100 pc, 500 pc, 1 kpc, 5 kpc, 10 kpc} \\
 \hline
-
-   $\tau_{a}$ & Mean temporal separation between consecutive
-   awakenings & 4 kyr & 200 kyr & 50\\ 
+   $\tau_{a}$ & Mean temporal separation between consecutive awakenings 
+	      & 5$\cdot10^2$~yr & 40500~yr & 21\\ 
+	    & & 5$\cdot$10$^5$~yr & 10$^6$~yr & 51\\ 
    $\tau_{s}$ & Mean lifetime of a \ceti{}
-   & 10 kyr & 500 kyr& 50\\ 
-   D$_{max}$ & Maximum reach of a message  &  \multicolumn{2}{l}{3 kpc, 12 kpc, 25 kpc} & 3\\
+	      & 5$\cdot10^2$~yr & 40500~yr & 21\\ 
+	    & & 5$\cdot$10$^5$~yr & 10$^6$~yr & 51\\ 
+	D$_{\rm max}$ & Maximum reach of a message  & \multicolumn{3}{c}{100 pc, 500 pc, 1 kpc, 5 kpc, 10 kpc} \\
 \hline
-
    \multicolumn{2}{l}{fixed variables} & \multicolumn{2}{c}{assumptions} &value \\
@@
    f$_p$ & panspermia or colonization &absent&&0\\
-   & shape of the Galactic Habitable Zone & 2--dimensional ring &&\\
+   & shape of the Galactic Habitable Zone & two-dimensional ring &&\\
 	R$_{GHZ}^{\rm min}$   & Inner radius of the GHZ  & \citet{lineweaver_galactic_2004} & & 7~kpc\\
    R$_{GHZ}^{\rm max}$   & Outer radius of the GHZ       & \citet{lineweaver_galactic_2004} & & 9~kpc\\
-	t$_{\rm max}$ & Time span of the simulation  & & \multicolumn{2}{r}{10$^6$~yr} \\
+	t$_{\rm max}$ & Time span of the simulation  & &
+   \multicolumn{2}{r}{10$^5$~yr - 10$^9$~yr} \\
     & \multicolumn{3}{l}{number of random realizations for each point
-    in the parameter space} & 50 \\
+    in the parameter space} & 1 - 50 \\
 \hline
@@
 For
-example, small $\tau_a$ and a large $\tau_s$ will produce a densely
+example, small $\tau_{\rm a}$ and a large $\tau_{\rm s}$ will produce a densely
 populated galaxy (in this context, a galaxy is an element in a 
@@
 Empirical cumulative distributions of the number contacts
-   for \cetis in different samples and with D$_{\rm max}$=10~kpc.
-%
-The upper panel shows the variation of the probability distribution
-   estimation as a function of the $\tau_a$ parameter.  All the curves
+   for \cetis{} in different samples and with D$_{\rm max}$=10~kpc.
+%
+The upper panel shows the variation of the distribution
+   as a function of the $\tau_{\rm a}$ parameter, including the values 
+   0.1, 0.5, 1, 2, and 5~kyr. All the curves
    correspond to models with 
-   $\tau_s=40~kyr$ and D$_{\rm max}=10\,kpc$.
-%
-The bottom panel corresponds to several values of $\tau_s$, with
-   $\tau_a=1~kyr$ and D$_{\rm max}=10\,kpc$.
+   $\tau_{\rm s}=$40~kyr and \mbox{D$_{\rm max}=10$~kpc}.
+%
+   The bottom panel corresponds to values of $\tau_{\rm s}$
+   5, 10, 20 and 50~kyr,
+   $\tau_{\rm a}=1$~kyr and D$_{\rm max}=10$~kpc.
 %
    Cumulative distributions can be used to visualise the probability
-   of a random node in a model of having more than M contacts.  For
+   of a random node of having more than M contacts in a given model.  For
    example, in the upper panel the probability of a random node having
    more than 10 contacts is nearly one for the model with the
-   smallest $\tau_a$ value, and nearly zero for the model with the
-   largest $\tau_a$ value.
+   smallest $\tau_{\rm a}$ value, and nearly zero for the model with the
+   largest $\tau_{\rm a}$ value.
    %
@@
 A discrete event simulation is performed for a given model, in this
-work $M(\tau_a, \tau_s,
+work $M(\tau_{\rm a}, \tau_{\rm s},
 D_{\rm max})$, by keeping track of a set of variables that change each time an
@@
 Rate of nodes that never make contact (listening) as a
-function of $\tau_a$ and $\tau_s$, for 
-$D_{\rm max}$=3~kpc (left panel),
-$D_{\rm max}$=12~kpc (middle panel), and
-$D_{\rm max}$=25~kpc (right panel)
+   function of $\tau_{\rm a}$ and $\tau_{\rm s}$, for 
+$D_{\rm max}$=1~kpc (left panel),
+$D_{\rm max}$=5~kpc (middle panel), and
+$D_{\rm max}$=10~kpc (right panel)
 %
@@
 Rate of nodes that listen at the moment of the
-awakening, as a funcion of $\tau_a$ and $\tau_s$, for
-$D_{\rm max}$=3~kpc (left panel),
-$D_{\rm max}$=12~kpc (middle panel), and
-$D_{\rm max}$=25~kpc (right panel)
+   awakening, as a funcion of $\tau_{\rm a}$ and $\tau_{\rm s}$, for
+$D_{\rm max}$=1~kpc (left panel),
+$D_{\rm max}$=5~kpc (middle panel), and
+$D_{\rm max}$=10~kpc (right panel)
 %
@@
 Also, the list of active \cetis{} is obtained as a function of time.
-%
-%Some fixed variables must be set in order to carry out the simulation:
-%the size of the Galactic Habitable Zone, the mean lifetime of a
-%\ceti{}, $\tau_s$, the mean waiting time for the appearance of another
-%\ceti{}, $\tau_a$, and the maximum distance a message can be detected.
-%
-% The Galactic Habitable Zone is set at fixed values of the inner and outer
-% radii, radial symmetry is assumed, and the with of the galactic disk
-% is neglected compared to the radial size.
-% %
-% The two radii are measured in light years, with the aim to maintain a
-% single and comprehensive unit for both space and time coordinates. 
-%
 
@@
 %
-% In order to achieve high performance we made use of vectorised methods
-% and array broadcasting properties provided by the NumPy package
-% \citep{oliphant_numpy_2006, vanderwalt_numpy_2011}
-% %
-% The combination of C-compiled Numpy routines for numerical
-% computations with a high-level language allows to obtain an easily
-% comprehensible and maintainable code.
-% %
 From the viewpoint of a user, \hs{} is an object-oriented package that
@@
 \footnote{https://hearsay.readthedocs.io/en/latest/}.
-
-% %Quality assurance
-% %
-% Regarding the quality assurance of the software, we endorse the PEP-8 style
-% guide for python (van Rossum et al. 2001) throughout the entire project, which
-% improves the code maintainability and helps to minimise the so called
-% ``code-entropy'' of future versions.
-% %
-% The deviations in style were analysed with the flake8 tool, which improves the
-% readability and consistency of the code.
-% %
-% Testing and code coverage have been applied to the package and the development
-% process as standard qualitative and quantitative metrics.
-% %
-% The different parts and functions can be validated to assert if they return the
-% expected results on specific test cases and testing scenarios
-% \citep{jazayeri_engeneering_2007}.
-% %
-% On the other hand, code coverage measures the fraction visited and executed in
-% the code by the testing procedure \citep{miller_testing_1963}.
-% %
-% The advantage of an exhaustive code-coverage is that it ensures that most of
-% the sources of errors have been considered.
-% %
-% In the \hs{} project we provide test tools for the distinct parts of the
-% simulation setup and run, reaching a 99\% of code coverage in the version used
-% for this work.
-% %
-% For every new version of the code, all these quality assurance methods and
-% metrics are run automatically by a continuous-integration service.
-% %
-% It is possible to reproduce the experiments performed in this work using the
-% \hs{} package.
-% %
-% The installation is available through a git repository and also on the Python
-% Package Index (PyPI).
-% %
+%
 Since the simulation setup is configurable, the time required for a simulation
@@
 %
-Left panel shows the histograms for several values of $\tau_a$, and
-$\tau_s$ in the range 17000-24000~yr.
-%
-Right panel shows the histograms for several values of $\tau_s$, and
-$\tau_a$ in the range 68000-100000~yr.
-%
-The shape of the histograms does not change when $\tau_a$ is
-modified, but significantly changes when $\tau_s$ is modified. 
+   Left panel shows the histograms for several values of $\tau_{\rm a}$, and
+   $\tau_{\rm s}$ in the range 17000-24000~yr.
+%
+   Right panel shows the histograms for several values of $\tau_{\rm
+   s}$, and
+   $\tau_{\rm a}$ in the range 68000-100000~yr.
+%
+   The shape of the histograms does not change when $\tau_{\rm a}$ is
+   modified, but significantly changes when $\tau_{\rm s}$ is modified. 
 %
@@
 
-%%% S E C T I O N - - - - - - - - - - - - - - - - - - - - - - -
 \section{Results: exploring the parameter space}\label{S_results}
@@
 The parameters for the temporal aspects of the simulation (the mean
-waiting time for the next awakening, $\left<\tau_a\right>$, and the
-mean lifetime, $\left<\tau_s\right>$) cover the ranges
+waiting time for the next awakening, $\left<\tau_{\rm a}\right>$, and the
+mean lifetime, $\left<\tau_{\rm s}\right>$) cover the ranges
 10$^2$-10$^6$~yr, with 
@@
 This partition was chosen after the requirement of the software to take linear bins, 
-aiming at a better sampling of the low $\tau_a$ and low $\tau_s$ region of the parameter space.
+aiming at a better sampling of the low $\tau_{\rm a}$ and low
+$\tau_{\rm s}$ region of the parameter space.
 %
@@
 %
-Some of them are directly derived from the discrete events, namely, the ID of
-emitting and receiving \cetis{} and the position in the galaxy.
-%
-We also save the times of each of the events that are relevant to keep track of
-the number of \cetis{} for each simulation, i.e., the times of the four types
-of events.
+Some of them are directly derived from the discrete events, namely,
+the ID of emitting and receiving \cetis{} and the position in the
+galaxy.
+%
+We also save the times of each of the events that are relevant to keep
+track of the number of \cetis{} for each simulation, i.e., the times
+of the four types of events.
 %
@@
 %
-We also obtain quantities that represent the properties of the \cetis{}, for
-example the total time elapsed between the A-type and the D-type of each node,
-which represent their corresponding lifetimes.
-%
-The time span of a \ceti{} listening another or being listened by another node
-can also be derived by keepig track of the times of the events in a simulation.
-%
-This way we can also compute the distribution in the simulated galaxy of
-\cetis{} that reach contact, the distributions of the waiting times until the
-first contact or the distributions of waiting times until the next contact.
-%
-The following properties of the population of \cetis{} can also be derived:
-%
-the fraction of the lifetime a node is listening to at least another node
-(i.e., within their light cone),
+We also obtain quantities that represent the properties of the
+\cetis{}, for example the total time elapsed between the A-type and
+the D-type of each node, which represent their corresponding
+lifetimes.
+%
+The time span of a \ceti{} listening another or being listened by
+another node can also be derived by keepig track of the times of the
+events in a simulation.
+%
+This way we can also compute the distribution in the simulated galaxy
+of \cetis{} that reach contact, the distributions of the waiting times
+until the first contact or the distributions of waiting times until
+the next contact.
+%
+The following properties of the population of \cetis{} can also be
+derived:
+%
+the fraction of the lifetime a node is listening to at least another
+node (i.e., within their light cone),
 %
@@
 %
-the fraction of \cetis{} where the first contact is given at the awakening,
+the fraction of \cetis{} where the first contact is given at the
+awakening,
 %
@@
 %
-the distribution of the number of contacts as a function of the age of the
-node,
+the distribution of the number of contacts as a function of the age of
+the node,
 %
@@
 %
-Another useful derived quantity is the duration of two-way communication
-channels or the fraction of contacts that admit a response.
-% 
-It is also possible to analyze the relations between the distance to \ceti{}
-vs. the time of two-way communications, the distance to \ceti{} vs. the age of
-contacted node, the age of a node and the maximum number of contacted nodes
-before the D--type event, or the lifespan of a node vs. the maximum number of
-contacts.
+Another useful derived quantity is the duration of two-way
+communication channels or the fraction of contacts that admit a
+response.
+% 
+It is also possible to analyze the relations between the distance to
+\ceti{} vs. the time of two-way communications, the distance to
+\ceti{} vs. the age of contacted node, the age of a node and the
+maximum number of contacted nodes before the D--type event, or the
+lifespan of a node vs. the maximum number of contacts.
 %
@@
 samples, including short and long lifetimes
-$\left(\,\left<\tau_s\right>=\right.$10000~yr and
-$\left<\tau_s\right>$=100000~yr, respectively$\left.\right)$, dense
+$\left(\,\left<\tau_{\rm s}\right>=\right.$10000~yr and
+$\left<\tau_{\rm s}\right>$=100000~yr, respectively$\left.\right)$, dense
 and sparse spatial distribution
-$\left(\,\left<\tau_a\right>=\right.$10000~yr and
-$\left<\tau_a\right>$= 100000~yr, respectively$\left.\right)$, and two
-different signal reach ranges (D$_{\rm max}$=12~kpc and
-D$_{\rm max}$=25~kpc).
-%
-The models M1 and M5 are not shown, since in both cases the total
-number of contacts is different from zero for less than the five per
-cent of the \cetis{}.
+$\left(\,\left<\tau_{\rm a}\right>=\right.$10000~yr and
+$\left<\tau_{\rm a}\right>$= 100000~yr, respectively$\left.\right)$, and two
+different signal reach ranges (D$_{\rm max}$=12~kpc and D$_{\rm
+max}$=25~kpc).
 %
@@
 %
-As it can be seen, the mean lifetime is more
-determinant than the mean awakening rate (dense and sparse,
-represented by a different shade) for the number of contacts.
-%
-As expected, of the samples with the longest $D_{\rm max}$ the model with
-a dense awakening in the timeline (low $\left<\tau_a\right>$) and a
-long lifetime (high $\left<\tau_s\right>$), is the model that
+As it can be seen, the mean lifetime is more determinant than the mean
+awakening rate (dense and sparse, represented by a different shade)
+for the number of contacts.
+%
+As expected, of the samples with the longest $D_{\rm max}$ the model
+with a dense awakening in the timeline (low $\left<\tau_{\rm a}\right>$) and
+a long lifetime (high $\left<\tau_{\rm s}\right>$), is the model that
 maximizes the number of contacts, reaching a maximum of order 20
@@
 %
-Similarly, for the models with long signal range
-($D_{\rm max}=25$~kpc), the one which is dense and with long lifetime
-has the maximum number of contacts, reaching nearly 100 contacts for
-each \ceti{} in its entire lifetime.
+Similarly, for the models with long signal range ($D_{\rm
+max}=25$~kpc), the one which is dense and with long lifetime has the
+maximum number of contacts, reaching nearly 100 contacts for each
+\ceti{} in its entire lifetime.
 %
@@
 %
-We should point out that this plot 
-has a logarithmic scale on the x-axis, and it is the cumulative, not
-differential, empirical distribution.
-%
-Therefore, the differences in the
-number of contacts for different models are relatively large.
-%
-Another interesting feature observed in the Fig.~\ref{F_number_of_contacts}, is that for the
-models analyzed, there are \cetis{} that have no causal contacts with
-any other \ceti{} during their whole lifetime.
+We should point out that this plot has a logarithmic scale on the
+x-axis, and it is the cumulative, not differential, empirical
+distribution.
+%
+Therefore, the differences in the number of contacts for different
+models are relatively large.
+%
+Another interesting feature observed in the
+Fig.~\ref{F_number_of_contacts}, is that for the models analyzed,
+there are \cetis{} that have no causal contacts with any other \ceti{}
+during their whole lifetime.
 %
 The fraction of \cetis{} with no contact ranges from nearly 20\% up to
-nearly 90\% (models M1 and M5, not shown), except for the dense and
-long lifetime models, for which almost all \cetis{} make at least one
-contact.
+nearly 90\%, except for the dense and long lifetime models, for which
+almost all \cetis{} make at least one contact.
 %
@@
 listen to another \ceti{}), as a function of the mean lifetime
-($\left<\tau_s\right>$, in the range 0--500 a) and the mean awakening
-time ($\left<\tau_a\right>$, in the range 0--200 a), for three
-different values of the maximum signal range, $D_{\rm max}$=3~kpc (left
-panel), $D_{\rm max}=12$~kpc (middle panel), and $D_{\rm max}=25$~kpc
-(right panel).
-%
-A clear pattern emerges, showing that the probability
-for a \ceti{} of making causal contact with at least another \ceti{}
-during their entire lifetime, increases with increasing $D_{\rm max}$,
-increasing $\left<\tau_s\right>$ and decreasing $\left<\tau_a\right>$,
-following a roughly linear dependence with the three parameters.
+($\left<\tau_{\rm s}\right>$, in the range 0--500 a) and the mean awakening
+time ($\left<\tau_{\rm a}\right>$, in the range 0--200 a), for three
+different values of the maximum signal range, $D_{\rm max}$=3~kpc
+(left panel), $D_{\rm max}=12$~kpc (middle panel), and $D_{\rm
+max}=25$~kpc (right panel).
+%
+A clear pattern emerges, showing that the probability for a \ceti{} of
+making causal contact with at least another \ceti{} during their
+entire lifetime, increases with increasing $D_{\rm max}$, increasing
+$\left<\tau_{\rm s}\right>$ and decreasing $\left<\tau_{\rm a}\right>$, following
+a roughly linear dependence with the three parameters.
 
@@
 as a function of the mean lifetime
-($\left<\tau_s\right>$, in the range 0--500 a) and the mean awakening
-time ($\left<\tau_a\right>$, in the range 0--200 a).
+($\left<\tau_{\rm s}\right>$, in the range 0--500 a) and the mean awakening
+time ($\left<\tau_{\rm a}\right>$, in the range 0--200 a).
 %
@@
 with the three parameters is roughly linear in both cases.
-%
-%% DISCUTIR ESTOS RESUOLTADOS SEGUN LA LITERATURA
 
@@
 Upper panel (a) panel shows the histograms for several values of the
-mean survival time $\tau_s$, with the mean awakening time $\tau_a$ in
+mean survival time $\tau_{\rm s}$, with the mean awakening time $\tau_{\rm
+a}$ in
 the range 68000-100000 a.
 %
-Lower panel (b) shows the histograms for several values of $\tau_a$,
-with $\tau_s$ in the range 17000-24000 a.
+Lower panel (b) shows the histograms for several values of $\tau_{\rm
+a}$,
+with $\tau_{\rm s}$ in the range 17000-24000 a.
 %
 As it can be seen, the shape of the histograms is almost the same when
-$\tau_a$ is modified (panel (b)), but significantly changes when
-$\tau_s$ is modified (panel (a)). 
-%
-For different values of $\tau_a$, the shape of the distribution is
+$\tau_{\rm a}$ is modified (panel (b)), but significantly changes when
+$\tau_{\rm s}$ is modified (panel (a)). 
+%
+For different values of $\tau_{\rm a}$, the shape of the distribution is
 similar, and the main difference is given by the variation on the number of
@@
 %
-Conversely, for different values of $\tau_s$ (at a fixed range for
-$\tau_a$), the shape of the
+Conversely, for different values of $\tau_{\rm s}$ (at a fixed range for
+$\tau_{\rm a}$), the shape of the
 distribution varies significantly.
@@
 respect to the total number of nodes that make causal contact at least
-one time in the time range from the A--type up to the D--type
-events.
-%
-From a frequentist approach, the cumulative fraction of contacts
-from the awakening
-($t=$0~yr) up to a given time are related to the estimation of the
-probability of listening during that time interval and reaching to
-causal contact region of at least another node.
+one time in the time range from the A--type up to the D--type events.
+%
+From a frequentist approach, the cumulative fraction of contacts from
+the awakening ($t=$0~yr) up to a given time are related to the
+estimation of the probability of listening during that time interval
+and reaching to causal contact region of at least another node.
 %
@@
 
-Remarkably, the differential probability of making a causal contact reach its
-maximum at the awakening event for all models.
-%
-That is, for a short period of time the initial moment is the most promising
-for making a (causal) contact, for a given technology.
+Remarkably, the differential probability of making a causal contact
+reach its maximum at the awakening event for all models.
+%
+That is, for a short period of time the initial moment is the most
+promising for making a (causal) contact, for a given technology.
 %
@@
 %
-This offers a new approach to SETI programs, where the
-search for new communication technologies or the exploration of
-new communication channels has a fundamental role.
+This offers a new approach to SETI programs, where the search for new
+communication technologies or the exploration of new communication
+channels has a fundamental role.
 %
@@
 %
-The probability of such a contact does not grows linearly
-with the observing time, but is greater during the first 100
-millennia.
+The probability of such a contact does not grows linearly with the
+observing time, but is greater during the first 100 millennia.
 
@@
 
-We have presented a stochastic model (SC3Net) to analyse the network of
-constrained causally connected nodes in a simplified Galaxy.
-%
-It represents an idealised scenario of perfectly efficient emitters and
-receivers with the restriction of a maximum distance separation.
+We have presented a stochastic model (SC3Net) to analyse the network
+of constrained causally connected nodes in a simplified Galaxy.
+%
+It represents an idealised scenario of perfectly efficient emitters
+and receivers with the restriction of a maximum distance separation.
 %
 These emitters and receivers correspond to nodes that form a
-communication network whose properties depend on their density
-and mean survival time.
+communication network whose properties depend on their density and
+mean survival time.
 %
@@
 three-dimensional parameter space, considering a grid of mean time
-separations between the activation of new nodes
-in the range 500--10$^6$~yr.
+separations between the activation of new nodes in the range
+500--10$^6$~yr.
 %
 Additionally, we explore a grid of the mean survival times between
-500~yr and 10$^6$~yr, and values for the maximum distance range of signals 
-of 100~pc, 500~pc, 1~kpc, 5~kpc and 10~kpc.
+500~yr and 10$^6$~yr, and values for the maximum distance range of
+signals of 100~pc, 500~pc, 1~kpc, 5~kpc and 10~kpc.
 %
@@
 
-% We have implemented a suite of simulations, following a stochastic
-% approach, to explore a three dimensional hypothesis space of a simple
-% model that accounts for the causal connections between
-% communicating civilisations in the Galaxy. The different models can be
-% generated with a minimal number of assumptions. We argue that three
-% parameters can describe a variety of situations, ranging from a galaxy
-% model where an intelligent civilization is very rare, to a galaxy
-% model populated with plenty of civilisations in causal contact. With
-% this approach, we can estimate several quantities as a function of the
-% parameters on the hypothesis space, so we explored the outcomes of
-% different models that can be derived from the simulations. For
-% instance, it is possible to estimate the timescale for an
-% uninterrupted SETI-equivalent effort to reach success, in terms of
-% different model parameters that reflect different, so far unknown,
-% scenarios for the appearance of intelligent life on the Galaxy.
-
-Although the simulations use several hypotheses, we
-argue that the model is not worthy of further complexity,
-given the limited knowledge about the origin and persistence of
-life in the Galaxy.
-%
-The implementation of more detailed or sophisticated models
-would increase the number of free parameters without any 
-improvement in their predictive power.
-%
-Thus, we take advantage of the simplicity of the model to explore
-the parameter space in order to gain insight on the consequences of
+Although the simulations use several hypotheses, we argue that the
+model is not worthy of further complexity, given the limited knowledge
+about the origin and persistence of life in the Galaxy.
+%
+The implementation of more detailed or sophisticated models would
+increase the number of free parameters without any improvement in
+their predictive power.
+%
+Thus, we take advantage of the simplicity of the model to explore the
+parameter space in order to gain insight on the consequences of
 different scenarios for the search of intelligent life.
 %
-Our analysis is not centred in obtaining the odds for the Earth
-to make contact with another civilization.
+Our analysis is not centred in obtaining the odds for the Earth to
+make contact with another civilisation.
 %
@@
 This causally connected nodes are sparsely distributed in both space
-and time, making analytical treatments difficult and justifying
-the simulation approach.
+and time, making analytical treatments difficult and justifying the
+simulation approach.
 
@@
 contact is extremely unlikely unless the galaxy is densely populated
-by intelligent civilizations with large average lifetimes.
-%
-On average, in order to have at least one contact in the entire lifetime
-of the \ceti{}, there must appear a mean of at least one \ceti{}
-every $\sim$20000 a, with a mean active period of at least
+by intelligent civilisations with large average lifetimes.
+%
+On average, in order to have at least one contact in the entire
+lifetime of the \ceti{}, there must appear a mean of at least one
+\ceti{} every $\sim$20000 a, with a mean active period of at least
 $\sim$30000 a.
@@
 several authors, which state that a contact between the Earth and
-another intelligent civilization in the Galaxy is quite unlikely,
+another intelligent civilisation in the Galaxy is quite unlikely,
 provided the maximum distance of the signal and the lifetime of the
@@
 hypothesis, it does consider catastrophic events implicitly in the
-mean value and distribution of the lifetime, $\left<\tau_s\right>$.
+mean value and distribution of the lifetime, $\left<\tau_{\rm s}\right>$.
 %
@@
 although they would be too slow to make a significant impact on the
-communication network among intelligent civilizations.
+communication network among intelligent civilisations.
 %
 Our results are also consistent with those presented by
-\citet{grimaldi_signal_2017}, who estimates an upper limit for the mean
-number of extraterrestrial civilizations that could contact Earth
-using Monte Carlo simulations, from a statistical model where the width
-of the Galactic disk is not negligible.
+\citet{grimaldi_signal_2017}, who estimates an upper limit for the
+mean number of extraterrestrial civilisations that could contact Earth
+using Monte Carlo simulations, from a statistical model where the
+width of the Galactic disk is not negligible.
 %
@@
 %
-Unlike most 
-of the studies that make use of statistical models or
+Unlike most of the studies that make use of statistical models or
 simulations \citep{cirkovic_temporal_2004, smith_broadcasting_2009,
@@
 %
-Thus, it does not need
-a detailed description or modeling of physical processes that give rise to
-intelligent life.
+Thus, it does not need a detailed description or modeling of physical
+processes that give rise to intelligent life.
 %
@@
 
-We have also fond that there 
-is a balance between the density of \cetis{} and the mean
-lifetime since, as expected, a lower density can be compensated by a
-longer active time period.
+We have also fond that there is a balance between the density of
+\cetis{} and the mean lifetime since, as expected, a lower density can
+be compensated by a longer active time period.
 %
@@
 the case of a less populated galaxy but with very ancient
-civilizations.
+civilisations.
 %
@@
 The short time interval between the
-rise and fall of civilizations, compared to the age and extension of
+rise and fall of civilisations, compared to the age and extension of
 our Galaxy, is a fundamental limitation for the number of contacts.
@@
 Plots and simulations were made with software developed by the authors
-using R and python languages. Plots were postprocessed with inkscape.
+the python language. 
+%
+Plots were postprocessed with inkscape.
+%
+Simulations were run on the Clemente cluster at the Instituto de
+Astronomía Teórica y Experimental (IATE).
 %

### 2020-06-04 submit_v03/manuscript.tex
submitted V3

@@
    %
-   The causal connection occurs between $t=$100~yr and $t=$130~yr
-   %
-   Meaningfull times are marked with vertical dashed lines.
+   The causal connection occurs between $t=$100~yr and $t=$130~yr.
+   %
+   Meaningful times are marked with vertical dashed lines.
    }
@@
 %
-The sphere of first contact (SFC) is the sphere centred on the emitter that
+The sphere of the first contact (SFC) is the sphere centred on the emitter that
 grows until its radius reaches the D$_{\rm max}$ distance, in which the
@@
 %
-The surface of last contact (SLC) is another sphere that grows from a ``D''
+The surface of the last contact (SLC) is another sphere that grows from a ``D''
 event (``Doomsday'').
@@
 %
-The node E$_1$ can listen signals from the node E$_2$, since
+The node E$_1$ can listen to signals from the node E$_2$, since
 the ``Contact'' event (t=C$_{12}$) up to the ``Blackout'' event
@@
 This is done in terms of three independent parameters, namely, the mean time
-span between the appearance of consecutive \cetis{}, $\left<\tau_{\rm
-a}\right>$,
-their mean lifetime $\left<\tau_{\rm s}\right>$, and the maximum distance a signal
+span between the appearance of consecutive \cetis{}, $\tau_{\rm a}$,
+their mean lifetime $\tau_{\rm s}$, and the maximum distance a signal
 can be detected by another \ceti{} ($D_{\rm max}$).
@@
 The later is related to the time span between the appearance of consecutive
-\cetis{} (since when $\left<\tau_{\rm a}\right>$ is shorter, it produces a greater
+\cetis{} (since when $\tau_{\rm a}$ is shorter, it produces a greater
 density of \cetis{}).
@@
 %
-The receiver node E$_1$ can listen signals from the emitter node
+The receiver node E$_1$ can listen to signals from the emitter node
 E$_2$, from the first contact event (t=C$_{12}$) until the last
@@
 %
-Similarly, the receiver node E$_2$ can listen signals from the emitter
+Similarly, the receiver node E$_2$ can listen to signals from the emitter
 node E$_1$, from t=C$_{21}$ up to t=B$_{21}$.
@@
 homogeneous over the GHZ.
+%
+The adopted geometry of the GHZ in all the simulations is given by a 
+two-dimensional annular region,
+with an inner radius of 7~kpc and an outer radius of 9~kpc
+\citep{lineweaver_galactic_2004}.
 %
@@
 imply extreme values for the parameters that represent the density of
-the nodes (i.e., $\left< \tau_{\rm a} \right>$).
+the nodes (i.e., $\tau_{\rm a}$).
 
@@
 	      & 5$\cdot10^2$~yr & 40500~yr & 21\\ 
-	    & & 5$\cdot$10$^5$~yr & 10$^6$~yr & 51\\ 
+	    & (linear grids) & 5$\cdot$10$^5$~yr & 10$^6$~yr & 51\\ 
    $\tau_{s}$ & Mean lifetime of a \ceti{}
 	      & 5$\cdot10^2$~yr & 40500~yr & 21\\ 
-	    & & 5$\cdot$10$^5$~yr & 10$^6$~yr & 51\\ 
+	    & (linear grids) & 5$\cdot$10$^5$~yr & 10$^6$~yr & 51\\ 
 	D$_{\rm max}$ & Maximum reach of a message  & \multicolumn{3}{c}{100 pc, 500 pc, 1 kpc, 5 kpc, 10 kpc} \\
@@
 %
-%% , benguigui_classificacion_2016}.
-%The power law behaviour has been observed in a variety of systems.
-%
 Additional examples include stock market fluctuations, sizes of computer files
@@
 More examples can be found in the literature \citep{martin_origin_2006,
-maccone_KLT_2010} \citep{barabasi_scale_2009, maccone_evolution_2014}
-\citep{maccone_lognormals_2014}
+maccone_KLT_2010, barabasi_scale_2009, maccone_evolution_2014, maccone_lognormals_2014}.
 %
 This distribution family is suitable for the statistical description
-of the $D_{\rm mar}$ parameter, although in this work we assume a uniform
+of the $D_{\rm max}$ parameter, although in this work we assume a uniform
 distribution for simplicity.
@@
 %
-   The bottom panel corresponds to values of $\tau_{\rm s}$
-   5, 10, 20 and 50~kyr,
-   $\tau_{\rm a}=1$~kyr and D$_{\rm max}=10$~kpc.
+   The bottom panel corresponds to 
+   $\tau_{\rm a}=1$~kyr and D$_{\rm max}=10$~kpc, and the following values
+   of $\tau_{\rm s}$: 5, 10, 20 and 50~kyr.
 %
@@
    %
-Rate of nodes that never make contact (listening) as a
+The fraction of constrained causal contact nodes that never make contact (listening) as a
    function of $\tau_{\rm a}$ and $\tau_{\rm s}$, for 
@@
 $D_{\rm max}$=5~kpc (middle panel), and
-$D_{\rm max}$=10~kpc (right panel)
+$D_{\rm max}$=10~kpc (right panel).
+%
+The values of $\tau_{\rm a}$ and $\tau_{\rm s}$ are in the range
+   5$\cdot10^5$ to 10$^6$~yr.
+%
+The matrix is shown as obtained from the simulations, and the level
+   curves are shown for the smoothed matrix.
 %
@@
    %
-Rate of nodes that listen at the moment of the
-   awakening, as a funcion of $\tau_{\rm a}$ and $\tau_{\rm s}$, for
+The fraction of constrained causal contact nodes that make contact at the moment of the
+   awakening (i.e., $t_{\rm A}=t_{\rm C}$), as a function of $\tau_{\rm a}$ and $\tau_{\rm s}$, for
 $D_{\rm max}$=1~kpc (left panel),
-$D_{\rm max}$=5~kpc (middle panel), and
-$D_{\rm max}$=10~kpc (right panel)
-%
+$D_{\rm max}$=5~kpc (middle panel) and
+$D_{\rm max}$=10~kpc (right panel).
+%
+The values of $\tau_{\rm a}$ and $\tau_{\rm s}$ are in the range
+   5$\cdot10^5$ to 10$^6$~yr.
+%
+The matrix is shown as obtained from the simulations, and the level
+   curves are shown for the smoothed matrix.
    }
@@
 
-\subsection{Model for \cetis{}}
-
-Since the method is based on events, the first step is to define the
-relationships between them and how events trigger changes on the state of the
-system.
-%
-Once an event is produced, the system is updated by saving the corresponding
-change in the related variables.
-%
-For example, a B-type event diminishes the number of active \cetis{} by one,
-and also could affect the number of ongoing communications.
-%
-We consider a first approach to the galactic habitable zone as a two-dimensional
-annular region.
-%
-The adopted values for the geometry of the GHZ are 7~kpc and 9~kpc, for the
-inner and outer radius, respectively \citep{lineweaver_galactic_2004}. 
-%
-This simple model does not take into account the variations in stellar density
-given by the spiral structure of the Galaxy.
-%
-The main task in setting a DE simulation is to define the events and what
-changes do these events produce on the system.
-%
-Although there are several possible approaches, we choose to follow the
-evolution of the system according to the following events:
-
-\begin{enumerate}
-   \item[(A)] Awakening: A new \ceti{} acquires the ability to emit
-      and receive messages and starts to actively broadcast and
-      lookout for messages.
-   \item[(C)] Contact: A new one-way causal contact is established.
-   \item[(B)] Blackout: An existing causal contact is interrupted,
-      with or without the occurrence of a doomsday event.
-   \item[(D)] Doomsday: An old \ceti{} disappears, or halts its
-      communication activities.
-\end{enumerate}
-
-We have chosen this nomenclature so that it can be easily remembered
-from the initials A, B, C, D.
-%
-The \blackout is produced when one of the two \cetis{} halts in its
-capability of emitting and receiving signals.
-%
-It can occur at the same time the \doomsday or before.
-%
-Similarly, a \ccontact can be produced at the time of the \aawakening
-or later, and there can be several \contacts for the same node.
-%
-The time marks for each event (A and D for each \ceti{} and C and B
-for each constrained causal contact) are stored as a result of the simulation.
-%
-Also, the list of active \cetis{} is obtained as a function of time.
-
-\subsection{Model implementation}
+\subsection{Implementation of the SC3Net model}
 
@@
 %
-In order to make a reproducible project, we developed the tools
-that allow to run the simulations and obtain the results shown in this work.
-%
-The simulations were implemented on a Python-3 code, dubbed \hs{} (Lares et al., in preparation),
-which is publicly accessible through the GitHub
-platform\footnote{https://github.com/mlares/hearsay} under the
-MIT-license.
+In order to make a reproducible project, we developed the tools that
+allow to run the simulations and obtain the results shown in this
+work.
+%
+The simulations were implemented on a Python-3 code, dubbed \hs{}
+(Lares et al., in preparation), which is publicly accessible through
+the GitHub platform\footnote{https://github.com/mlares/hearsay} under
+the MIT-license.
 %
@@
 %
-Since the simulation setup is configurable, the time required for a simulation
-to complete depends on the simulation parameters.
-%
-It also depends on the hardware that is used to make the run, and on whether
-the parallel option is set.
+Since the simulation setup is configurable, the time required for a
+simulation to complete depends on the simulation parameters.
+%
+It also depends on the hardware that is used to make the run, and on
+whether the parallel option is set.
 %
 However, it is simple to make a local run of limited versions of the
-experiments to have a sense of the time required by the code to complete the
-simulations.
-%
-More information on this can be found on the documentation of the software.
+experiments to have a sense of the time required by the code to
+complete the simulations.
+%
+More information on this can be found on the documentation of the
+software.
 
@@
 %
-   Left panel shows the histograms for several values of $\tau_{\rm a}$, and
-   $\tau_{\rm s}$ in the range 17000-24000~yr.
-%
-   Right panel shows the histograms for several values of $\tau_{\rm
-   s}$, and
-   $\tau_{\rm a}$ in the range 68000-100000~yr.
-%
-   The shape of the histograms does not change when $\tau_{\rm a}$ is
-   modified, but significantly changes when $\tau_{\rm s}$ is modified. 
+   Upper panel shows the histograms for several values of $\tau_{\rm a}$, and
+   $\tau_{\rm s}$=10~kyr and D$_{\rm max}$=10~kpc.
+%
+   Bottom panel shows the histograms for several values of $\tau_{\rm
+   s}$, with $\tau_{\rm a}$=10~kyr and D$_{\rm max}$=10~kpc. The
+   arrows indicate the sample and highlight the values of the first
+   waiting time bin, which has a clear excess for the models with
+   larger $\tau_{\rm s}$ values.
+%
+   The choice of these models was made in order to show the trends in
+   the results as a function of the two temporal parameters.
 %
@@
 
-We implemented the simulation of a regular grid of models varying over
-the parameter space, which covers 15000 different models.
+We implemented the simulation of a regular grids of models varying over
+the parameter space, which covers 5204 models.
 %
 For each model, we simulated several realizations with different random
-seeds, in order to perform a MonteCarlo estimation of the variance.
-%
-The number of random realzations varies from one (for the densel populated models)
-up to 20 (for the sparsely populated models).
+seeds, adding up a total of 
+158546 simulation runs.
+%
+The number of random realzations varies from one (for the densest populated models)
+up to 50 (for the sparsely populated models).
 %
 The parameters for the temporal aspects of the simulation (the mean
-waiting time for the next awakening, $\left<\tau_{\rm a}\right>$, and the
-mean lifetime, $\left<\tau_{\rm s}\right>$) cover the ranges
+waiting time for the next awakening, $\tau_{\rm a}$, and the
+mean lifetime, $\tau_{\rm s}$) cover the ranges
 10$^2$-10$^6$~yr, with 
@@
 values 100~pc, 500~pc, 1~kpc, 5~kpc and 10~kpc.
-%
-This setup makes a total of 750000 simulations, each covering a time
-range of ten million years.
 %
@@
 distributions of the number contacts for \cetis{} in six different
-samples, including short and long lifetimes
-$\left(\,\left<\tau_{\rm s}\right>=\right.$10000~yr and
-$\left<\tau_{\rm s}\right>$=100000~yr, respectively$\left.\right)$, dense
-and sparse spatial distribution
-$\left(\,\left<\tau_{\rm a}\right>=\right.$10000~yr and
-$\left<\tau_{\rm a}\right>$= 100000~yr, respectively$\left.\right)$, and two
-different signal reach ranges (D$_{\rm max}$=12~kpc and D$_{\rm
-max}$=25~kpc).
-%
-The areas between models with the same mean lifetime and mean
-awakening time are shaded for visualization purposes.
+samples, including short and long lifetimes, dense and sparse spatial
+distribution and D$_{\rm max}$=10~kpc.
 %
@@
 %
-As expected, of the samples with the longest $D_{\rm max}$ the model
-with a dense awakening in the timeline (low $\left<\tau_{\rm a}\right>$) and
-a long lifetime (high $\left<\tau_{\rm s}\right>$), is the model that
-maximizes the number of contacts, reaching a maximum of order 20
+The model with a dense awakening in the timeline (low $\tau_{\rm a}$)
+maximizes the number of contacts, reaching a maximum of more than 300
 contacts for a single \ceti{}.
 %
-Similarly, for the models with long signal range ($D_{\rm
-max}=25$~kpc), the one which is dense and with long lifetime has the
-maximum number of contacts, reaching nearly 100 contacts for each
-\ceti{} in its entire lifetime.
+This case, however, requires that a new node appears in the Galaxy
+every 100 years on average.
+%
+Similarly, the model with long lifetimes has the maximum number of
+contacts, reaching nearly 100 contacts for each \ceti{} in its entire
+lifetime.
 %
@@
 %
+It is expected then that a model where nodes appear with a high
+frecuency and have very long lifetimes can reach tens of contacts on
+the full time period between $t_{\rm A}$ and $t_{\rm D}$.
+%
+On the other side, a model where the activation of new nodes requires
+a large waiting time and the survival time is short, contacts are
+extremely rare.
+%
 We should point out that this plot has a logarithmic scale on the
@@
 Therefore, the differences in the number of contacts for different
-models are relatively large.
-%
-Another interesting feature observed in the
-Fig.~\ref{F_number_of_contacts}, is that for the models analyzed,
-there are \cetis{} that have no causal contacts with any other \ceti{}
-during their whole lifetime.
-%
-The fraction of \cetis{} with no contact ranges from nearly 20\% up to
-nearly 90\%, except for the dense and long lifetime models, for which
-almost all \cetis{} make at least one contact.
-%
-Since these results depend on just eight models, we analyze the
-properties of the frequency of contacts as a function of both the mean
-lifetime and the mean awakening time.
+models are large.
+%
+This is a consequence of the wide range in both $\tau_{\rm a}$ and
+$\tau_{\rm s}$ covered in the simulations.
+%
+With this ranges, the fraction of \cetis{} with no contact ranges from
+nearly zero up to one.
+%
+This analysis is made in order to explore the behavior of the SC3Net
+model.
 %
@@
 %
-On Fig.~\ref{F_never_contact} we show 2D color maps with the fraction
-of \cetis{} in the simulations that never make contact (i.e., never
-listen to another \ceti{}), as a function of the mean lifetime
-($\left<\tau_{\rm s}\right>$, in the range 0--500 a) and the mean awakening
-time ($\left<\tau_{\rm a}\right>$, in the range 0--200 a), for three
-different values of the maximum signal range, $D_{\rm max}$=3~kpc
-(left panel), $D_{\rm max}=12$~kpc (middle panel), and $D_{\rm
-max}=25$~kpc (right panel).
+On the Fig.~\ref{F_never_contact} we show 2D color maps with the
+fraction of \cetis{} in the simulations that never make contact (i.e.,
+never listen to another \ceti{}), as a function of the mean lifetime
+($\tau_{\rm s}$, in the range 5$\cdot 10^4$ to 10$^6$ yr) and the mean
+awakening time ($\tau_{\rm a}$, in the range 5$\cdot 10^4$ to 10$^6$
+yr), for three different values of the maximum signal range, $D_{\rm
+max}$=1~kpc (left panel), $D_{\rm max}=5$~kpc (middle panel), and
+$D_{\rm max}=10$~kpc (right panel).
 %
@@
 entire lifetime, increases with increasing $D_{\rm max}$, increasing
-$\left<\tau_{\rm s}\right>$ and decreasing $\left<\tau_{\rm a}\right>$, following
-a roughly linear dependence with the three parameters.
-
-The number of \cetis{} that do not succeed in reaching the
-causal contact regions of other nodes is a useful indicator of
-the
-degree of isolation.
-%
-On the other hand, there is also the chance that a number of
-\cetis{} are already in the causal contact region of other nodes,
-at the time of their A-type events.
+$\tau_{\rm s}$ and decreasing $\tau_{\rm a}$, following a roughly
+linear dependence with the three parameters.
+%
+The results of the simulations that comprise the full range of values
+for $\tau_{\rm a}$ and $\tau_{\rm s}$ from 5$\cdot10^2$~yr to $10^6$~yr
+(not shown in the Figures) maintain a similar trend.
+%
+The number of \cetis{} that do not succeed in reaching the causal
+contact regions of other nodes is a useful indicator of the degree of
+isolation.
+%
+On the other hand, there is also the chance that a number of \cetis{}
+are already in the causal contact region of other nodes, at the time
+of their A-type events.
 %
@@
 The fraction of \cetis{} that make the first contact at the awakening
-event is shown in the Fig.~\ref{F_C_at_A}, 
-as a function of the mean lifetime
-($\left<\tau_{\rm s}\right>$, in the range 0--500 a) and the mean awakening
-time ($\left<\tau_{\rm a}\right>$, in the range 0--200 a).
+event is shown in the Fig.~\ref{F_C_at_A}, as a function of the mean
+lifetime ($\tau_{\rm s}$, in the range 5$\cdot 10^4$ to 10$^6$ yr) and
+the mean awakening time ($\tau_{\rm a}$, in the range 5$\cdot 10^4$ to
+10$^6$ yr).
 %
 The three panels correspond to different values of the maximum signal
-range ($D_{\rm max}$):
-%
-$D_{\rm max}=3$~kpc (left
-panel), $D_{\rm max}=12$~kpc (middle panel), and $D_{\rm max}=25$~kpc
-(right panel).
-%
-Although they are not complementary, it is clear that the dependence
-with the three parameters is roughly linear in both cases.
+range:
+%
+$D_{\rm max}=1$~kpc (left panel), $D_{\rm max}=5$~kpc (middle panel),
+and $D_{\rm max}=10$~kpc (right panel).
+%
+The dependence of this metric with the three parameters is roughly
+linear in all cases.
 
@@
 Such distribution can be considered to compute the probability for a
-random \ceti{} to carry out a SETI project and spend a given time until
+random \ceti{} to carry out the searching of other nodes and spend a given time until
 the first contact is made, under the hypotheses of the experiment.
@@
 Upper panel (a) panel shows the histograms for several values of the
-mean survival time $\tau_{\rm s}$, with the mean awakening time $\tau_{\rm
-a}$ in
-the range 68000-100000 a.
-%
-Lower panel (b) shows the histograms for several values of $\tau_{\rm
-a}$,
-with $\tau_{\rm s}$ in the range 17000-24000 a.
-%
-As it can be seen, the shape of the histograms is almost the same when
-$\tau_{\rm a}$ is modified (panel (b)), but significantly changes when
-$\tau_{\rm s}$ is modified (panel (a)). 
-%
-For different values of $\tau_{\rm a}$, the shape of the distribution is
-similar, and the main difference is given by the variation on the number of
-contacts on the different models.
-%
-Conversely, for different values of $\tau_{\rm s}$ (at a fixed range for
-$\tau_{\rm a}$), the shape of the
-distribution varies significantly.
-
+mean awakening time $\tau_{\rm a}$, with the mean survival time $\tau_{\rm
+s}$ of 10~kyr.
+%
+Lower panel (b) shows the histograms for several values of $\tau_{\rm s}$,
+with $\tau_{\rm a}$=10~kyr.
+%
+In both panels the value of D$_{\rm max}$ is set to 10~kpc.
+%
+As it can be seen, there is a clear trend where the number of nodes
+that require a time $t_{\rm 1}$ to make the first contact 
+decreases exponentially with the time $t_{\rm 1}$.
+%
 The fraction of the \cetis{} that have made at least one contact, as a
@@
 large time periods, meaning the a \ceti{} will eventually enter into
-the light cone of other \cetis{}
-
-Remarkably, the differential probability of making a causal contact
-reach its maximum at the awakening event for all models.
+the light cone of other \cetis{}.
+%
+Remarkably, in densely populated models there is a clear excess at
+$t_{\rm 1}$=0~yr.
 %
@@
 %
-This allows to consider the fact that for the nearly 50 years SETI
-programs have been active, no contact event has been reported yet.
+This is given by the fact that at the awakening event, many nodes are
+already on the light cones of other nodes, so that the awakening time
+offers the best chance of making contacts.
 %
@@
 communication technologies or the exploration of new communication
-channels has a fundamental role.
-%
-Moreover, as it can be seen in the Fig.~\ref{F_waiting_for_1C}, there
-is a maximum probability of making the first causal contact at nearly
-10$^5$ years.
-%
-The probability of such a contact does not grows linearly with the
-observing time, but is greater during the first 100 millennia.
+channels has a fundamental role and could be more efficient than long
+observation programs.
 
@@
 derived quantities.
-%
-The full experiment comprises 46485 simulations for a total of 5184
-unique points in the parameter space.
 
@@
 by intelligent civilisations with large average lifetimes.
-%
-On average, in order to have at least one contact in the entire
-lifetime of the \ceti{}, there must appear a mean of at least one
-\ceti{} every $\sim$20000 a, with a mean active period of at least
-$\sim$30000 a.
 %
@@
 %
-According to the timescales involved, the results could explain the
+According to the timescales involved, their results could explain the
 Fermi paradox.
@@
 hypothesis, it does consider catastrophic events implicitly in the
-mean value and distribution of the lifetime, $\left<\tau_{\rm s}\right>$.
+mean value and distribution of the lifetime, $\tau_{\rm s}$.
 %
@@
 %
-On the other hand, taking into account the probability of contact with
-radio signals not sent intentionally \citep{horvat_calculating_2007},
-our analysis would imply a highly populated solar neighborhood.
-%
 Unlike most of the studies that make use of statistical models or
@@
 %
-Thus, it does not need a detailed description or modeling of physical
+Thus, it does not need a detailed description or modeling of the physical
 processes that give rise to intelligent life.
@@
 However, we argue that it is a valid empirical formulation to discuss
-the probabilities of contact, and the time scales involed in the
+the probabilities of contact and the time scales involed in the
 problem.
@@
 Ours is an alternative to the method proposed by
-\citet{balbi_impact_2018}, who perform an analysis based on the Earth
-and use a different model, although our results agree in general.
+\citet{balbi_impact_2018}, who performs an analysis based on the Earth
+with a different model.
+%
+Despite those differences, their results and ours are in general agreement.
 
@@
 parameter space, reducing the problem to only three parameters and a
-few simple hypothesis to perform a complete model for the population and
+few simple hypotheses to perform a complete model for the population and
 communication network in the galaxy.
@@
 %
+If the time intervals between the rise and fall of civilisations are short
+compared to the time required for an electromagnetic signal to travel the 
+large distances in the Galaxy, 
+then the number of contacts would be limited to a low number.
+
 The short time interval between the
@@
 %
-Thus, the temporal dimension, which is missing in the Drake
+The temporal dimension, which is missing in the Drake
 equation, is a key factor to understand the network of contacts on

### 2020-07-17 submit_v05/manuscript.tex
final version

--- 2020-06-04 submit_v03/manuscript.tex
@@
 \usepackage{url}
-
+\usepackage{natbib}
 \usepackage{boites}
@@
 \jname{International Journal of Astrobiology}%
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{condition}{Condition}
 
@@
 
-\newcommand{\awakening}{A event }
-\newcommand{\contact}{C event }
-\newcommand{\blackout}{B event }
-\newcommand{\doomsday}{D event }
-
-\newcommand{\aawakening}{Awakening event }
-\newcommand{\ccontact}{Contact event }
-\newcommand{\bblackout}{Blackout event }
-\newcommand{\ddoomsday}{Doomsday event }
-
-\newcommand{\awakenings}{A events }
-\newcommand{\contacts}{C events }
-\newcommand{\blackouts}{B events }
-\newcommand{\doomsdays}{D events }
-
-\newcommand{\ff}{{\color{red}\rule{10pt}{10pt}}}
-
-%\newcommand{\ffn}[1]{ \colorbox{red}{FIG {#1}} }    
-\newcommand{\ffn}[1]{}
-
-%\newcommand{\ttn}[1]{ \colorbox{green}{TABLE {#1}} }    
-\newcommand{\ttn}[1]{}
-
 \begin{document}
@@
 \keywords{SETI, Computer simulations, Statistics}
-
-%\JELclassification{XX; XX}
-%\MSCcodes{XX; XX}
 
@@
 %
-A comprehensive review and an analysis of each term of the equation are
-presented in \citet{vakoch_drake_2015}.
+A comprehensive review and an analysis of each term of the equation
+are presented in \citet{vakoch_drake_2015}.
 %
 Optimistic estimates from the Drake equation contrast with the
-so-called Fermi paradox, which states the apparent contradiction between the
-expected abundance of life in the Galaxy and the lack of evidence for
-it \citep[e.g. ][]{hart_explanation_1975, brin_great_1983,
-barlow_galactic_2012, forgan_galactic_2016, anchordoqui_is_2017,
-Sotos_biotechnology_2019, carroll_nellemback_fermi_2019}.
+so-called Fermi paradox, which states the apparent contradiction
+between the expected abundance of life in the Galaxy and the lack of
+evidence for it \citep[e.g. ][]{hart_explanation_1975,
+brin_great_1983, barlow_galactic_2012, forgan_galactic_2017,
+anchordoqui_is_2017, Sotos_biotechnology_2019,
+carroll_nellemback_fermi_2019}.
 %
 There are many propositions aimed at solving this paradox, which make
-use of statistical \citep{solomonides_probabilistic_2016, horvat_calculating_2007,
-maccone_statistical_2015} or stochastic approaches
-\citep{forgan_numerical_2009, bloetscher_using_2019,
+use of statistical \citep{solomonides_probabilistic_2016,
+horvat_calculating_2007, maccone_statistical_2015} or stochastic
+approaches \citep{forgan_numerical_2009, bloetscher_using_2019,
 glade_stochastic_2011, forgan_numerical_2010}.
@@
 The absence of detections of extraterrestrial intelligent signals
-could be explained by
-astrophysical phenomena that makes life difficult to develop
-\citep{annis_astrophysical_1999}.
-%
-Besides the possible scarcity of life,
-alternative scenarios have also been discussed
-\citep{barlow_galactic_2013, lampton_information_2013,
+could be explained by astrophysical phenomena that makes life
+difficult to develop \citep{annis_astrophysical_1999}.
+%
+Besides the possible scarcity of life, alternative scenarios have also
+been discussed \citep{barlow_galactic_2013, lampton_information_2013,
 conway_three_2018, forgan_galactic_2017}.
@@
 The large distances in the Galaxy and the likely limited lifetime of
-civilisations may play an important role in determining how difficult it
-would be to obtain evidence for other inhabited worlds.
+civilisations may play an important role in determining how difficult
+it would be to obtain evidence for other inhabited worlds.
 %
@@
 The absence of detections has also motivated alternative ideas for new
-SETI strategies \citep{forgan_exoplanet_2017, balbi_impact_2018,
+SETI strategies \citep{forgan_exoplanet_2019, balbi_impact_2018,
 loeb_eavesdropping_2006, maccone_KLT_2010, tarter_advancing_2009,
@@
 \citep{fogg_temporal_1987, forgan_spatiotemporal_2011,
-balbi_impact_2018, balb_spatiotemporal_2018, horvat_impact_2011}, as
+balbi_impact_2018, balbi_spatiotemporal_2018, horvat_impact_2011}, as
 well as efforts on considering the stochastic nature of the Drake
@@
 %
-In a recent work, \citet{balbi_impact_2018} uses a statistical model to
-analyse the occurrence of causal contacts between civilisations in the
-Galaxy.
+In a recent work, \citet{balbi_impact_2018} uses a statistical model
+to analyse the occurrence of causal contacts between civilisations in
+the Galaxy.
 %
@@
 \citet{cirkovic_temporal_2004} also emphasises the lack of temporal
-structure in the Drake equation and, in particular, the limitations
-of this expression to estimate the required timescale of a SETI
-program to succeed in the detection of intelligent signals.
+structure in the Drake equation and, in particular, the limitations of
+this expression to estimate the required timescale of a SETI program
+to succeed in the detection of intelligent signals.
 % 
@@
 %
-The distances are drawn from a uniform model 
-for the
-positions of civilisations within the plane of the Galaxy.
+The distances are drawn from a uniform model for the positions of
+civilisations within the plane of the Galaxy.
 %
 For the distribution of the characteristic time of appearance, the
-author explores exponential and truncated Gaussian functions,
-while for the lifespans chooses an exponential behaviour.
+author explores exponential and truncated Gaussian functions, while
+for the lifespans chooses an exponential behaviour.
 %
@@
 %
-For all analysed distributions, the author concludes that the
-fraction of emitters that are listened is low if they are spreaded in
-time and with limited lifetimes.
-%
-An analytical explanation of these concepts are presented 
-in \citet{grimaldi_signal_2017}, who considers a statistical model for the
-probability of the Earth contacting other intelligent civilisations,
-taking into account the finite lifetime of signal emitters, and based
-on the fractional volume occupied by all signals reaching our planet.
-
-The quest for a formal statistical theory has also lead 
-to important progress in the mathematical foundations of SETI.
+For all analysed distributions, the author concludes that the fraction
+of emitters that are listened is low if they are spreaded in time and
+with limited lifetimes.
+%
+An analytical explanation of these concepts are presented in
+\citet{grimaldi_signal_2017}, who considers a statistical model for
+the probability of the Earth contacting other intelligent
+civilisations, taking into account the finite lifetime of signal
+emitters, and based on the fractional volume occupied by all signals
+reaching our planet.
+
+The quest for a formal statistical theory has also lead to important
+progress in the mathematical foundations of SETI.
 %
 Recently, \citet{bloetscher_using_2019} considers a Bayesian approach,
-still motivated by the Drake equation, to estimate
-the number of civilisations in the Galaxy.
-%
-To that end, the authors employ Monte Carlo Markov Chains over each factor
-of the Drake equation, and combine the mean values to reach a probabilistic
-result.
+still motivated by the Drake equation, to estimate the number of
+civilisations in the Galaxy.
+%
+To that end, the authors employ Monte Carlo Markov Chains over each
+factor of the Drake equation, and combine the mean values to reach a
+probabilistic result.
 %
@@
 %
-The author stresses that 
-the maximum broadcasting distance and the lifetime of civilisations
-come into play to produce the possible network of connections.
+The author stresses that the maximum broadcasting distance and the
+lifetime of civilisations come into play to produce the possible
+network of connections.
 
@@
 does not account for the network of causal contacts among
-technological civilisations, it offers a tool to think SETI searches
-from a novel point of view.
-
-% INTRODUCTION OF *THIS* WORK
+technological civilisations, it offers a tool to think SETI from a
+novel point of view.
 
@@
 %
-A node is causally connected to other node if it is within their light
-cone.
+A node is causally connected to other nodes if it is within their
+light cone.
 %
@@
 on the statistical distributions of the duration of causal contacts in
-and the distribution of time intervals of waiting for the first contact.
+and the distribution of time intervals of waiting for the first
+contact.
 %
@@
 
-Simulations are suitable tools to analyse systems that evolve with time and
-involve randomness.
-%
-An advantage of a numerical approach is that it usually requires fewer assumptions
-and simplifications, and can be applied to systems where analytical models are
-hard or impossible to develop.
-%
-In particular, a suitable tool to model complex stochastic processes through
-the changes in the state of a system is the discrete--event (hereafter, DE)
-simulation approach.
-%
-A system described with the DE paradigm is characterised by a set of actors and
-events.
-%
-Actors interact causally through a series of events on a timeline and process
-them in chronological order \citep{ptolemaeus_system_2014,
+Simulations are suitable tools to analyse systems that evolve with
+time and involve randomness.
+%
+An advantage of a numerical approach is that it usually requires fewer
+assumptions and simplifications, and can be applied to systems where
+analytical models are hard or impossible to develop.
+%
+In particular, a suitable tool to model complex stochastic processes
+through the changes in the state of a system is the discrete--event
+(hereafter, DE) simulation approach.
+%
+A system described with the DE paradigm is characterised by a set of
+actors and events.
+%
+Actors interact causally through a series of events on a timeline and
+process them in chronological order \citep{ptolemaeus_system_2014,
 chung_simulation_2003, ross_simulation_2012}.
@@
 This method is well suited for the particular case of the diffusion of
-intelligent signals in the galaxy and allows to explore several models easily.
-%
-We simulate the statistical properties of a set of points in space and time
-that have a causal connection at light speed and are separated
-by a maximum distance.
+intelligent signals in the galaxy and allows to explore several models
+easily.
+%
+We simulate the statistical properties of a set of points in space and
+time that have a causal connection at light speed and are separated by
+a maximum distance.
 %
@@
 %
-We choose this generic name in order to stress the fact that in this analysis 
-only the causal contact is considered, independently of any broadcasting or
-lookout activity.
-%
-The system we propose is ideal, in the sense that it considers the special case
-of a fully efficient node that emits and receives isotropically.
-%
-A causal contact node, considered as a broadcasting station that has the
-ability to detect signals through an active search program, could have a
-lesser than one efficiency factor, which is not included in the current
-analysis.
+We choose this generic name in order to stress the fact that in this
+analysis  only the causal contact is considered, independently of
+any broadcasting or lookout activity.
+%
+The system we propose is ideal, in the sense that it considers the
+special case of a fully efficient node that emits and receives
+isotropically.
+%
+A causal contact node, considered as a broadcasting station that has
+the ability to detect signals through an active search program, could
+have a lesser than one efficiency factor, which is not included in the
+current analysis.
 %
@@
 %
-It can be associated with a planet where life has developed,
-became intelligent, reached the skills required to find the right communication
-channel, sustained a search and established a contact.
-%
-Alternative message processing entities could be considered, for example
-interstellar beacons where intelligent life has ceased to exist but continue with its
-emission, or communication stations established by probes or left by
-intelligent beings \citep[see, e.g., ][]{peters_outer_2018,
-barlow_galactic_2013}.
-%
-In principle, these strategies could affect our results since it would be
-easier to configure a cluster of \cetis{} that spread in time.
-%
-However, we do not consider these speculative alternatives at this point.
-%
-For the purpose of the present analysis, only the communication capability is
-relevant, since we study the causal contacts between the locations.
-%
-The system is defined by a number of actors that represent \cetis{} and appear
-at different instants in time, generating events that produce meaningful
-changes in the variables that describe the system, i.e., in the arrangement of
-\cetis{} and their network of causal contacts.
-%
-For example, the appearance of a new \ceti{} in a region filled with a signal
-emitted by another node, will increase the number of active \cetis{} and the
-number of pairs of \cetis{} in causal contact.
-%
-Assuming some simple hypotheses, the discrete events method can be performed
-taking into account a small number of variables, which allow to analyze the
-variation of the results in the SC3Net model parameter space.
-
-In what follows, we outline the experimental setup adopted to estimate the
-probabilities of causal contacts and several derived quantities.
-%
-This is done in terms of three independent parameters, namely, the mean time
-span between the appearance of consecutive \cetis{}, $\tau_{\rm a}$,
-their mean lifetime $\tau_{\rm s}$, and the maximum distance a signal
-can be detected by another \ceti{} ($D_{\rm max}$).
-%
-Intuitively, the probability of the existence of causal contacts between pairs
-of \cetis{} would be larger for smaller $\tau_{\rm a}$ parameters,
-higher $\tau_{\rm s}$ or
-higher $D_{\rm max}$ parameter values.
-%
-We also propose theoretical distribution functions for both the lifespans
-($\tau_{\rm s}$) and the number of \cetis{} per unit time
-\citep{maccone_evolution_2014, Sotos_biotechnology_2019}.
-%
-The later is related to the time span between the appearance of consecutive
-\cetis{} (since when $\tau_{\rm a}$ is shorter, it produces a greater
-density of \cetis{}).
-%
-The analytical expresions for these distributions are set to a fixed law, as
-discussed in Sec.~\ref{SS_PDF_shape}.
-
-\ffn{1} Space--time diagrams, where time and space are represented on
-the horizontal and vertical axes, respectively, are suitable tools to
+It can be associated with a planet where life has developed, became
+intelligent, reached the skills required to find the right
+communication channel, sustained a search and established a contact.
+%
+Alternative message processing entities could be considered, for
+example interstellar beacons where intelligent life has ceased to
+exist but continue with its emission, or communication stations
+established by probes or left by intelligent beings \citep[see, e.g.,
+][]{peters_outer_2018, barlow_galactic_2013}.
+%
+In principle, these strategies could affect our results since it would
+be easier to configure a cluster of \cetis{} that spread in time.
+%
+However, we do not consider these speculative alternatives at this
+point.
+%
+For the purpose of the present analysis, only the communication
+capability is relevant, since we study the causal contacts between the
+locations.
+%
+The system is defined by a number of actors that represent \cetis{}
+and appear at different instants in time, generating events that
+produce meaningful changes in the variables that describe the system,
+i.e., in the arrangement of \cetis{} and their network of causal
+contacts.
+%
+For example, the appearance of a new \ceti{} in a region filled with a
+signal emitted by another node, will increase the number of active
+\cetis{} and the number of pairs of \cetis{} in causal contact.
+%
+Assuming some simple hypotheses, the discrete events method can be
+performed taking into account a small number of variables, which allow
+to analyze the variation of the results in the SC3Net model parameter
+space.
+
+In what follows, we outline the experimental setup adopted to estimate
+the probabilities of causal contacts and several derived quantities.
+%
+This is done in terms of three independent parameters, namely, the
+mean time span between the appearance of consecutive \cetis{},
+$\tau_{\rm a}$, their mean lifetime $\tau_{\rm s}$, and the maximum
+distance a signal can be detected by another \ceti{} ($D_{\rm max}$).
+%
+Intuitively, the probability of the existence of causal contacts
+between pairs of \cetis{} would be larger for smaller $\tau_{\rm a}$
+parameters, higher $\tau_{\rm s}$ or higher $D_{\rm max}$ parameter
+values.
+%
+We also propose theoretical distribution functions for both
+the lifespans ($\tau_{\rm s}$) and the number of \cetis{} per unit
+time \citep{maccone_evolution_2014, Sotos_biotechnology_2019}.
+%
+The later is related to the time span between the appearance of
+consecutive \cetis{} (since when $\tau_{\rm a}$ is shorter, it
+produces a greater density of \cetis{}).
+%
+The analytical expresions for these distributions are set to a fixed
+law, as discussed in Sec.~\ref{SS_PDF_shape}.
+
+Space--time diagrams, where time and space are represented on the
+horizontal and vertical axes, respectively, are suitable tools to
 describe the causal connections among different nodes.
@@
 In this scheme, a light pulse would follow a trajectory represented by
-a line at 45 degrees from the axes, given the units of time and
-space axes are years and light years, respectively.
-%
-Panel (a) of the represents the region $R$ (shaded polygon)
-which is causally connected to the emitter (left vertex).
+a line at 45 degrees from the axes, given the units of time and space
+axes are years and light years, respectively.
+%
+Panel (a) represents the region $R$ (shaded polygon) which is causally
+connected to the emitter (left vertex).
 %
@@
 %
-This sphere grows until its radius reaches the D$_{\rm max}$ distance, in
-which it would be no longer detectable due to the decrease in the
+This sphere grows until its radius reaches the D$_{\rm max}$ distance,
+in which it would be no longer detectable due to the decrease in the
 energy per unit area, which falls under the assumed detectability
@@
 This polygon is limited in the spatial axis by the maximum distance of
-the signal $D_{\rm max}$ and in the time axis by the time span between the
-A-type event (concave vertex) and the D-type event (convex vertex).
+the signal $D_{\rm max}$ and in the time axis by the time span between
+the A-type event (concave vertex) and the D-type event (convex
+vertex).
 %
@@
 Similarly, the instants in time where the \cetis{} lose their
-communicating capacity are dubbed ‘’Doomsday’’ events ($D_1$ and $D_2$).
+communicating capacity are dubbed ‘’Doomsday’’ events ($D_1$ and
+$D_2$).
 %
@@
 %
-Similarly, the receiver node E$_2$ can listen to signals from the emitter
-node E$_1$, from t=C$_{21}$ up to t=B$_{21}$.
+Similarly, the receiver node E$_2$ can listen to signals from the
+emitter node E$_1$, from t=C$_{21}$ up to t=B$_{21}$.
 %
@@
 
-The temporal structure that emerges from the experimental setup implies that,
-as a fundamental property of the simulations, a causal connection can be
-produced without  requiring that the two nodes are active at the same time.
-%
-This property of the system arises as a consequence of the large spatial and
-temporal scales, where a message is transmitted at the (relatively small) light
-speed.
-%
-Although a \ceti{} could be active for a large enough period to transmit a
-message at large distances, the limited power of the message and the dilution
-that depends on the squared distance from the source imposes a detectability
-limit.
-%
-As a consequence of this limitation and of their finite lifetime, considered as
-the period between the acquisition and loss of communicating capacity, each
-\ceti{} will fill a spherical shell region of the galaxy, limited by two
-concentric spherical surfaces.
-%
-The leading front, or surface of first contact (SFC) grows from the central
-\ceti{} until it reaches the maximum distance $D_{\rm max}$.
-%
-Following the end of the civilisation, there is still a region which is filled
-with the emitted signals.
-%
-This approach has been also considered in other statistical models \citep[e.g.,
-][]{smith_broadcasting_2009, grimaldi_signal_2017, Grimaldi2018}.
-%
-The trailing front, or surface of last contact (SLC) also grows from the
-central \ceti{}, with a delay with respect to the SFC equivalent to the
-lifetime of the node, and produces a spherical shell region.
-%
-Any other \ceti{} within this region will be in causal contact with the originating node,
-even if it has disappeared before the time of contact.
-%
-This region will grow if the surface of first contact has not yet reached the
-maximum distance $D_{\rm max}$, and will shrink after a D-type event until the
-surface of last contact reaches $D_{\rm max}$, producing as a result the loss
-of all signal from the central \ceti{}.
-%
-In our approach, we consider a model galaxy where the width of the disk is
-negligible with respect to the radius of the disk.
-%
-In the 2D simulation only the intersection of the communicating spherical
-shells with the plane of the galaxy is relevant, and produce the corresponding
-circles or rings for the filled spheres or annular regions of the spherical
-shells, respectively.
-%
-The initially growing communicating sphere is shown over space--time diagrams,
-where time is represented on the horizontal direction, and space is represented
-in the vertical direction.
-%
-In the Fig.~\ref{F_scheme} the two emitters in panel (b), E$_1$ and E$_2$,
-reach each other at different times.
+The temporal structure that emerges from the experimental setup
+implies that, as a fundamental property of the simulations, a causal
+connection can be produced without  requiring that the two nodes are
+active at the same time.
+%
+This property of the system arises as a consequence of the large
+spatial and temporal scales, where a message is transmitted at the
+(relatively small) light speed.
+%
+Although a \ceti{} could be active for a large enough period to
+transmit a message at large distances, the limited power of the
+message and the dilution that depends on the squared distance from the
+source imposes a detectability limit.
+%
+As a consequence of this limitation and of their finite lifetime,
+considered as the period between the acquisition and loss of
+communicating capacity, each \ceti{} will fill a spherical shell
+region of the galaxy, limited by two concentric spherical surfaces.
+%
+The leading front, or surface of first contact (SFC) grows from the
+central \ceti{} until it reaches the maximum distance $D_{\rm max}$.
+%
+Following the end of the civilisation, there is still a region which
+is filled with the emitted signals.
+%
+This approach has been also considered in other statistical models
+\citep[e.g., ][]{smith_broadcasting_2009, grimaldi_signal_2017,
+Grimaldi2018}.
+%
+The trailing front, or surface of last contact (SLC) also grows from
+the central \ceti{}, with a delay with respect to the SFC equivalent
+to the lifetime of the node, and produces a spherical shell region.
+%
+Any other \ceti{} within this region will be in causal contact with
+the originating node, even if it has disappeared before the time of
+contact.
+%
+This region will grow if the surface of first contact has not yet
+reached the maximum distance $D_{\rm max}$, and will shrink after a
+D-type event until the surface of last contact reaches $D_{\rm max}$,
+producing as a result the loss of all signal from the central \ceti{}.
+%
+In our approach, we consider a model galaxy where the width of the
+disk is negligible with respect to the radius of the disk.
+%
+In the 2D simulation only the intersection of the communicating
+spherical shells with the plane of the galaxy is relevant, and produce
+the corresponding circles or rings for the filled spheres or annular
+regions of the spherical shells, respectively.
+%
+The initially growing communicating sphere is shown over space--time
+diagrams, where time is represented on the horizontal direction, and
+space is represented in the vertical direction.
+%
+In the Fig.~\ref{F_scheme} the two emitters in panel (b), E$_1$ and
+E$_2$, reach each other at different times.
 %
@@
 %
-The type and length of causal contact in both directions depend on the distance
-and time lag between the awakening events, the maximum distance that a signal
-can reach and the time period in which each emitter is active.
-
-In our experimental configuration the simulation starts assuming that the
-stochastic process is already stable, and finishes before any galactic
-evolution effect could modify the fixed values of the variables.
-%
-Likewise, we assume that the probability for the appearance of a \ceti{} is
-homogeneous over the GHZ.
+The type and length of causal contact in both directions depend on the
+distance and time lag between the awakening events, the maximum
+distance that a signal can reach and the time period in which each
+emitter is active.
+
+In our experimental configuration the simulation starts assuming that
+the stochastic process is already stable, and finishes before any
+galactic evolution effect could modify the fixed values of the
+variables.
+%
+Likewise, we assume that the probability for the appearance of a
+\ceti{} is homogeneous over the GHZ.
 %
 The adopted geometry of the GHZ in all the simulations is given by a
-two-dimensional annular region,
-with an inner radius of 7~kpc and an outer radius of 9~kpc
-\citep{lineweaver_galactic_2004}.
-%
-Although the Galaxy has a well-known spiral structure, the nodes are assumed
-to be sparse (otherwise the Galaxy would be full of life) and the spiral
-structure would not, in that case, produce significant differences.
-%
-If the distribution of nodes is not sparse, as it could be the case if the
-spiral arms host most of the \cetis{}, then contacts would be more frequent
-between closely located nodes.
-%
-In such a case our results underestimate the number of contacts between
-close pairs of nodes within the same spiral arm, and conversely, overestimate
-the number of contacts between separated nodes.
-%
-We also limit the possibilities of life or other types of civilisations to the
-usually stated hypotheses for the definition of the GHZ
-\citep{dayal_habitability_2016, gonzalez_galactic_2001,
-lineweaver_galactic_2004, gonzalez_habitable_2005, morrison_extending_2015,
-haqq-misra_evolution_2019, rahvar_cosmic_2016, gobat_evolution_2016,
-rahvar_cosmic_2016} and consider that habitability remains constant over time
-\citep[see, however, ][]{gonzalez_habitable_2005, dayal_habitability_2016,
+two-dimensional annular region, with an inner radius of 7~kpc and an
+outer radius of 9~kpc \citep{lineweaver_galactic_2004}.
+%
+Although the Galaxy has a well-known spiral structure, the nodes are
+assumed to be sparse (otherwise the Galaxy would be full of life) and
+the spiral structure would not, in that case, produce significant
+differences.
+%
+If the distribution of nodes is not sparse, as it could be the case if
+the spiral arms host most of the \cetis{}, then contacts would be
+more frequent between closely located nodes.
+%
+In such a case our results underestimate the number of contacts
+between close pairs of nodes within the same spiral arm, and
+conversely, overestimate the number of contacts between separated
+nodes.
+%
+We also limit the possibilities of life or other types of
+civilisations to the usually stated hypotheses for the definition of
+the GHZ \citep{dayal_habitability_2016, gonzalez_galactic_2001,
+lineweaver_galactic_2004, gonzalez_habitable_2005,
+morrison_extending_2015, haqq-misra_evolution_2019,
+rahvar_cosmic_2016, gobat_evolution_2016, rahvar_cosmic_2016} and
+consider that habitability remains constant over time \citep[see,
+however, ][]{gonzalez_habitable_2005, dayal_habitability_2016,
 gobat_evolution_2016}.
 %
-This means that we set aside civilisations that could survive in severe
-conditions or unstable systems, which would prevent the appearance of life as
-we know it.
-
-We stress the fact that we are
-considering causal contacts instead of actual contacts.
-%
-In more realistic scenarios, there are several sources of ``signal loss'' with
-respect to the ideal case.
-%
-Among them, we can mention temporal and signal power aspects and direction
-dependent communication capabilities.
-%
-The results must then be interpreted as the case of ideal nodes, with a perfect
-efficiency in the emision and reception of signals.
-%
-The probabilities of contacts presented here are then upper limits to the
-number of communications between civilisations in the Galaxy, given the
-implemented hipotheses in the model.
-%
-The use of light cones as causal contact regions is inspired by the fact that
-light-speed traveling messengers like electromagnetic radiation, gravitational
-radiation or neutrinos are often considered as possible message carriers
-\citep{hippke_interstellar_2017, wright_how_2018}.
+This means that we set aside civilisations that could survive in
+severe conditions or unstable systems, which would prevent the
+appearance of life as we know it.
+
+We stress the fact that we are considering causal contacts instead of
+actual contacts.
+%
+In more realistic scenarios, there are several sources of ``signal
+loss'' with respect to the ideal case.
+%
+Among them, we can mention temporal and signal power aspects and
+direction dependent communication capabilities.
+%
+The results must then be interpreted as the case of ideal nodes, with
+a perfect efficiency in the emision and reception of signals.
+%
+The probabilities of contacts presented here are then upper limits to
+the number of communications between civilisations in the Galaxy,
+given the implemented hipotheses in the model.
+%
+The use of light cones as causal contact regions is inspired by the
+fact that light-speed traveling messengers like electromagnetic
+radiation, gravitational radiation or neutrinos are often considered
+as possible message carriers \citep{hippke_interstellar_2017,
+wright_how_2018}.
 %
 This excludes messages sent with mechanical means or physical objects
-\citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some unknown
-technology that violates the known laws of physics.
-
-As part of this benchmark, we assume that the capacity to emit and receive
-signals occur at the same time.
-%
-Although there are several reasons to think that this could not be the case, at
-large time scales it can be considered that both abilities occur roughly at
-coincident epochs.
-%
-In the ideal setup, this would be equivalent to nodes that send messages
-isotropically and scan the local skies on all directions with a perfect
-eficiency.
-%
-Another essential assumption is that all \cetis{} use the same signal power, so
-that there is a maximum distance out to which it can be detected.
-%
-In such a system we compute probabilities of a random node making contact with
-another node, i.e., they are not specific for the case of contacts with the
-Earth.
-%
-Regarding the extent of the signals, we know that the distance from which a
-signal from Earth could be detected using the current technology is about few
-parsecs, given that the signal was sent to a specific direction.
-%
-It is straightforward to propose and implement a distribution of maximum
-distances, although this would increase the model complexity at the cost of a
-larger uncertainty.
+\citep[e.g., ][]{Armstrong2013, barlow_galactic_2013}, or through some
+unknown technology that violates the known laws of physics.
+
+As part of this benchmark, we assume that the capacity to emit and
+receive signals occur at the same time.
+%
+Although there are several reasons to think that this could not be the
+case, at large time scales it can be considered that both abilities
+occur roughly at coincident epochs.
+%
+In the ideal setup, this would be equivalent to nodes that send
+messages isotropically and scan the local skies on all directions with
+a perfect efficiency.
+%
+Another essential assumption is that all \cetis{} use the same signal
+power, so that there is a maximum distance out to which it can be
+detected.
+%
+In such a system we compute probabilities of a random node making
+contact with another node, i.e., they are not specific for the case of
+contacts with the Earth.
+%
+Regarding the extent of the signals, we know that the distance from
+which a signal from Earth could be detected using the current
+technology is about few parsecs, given that the signal was sent to a
+specific direction.
+%
+It is straightforward to propose and implement a distribution of
+maximum distances, although this would increase the model complexity
+at the cost of a larger uncertainty.
 
@@
 %
-In particular, we do not consider the
-possibility of stellar colonisation
-\citep[e.g.][]{newman_galactic_1981, walters_interstellar_1980,
-starling_virulence_2013, barlow_galactic_2012, jeong_large_2000,
-maccone_mathematical_2011}.
-%
-We also assume that communication is
-equally likely in all directions, i.e., we assume isotropic
-communication in all cases.
-%
-Different communication efficiencies or
-detection methods are straightforward to carry out in the simulations
-for more detailed and complex scenarios.
-%
-This approach, however, is
-beyond this work because it obscures the experimental setup and make
-the results less clear.
-%
-The assumptions we accept imply that the
-results are independent of whether intelligent agents are organic or
-artificial.
-%
-Moreover, the causes of the limited lifetime of a
-civilisation can be natural (astrophysical phenomena), caused by auto
-destruction or by external factors, to name a few.
-%
-However, we assume
-that these events are sufficiently numerous in the Galaxy so that a
-statistical model is plausible. The failure of this hypothesis would
-imply extreme values for the parameters that represent the density of
-the nodes (i.e., $\tau_{\rm a}$).
+In particular, we do not consider the possibility of stellar
+colonisation \citep[e.g.][]{newman_galactic_1981,
+walters_interstellar_1980, starling_virulence_2013,
+barlow_galactic_2012, jeong_large_2000, maccone_mathematical_2011}.
+%
+We also assume that communication is equally likely in all directions,
+i.e., we assume isotropic communication in all cases.
+%
+Different communication efficiencies or detection methods are
+straightforward to carry out in the simulations for more detailed and
+complex scenarios.
+%
+This approach, however, is beyond this work because it obscures the
+experimental setup and make the results less clear.
+%
+The assumptions we accept imply that the results are independent
+of whether intelligent agents are organic or artificial.
+%
+Moreover, the causes of the limited lifetime of a civilisation can be
+natural (astrophysical phenomena), caused by auto destruction or by
+external factors, to name a few.
+%
+However, we assume that these events are sufficiently numerous in the
+Galaxy so that a statistical model is plausible. The failure of this
+hypothesis would imply extreme values for the parameters that
+represent the density of the nodes (i.e., $\tau_{\rm a}$).
 
@@
 \hline
-   A event & Awakening: a \ceti{} starts its communication
-   capabilities &Number of active \cetis{}\\
-   B event & Blackout: the end of the communication channel stops
-   &Number causal channels\\
+   A event & Awakening: a \ceti{} starts its communication capabilities &Number of active \cetis{}\\
+   B event & Blackout: the end of the communication channel stops &Number causal channels\\
    C event & Contact: a new causal contact is produced &Number of causal channels\\
@@
 \end{tabular}
-\caption{Definition of independent variables and adopted values for 
-   fixed parameters 
-   that are part of the simulation.  Variable parameters define the
-   spatial and temporal structure
-   of the process and the maximum reach of the messages.}
+\caption{
+   %
+   Definition of independent variables and adopted values for fixed
+   parameters that are part of the simulation.  Variable parameters
+   define the spatial and temporal structure of the process and the
+   maximum reach of the messages.}
+   %
 \label{T_simu_hypotheses}
@@
 The temporal structure of the process is defined by two distribution
-parameters. One of them represent the mean time interval that a node can
-emit and receive signals (its lifetime), and the other the mean time interval between the
-emergence of consecutive \cetis{}.
-%
-The spatial structure of the simulation is
-given by the size and shape of the Galactic Habitable Zone and the maximum
-distance a signal can travel to be detected (D$_{\rm max}$).
-%
-The parameters for
-the temporal distributions also determine the spatial properties, since the
-density of active \cetis{} in the galaxy depend on these two parameters.
-%
-For
-example, small $\tau_{\rm a}$ and a large $\tau_{\rm s}$ will produce a densely
-populated galaxy (in this context, a galaxy is an element in a 
-statistical ensemble, not the Milky Way).
-%
-Also, some hypotheses regarding the shapes of the
-distributions of the temporal parameters must be made in order to complete the
-simulation.
+parameters. One of them represent the mean time interval that a node
+can emit and receive signals (its lifetime), and the other the mean
+time interval between the emergence of consecutive \cetis{}.
+%
+The spatial structure of the simulation is given by the size and shape
+of the Galactic Habitable Zone and the maximum distance a signal can
+travel to be detected (D$_{\rm max}$).
+%
+The parameters for the temporal distributions also determine
+the spatial properties, since the density of active \cetis{} in the
+galaxy depend on these two parameters.
+%
+For example, small $\tau_{\rm a}$ and a large $\tau_{\rm s}$ will
+produce a densely populated galaxy (in this context, a galaxy is an
+element in a statistical ensemble, not the Milky Way).
+%
+Also, some hypotheses regarding the shapes of the distributions of the
+temporal parameters must be made in order to complete the simulation.
 %
 \citet{forgan_spatiotemporal_2011} argues that the times at which
-different civilisations become intelligent follow a Gaussian distribution, and
-then the distribution of inter-arrival times is an inverse exponential.
-%
-We
-assume that the distribution of the times of A events is a stationary Poisson
-process, and then the distribution of the times between the appearance of new
-consecutive nodes is exponential.
-%
-Regarding the duration of a node, we
-propose that its distribution is a stationary
-exponential distribution.
-%
-There are no clear arguments to conclude a
-statistical law for the later distribution, so that we propose it as a working
-hipothesis.  
-%
-This heuristic does not make any consideration about the origin of life, although
-different approaches are possible.
-%
-For example,
-\citet{maccone_lognormals_2014} argues that this distribution should be a log-normal. 
-%
-Preferentially, a theoretical statistical distribution of the lifespan of
-civilisations would rely on the basis of the underlying
+different civilisations become intelligent follow a Gaussian
+distribution, and then the distribution of inter-arrival times is an
+inverse exponential.
+%
+We assume that the distribution of the times of A events is a
+stationary Poisson process, and then the distribution of the times
+between the appearance of new consecutive nodes is exponential.
+%
+Regarding the duration of a node, we propose that its distribution is
+a stationary exponential distribution.
+%
+There are no clear arguments to conclude a statistical law for the
+later distribution, so that we propose it as a working hipothesis.  
+%
+This heuristic does not make any consideration about the origin of
+life, although different approaches are possible.
+%
+For example, \citet{maccone_lognormals_2014} argues that this
+distribution should be a log-normal. 
+%
+Preferentially, a theoretical statistical distribution of the lifespan
+of civilisations would rely on the basis of the underlying
 astrophysical and biological processes \citep{balbi_impact_2018}.
 
-The power law and exponential statistical distributions are among the most
-common patterns found in natural phenomena.
-%
-For example, the distribution of the frequency of words in many languages is
-known to follow the law of Zipf (which is a power law).
-%
-These distributions arise any time a phenomenon is characterised by commonly
-occurring small events and rarely produced large events \citep[e.g.
-][]{adamic_zipf_2000}.
-%
-Zipf law also describes population ranks of cities in various countries,
-corporation sizes, income rankings, ranks of the number of people watching the
-same TV channel, etc.
-%
-The magnitudes of earthquakes, hurricanes, volcanic eruptions and floods; the
-sizes of meteorites or the losses caused by business interruptions from
-accidents, are also well described by power laws
+The power law and exponential statistical distributions are among the
+most common patterns found in natural phenomena.
+%
+For example, the distribution of the frequency of words in many
+languages is known to follow the law of Zipf (which is a power law).
+%
+These distributions arise any time a phenomenon is characterised by
+commonly occurring small events and rarely produced large
+events \citep[e.g.  ][]{adamic_zipf_2000}.
+%
+Zipf law also describes population ranks of cities in
+various countries, corporation sizes, income rankings, ranks of
+the number of people watching the same TV channel, etc.
+%
+The magnitudes of earthquakes, hurricanes, volcanic eruptions and
+floods; the sizes of meteorites or the losses caused by business
+interruptions from accidents, are also well described by power laws
 \citep{sornette_critical_2006}.
 %
-Additional examples include stock market fluctuations, sizes of computer files
-or word frequency in languages \citep{mitzenmacher_brief_2004,
-newman_power_2005, simkin_theory_2006}. 
+Additional examples include stock market fluctuations, sizes of
+computer files or word frequency in languages
+\citep{mitzenmacher_brief_2004, newman_power_2005,
+simkin_theory_2006}. 
 %
@@
 %
-Some examples are the analysis of connectivity patterns in metabolic networks
-\citep{jeong_large_2000} or the number of species observed per unit area in
-ecology \citep{martin_origin_2006, frank_common_2009}.
-%
-More examples can be found in the literature \citep{martin_origin_2006,
-maccone_KLT_2010, barabasi_scale_2009, maccone_evolution_2014, maccone_lognormals_2014}.
+Some examples are the analysis of connectivity patterns in metabolic
+networks \citep{jeong_large_2000} or the number of species observed
+per unit area in ecology \citep{martin_origin_2006,
+frank_common_2009}.
+%
+More examples can be found in the literature
+\citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009,
+maccone_evolution_2014, maccone_lognormals_2014}.
 %
 This distribution family is suitable for the statistical description
-of the $D_{\rm max}$ parameter, although in this work we assume a uniform
-distribution for simplicity.
-%
-The power law family of functions is also a good candidate for the description of
-the temporal variables in the model.
+of the $D_{\rm max}$ parameter, although in this work we assume a
+uniform distribution for simplicity.
+%
+The power law family of functions is also a good candidate for the
+description of the temporal variables in the model.
 %
@@
 %
-The exponential distribution of lifespan and waiting times is justified by
-considering the hypothesis that the process of appearance of life in the galaxy
-is homogeneous and stationary.
-%
-That is, there is not a preferred region within
-the GHZ for the spontaneous appearance of life, and the emergence of a node is
+The exponential distribution of lifespan and waiting times is
+justified by considering the hypothesis that the process of appearance
+of life in the galaxy is homogeneous and stationary.
+%
+That is, there is not a preferred region within the GHZ for the
+spontaneous appearance of life, and the emergence of a node is
 independent of the existence of previous nodes in the galaxy.
 %
-These seem to be simple conditions, and allow to propose a distribution family
-without knowing the details of the underlying process.
-%
-The exponential distribution for the separations in time is equivalent to
-proposing a Poisson process for the emergence of nodes, given the relation
-between the number of events in time or space and the waiting time or
-separation, respectively \citep[e.g., ][]{ross_simulation_2012}.
-%
-That is, these
-are two alternative approaches to describing the same process, a Poisson
-distribution for the number of events implies an exponential distribution for
-their separations, and vice versa.
-%
-It should be emphasised that the exponential
-laws used in this work are assumed as part of the working hypothesis, and
-instead of analysing results from a particular parameter chosen ad hoc, we
-explore the parameter space and analyse the impact of the values of these
-parameters on the results.
+These seem to be simple conditions, and allow to propose a
+distribution family without knowing the details of the underlying
+process.
+%
+The exponential distribution for the separations in time is equivalent
+to proposing a Poisson process for the emergence of nodes, given the
+relation between the number of events in time or space and the waiting
+time or separation, respectively \citep[e.g.,
+][]{ross_simulation_2012}.
+%
+That is, these are two alternative approaches to describing the same
+process, a Poisson distribution for the number of events implies an
+exponential distribution for their separations, and vice versa.
+%
+It should be emphasised that the exponential laws used in this work
+are assumed as part of the working hypothesis, and instead of
+analysing results from a particular parameter chosen ad hoc, we
+explore the parameter space and analyse the impact of the values of
+these parameters on the results.
 
@@
 
-In this Section we discuss the degree of complexity in the model, considering a
-compromise between the accuracy of the model and the number of parameters that
-are free or with a high uncertainty.
-%
-Firstly, we emphasise that the odds of a causal contact between two nodes
-should not be considered as the odds of a contact between two intelligent
-civilisations, and in fact the latter could be much lesser than the former.
-%
-Indeed, in order to establish a contact between any two entities, a minimum
-degree of compatibility must be accomplished without any previous agreement,
-making the possibility of a contact with a message that could be deciphered
-highly rare \citep[see e.g. ][]{forgan_collimated_2014}.
-%
-Besides the trade--off between the simplicity and the complexity of the
-experiments, further analysis could be performed following this framework in
-order to explore possible implications of the results for more detailed
-configurations.
-%
-For example, the communication method (isotropic, collimated, serendipitous)
-can affect the observables, making it necessary to implement a correction
-factor.
-%
-Taking this into account, our results regarding the probability of causal
-contact should be considered as upper limits for effective contacts, since they
-depend on the efficiency of both the emitter and the receiver to broadcast and
-scan the sky for intelligent signals, respectively
+In this Section we discuss the degree of complexity in the model,
+considering a compromise between the accuracy of the model and the
+number of parameters that are free or with a high uncertainty.
+%
+Firstly, we emphasise that the odds of a causal contact between two
+nodes should not be considered as the odds of a contact between two
+intelligent civilisations, and in fact the latter could be much
+lesser than the former.
+%
+Indeed, in order to establish a contact between any two entities, a
+minimum degree of compatibility must be accomplished without any
+previous agreement, making the possibility of a contact with a message
+that could be deciphered highly rare \citep[see e.g.
+][]{forgan_collimated_2014}.
+%
+Besides the trade--off between the simplicity and the complexity of
+the experiments, further analysis could be performed following
+this framework in order to explore possible implications of the
+results for more detailed configurations.
+%
+For example, the communication method (isotropic, collimated,
+serendipitous) can affect the observables, making it necessary
+to implement a correction factor.
+%
+Taking this into account, our results regarding the probability of
+causal contact should be considered as upper limits for effective
+contacts, since they depend on the efficiency of both the emitter and
+the receiver to broadcast and scan the sky for intelligent signals,
+respectively \citep{grimaldi_signal_2017}.
+%
+A correction by a coverage ratio in the detection and by a targeting
+ratio in the emission could be easily implemented in the simulation,
+although the effect of reducing the probability of contact is
+basically the product of the efficiency ratios and thus such
+implementation is not necessary.
+%
+Therefore, the values of the probabilities could be modified by a
+constant correction factor equal to the combined emission/reception
+efficiency \citep{smith_broadcasting_2009, anchordoqui_upper_2019,
+forgan_collimated_2014} or for beam-like transmissions
 \citep{grimaldi_signal_2017}. 
 %
-A correction by a coverage ratio in the detection and by a targeting ratio in the
-emission could be easily implemented in the simulation, although the effect of
-reducing the probability of contact is basically the product of the efficiency
-ratios and thus such implementation is not necessary.
-%
-Therefore, the values of the probabilities could be modified by a constant
-correction factor equal to the combined emission/reception efficiency
-\citep{smith_broadcasting_2009, anchordoqui_upper_2019, forgan_collimated_2014}
-or for beam-like transmissions \citep{grimaldi_signal_2017}. 
-%
-Other considerations include the effects of alignments, the use of stars as
-sources or amplifiers \citep{Edmondson2003, borra_searching_2012}, or the
-nature of the message carrier.
-%
-Another improvement could be the use of a spatial distribution that resembles
-the spiral shape and the width of the disc of the Galaxy.
-%
-Regarding the distributions of the parameters, different distribution functions
-for the mean lifetime of \cetis{} can be implemented as alternatives.
+Other considerations include the effects of alignments, the use of
+stars as sources or amplifiers \citep{Edmondson2003,
+borra_searching_2012}, or the nature of the message carrier.
+%
+Another improvement could be the use of a spatial distribution that
+resembles the spiral shape and the width of the disc of the Galaxy.
+%
+Regarding the distributions of the parameters, different distribution
+functions for the mean lifetime of \cetis{} can be implemented as
+alternatives.
 %
 Regarding the degree of idealisation, the model admits different
-efficiencies for the sphere of the causal region of each \ceti{}, featuring
-different searching strategies \citep{hippke_interstellar_2017}. 
-%
-It is also possible to consider that $D_{\rm max}$ is different for different
-\cetis{}.
-%
-For example, a power law where a powerful emission is rare and a low emission
-is common could be an improvement to the model.
-%
-In this work we choose not to implement this for the sake of simplicity.
-%
-Finally, the role of the message contents could influence on the lifespan of a
-\ceti{} that receives a message, although the implementation of such behaviour
-would increase the number of free parameters and would be more speculative.   
+efficiencies for the sphere of the causal region of each \ceti{},
+featuring different searching strategies
+\citep{hippke_interstellar_2017}. 
+%
+It is also possible to consider that $D_{\rm max}$ is different for
+different \cetis{}.
+%
+For example, a power law where a powerful emission is rare and a low
+emission is common could be an improvement to the model.
+%
+In this work we choose not to implement this for the sake of
+simplicity.
+%
+Finally, the role of the message contents could influence on the
+lifespan of a \ceti{} that receives a message, although the
+implementation of such behaviour would increase the number of free
+parameters and would be more speculative.   
 %
@@
 %
-Once the model has been defined, it can be implemented as a discrete event
-simulation, as described in detail in the next Section.
+Once the model has been defined, it can be implemented as a discrete
+event simulation, as described in detail in the next Section.
 
@@
 A discrete event simulation is performed for a given model, in this
-work $M(\tau_{\rm a}, \tau_{\rm s},
-D_{\rm max})$, by keeping track of a set of variables that change each time an
-event happens. 
-%
-In addition, the model comprises elements that are fixed for all simulations,
-for example, the functional forms of the statistical distributions and the
-adopted values of particular variables (see \ref{T_simu_hypotheses}).
+work $M(\tau_{\rm a}, \tau_{\rm s}, D_{\rm max})$, by keeping track of
+a set of variables that change each time an event happens. 
+%
+In addition, the model comprises elements that are fixed for all
+simulations, for example, the functional forms of the statistical
+distributions and the adopted values of particular variables (see
+\ref{T_simu_hypotheses}).
 %
 The main variables that follow the evolution of the simulation are:
-the positions of stars, which are sampled randomly within the GHZ; the time of
-the awakening of each node (A event); and its time of disappearance (D event).
-%
-The variables that can be deduced from the previous ones include: the number of
-\cetis{} in casual contact with at least another \ceti{} at a given time; the
-number of \cetis{} as a function of time; the number of \cetis{} that receive at
-least one message; the number of \cetis{} that receive a message at
-least one time and successfully deliver an answer; and the number distribution
-of waiting times to receive a message.
-%
-All these quantities are updated each time one of the four events (A, B, C, D)
-occurs.
+the positions of stars, which are sampled randomly within the GHZ; the
+time of the awakening of each node (A event); and its time of
+disappearance (D event).
+%
+The variables that can be deduced from the previous ones include:
+the number of \cetis{} in casual contact with at least another \ceti{}
+at a given time; the number of \cetis{} as a function of time; the
+number of \cetis{} that receive at least one message; the number of
+\cetis{} that receive a message at least one time and successfully
+deliver an answer; and the number distribution of waiting times to
+receive a message.
+%
+All these quantities are updated each time one of the four events (A,
+B, C, D) occurs.
 
@@
 
-% hearsay 
-%
 In order to make a reproducible project, we developed the tools that
@@
 
-We implemented the simulation of a regular grids of models varying over
-the parameter space, which covers 5204 models.
-%
-For each model, we simulated several realizations with different random
-seeds, adding up a total of 
-158546 simulation runs.
-%
-The number of random realzations varies from one (for the densest populated models)
-up to 50 (for the sparsely populated models).
+We implemented the simulation of a regular grids of models varying
+over the parameter space, which covers 5204 models.
+%
+For each model, we simulated several realizations with different
+random seeds, adding up a total of 158546 simulation runs.
+%
+The number of random realzations varies from one (for the densest
+populated models) up to 50 (for the sparsely populated models).
 %
 The parameters for the temporal aspects of the simulation (the mean
-waiting time for the next awakening, $\tau_{\rm a}$, and the
-mean lifetime, $\tau_{\rm s}$) cover the ranges
-10$^2$-10$^6$~yr, with 
-two linear partitions of 21 values (10$^2$-40500~yr) and 51 values
-($5\cdot10^5$-10$^6$~yr)
-for each parameter.
-%
-This partition was chosen after the requirement of the software to take linear bins, 
-aiming at a better sampling of the low $\tau_{\rm a}$ and low
-$\tau_{\rm s}$ region of the parameter space.
-%
-For the D$_{\rm max}$ parameter, we take the 
-values 100~pc, 500~pc, 1~kpc, 5~kpc and 10~kpc.
-%
-\ttn{1}
+waiting time for the next awakening, $\tau_{\rm a}$, and the mean
+lifetime, $\tau_{\rm s}$) cover the ranges 10$^2$-10$^6$~yr, with two
+linear partitions of 21 values (10$^2$-40500~yr) and 51 values
+($5\cdot10^5$-10$^6$~yr) for each parameter.
+%
+This partition was chosen after the requirement of the software to
+take linear bins, aiming at a better sampling of the low $\tau_{\rm
+a}$ and low $\tau_{\rm s}$ region of the parameter space.
+%
+For the D$_{\rm max}$ parameter, we take the values 100~pc, 500~pc,
+1~kpc, 5~kpc and 10~kpc.
 %
@@
 %
-We also show the set of fixed parameters that take part in the simulation,
-their values and the hypotheses that they represent.
+We also show the set of fixed parameters that take part in the
+simulation, their values and the hypotheses that they represent.
 
@@
 the fraction of the lifetime a node is listening to at least another
-node (i.e., within their light cone),
-%
-the age of contacted \cetis{} at first contact,
-%
-the fraction of \cetis{} where the first contact is given at the
-awakening,
-%
-the distribution of the number of contacts for each \ceti{},
-%
-the distribution of the number of contacts as a function of the age of
-the node,
-%
-the number of contacts as a function of time in the galaxy,
-%
-the fractions of nodes that succeed in making contact,
-%
+node (i.e., within their light cone), the age of contacted \cetis{} at
+first contact, the fraction of \cetis{} where the first contact is
+given at the awakening, the distribution of the number of contacts for
+each \ceti{}, the distribution of the number of contacts as a function
+of the age of the node, the number of contacts as a function of time
+in the galaxy, the fractions of nodes that succeed in making contact,
 and the distribution of distances between contacted nodes.
@@
 
-\ffn{3}
-%
 In Fig.~\ref{F_number_of_contacts} we show the empirical cumulative
@@
 model.
-%
-\ffn{4}
 %
@@
 The results of the simulations that comprise the full range of values
-for $\tau_{\rm a}$ and $\tau_{\rm s}$ from 5$\cdot10^2$~yr to $10^6$~yr
-(not shown in the Figures) maintain a similar trend.
+for $\tau_{\rm a}$ and $\tau_{\rm s}$ from 5$\cdot10^2$~yr to
+$10^6$~yr (not shown in the Figures) maintain a similar trend.
 %
@@
 of their A-type events.
-%
-\ffn{5}
 %
@@
 
-In this subsection we analyze the distribution of the waiting times for
-a first contact.
+In this subsection we analyze the distribution of the waiting times
+for a first contact.
 %
 Such distribution can be considered to compute the probability for a
-random \ceti{} to carry out the searching of other nodes and spend a given time until
-the first contact is made, under the hypotheses of the experiment.
-%
-\ffn{6}
+random \ceti{} to carry out the searching of other nodes and spend a
+given time until the first contact is made, under the hypotheses of
+the experiment.
 %
@@
 Upper panel (a) panel shows the histograms for several values of the
-mean awakening time $\tau_{\rm a}$, with the mean survival time $\tau_{\rm
-s}$ of 10~kyr.
-%
-Lower panel (b) shows the histograms for several values of $\tau_{\rm s}$,
-with $\tau_{\rm a}$=10~kyr.
+mean awakening time $\tau_{\rm a}$, with the mean survival time
+$\tau_{\rm s}$ of 10~kyr.
+%
+Lower panel (b) shows the histograms for several values of $\tau_{\rm
+s}$, with $\tau_{\rm a}$=10~kyr.
 %
@@
 As it can be seen, there is a clear trend where the number of nodes
-that require a time $t_{\rm 1}$ to make the first contact 
-decreases exponentially with the time $t_{\rm 1}$.
+that require a time $t_{\rm 1}$ to make the first contact decreases
+exponentially with the time $t_{\rm 1}$.
 %
@@
 %
-Thus, it does not need a detailed description or modeling of the physical
-processes that give rise to intelligent life.
+Thus, it does not need a detailed description or modeling of the
+physical processes that give rise to intelligent life.
 %
@@
 %
-Despite those differences, their results and ours are in general agreement.
+Despite those differences, their results and ours are in general
+agreement.
 
@@
 
-In the approach of this work, we have used computer simulations to address
-the problem of the probabilities of causal contacts between locations
-in the Galaxy with the possibility of sending and receiving messages.
+In the approach of this work, we have used computer simulations to
+address the problem of the probabilities of causal contacts between
+locations in the Galaxy with the possibility of sending and receiving
+messages.
 %
@@
 parameter space, reducing the problem to only three parameters and a
-few simple hypotheses to perform a complete model for the population and
-communication network in the galaxy.
+few simple hypotheses to perform a complete model for the population
+and communication network in the galaxy.
 %
@@
 %
-If the time intervals between the rise and fall of civilisations are short
-compared to the time required for an electromagnetic signal to travel the 
-large distances in the Galaxy, 
-then the number of contacts would be limited to a low number.
-
-The short time interval between the
-rise and fall of civilisations, compared to the age and extension of
-our Galaxy, is a fundamental limitation for the number of contacts.
-%
-The temporal dimension, which is missing in the Drake
-equation, is a key factor to understand the network of contacts on
-different scenarios.
+If the time intervals between the rise and fall of civilisations are
+short compared to the time required for an electromagnetic signal to
+travel the large distances in the Galaxy, then the number of contacts
+would be limited to a low number.
+
+The short time interval between the rise and fall of civilisations,
+compared to the age and extension of our Galaxy, is a fundamental
+limitation for the number of contacts.
+%
+The temporal dimension, which is missing in the Drake equation, is a
+key factor to understand the network of contacts on different
+scenarios.
 
@@
 Simulations were run on the Clemente cluster at the Instituto de
-Astronomía Teórica y Experimental (IATE).
+Astronom\'{\i}a Te\'orica y Experimental (IATE).
 %
@@
 
-\bibliographystyle{mn2e}
-\setlength{\bibsep}{0.0pt}
-%\bibliography{biblio_seti, biblio_software}
-
-\begin{thebibliography}{95}
-\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
-
-\bibitem[{Adamic(2000)}]{adamic_zipf_2000}
-Adamic, L.A., 2000. Zipf, power-laws, and pareto-a ranking tutorial. Xerox Palo Alto Research Center, Palo Alto, CA, http://ginger. hpl. hp. com/shl/papers/ranking/ranking. html.
-
-\bibitem[{{Allen} {et~al}\mbox{.}(2020){Allen}, {Berriman}, {DuPrie}, {Mink},
-  {Nemiroff}, {Schmidt}, {Shamir}, {Shortridge}, {Taylor}, \&
-  {Teuben}}]{2020ASPC..522..731A}
-{Allen} A. {et~al.}, 2020, in Astronomical Society of the Pacific Conference
-  Series, Vol. 522, Astronomical Data Analysis Software and Systems XXVII,
-  {Ballester} P., {Ibsen} J., {Solar} M., {Shortridge} K., eds., p. 731
-
-\bibitem[{{Allen} \& {Schmidt}(2015)}]{2015JORS....3E..15A}
-{Allen} A., {Schmidt} J., 2015, Journal of Open Research Software, 3, E15
-
-\bibitem[{Anchordoqui \& Weber(2019)}]{anchordoqui_upper_2019}
-Anchordoqui L.~A., Weber S.~M., 2019, eprint arXiv:1908.01335
-
-\bibitem[{Annis(1999)}]{annis_astrophysical_1999}
-Annis J., 1999, JBIS - Journal of the British Interplanetary Society, 52, 19
-
-\bibitem[{Armstrong \& Sandberg(2013)}]{Armstrong2013}
-Armstrong S., Sandberg A., 2013, Acta Astronautica, 89, 1
-
-\bibitem[{Balb(2018)}]{balb_spatiotemporal_2018}
-Balb A., 2018, Mem. della Soc. Astron. Ital., 89, 425
-
-\bibitem[{Balbi(2018)}]{balbi_impact_2018}
-Balbi A., 2018, Astrobiology, 18, 54
-
-\bibitem[{Barab{\'{a}}si(2009)}]{barabasi_scale_2009}
-Barab{\'{a}}si A.~L., 2009, Science, 325, 412
-
-\bibitem[{Barlow(2013{\natexlab{a}})}]{barlow_galactic_2012}
-Barlow M.~T., 2013{\natexlab{a}}, Int. J. Astrobiol., 12, 63
-
-\bibitem[{Barlow(2013{\natexlab{b}})}]{barlow_galactic_2013}
-Barlow M.~T., 2013{\natexlab{b}}, Int. J. Astrobiol., 12, 63
-
-\bibitem[{Bloetscher(2019)}]{bloetscher_using_2019}
-Bloetscher F., 2019, Acta Astronaut., 155, 118
-
-\bibitem[{Borra(2012)}]{borra_searching_2012}
-Borra E.~F., 2012, AJ, 144, 181
-
-\bibitem[{Brin(1983)}]{brin_great_1983}
-Brin G., 1983, Quarterly Journal of the Royal Astronomical Society, 24, 283
-
-\bibitem[{Burchell(2006)}]{burchell_whither_2006}
-Burchell M.~J., 2006, Int. J. Astrobiol., 5, 243
-
-\bibitem[{Carroll-Nellenback {et~al}\mbox{.}(2019)Carroll-Nellenback, Frank,
-  Wright, \& Scharf}]{carroll_nellemback_fermi_2019}
-Carroll-Nellenback J., Frank A., Wright J., Scharf C., 2019, AJ, 158, 117
-
-\bibitem[{Chung(2003)}]{chung_simulation_2003}
-Chung C.~A., 2003, {Simulation Modeling Handbook: A Practical Approach
-  (Industrial and Manufacturing Engineering Series)}. CRC press
-
-\bibitem[{{\'{C}}irkovi{\'{c}}(2004)}]{cirkovic_temporal_2004}
-{\'{C}}irkovi{\'{c}} M.~M., 2004, Astrobiology, 4, 225
-
-\bibitem[{{Conway Morris}(2018)}]{conway_three_2018}
-{Conway Morris} S., 2018, Int. J. Astrobiol., 17, 287
-
-\bibitem[{Dayal {et~al}\mbox{.}(2016)Dayal, Ward, \&
-  Cockell}]{dayal_habitability_2016}
-Dayal P., Ward M., Cockell C., 2016, eprint arXiv:1606.09224
-
-\bibitem[{Do{\v{s}}ovi{\'{c}} {et~al}\mbox{.}(2019)Do{\v{s}}ovi{\'{c}},
-  Vukoti{\'{c}}, \& {\'{C}}irkovi{\'{c}}}]{Dosovic2019}
-Do{\v{s}}ovi{\'{c}} V., Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2019,
-  A\&A, 625, A98
-
-\bibitem[{Drake(1962)}]{drake_intelligent_1962}
-Drake F., 1962, ``Intelligent life in space'', Macmillan Ed.
-
-\bibitem[{Edmondson \& Stevens(2003)}]{Edmondson2003}
-Edmondson W.~H., Stevens I.~R., 2003, Int. J. Astrobiol., 2,
-  231
-
-\bibitem[{Enriquez {et~al}\mbox{.}(2017)Enriquez, Siemion, Foster, Gajjar,
-  Hellbourg, Hickish, Isaacson, Price, Croft, DeBoer, Lebofsky, MacMahon, \&
-  Werthimer}]{enriquez_breakthrough_2017}
-Enriquez J.~E. {et~al.}, 2017, ApJ, 849, 104
-
-\bibitem[{Fogg(1987)}]{fogg_temporal_1987}
-Fogg M.~J., 1987, Icarus, 69, 370
-
-\bibitem[{Forgan {et~al}\mbox{.}(2016)Forgan, Dayal, Cockell, \&
-  Libeskind}]{forgan_evaluating_2015}
-Forgan D., Dayal P., Cockell C., Libeskind N., 2016, Int. J. Astrobiol., 16, 60
-
-\bibitem[{Forgan(2009)}]{forgan_numerical_2009}
-Forgan D.~H., 2009, Int. J. Astrobiol., 8, 121
-
-\bibitem[{Forgan(2011)}]{forgan_spatiotemporal_2011}
-Forgan D.~H., 2011, Int. J. Astrobiol., 10, 341
-
-\bibitem[{Forgan(2013)}]{forgan_possibility_2013}
-Forgan D.~H., 2013, JBIS - J. Br. Interplanet. Soc., 66, 144
-
-\bibitem[{Forgan(2014)}]{forgan_collimated_2014}
-Forgan D.~H., 2014, JBIS - J. Br. Interplanet. Soc., 67, 232
-
-\bibitem[{Forgan(2017{\natexlab{a}})}]{forgan_galactic_2016}
-Forgan D.~H., 2017{\natexlab{a}}, Int. J. Astrobiol., 16, 349
-
-\bibitem[{Forgan(2017{\natexlab{b}})}]{forgan_galactic_2017}
-Forgan D.~H., 2017{\natexlab{b}}, Int. J. Astrobiol., 16, 349
-
-\bibitem[{Forgan(2019)}]{forgan_exoplanet_2017}
-Forgan D.~H., 2019, Int. J. Astrobiol., 18, 189
-
-\bibitem[{Forgan \& Rice(2010)}]{forgan_numerical_2010}
-Forgan D.~H., Rice K., 2010, Int. J. Astrobiol., 9, 73
-
-\bibitem[{Frank(2009)}]{frank_common_2009}
-Frank S.~A., 2009, {The common patterns of nature} Journal of Evolutionary Biology, 22: 1563-1585. doi:10.1111/j.1420-9101.2009.01775.x
-
-\bibitem[{Funes {et~al}\mbox{.}(2019)Funes, Florio, Lares, \&
-  Asla}]{funes_searching_2019}
-Funes J.~G., Florio L., Lares M., Asla M., 2019, Theol. Sci., 0, 1
-
-\bibitem[{Galera {et~al}\mbox{.}(2019)Galera, Galanti, \&
-  Kinouchi}]{galera_invasion_2019}
-Galera E., Galanti G.~R., Kinouchi O., 2019, Int. J. Astrobiol., 18, 316
-
-\bibitem[{Glade {et~al}\mbox{.}(2012)Glade, Ballet, \&
-  Bastien}]{glade_stochastic_2011}
-Glade N., Ballet P., Bastien O., 2012, Int. J. Astrobiol., 11, 103
-
-\bibitem[{Gobat \& Hong(2016)}]{gobat_evolution_2016}
-   Gobat R., Hong S.~E., 2016, A\&A, 592
-
-\bibitem[{Gonzalez(2005)}]{gonzalez_habitable_2005}
-Gonzalez G., 2005, Orig. Life Evol. Biosph., 35, 555
-
-\bibitem[{Gonzalez {et~al}\mbox{.}(2001)Gonzalez, Brownlee, \&
-  Ward}]{gonzalez_galactic_2001}
-Gonzalez G., Brownlee D., Ward P., 2001, Icarus, 152, 185
-
-\bibitem[{Grimaldi(2017)}]{grimaldi_signal_2017}
-Grimaldi C., 2017, Sci. Rep., 7, 46273
-
-\bibitem[{Grimaldi {et~al}\mbox{.}(2018)Grimaldi, Marcy, Tellis, \&
-  Drake}]{Grimaldi2018}
-Grimaldi C., Marcy G.~W., Tellis N.~K., Drake F., 2018, PASP, 130, 987, 054101
-
-\bibitem[{Haqq-Misra(2019)}]{haqq-misra_evolution_2019}
-Haqq-Misra J., 2019, Astrobiology, 19, 10, 1292
-
-\bibitem[{Haqq-Misra \& Kopparapu(2018)}]{haqq-misra_drake_2017}
-Haqq-Misra J., Kopparapu R.~K., 2018, preprint arXiv:1705.07816
-
-\bibitem[{Harp {et~al}\mbox{.}(2018)Harp, Ackermann, Astorga, Arbunich,
-  Barrios, Hightower, Meitzner, Barott, Nolan, Messerschmitt, Vakoch, Shostak,
-  \& Tarter}]{harp_application_2018}
-Harp G.~R. {et~al.}, 2018, ApJ, 869, 66
-
-\bibitem[{Hart(1975)}]{hart_explanation_1975}
-Hart M.~H., 1975, Quarterly Journal of the Royal Astronomical Society, 16, 128
-
-\bibitem[{Hinkel {et~al}\mbox{.}(2019)Hinkel, Hartnett, Lisse, \&
-  Young}]{hinkel_interdisciplinary_2019}
-Hinkel N., Hartnett H., Lisse C., Young P., 2019, Bull. Am. Astron. Soc., 51,
-  497
-
-\bibitem[{Hippke(2017)}]{hippke_interstellar_2017}
-Hippke, M., 2018. Benchmarking information carriers. Acta Astronautica, 151, pp.53-62.
-
-\bibitem[{Horvat(2006)}]{horvat_calculating_2007}
-Horvat M., 2006, Int. J. Astrobiol., 5, 143
-
-\bibitem[{Horvat {et~al}\mbox{.}(2011)Horvat, Naki{\'{c}}, \&
-  Oto{\v{c}}an}]{horvat_impact_2011}
-Horvat M., Naki{\'{c}} A., Oto{\v{c}}an I., 2011, Int. J. Astrobiol., 11, 51
-
-\bibitem[{{Jazayeri}(2007)}]{jazayeri_engeneering_2007}
-{Jazayeri} M., 2007, in Future of Software Engineering (FOSE '07), pp. 199--213
-
-\bibitem[{Jeong {et~al}\mbox{.}(2000)Jeong, Tombor, Albert, Oltvai, \&
-  Barabasi}]{jeong_large_2000}
-Jeong H., Tombor B., Albert R., Oltvai Z.~N., Barabasi A.~L., 2000, Nature,
-  407, 651
-
-\bibitem[{Lampton(2013)}]{lampton_information_2013}
-Lampton M., 2013, Int. J. Astrobiol., 12, 312
-
-\bibitem[{Lineweaver {et~al}\mbox{.}(2004)Lineweaver, Fenner, \&
-  Gibson}]{lineweaver_galactic_2004}
-Lineweaver C.~H., Fenner Y., Gibson B.~K., 2004, Science, 303, 59
-
-\bibitem[{Lingam \& Loeb(2018)}]{lingam_relative_2019}
-Lingam M., Loeb A., 2018, Astrobiology, 19, 28
-
-\bibitem[{Loeb {et~al}\mbox{.}(2016)Loeb, Batista, \&
-  Sloan}]{loeb_relative_2016}
-Loeb A., Batista R.~A., Sloan D., 2016, J. Cosmol. Astropart. Phys., 2016, 040
-
-\bibitem[{Loeb \& Zaldarriaga(2007)}]{loeb_eavesdropping_2006}
-Loeb A., Zaldarriaga M., 2007, J. Cosmol. Astropart. Phys., 1, id. 020
-
-\bibitem[{Maccone(2010)}]{maccone_KLT_2010}
-Maccone C., 2010, Acta Astronaut., 67, 1427
-
-\bibitem[{Maccone(2011)}]{maccone_mathematical_2011}
-Maccone C., 2011, Orig. Life Evol. Biosph., 41, 609
-
-\bibitem[{MacCone(2011)}]{maccone_SETI_2011}
-MacCone C., 2011, Acta Astronaut., 68, 63
-
-\bibitem[{MacCone(2013)}]{maccone_SETI_2013}
-MacCone C., 2013, Int. J. Astrobiol., 12, 218
-
-\bibitem[{Maccone(2014{\natexlab{a}})}]{maccone_evolution_2014}
-Maccone C., 2014{\natexlab{a}}, Int. J. Astrobiol., 13, 290
-
-\bibitem[{Maccone(2014{\natexlab{b}})}]{maccone_lognormals_2014}
-Maccone C., 2014{\natexlab{b}}, Acta Astronaut., 105, 538
-
-\bibitem[{Maccone(2015)}]{maccone_statistical_2015}
-Maccone C., 2015, Acta Astronaut., 115, 277
-
-\bibitem[{Mart{\'{i}}n \& Goldenfeld(2006)}]{martin_origin_2006}
-Mart{\'{i}}n H.~G., Goldenfeld N., 2006, Proc. Natl. Acad. Sci. U. S. A., 103,
-  10310
-
-\bibitem[{Miller \& Maloney(1963)}]{miller_testing_1963}
-Miller J.~C., Maloney C.~J., 1963, Commun. ACM, 6, 58–63
-
-\bibitem[{Mitzenmacher(2004)}]{mitzenmacher_brief_2004}
-Mitzenmacher M., 2004, Internet Math., 1, 226
-
-\bibitem[{Morrison \& Gowanlock(2015)}]{morrison_extending_2015}
-Morrison I.~S., Gowanlock M.~G., 2015, Astrobiology, 15, 683
-
-\bibitem[{Murante {et~al}\mbox{.}(2015)Murante, Monaco, Borgani, Tornatore,
-  Dolag, \& Goz}]{murante_simulating_2015}
-Murante G., Monaco P., Borgani S., Tornatore L., Dolag K., Goz D., 2015, Mon.
-  Not. R. Astron. Soc., 447, 178
-
-\bibitem[{Newman(2005)}]{newman_power_2005}
-Newman M., 2005, Contemp. Phys., 46, 323
-
-\bibitem[{Newman \& Sagan(1981)}]{newman_galactic_1981}
-Newman W.~I., Sagan C., 1981, Icarus, 46, 293
-
-\bibitem[{Oliphant(2006)}]{oliphant_numpy_2006}
-Oliphant T.~E., 2006, A guide to NumPy, Vol.~1. Trelgol Publishing USA
-
-\bibitem[{Peters(2018)}]{peters_outer_2018}
-Peters T., 2018, Int. J. Astrobiol., 17, 282
-
-\bibitem[{Prantzos(2013)}]{prantzos_joint_2013}
-Prantzos N., 2013, Int. J. Astrobiol., 12, 246
-
-\bibitem[{Ptolemaeus(2014)}]{ptolemaeus_system_2014}
-Ptolemaeus C., 2014, {System Design, Modeling, and Simulation. Using Ptolemy
-  II}. Ptolemy.org, Berkeley, p. 690
-
-\bibitem[{Rahvar(2017)}]{rahvar_cosmic_2016}
-Rahvar S., 2017, Mon. Not. R. Astron. Soc., 470, 3095
-
-\bibitem[{Ramirez {et~al}\mbox{.}(2017)Ramirez, G{\'{o}}mez-Mu{\~{n}}oz,
-  V{\'{a}}zquez, \& N{\'{u}}{\~{n}}ez}]{ramirez_new_2017}
-Ramirez R., G{\'{o}}mez-Mu{\~{n}}oz M.~A., V{\'{a}}zquez R., N{\'{u}}{\~{n}}ez
-  P.~G., 2017, Int. J. Astrobiol., 1, 34
-
-\bibitem[{Ross(2012)}]{ross_simulation_2012}
-Ross S.~M., 2012, {Simulation}. Elsevier Science Publishing Co Inc.
-
-\bibitem[{Simkin \& Roychowdhury(2006)}]{simkin_theory_2006}
-Simkin M.~V., Roychowdhury V.~P., 2006, J. Math. Sociol., 30, 33
-
-\bibitem[{Smith(2009)}]{smith_broadcasting_2009}
-Smith R.~D., 2009, Int. J. Astrobiol., 8, 101
-
-\bibitem[{Solomonides {et~al}\mbox{.}(2016)Solomonides, Kaltenegger, \&
-  Terzian}]{solomonides_probabilistic_2016}
-Solomonides E., Kaltenegger L., Terzian Y., 2016, 228th AAS, San Diego, 228, 1
-
-\bibitem[{Sornette(2006)}]{sornette_critical_2006}
-Sornette D., 2006, {Critical phenomena in natural sciences: Chaos, fractals,
-  selforganization and disorder: Concepts and tools}. Springer
-
-\bibitem[{Sotos(2019)}]{Sotos_biotechnology_2019}
-Sotos, J. (2019) Int. J. Astrobiol., 18(5), 445
-
-\bibitem[{Starling \& Forgan(2014)}]{starling_virulence_2013}
-Starling J., Forgan D.~H., 2014, Int. J. Astrobiol., 13, 45
-
-\bibitem[{Tarter(2001)}]{tarter_search_2001}
-Tarter J., 2001, Annu. Rev. Astron. Astrophys., 39, 511
-
-\bibitem[{Tarter {et~al}\mbox{.}(2009)Tarter, Backus, Blair, Cordes, Harp,
-  Henry, Horowitz, Howard, Kilsdonk, Korpela, Lasio, Levin, Shostack, \&
-  Werthimer}]{tarter_advancing_2009}
-Tarter J. {et~al.}, 2009, Astro2010 Astron. Astrophys. Decad. Surv. Sci. White
-  Pap. no. 294, 2010
-
-\bibitem[{Vakoch \& Dowd(2015)}]{vakoch_drake_2015}
-Vakoch D.~A., Dowd M.~F., 2015, {The drake equation: Estimating the prevalence
-  of extraterrestrial life through the ages}, Vakoch D.~A., Dowd M.~F., eds.
-  Cambridge University Press, Cambridge, pp. 1--319
-
-\bibitem[{{van der Walt} {et~al}\mbox{.}(2011){van der Walt}, {Colbert}, \&
-  {Varoquaux}}]{vanderwalt_numpy_2011}
-{van der Walt} S., {Colbert} S.~C., {Varoquaux} G., 2011, Computing in Science
-  Engineering, 13, 22
-
-\bibitem[\protect\citeauthoryear{Anchordoqui, Weber \& Fernandez
-   Soriano}{2017}]{anchordoqui_is_2017} Anchordoqui L.~A., Weber S.~M., Fernandez Soriano J., 2017, ICRC, 301, 254, ICRC...35
-
-\bibitem[{Vukoti{\'{c}} \&
-  {\'{C}}irkovi{\'{c}}(2012)}]{vukotic_astrobiological_2012}
-Vukoti{\'{c}} B., {\'{C}}irkovi{\'{c}} M.~M., 2012, Orig. Life Evol. Biosph.,
-  42, 347
-
-\bibitem[{Vukoti{\'{c}} {et~al}\mbox{.}(2016)Vukoti{\'{c}}, Steinhauser,
-  Martinez-Aviles, {\'{C}}irkovi{\'{c}}, Micic, \&
-  Schindler}]{vukotic_grandeur_2016}
-Vukoti{\'{c}} B., Steinhauser D., Martinez-Aviles G., {\'{C}}irkovi{\'{c}}
-  M.~M., Micic M., Schindler S., 2016, MNRAS, 459, 3512
-
-\bibitem[{Walters {et~al}\mbox{.}(1980)Walters, Hoover, \&
-  Kotra}]{walters_interstellar_1980}
-Walters C., Hoover R.~A., Kotra R., 1980, Icarus, 41, 193
-
-\bibitem[{Wright {et~al}\mbox{.}(2015)Wright, Cartier, Zhao, Jontof-Hutter, \&
-  Ford}]{wright_theGsearch_2015}
-Wright J.~T., Cartier K. M.~S., Zhao M., Jontof-Hutter D., Ford E.~B., 2015,
-  ApJ, 816, 17
-
-\bibitem[{Wright {et~al}\mbox{.}(2018)Wright, Kanodia, \&
-  Lubar}]{wright_how_2018}
-Wright J.~T., Kanodia S., Lubar E., 2018, AJ, 156, 260
+\begin{thebibliography}{90}
+\providecommand{\natexlab}[1]{#1}
+\providecommand{\url}[1]{\texttt{#1}}
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{doi: #1}\else
+  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
+
+\bibitem[Adamic(2000)]{adamic_zipf_2000}
+{Adamic, L.A.} 2000,
+\newblock {Zipf, Power-laws, and Pareto - a ranking tutorial}.
+\newblock Information Dynamics Lab, HP Labs
+\newblock URL \url{https://www.hpl.hp.com/research/idl/papers/ranking/ranking.html}.
+
+\bibitem[{Allen} \& {Schmidt}(2015)]{2015JORS....3E..15A}
+\newblock {{Allen, A., Schmidt, J.}} 2015,
+\newblock {{Looking before Leaping: Creating a Software Registry}}.
+\newblock \textit{Journal of Open Research Software}, \textbf{3},  E15, 2015.
+
+\bibitem[{Allen} et~al. (2020)]{2020ASPC..522..731A}
+\newblock {Allen, A., Berriman, G.B., DuPrie, K., Mink, J., Nemiroff,
+      R., Schmidt, J., Shamir, L., Shortridge, K., Taylor, M.B.,
+      Teuben, P.} 2020,
+\newblock {The Astrophysics Source Code Library: What's New, What's Coming}.
+\newblock In Pascal {Ballester}, Jorge {Ibsen}, Mauricio {Solar}  \& Keith
+  {Shortridge}, editors, \textit{Astronomical Data Analysis Software and
+  Systems XXVII}, volume 522 of \textit{Astronomical Society of the Pacific
+  Conference Series}, page 731, April 2020.
+
+\bibitem[Anchordoqui et al. (2017)]{anchordoqui_is_2017}
+\newblock {Anchordoqui, L.A., Weber, S.M., Fernandez Soriano, J.} 2017,
+\newblock {Is there anybody out there?}
+\newblock \textit{35th International Cosmic Ray Conference (ICRC2017)} 254.
+
+\bibitem[Anchordoqui \& Weber (2019)]{anchordoqui_upper_2019}
+\newblock {Anchordoqui, L.A., Weber, S.M.} 2019,
+\newblock {{Upper limit on the fraction of alien civilizations that develop communication technology}}.
+\newblock \textit{International Journal of High School Research (accepted).}
+\newblock arXiv e-prints, 1908.01335.
+
+\bibitem[Annis(1999)]{annis_astrophysical_1999}
+\newblock {Annis, J.} 1999,
+\newblock {{An astrophysical explanation for the "GREAT SILENCE"}}.
+\newblock \textit{Journal of the British Interplanetary Society}, \textbf{52} (1),  19--22, 1999.
+
+\bibitem[Armstrong \& Sandberg(2013)]{Armstrong2013}
+\newblock {Armstrong, S. \& Sandberg, A.} 2013,
+\newblock {{Eternity in six hours: Intergalactic spreading of intelligent life and sharpening the Fermi paradox}}.
+\newblock \textit{Acta Astronautica}, \textbf{89},  1--13, 2013.
+
+\bibitem[Balbi (2018a)]{balbi_impact_2018}
+\newblock {Balbi, A.} 2018a,
+\newblock {{The Impact of the Temporal Distribution of Communicating
+  Civilizations on Their Detectability}}.
+\newblock \textit{Astrobiology}, \textbf{18} (1),  54--58, 2018.
+
+\bibitem[Balbi (2018b)]{balbi_spatiotemporal_2018}
+\newblock {Balbi, A.} 2018b,
+\newblock {{The spatiotemporal aspects of SETI.}}
+\newblock \textit{Mem. della Soc. Astron. Ital.}, \textbf{89},  425, 2018.
+
+\bibitem[Barab{\'{a}}si(2009)]{barabasi_scale_2009}
+{Barab{\'{a}}si, A.L.} 2009,
+\newblock {Scale-free networks: A decade and beyond}.
+\newblock \textit{Science} \textbf{80}, 325 (5939),  412--413, 2009.
+
+\bibitem[Barlow (2013{\natexlab{a}})]{barlow_galactic_2012}
+{Barlow, M.T.} 2013a,
+\newblock {{Galactic exploration by directed self-replicating probes,
+  and its implications for the Fermi paradox}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{12} (1),  63--68, 2013{\natexlab{a}}.
+
+\bibitem[Barlow (2013{\natexlab{b}})]{barlow_galactic_2013}
+\newblock {Barlow, M.T.} 2013b,
+\newblock {{Galactic exploration by directed self-replicating probes,
+  and its implications for the Fermi paradox}}.
+\newblock \textit{International Journal Astrobiology}, \textbf{12} (1),  63--68, 2013{\natexlab{b}}.
+
+\bibitem[Bloetscher (2019)]{bloetscher_using_2019}
+{Bloetscher, F.} 2019, 
+\newblock Frederick Bloetscher.
+\newblock {{Using predictive Bayesian Monte Carlo- Markov Chain methods
+  to provide a probablistic solution for the Drake equation}}.
+\newblock \textit{Acta Astronautica}, \textbf{155},  118--130, 2019.
+
+\bibitem[Borra (2012)]{borra_searching_2012}
+\newblock {Borra, E.F.} 2012,
+\newblock {{Searching for extraterrestrial intelligence signals in
+  astronomical spectra, including existing data}}.
+\newblock \textit{Astronomical Journal}, \textbf{144} (6), 2012.
+
+\bibitem[Brin (1983)]{brin_great_1983}
+\newblock {Brin, G.} 1983,
+\newblock {{The great silence: the controversy concerning
+  extraterrestrial intelligent life}}.
+\newblock \textit{Quarterly Journal of the Royal Astronomical Society}, \textbf{24} (3), 283--309, 1983.
+
+\bibitem[Burchell (2006)]{burchell_whither_2006}
+\newblock {Burchell, M.J.} 2006,
+\newblock {{W(h)ither the Drake equation?}}
+\newblock \textit{International Journal of Astrobiology}, \textbf{5} (3),  243--250, 2006.
+
+\bibitem[Carroll-Nellenback et~al.(2019)Carroll-Nellenback, Frank, Wright, \& Scharf]{carroll_nellemback_fermi_2019}
+\newblock {Carroll-Nellenback, J., Frank, A., Wright, J. \& Scharf, C.}, {aug 2019}.
+\newblock {{The Fermi Paradox and the Aurora Effect: Exo-civilization Settlement, Expansion, and Steady States}}.
+\newblock \textit{The Astronomical Journal}, \textbf{158} (3),  117, 2019.
+
+\bibitem[Chung(2003)]{chung_simulation_2003}
+\newblock {Chung, C.A.} 2003,
+\newblock Christopher~A. Chung.
+\newblock {{Simulation Modeling Handbook: A Practical Approach (Industrial and Manufacturing Engineering Series)}}.
+\newblock CRC press, 2003.
+
+\bibitem[{\'{C}}irkovi{\'{c}}(2004)]{cirkovic_temporal_2004}
+\newblock {{\'{C}}irkovi{\'{c}}, M.M.} 2004,
+\newblock Milan~M. {\'{C}}irkovi{\'{c}}.
+\newblock {{The temporal aspect of the Drake equation and SETI}}.
+\newblock \textit{Astrobiology}, \textbf{4} (2),  225--231, 2004.
+
+\bibitem[{Conway Morris}(2018)]{conway_three_2018}
+\newblock {{Conway Morris}, S.} 2018,
+\newblock Simon {Conway Morris}.
+\newblock {{Three explanations for extraterrestrials: sensible, unlikely, mad}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{17} (4),  287--293, 2018.
+
+\bibitem[Dayal et~al.(2016)Dayal, Ward, \& Cockell]{dayal_habitability_2016}
+\newblock {Dayal P., Ward, M., Cockell, C.} 2016,
+\newblock Pratika Dayal, Martin Ward  \& Charles Cockell.
+\newblock {{The habitability of the Universe through 13 billion years of cosmic time.}}.
+\newblock arXiv e-prints, 1606.09224.
+
+\bibitem[Do{\v{s}}ovi{\'{c}} et~al.(2019)Do{\v{s}}ovi{\'{c}}, Vukoti{\'{c}}, \& {\'{C}}irkovi{\'{c}}]{Dosovic2019}
+\newblock {Do{\v{s}}ovi{\'{c}}, V., Vukoti{\'{c}}, B., {\'{C}}irkovi{\'{c}}, M.M.} 2019,
+\newblock Vladimir Do{\v{s}}ovi{\'{c}}, Branislav Vukoti{\'{c}}  \& Milan~M.  {\'{C}}irkovi{\'{c}}.
+\newblock {{Advanced aspects of Galactic habitability}}.
+\newblock \textit{Astronomy \& Astrophysics}, \textbf{625}, 2019.
+
+\bibitem[Drake(1962)]{drake_intelligent_1962}
+\newblock {Drake, F.D.} 1962,
+\newblock {{Intelligent life in space}},
+\newblock Macmillan
+
+\bibitem[Edmondson \& Stevens(2003)]{Edmondson2003}
+\newblock {Edmondson, W.H., Stevens, I.R.}, {2003}.
+\newblock William~H. Edmondson \& Ian~R. Stevens.
+\newblock {{The utilization of pulsars as SETI beacons}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{2} (4),  231--271, 2003.
+
+\bibitem[Enriquez et~al.(2017)Enriquez, Siemion, Foster, Gajjar, Hellbourg,
+   Hickish, Isaacson, Price, Croft, DeBoer, Lebofsky, MacMahon, \&
+  Werthimer]{enriquez_breakthrough_2017}
+\newblock {Enriquez, J.E., Siemion, A., Foster, G, Gajjar, V.,
+  Hellbourg, G, Hickish, J., Isaacson, H., Price, D.C., Croft, S.
+  DeBoer, D., Lebofsky, M., MacMahon, D., Werthimer, D.} 2017, 
+\newblock {{The Breakthrough Listen Search for Intelligent Life: 1.1-1.9
+  GHz observations of 692 Nearby Stars}}.
+\newblock \textit{Astrophysical Journal}, \textbf{849} (2),  104, 2017.
+
+\bibitem[Fogg(1987)]{fogg_temporal_1987}
+\newblock {Fogg, M.J.} 1987,
+\newblock {{Temporal aspects of the interaction among the first galactic
+  civilizations: The "interdict hypothesis"}}.
+\newblock \textit{Icarus}, \textbf{69} (2),  370--384, 1987.
+
+\bibitem[Forgan(2009)]{forgan_numerical_2009}
+\newblock {Forgan, D.H.} 2009,
+\newblock {{A numerical testbed for hypotheses of extraterrestrial life
+  and intelligence}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{8} (2),  121--131, 2009.
+
+\bibitem[Forgan \& Rice(2010)]{forgan_numerical_2010}
+\newblock {Forgan, D.H., Rice, K.} 2010,
+\newblock {{Numerical testing of the Rare Earth Hypothesis using Monte
+  Carlo realization techniques}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{9} (2),  73--80, 2010.
+
+\bibitem[Forgan(2011)]{forgan_spatiotemporal_2011}
+\newblock {Forgan, D.H.} 2011,
+\newblock {{Spatio-temporal constraints on the zoo hypothesis, and the
+  breakdown of total hegemony}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{10} (4),  341--347, 2011.
+
+\bibitem[Forgan(2013)]{forgan_possibility_2013}
+\newblock {Forgan, D.H.} 2013,
+\newblock {{On the possibility of detecting class a stellar engines
+  using exoplanet transit curves}}.
+\newblock \textit{J. Br. Interplanet. Soc.}, \textbf{66} (5-6),  144--154, 2013.
+
+\bibitem[Forgan(2014)]{forgan_collimated_2014}
+\newblock {Forgan, D.H.} 2014,
+\newblock {{Can Collimated Extraterrestrial Signals be Intercepted?}}
+\newblock \textit{J. Br. Interplanet. Soc.}, \textbf{67} (6),  232--236, 2014.
+
+\bibitem[Forgan et~al.(2016)Forgan, Dayal, Cockell, \&
+  Libeskind]{forgan_evaluating_2015}
+\newblock {Forgan, D.H., Dayal, P., Cockell, C., Libeskind, N.} 2016,
+\newblock {{Evaluating galactic habitability using high-resolution
+  cosmological simulations of galaxy formation}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{16} (1),  60--73, 2016.
+
+\bibitem[Forgan (2017)]{forgan_galactic_2017}
+\newblock {Forgan, D.H.} 2017,
+\newblock {{The Galactic Club or Galactic Cliques? Exploring the limits
+  of interstellar hegemony and the Zoo Hypothesis}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{16} (4),  349--354, 2017
+
+\bibitem[Forgan (2019)]{forgan_exoplanet_2019}
+\newblock {Forgan, D.H.} 2019,
+\newblock {{Exoplanet transits as the foundation of an interstellar communications network}}.
+\newblock \textit{Internationa Journal of Astrobiology}, \textbf{18} (3),  189--198, 2019.
+
+\bibitem[Frank(2009)]{frank_common_2009}
+\newblock {Frank, S.A.} 2009,
+\newblock {{The common patterns of nature}}, 2009.
+\newblock \textit{Journal of Evolutionary Biology}, \textbf{22}: 1563-1585
+
+\bibitem[Funes et~al.(2019)Funes, Florio, Lares, \& Asla]{funes_searching_2019}
+\newblock {Funes, J.G., Florio, L., Lares, M., Asla, M.} 2019,
+\newblock {{Searching for Spiritual Signatures in SETI Research}}.
+\newblock \textit{Theology \& Science}, \textbf{0} (0),  1--9, 2019.
+
+\bibitem[Galera et~al.(2019)Galera, Galanti, \& Kinouchi]{galera_invasion_2019}
+\newblock {Galera, E., Galanti, G.R., Kinouchi, O.} 2019,
+\newblock {{Invasion percolation solves Fermi Paradox but challenges SETI projects}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{18} (4), 316--322, 2019.
+
+\bibitem[Glade et~al.(2012)Glade, Ballet, \& Bastien]{glade_stochastic_2011}
+\newblock {Glade, N., Ballet, P., Bastien, O.} 2012,
+\newblock {{A stochastic process approach of the drake equation parameters}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{11} (2),  103--108, 2012.
+
+\bibitem[Gobat \& Hong(2016)]{gobat_evolution_2016}
+\newblock {Gobat, R., Hong, S.E.} 2016,
+\newblock {{Evolution of galaxy habitability}}.
+\newblock \textit{Astronomy \& Astrophysics}, \textbf{592}, 2016.
+
+\bibitem[Gonzalez et~al.(2001)Gonzalez, Brownlee, \& Ward]{gonzalez_galactic_2001}
+\newblock {Gonzalez, G., Brownlee, D., Ward, P.} 2001,
+\newblock {{The Galactic Habitable Zone: Galactic Chemical Evolution}}.
+\newblock \textit{Icarus}, \textbf{152} (1),  185--200, 2001.
+
+\bibitem[Gonzalez(2005)]{gonzalez_habitable_2005}
+\newblock {Gonzalez, G.} 2005,
+\newblock {{Habitable Zones in the Universe}}.
+\newblock \textit{Orig. Life Evol. Biosph.}, \textbf{35} (6),  555--606, 2005.
+
+\bibitem[Grimaldi(2017)]{grimaldi_signal_2017}
+\newblock {Grimaldi, C.} 2017,
+\newblock {{Signal coverage approach to the detection probability of
+  hypothetical extraterrestrial emitters in the Milky Way}}.
+\newblock \textit{Sci. Rep.}, \textbf{7} (1),  46273, 2017.
+
+\bibitem[Grimaldi et~al.(2018)Grimaldi, Marcy, Tellis, \& Drake]{Grimaldi2018}
+\newblock {Grimaldi, C., Marcy, G.W., Tellis, N.K., Drake, F.} 2018,
+\newblock {{Area coverage of expanding E.T. signals in the galaxy: SETI and Drake's N}}.
+\newblock \textit{Publications of the Astronomical Society of the Pacific}, \textbf{130} (987), 2018.
+
+\bibitem[Haqq-Misra \& Kopparapu(2018)]{haqq-misra_drake_2017}
+\newblock {Haqq-Misra, J., Kopparapu, R.K.} 2018,
+\newblock {{The Drake Equation as a Function of Spectral Type and Time}}.
+\newblock \textit{Habitability of the Universe Before Earth}, pages 307--319, 2018.
+
+\bibitem[Haqq-Misra(2019)]{haqq-misra_evolution_2019}
+\newblock {Haqq-Misra, J.} 2019,
+\newblock {{Does the evolution of complex life depend on the stellar spectral energy distribution?}}
+\newblock Accepted for publication in Astrobiology; doi:10.1089/ast.2018.1946
+
+\bibitem[Harp et~al.(2018)Harp, Ackermann, Astorga, Arbunich, Barrios,
+  Hightower, Meitzner, Barott, Nolan, Messerschmitt, Vakoch, Shostak, \&
+  Tarter]{harp_application_2018}
+\newblock {Harp, G.R., Ackermann, R.F., Astorga, A., Arbunich, J., Barrios, J.,
+Hightower, K., Meitzner, S., Barott, W.C., Nolan, M.C., Messerschmitt,
+D.G., Vakoch, D.A., Shostak, S.  Tarter, J.C.} 2018,
+\newblock {{The Application of Autocorrelation SETI Search Techniques in
+  an ATA Survey}}.
+\newblock \textit{Astrophysical Journal}, \textbf{869} (1),  66, 2018.
+
+\bibitem[Hart(1975)]{hart_explanation_1975}
+\newblock {Hart, M.H.} 1975,
+\newblock {{An Explanation for the Absence of Extraterrestrials on
+  Earth}}.
+\newblock \textit{Quarterly Journal of the Royal Astronomical Society}, \textbf{16},  128, 1975.
+
+\bibitem[Hinkel et~al.(2019)Hinkel, Hartnett, Lisse, \& Young]{hinkel_interdisciplinary_2019}
+\newblock {Hinkel, N., Hartnett, H., Lisse, C., Young, P.} 2019,
+\newblock {{An Interdisciplinary Perspective on Elements in Astrobiology: From Stars to Planets to Life}}.
+\newblock \textit{Bull. Am. Astron. Soc.}, \textbf{51} (3),  497, 2019.
+
+\bibitem[Hippke(2017)]{hippke_interstellar_2017}
+\newblock {Hippke, M.} 2017,
+\newblock Michael Hippke.
+\newblock {{Interstellar communication. IV. Benchmarking information carriers}}.
+\newblock \textit{Acta Astronáutica, (accepted),}
+\newblock arXiv e-prints, 1711.07962
+
+\bibitem[Horvat(2006)]{horvat_calculating_2007}
+\newblock {Horvat, M.} 2006,
+\newblock Marko Horvat.
+\newblock {{Calculating the probability of detecting radio signals from alien civilizations}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{5} (2),  143--149, 2006.
+
+\bibitem[Horvat et~al.(2011)Horvat, Naki{\'{c}}, \&
+  Oto{\v{c}}an]{horvat_impact_2011}
+\newblock {Horvat, M., Naki{\'{c}}, A., Oto{\v{c}}an, I.} 2011,
+\newblock {{Impact of technological synchronicity on prospects for CETI}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{11} (1),  51--59, 2011.
+
+\bibitem[Jeong et~al.(2000)Jeong, Tombor, Albert, Oltvai, \&
+  Barabasi]{jeong_large_2000}
+\newblock {Jeong, H., Tombor, B., Albert, R., Oltivai, Z.N., Barabasi, A.L.} 2000,
+\newblock {{The large-scale organization of metabolic networks}}.
+\newblock \textit{Nature}, \textbf{407} (6804),  651--654, 2000.
+
+\bibitem[Lampton(2013)]{lampton_information_2013}
+\newblock {Lampton, M.} 2013,
+\newblock {{Information-driven societies and Fermi's paradox}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{12} (4),  312--313, 2013.
+
+\bibitem[Lineweaver et~al.(2004)Lineweaver, Fenner, \& Gibson]{lineweaver_galactic_2004}
+\newblock {Lineweaver, C.H., Fenner, Y., Gibson, B.K.} 2004,
+\newblock {{The Galactic Habitable Zone and the Age Distribution of Complex Life in the Milky Way}}.
+\newblock \textit{Science} \textbf{80}, 303 (5654),  59--62, 2004.
+
+\bibitem[Loeb \& Zaldarriaga(2007)]{loeb_eavesdropping_2006}
+\newblock {Loeb, A., Zaldarriaga, M.} 2007,
+\newblock {{Eavesdropping on radio broadcasts from galactic
+  civilizations with upcoming observatories for redshifted 21 cm radiation}}.
+\newblock \textit{J. Cosmol. Astropart. Phys.}, \textbf{1} (1), 2007.
+
+\bibitem[Loeb et~al.(2016)Loeb, Batista, \& Sloan]{loeb_relative_2016}
+\newblock {Loeb, A., Batista, R.A., Sloan, D.} 2016,
+\newblock {{Relative likelihood for life as a function of cosmic time}}.
+\newblock \textit{J. Cosmol. Astropart. Phys.}, \textbf{8},  040--040, 2016.
+
+\bibitem[Lingam \& Loeb(2018)]{lingam_relative_2019}
+\newblock {Lingam, M., Loeb, A.} 2018,
+\newblock {{Relative Likelihood of Success in the Search for Primitive
+  versus Intelligent Extraterrestrial Life}}.
+\newblock \textit{Astrobiology}, \textbf{19} (1),  28--39, 2018.
+
+\bibitem[Maccone(2010)]{maccone_KLT_2010}
+\newblock {Maccone, C.} 2010,
+\newblock {{The KLT (Karhunen–Lo{\`{e}}ve Transform) to extend SETI
+  searches to broad-band and extremely feeble signals}}.
+\newblock \textit{Acta Astronautica}, \textbf{67} (11-12),  1427--1439, 2010.
+
+\bibitem[Maccone(2011a)]{maccone_SETI_2011}
+\newblock {Maccone, C.} 2011a,
+\newblock {{SETI and SEH (Statistical Equation for Habitables)}}.
+\newblock \textit{Acta Astronautica}, \textbf{68} (1-2),  63--75, 2011.
+
+\bibitem[Maccone(2011b)]{maccone_mathematical_2011}
+\newblock {Maccone, C.} 2011b,
+\newblock {{A Mathematical Model for Evolution and SETI}}.
+\newblock \textit{Orig. Life Evol. Biosph.}, \textbf{41} (6),  609--619, 2011.
+
+\bibitem[Maccone(2013)]{maccone_SETI_2013}
+\newblock {Maccone, C.} 2013,
+\newblock {{SETI, evolution and human history merged into a mathematical
+  model}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{12} (3),  218--245, 2013.
+
+\bibitem[Maccone(2014{\natexlab{a}})]{maccone_evolution_2014}
+\newblock {Maccone, C.} 2014a,
+\newblock Claudio Maccone.
+\newblock {{Evolution and mass extinctions as lognormal stochastic
+  processes}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{13} (4),  290--309, 2014{\natexlab{a}}.
+
+\bibitem[Maccone(2014{\natexlab{b}})]{maccone_lognormals_2014}
+\newblock {Maccone, C.} 2014b,
+\newblock Claudio Maccone.
+\newblock {{Lognormals for SETI, Evolution and mass extinctions}}.
+\newblock \textit{Acta Astronautica}, \textbf{105} (2),  538--546, 2014{\natexlab{b}}.
+
+\bibitem[Maccone(2015)]{maccone_statistical_2015}
+\newblock {Maccone, C.} 2015,
+\newblock {{Statistical Drake-Seager Equation for exoplanet and SETI searches}}.
+\newblock \textit{Acta Astronautica}, \textbf{115},  277--285, 2015.
+
+\bibitem[Mart{\'{i}}n \& Goldenfeld(2006)]{martin_origin_2006}
+\newblock    {Mart{\'{i}}n, H.G., Goldenfeld, N.} 2006,
+\newblock {{On the origin and robustness of power-law species-area relationships in ecology}}.
+\newblock \textit{Proc. Natl. Acad. Sci. U. S. A.}, \textbf{103} (27),  10310--10315, 2006.
+
+\bibitem[Mitzenmacher(2004)]{mitzenmacher_brief_2004}
+\newblock {Mitzenmacher, M.} 2004,
+\newblock {{A brief history of generative models for power law and lognormal distributions}}.
+\newblock \textit{Internet Math.}, \textbf{1} (2),  226--251, 2004.
+
+\bibitem[Morrison \& Gowanlock(2015)]{morrison_extending_2015}
+\newblock    {Morrison, I.S., Gowanlock, M.G.} 2015,
+\newblock {{Extending Galactic Habitable Zone Modeling to Include the Emergence of Intelligent Life}}.
+\newblock \textit{Astrobiology}, \textbf{15} (8),  683--696, 2015.
+
+\bibitem[Murante et~al.(2015)Murante, Monaco, Borgani, Tornatore, Dolag, \& Goz]{murante_simulating_2015}
+\newblock {Murante, G., Moncaco, P., Borgani, S., Tornatore, L., Dolag, K., Goz, D.} 2015,
+\newblock {{Simulating realistic disc galaxies with a novel sub-resolution ISM model}}.
+\newblock \textit{Mon. Not. R. Astron. Soc.}, \textbf{447} (1),  178--201, 2015.
+
+\bibitem[Newman(2005)]{newman_power_2005}
+\newblock {Newman, M.E.J.} 2005,
+\newblock {{Power laws, Pareto distributions and Zipf's law}}.
+\newblock \textit{Contemp. Phys.}, \textbf{46} (5),  323--351, 2005.
+
+\bibitem[Newman \& Sagan(1981)]{newman_galactic_1981}
+\newblock {Newman, W.I., Sagan, C.} 1981,
+\newblock {{Galactic civilizations: Population dynamics and interstellar diffusion}}.
+\newblock \textit{Icarus}, \textbf{46} (3),  293--327, 1981.
+
+\bibitem[Peters(2018)]{peters_outer_2018}
+\newblock {Peters, T.} 2018,
+\newblock {{Outer space and cyber space: meeting ET in the cloud}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{17} (4),  282--286, 2018.
+
+\bibitem[Prantzos(2013)]{prantzos_joint_2013}
+\newblock {Prantzos, N.} 2013,
+\newblock {{A joint analysis of the Drake equation and the Fermi paradox}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{12} (3),  246--253, 2013.
+
+\bibitem[Ptolemaeus(2014)]{ptolemaeus_system_2014}
+\newblock {Ptolemaeus, C.} 2014,
+\newblock \textit{{System Design, Modeling, and Simulation. Using Ptolemy II}}.
+\newblock Ptolemy.org, Berkeley, 2014.
+\newblock ISBN 9781304421067.
+
+\bibitem[Rahvar(2017)]{rahvar_cosmic_2016}
+\newblock {Rahvar, S.} 2017,
+\newblock Sohrab Rahvar.
+\newblock {{Cosmic initial conditions for a habitable universe}}.
+\newblock \textit{Mon. Not. R. Astron. Soc.}, \textbf{470} (3),  3095--3102, 2017.
+
+\bibitem[Ramirez et~al.(2017)Ramirez, G{\'{o}}mez-Mu{\~{n}}oz, V{\'{a}}zquez, \& N{\'{u}}{\~{n}}ez]{ramirez_new_2017}
+\newblock {Ramirez, R., G{\'{o}}mez-Mu{\~{n}}oz, M.A., V{\'{a}}zquez, R., N{\'{u}}{\~{n}}ez, P.g.} 2017,
+\newblock {{New numerical determination of habitability in the Galaxy: the SETI connection}}.
+\newblock \textit{International Journal of Astrobiology}, pages 1--10, 2017.
+
+\bibitem[Ross(2012)]{ross_simulation_2012}
+\newblock {Ross, S.} 2012,
+\newblock \textit{{Simulation}}.
+\newblock Elsevier Science Publishing Co Inc, 2012.
+
+\bibitem[Simkin \& Roychowdhury(2006)]{simkin_theory_2006}
+\newblock {Simkin, M.V., Roychowdhury, V.P.} 2006,
+\newblock {{Theory of Aces: Fame by Chance or Merit?}}
+\newblock \textit{J. Math. Sociol.}, \textbf{30} (1),  33--42, 2006.
+
+\bibitem[Smith(2009)]{smith_broadcasting_2009}
+\newblock {Smith, R.D.} 2009,
+\newblock {{Broadcasting but not receiving: Density dependence considerations for SETI signals}}.
+\newblock \textit{International Journal of Astrobiology}, \textbf{8} (2),  101--105, 2009.
+
+\bibitem[Solomonides et~al.(2016)Solomonides, Kaltenegger, \& Terzian]{solomonides_probabilistic_2016}
+\newblock {Solomonides, E., Kaltenegger, L., Terzian, Y.} 2016,
+\newblock {{A Probabilistic Analysis of the Fermi Paradox}}.
+\newblock \textit{228th AAS, San Diego}, \textbf{228},  1--14, 2016.
+
+\bibitem[Sornette(2006)]{sornette_critical_2006}
+\newblock {Sornette, D.} 2006,
+\newblock \textit{{Critical phenomena in natural sciences: Chaos, fractals, selforganization and disorder: Concepts and tools}}.
+\newblock Springer, 2006.
+\newblock ISBN 9783540308829.
+\newblock \doi{10.1088/0305-4470/37/40/b03}.
+
+\bibitem[Sotos(2019)]{Sotos_biotechnology_2019}
+\newblock {Sotos, J.G.} 2019,
+\newblock \textit{{Biotechnology and the lifetime of technical civilizations}}, 2019.
+
+\bibitem[Starling \& Forgan(2014)]{starling_virulence_2013}
+\newblock {Starling, J., Forgan, D.H.} 2014,
+\newblock {{Virulence as a model for interplanetary and interstellar colonization-parasitism or mutualism?}}
+\newblock \textit{International Journal of Astrobiology}, \textbf{13} (1),  45--52, 2014.
+
+\bibitem[Tarter(2001)]{tarter_search_2001}
+\newblock {Tarter, J.} 2001,
+\newblock {{The Search for Extraterrestrial Intelligence (SETI)}}.
+\newblock \textit{Annu. Rev. Astron. Astrophys.}, \textbf{39} (1),  511--548, 2001.
+
+\bibitem[Tarter et~al.(2009)Tarter, Backus, Blair, Cordes, Harp, Henry,
+   Horowitz, Howard, Kilsdonk, Korpela, Lasio, Levin, Shostack, \&
+  Werthimer]{tarter_advancing_2009}
+\newblock Tarter, J.,
+Backus, P., Blair, S., Cordes, J., Harp, G.,
+Henry, R.C., Horowitz, P., Howard, A.W., Kilsdonk, T., Korpela, E.J., 
+Lasio, J., Levin, S., Shostack, G.S., Werthimer, D. 2009,
+\newblock {{Advancing the Search for Extraterrestrial Intelligence}}.
+\newblock \textit{Astro2010 Astron. Astrophys. Decad. Surv. Sci. White Pap.} no. 294, 2010, 2009.
+\newblock URL \url{http://adsabs.harvard.edu/abs/2009astro2010S.294T}.
+
+\bibitem[Vakoch \& Dowd(2015)]{vakoch_drake_2015}
+\newblock {Vakoch, D.A., Dowd, M.F.} 2015,
+\newblock \textit{{The drake equation: Estimating the prevalence of extraterrestrial life through the ages}}.
+\newblock Cambridge University Press, Cambridge, 2015.
+
+\bibitem[Vukoti{\'{c}} \&
+\newblock {\'{C}}irkovi{\'{c}}(2012)]{vukotic_astrobiological_2012} {Vukoti{\'{c}}, B., {\'{C}}irkovi{\'{c}}, M.M.} 2012,
+\newblock \textit{{Astrobiological Complexity with Probabilistic Cellular Automata}}.
+\newblock Orig. Life Evol. Biosph., \textbf{42} (4),  347--371, 2012.
+
+\bibitem[Vukoti{\'{c}} et~al.(2016)Vukoti{\'{c}}, Steinhauser, Martinez-Aviles,
+   {\'{C}}irkovi{\'{c}}, Micic, \& Schindler]{vukotic_grandeur_2016}
+\newblock {Vukoti{\'{c}}, B., Steinhauser, D., Martinez-Aviles, G., {\'{C}}irkovi{\'{c}}, M.M., Micic, M., Schindler, S.} 2016,
+\newblock {{‘Grandeur in this view of life': N -body simulation models of the Galactic habitable zone}}.
+\newblock \textit{Mon. Not. R. Astron. Soc.}, \textbf{459} (4),  3512--3524, 2016.
+
+\bibitem[Walters et~al.(1980)Walters, Hoover, \& Kotra]{walters_interstellar_1980}
+\newblock {Walters, C., Hoover, R.A., Kotra, R.K.} 1980,
+\newblock \textit{{Interstellar colonization: A new parameter for the Drake equation?}}
+\newblock Icarus, \textbf{41} (2),  193--197, 1980.
+
+\bibitem[Wright et~al.(2015)Wright, Cartier, Zhao, Jontof-Hutter, \& Ford]{wright_theGsearch_2015}
+\newblock {Wright, J.T., Cartier, K.M.S., Zhao, M., Jontof-Hutter, D., Ford, E.B.} 2015,
+\newblock {{the Ĝ Search for Extraterrestrial Civilizations With Large
+  Energy Supplies. Iv. the Signatures and Information Content of Transiting Megastructures}}.
+\newblock \textit{Astrophysical Journal}, \textbf{816} (1),  17, 2015.
+
+\bibitem[Wright et~al.(2018)Wright, Kanodia, \& Lubar]{wright_how_2018}
+\newblock {Wright, J.T., Kanodia, S., Lubar, E.} 2018,
+\newblock {{How Much SETI Has Been Done? Finding Needles in the n-dimensional Cosmic Haystack}}.
+\newblock \textit{The Astronomical Journal}, \textbf{156} (6),  260, 2018.
 
